### 2017-09-16 paper/DPC_T&A.tex
paper first commit

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\journal{Journal of \LaTeX\ Templates}

%%%%%%%%%%%%%%%%%%%%%%%
%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{Data Predictive Control: Theory and Applications\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{The research leading to these results has received funding from the Italian Government under Cipe resolution n.135 (Dec. 21, 2012), project \emph{INnovating City Planning through Information and Communication Technologies} (INCIPICT).}

\author[FSaddress,EC]{Francesco Smarra\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{francesco.smarra@univaq.it@univaq.it}

\author[AJaddress,EC]{Achin Jain}
\ead{achinj@seas.unpenn.edu}
\author[TDRaddress,EC]{Tullio De Rubeis}
\ead{tullio.derubeis@graduate.univaq.it}
\author[TDRaddress]{Dario Ambrosini}
\ead{dario.ambrosini@ing.univaq.it}
\author[FSaddress]{Alessandro D'Innocenzo}
\ead{alessandro.dinnocenzo@univaq.it}
\author[AJaddress]{Rahul Mangharam}
\ead{rahulm@seas.upenn.edu}

\address[EC]{Equal contribution}
\address[FSaddress]{Department of Information Engineering, Computer Science and Mathematics, Università degli Studi dell’Aquila, L’Aquila, Italy}
\address[TDRaddress]{Department of Industrial and Information Engineering and Economics, Università degli Studi dell’Aquila, L’Aquila, Italy}
\address[AJaddress]{Department of Electrical and Systems Engineering, University of Pennsylvania, Philadelphia, USA}

\begin{abstract}
TBD
\end{abstract}

\begin{keyword}
Energy efficiency, predictive control, machine learning.
\end{keyword}

\end{frontmatter}

%\linenumbers

\section{Introduction}

\paragraph{Motivations} TBD

\paragraph{State of the art} TBD

\paragraph{Contributions} TBD

\paragraph{Paper organization} TBD

\section{Problem definition and objectives} TBD

\section{Data Predictive Control using Random Forests} TBD

\subsection{Modeling} TBD

\subsection{Predictive Control} TBD

\section{Case study 1: bilinear model} TBD

\section{Case study 2: EnergyPlus model} TBD

\section{Case study 3: real house} TBD

\subsection{Description of the house} TBD

\subsection{EnergyPlus model} TBD

\subsection{Simulation Results} TBD

\paragraph{Energy-efficiency improvements}

\section{Conclusion and future work} TBD

\section*{References}

\bibliography{mybibfile}

\end{document}

### 2017-09-17 code/Yalmip/yalmip/@sdpvar/log.m
added code

@@ -1,0 +1,124 @@
+function varargout = log(varargin)
+%LOG (overloaded)
+
+% Author Johan Löfberg
+% $Id: log.m,v 1.21 2009-05-05 11:42:00 joloef Exp $
+switch class(varargin{1})
+
+    case 'double'
+        error('Overloaded SDPVAR/LOG CALLED WITH DOUBLE. Report error')
+
+    case 'sdpvar'
+         % Try to detect logsumexp construction etc
+        varargout{1} = check_for_special_cases(varargin{:});
+        % Nope, then just define this logarithm
+        if isempty(varargout{1})
+            varargout{1} = InstantiateElementWise(mfilename,varargin{:});
+        end
+        
+    case 'char'
+
+        X = varargin{3};      
+        F = set(X >= 1e-8);
+
+        operator = struct('convexity','concave','monotonicity','increasing','definiteness','none','model','callback');
+        operator.inverse = 'exp';
+        operator.convexhull = @convexhull;
+        operator.bounds = @bounds;
+        operator.domain = [0 inf];
+        operator.derivative = @(x)(1./(abs(x)+eps));
+
+        varargout{1} = F;
+        varargout{2} = operator;
+        varargout{3} = X;
+
+    otherwise
+        error('SDPVAR/LOG called with CHAR argument?');
+end
+
+function [L,U] = bounds(xL,xU)
+if xL <= 0
+    % The variable is not bounded enough yet
+    L = -inf;
+else
+    L = log(xL);
+end
+if xU < 0
+    % This is an infeasible problem
+    L = inf;
+    U = -inf;
+else
+    U = log(xU);
+end
+
+function [Ax, Ay, b, K] = convexhull(xL,xU)
+K = [];
+if xL <= 0
+    fL = inf;
+else
+    fL = log(xL);
+end
+fU = log(xU);
+dfL = 1/(xL);
+dfU = 1/(xU);
+xM = (xU - xL)/(fU-fL);
+fM = log(xM);
+dfM = 1/xM;
+
+[Ax,Ay,b] = convexhullConcave(xL,xM,xU,fL,fM,fU,dfL,dfM,dfU);
+remove = isinf(b) | isinf(Ax) | isnan(b);
+if any(remove)
+    remove = find(remove);
+    Ax(remove)=[];
+    b(remove)=[];
+    Ay(remove)=[];
+end
+
+
+function f = check_for_special_cases(x)
+f = [];
[…]

### 2017-09-17 code/Yalmip/yalmip/@sdpvar/model.m

@@ -1,0 +1,125 @@
+function [properties,F,arguments,fcn]=model(X,method,options,extstruct,w)
+%MODEL  Internal function to extracts nonlinear operator models
+%
+% [properties,F] = model(x)
+%
+% MODEL returns the constraints needed to model a variable related to an
+% extended operator such as min, max, abs, norm, geomean, ...
+%
+% Examples :
+%
+% sdpvar x y;
+% t = min(x,y);
+% [properties,F] = model(t)
+% Gives F = [t<x, t<y]
+
+% Author Johan Löfberg
+% $Id: model.m,v 1.62 2009-05-05 07:20:46 joloef Exp $
+
+
+extvar = getvariables(X);
+arguments   = cell(1,length(extvar));
+properties  = cell(1,length(extvar));
+
+if nargin<2
+    method = 'graph';
+end
+
+if nargin < 3
+    options = [];
+end
+
+if nargin < 5
+    % Not used
+    w = [];
+end
+
+if nargin<4
+    extstruct = yalmip('extstruct',extvar);
+elseif isempty(extstruct)
+    extstruct = yalmip('extstruct',extvar);
+end
+
+if isempty(extstruct)
+    error('This is not a nonlinear operator variable');
+end
+
+fcn = extstruct.fcn;
+try
+    n = yalmip('nvars');
+    [F,properties,arguments] = feval(fcn,method,extstruct.var,extstruct.arg{1:end-1});
+    newAux = n+1:yalmip('nvars');
+    involved = [];
+    for i = 1:length(extstruct.arg)-1
+        involved = union(involved,getvariables(extstruct.arg{i}));
+    end
+    if ~isempty(options)
+        if ~(strcmp(options.robust.auxreduce,'none'))
+            % This info is only needed when we do advanced Robust optimization
+            yalmip('setdependence',[getvariables(extstruct.var) newAux],involved);
+            yalmip('setdependence',[getvariables(extstruct.var)],newAux);
+        end
+    end
+catch
+    error(['Failed when trying to create a model for the "' extstruct.fcn '" operator']);
+end
+
+% Make sure all operators have these properties
+if ~isempty(properties)
+    if ~iscell(properties)
+        properties = {properties};
+    end
+    for i = 1:length(properties)
+        properties{i}.name = fcn;
+        properties{i} = assertProperty(properties{i},'definiteness','none');
+        properties{i} = assertProperty(properties{i},'convexity','none');
[…]

### 2017-09-17 code/Yalmip/yalmip/@sdpvar/ne.m

@@ -1,0 +1,44 @@
+function F = ne(X,Y)
+%NE (overloaded)
+%
+%    F = set(ne(x,y))
+%
+%   See also SDPVAR/AND, SDPVAR/OR, BINVAR, BINARY
+
+% Author Johan Löfberg
+% $Id: ne.m,v 1.4 2009-02-25 12:29:25 joloef Exp $
+
+% Models NE using logic constraints
+
+% bin1 = isa(X,'sdpvar') | isa(X,'double');
+% bin2 = isa(Y,'sdpvar') | isa(Y,'double');
+%
+% if ~(bin1 & bin2)
+%     error('Not equal can only be applied to integer data')
+% end
+
+if isa(X,'sdpvar') & isa(Y,'sdpvar') & is(X,'binary') & is(Y,'binary')
+    F = set(X + Y == 1);
+elseif 0%isa(X,'sdpvar') & is(X,'binary') &  isa(Y,'double') &  ismember(Y,[0 1])
+    zv = find((Y == 0));
+    ov = find((Y == 1));
+    lhs = 0;
+    if ~isempty(zv)
+        lhs = lhs + sum(extsubsref(X,zv));
+    end
+    if ~isempty(ov)
+        lhs = lhs + sum(1-extsubsref(X,ov));
+    end
+    F = set(lhs >=1);
+elseif isa(X,'sdpvar') & is(X,'integer') &  isa(Y,'double')
+    X = reshape(X,[],1);
+    Y = reshape(Y,[],1);
+    if length(X)>1 & length(Y)==1
+        Y = repmat(Y,length(X),1);
+    end
+    less = binvar(length(X),1);
+    larger = binvar(length(X),1);
+    F = [implies(less,X<=Y-1), implies(larger,X>=Y+1), sum(less) + sum(larger) >= 1];
+else
+    F = set((X<=Y-0.5) | (X>=Y+0.5));
+end

### 2017-09-17 code/Yalmip/yalmip/@sdpvar/sdpvar.m

@@ -1,0 +1,796 @@
+function sys = sdpvar(varargin)
+%SDPVAR Create symbolic decision variable
+%
+%   You can create a sdpvar variable by:
+%     X = SDPVAR(n)               Symmetric nxn matrix
+%     X = SDPVAR(n,n)             Symmetric nxn matrix
+%     X = SDPVAR(n,m)             Full nxm matrix (n~=m)
+%
+%   Definition of multiple scalars can be simplified
+%     SDPVAR x y z w
+%
+%   The parametrizations supported are
+%     X = SDPVAR(n,n,'full')      Full nxn matrix
+%     X = SDPVAR(n,n,'symmetric') Symmetric nxn matrix
+%     X = SDPVAR(n,n,'diagonal')  Diagonal matrix
+%     X = SDPVAR(n,n,'toeplitz')  Symmetric Toeplitz
+%     X = SDPVAR(n,n,'hankel')    Unsymmetric Hankel (zero below the first anti-diagonal)
+%     X = SDPVAR(n,n,'rhankel')   Symmetric Hankel
+%     X = SDPVAR(n,n,'skew')      Skew-symmetric
+%     X = SDPVAR(n,n,'diagonal')  Diagonal
+%
+%   The letters 'sy','f','ha', 't' and 'sk' are searched for in the third argument
+%   hence sdpvar(n,n,'toeplitz') gives the same result as sdpvar(n,n,'t')
+%
+%   Only square Toeplitz and Hankel matries are supported
+%
+%   A scalar is defined as a 1x1 matrix
+%
+%   Higher-dimensional matrices are also supported, although this currently
+%   is an experimental feature with limited use. The type flag applies to
+%   the lowest level slice.
+%
+%     X = SDPVAR(n,n,n,'full')      Full nxnxn matrix
+%
+%   In addition to the matrix type, a fourth argument
+%   can be used to obtain a complex matrix. All the
+%   matrix types above apply to a complex matrix, and
+%   in addition a Hermitian type is added
+%
+%     X = SDPVAR(n,n,'hermitian','complex') Complex Hermitian nxn matrix (X=X'=conj(X.'))
+%
+%   The other types are obtained as above
+%     X = SDPVAR(n,n,'symmetric','complex') Complex symmetric nxn matrix (X=X.')
+%     X = SDPVAR(n,n,'full','complex')      Complex full nxn matrix
[…]

### 2017-09-17 code/Yalmip/yalmip/@sdpvar/subsref.m

@@ -1,0 +1,313 @@
+function varargout = subsref(varargin)
+%SUBSREF (overloaded)
+
+% Author Johan Löfberg
+% $Id: subsref.m,v 1.23 2009-09-28 07:36:25 joloef Exp $
+
+
+% Stupid first slice call (supported by MATLAB)
+% x = sdpvar(2);x(1,:,:)
+if length(varargin{2})==1
+    if  length(varargin{2}.subs) > 2 & isequal(varargin{2}.type,'()')
+        i = 3;
+        ok = 1;
+        while ok & (i <= length(varargin{2}.subs))
+            ok = ok & (isequal(varargin{2}.subs{i},1) | isequal(varargin{2}.subs{i},':'));
+            i = i + 1;
+        end
+        if ok
+            varargin{2}.subs = {varargin{2}.subs{1:2}};
+        else
+            error('??? Index exceeds matrix dimensions.');
+        end
+    end
+end
+
+if (isequal(varargin{2}.type,'()') & ((isa(varargin{2}.subs{1},'sdpvar')) | (length(varargin{2}.subs)==2 & isa(varargin{2}.subs{2},'sdpvar'))))
+    % *****************************************
+    % Experimental code for varaiable indicies
+    % *****************************************
+    varargout{1} = milpsubsref(varargin{:});
+    return
+else
+    X = flush(varargin{1});
+    Y = varargin{2};
+end
+
+try
+    switch Y(1).type
+        case '()'
+            if  isa(Y(1).subs{1},'constraint')
+                error('Conditional indexing not supported.');
+            end
+            % Check for simple cases to speed things up (yes, ugly but we all want speed don't we!)
+            switch size(Y(1).subs,2)
+                case 1
+                    if isa(Y(1).subs{1},'sdpvar')
+                        varargout{1} = yalmip('addextendedvariable',mfilename,varargin{:});
+                        return
+                    else
+                        y = subsref1d(X,Y(1).subs{1});
+                    end
+                case 2
+                    y = subsref2d(X,Y.subs{1},Y(1).subs{2});
+                otherwise
+                    error('Indexation error.');
+            end
+        case '{}'
+            varargout{nargout} = [];
+            
+            % it could be the case that we have an extended variable
+            % This is a bit tricky, so we do the best we can; assume that
+            % we want to replace the internal argument wih the new
+            % expression
+            OldArgument = recover(depends(X));
[…]

### 2017-09-17 code/Yalmip/yalmip/@sdpvar/times.m

@@ -1,0 +1,264 @@
+function y = times(X,Y)
+%TIMES (overloaded)
+
+% Author Johan Löfberg
+% $Id: times.m,v 1.30 2009-09-28 12:53:16 joloef Exp $
+
+% Check dimensions
+[n,m]=size(X);
+if ~((prod(size(X))==1) | (prod(size(Y))==1))
+    if ~((n==size(Y,1) & (m ==size(Y,2))))
+        error('Matrix dimensions must agree.')
+    end
+end;
+
+% Convert block objects
+if isa(X,'blkvar')
+    X = sdpvar(X);
+end
+
+if isa(Y,'blkvar')
+    Y = sdpvar(Y);
+end
+
+
+if isempty(X)
+    YY = full(reshape(Y.basis(:,1),Y.dim(1),Y.dim(2)));
+    y = X.*YY;
+    return
+elseif isempty(Y)
+    XX = full(reshape(X.basis(:,1),X.dim(1),X.dim(2)));
+    y = XX.*Y;
+    return
+end
+
+if (isa(X,'sdpvar') & isa(Y,'sdpvar'))
+    X = flush(X);
+    Y = flush(Y);
+    if (X.typeflag==5) & (Y.typeflag==5)
+        error('Product of norms not allowed');
+    end
+
+    try
+        
+        y = check_for_special_case(Y,X);
+        if ~isempty(y)
+            return
+        end
+        
+        
+        x_isscalar =  (X.dim(1)*X.dim(2)==1);
+        y_isscalar =  (Y.dim(1)*Y.dim(2)==1);
+
+        all_lmi_variables = uniquestripped([X.lmi_variables Y.lmi_variables]);
+        Z = X;Z.dim(1) = 1;Z.dim(2) = 1;Z.lmi_variables = all_lmi_variables;Z.basis = [];
+
+        % Awkward code due to bug in ML6.5
+        Xbase = reshape(X.basis(:,1),X.dim(1),X.dim(2));
+        Ybase = reshape(Y.basis(:,1),Y.dim(1),Y.dim(2));
+        if x_isscalar
+            Xbase = sparse(full(Xbase));
+        end
+        if y_isscalar
+            Ybase = sparse(full(Ybase));
+        end
+
+
+        index_Y = zeros(length(all_lmi_variables),1);
+        index_X = zeros(length(all_lmi_variables),1);
+        for j = 1:length(all_lmi_variables)
+            indexy = find(all_lmi_variables(j)==Y.lmi_variables);
[…]

### 2017-09-17 code/Yalmip/yalmip/Contents.m

@@ -1,0 +1,19 @@
+% YALMIP
+% Version 13-Feb-2013
+%
+% Information
+%
+% Variables.
+%   sdpvar       - Create SDPVAR variable 
+%   ...find out more in the Wiki!
+
+% Inequalites and constraints.
+%   ...check the Wiki
+%
+% Optimization related.
+%   solvesdp     - Computes solution to optimization problem
+%   ...
+%
+
+% Author Johan Löfberg
+

### 2017-09-17 code/Yalmip/yalmip/extras/@lmi/checkset.m

@@ -1,0 +1,236 @@
+function [pres,dres] = checkset(F)
+% CHECKSET(F)  Displays/calculates constraint residuals on constraint F
+%
+% [pres,dres] = checkset(F)
+%
+% pres : Primal constraint residuals
+% dres : Dual constraint residuals
+%
+% If no output argument is supplied, tabulated results are displayed
+%
+% Primal constraint residuals are calculated as:
+%
+%  Semidefinite constraint F(x)>0 : min(eig(F))
+%  Element-wise constraint F(x)>0 : min(min(F))
+%  Equality constraint F==0       : -max(max(abs(F)))
+%  Second order cone t>||x||      : t-||x||
+%  Integrality constraint on x    : max(abs(x-round(x)))
+%  Rank constraint rank(X) < r     : r-rank(X)
+%  Sum-of-square constraint       : Minus value of largest (absolute value) coefficient 
+%                                   in the polynomial p-v'*v
+%
+% Dual constraints are evaluated similarily.
+%
+%    See also   SET, SOLVESDP, SOLVESOS, SOSD, DUAL
+
+% Author Johan Löfberg
+% $Id: checkset.m,v 1.19 2008-02-21 15:32:09 joloef Exp $
+
+% Check if solution avaliable
+currsol = evalin('caller','yalmip(''getsolution'')');
+if isempty(currsol)
+    disp('No solution available.')
+    return
+end
+
+nlmi = size(F.clauses,2);
+spaces = ['                                    '];
+if (nlmi == 0)
+    if nargout == 0
+        disp('empty LMI')
+    else
+        pres = [];
+        dres = [];
+    end 
+    return
+end
+
+lmiinfo{1} = 'Matrix inequality';
+lmiinfo{2} = 'Elementwise inequality';
+lmiinfo{3} = 'Equality constraint';
+lmiinfo{4} = 'Second order cone constraint';
+lmiinfo{5} = 'Rotated Lorentz constraint';
+lmiinfo{7} = 'Integer constraint';
+lmiinfo{8} = 'Binary constraint';
+lmiinfo{9} = 'KYP constraint';
+lmiinfo{10} = 'Eigenvalue constraint';
+lmiinfo{11} = 'SOS constraint';
+lmiinfo{15} = 'Uncertainty declaration';
+lmiinfo{55} = 'Complementarity constraint';
+header = {'ID','Constraint','Type','Primal residual','Dual residual','Tag'};
+
+if nargout==0
+    disp(' ');
[…]

### 2017-09-17 code/Yalmip/yalmip/extras/@lmi/display.m

@@ -1,0 +1,129 @@
+function sys = display(X)
+%display           Displays a SET object.
+
+% Author Johan Löfberg
+% $Id: display.m,v 1.12 2009-05-29 08:05:12 joloef Exp $
+
+nlmi = size(X.clauses,2);
+nlmi = length(X.LMIid);
+
+if (nlmi == 0)
+    disp('empty SET')
+    return
+end
+
+lmiinfo{1} = 'Matrix inequality';
+lmiinfo{2} = 'Element-wise inequality';
+lmiinfo{3} = 'Equality constraint';
+lmiinfo{4} = 'Second order cone constraint';
+lmiinfo{5} = 'Rotated Lorentz constraint';
+lmiinfo{7} = 'Integrality constraint';
+lmiinfo{8} = 'Binary constraint';
+lmiinfo{9} = 'KYP constraint';
+lmiinfo{10}= 'Eigenvalue constraint';
+lmiinfo{11}= 'Sum-of-square constraint';
+lmiinfo{12}= 'Logic constraint';
+lmiinfo{13}= 'Parametric declaration';
+lmiinfo{14}= 'Low-rank data declaration';
+lmiinfo{15}= 'Uncertainty declaration';
+lmiinfo{16}= 'Distribution declaration';
+lmiinfo{20}= 'Power cone constraint';
+lmiinfo{30}= 'User defined compilation';
+lmiinfo{40}= 'Generalized KYP constraint';
+lmiinfo{50}= 'Special ordered set of type 2';
+lmiinfo{51}= 'Special ordered set of type 1';
+lmiinfo{52}= 'Semi-continuous variable';
+lmiinfo{53}= 'Semi-integer variable';
+lmiinfo{54}= 'Second order cone constraints';
+lmiinfo{55}= 'Complementarity constraint';
+lmiinfo{56}= 'Meta constraint';
+
+headers = {'ID','Constraint','Type','Tag'};
+rankVariables = yalmip('rankvariables');
+extVariables = yalmip('extvariables');
+if nlmi>0
+    for i = 1:nlmi
+        
+        data{i,1} = ['#' num2str(i)];
+        data{i,2} = X.clauses{i}.symbolic;
+        data{i,3} = lmiinfo{X.clauses{i}.type};
+        data{i,4} = '';
+        if length(getvariables(X.clauses{i}.data)) == 1
+            if any(ismember(getvariables(X.clauses{i}.data),rankVariables))
+                 data{i,3} = 'Rank constraint';                
+            end
+        end
[…]

### 2017-09-17 code/Yalmip/yalmip/extras/@lmi/envelope.m

@@ -1,0 +1,136 @@
+function [E,P] = envelope(C,x)
+%ENVELOPE Create linear approximation of envelope of nonlinear constraint
+%
+%     [E,P] = envelope(C,x)
+%
+% C: Constraint object involving nonlinear expresions
+% x: Optional: The linear variables of interest (projection onto these will be performed)
+% E: Constraint object representing the envelope approximation
+% P: Optional: Polyhedral object representing E (MPT Toolbox object)
+%
+% Examples
+%
+% In order to derive the outer approximation in the original linear
+% variables, the model has to be projected onto the linear variables
+% This projetion based format requires MPT3
+%   sdpvar x u
+%   E = envelope([-1 <= x <= 1, u == x^3],[x;u]);
+%   plot(E)
+%   xx = (-1:0.01:1);hold on;plot(xx,xx.^3)
+%
+%   E = envelope([-1 <= x <= 1,x+sin(pi*x) <= u <= 4-x^2],[x;u]);
+%   plot(E)
+%   xx = (-1:0.01:1);hold on;plot(xx,xx+sin(pi*xx),xx,4-xx.^2)
+%
+% Alternatively, we can create a model which adds the outer approximation
+% cuts for the envelopes of the nonlinear variables, but keep the nonlinear
+% variables, and then plot the projection of the outer approximation,
+% keeping in mind that we now have to relax the nonlinear variables (this
+% is the model which thus would be used in a branch&bound scheme)
+%   E = envelope([-1 <= x <= 1,x+sin(pi*x) <= u <= 4-x^2]);
+%   plot(E,[x;u],[],[],sdpsettings('relax',1))
+%   xx = (-1:0.01:1);hold on;plot(xx,xx+sin(pi*xx),xx,4-xx.^2)
+
+% Author Johan Löfberg
+
+[aux1,aux2,aux3,p] = export(C,[],sdpsettings('solver','bmibnb'));
+
+if isempty(p)
+    if ~isempty(aux3)
+        aux3.info
+    end
+    error('Failed to export a model')
+end
+
+% Copied from bmibnb
+p.EqualityConstraintState = ones(p.K.f,1);
+p.InequalityConstraintState = ones(p.K.l,1);
[…]

### 2017-09-17 code/Yalmip/yalmip/extras/@lmi/hull.m

@@ -1,0 +1,78 @@
+function [Fhull,t,y] = hull(varargin)
+% HULL  Construct a model of the convex hull
+%
+% H = hull(F1,F2,...)
+%
+% OUTPUT
+%   H   : SET object describing the convex hull of the input constraints
+%
+% INPUT
+%   Fi  : SET objects with constraints
+%
+% Note that the convex representation of the convex hull requires a lifting
+% (introduction of auxially variables). Hence, if you have many set of
+% constraints, your problem rapidly grows large.
+
+if nargin==1
+    Fhull = varargin{1};
+    t = [];
+end
+% Pre-process to convert convex quadratic constraints to socp constraints.
+% This makes the perspective code easier.
+% We also wexpand all graph-operators etc to find out all involved
+% variables and constraints
+
+for i = 1:nargin
+    varargin{i} = convertquadratics(varargin{i});
+    varargin{i} = expandmodel(varargin{i},[]);
+end
+
+variables = [];
+for i = 1:nargin
+    % quick fix
+    if isa(varargin{i},'constraint')
+        varargin{i} = set(varargin{i});
+    end
+    if ~(isa(varargin{i},'lmi') | isa(varargin{i},'socc'))
+        error('Hull can only be applied to linear constraints');
+    elseif ~(islinear(varargin{i}))
+        error('Hull can only be applied to linear and convex quadratic constraints');
+    end
+    variables = unique([variables depends(varargin{i})]);
+end
+
+if nargin == 1
+    Fhull = varargin{1};
+    return
+end
+
+% Define variables. To allow users to easily plot convex hulls, we mark the
+% internally defined variables as auxilliary. YALMIP will then not plot
+% w.r.t these variables
+nInitial = yalmip('nvars');
+y = sdpvar(repmat(length(variables),1,nargin),repmat(1,1,nargin));
+t = sdpvar(nargin,1);
+nNow = yalmip('nvars');
+yalmip('addauxvariables',nInitial+1:nNow);
+
+Fhull = set([]);
+for i = 1:nargin
+    Fi = varargin{i};
+    tvariable = getvariables(t(i));
+    for j = 1:length(Fi.clauses)
+        local_variables = getvariables(Fi);
+        Xi = Fi.clauses{j}.data;
+        local_variables = getvariables(Xi);
+        local_index = find(ismember(variables,local_variables));
[…]

### 2017-09-17 code/Yalmip/yalmip/extras/@lmi/kkt.m

@@ -1,0 +1,283 @@
+function [KKTConstraints, details] = kkt(F,h,parametricVariables,ops);
+%KKT Create KKT system
+%
+% [KKTConstraints, details] = kkt(Constraints,Objective,parameters,options)
+
+% Author Johan Löfberg
+
+[aux1,aux2,aux3,model] = export(F,h,sdpsettings('solver','quadprog','relax',2));
+if isempty(model)
+    error('KKT system can only be derived for LPs or QPs');
+end
+model.problemclass.constraint.binary = 0;
+model.problemclass.constraint.integer = 0;
+model.problemclass.constraint.semicont = 0;
+model.problemclass.constraint.sos1 = 0;
+model.problemclass.constraint.sos2 = 0;
+if ~ismember(problemclass(model), {'LP', 'Convex QP', 'Nonconvex QP'})
+    error('KKT system can only be derived for LPs or QPs');
+end
+if ~isempty(model.binary_variables) | ~isempty(model.integer_variables) | ~isempty(model.semicont_variables)
+    comb = [model.used_variables(model.binary_variables) model.used_variables(model.integer_variables) model.used_variables(model.semicont_variables) ];
+    if ~isempty(intersect(comb,getvariables(decisionvariables)))
+        error('KKT system cannot be derived due to binary or integer decision variables');
+    end
+end
+
+if nargin < 3
+    x = recover(model.used_variables);
+    parameters = [];
+else
+    % Make sure they are sorted    
+    parameters = getvariables(parametricVariables);
+    x = recover(setdiff(model.used_variables,parameters));
+    notparameters = find(ismember(model.used_variables,getvariables(x)));%parameters);
+    parameters = setdiff(1:length(model.used_variables),notparameters);
+    if ~isempty(parameters)
+        y = recover(model.used_variables(parameters));
+    end
+end
+
+if nargin < 4
+    ops = sdpsettings;
+end
+if isempty(ops)
+    ops = sdpsettings;
+end
+
+% Ex==f
+E = model.F_struc(1:model.K.f,2:end);
+f = -model.F_struc(1:model.K.f,1);
[…]

### 2017-09-17 code/Yalmip/yalmip/extras/@lmi/lmi2sedumistruct.m

@@ -1,0 +1,578 @@
+function [F_struc,K,KCut,schur_funs,schur_data,schur_variables] = lmi2sedumistruct(F)
+%lmi2sedumistruct   Internal function: Converts LMI to format needed in SeDuMi
+%
+% % Author Johan Löfberg
+% % $Id: lmi2sedumistruct.m,v 1.35 2009-09-29 12:02:40 joloef Exp $
+
+nvars = yalmip('nvars'); %Needed lot'sa times...
+
+% We first browse to see what we have got and the
+% dimension of F_struc (massive speed improvement)
+type_of_constraint = zeros(length(F.LMIid),1);%zeros(size(F.clauses,2),1);
+any_cuts = 0;
+for i = 1:length(F.LMIid)%size(F.clauses,2)
+    type_of_constraint(i) = F.clauses{i}.type;
+    if F.clauses{i}.cut
+        any_cuts = 1;
+    end
+end
+
+F_struc = [];
+
+schur_data = [];
+schur_funs = [];
+schur_variables = [];
+
+sdp_con = find(type_of_constraint == 1 | type_of_constraint == 9 | type_of_constraint == 40);
+lin_con = find(type_of_constraint == 2 | type_of_constraint == 12);
+equ_con = find(type_of_constraint == 3);
+qdr_con = find(type_of_constraint == 4);
+mqdr_con = find(type_of_constraint == 54);
+rlo_con = find(type_of_constraint == 5);
+pow_con = find(type_of_constraint == 20);
+sos2_con = find(type_of_constraint == 50);
+sos1_con = find(type_of_constraint == 51);
+
+% SeDuMi struct
+K.f = 0; % Linear equality
+K.l = 0; % Linear inequality
+K.q = 0; % SOCP
+K.r = 0; % Rotated SOCP (obsolete)
+K.p = 0; % Power cone
+K.s = 0; % SDP cone
+K.rank = [];
+K.dualrank = [];
+K.scomplex = [];
+K.xcomplex = [];
+
+KCut.f = [];
+KCut.l = [];
+KCut.q = [];
+KCut.r = [];
+KCut.p = [];
+KCut.s = [];
+
+top = 1;
+localtop = 1;
+
+% In the first part of the code, we will work with a transposed version of
+% the vectorized constraints, i.e. constraints are added by adding columns,
+% wheras the final output will be transposed
+
+% Linear equality constraints
+for i = 1:length(equ_con)
+    constraints = equ_con(i);
[…]

### 2017-09-17 code/Yalmip/yalmip/extras/@lmi/plot.m

@@ -1,0 +1,342 @@
+function varargout = plot(varargin)
+%PLOT  Plots the feasible region of a set of constraints
+%
+% p = plot(C,x,c,n,options)
+%
+% Note that only convex sets are allowed, or union of convex sets
+% represented using binary variables (either defined explictly or
+% introduced by YALMIP when modelling, e.g., mixed integer linear
+% programming represetable operators)
+%
+% C:  SET object
+% x:  Plot variables [At most three variables]
+% c:  color [double] ([r g b] format) or char from 'rymcgbk'
+% n:  #vertices [double ] (default 50 in 2D and 200 otherwise)
+% options: options structure from sdpsettings
+
+% Author Johan Löfberg
+% $Id: plot.m,v 1.18 2010-03-31 14:27:39 joloef Exp $
+
+% Get the onstraints
+if nargin<1
+    return
+end
+
+F = varargin{1};
+
+if length(F)==0
+    return;
+end
+
+if nargin < 5
+    opts = sdpsettings('verbose',0);
+else
+    opts = varargin{5};
+    if isempty(opts)
+        opts = sdpsettings('verbose',0);
+    end
+end
+opts.verbose = max(opts.verbose-1,0);
+
+if any(is(F,'uncertain'))
+    F = robustify(F,[],opts);
+end
+
+if nargin < 3
+    color=['rymcgbk']';
+else
+    color = varargin{3};
+    if isa(color,'sdpvar')
+        error('The variables should be specified in the second argument.');
+    end
+    color = color(:)';
+    if isempty(color)
+        color=['rymcgbk']';
+    end
+end
+
+% Plot onto this projection (at most in 3D)
+if nargin < 2
+    x = [];
+else
+    x = varargin{2};
+    if ~isempty(x)
+        if ~isa(x,'sdpvar')
+            error('Second argument should be empty or an SDPVAR');
+        end
+        x = x(:);
+        x = x(1:min(3,length(x)));
+    end    
+end
+
+if isempty(F)
+    return
+end
+
+if any(is(F,'sos'))
+    % Use image representation, feels safer (I'm not sure about the logic
+    % in the code at the moment. Can the dualization mess up something
+    % otherwise...)
+    if ~(opts.sos.model == 1)
+        opts.sos.model = 2;
+    end
+    % Assume the variables we are plotting are the parametric
+    F = compilesos(F,[],opts,x);
+end
+
[…]

### 2017-09-17 code/Yalmip/yalmip/extras/@optimizer/optimizer.m

@@ -1,0 +1,272 @@
+function sys = optimizer(Constraints,Objective,options,x,u)
+%OPTIMIZER  Container for optimization problem
+%
+%   OPT = OPTIMIZER(Constraints,Objective,options,x,u) exports an object that
+%   contains precompiled numerical data to be solved for varying arguments
+%   x, returning the optimal value of the expression u.
+%
+%   OPTIMIZER typically only works efficiently if the varying data x enters
+%   the optmization problem affinely. If not, the precompiled problems will
+%   be nonconvex, despite the problem being simple for a fixed value of the
+%   parameter (see Wiki for beta support of a much more general optimizer)
+%
+%   In principle, if an optimization problem has a parameter x, and we
+%   repeatedly want to solve the problem for varying x to compute a
+%   variable u, we can, instead of repeatedly constructing optimization
+%   problems for fixed values of x, introduce a symbolic x, and then
+%   simply add an equality 
+%    solvesdp([Constraints,x == value],Objective);
+%    uopt = double(u)
+%   There will still be overhead from the SOLVESDP call, so we can
+%   precompile the whole structure, and let YALMIP handle the addition of
+%   the equality constraint for the fixed value, and automatically extract
+%   the solution variables we are interested in
+%    OPT = optimizer(Constraints,Objective,options,x,u)
+%    uopt1 = OPT{value1}
+%    uopt2 = OPT{value2}
+%    uopt3 = OPT{value3}
+%
+%   By default, display is turned off (since optimizer is used in
+%   situations where many problems are solved repeatedly. To turn on
+%   display, set the verbose option in sdpsetting to 2.
+%
+%   Example
+%
+%    The following problem creates an LP with varying upper and lower
+%    bounds on the decision variable.
+%
+%    The optimizing argument is obtained by indexing (with {}) the optimizer 
+%    object with the point of interest. The argument should be a column
+%    vector (if the argument has a width larger than 1, YALMIP assumes that
[…]

### 2017-09-17 code/Yalmip/yalmip/extras/@optimizer/subsref.m

@@ -1,0 +1,170 @@
+function varargout = subsref(self,subs)
+
+if isequal(subs.type,'()')
+    
+    if size(self.diminOrig,1) > 1
+        error('Can index in cell-based OPTIMIZER. Perhaps you mean {}');
+    end
+    if isempty(self.output.z)
+        output_length = length(self.output.expression);
+    else
+        output_length = length(self.map);
+    end
+    if ~isequal(subs.subs{1},round(subs.subs{1})) | min(subs.subs{1})<1 |   max(subs.subs{1}) > output_length
+        error('Beware of syntax change in optimizer. {} is now used to obtained solution ??? Subscript indices must either be real positive integers or logicals.')
+    end
+    
+    % Create a new function with extracted outputs
+    if isempty(self.output.z)
+        self.map = self.map(subs.subs{1});
+        self.output.expression = self.output.expression(subs.subs{1});
+    else
+     %   self.map = self.map(subs.subs{1});
+        self.output.expression = self.output.expression(subs.subs{1});
+    end
+    self.dimoutOrig = size(self.output.expression);
+    self.dimout = [prod(size(self.output.expression)) 1];
+    varargout{1} = self;
+    
+elseif isequal(subs.type,'.')
+    
+    switch subs.subs
+        case 'model'
+            varargout{1} = self.model;
+        otherwise
+            error('Field not accesible')
+    end
+    
+elseif isequal(subs.type,'{}')
+
+    % This is not really supported yet...
+    if isa(subs.subs{1},'sdpvar')
+        varargout{1} = yalmip('definemulti','optimizer_operator',subs(1).subs{1},self,self.dimout);
+        return
+    end
+        
+    if size(self.diminOrig,1) > 1
+        % Concatenate inputs?
+        if isa(subs.subs{1},'cell')
+            if length(subs.subs{1})~=size(self.diminOrig,1)
+                error('The number of cell elements in input does not match OPTIMIZER declaration');
+            end
+            vecIn = [];
+            for i = 1:length(subs.subs{1})
+                arg = subs.subs{1}{i};
+                if ~isequal(size(arg),self.diminOrig(i,:))
+                    error('Size mismatch on argument in cell');
+                end
+                temp = arg(:);
+                temp = temp(self.mask{i});
+                vecIn = [vecIn;temp];
+            end    
+            subs.subs{1} = vecIn;
+        else
+            error('OPTIMIZER defined with cell input format, but you sent a vector');
+        end
+    else
[…]

### 2017-09-17 code/Yalmip/yalmip/extras/@optproblem/kkt.m

@@ -1,0 +1,12 @@
+function [KKTConstraints, details] = kkt(P,x,ops)
+%KKT Create KKT system for optimization system P with parametric variables x
+%
+% [KKTConstraints, details] = kkt(P,x)
+
+if nargin < 2
+    [KKTConstraints, details] = kkt(P.Constraints,P.Objective);
+elseif nargin < 3
+    [KKTConstraints, details] = kkt(P.Constraints,P.Objective,x);
+else
+    [KKTConstraints, details] = kkt(P.Constraints,P.Objective,x,ops);
+end

### 2017-09-17 code/Yalmip/yalmip/extras/InstantiateElementWise.m

@@ -1,0 +1,29 @@
+function y = InstantiateElementWise(funName,varargin);
+
+if length(varargin{1}) == 1
+    y = yalmip('define',funName,varargin{:});
+else
+    y = [];
+    args = varargin{1};
+    dims = size(args);
+    args = args(:);
+    Base = getbase(args);
+    iD = ~any(Base(:,2:end),2);
+    iS = any(Base(:,2:end),2);
+    if nnz(iD)>0
+        iD = find(iD);
+        varargin{1} = args(iD);
+        yDoubles = feval(funName,varargin{:});    
+    else
+        yDoubles = [];
+        iD = [];
+    end
+    iS = find(iS);
+    args = args(iS);
+    for i = 1:length(args)
+        varargin{1} = args(i);
+        y = [y;yalmip('define',funName,varargin{:})];    
+    end
+    y = sparse([iD;iS],ones(length(iD)+length(iS),1),[yDoubles;y],prod(dims),1);
+    y = reshape(y,dims);
+end

### 2017-09-17 code/Yalmip/yalmip/extras/append_normalized_socp.m

@@ -1,0 +1,40 @@
+function [F_struc,K,c,Q,UB,LB,Qi,Li,ri] = append_normalized_socp(F_struc,K,c,Q,UB,LB)
+
+if K.q(1)>0
+    % To simplify code, we currently normalize everything to z'*z<z0^2
+    % This done by writting [c^Tx+d;Ax+b] in cone as
+    % [c^Tx+d;Ax+b]==[z0;z],  [z0;z] in cone
+    nNew = sum(K.q);
+    nOriginal = length(c);
+    Ftemp = F_struc(1+K.f+K.l:end,:);
+    F_strucSOCP = [Ftemp -speye(nNew)];
+    F_struc = [F_struc(1:K.f+K.l,:) spalloc(K.f+K.l,nNew,0)];
+    F_struc = [F_strucSOCP;F_struc];
+    K.f = K.f + nNew;
+    c = [c;spalloc(nNew,1,0)];
+    Q = blkdiag(Q,spalloc(nNew,nNew,0));
+    UB = [UB;inf(nNew,1)];
+    for i = 1:length(K.q);
+        LB = [LB;0;-inf(K.q(i)-1,1)];
+    end
+    if nargout > 6
+        % Cplex interface needs explicit representations of z'Q*z+Lz+r
+        top = nOriginal+1;
+        ri = zeros(1,length(K.q));
+        Li = spalloc(nOriginal+nNew,length(K.q),0);
+        nTotal = nOriginal+nNew;
+        for i = 1:length(K.q);
+            ind = top:top+K.q(i)-1;
+            vals = [-1 ones(1,K.q(i)-1)];
+            aux = diag(sparse(ind,1,vals,nTotal,1));
+            Qi{i} = aux;
+            top = top + K.q(i);
+        end
+    end
+else
+    Qi = [];
+    ri = [];
+    Li = [];
+end
+
+

### 2017-09-17 code/Yalmip/yalmip/extras/apply_recursive_differentiation.m

@@ -1,0 +1,196 @@
+function dX = apply_recursive_differentiation(model,x,requested,recursivederivativeprecompute);
+
+% Compute all evaluation-based derivatives df(x)
+dxi = [];
+dxj = [];
+dxs = [];
+for i = 1:length(model.evaluation_scheme)
+    if isequal(model.evaluation_scheme{i}.group,'eval')
+        for j = model.evaluation_scheme{i}.variables
+            k = model.evalMap{j}.variableIndex;
+            if any(requested(model.evalMap{j}.computes))
+                derivative = model.evalMap{j}.properties.derivative(x(k));
+                z{i,j} = derivative;
+                if i == 1
+                    dxj = [dxj model.evalMap{j}.variableIndex];
+                    if length(derivative)>1 & length(model.evalMap{j}.computes)==1
+                         dxi = [dxi repmat(model.evalMap{j}.computes,1,length(derivative))]; 
+                    else
+                        dxi = [dxi model.evalMap{j}.computes];
+                    end
+                    dxs = [dxs;derivative(:)];
+                end
+            end
+        end
+    end
+end
+
+% Apply chain-rule. This code is horrible
+ss = model.deppattern(requested,model.linearindicies);
+ss = ss | ss;
+ss = sum(ss,1)>1;
+
+monomTablePattern = sparse(double(model.monomtable | model.monomtable));
+mtT = model.monomtable';
+if 1 % USE NEW
+    [~,dxj] = ismember(dxj,model.linearindicies);
+    dxi = [dxi model.linearindicies];
+    dxj = [dxj 1:length(model.linearindicies)];
+    dxs = [dxs(:)' ones(length(model.linearindicies),1)'];
+    dX = sparse(dxi,dxj,dxs,length(model.c),length(model.linearindicies));
+    % dX = sparse(model.linearindicies,1:length(model.linearindicies),ones(length(model.linearindicies),1),length(model.c),length(model.linearindicies));
+    newMonoms = [];
+    for i = 1:length(model.evaluation_scheme)
+        switch model.evaluation_scheme{i}.group
+            case 'eval'
+                if i>1
+                    for variable = 1:length(model.linearindicies)
+                        if ss(variable)
+                            for j = recursivederivativeprecompute{variable,i}
+                                k = model.evalMap{j}.variableIndex;
+                                if length(model.evalMap{j}.computes) == 1
+                                    dX(model.evalMap{j}.computes,variable) = dX(k,variable)'*z{i,j};
[…]

### 2017-09-17 code/Yalmip/yalmip/extras/compileinterfacedata.m

@@ -1,0 +1,1333 @@
+function [interfacedata,recoverdata,solver,diagnostic,F,Fremoved] = compileinterfacedata(F,aux_obsolete,logdetStruct,h,options,findallsolvers,parametric)
+
+persistent CACHED_SOLVERS
+persistent EXISTTIME
+persistent NCHECKS
+
+%% Initilize default empty outputs
+diagnostic = [];
+interfacedata = [];
+recoverdata = [];
+solver = [];
+Fremoved  = [];
+        
+%% Did we make the call from SOLVEMP
+if nargin<7
+    parametric = 0;
+end
+
+%% Clean objective to default empty
+if isa(h,'double')
+    h = [];
+end
+
+% *************************************************************************
+%% Exit if LOGDET objective is nonlinear
+% *************************************************************************
+if ~isempty(logdetStruct)
+    for i = 1:length(logdetStruct.P)
+        if ~is(logdetStruct.P{i},'linear')
+            diagnostic.solvertime = 0;
+            diagnostic.problem = -2;
+            diagnostic.info = yalmiperror(diagnostic.problem,'');
+            return
+        end
+    end
+end
+
+% *************************************************************************
+%% EXTRACT LOW-RANK DESCRIPTION
+% *************************************************************************
+lowrankdetails = getlrdata(F);
+if ~isempty(lowrankdetails)
+    F = F(~is(F,'lowrank'));
+end
+
+% *************************************************************************
+%% PERTURB STRICT INEQULAITIES
+% *************************************************************************
+if isa(options.shift,'sdpvar') | (options.shift~=0)
+    F = shift(F,options.shift);
+end
+
+% *************************************************************************
+%% ADD RADIUS CONSTRAINT
+% *************************************************************************
+if isa(options.radius,'sdpvar') | ~isinf(options.radius)
+    x = recover(unique(union(depends(h),depends(F))));
+    if length(x)>1
+        F = F + set(cone(x,options.radius));
+    else
+        F = F + set(-options.radius <= x <= options.radius);
+    end
+end
+
+% *************************************************************************
+%% CONVERT LOGIC CONSTRAINTS
+% *************************************************************************
+[F,changed] = convertlogics(F);
+if changed
+    options.saveduals = 0; % Don't calculate duals since we changed the problem
+end
+
[…]

### 2017-09-17 code/Yalmip/yalmip/extras/compress_evaluation_scheme.m

@@ -1,0 +1,27 @@
+function model = compress_evaluation_scheme(model);
+scalars = {'exp','log','sin','cos','log2','log10','slog','inverse_internal2'};
+for i = 1:length(model.evaluation_scheme)
+    if strcmp(model.evaluation_scheme{i}.group,'eval')
+        clear fun
+        for k = 1:length(scalars)
+            for j = 1:length(model.evaluation_scheme{i}.variables)
+                fun{k}(j) = strcmp(model.evalMap{model.evaluation_scheme{i}.variables(j)}.fcn,scalars{k});
+            end
+        end
+        for k = 1:length(scalars)
+            fun_i = find(fun{k});
+            if length(fun_i) > 1
+                all_outputs = [];
+                all_inputs  = [];
+                for j = fun_i
+                    all_outputs = [all_outputs model.evalMap{model.evaluation_scheme{i}.variables(j)}.computes];
+                    all_inputs  = [all_inputs model.evalMap{model.evaluation_scheme{i}.variables(j)}.variableIndex];
+                end
+                model.evalMap{model.evaluation_scheme{i}.variables(fun_i(1))}.computes = all_outputs;
+                model.evalMap{model.evaluation_scheme{i}.variables(fun_i(1))}.variableIndex = all_inputs;
+                model.evaluation_scheme{i}.variables(fun_i(2:end)) = nan;
+            end
+        end
+        model.evaluation_scheme{i}.variables(isnan(model.evaluation_scheme{i}.variables)) = [];
+    end
+end

### 2017-09-17 code/Yalmip/yalmip/extras/convert_sigmonial_to_sdpfun.m

@@ -1,0 +1,321 @@
+function [model,changed] = convert_sigmonial_to_sdpfun(model)
+
+% Always add this dummy struct
+model.high_monom_model = [];
+
+% Assume we don't do anything
+changed = 0;
+
+found_and_converted = [];
+if any(model.variabletype > 3)
+    % Bugger...
+    changed = 1;
+
+    % Find a higher order term
+    sigmonials = find(model.variabletype == 4);
+
+    model = update_monomial_bounds(model);
+
+    monosig = sigmonials(find(sum(model.monomtable(sigmonials,:) | model.monomtable(sigmonials,:),2)==1));
+    if ~isempty(monosig)
+        % These are just monomial terms such as x^0.4 etc
+        for i = 1:length(monosig)
+            variable = find(model.monomtable(monosig(i),:));
+            power = model.monomtable(monosig(i),variable);
+            model = add_sigmonial_eval(model,monosig(i),variable,power);
+            found_and_converted = [found_and_converted;variable power monosig(i)];
+        end
+    end
+end
+if any(model.variabletype > 3)
+    % Bugger...we have mixed terms such as x/y etc
+
+    % Find a higher order term
+    sigmonials = find(model.variabletype == 4);
+
+    for i = 1:length(sigmonials)
+        n_old_monoms = size(model.monomtable,1);
+        monoms = model.monomtable(sigmonials(i),:);
+
+        % Which variables have fractional or negative powers
+        sigs = find((monoms ~= fix(monoms)) | monoms<0);
+        powers =  monoms(sigs);
+        if ~isempty(found_and_converted)
+            % Maybe some of the terms have already been defined as new
+            % variables
+            for j = 1:length(sigs)
+                old_index = findrows(found_and_converted(:,1:2),[sigs(j) powers(j)]);
+                if ~isempty(old_index)
+                    corresponding_variable = found_and_converted(old_index,3);
+                    model.monomtable(sigmonials(i),sigs(j)) = 0;
[…]

### 2017-09-17 code/Yalmip/yalmip/extras/convertquadratics.m

@@ -1,0 +1,101 @@
+function [Fconv,no_changed,infeasible] = convertquadratics(F)
+%CONVERTQUADRATICS Internal function to extract quadratic constraints
+
+% Author Johan Löfberg
+% $Id: convertquadratics.m,v 1.17 2009-05-15 10:32:42 joloef Exp $
+
+% ******************************
+% LINEAR?
+% ******************************
+
+infeasible = 0;
+itslinear = islinear(F);
+if itslinear
+    Fconv = F;
+    no_changed = 0;
+    return
+end
+
+itssigmonial = issigmonial(F);
+if itssigmonial
+    Fconv = F;
+    no_changed = 0;
+    return
+end
+
+Fconv = lmi;
+no_changed = 0;
+for i = 1:1:length(F)
+    if is(F(i),'element-wise') & ~is(F(i),'linear') & ~is(F(i),'sigmonial')
+        % f-c'*x-x'*Q*x>0
+        fi = sdpvar(F(i));fi = fi(:);
+        %[Qs,cs,fs,x,info] = vecquaddecomp(fi);
+        for j = 1:length(fi)
+            fij = fi(j);
+            if isa(fij,'double')
+                if fij < 0
+                    infeasible = 1;                    
+                    warning('The problem is trivially infeasible. You have a term POSITIVE NUMBER < 0.');
+                    return
+                end
+            end
+           % Q = Qs{j};c = cs{j};f = fs{j};
+            [Q,c,f,x,info] = quaddecomp(fij);
+            if info==0
+                if nnz(Q)==0
+                    % Oh, linear,...
+                    Fconv = Fconv + set(fi(j));
+                else
+                    % Yes, quadratic, but convex?
+                    % Change sign definitions
+                    Q = -Q;
+                    c = -c;
+                    f = -f;
+                    % Semi-definite case when only part of x in Q
+                    % Occurs, e.g, in constraints like y'*Q*y < t
+                    used = find(any(Q));Qred=Q(:,used);Qred = Qred(used,:);xred = x(used);
+                    [R,p]=chol(Qred);
+                    if p
+                        % Safety check to account for low rank problems
+%                        try
+                        if all(eig(full(Qred))>=-1e-12)
+                            [u,s,v]=svd(full(Qred));
+                            r=find(diag(s)>1e-12);
+                            R=(u(:,r)*sqrt(s(r,r)))';
+                            p=0;
+                        end
+ %                       catch
+ %                           1
+ %                       end
+                    end
[…]

### 2017-09-17 code/Yalmip/yalmip/extras/dualize.m

@@ -1,0 +1,754 @@
+function [Fdual,objdual,X,t,err] = dualize(F,obj,auto,extlp,extend)
+% DUALIZE Create the dual of an SDP given in primal form
+%
+% [Fd,objd,X,t,err] = dualize(F,obj,auto)
+%
+% Input
+%  F    : Primal constraint in form AX=b+dt, X>0, t free.
+%  obj  : Primal cost CX+ct
+%  auto : If set to 0, YALMIP will not automatically handle variables
+%        and update variable values when the dual problem is solved.
+%  extlp: If set to 0, YALMIP will not try to perform variables changes in
+%         order to convert simple translated LP cones (as in x>1) to
+%         standard unit cone constraints (x>0)
+%
+% Output
+%  Fd  : Dual constraints in form C-A'y>0, c-dy==0
+%  obj : Dual cost b'y (to be MAXIMIZED!)
+%  X   : The detected primal cone variables
+%  t   : The detected primal free variables
+%  err : Error status (returns 0 if no problems)
+%
+% Example
+%  See the HTML help.
+%
+% See also DUAL, SOLVESDP, PRIMALIZE
+
+% Author Johan Löfberg
+% $Id: dualize.m,v 1.62 2007-05-15 12:06:03 joloef Exp $
+
+% Check for unsupported problems
+
+if isempty(F)
+    F = set([]);
+end
+
+LogDetTerm = 0;
+if nargin < 2
+    obj = [];
+elseif isa(obj,'logdet')
+    Plogdet  = getP(obj);
+    gainlogdet = getgain(obj);
+    if ~all(gainlogdet <= 0)
+        error('There are nonconvex logdet terms in the objective')
+    end
+    if ~all(gainlogdet == -1)
+        error('DUALIZE does currently not support coefficients before LOGDET terms')
+    end
+    obj = getcx(obj);
+    if ~isempty(obj)
+        if ~is(obj,'linear')
+            error('DUALIZE does not support nonlinear terms in objective (except logdet terms)')
+        end
+    end
+    LogDetTerm = 1;
+    Ftemp = set([]);
+    for i = 1:length(Plogdet)
+        Ftemp = Ftemp + set(Plogdet{i} >= 0,'LOGDET');
+    end
+    F = Ftemp + F;
+end
+
+err = 0;
+p1 = ~(isreal(F) & isreal(obj));
+p2 = ~(islinear(F) & islinear(obj));
[…]

### 2017-09-17 code/Yalmip/yalmip/extras/eliminatevariables.m

@@ -1,0 +1,244 @@
+function [model,keptvariables,infeasible] = eliminatevariables(model,varindex,value)
+
+keptvariables = 1:length(model.c);
+
+% The first set of equalities are just used when the model is linear in the
+% parametric variables, hence we delete them 
+model.F_struc(1:length(varindex),:) = [];
+model.K.f = model.K.f - length(varindex);
+
+newmonomtable = model.monomtable;
+rmvmonoms = newmonomtable(:,varindex);
+newmonomtable(:,varindex) = 0;
+
+ss = repmat(value(:)',size(rmvmonoms,1),1);
+aux = ss.^rmvmonoms;
+monomvalue = prod(aux,2);
+
+removethese = find(~any(newmonomtable,2));
+keepingthese = find(any(newmonomtable,2));
+
+value = monomvalue(removethese);
+monomgain = monomvalue;monomgain(removethese) = [];
+
+if ~isempty(model.F_struc)
+    model.F_struc(:,1) = model.F_struc(:,1)+model.F_struc(:,1+removethese)*value;
+    model.F_struc(:,1+removethese) = [];
+   % model.F_struc = model.F_struc*sparse(diag([1;monomgain]));
+    model.F_struc = model.F_struc*diag(sparse([1;monomgain]));
+end
+infeasible = 0;
+if model.K.f > 0
+    candidates = find(sum(abs(model.F_struc(1:model.K.f,2:end)),2) == 0);
+    if ~isempty(candidates)
+        infeasibles = find(model.F_struc(candidates,1)~=0);
+        if ~isempty(infeasibles)
+            infeasible = 1;
+            return
+        else
+            model.F_struc(candidates,:) = [];
+            model.K.f = model.K.f - length(candidates);
+        end
+    end
+end
+if model.K.l > 0
+    candidates = find(sum(abs(model.F_struc(model.K.f + (1:model.K.l),2:end)),2) == 0);
+    if ~isempty(candidates)
+        infeasibles = find(model.F_struc(model.K.f + candidates,1)<0);
+        if ~isempty(infeasibles)
+            infeasible = 1;
+            return
+        else
+            model.F_struc(model.K.f + candidates,:) = [];
+            model.K.l = model.K.l - length(candidates);
+        end
+    end
+end
+if model.K.q(1) > 0
+    removeqs = [];
+    removeRows = [];
+    top = model.K.f + model.K.l + 1;  
+    F_struc = model.F_struc(top:top+sum(model.K.q)-1,:);   
+    top = 1;
[…]

### 2017-09-17 code/Yalmip/yalmip/extras/export.m

@@ -1,0 +1,425 @@
+function [model,recoverdata,diagnostic,interfacedata] = export(varargin)
+%EXPORT  Exports YALMIP problem to solver specific format
+%
+%   [MODEL,RECOVERYMODEL,DIAGNOSTIC,INTERNAL] = EXPORT(F,h,options) is the common command to
+%   export optimization problems of the following kind
+%
+%    min        h
+%    subject to
+%            F >(=) 0
+%
+%
+%   The MODEL is exported in the format defined by the solver chosen
+%   in the options structure, or automatically chosen by YALMIP.
+%
+%   If the solver format not is support by EXPORT,the YALMIP model used to
+%   call the solver is returned)
+%
+%   If YALMIP by some reason failed to generate a model, the DIAGNOSTIC 
+%   variable will be non-empty.
+%
+%   The fourth output is the internal model used by YALMIP to communicate
+%   with the generalized solver interfaces.
+%
+%   The RECOVERYMODEL is used to relate a solution of the exported model
+%   to the original variables in YALMIP.
+
+% Arrrgh, new format with logdet much better, but we have to
+% take care of old code, requires some testing...
+varargin = combatible({varargin{:}});
+nargin = length(varargin);
+% *********************************
+% CHECK INPUT
+% *********************************
+if nargin<1
+    help export
+    return
+else
+    F = varargin{1};
+    % Check for wrong syntax
+    if ~isempty(F) & ~isa(F,'lmi') & ~isa(F,'constraint')
+        error('First argument should be a SET object')
+    end
+
+    if isa(F,'constraint')
+        F = set(F);
+    end
+end
+
+model = [];
+recoverdata = [];
+diagnostic = [];
+interfacedata = [];
+
+if nargin>=2
+    h = varargin{2};
+    if isa(h,'double')
+        h = [];
+    end
+    if ~(isempty(h) | isa(h,'sdpvar') | isa(h,'logdet'))
+        error('Second argument (the objective function h) should be an sdpvar or logdet object (or empty).');
+    end
+    if isa(h,'logdet')
+        logdetStruct.P  = getP(h);
+        logdetStruct.gain  = getgain(h);
+        if any(logdetStruct.gain>0)
+            warning('Perhaps you mean -logdet(P)...')
+            diagnostic.yalmiptime = etime(clock,yalmiptime);
[…]

### 2017-09-17 code/Yalmip/yalmip/extras/fmincon_con.m

@@ -1,0 +1,112 @@
+function [g,geq,dg,dgeq,xevaled] = fmincon_con(x,model,xevaled)
+
+global latest_xevaled
+global latest_x_xevaled
+% Early bail for linear problems
+g = [];
+geq = [];
+dg = [];
+dgeq = [];
+if model.linearconstraints
+    xevaled = [];
+    return
+end
+
+if nargin<3
+    if isequal(x,latest_x_xevaled)
+        xevaled = latest_xevaled;
+    else
+        xevaled = zeros(1,length(model.c));
+        xevaled(model.linearindicies) = x;
+        xevaled = apply_recursive_evaluation(model,xevaled);
+        latest_x_xevaled = x;
+        latest_xevaled = xevaled;
+    end
+end
+
+if model.nonlinearinequalities
+    g = full(model.Anonlinineq*xevaled(:)-model.bnonlinineq);
+end
+
+if model.nonlinearequalities
+    geq = full(model.Anonlineq*xevaled(:)-model.bnonlineq);
+end
+
+dgAll_test = [];
+
+if nargout == 2 || ~model.derivative_available
+    return
+elseif ~isempty(dgAll_test) & isempty(model.evalMap)
+    dgAll = dgAll_test;
+elseif isempty(model.evalMap) & (model.nonlinearinequalities | model.nonlinearequalities)
+   % allA = [model.Anonlineq;model.Anonlinineq];
+   % dgAll = [];
+    n = length(model.c);
+    linearindicies = model.linearindicies;
+    mtNonlinear = model.monomtable(model.nonlinearindicies,:);
+    xevaled = zeros(1,n);
+    xevaled(linearindicies) = x;
+    % FIXME: This should be vectorized
+    
+    news = model.fastdiff.news;
+    allDerivemt = model.fastdiff.allDerivemt;
+    c = model.fastdiff.c;
+    
+    X = repmat(x(:)',length(c),1);
+    O = ones(length(c),length(x));    
+    nz = find(allDerivemt);
+    O(nz) = X(nz).^allDerivemt(nz);
+    zzz = c.*prod(O,2);
+   
+    if 1
+        a1 = news(1:length(c),2);
+        a2 = model.nonlinearindicies(news(1:length(c),1))';
+        nn = max(max(length(linearindicies)),max(news(:,2)));
+        mm = max(max(linearindicies),max(model.nonlinearindicies(news(:,1))));
[…]

### 2017-09-17 code/Yalmip/yalmip/extras/hessian.m

@@ -1,0 +1,38 @@
+function d2fdx2 = hessian(f,x)
+% HESSIAN Hessian of scalar polynomial SDPVAR object
+%
+% J = HESSIAN(p)    Hessian w.r.t all variables in p
+% J = HESSIAN(p,x)  Hessian w.r.t the SDPVAR variables x
+%
+% See also SDPVAR, JACOBIAN, LINEARIZE
+
+% Author Johan Löfberg
+% $Id: hessian.m,v 1.2 2004-07-02 08:17:31 johanl Exp $
+
+
+if nargin==1
+    if isa(f,'sdpvar')
+        x = recover(depends(f));
+    else
+        x = 0;
+    end
+else
+    if length(getvariables(x))<length(x)
+        error('x should be a vector of scalar independant variables');
+    end
+end
+
+if prod(size(f))>1
+    error('Hessian only defined for scalars.')
+end
+
+if any(ismember([getvariables(f) depends(f)],yalmip('extvariables')))
+    error('Hessian is only applicable to polynomial expressions');
+end
+
+if isa(f,'double')
+    d2fdx2 = zeros(length(x));
+    return
+end
+
+d2fdx2 = jacobian(jacobian(f,x)',x);

### 2017-09-17 code/Yalmip/yalmip/extras/ipopt_callback_dg.m

@@ -1,0 +1,32 @@
+function G = ipopt_callback_dg(x,model)%returnStructOnly,model)
+
+global latest_x_g
+global latest_G
+global latest_g
+
+% if returnStructOnly
+%     G = sparse(ones(length(model.beq)+length(model.b)+length(model.bnonlineq)+length(model.bnonlinineq),length(model.linearindicies)));
+%     return
+% end
+
+x = x(:);
+if isequal(x,latest_x_g)
+    % Jacobian was computed already in the call for constraints
+    G = latest_G;
+else
+    % Compute the nonlinear terms in the constraints
+    [g,geq,dg,dgeq] = fmincon_con(x,model);
+
+    % Append with linear terms
+    G = [dg';dgeq'];
+    if ~isempty(model.A)
+        G = [G;model.A];
+    end
+    if ~isempty(model.Aeq)
+        G = [G;model.Aeq];
+    end
+    G = sparse(G);
+end
+if model.dense
+    G = full(G);
+end

### 2017-09-17 code/Yalmip/yalmip/extras/sdpsettings.m

@@ -1,0 +1,1655 @@
+function options = sdpsettings(varargin)
+%SDPSETTINGS Create/alter solver options structure.
+%
+%   OPTIONS = SDPSETTINGS with no input arguments returns
+%   setting structure with default values
+%c
+%   OPTIONS = SDPSETTINGS('NAME1',VALUE1,'NAME2',VALUE2,...) creates a
+%   solution options structure OPTIONS in which the named properties have
+%   the specified values.  Any unspecified properties have default values.
+%   It is sufficient to type only the leading characters that uniquely
+%   identify the property.  Case is ignored for property names.
+%
+%   OPTIONS = SDPSETTINGS(OLDOPTS,'NAME1',VALUE1,...) alters an existing options
+%   structure OLDOPTS.
+%
+%
+%   SDPSETTINGS PROPERTIES
+%
+%   GENERAL
+%
+%    solver             - Specify solver [''|sdpt3|sedumi|sdpa|pensdp|penbmi|csdp|dsdp|maxdet|lmilab|cdd|cplex|xpress|mosek|nag|quadprog|linprog|bnb|bmibnb|kypd|mpt|none ('')]
+%    verbose            - Display-level [0|1|2|...(0)] (0-silent, 1-normal, >1-loud)
+%    usex0              - Use the current values obtained from double as initial iterate [0|1 (0)]
+%    dimacs             - Compute DIMACS error measures [0|1(0)]
+%
+%    showprogress       - Show progress of YALMIP (suitable for large problems) [0|1 (0)]
+%    cachesolvers       - Check for available solvers only first time solvesdp is called [0|1 (0)]
+%    warning            - Shows a warning if a problems occurs when solving a problem (infeasibility, numerical problem etc.) [0|1 (1)]
+%    beeponproblem      - Beeps when certain warning/error occurs [ integers -2|-1|1|2|3|4|5|6|7|8|9|10|11]
[…]

### 2017-09-17 code/Yalmip/yalmip/extras/selectsolver.m

@@ -1,0 +1,476 @@
+function [solver,problem,forced_choice] = selectsolver(options,ProblemClass,solvers,socp_are_really_qc);
+%SELECTSOLVER Internal function to select solver based on problem category
+
+% Author Johan Löfberg
+% $Id: selectsolver.m,v 1.18 2008-04-24 11:15:13 joloef Exp $
+
+problem = 0;
+
+% UNDOCUMENTED
+force_solver = yalmip('solver');
+if length(force_solver)>0
+    options.solver = force_solver;
+end
+% ***************************************************
+% Maybe the user is stubborn and wants to pick solver
+% ***************************************************
+forced_choice = 0;
+if length(options.solver)>0 & isempty(findstr(options.solver,'*'))
+    
+    if strfind(options.solver,'+')
+        forced_choice = 1;
+        options.solver = strrep(options.solver,'+','');
+    end
+    % Create tags with version also        
+    temp = solvers;
+    for i = 1:length(temp)
+        if length(temp(i).version)>0
+            temp(i).tag = lower([temp(i).tag '-' temp(i).version]);
+        end
+    end
+    
+    opsolver = lower(options.solver);
+    splits = findstr(opsolver,',');
+    if isempty(splits)
+        names{1} = opsolver;
+    else
+        start = 1;
+        for i  = 1:length(splits)
+            names{i} = opsolver(start:splits(i)-1);
+            start = splits(i)+1;
+        end
+        names{end+1} = opsolver(start:end);
+    end
+        
+    index1 = [];
+    index2 = [];
+    for i = 1:length(names)
+        index1 = [index1 find(strcmp(lower({solvers.tag}),names{i}))];
+        index2 = [index1 find(strcmp(lower({temp.tag}),names{i}))];
+    end
+    if isempty(index1) & isempty(index2)
+        solver = [];
+        problem = -3;
+        return;
+    else
+        solvers = solvers(union(index1,index2));
+    end    
+end
+
+% if forced_choice
+%     solver = solvers(end);
+%     problem = 0;
+%     return
+% end
+
+% ************************************************
+% Prune based on objective
+% ************************************************
+if ProblemClass.objective.sigmonial & ~forced_choice
+    keep = ones(length(solvers),1);
+    for i = 1:length(solvers)
+        keep(i) = solvers(i).objective.sigmonial;                         
+    end
+    solvers = solvers(find(keep));
+end    
[…]

### 2017-09-17 code/Yalmip/yalmip/extras/setup_fmincon_params.m

@@ -1,0 +1,288 @@
+function model = setup_fmincon_params(model)
+
+monomtable = model.monomtable;
+nonlinearindicies = model.nonlinearindicies;
+linearindicies = model.linearindicies;
+
+model.evalinobjective  = 0;
+
+if ~isempty(model.evalMap)
+    [i,j,k] = find(model.monomtable(:,model.evalVariables));
+    evalInvolved = union(model.evalVariables,i);
+    model.evalinobjective = any(model.c(evalInvolved)) | any(model.Q(evalInvolved,1));
+    if size(model.Anonlineq,1)>0
+        if nnz(model.Anonlineq(:,evalInvolved))>0
+            model.evalinconstraint = 1;
+        end
+    end
+    % Speed up callbacks by removing last marker argument (only used in
+    % expandmodel etc) and figuring out what R^m -> R^n operators computes
+    for i = 1:length(model.evalMap)
+        %model.evalMap{i}.prearg = {model.evalMap{i}.fcn,[],model.evalMap{i}.arg{2:end-1}};
+        model.evalMap{i}.prearg = {model.evalMap{i}.fcn,model.evalMap{i}.arg{1:end-1}};
+        model.evalMap{i}.prearg{1+model.evalMap{i}.argumentIndex} = [];
+        if isfield(model.evalMap{i},'computes')
+%             temp = zeros(1,length(model.evalMap{i}.computes));
+%             for j = 1:length(model.evalMap{i}.computes)
+%                 temp(j) = find(model.evalMap{i}.computes(j) == model.used_variables);
+%             end
+%             model.evalMap{i}.computes = temp;
+        else
+            model.evalMap{i}.computes = model.evalVariables(i);
+        end
+    end
+end
+
+% Figure out if YALMIP easily can compute the gradient of the objective
+% This will done completely general later
+model.SimpleLinearObjective = 0;    % Linear
+model.SimpleQuadraticObjective = 0; % Quadratic
+model.SimpleNonlinearObjective = 1; % Polynomial
+if isempty(model.evalMap)
+    if nnz(model.c(nonlinearindicies)) == 0
+        if (nnz(model.Q)==0)
+            model.SimpleLinearObjective = 1;
+        else
+            if nnz(model.Q(nonlinearindicies,nonlinearindicies))==0
+                model.SimpleQuadraticObjective = 1;
+            end
+        end
+    end
+elseif ~model.evalinobjective
[…]

### 2017-09-17 code/Yalmip/yalmip/extras/yalmip.m

@@ -1,0 +1,1052 @@
+function  varargout = yalmip(varargin)
+%YALMIP Returns various information about YALMIP
+%
+%   YALMIP can be used to check version numbers and
+%   find the SDPVAR and SET objects available in workspace
+%
+%   EXAMPLES
+%    V = YALMIP('version')     % Returns version
+%    YALMIP('nvars')           % Returns total number of declared variables
+%    YALMIP('info')            % Display basic info.
+%    YALMIP('solver','tag')    % Sets the solver 'solvertag' (see sdpsettings) as default solver
+%
+%
+% If you want information on how to use YALMIP, you are advised to check out
+% http://users.isy.liu.se/johanl/yalmip/
+%
+%   See also YALMIPTEST, YALMIPDEMO
+
+% Author Johan Löfberg
+
+persistent prefered_solver internal_sdpvarstate internal_setstate
+
+if nargin==0
+    help yalmip
+    return
+end
+
+if isempty(internal_sdpvarstate)
+    internal_sdpvarstate.monomtable = spalloc(0,0,0);   % Polynomial powers table
+    internal_sdpvarstate.hashedmonomtable = [];         % Hashed polynomial powers table
+    internal_sdpvarstate.hash = [];
+    internal_sdpvarstate.boundlist = [];
+    internal_sdpvarstate.variabletype = spalloc(0,0,0); % Pre-calc linear/quadratic/polynomial/sigmonial
+    internal_sdpvarstate.intVariables = [];   % ID of integer variables
+    internal_sdpvarstate.binVariables = [];   % ID of binary variables
+    internal_sdpvarstate.semicontVariables = [];
+    internal_sdpvarstate.uncVariables = [];   % ID of uncertain variables (not used)
+    internal_sdpvarstate.parVariables = [];   % ID of parametric variables (not used)
+    internal_sdpvarstate.extVariables = [];   % ID of extended variables (for max,min,norm,sin, etc)
+    internal_sdpvarstate.auxVariables = [];   % ID of auxilliary variables (introduced when modelling extended variables)
+    internal_sdpvarstate.auxVariablesW = [];   % ID of uncertain auxilliary variables (introduced when modelling uncertain extended variables)
+    internal_sdpvarstate.logicVariables = []; % ID of extended logic variables (for or, nnz, alldifferent etc)
+    internal_sdpvarstate.complexpair = [];
+    internal_sdpvarstate.internalconstraints = [];
[…]

### 2017-09-17 code/Yalmip/yalmip/extras/yalmip2nonlinearsolver.m

@@ -1,0 +1,236 @@
+function model = yalmip2nonlinearsolver(model)
+
+model.dense = 0;
+if ~model.equalitypresolved
+    model = propagate_bounds_from_equalities(model);
+end
+
+K = model.K;
+lb = model.lb;
+ub = model.ub;
+x0 = model.x0;
+c = model.c;
+
+if isempty(model.evaluation_scheme)
+    model = build_recursive_scheme(model);
+end
+model = compress_evaluation_scheme(model);
+
+% Do some pre-calc to be used in calls from fmincon
+nonlinearindicies = union(find(model.variabletype~=0),model.evalVariables);
+linearindicies    = setdiff(find(model.variabletype==0),nonlinearindicies);
+model.nonlinearindicies = nonlinearindicies;
+model.linearindicies    = linearindicies;
+
+model.Anonlinineq = [];
+model.bnonlinineq = [];
+model.Anonlineq = [];
+model.bnonlineq = [];
+
+% Extract linear and nonlinear equality constraints
+if K.f>0
+    Aeq = -model.F_struc(1:1:K.f,2:end);
+    beq = model.F_struc(1:1:model.K.f,1);
+    
+    nonlinear_equalities_indicies = find(any(Aeq(:,nonlinearindicies),2));
+    model.Anonlineq = Aeq(nonlinear_equalities_indicies,:);
+    model.bnonlineq = beq(nonlinear_equalities_indicies);
+    
+    Aeq(nonlinear_equalities_indicies,:) = [];
+    beq(nonlinear_equalities_indicies,:) = [];
+    Aeq(:,nonlinearindicies) = [];
+    model.F_struc(1:model.K.f,:) = [];
+    model.K.f = 0;
+else
+    Aeq = [];
+    beq = [];
+end
+
+% Find nonlinear eualities implied by lower and upper bounds
+if ~isempty(ub) & ~isempty(lb)
+    nonlinearequality = find(lb(nonlinearindicies) == ub(nonlinearindicies));
+    if ~isempty(nonlinearequality)
+        for i = 1:length(nonlinearequality)
+            model.Anonlineq = [model.Anonlineq;eyev(length(c),nonlinearindicies(nonlinearequality(i)))'];
+            model.bnonlineq = [model.bnonlineq;lb(nonlinearindicies(nonlinearequality(i)))];
+        end
+    end
+end
+
+% Extract linear and nonlinear inequality constraints
[…]

### 2017-09-17 code/Yalmip/yalmip/modules/global/bnb.m

@@ -1,0 +1,1414 @@
+function output = bnb(p)
+%BNB          General branch-and-bound scheme for conic programs
+%
+% BNB applies a branch-and-bound scheme to solve mixed integer
+% conic programs (LP, QP, SOCP, SDP) and mixed integer geometric programs.
+%
+% BNB is never called by the user directly, but is called by
+% YALMIP from SOLVESDP, by choosing the solver tag 'bnb' in sdpsettings.
+%
+% BNB is used if no other mixed integer solver is found, and
+% is only useful for very small problems, due to its simple
+% and naive implementation.
+%
+% The behaviour of BNB can be altered using the fields
+% in the field 'bnb' in SDPSETTINGS
+%
+% bnb.branchrule   Deceides on what variable to branch
+%                   'max'     : Variable furthest away from being integer
+%                   'min'     : Variable closest to be being integer
+%                   'first'   : First variable (lowest variable index in YALMIP)
+%                   'last'    : Last variable (highest variable index in YALMIP)
+%                   'weight'  : See manual
+%
+% bnb.method       Branching strategy
+%                   'depth'   : Depth first
+%                   'breadth' : Breadth first
+%                   'best'    : Expand branch with lowest lower bound
+%                   'depthX'  : Depth until integer solution found, then X (e.g 'depthbest')
+%
+% solver           Solver for the relaxed problems (standard solver tag, see SDPSETTINGS)
+%
+% maxiter          Maximum number of nodes explored
+%
+% inttol           Tolerance for declaring a variable as integer
+%
+% feastol          Tolerance for declaring constraints as feasible
+%
+% gaptol           Exit when (upper bound-lower bound)/(1e-3+abs(lower bound)) < gaptol
+%
+% round            Round variables smaller than bnb.inttol
+%
+%
+% See also SOLVESDP, BINVAR, INTVAR, BINARY, INTEGER
+
+% Author Johan Löfberg
+% $Id: bnb.m,v 1.60 2010-03-09 14:51:21 joloef Exp $
+
+% ********************************
+%% INITIALIZE DIAGNOSTICS IN YALMIP
+% ********************************
+bnbsolvertime = clock;
[…]

### 2017-09-17 code/Yalmip/yalmip/modules/global/branch_and_bound.m

@@ -1,0 +1,742 @@
+function [x_min,solved_nodes,lower,upper,lower_hist,upper_hist,timing] = branch_and_bound(p,x_min,upper,timing)
+
+% *************************************************************************
+% Create handles to solvers
+% *************************************************************************
+lowersolver = p.solver.lowersolver.call; % For relaxed lower bound problem
+uppersolver = p.solver.uppersolver.call; % Local nonlinear upper bound
+lpsolver    = p.solver.lpsolver.call;    % LP solver for bound propagation
+
+% *************************************************************************f
+% GLOBAL PROBLEM DATA (these variables are the same in all nodes)
+% *************************************************************************
+c       = p.c;
+Q       = p.Q;
+f       = p.f;
+K       = p.K;
+options = p.options;
+
+% *************************************************************************
+% DEFINE UPPER BOUND PROBLEM. Basically just remove the cuts
+% *************************************************************************
+p_upper = cleanuppermodel(p);
+
+% *************************************************************************
+% Active constraints in main model
+% 0   : Inactive constraint (i.e. a cut which unused)
+% 1   : Active constraint
+% inf : Removed constraint  (found to be redundant)
+% *************************************************************************
+p.InequalityConstraintState = ones(p.K.l,1);
+p.InequalityConstraintState(p.KCut.l,1) = 0;
+p.EqualityConstraintState = ones(p.K.f,1);
+
+% *************************************************************************
+% LPs ARE USED IN  BOX-REDUCTION
+% *************************************************************************
+p.lpcuts = p.F_struc(1+p.K.f:1:p.K.l+p.K.f,:);
+p.cutState = ones(p.K.l,1);
+p.cutState(p.KCut.l,1) = 0; % Don't use to begin with
+
+% *************************************************************************
+% INITIALITAZION
+% *************************************************************************
+p.depth = 0;        % depth in search tree
+p.dpos  = 0;        % used for debugging
+p.lower = NaN;
+lower   = NaN;
+gap     = inf;
+stack   = [];
+solved_nodes = 0;
+numGlobalSolutions = 0;
+
+% *************************************************************************
+% Silly hack to speed up solver calls
+% *************************************************************************
+p.getsolvertime = 0;
+
+if options.bmibnb.verbose>0
+    disp('* Starting YALMIP global branch & bound.');
[…]

### 2017-09-17 code/Yalmip/yalmip/modules/global/cutsdp.m

@@ -1,0 +1,491 @@
+function output = cutsdp(p)
+% CUTSDP
+%
+% See also SOLVESDP, BNB, BINVAR, INTVAR, BINARY, INTEGER, LMI
+
+% Author Johan Löfberg
+% $Id: cutsdp.m,v 1.10 2010-03-09 14:51:21 joloef Exp $
+
+% *************************************************************************
+%% INITIALIZE DIAGNOSTICS IN YALMIP
+% *************************************************************************
+bnbsolvertime = clock;
+showprogress('Cutting plane solver started',p.options.showprogress);
+
+% *************************************************************************
+%% If we want duals, we may not extract bounds. However, bounds must be
+% extracted in discrete problems.
+% *************************************************************************
+if p.options.cutsdp.recoverdual
+    warning('Dual recovery not implemented yet in CUTSDP')
+end
+
+% *************************************************************************
+%% Define infinite bounds
+% *************************************************************************
+if isempty(p.ub)
+    p.ub = repmat(inf,length(p.c),1);
+end
+if isempty(p.lb)
+    p.lb = repmat(-inf,length(p.c),1);
+end
+
+% *************************************************************************
+%% ADD CONSTRAINTS 0<x<1 FOR BINARY
+% *************************************************************************
+if ~isempty(p.binary_variables)
+    p.ub(p.binary_variables) =  min(p.ub(p.binary_variables),1);
+    p.lb(p.binary_variables) =  max(p.lb(p.binary_variables),0);
+end
+
+% *************************************************************************
+%% Extract better bounds from model
+% *************************************************************************
+if ~isempty(p.F_struc)
+    [lb,ub,used_rows_eq,used_rows_lp] = findulb(p.F_struc,p.K);
+    if ~isempty([used_rows_eq(:);used_rows_lp(:)])
+        lower_defined = find(~isinf(lb));
+        if ~isempty(lower_defined)
+            p.lb(lower_defined) = max(p.lb(lower_defined),lb(lower_defined));
+        end
+        upper_defined = find(~isinf(ub));
+        if ~isempty(upper_defined)
+            p.ub(upper_defined) = min(p.ub(upper_defined),ub(upper_defined));
+        end
[…]

### 2017-09-17 code/Yalmip/yalmip/modules/global/global_solve_upper.m

@@ -1,0 +1,61 @@
+function [output,timing] = global_solve_upper(p,p_original,x,options,uppersolver,timing)
+
+if ~isempty(p.binary_variables)
+    local_gave_good = find(abs(x(p_original.binary_variables)-fix(x(p_original.binary_variables)))< options.bnb.inttol);
+    p.lb(p.binary_variables(local_gave_good)) = fix(x(p.binary_variables(local_gave_good)));
+    p.ub(p.binary_variables(local_gave_good)) = fix(x(p.binary_variables(local_gave_good)));  
+    for i = 1:3
+        p = update_eval_bounds(p);
+        p = propagate_bounds_from_equalities(p);  
+        p = update_monomial_bounds(p);
+    end
+end
+
+% The bounds and relaxed solutions have been computed w.r.t to the relaxed
+% bilinear model. We only need the original bounds and variables.
+p.lb = p.lb(1:length(p_original.c));
+p.ub = p.ub(1:length(p_original.c));
+x = x(1:length(p_original.c));
+
+ if ~isempty(p_original.integer_variables)
+     % disp('Report bug: FIX ME in global_solve_upper at 16')
+     local_gave_good = find(abs(x(p.integer_variables)-fix(x(p.integer_variables)))< options.bnb.inttol);
+     p.lb((p.integer_variables(local_gave_good))) = fix(x(p.integer_variables(local_gave_good)));
+     p.ub((p.integer_variables(local_gave_good))) = fix(x(p.integer_variables(local_gave_good)));
+ end
+%         
+p_upper = p_original;
+
+p_upper.lb = p.lb;
+p_upper.ub = p.ub;
+
+% Pick an initial point (this can be a bit tricky...)
+lbinfbounds = find(isinf(p.lb));
+ubinfbounds = find(isinf(p.ub));
+switch p.options.bmibnb.localstart
+    case 'relaxed'
+        p_upper.x0 = x;
+    case 'boxcenter'
+        p_upper.x0 = x;        
+        p_upper.x0(lbinfbounds)=0;
+        p_upper.x0(ubinfbounds)=0;
+    case 'zero'       
+        p_upper.x0 = zeros(length(p.lb),1);
+    otherwise
+        error('Unknown local starting point option');
+end
+
+% Save time
+p_upper.options.saveduals = 0;
+
+% Solve upper bounding problem
+p_upper.options.usex0 = 1;
+tstart = tic;
[…]

### 2017-09-17 code/Yalmip/yalmip/modules/global/propagate_bounds_from_equalities.m

@@ -1,0 +1,373 @@
+function p = propagate_bounds_from_equalities(p)
+
+% % Simple extraction first
+% if p.K.f > 0
+%     [p.lb,p.ub] = remove_bounds_from_Aeqbeq(-p.F_struc(1:p.K.f,2:end),p.F_struc(1:p.K.f,1),p.lb,p.ub);
+% end
+
+%p.equalitypresolved = 1;
+LU = [p.lb p.ub];
+
+p_F_struc = p.F_struc;
+n_p_F_struc_cols = size(p_F_struc,2);
+
+fixedVars = find(p.lb == p.ub & p.variabletype(:) == 0);
+if ~isempty(fixedVars)
+    p_F_struc_forbilin = p_F_struc;
+    p_F_struc_forbilin(:,1) = p_F_struc(:,1) + p_F_struc(:,1+fixedVars)*p.lb(fixedVars);
+    p_F_struc_forbilin(:,1+fixedVars) = 0;
+else
+    p_F_struc_forbilin=p_F_struc;
+end
+
+if p.K.f >0
+    
+    for j = 1:p.K.f
+        thisrow = p_F_struc(j,:);
+        if thisrow(1)<0
+            thisrow = -thisrow;
+        end
+        if thisrow(1)>0
+            [row,col,val] = find(thisrow);
+            % Find bounds from sum(xi) = 1, xi>0
+            if all(val(2:end) < 0)
+                usedVars = col(2:end)-1;
+                if all(p.lb(usedVars)>=0)
+                    p.ub(usedVars) = min( p.ub(usedVars) , val(1)./abs(val(2:end)'));
+                end
+            end            
+        end
+    end
+    
+    % Presolve from bilinear x*y == k
+    for j = 1:p.K.f
+        if p_F_struc_forbilin(j,1)~=0
+            [row,col,val] = find(p_F_struc_forbilin(j,:));
+            
+            % Find bounds from sum(xi) = 1, xi>0
+            if length(col)==2
+                val = val/val(2); % val(1) + x==0
+                var = col(2)-1;
+                if p.variabletype(var)==1
+                    [ij] = find(p.monomtable(var,:));
+                    if p.lb(ij(1))>=0 & p.lb(ij(2))>=0
+                        % xi*xj == val(1)
+                        if -val(1)<0
+                            p.feasible = 0;
[…]

### 2017-09-17 code/Yalmip/yalmip/modules/global/solvelower.m

@@ -1,0 +1,239 @@
+function [output,cost,psave,timing] = solvelower(p,options,lowersolver,xmin,upper,timing)
+
+psave = p;
+removeThese = find(p.InequalityConstraintState==inf);
+p.F_struc(p.K.f + removeThese,:) = [];
+p.K.l = p.K.l - length(removeThese);
+
+removeThese = find(p.EqualityConstraintState==inf);
+p.F_struc(removeThese,:) = [];
+p.K.f = p.K.f - length(removeThese);
+
+if p.options.bmibnb.cut.bilinear
+    p_cut = addBilinearVariableCuts(p);
+end
+if p.options.bmibnb.cut.evalvariable
+    p_cut = addEvalVariableCuts(p_cut);
+    psave.evalMap = p_cut.evalMap;
+end
+if p.options.bmibnb.cut.monomial
+    p_cut = addMonomialCuts(p_cut);
+end
+if p.options.bmibnb.cut.multipliedequality
+    p_cut = addMultipliedEqualityCuts(p_cut);
+end
+
+% **************************************
+% SOLVE NODE PROBLEM
+% **************************************
+if any(p_cut.ub+1e-8<p_cut.lb)
+    output.problem=1;
+    cost = inf;
+else
+    % We are solving relaxed problem (penbmi might be local solver)
+    p_cut.monomtable = eye(length(p_cut.c));
+    
+    if p.solver.lowersolver.objective.quadratic.convex
+        % Setup quadratic
+        for i = 1:size(p.bilinears,1)
+            if p_cut.c(p.bilinears(i,1))
+                p_cut.Q(p.bilinears(i,2),p.bilinears(i,3)) = p_cut.c(p.bilinears(i,1))/2;
+                p_cut.Q(p.bilinears(i,3),p.bilinears(i,2)) = p_cut.Q(p.bilinears(i,3),p.bilinears(i,2))+p_cut.c(p.bilinears(i,1))/2;
+                p_cut.c(p.bilinears(i,1)) = 0;
+            end
+        end
+        
+        if ~all(eig(full(p_cut.Q))>-1e-12)
+            p_cut.Q = p.Q;
+            p_cut.c = p.c;
+        end
+    end
+    
+    fixed = p_cut.lb >= p_cut.ub;
+    if nnz(fixed) == length(p.c)
+        % All variables are fixed to a bound
+        output.Primal = p.lb;
+        res = constraint_residuals(p,output.Primal);
[…]

### 2017-09-17 code/Yalmip/yalmip/modules/global/update_integer_bounds.m

@@ -1,0 +1,16 @@
+function model = update_integer_bounds(model);
+
+if ~isempty(model.integer_variables)
+    model.lb(model.integer_variables) = fix(model.lb(model.integer_variables)-1e-4);
+    model.ub(model.integer_variables) = fix(model.ub(model.integer_variables)+1e-4);
+end
+if ~isempty(model.binary_variables)
+    model.lb(model.binary_variables) = fix(model.lb(model.binary_variables)-1e-4);
+    model.ub(model.binary_variables) = fix(model.ub(model.binary_variables)+1e-4);
+end
+if any(model.lb(model.binary_variables) > model.ub(model.binary_variables))
+    model.feasible = 0;
+end
+if any(model.lb(model.integer_variables) > model.ub(model.integer_variables))
+    model.feasible = 0;
+end

### 2017-09-17 code/Yalmip/yalmip/modules/global/update_monomial_bounds.m

@@ -1,0 +1,19 @@
+function model = update_monomial_bounds(model,these);
+
+if nargin == 1
+    polynomials = find((model.variabletype ~= 0));
+else
+    polynomials = find((model.variabletype ~= 0));
+    polynomials = polynomials(find(ismember(polynomials,these)));
+end
+
+for i = 1:length(polynomials)
+    j = polynomials(i);
+    if j<=length(model.lb)
+        bound = powerbound(model.lb,model.ub,model.monomtable(j,:));
+        model.lb(j) = max(model.lb(j),bound(1));
+        model.ub(j) = min(model.ub(j),bound(2));
+    end
+end
+
+

### 2017-09-17 code/Yalmip/yalmip/modules/sos/compilesos.m

@@ -1,0 +1,1201 @@
+function [F,obj,m] = solvesos(F,obj,options,params,candidateMonomials)
+%COMPILESOS Sum of squares decomposition
+%
+%    [F,obj,m] = compilesos(F,h,options,params,monomials) derives the SOS
+%    problem without actually solving it
+% 
+% See also SOLVESOS
+
+%% Time YALMIP
+yalmip_time = clock;
+
+% ************************************************
+%% Check #inputs
+% ************************************************
+if nargin<5
+    candidateMonomials = [];
+    if nargin<4
+        params = [];
+        if nargin<3
+            options = sdpsettings;
+            if nargin<2
+                obj = [];
+                if nargin<1
+                    help solvesos
+                    return
+                end
+            end
+        end
+    end
+end
+
+if isempty(options)
+    options = sdpsettings;
+end
+
+% Lazy syntax (not official...)
+if nargin==1 & isa(F,'sdpvar')
+    F = set(sos(F));
+end
+
+if ~isempty(options)
+    if options.sos.numblkdg
+        [sol,m,Q,residuals,everything] = solvesos_find_blocks(F,obj,options,params,candidateMonomials);
+        return
+    end
+end
+
+% *************************************************************************
+%% Extract all SOS constraints and candidate monomials
+% *************************************************************************
+if ~any(is(F,'sos'))
+    error('At-least one constraint should be an SOS constraints!');
+end
+p = [];
+ranks = [];
+for i = 1:length(F)
+    if is(F(i),'sos')
+        pi = sdpvar(F(i));
+        p{end+1} = pi;
+        ranks(end+1) = getsosrank(pi); % Desired rank : Experimental code
+    end
+end
+if isempty(candidateMonomials)
+    for i = 1:length(F)
+        candidateMonomials{i}=[];
+    end
+elseif isa(candidateMonomials,'sdpvar')
+    cM=candidateMonomials;
+    candidateMonomials={};
+    for i = 1:length(p)
+        candidateMonomials{i}=cM;
+    end
+elseif isa(candidateMonomials,'cell')
+    if length(p)~=length(candidateMonomials)
+        error('Dimension mismatch between the candidate monomials and the number of SOS constraints');
+    end
+end
+
+% *************************************************************************
+%% Get the parametric constraints
+% *************************************************************************
+F_original = F;
+F_parametric = F(find(~is(F,'sos')));
+if isempty(F_parametric)
+    F_parametric = set([]);
+end
+
+% *************************************************************************
[…]

### 2017-09-17 code/Yalmip/yalmip/modules/sos/solvesos.m

@@ -1,0 +1,1530 @@
+function [sol,m,Q,residuals,everything] = solvesos(F,obj,options,params,candidateMonomials)
+%SOLVESOS Sum of squares decomposition
+%
+%    [sol,m,B,residual] = solvesos(F,h,options,params,monomials) is used
+%    for finding SOS decompositions of polynomials.
+%
+%    The coefficients of the polynomials are assumed linear w.r.t a set of
+%    decision variables params and polynomial with respect to a variable x.
+%    
+%    An extension with a nonlinear parameterization in params is possible.
+%    Note though that this gives BMIs or PMIs, solvable (locally) only if
+%    PENBMI is installed, or by specifying 'moment' as solver to try to
+%    solve the nonconvex semidefinite programming problem using a
+%    semidefinite relaxation based on moments.
+%
+%    The SOS problem can be formulated as
+%
+%              min h(params)
+%
+%       subject to  F(i) >(=) 0 or F(i) is SOS w.r.t x
+%
+%   INPUT
+%    F         : SET object with SOS constrained polynomials and constraints on variables params
+%    h         : scalar SDPVAR object (can be [])
+%    options   : options structure obtained from SDPSETTINGS (can be [])
+%    params    : SDPVAR object defining parametric variables (can be [])
+%    monomials : SDPVAR object with user-specified monomials for decomposition (can be [])
+%
+%   OUTPUT
+%    sol       : Solution diagnostic from SDP problem
+%    v         : Cell with monomials used in decompositions
+%    Q         : Cell with Gram matrices, p = v{i}'*Q{i}*v{i}, where p is the ith SOS polynomial in your model.
+%    residuals : Mismatch between p and decompositions. Same values (modulo numerical issue) as checkset(find(is(F,'sos')))
+%                Warning, these residuals are not computed on matrix sos-of-squares
+%
+%   EXAMPLE
+%    x = sdpvar(1);solvesos(set(sos(x^4+x^3+1)));                    % Simple decompositions
+%    x = sdpvar(1);t = sdpvar(1);solvesos(set(sos(x^4+x^3+1-t)),-t); % Lower bound by maximizing t
+%
+%   NOTES
+%
+%    Variables not part of params, but part of non-SOS constraints in F
+%    or objective h will automatically be appended to the params list.
+%
+%    To extract SOS decomposition, use the command SOSD (or compute from use v and Q)
+%
+%    If the 5th input argument is used, no additional monomial reduction is
+%    performed (Newton, inconstency, congruence). It is thus assumed that
[…]

### 2017-09-17 code/Yalmip/yalmip/operators/implies.m

@@ -1,0 +1,80 @@
+function varargout = implies(varargin)
+%IMPLIES Logical implication
+%
+% IMPLIES(X,Y) creates a mixed integer representation of
+% the constraint X --> Y, i.e. Y is true if X is true.
+%
+% Syntax
+%   F = implies(X,Y,tol)
+%
+% Input
+%   X : binary SDPVAR variable or a constraint
+%   Y : binary SDPVAR variable or a constraint
+%  tol: Optional threshhold for defining zero (see NOTE)
+%
+% Output
+%   F : SET object
+%
+% Examples
+%
+%  binvar X,Y; F = implies(X,Y);
+%  binvar X;sdpvar Y; F = implies(X,Y>=5);
+%  binvar X;Y=sdpvar(3,1); F = set(implies(X,[sum(Y);Y(2)]>=[5;0]));
+%
+% Note
+%  Using implies with X non-binary is highly sensitive numerically.
+%  The problem comes from the definition of 0 in a floating-point
+%  environment, and precision in the solver. To account for this,
+%  the user can supply a third argument to define a dead-zone around
+%  zero, i.e Implies(X<=0,Y) will be replaced with IMPLIES(X<=-tol,Y)
+%  Note, you typically need to tweak this number for your
+%  application/solver. By default, YALMIP uses tol = 0, which means you
+%  easily can get garbage... A positive number means YALMIP is cautious in
+%  terms of activating the condition, while a negative number means YALMIP
+%  will be aggressive  in activating the condition. 
+%
+%   See also @SDPVAR/AND, @SDPVAR/OR, IFF
+
+% Author Johan Löfberg
+% $Id: implies.m,v 1.6 2009-05-15 10:32:42 joloef Exp $
+
+
+% There are some cases to take care of...
+%
+% X --> Y     binary/binary :                     Implemented
+% X --> Y     binary/(LP,equality,sdp)            Implemented
+% X --> Y     (LP,equality,sdp)/binary            Not implemented
+% X --> Y     (LP,equality,sdp)/(LP,equality,sdp) Not implemented
+
+X = varargin{1};
+Y = varargin{2};
+
+% % Call recursicely on X -> (A,B,...)
+% if isa(varargin{1},'sdpvar') & (isa(varargin{2},'lmi') | isa(varargin{2},'constraint'))
+%     if length(varargin{1})==1 & length(varargin{2})>1
+%         F = set([]);
+%         for i = 1:length(varargin{2})
+%             if nargin == 3
[…]

### 2017-09-17 code/Yalmip/yalmip/solvers/call_cplexibm_miqp.m

@@ -1,0 +1,162 @@
+function output = call_cplexibm_miqp(interfacedata)
+
+% Author Johan Löfberg
+
+options = interfacedata.options;
+model = yalmip2cplex(interfacedata);
+
+if interfacedata.options.savedebug
+    save cplexdebug model
+end
+
+solvertime = clock;
+[x,fval,exitflag,output,lambda] = localSolverCall(model);
+if output.cplexstatus == 4 | output.cplexstatus == 119
+    % CPLEX reports infeasible OR unbounded
+    % Remove objective and resolve
+    model.H = model.H*0;
+    model.f = model.f*0;
+    [x,fval,exitflag,output,lambda] = localSolverCall(model);
+    switch output.cplexstatus
+        case {1,101,102} % It was ok, hence it must have been unbounded
+            output.cplexstatus = 2;
+        case {3,103,106} % Infeasible, so original was infeasible
+            output.cplexstatus = 3; % Infeasible
+        otherwise
+            output.cplexstatus = 4; % I give up
+    end
+end
+if interfacedata.getsolvertime solvertime = etime(clock,solvertime);else solvertime = 0;end
+
+% Inconstency in early version of CPLEX
+if str2num(interfacedata.solver.subversion)>=12.3
+    the_sign = 1;
+else
+    the_sign = -1;
+end
+if ~isempty(lambda)
+    D_struc = [the_sign*lambda.eqlin;the_sign*lambda.ineqlin];
+else
+    D_struc = [];    
+end
+
+if length(x) == length(model.f)
+    if ~isempty(model.NegativeSemiVar)
+        x(model.NegativeSemiVar) = -x(model.NegativeSemiVar);
+    end
+end
+
+showprogress('Calling CPLEX-IBM',options.showprogress);
+% Check, currently not exhaustive...
+switch output.cplexstatus
+    case {1,101,102}
+        problem = 0;
+    case {3,103,106}
+        problem = 1; % Infeasible
+    case {2,20,21,118}
+        problem = 2; % Unbounded
+    case {4,119}
+        problem = 15;    
+    case {10,11,104,105,107,108,111,112}
+        problem = 3; % Iteration/time
+    case 119
+        problem = 15;
+    case {5,6,109,110}
+        problem = 4; % Numerics
+    otherwise
+        problem = -1;
+end
[…]

### 2017-09-17 code/Yalmip/yalmip/solvers/callbonmin.m

@@ -1,0 +1,155 @@
+function output = callbonmin(model)
+
+% Author Johan Löfberg
+
+model = yalmip2nonlinearsolver(model);
+options = model.options.bonmin;
+options.ipopt.print_level = model.options.verbose;  
+
+if ~model.derivative_available
+    disp('Derivate-free call to bonmin/ipopt not yet implemented')
+    error('Derivate-free call to bonmin/ipopt not yet implemented')
+end
+
+if model.options.savedebug
+    save bonmindebug model
+end
+
+Fupp = [ repmat(0,length(model.bnonlinineq),1);
+    repmat(0,length(model.bnonlineq),1);
+    repmat(0,length(model.b),1);
+    repmat(0,length(model.beq),1)];
+
+Flow = [ repmat(-inf,length(model.bnonlinineq),1);
+    repmat(0,length(model.bnonlineq),1);
+    repmat(-inf,length(model.b),1);
+    repmat(0,length(model.beq),1)];
+
+if isempty(Flow)
+    Flow = [];
+    Fupp = [];
+end
+
+% Since ipopt react strangely on lb>ub, we should bail if that is detected
+% (ipopt creates an exception)
+if ~isempty(model.lb)
+    if any(model.lb>model.ub)
+        problem = 1;   
+        solverinput = [];
+        solveroutput = [];  
+        output = createoutput(model.lb*0,[],[],problem,'BONMIN',solverinput,solveroutput,0);
+        return
+    end
+end
+
+% These are needed to avoid recomputation due to ipopts double call to get
+% f and df, and g and dg
+global latest_x_f
+global latest_x_g
+global latest_df
+global latest_f
+global latest_G
+global latest_g
+global latest_xevaled
+global latest_x_xevaled
+latest_G= [];
+latest_g = [];
+latest_x_f = [];
+latest_x_g = [];
+latest_xevaled = [];
+latest_x_xevaled = [];
+
+funcs.objective = @(x)ipopt_callback_f(x,model);
+funcs.gradient = @(x)ipopt_callback_df(x,model);
+if ~isempty(Fupp)
+    funcs.constraints = @(x)ipopt_callback_g(x,model);
+    funcs.jacobian  = @(x)ipopt_callback_dg(x,model);
+end
+
+options.lb = model.lb(:)';
+options.ub = model.ub(:)';
+if ~isempty(Fupp)
+    options.cl = Flow;
+    options.cu = Fupp;
+end
+
+if ~isempty(Fupp)
+    m = length(model.lb);    
[…]

### 2017-09-17 code/Yalmip/yalmip/solvers/callgurobi.m

@@ -1,0 +1,91 @@
+function output = callgurobi(interfacedata)
+
+% Author Johan Löfberg 
+
+options = interfacedata.options;
+nOriginal = length(interfacedata.c);
+model = yalmip2gurobi(interfacedata);
+
+if interfacedata.options.savedebug
+    save gurobidebug model
+end
+
+if options.showprogress;showprogress('Calling GUROBI',options.showprogress);end
+solvertime = clock; 
+result = gurobi(model,model.params);
+if interfacedata.getsolvertime solvertime = etime(clock,solvertime);else solvertime = 0;end
+
+% Gurobi assumes semi-continuous variables only can take negative values so
+% we negate semi-continuous violating this
+if isfield(result,'x')
+    x = result.x;
+    if length(x) == length(model.obj)
+        if ~isempty(model.NegativeSemiVar)
+            x(model.NegativeSemiVar) = -x(model.NegativeSemiVar);
+        end
+    end
+    x = x(1:nOriginal);
+else
+    x = zeros(nOriginal,1);
+end
+
+problem = 0;
+if isfield(result,'pi')
+    if length(model.obj) == nOriginal
+        D_struc = -result.pi;
+    else
+    D_struc = [];
+    end
+else
+    D_struc = [];
+end
+
+switch result.status
+    case 'OPTIMAL'
+        problem =0;
+    case 'INFEASIBLE'
+        problem = 1;
+    case 'UNBOUNDED'
+        problem = 2;
+    case {'ITERATION_LIMIT','TIME_LIMIT','NODE_LIMIT'}
+        problem = 3;
+    case {'NUMERIC','SUBOPTIMAL'}
+        problem = 4;
+    case 'INF_OR_UNBD'
+        problem = 12;
+    otherwise
+        problem = 9;
+end
+
+% Save all data sent to solver?
+if interfacedata.options.savesolverinput
+	solverinput.model = model;
+else
+	solverinput = [];
+end
+
+% Save all data from the solver?
+if interfacedata.options.savesolveroutput    
+	solveroutput.result = result; 
+else
+	solveroutput = [];
+end
+
+% Standard interface 
+output.Primal      = x;
+output.Dual        = D_struc;
+output.Slack       = [];
+output.problem     = problem;
+output.solverinput = solverinput;
+output.solveroutput= solveroutput;
+output.solvertime  = solvertime;
+
+
+
+
+
+
+
+
+
+

### 2017-09-17 code/Yalmip/yalmip/solvers/callipopt.m

@@ -1,0 +1,138 @@
+function output = callipopt(model)
+
+% Author Johan Löfberg
+% $Id: callipopt.m,v 1.6 2009-09-29 10:30:15 joloef Exp $
+
+options = [];
+options.ipopt = model.options.ipopt;
+options.ipopt.print_level = 2+model.options.verbose;
+
+model = yalmip2nonlinearsolver(model);
+
+if ~model.derivative_available
+    disp('Derivate-free call to ipopt not yet implemented')
+    error('Derivate-free call to ipopt not yet implemented')
+end
+if model.options.savedebug
+    save ipoptdebug model
+end
+showprogress('Calling IPOPT',model.options.showprogress);
+
+Fupp = [ repmat(0,length(model.bnonlinineq),1);
+    repmat(0,length(model.bnonlineq),1);
+    repmat(0,length(model.b),1);
+    repmat(0,length(model.beq),1)];
+
+Flow = [ repmat(-inf,length(model.bnonlinineq),1);
+    repmat(0,length(model.bnonlineq),1);
+    repmat(-inf,length(model.b),1);
+    repmat(0,length(model.beq),1)];
+
+if isempty(Flow)
+    Flow = [];
+    Fupp = [];
+end
+
+% Since ipopt react strangely on lb>ub, we should bail if that is detected
+% (ipopt creates an exception)
+if ~isempty(model.lb)
+    if any(model.lb>model.ub)
+        problem = 1;   
+        solverinput = [];
+        solveroutput = [];  
+        output = createoutput(model.lb*0,[],[],problem,'IPOPT',solverinput,solveroutput,0);
+        return
+    end
+end
+
+% These are needed to avoid recomputation due to ipopts double call to get
+% f and df, and g and dg
+global latest_x_f
+global latest_x_g
+global latest_df
+global latest_f
+global latest_G
+global latest_g
+global latest_xevaled
+global latest_x_xevaled
+latest_G= [];
+latest_g = [];
+latest_x_f = [];
+latest_x_g = [];
+latest_xevaled = [];
+latest_x_xevaled = [];
+
+funcs.objective = @(x)ipopt_callback_f(x,model);
+funcs.gradient = @(x)ipopt_callback_df(x,model);
+if ~isempty(Fupp)
+    funcs.constraints = @(x)ipopt_callback_g(x,model);
+    funcs.jacobian  = @(x)ipopt_callback_dg(x,model);
+end
+
[…]

### 2017-09-17 code/Yalmip/yalmip/solvers/callmosek.m

@@ -1,0 +1,424 @@
+function output = callmosek(interfacedata)
+
+% Author Johan Löfberg 
+% $Id: callmosek.m,v 1.18 2009-03-11 09:45:32 joloef Exp $
+
+% Retrieve needed data
+options = interfacedata.options;
+F_struc = interfacedata.F_struc;
+c       = interfacedata.c;
+Q       = interfacedata.Q;
+K       = interfacedata.K;
+x0      = interfacedata.x0;
+integer_variables = interfacedata.integer_variables;
+binary_variables = interfacedata.binary_variables;
+extended_variables = interfacedata.extended_variables;
+ub      = interfacedata.ub;
+lb      = interfacedata.lb;
+mt      = interfacedata.monomtable;
+
+% *********************************
+% What type of variables do we have
+% *********************************
+linear_variables = find((sum(abs(mt),2)==1) & (any(mt==1,2)));
+nonlinear_variables = setdiff((1:size(mt,1))',linear_variables);
+sigmonial_variables = find(any(0>mt,2) | any(mt-fix(mt),2));
+
+if K.s(1)>0
+    [x,D_struc,problem,res,solvertime,prob] = call_mosek_sdp(interfacedata);
+elseif ~isempty(sigmonial_variables) | isequal(interfacedata.solver.version,'GEOMETRIC')
+    [x,D_struc,problem,res,solvertime,prob] = call_mosek_geometric(options,F_struc,c,Q,K,ub,lb,mt,linear_variables,extended_variables);          
+else
+    [x,D_struc,problem,res,solvertime,prob] = call_mosek_lpqp(options,F_struc,c,Q,K,ub,lb,mt,linear_variables,integer_variables,x0);        
+end
+
+infostr = yalmiperror(problem,'MOSEK');	
+
+% Save all data sent to solver?
+if options.savesolverinput
+    solverinput.prob = prob;
+else
+    solverinput = [];
+end
+
+% Save all data from the solver?
+if options.savesolveroutput
+    solveroutput.res = res;
+else
+    solveroutput = [];
+end
+
+% Standard interface 
+output.Primal      = x;
+output.Dual        = D_struc;
+output.Slack       = [];
+output.problem     = problem;
[…]

### 2017-09-17 code/Yalmip/yalmip/solvers/definesolvers.m

@@ -1,0 +1,1275 @@
+function solver = definesolvers
+
+% ****************************
+% Create a default solver
+% ****************************
+emptysolver.tag     = '';
+emptysolver.version = '';
+emptysolver.subversion = '';
+emptysolver.checkfor= {''};
+emptysolver.testcode= {''};
+emptysolver.call    = '';
+emptysolver.subcall = '';
+emptysolver.show    = 1;
+emptysolver.usesother = 0;
+emptysolver.supportsinitial = 1;
+
+emptysolver.objective.linear = 0;
+emptysolver.objective.quadratic.convex = 0;
+emptysolver.objective.quadratic.nonconvex = 0;
+emptysolver.objective.polynomial = 0;
+emptysolver.objective.maxdet.convex = 0;
+emptysolver.objective.maxdet.nonconvex = 0;
+emptysolver.objective.sigmonial = 0;
+
+emptysolver.constraint.equalities.linear     = 0;
+emptysolver.constraint.equalities.quadratic  = 0;
+emptysolver.constraint.equalities.polynomial = 0;
+emptysolver.constraint.equalities.sigmonial  = 0;
+
+emptysolver.constraint.inequalities.elementwise.linear = 0;
+emptysolver.constraint.inequalities.elementwise.quadratic.convex = 0;
+emptysolver.constraint.inequalities.elementwise.quadratic.nonconvex = 0;
+emptysolver.constraint.inequalities.elementwise.polynomial = 0;
+emptysolver.constraint.inequalities.elementwise.sigmonial = 0;
+
+emptysolver.constraint.inequalities.semidefinite.linear = 0;
+emptysolver.constraint.inequalities.semidefinite.quadratic = 0;
+emptysolver.constraint.inequalities.semidefinite.polynomial = 0;
+emptysolver.constraint.inequalities.semidefinite.sigmonial = 0;
+emptysolver.constraint.inequalities.rank = 0;
+
+emptysolver.constraint.inequalities.secondordercone = 0;
+emptysolver.constraint.inequalities.rotatedsecondordercone = 0;
+emptysolver.constraint.inequalities.powercone = 0;
+
[…]

### 2017-09-17 code/Yalmip/yalmip/solvers/yalmip2cplex.m

@@ -1,0 +1,123 @@
+function model = yalmip2cplex(interfacedata)
+
+% Author Johan Löfberg
+
+% Retrieve needed data
+options = interfacedata.options;
+F_struc = interfacedata.F_struc;
+c       = interfacedata.c;
+K       = interfacedata.K;
+x0      = interfacedata.x0;
+integer_variables = interfacedata.integer_variables;
+binary_variables = interfacedata.binary_variables;
+semicont_variables = interfacedata.semicont_variables;
+
+ub      = interfacedata.ub;
+lb      = interfacedata.lb;
+H       = 2*interfacedata.Q;
+[F_struc,K,lb,ub,semicont_variables] = extractSemiContBounds(F_struc,K,lb,ub,semicont_variables);
+[F_struc,K,c,H,ub,lb,Qi,Li,ri] = append_normalized_socp(F_struc,K,c,H,ub,lb);
+
+if any(interfacedata.variabletype)
+    error('Nonlinear monomials remain when calling CPLEX. If you are using OPTIMIZER, please report this and ask for a feature improvement.');
+end
+
+% Notation used
+f = c;
+if ~isempty(F_struc)
+    Aineq = -F_struc(:,2:end);
+    bineq = F_struc(:,1);
+else
+    Aineq = [];
+    bineq = [];
+end
+
+% CPLEX assumes semi-continuous variables only can take negative values so
+% we negate semi-continuous violating this
+[NegativeSemiVar,H,f,Aineq,lb,ub,semicont_variables] = negateNegativeSemiContVariables(H,f,Aineq,lb,ub,semicont_variables,[]);
+
+if K.f > 0
+    Aeq = Aineq(1:K.f,:);
+    beq = bineq(1:K.f);
+    % Code around performance bugs in older version of MATLAB
+    [ii,jj,ss] = find(Aineq);keeps = ii>K.f;
+    Aineq = sparse(ii(keeps)-K.f,jj(keeps),ss(keeps),size(Aineq,1)-K.f,size(Aineq,2));
+    [ii,jj,ss] = find(bineq);keeps = ii>K.f;
+    bineq = sparse(ii(keeps)-K.f,jj(keeps),ss(keeps),length(bineq)-K.f,1);
+else
+    Aeq = [];
+    beq = [];
+end
+
+if all(isinf(lb))
+    lb = [];
+end
+if all(isinf(ub))
+    ub = [];
+end
+
+ctype = char(ones(length(f),1)*67);
[…]

### 2017-09-17 code/Yalmip/yalmip/solvers/yalmip2gurobi.m

@@ -1,0 +1,133 @@
+function model = yalmip2gurobi(interfacedata)
+
+% Retrieve needed data
+F_struc = interfacedata.F_struc;
+Q       = interfacedata.Q;
+c       = interfacedata.c;
+K       = interfacedata.K;
+integer_variables = interfacedata.integer_variables;
+binary_variables = interfacedata.binary_variables;
+semicont_variables = interfacedata.semicont_variables;
+ub      = interfacedata.ub;
+lb      = interfacedata.lb;
+x0      = interfacedata.x0;
+n = length(c);
+
+if ~isempty(ub)
+    LB = lb;
+    UB = ub;
+    if ~isempty(binary_variables)
+        LB(binary_variables)  = round(LB(binary_variables));
+        UB(binary_variables)  = round(UB(binary_variables));
+    end
+    if ~isempty(integer_variables)
+        LB(integer_variables) = round(LB(integer_variables));
+        UB(integer_variables) = round(UB(integer_variables));
+    end
+else
+    LB = -repmat(inf,n,1);
+    UB = repmat(inf,n,1);
+end
+
+%[F_struc,K,LB,UB,semicont_variables] = extractSemiContBounds(F_struc,K,LB,UB,semicont_variables);
+
+if ~isempty(semicont_variables)
+    % Bounds must be placed in LB/UB
+    [LB,UB,cand_rows_eq,cand_rows_lp] = findulb(F_struc,K,LB,UB);
+    F_struc(K.f+cand_rows_lp,:)=[];
+    F_struc(cand_rows_eq,:)=[];
+    K.l = K.l-length(cand_rows_lp);
+    K.f = K.f-length(cand_rows_eq);    
+    redundant = find(LB<=0 & UB>=0);
+    semicont_variables = setdiff(semicont_variables,redundant);        
+end
+
+n_original = length(c);
+[F_struc,K,c,Q,UB,LB] = append_normalized_socp(F_struc,K,c,Q,UB,LB);
+
+if size(F_struc,1)>0
+    B = full(F_struc(:,1));         % Must be full  
+    A =-F_struc;
+    A(:,1)=[];
+else
+    B = [];
+    A = [];
+end
+
+% Optimized code, make a lot of difference when you make this call 10000
+% times in a branch and bound setting...
+CTYPE = char([ones(K.f,1)*61; ones(K.l,1)*60]);
+VARTYPE = char(ones(length(c),1)*67);
+if isempty(semicont_variables)
[…]

### 2017-09-17 code/Yalmip/yalmip/solvesdp.m

@@ -1,0 +1,524 @@
+function diagnostic = solvesdp(varargin)
+%SOLVESDP Computes solution to optimization problem
+%
+%   DIAGNOSTIC = SOLVESDP(F,h,options) is the common command to
+%   solve optimization problems of the following kind
+%
+%    min        h
+%    subject to
+%            F >=(<=,==) 0
+%
+%   NOTES
+%    Despite the name, SOLVESDP is the interface for solving all
+%    supported problem classes (LP, QP, SOCP, SDP, BMI, MILP, MIQP,...)
+%
+%    To obtain solution for a variable, use DOUBLE.
+%
+%    To obtain dual variable for a constraint, use DUAL.
+%
+%    See YALMIPERROR for error codes returned in output.
+%
+%   OUTPUT
+%     diagnostic : Diagnostic information
+%
+%   INPUT
+%     F          : Object describing the constraints. Can be [].
+%     h          : SDPVAR object describing the objective h(x). Can be [].
+%     options    : Options structure. See SDPSETTINGS. Can be [].
+%
+%   EXAMPLE
+%    A = randn(15,5);b = rand(15,1)*5;c = randn(5,1);
+%    x = sdpvar(5,1);
+%    solvesdp(set(A*x<=b),c'*x);double(x)
+%
+%   See also DUAL, @SDPVAR/DOUBLE, SDPSETTINGS, YALMIPERROR
+
+% Author Johan Löfberg
+% $Id: solvesdp.m,v 1.75 2010-04-06 06:32:49 joloef Exp $
+
+yalmiptime = clock; % Let us see how much time we spend
+
+% Avoid warning
+if length(varargin)>=2 & isa(varargin{2},'double')
+	varargin{2} = [];
+end
+
+if length(varargin)>=2 & isa(varargin{2},'sdpvar') & prod(size(varargin{2}))>1
+    % Several objectives
+     diagnostic = solvesdp_multiple(varargin{:});
+     return
+end
+
+% Arrrgh, new format with logdet much better, but we have to
+% take care of old code, requires some testing...
+varargin = combatible({varargin{:}});
+nargin = length(varargin);
+
+% *********************************
+% CHECK INPUT
+% *********************************
+
+
+if nargin<1
+    help solvesdp
+    return
+else
+    F = varargin{1};
+    if isa(F,'constraint')
+        F = set(F);
+    end
+    
+    if isa(F,'sdpvar')
[…]

### 2017-09-17 code/Yalmip/yalmip/yalmiptest.m

@@ -1,0 +1,826 @@
+function out = yalmiptest(prefered_solver,auto)
+%YALMIPTEST Runs a number of test problems.
+%
+%   YALMIPTEST is recommended when a new solver or a new version
+%   of YALMIP installed.
+%
+%   EXAMPLES
+%    YALMIPTEST               % Without argument, default solver used
+%    YALMIPTEST('solver tag') % Test with specified solver
+%    YALMIPTEST(options)      % Test with specific options structure from
+%
+%   See also SDPSETTINGS, YALMIPDEMO
+
+% Author Johan Löfberg 
+% $Id: yalmiptest.m,v 1.28 2008-05-09 13:51:41 joloef Exp $
+
+
+if ~exist('sedumi2pen.m')
+    disp('Add /yalmip/extras to your path first...')
+    return
+end
+
+if ~exist('bmibnb.m')
+    disp('Add /yalmip/modules/ with sub-directories to your path first...')
+    return
+end
+
+if ~exist('pwa_yalmip.m')
+    disp('Add /yalmip/operators to your path first...')
+    return
+end
+
+if ~exist('callsedumi.m')
+    disp('Add /yalmip/solvers to your path first...')
+    return
+end
+
+detected = which('yalmip.m','-all');
+if length(detected)>1
+    disp('You seem to have multiple installations of YALMIP in your path. Please correct this...');
+    detected
+    return
+end
+
+donttest = 0;
+if (nargin==1) & isa(prefered_solver,'char') & strcmp(prefered_solver,'test')
+    donttest = 0;
+    prefered_solver = '';    
+else
+    donttest = 1;
+end
+
+if nargin==0
+    prefered_solver = '';
+else
+    if ~(isa(prefered_solver,'struct') | isa(prefered_solver,'char'))
+        error('Argument should be a solver tag, or a sdpsettings structure');
+    end
+    if isa(prefered_solver,'char') 
+        donttest = 1;
+    end
+end
+
+if ~(exist('callsedumi')==2)
+    disp('The directory yalmip/solvers is not in your path.')
[…]

### 2017-09-17 code/data/1 inputs/CreateDPCdata.m

@@ -1,0 +1,52 @@
+%% Convert data to adapt to dpc code
+
+clear all
+close all
+clc
+
+load('Data_PreProccessing/cetemps_interp_10min.mat')
+load('Data_PreProccessing/et_interp_10min.mat')
+load('Data_PreProccessing/ti_interp_10min.mat')
+
+sTrain = 67; % Training start index
+eTrain = 6734; % Training end index
+sTest = 7346; % Testing start index
+eTest = 9505; % Testing end index
+
+%% Creating "tod - time of day" and "dow - day of week" for the training and testing processes
+tTrain = datenum(ti_interp_10min(sTrain:eTrain,1)); % Time vector for the train dataset
+temp = datevec(tTrain);
+todTrain = temp(:,4)';
+[dowTrain ~] = weekday(tTrain);
+dowTrain = dowTrain';
+
+tTest = datenum(ti_interp_10min(sTest:eTest,1));
+temp = datevec(tTest);
+todTest = temp(:,4)';
+[dowTest ~] = weekday(tTest);
+dowTest = dowTest';
+
+clear temp
+
+%% Creates data sets
+
+% Creates training data
+
+x = cell2mat(ti_interp_10min(sTrain:eTrain,2:end))';
+d = cell2mat(cetemps_interp_10min(sTrain:eTrain,2:end))';
+d = [d;todTrain;dowTrain];
+u = cell2mat(et_interp_10min(sTrain:eTrain,[5]))';
+y = cell2mat(et_interp_10min(sTrain:eTrain,6))';
+
+save('train.mat','d','u','x','y');
+save('tTrain','tTrain')
+% Creates testing data
+
+x = cell2mat(ti_interp_10min(sTest:eTest,2:end))';
+d = cell2mat(cetemps_interp_10min(sTest:eTest,2:end))';
+d = [d;todTest;dowTest];
+u = cell2mat(et_interp_10min(sTest:eTest,[5]))';
+y = cell2mat(et_interp_10min(sTest:eTest,6))';
+
+save('test.mat','d','u','x','y');
+save('tTest','tTest')

### 2017-09-17 code/data/2 inputs/CreateDPCdata.m

@@ -1,0 +1,51 @@
+%% Convert data to adapt to dpc code
+
+clear all
+close all
+clc
+
+load('Data_PreProccessing/cetemps_interp_10min.mat')
+load('Data_PreProccessing/et_interp_10min.mat')
+load('Data_PreProccessing/ti_interp_10min.mat')
+
+sTrain = 67; % Training start index
+eTrain = 6734; % Training end index
+sTest = 7346; % Testing start index
+eTest = 9505; % Testing end index
+
+%% Creating "tod - time of day" and "dow - day of week" for the training and testing processes
+tTrain = datenum(ti_interp_10min(sTrain:eTrain,1)); % Time vector for the train dataset
+temp = datevec(tTrain);
+todTrain = temp(:,4)';
+[dowTrain ~] = weekday(tTrain);
+dowTrain = dowTrain';
+
+tTest = datenum(ti_interp_10min(sTest:eTest,1));
+temp = datevec(tTest);
+todTest = temp(:,4)';
+[dowTest ~] = weekday(tTest);
+dowTest = dowTest';
+
+clear temp
+
+%% Creates data sets
+
+% Creates training data
+
+x = cell2mat(ti_interp_10min(sTrain:eTrain,2:end))';
+d = cell2mat(cetemps_interp_10min(sTrain:eTrain,2:end))';
+d = [d;todTrain;dowTrain];
+u = cell2mat(et_interp_10min(sTrain:eTrain,[3 5]))';
+y = cell2mat(et_interp_10min(sTrain:eTrain,6))';
+
+save('train.mat','d','u','x','y');
+
+% Creates testing data
+
+x = cell2mat(ti_interp_10min(sTest:eTest,2:end))';
+d = cell2mat(cetemps_interp_10min(sTest:eTest,2:end))';
+d = [d;todTest;dowTest];
+u = cell2mat(et_interp_10min(sTest:eTest,[3 5]))';
+y = cell2mat(et_interp_10min(sTest:eTest,6))';
+
+save('test.mat','d','u','x','y');

### 2017-09-17 code/dpcCVX.m

@@ -1,0 +1,116 @@
+function [U,epsilonm,epsilonM] = dpcCVX(model, leafmodels, k, xtot, uprev, d, orderAR, N, umin, umax, nc, numRooms, tref,P,Q)
+
+%% calculate leaf coeffs
+nx = numRooms+1;
+d_cur = d(:,k-N+1:k); % features corr to disturbances
+
+% Creating empty cell for coefficients: 1 for the power and one for each room
+coeff = cell(numRooms+1,1);
+for ii = 1:numRooms+1
+    coeff{ii} = zeros(nc*N+1,N);
+end
+
+% Computing coefficients for room models
+for ii = 1:numRooms
+    % prepare features_d to find an appropriate leaf for rooms
+    x_cur = fliplr(xtot(ii,k-orderAR+1:k)); % features corr to indoor temp.
+    features_d = [d_cur(:)', x_cur];
+    for idm = 1:N % Because the first column contains a0+a1*u(0), the second b0+b1*u(0)+b2*u(1) and so on
+        coeff{ii}(1:nc*idm+1,idm) = find_linearmodel_in_leaves(model{ii,idm}, leafmodels{ii,idm}, features_d);
+    end
+end
+% prepare features_d to find an appropriate leaf for power
+d_cur = d(:,k:k+N-1); % features corr to disturbances
+x_cur = [fliplr(xtot(1,k-orderAR+1:k)),...
+         fliplr(xtot(2,k-orderAR+1:k)),...
+         fliplr(xtot(3,k-orderAR+1:k)),...
+         fliplr(xtot(4,k-orderAR+1:k))]; % features corr to indoor temp.
+features_d = [d_cur(:)', x_cur];
+
+% Computing coefficients for power
+for idm = 1:N % Because the first column contains a0+a1*u(0), the second b0+b1*u(0)+b2*u(1) and so on
+   coeff{numRooms+1}(1:nc*idm+1,idm) = find_linearmodel_in_leaves(model{numRooms+1,idm}, leafmodels{numRooms+1,idm}, features_d);
+end
+
+% tref1 = tref(1,k+1:k+N);
+% tref2 = tref(2,k+1:k+N);
+% tref3 = tref(3,k+1:k+N);
+% tref4 = tref(4,k+1:k+N);
+% tref1 = tref(1);
+% tref2 = tref(2);
+% tref3 = tref(3);
+% tref4 = tref(4);
+tref3 = tref(:,k+1:k+N);
[…]

### 2017-09-17 code/find_linearmodel_in_leaves.m

@@ -1,0 +1,36 @@
+function coeff = find_linearmodel_in_leaves(model, leafmodels, features_d)
+
+NTrees = model.NTrees;
+coeff = zeros(length(leafmodels{1}(1).coeff),NTrees);
+
+for idt = 1:NTrees
+    
+    outputs = predict(model.Trees{idt}, features_d);
+
+    % find the corresponding leaf index
+    for idl = 1:length(leafmodels{idt})
+        if abs(outputs - leafmodels{idt}(idl).mean)<1e-8
+            break;
+        end
+    end
+
+%     % litmus test
+%     [testypred,testleaf] = predict(model.Trees{idt},features_d);
+%     [~, testnode] = predict( model.Trees{idt}, model.X(~model.OOBIndices(:,idt),:) );
+%     % if mean(model.Y(find(testnode==testleaf)))-testypred>1e-10
+%     %     keyboard;
+%     % end
+%     if find(testnode==testleaf)~=leafmodels{idt}(idl).leaves{1,1}
+%         keyboard;
+%     end
+
+    % extract the linear model coefficients from that leaf
+    coeff(:,idt) = leafmodels{idt}(idl).coeff;
+    
+    if coeff(1,idt)<0 || coeff(1,idt)>50
+        coeff(:,idt) = NaN;
+    end
+    
+end
+coeff(:,isnan(coeff(1,:))) = [];
+coeff = mean(coeff,2);

### 2017-09-17 code/load_data_new_power.m

@@ -1,0 +1,28 @@
+% idm is the tree number on the horizon we want to load disturbance for
+
+function [features_d, features_c, outputs] = load_data_new_power(argStr, orderAR, ctrlHzn, idm, nc)
+
+if strcmp(argStr, 'train')
+    load(['Progetto Tullio/DPC/data/',num2str(nc),' inputs/train.mat']);
+elseif strcmp(argStr, 'test')
+    load(['Progetto Tullio/DPC/data/',num2str(nc),' inputs/test.mat']);
+else
+    error('input string should be either ''train'' or ''test.');
+end
+
+features_d = [lagmatrix(d', 0-idm+1:1:ctrlHzn-1-idm+1),...
+              lagmatrix(x(1,:), 0:1:orderAR-1),...
+              lagmatrix(x(2,:), 0:1:orderAR-1),...
+              lagmatrix(x(3,:), 0:1:orderAR-1),...
+              lagmatrix(x(4,:), 0:1:orderAR-1)]; %#ok<*NODEF>
+
+features_c = lagmatrix(u', -(0:1:ctrlHzn-1));
+
+outputs = lagmatrix(y', -(1:1:ctrlHzn));
+
+features_d(1:orderAR-1,:) = [];
+features_c(1:orderAR-1,:) = [];
+outputs(1:orderAR-1,:) = [];
+features_d(end-ctrlHzn:end,:) = [];
+features_c(end-ctrlHzn:end,:) = [];
+outputs(end-ctrlHzn:end,:) = [];

### 2017-09-17 code/load_data_new_states.m

@@ -1,0 +1,24 @@
+% idm is the tree number on the horizon we want to load disturbance for
+
+function [features_d, features_c, outputs] = load_data_new_states(argStr, orderAR, ctrlHzn, idm, ii,nc)
+
+if strcmp(argStr, 'train')
+    load(['Progetto Tullio/DPC/data/',num2str(nc),' inputs/train.mat']);
+elseif strcmp(argStr, 'test')
+    load(['Progetto Tullio/DPC/data/',num2str(nc),' inputs/test.mat']);
+else
+    error('input string should be either ''train'' or ''test.');
+end
+
+features_d = [lagmatrix(d', 0-idm+1:1:ctrlHzn-1-idm+1), lagmatrix(x(ii,:), 0:1:orderAR-1)]; %#ok<*NODEF>
+
+features_c = lagmatrix(u', -(0:1:ctrlHzn-1));
+
+outputs = lagmatrix(x(ii,:)', -(1:1:ctrlHzn));
+
+features_d(1:orderAR-1,:) = [];
+features_c(1:orderAR-1,:) = [];
+outputs(1:orderAR-1,:) = [];
+features_d(end-ctrlHzn:end,:) = [];
+features_c(end-ctrlHzn:end,:) = [];
+outputs(end-ctrlHzn:end,:) = [];

### 2017-09-17 code/main.m

@@ -1,0 +1,205 @@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Things to do before compiling: comment or add input constraints for u3 in dpcCVX
+%                                change nc = 1 or nc = 2 or nc = 3
+%                                check orderAR and ctrlHzn
+%                                reload models
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+
+
+% Before running this function, make sure you run model_training
+clearvars -except model leafmodels power roomT1 roomT2 roomT3 roomT4
+close all
+clc
+addpath(genpath('D:\PostDoc\Matlab\cvx'))
+cvx_solver gurobi
+
+% addpath(genpath('C:\Program Files\IBM'))
+
+orderAR = 2; % This has to be the same used in "model_training.m"
+ctrlHzn = 4; % This has to be the same used in "model_training.m"
+index = max(orderAR,ctrlHzn); % Used to create the initial state and input vectors of appropriate dimensions
+start = 1;
+nc = 1;
+numRooms = 4;
+simTime = 2150;
+xtot = zeros(numRooms+1,simTime);
+utot = zeros(nc,simTime);
+epsilontotm = zeros(4,simTime);
+epsilontotM = zeros(4,simTime);
+temptot = zeros(4,simTime);
+%% Load models
+% disp('Loading ramdom forest models')
+% load(['data/',num2str(nc),' inputs/models_random_forest_N=4_AR=2.mat']);
+% disp('Loading linear models in the leaves')
+% load(['data/',num2str(nc),' inputs/lin_random_forest_N=4_AR=2.mat']);
+% disp('Loading testing data')
+load(['data/',num2str(nc),' inputs/test.mat']);
+load(['data/',num2str(nc),' inputs/tTest.mat']);
+
+% disp('Loading real plant data')
[…]

### 2017-09-17 code/model_training.m

@@ -1,0 +1,95 @@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Things to do before compiling: change nc - Choose between 2 inputs and 3 inputs
+%                                check orderAR and ctrlHzn
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+
+clear all
+close all
+clc
+
+% set params
+orderAR = 2; % Autoregression order
+ctrlHzn = 1; % Control horizon
+nc = 1;
+
+%% Creating random forest models
+numRooms = 4;
+model = cell(numRooms+1,ctrlHzn); % The +1 is for the power component
+
+warning('off','all')
+leafmodels = cell(numRooms+1,ctrlHzn);
+
+% Creating models for the rooms
+for ii = 1:numRooms %% 5 is the number of rooms in ti_intermp_10min
+   
+    % Create random forests
+    catcol = sort([7*(1:ctrlHzn)-1 7*(1:ctrlHzn)]);
+    leafsize = 20;
+    disp('Room models computation');
+    for idm = 1:ctrlHzn % separate regression tree for each output
+        disp(['Loading data for room: ',num2str(ii)]);
+        [xTrain_d, xTrain_c, yTrain] = load_data_new_states('train', orderAR, ctrlHzn,idm,ii,nc);
+        nc = size(xTrain_c,2)/ctrlHzn; % Number of control inputs
+    
+            % Remove NaN values
+            total = [yTrain,xTrain_d,xTrain_c];
+            [m,n] = size(xTrain_d);
+            [m,p] = size(total);
+            total(any(isnan(total),2),:) = [];
+            yTrain = total(:,1:ctrlHzn);
+            xTrain_d = total(:,ctrlHzn + 1:ctrlHzn + n);
+            xTrain_c = total(:,ctrlHzn + n + 1:p);
+        
+        % Create random forest
[…]

### 2017-09-17 code/old functions/dpc.m

@@ -1,0 +1,72 @@
+function [input] = dpc(model, leafmodels, k, xtot, uprev, d, orderAR, N, umin, umax, xmin, xmax, nc, numRooms, tref,P,Q)
+
+%% calculate leaf coeffs
+nx = numRooms+1;
+d_cur = d(:,k-N+1:k); % features corr to disturbances
+
+% Creating empty cell for coefficients: 1 for the power and one for each room
+coeff = cell(numRooms+1,1);
+for ii = 1:numRooms+1
+    coeff{ii} = zeros(nc*N+1,N);
+end
+
+% Computing coefficients for room models
+for ii = 1:numRooms
+    % prepare features_d to find an appropriate leaf for rooms
+    x_cur = fliplr(xtot(ii,k-orderAR+1:k)); % features corr to indoor temp.
+    features_d = [d_cur(:)', x_cur];
+    for idm = 1:N % Because the first column contains a0+a1*u(0), the second b0+b1*u(0)+b2*u(1) and so on
+        coeff{ii}(1:nc*idm+1,idm) = find_linearmodel_in_leaves(model{ii,idm}, leafmodels{ii,idm}, features_d);
+    end
+end
+% prepare features_d to find an appropriate leaf for power
+d_cur = d(:,k:k+N-1); % features corr to disturbances
+x_cur = [fliplr(xtot(1,k-orderAR+1:k)),...
+         fliplr(xtot(2,k-orderAR+1:k)),...
+         fliplr(xtot(3,k-orderAR+1:k)),...
+         fliplr(xtot(4,k-orderAR+1:k))]; % features corr to indoor temp.
+features_d = [d_cur(:)', x_cur];
+
+% Computing coefficients for power
+for idm = 1:N % Because the first column contains a0+a1*u(0), the second b0+b1*u(0)+b2*u(1) and so on
+   coeff{numRooms+1}(1:nc*idm+1,idm) = find_linearmodel_in_leaves(model{numRooms+1,idm}, leafmodels{numRooms+1,idm}, features_d);
+end
+
+tref1 = tref(1,k+1:k+N);
+tref2 = tref(2,k+1:k+N);
[…]

### 2017-09-17 code/old functions/dpcYalmip.m

@@ -1,0 +1,114 @@
+function [input] = dpcYalmip(model, leafmodels, k, xtot, uprev, d, orderAR, N, umin, umax, xmin, xmax, nc, numRooms, tref,P,Q)
+
+%% calculate leaf coeffs
+nx = numRooms+1;
+d_cur = d(:,k-N+1:k); % features corr to disturbances
+
+% Creating empty cell for coefficients: 1 for the power and one for each room
+coeff = cell(numRooms+1,1);
+for ii = 1:numRooms+1
+    coeff{ii} = zeros(nc*N+1,N);
+end
+
+% Computing coefficients for room models
+for ii = 1:numRooms
+    % prepare features_d to find an appropriate leaf for rooms
+    x_cur = fliplr(xtot(ii,k-orderAR+1:k)); % features corr to indoor temp.
+    features_d = [d_cur(:)', x_cur];
+    for idm = 1:N % Because the first column contains a0+a1*u(0), the second b0+b1*u(0)+b2*u(1) and so on
+        coeff{ii}(1:nc*idm+1,idm) = find_linearmodel_in_leaves(model{ii,idm}, leafmodels{ii,idm}, features_d);
+    end
+end
+% prepare features_d to find an appropriate leaf for power
+d_cur = d(:,k:k+N-1); % features corr to disturbances
+x_cur = [fliplr(xtot(1,k-orderAR+1:k)),...
+         fliplr(xtot(2,k-orderAR+1:k)),...
+         fliplr(xtot(3,k-orderAR+1:k)),...
+         fliplr(xtot(4,k-orderAR+1:k))]; % features corr to indoor temp.
+features_d = [d_cur(:)', x_cur];
+
+% Computing coefficients for power
+for idm = 1:N % Because the first column contains a0+a1*u(0), the second b0+b1*u(0)+b2*u(1) and so on
+   coeff{numRooms+1}(1:nc*idm+1,idm) = find_linearmodel_in_leaves(model{numRooms+1,idm}, leafmodels{numRooms+1,idm}, features_d);
+end
+
+tref1 = tref(1,k+1:k+N);
+tref2 = tref(2,k+1:k+N);
+tref3 = tref(3,k+1:k+N);
+tref4 = tref(4,k+1:k+N);
+% tref1 = tref(1);
+% tref2 = tref(2);
+% tref3 = tref(3);
+% tref4 = tref(4);
+%% solve optimization problem
+addpath(genpath('D:\PostDoc\Matlab\cvx'))
[…]

### 2017-09-17 code/old functions/dpc_original.m

@@ -1,0 +1,63 @@
+function [u,fval,iter,time] = mydpc(model, leafmodels, k, x, d, Q, R, orderAR, N, umin, umax, uprev, udotmax, solver, xref)
+
+%% calculate leaf coeffs
+
+% prepare features_d to find an appropriate leaf
+d_cur = d(:,k:k+N-1); % features corr to disturbances
+x_cur = fliplr(x(4,k-orderAR+1:k)); % features corr to indoor temp.
+features_d = [d_cur(:)', x_cur];
+
+coeff = zeros(N+1,N);
+for idm = 1:N
+    coeff(1:idm+1,idm) = find_linearmodel_in_leaves(model{idm}, leafmodels{idm}, features_d);
+end
+
+% objective function
+QQ = diag(Q(4,4)*ones(N,1));
+RR=kron(eye(N),R);
+xref = xref(4)*ones(N,1);
+
+% input rate change constraint
+UB = uprev+udotmax;
+LB = uprev-udotmax;
+umin = max(umin,LB);
+umax = min(UB,umax);
+
+%% solve optimization problem
+
+switch solver
+    
+    case 'cvx' % too slow
+        
+        tic;
+        cvx_begin quiet
+            variables U(N)
+            minimize( U'*RR*U + ([1, U']*coeff-xref')*QQ*([1, U']*coeff-xref')' )
+            subject to
+                U>=umin; %#ok<*VUNUS>
+                U<=umax;
+        cvx_end
+        time = toc;
+        fval = 0;
+        iter = 0;
+        
+    case 'quadprog'
+        
+        C1 = coeff(1,:);
+        C2 = coeff(2:end,:);
+        H = 2*( RR +  C2*QQ*C2' );
+        f = 2*C2*QQ*C1'-2*C2*QQ*xref;
+        c = C1*QQ*C1'-C1*QQ*xref-xref'*QQ*C1'+xref'*QQ*xref;
+        lb = repmat(umin,N,1);
+        ub = repmat(umax,N,1);
+        options=optimset;
+        options.Algorithm='interior-point-convex'; % to get rid of some warnings.
+        options.Display='off';
+        tic;
+        [U, fval] = quadprog(H,f,[],[],[],[],lb,ub,[],options);
+        time=toc;
+        iter=nan;
+        
+end
+    
+u = U(1);

### 2017-09-17 code/old functions/load_data_power.m

@@ -1,0 +1,30 @@
+function [features_d, features_c, outputs] = load_data_power(argStr, orderAR, ctrlHzn)
+
+if strcmp(argStr, 'train')
+    load('data/train.mat');
+elseif strcmp(argStr, 'test')
+    load('data/test.mat');
+else
+    error('input string should be either ''train'' or ''test.');
+end
+
+features_d = [lagmatrix(d', -(0:1:ctrlHzn-1)),...
+              lagmatrix(x(1,:), 0:1:orderAR-1),...
+              lagmatrix(x(2,:), 0:1:orderAR-1),...
+              lagmatrix(x(3,:), 0:1:orderAR-1),...
+              lagmatrix(x(4,:), 0:1:orderAR-1)]; %#ok<*NODEF>
+          
+          
+features_d = [lagmatrix(d', -(0:1:ctrlHzn-1)), lagmatrix(x(4,:), 0:1:orderAR-1)]; %#ok<*NODEF>
+
+
+features_c = lagmatrix(u', -(0:1:ctrlHzn-1));
+
+outputs = lagmatrix(y', -(1:1:ctrlHzn));
+
+features_d(1:orderAR-1,:) = [];
+features_c(1:orderAR-1,:) = [];
+outputs(1:orderAR-1,:) = [];
+features_d(end-ctrlHzn:end,:) = [];
+features_c(end-ctrlHzn:end,:) = [];
+outputs(end-ctrlHzn:end,:) = [];

### 2017-09-17 code/old functions/load_data_states.m

@@ -1,0 +1,23 @@
+function [features_d, features_c, outputs] = load_data_states(argStr, orderAR, ctrlHzn,ii)
+
+if strcmp(argStr, 'train')
+    load('data/train.mat');
+elseif strcmp(argStr, 'test')
+    load('data/test.mat');
+else
+    error('input string should be either ''train'' or ''test.');
+end
+
+
+features_d = [lagmatrix(d', -(0:1:ctrlHzn-1)), lagmatrix(x(ii,:), 0:1:orderAR-1)]; %#ok<*NODEF>
+
+features_c = lagmatrix(u', -(0:1:ctrlHzn-1));
+
+outputs = lagmatrix(x(ii,:)', -(1:1:ctrlHzn));
+
+features_d(1:orderAR-1,:) = [];
+features_c(1:orderAR-1,:) = [];
+outputs(1:orderAR-1,:) = [];
+features_d(end-ctrlHzn:end,:) = [];
+features_c(end-ctrlHzn:end,:) = [];
+outputs(end-ctrlHzn:end,:) = [];

### 2017-09-17 code/plot_ml_results.m

@@ -1,0 +1,19 @@
+function plot_ml_results(yTrue, yPred, argStr, N)
+
+switch argStr
+    
+    case 'true-predicted'
+        % true-predicted plot
+        figure; grid on;
+        plot(yTrue(:,1), yPred(1,:)', '.');
+        
+    case 'predictive'
+        % quantitative output comparison
+        figure; hold on; grid on;
+        l = length(yTrue(:,1));
+        h1 = plot(1:l, yPred(1,:), 'LineWidth', 2);
+        h3 = plot((1:l)+N-1, yPred(N,:), 'LineWidth', 2);
+        h4 = plot(1:l, yTrue(:,1), 'LineWidth', 2);
+        legend([h1, h3, h4], 'T', 'T+N', 'ground truth')
+        
+end

### 2017-09-17 code/test_performance.m

@@ -1,0 +1,84 @@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Things to do before compiling: change path in the load_data_new_states and power - Choose between 2 inputs folder and 3 inputs folder
+%                                check orderAR and ctrlHzn
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+orderAR = 2;
+ctrlHzn = 1;
+nc = 1;
+modelIndex = 3; % 1 for roomT1,...,4 for roomT4, 5 for power
+if ((modelIndex > 0) && (modelIndex < 5))
+    [xTest_d, xTest_c, yTest] = load_data_new_states('test', orderAR, ctrlHzn,1,modelIndex,nc);
+elseif modelIndex == 5
+    [xTest_d, xTest_c, yTest] = load_data_new_power('test', orderAR, ctrlHzn,1,nc);
+end
+nc = size(xTest_c,2)/ctrlHzn;
+
+n = 1000;
+xTest_d = xTest_d(1:n,:);
+xTest_c = xTest_c(1:n,:);
+yTest = yTest(1:n,:);
+
+% calculate root mean square error with our definition
+yTrue = yTest;
+yPred = zeros(size(xTest_d,1),ctrlHzn);
+
+disp('Loading ramdom forest models')
+load(['data/',num2str(nc),' inputs/models_random_forest_N=1_AR=2.mat']);
+disp('Loading linear models in the leaves')
+load(['data/',num2str(nc),' inputs/lin_random_forest_N=1_AR=2.mat']);
+disp('Loading done!')
+
+% find linear models
+for idn = 1:size(xTest_d,1)
+    for idm = 1
+        coeff = find_linearmodel_in_leaves(model{modelIndex,idm}, leafmodels{modelIndex,idm}, xTest_d(idn,:));
+        yPred(idn,idm) = [1, xTest_c(idn,1:nc*idm)]*coeff;
+    end
[…]

### 2017-09-17 code/train_linearmodel_in_leaves.m

@@ -1,0 +1,64 @@
+function linearmodels = train_linearmodel_in_leaves(model, xTrain_c_all, yTrain_all, numb,nxc)
+
+xTrain_d = model.X;
+NTrees = model.NTrees;
+
+h0 = waitbar(0, 'Synthesis Forest');
+
+for i = 1:NTrees
+    
+    leaf_index = find((model.Trees{i}.Children(:,1)==0)&(model.Trees{i}.Children(:,2)==0));
+    numleafs = length(leaf_index);
+    [~,node] = predict(model.Trees{i}, xTrain_d(~model.OOBIndices(:,i),:));
+    xTrain_c = xTrain_c_all(~model.OOBIndices(:,i),:);
+    yTrain = yTrain_all(~model.OOBIndices(:,i),:);
+    h = waitbar(0, 'Sythesis Tree');
+    
+    for j=1:numleafs
+        
+        % find indices of samples which end up in this leaf
+        leafmodels(j).leaves = {find(node==leaf_index(j))}; %#ok<*AGROW>
+        
+        % mean prediction at this leaf
+        leafmodels(j).mean = model.Trees{i}.NodeMean(leaf_index(j));
+        
+        % the control variables sample values which contribute to this leaf (support)
+        leafmodels(j).xdata = {xTrain_c(leafmodels(j).leaves{1,1},:)};
+        
+        % the response variable value which contribute to this leaf
+        leafmodels(j).ydata = {yTrain(leafmodels(j).leaves{1,1})};
+        
+        % train a linear model
+        X = [ones(size(leafmodels(j).ydata{1},1),1), leafmodels(j).xdata{1}];
+        Y = leafmodels(j).ydata{1,1};
+        
+        Y_t = Y;
+        X_t = X(:,1:nxc*numb+1);
+        if size(X_t,1)<size(X_t,2)
+            % keyboard;
+            leafmodels(j).coeff = nan(nxc*numb+1,1);
+        else
+            options=optimset;
+            options.Algorithm='interior-point'; % to get rid of some warnings
+            options.Display='off';
+            Aineq = diag([0;-ones(nxc*numb,1)]);
+            bineq = zeros(nxc*numb+1,1);
+            lb=zeros(1,nxc*numb+1);
+            coeff = lsqlin(X_t,Y_t,Aineq,bineq,[],[],lb,[],[],options);
+            leafmodels(j).coeff = coeff;
+            leafmodels(j).xdata = [];
[…]

### 2017-09-17 code/updateModel.m

@@ -1,0 +1,32 @@
+function x = updateModel(power,roomT1,roomT2,roomT3,roomT4,d,u,xtot,model,leafmodels)
+
+Xtest = [d' u' xtot'];
+powerUpdate  = predict(power,Xtest);
+roomT1Update = predict(roomT1,Xtest);
+roomT2Update = predict(roomT2,Xtest);
+roomT3Update = predict(roomT3,Xtest);
+roomT4Update = predict(roomT4,Xtest);
+
+x = [roomT1Update;roomT2Update;roomT3Update;roomT4Update;powerUpdate];
+% idm = 1;
+% Xtest1 = [d' xtot(1) xtot(1)];
+% coeff = find_linearmodel_in_leaves(model{1,idm}, leafmodels{1,idm}, Xtest1);
+% roomT1Update = [1, u]*coeff;
+% 
+% Xtest2 = [d' xtot(2) xtot(2)];
+% coeff = find_linearmodel_in_leaves(model{2,idm}, leafmodels{2,idm}, Xtest2);
+% roomT2Update = [1, u]*coeff;
+% 
+% Xtest3 = [d' xtot(3) xtot(3)];
+% coeff = find_linearmodel_in_leaves(model{3,idm}, leafmodels{3,idm}, Xtest3);
+% roomT3Update = [1, u]*coeff;
+% 
+% Xtest4 = [d' xtot(4) xtot(4)];
+% coeff = find_linearmodel_in_leaves(model{4,idm}, leafmodels{4,idm}, Xtest4);
+% roomT4Update = [1, u]*coeff;
+% 
+% XtestP = [d' xtot(1) xtot(1) xtot(2) xtot(2) xtot(3) xtot(3) xtot(4) xtot(4)];
+% coeff = find_linearmodel_in_leaves(model{5,idm}, leafmodels{5,idm}, XtestP);
+% powerUpdate = [1, u]*coeff;
+% 
+% x = [roomT1Update;roomT2Update;roomT3Update;roomT4Update;powerUpdate];

### 2017-09-17 paper/abstract.tex
dumped CDC paper

Decisions on how to best operate large complex plants such as natural gas processing, oil refineries, and energy efficient buildings are becoming ever so complex that model-based predictive control (MPC) algorithms must play an important role. However, a key factor prohibiting the widespread adoption of MPC, is the cost, time, and effort associated with learning first-principles dynamical models of the underlying physical system. An alternative approach is to employ learning algorithms to build black-box models which rely only on real-time data from the sensors. Machine learning is widely used for regression and classification, but thus far data-driven models have not been used for closed-loop control. We present novel Data Predictive Control (DPC) algorithms that use Regression Trees and Random Forests for receding horizon control. 
%DPC also utilizes Gaussian Processes for smart experiment design by recommending how the accuracy of the current black-box models can be improved. 
We demonstrate the strength of our approach with a case study on a bilinear building model identified using real weather data and sensor measurements. 
In a one-to-one comparison, we show that DPC explains 70\% variation in the MPC controller.
%We validate our results by a one-to-one comparison against a benchmark MPC controller. 
We further apply DPC to a large scale multi-story EnergyPlus building model to curtail total power consumption in a Demand Response setting. 
In such cases, when the model-based controllers fail due to modeling cost, complexity and scalability, our results show that DPC curtails the desired power usage with high confidence.

### 2017-09-17 paper/casestudyBilinear.tex
dumped CDC paper

\section{COMPARISON WITH MPC}
\label{S:proof}

We consider a bilinear building model developed at Automatic Control Laboratory, ETH Zurich.
It captures the essential dynamics governing the zone-level operation while considering the external and the internal thermal disturbances. 
By Swiss standards, the model used for this study is of a heavyweight construction with a high window area fraction on one facade and high internal gains due to occupancy and equipments \cite{Gyalistras2010a}.

The bilinear model is a standard building model used for practical considerations \cite{Ma2015,Oldewurtel2011,Oldewurtel2012} as it is detailed enough and suitable for model-based control unlike the ones obtained from simulation software like EnergyPlus. We specifically consider this model to show a comparison against MPC. 
MPC of EnergyPlus models can be cost and time prohibitive, making them unsuitable for control. In Sec.~\ref{S:casestudy}, we show how DPC scales easily to such large scale models.

\subsection{Bilinear Model}
\label{SS:model}
\input{cs-model.tex}

\subsection{Model Predictive Control}
\label{SS:mpc}
\input{cs-mpc.tex}

\subsection{Data Predictive Control}
\label{SS:dpc}
\input{cs-dpc.tex}

\subsection{Comparison}
\label{SS:comp}
\input{cs-comparison.tex}

### 2017-09-17 paper/casestudyEPlus.tex
dumped CDC paper

\section{APPLICATION: DEMAND RESPONSE}
\label{S:casestudy}

In January 2014, the east coast (PJM) electricity grid experienced an 86x increase in the price of electricity from \$31/MWh to \$2,680/MWh in a matter of 10 minutes. Similarly, the price spiked 32x from an average of \$25/MWh to \$800/MWh in July of 2015. This extreme price volatility has become the new norm in our electric grids. Building additional peak generation capacity is not environmentally or economically sustainable. Furthermore, the traditional view of energy efficiency does not address this need for \emph{Energy Flexibility}. The solution lies with Demand Response (DR) from the customer side - curtailing demand during peak capacity for financial incentives. However, this is a very hard problem for commercial, industrial and institutional plants, the largest electricity consumers.
%the cost to model each building is over \$100K and buildings are all unique, they cannot decide which of the 100,000's of control knobs to turn as it is too complex, they must rely on rule-based curtailment approaches which are ad hoc, inefficient and do not provide any guarantees for energy reduction and are consequently exposed to a huge financial risk.

Thus, the problem of energy management during a DR event makes an ideal case for DPC. In the following sections, we apply DPC-En to a large scale EnergyPlus model to show how effectively DPC can provide a desired power curtailment as well as a desired thermal comfort. DPC builds predictive models of a building based on historical weather, schedule, set-points and electricity consumption data, while also learning from the actions of the building operator. These models are then used for synthesizing recommendations about the control actions that the operator needs to take, during a DR event, to obtain a given load curtailment while providing guarantees on occupant comfort and operations.

\subsection{EnergyPlus Model}
We use the DoE Commercial Reference Building (DoE CRB) simulated in EnergyPlus \cite{Deru2011} as the virtual test-bed building.
This is a large 6 story hotel building consisting of 22 zones with a total area of 122,120 sq.ft. 
During peak load conditions the building can consume up to 400 kW of power. 
For the simulation of the DoE CRB building we use actual meteorological year data from Chicago for the years 2012 and 2013.

\subsection{Model training for DPC}

In the following simulations, we consider a long DR event from 7am - 2pm when the end-users are expected to follow/track the reference power signal sent by the utility. This is indeed common in Demand Tracking Control. 
During offline training, we sample data every 15 min to learn 2 kinds of forests. (1) Power forests are built using output as the total building power consumption, and (2) Temperature forests with output as temperature of one of the 22 zones.

The training data set contains the following types of features. (1) The \textit{weather data} which includes measurements of the outside air temperature and relative humidity. Since we are interested in predicting the power consumption or the zone temperature for a finite horizon, we include the weather forecast of the complete horizon in the training features. (2) The \textit{schedule data} includes the proxy variables which correlate with repeated patterns of electricity consumption e.g., due to occupancy or equipment schedules. Day of Week is a categorical predictor which takes values from 1-7 depending on the day of the week. This variable can capture any power consumption patterns which occur on specific days of the week. Likewise, Time of Day is quite an important predictor of power consumption as it can adequately capture daily patterns of occupancy, lighting and appliance use without directly measuring any one of them. Besides using proxy schedule predictors, actual building equipment schedules can also be used as training data for building the trees. (3) The \textit{building data} include (i) cooling set points for  the guest rooms, kitchen and corridors, (ii) supply air temperature, and (iii) chilled water temperature.

For the following simulations, we use five control variables (i) cooling set point for corridors $\mathsf{ClgSP}$, (ii) cooling set point for guest rooms $\mathsf{GuestSP}$, (iii) cooling set point for kitchen $\mathsf{KitchenSP}$, (iv) chilled water supply temperature $\mathsf{ChwSP}$, and (v) supply air temperature  $\mathsf{SupplyAirSP}$, so $\tX^c = [\mathsf{ClgSP},\mathsf{GuestClgSP},\mathsf{KitchenClgSP},\mathsf{SupplyAirSP}, \mathsf{ChwSP}]$. 
The power forest $\mathcal{R}_p$ is built using the total building power consumption $\tP$. Its features $\tX^d$ include the weather variables, their lag terms and their forecast over the horizon, the schedule variables, and finally the lag terms of the power consumption.
The temperature forest $\mathcal{R}_t$ is built with zone temperature $\tT$ as the output. Except for the lag terms corresponding to the same zone temperature, all other features are same in $\tX^d$.

Fig.~\ref{F:sepvars} shows the prediction accuracy for the power forest, and also explains the two level training approach introduced in Sec.~\ref{SS:sepvar}. During S1, the forests are trained using only disturbances as the features. Then in S2, the local effects of the control variables are accounted for by the linear models in the leaves. We observe how the accuracy is drastically improved after including the linear models in the predictions.

\begin{figure}[h!]
	\centering
	\hspace{-0.3cm}\includegraphics[width=22pc]{figures/eplus_validation.eps}	
	\caption{Model accuracy during training: The prediction made by forest using only $\tX^d$ (red) captures the effect due to disturbances. The linear models in the leaves capture the local effects (green) due to the control inputs $\tX^c$ and improve the model accuracy.}
	\label{F:sepvars}
	\captionsetup{justification=centering}
\end{figure}

\subsection{Power Management}
\label{SS:powermanagement}
Typically, the end customer receives a notification to curtail the power by some fraction. 
In this example on power management, we show how DPC can generate optimal inputs to track a desired power signal within a small allowance while maintaining the zone level thermal comfort. It may not be possible to have the same thermal comfort level in all the zones due to power curtailment, so we choose one zone (for example CEO's office) where the constraints must be met.
This is done by solving the following optimization problem with control variables $\tX^c = [\mathsf{ClgSP},\mathsf{GuestClgSP},\mathsf{KitchenClgSP},\mathsf{SupplyAirSP}, \mathsf{ChwSP}]$ as defined before:
\begin{align}
\begin{aligned}
\text{min } & \ \ \ \ \ \ \ \sum_{j=1}^{N} ({\tP}_{\mathrm{k+j|k}} - \tP_{\mathrm{ref}})^2 +  \lambda\epsilon_j + \nu \delta_j \\
\text{s.~t. } & \ \ \ \ \ \tP_{\mathrm{k+j|k}} =  \hat{\Theta}^T_{\tP_j} [ 1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
& \ \ \ \ \ \tT_{\mathrm{k+j|k}} =  \hat{\Theta}^T_{\tT_j} [ 1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tP}-\epsilon_j \leq \tP_{\mathrm{k+j|k}} \leq \bar{\tP} + \epsilon_j\\
& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tT}-\delta_j \leq \tT_{\mathrm{k+j|k}} \leq \bar{\tT} + \delta_j\\
& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tX}^c \leq \tX^c_{\mathrm{k+j-1|k}} \leq \bar{\tX}^c\\ 
& \ \ \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0,\delta_j \geq 0, \ j = 1,\dots,N.
\end{aligned}
\label{E:ptrack}
\end{align}

\begin{figure}[h]
	\subfigure[Optimal inputs calculated by DPC-En. At first, the inputs are changed rapidly because of a significant difference between the desired and the actual power consumption. Then gradual adjustments are made to follow the desired reference.]{
		\label{F:eplus_inputs}
		\centering
		\includegraphics[width=22pc]{figures/eplus_control.eps}
	}
	\subfigure[Power tracking by DPC-En at 1.1 MW: The difference in closed-loop simulation and prediction is due to model mismatch.]{
		\label{F:eplus_power}
		\centering
		\includegraphics[width=22pc]{figures/eplus_closedloop.eps}
	}
	\caption{Power management using DPC. The controller is active between 7am - 2pm. This region is marked in dashed red lines.}
	\captionsetup{justification=centering}
	\label{F:eplus_track}
\end{figure}

Here, the temperature forests are used to enforce thermal constraints in the zone of interest. The setup of optimization problem is flexible to include even other variables in the cost or the constraints. For example, we are currently looking at including the dynamic pricing of electricity in the cost since the customers can more directly relate to the financial incentives.

The results are shown in Fig.~\ref{F:eplus_track}. 
The DPC controller is active between 7am - 2pm. Before 7am and after 2pm, the building is using a predefined rule-based control strategy.
The optimal control inputs from DPC-En are shown in Fig.~\ref{F:eplus_inputs}. It is observed that, with the optimal inputs generated by DPC, we can track the reference power consumption signal closely. In fact, the average tracking error between 7am - 2pm is 3\%. The difference between the predicted power consumption and that in the closed-loop simulation in Fig.~\ref{F:eplus_power} is due to model mismatch between the EnergyPlus model and the power forest used in the optimization \eqref{E:ptrack}. Due to this inaccuracy, the actual power consumption is on an average 7 kW higher.

Thus, DPC-En successfully tracks a given power reference signal with an average $\sim$ 3\% error for such a complex building which would require several years of efforts to develop a physics based model.

\subsection{Practical Challenges and Future Work}
\label{SS:challenges}
\input{cs-challenges.tex}

### 2017-09-17 paper/casestudyReal.tex
dumped CDC paper

--- 2017-09-16 paper/DPC_T&A.tex
@@
-\documentclass[review]{elsarticle}
-
-\usepackage{lineno,hyperref}
-\modulolinenumbers[5]
-
-\journal{Journal of \LaTeX\ Templates}
-
-%%%%%%%%%%%%%%%%%%%%%%%
-%% `Elsevier LaTeX' style
-\bibliographystyle{elsarticle-num}
-%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-
-\begin{frontmatter}
-
-\title{Data Predictive Control: Theory and Applications\tnoteref{mytitlenote}}
-\tnotetext[mytitlenote]{The research leading to these results has received funding from the Italian Government under Cipe resolution n.135 (Dec.
-21, 2012), project \emph{INnovating City Planning through Information and Communication Technologies} (INCIPICT).}
-
-\author[FSaddress,EC]{Francesco Smarra\corref{mycorrespondingauthor}}
-\cortext[mycorrespondingauthor]{Corresponding author}
-\ead{francesco.smarra@univaq.it@univaq.it}
-
-\author[AJaddress,EC]{Achin Jain}
-\ead{achinj@seas.unpenn.edu}
-\author[TDRaddress,EC]{Tullio De Rubeis}
-\ead{tullio.derubeis@graduate.univaq.it}
-\author[TDRaddress]{Dario Ambrosini}
-\ead{dario.ambrosini@ing.univaq.it}
-\author[FSaddress]{Alessandro D'Innocenzo}
-\ead{alessandro.dinnocenzo@univaq.it}
-\author[AJaddress]{Rahul Mangharam}
-\ead{rahulm@seas.upenn.edu}
-
-\address[EC]{Equal contribution}
-\address[FSaddress]{Department of Information Engineering, Computer Science and Mathematics, Università degli Studi dell’Aquila, L’Aquila, Italy}
-\address[TDRaddress]{Department of Industrial and Information Engineering and Economics, Università degli Studi dell’Aquila, L’Aquila, Italy}
-\address[AJaddress]{Department of Electrical and Systems Engineering, University of Pennsylvania, Philadelphia, USA}
-
-\begin{abstract}
-TBD
-\end{abstract}
-
-\begin{keyword}
-Energy efficiency, predictive control, machine learning.
-\end{keyword}
-
-\end{frontmatter}
-
-%\linenumbers
-
-\section{Introduction}
-
-\paragraph{Motivations} TBD
-
-\paragraph{State of the art} TBD
-
-\paragraph{Contributions} TBD
-
-\paragraph{Paper organization} TBD
-
-\section{Problem definition and objectives} TBD
-
-\section{Data Predictive Control using Random Forests} TBD
-
-\subsection{Modeling} TBD
-
-\subsection{Predictive Control} TBD
-
-\section{Case study 1: bilinear model} TBD
-
-\section{Case study 2: EnergyPlus model} TBD
-
 \section{Case study 3: real house} TBD
@@
 \paragraph{Energy-efficiency improvements}
-
-\section{Conclusion and future work} TBD
-
-\section*{References}
-
-\bibliography{mybibfile}
-
-\end{document}

### 2017-09-17 paper/conclusion.tex
dumped CDC paper

We present two algorithms based on trees and random forests for receding horizon control with data-driven models. 
We compare the performance of our Data Predictive Control to MPC on a multivariable bilinear building model. We establish that DPC with random forests shows a remarkable similarity to MPC in the optimal control strategies explaining 70\% variance. On the other hand, DPC with regression trees suffers from practical limitations due to model overfitting.
We further apply DPC with random forests to a large scale 6 story EnergyPlus model with 22 zones for which the traditional model-based control is largely unsuitable due to complex dynamics and the cost of model identification. We show that DPC, relying only on the sensor data, can provide significant energy savings while maintaining thermal comfort. Our results demonstrate that even for such complex system, DPC tracks a reference signal with a mean error of 3\%.

DPC has applications which go beyond buildings and energy systems, to industrial process control, and controlling large critical infrastructures like water networks, district heating \& cooling. DPC is immensely valuable in situations where first principles based modeling cost is extremely high.

### 2017-09-17 paper/cs-challenges.tex
dumped CDC paper

\textit{Data Availability:} The main practical challenge for DPC lies in the availability of data for training and we require answers to questions like how much data (functional testing) is required, and how should the sampling be done? Therefore, the procedure for optimal experiment design, and model improvement with estimation of variance in predictions is one of the main focus of our ongoing work.

\textit{Stability:} While the buildings are inherently stable, many other applications require stability guarantees. In our ongoing work, we are working towards proving asymptotic stability to origin with DPC-RT and DPC-En by using concept of switched LTI systems. This will make DPC useful for systems with faster dynamics.

\textit{Robustness:} Another direction of work is on handling uncertainties in the DPC framework, namely an extension to Scenario DPC to account for the disturbance uncertainty. This will help us in quantifying the robustness of DPC.

### 2017-09-17 paper/cs-comparison.tex
dumped CDC paper

We compare the performance of DPC \eqref{E:dpc} against an equivalent MPC formulation \eqref{E:mpc}. The solution obtained from MPC sets the benchmark that we compare to. Note that the MPC implementation uses the exact knowledge of the plant dynamics. Therefore, the associated control strategy is indeed the optimal strategy for the plant.

The performance is compared for 3 days in winter, i.e. January 28-31 and 3 days in summer, i.e. May 1-3. These are shown on the same plots in Fig.~\ref{F:comparison}.
The sampling time in the simulations is 1 hr. The control horizon $N$ and the order of autoregression are both 6 hrs. The training procedure required a few minutes in the case of trees and 2 hrs for forests on a Win 10 machine with an i7 processor and 8GB memory.
The cooling usage factor $\mathsf{C}$ is constrained in $[0,1]$, the heat input in $[0,23]\ \mathrm{W/m^2}$, and the room temperature in $[19,25]\ \mathrm{^oC}$ during the winters and $[20,26]\ \mathrm{^oC}$ during the summers.
The optimization is solved using CPLEX \cite{IBM}.

The external disturbances - solar gains, internal gain due to equipment and dry-bulb temperature during the chosen periods are shown in Fig.~\ref{F:dist}. The internal gain due to occupancy was proportional to the gain due to equipment. 
The reference temperature is chosen to be 22 $\mathrm{^oC}$. Due to cold weather, which is evident from the dry-bulb temperature, the heater is switched on during the night to maintain the thermal comfort requirements. When the building is occupied during the day, due to excessive internal gains, the building requires cooling. The lighting in the building is adjusted to meet the minimum light requirements.
The optimal cooling usage factor and the radiator power for MPC, DPC-En and DPC-RT are shown in Fig.~\ref{F:control1} and Fig.~\ref{F:control2}, respectively. The control strategy with DPC-En shows a remarkable similarity to MPC, switching on/off the equipments at the same time with similar usage. However, the performance with DPC-RT is much different and worse. DPC-RT inherently suffers from high variance which is also evident in the control strategy, thus making it unsuitable for practical purposes. 
Although it seems like that adding the rate constraints to DPC-En would smoothen its behavior, this was avoided because the sampling time of the system is 1 hr which is already too high. The room temperature profile in Fig.~\ref{F:state} is close to the reference in the case of DPC-En as well as MPC. 
Fig.~\ref{F:obj} shows that the cumulative cost of the objective function is, as expected, minimum for MPC, and a bit higher for DPC-En. The cost for DPC-RT blows up around 12 noon on 30$^{th}$ January as one of the slack variables is non-zero, which happens due to high model inaccuracy.

The quantitative performance comparison is shown in Tab.~\ref{T:comparison}. MPC tracks the reference more closely at the expense of higher input costs in comparison to DPC-En. The higher cost of the inputs in MPC is also due to lighting. DPC-En explains 70.1\% variation in the optimal control strategies obtained from MPC while DPC-RT explains only 1.8\%. The mean optimal cost of DPC-En is more than MPC, and is maximum for DPC-RT due to a constraint violation.

\begin{table}[h!]
	\centering
	\caption{Quantitative comparison of explained variance, mean value of objective function, mean input cost $c^Tu$ and mean deviance from the reference temperature $|\mathsf{T}-\mathsf{T}_{\mathrm{ref}}|$.}
	\captionsetup{justification=centering}
	\begin{tabular}{c|c|c|c|c}
		\toprule
		& explained & mean objective& mean input  & mean  \\
		&  variance$[\mathrm{-}]$ & value $[\mathrm{-}]$ & cost $[-]$ & deviance $[\mathrm{^oC}]$ \\     
		\midrule
		MPC    &  $\mathrm{-}$ &  22.60 & 17.16  &  0.26  \\
		DPC-En   & 70.1\% &  39.26  & 15.12 &  0.48 \\
		DPC-RT  & 1.8\% & 204.55 & 16.84 &  0.57 \\
		\bottomrule
	\end{tabular}
	\label{T:comparison}
\end{table}
\begin{figure}[h!]
	\subfigure[External disturances: solar gains, internal gain due to equipment and dry-bulb temperature.]{
		\label{F:dist}
		\centering
		\includegraphics[width=20pc]{figures/disturbances.eps}
	}
	\subfigure[Optimal control input: cooling usage factor $\mathsf{C}$ with  $0 \leq \mathsf{C} \leq 1$. DPC-En generates a control strategy very simular to MPC.]{
		\label{F:control1}
		\centering
		\includegraphics[width=20pc]{figures/input3.eps}
	}
	\subfigure[Optimal control input: radiator heat $\mathsf{H}$ with  $0 \leq \mathsf{H} \leq 23 \ \mathrm{W/m^2}$. Again, DPC-En generates a control strategy very simular to MPC.]{
		\label{F:control2}
		\includegraphics[width=20pc]{figures/input4.eps}
	}
	\subfigure[Room temperature has time varying bounds. When the building is occuped the constraints are relaxed, else $19(20) \leq \mathsf{T}_{\mathrm{in}} \leq 25(26) \ \mathrm{^oC}$ in January(May). MPC and DPC-En are able to track the reference temperature ($22 \ \mathrm{^oC}$) closely.]{
		\label{F:state}
		\centering
		\includegraphics[width=20pc]{figures/state.eps}
	}
	\subfigure[Cumulative optimal cost after solving optimization. MPC serves as the benchmark with the minimum cost, followed by DPC-En and then DPC-RT.]{
		\label{F:obj}
		\centering
		\includegraphics[width=20pc]{figures/cumsumcost.eps}
	}
	%\vspace{-1cm}
	\caption{Comparison of optimal performance obtained with MPC, DPC-En and DPC-RT for 3 days in January and 3 days in May.}
	\label{F:comparison}
	\captionsetup{justification=centering}
\end{figure}

Thus, we have shown that DPC-En provides a comparable performance to MPC without using the physical model.
However, one major limitation of the bilinear model is that the information about the building power consumption is not available. Much nonlinearities in the system are due to equipment efficiencies which are not considered in the bilinear case but are very important for practical purposes.

Therefore, our next goal is to apply DPC-En on even more complex and realistic EnergyPlus model for which building a model predictive controller is time and cost prohibitive \cite{Sturzenegger2016}. This is because we would need to model intricate details like the geometry and construction layouts, the equipment design and layout plans, material properties, equipment and operational schedules etc.

### 2017-09-17 paper/cs-dpc.tex
dumped CDC paper

In this section, we explain how DPC can be applied to this case study. We begin with a description of features $\tX$ and output $\tY$ used for training.

\subsubsection{Training Data} 
\label{SSS:dpc_data}

The fundamental reason why DPC  is suitable for such a problem is that when the complexity rises, there is a huge cost to model all the states given by the dynamical system \eqref{E:bilinear1}. For example, states in the bilinear model also include slab temperatures which require modeling of structural and material properties in detail and often we also need to install new sensors to capture additional states. Thus, DPC is based solely on one state of the model i.e. the zone temperature that can be easily measured with a thermostat. This serves as the output variable $\tY$ of interest for which we build $N$ trees and $N$ forests as described in Sec.~\ref{SS:dpcrt} and \ref{SS:dpcrf}, respectively. Therefore, $\tY_{\mathrm{k+j|k}}:=x_{\mathrm{k+j|k}}^1$, where $x^1$ is the first component of $x$.
Next, we define the non-manipulated features $\tX^d_{\mathrm{k|k}}$. At time $k$, for the tree $\mathcal{T}_j$ and the forest $\mathcal{R}_j$, we base these features to include weather disturbances, external disturbances due to occupancy and equipments, and autoregressive terms of the room temperature, i.e.
$\tX^d_{\mathrm{k|k}}:=[ d_{\mathrm{k+j-N|k}} ,\dots,d_{\mathrm{k+j-1|k}}, x_{\mathrm{k|k}}^1,\dots, x_{\mathrm{k-\delta|k}}^1 ]$, where $\delta$ is the order of autoregression.
Finally, the inputs in DPC are exactly same as in MPC. i.e. $\tX^c_{\mathrm{k+j-1|k}}:=u_{\mathrm{k+j-1|k}}$.

The training data in the above format was generated by simulating the bilinear model with rule-based strategies for 10 months in 2007. January and May were deliberately excluded for testing the DPC implementation.

\subsubsection{Optimization} 
\label{SSS:dpc_opt}
For a fair comparison with MPC, we cast DPC optimization problem as follows:
\begin{subequations}
\begin{align}
\text{min } & \sum_{j=1}^{N} {\tY}_{\mathrm{k+j|k}} \mathcal{Q}^{(1,1)} {\tY}_{\mathrm{k+j|k}} + c^T \tX^c_{\mathrm{k+j-1|k}}+  \lambda\epsilon_j\\
\text{s.~t. } & \ \ \ \ \ \tY_{\mathrm{k+j|k}} =  \alpha_j^T \left[1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} \right]^T \label{SE:dpc}\\
& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tX}^c \leq \tX^c_{\mathrm{k+j-1|k}} \leq \bar{\tX}^c\\ 
& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tY}-\epsilon_j \leq \tY_{\mathrm{k+j|k}} \leq \bar{\tY} + \epsilon_j\\\
& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0, \ j = 1,\dots,N.
\end{align} \label{E:dpc}
\end{subequations}
\noindent Here $\alpha = \beta$ for DPC-RT and $\alpha = \hat{\Theta}$ for DPC-En.
Note that, \eqref{E:dpc} is DPC analog of \eqref{E:mpc}. The only difference is the state dynamics \eqref{SE:mpc1} and \eqref{SE:mpc2} are now replaced with \eqref{SE:dpc}.

\subsubsection{Validation} 
\label{SSS:dpc_val}

We compare the prediction for the first time step $\tY_{\mathrm{k+1|k}}$ and the 6-hour ahead prediction $\tY_{\mathrm{k+6|k}}$ for a week in the month of May in Fig.~\ref{F:validation}. It is visible how trees have a high variance, and the forests are more accurate. Note that data from January and May was not used for training. The quantitative summary of the accuracy is given in Tab.~\ref{T:validation}. We can see that the random forests are better in all respects.
\begin{table}[h!]
	\centering
	\caption{Quantitative comparison of root mean square error (RMSE), $R^2$ score, and explained variance (EV) for trees and forests for different predictions steps.}
	\captionsetup{justification=centering}
	\begin{tabular}{c|c|c|c}
		\toprule
		& RMSE & $R^2$ score & EV  \\ 
		\midrule
		tree-$\tY_{\mathrm{k+1|k}}$    &  0.42 &  0.75 & 0.76    \\
		tree-$\tY_{\mathrm{k+6|k}}$  & 0.64 &  0.41  & 0.42 \\
		forest-$\tY_{\mathrm{k+1|k}}$  & 0.29 & 0.87  & 0.88 \\
		forest-$\tY_{\mathrm{k+6|k}}$  & 0.38 & 0.78 & 0.80 \\
		\bottomrule
	\end{tabular}
	\label{T:validation}
\end{table}

\begin{figure}[h!]
	\centering
	\includegraphics[width=20pc]{figures/validation-s1.eps}
	\includegraphics[width=20pc]{figures/validation-s6.eps}
	\caption{Temperature predictions from a tree and a forest for first step prediction (top) and the 6-hour ahead prediction (bottom). Ensemble method shows a relatively higher accuracy.}
	\captionsetup{justification=centering}
	\label{F:validation}
\end{figure}

%\begin{figure}[t!]
%	\centering
%	\includegraphics[width=20pc]{figures/validation-s6.eps}
%	\caption{Temperature predictions from a tree and a forest for 6 hour ahead prediction.}
%	\captionsetup{justification=centering}
%	\label{F:validation-s6}
%\end{figure}

### 2017-09-17 paper/cs-model.tex
dumped CDC paper

%The building is treated as a lumped-parameter system, which assumes nodes in the room as well as in the walls, floor, and ceiling describing the respective temperatures.
%Considering the heat transfer rate between the nodes, an equivalent thermal RC network is identified. 
%The influence of the actuators is modeled according to their heat transfer properties, e.g. the heating input from the radiator goes directly into the room node, whereas floor heating affects the room with some delay and therefore the control input from floor heating goes into the node in the floor.

The bilinear model has 12 internal states including the inside zone temperature $\mathsf{T}_{\mathrm{in}}$, the slab temperatures $\mathsf{T}_{\mathrm{sb}}$, the inner wall $\mathsf{T}_{\mathrm{iw}}$ and the outside wall temperature $\mathsf{T}_{\mathrm{ow}}$. The state vector is defined as $x:=[\mathsf{T}_{\mathrm{in}}, \mathsf{T}_{\mathrm{sb}}^{(1:5)}, \mathsf{T}_{\mathrm{ef}}^{(1:3)}, \mathsf{T}_{\mathrm{in}}^{(1:3)}]^T$.

There are 4 control inputs including the blind position $\mathsf{B}$, the gains due to electric lighting $\mathsf{L}$, the evaporative cooling usage factor $\mathsf{C}$, and the heat from the radiator $\mathsf{H}$ such that $u:=[\mathsf{B},\mathsf{L},\mathsf{H},\mathsf{C}]^T$. $\mathsf{B}$ and $\mathsf{L}$ affect both room illuminance and temperature due to heat transfer whereas $\mathsf{C}$ and $\mathsf{H}$ affect only temperature.

The model is subject to 5 weather disturbances: solar gains with fully closed blinds $\mathsf{Q}_{\mathrm{sc}}$ and with open blinds $\mathsf{Q}_{\mathrm{so}}$, daylight illuminance with open blinds $\mathsf{I}_{\mathrm{o}}$, external dry-bulb temperature $\mathsf{T}_{\mathrm{db}}$ and external wet-bulb temperature $\mathsf{T}_{\mathrm{wb}}$. 
The hourly weather forecast, provided by MeteoSwiss, was updated every 12 hrs. Therefore, to improve the forecast,  an autoregressive model of the uncertainty was considered.
Other disturbances come from the internal gains due to occupancy $\mathsf{Q}_{\mathrm{io}}$ and due to equipments $\mathsf{Q}_{\mathrm{ie}}$ which were assumed as per the Swiss standards \cite{Merkblatt2006}. We define $d:=[\mathsf{Q}_{\mathrm{sc}},\mathsf{Q}_{\mathrm{so}},\mathsf{I}_{\mathrm{o}},\mathsf{Q}_{\mathrm{io}},\mathsf{Q}_{\mathrm{ie}},\mathsf{T}_{\mathrm{db}},\mathsf{T}_{\mathrm{wb}}]^T$. For further details, we refer the reader to \cite{Oldewurtel2011}.

The model dynamics is given below. The bilinearity is present in both input-state, and input-disturbance.
\begin{gather}
\label{E:bilinear1}
x_{k+1} = Ax_{k}+(B_u +B_{xu}[x_k] + B_{du}[d_k]) u_k+B_dd_k \\
x_{k} \in \mathbb{R}^{12}, u_{k} \in \mathbb{R}^{4}, d_{k} \in \mathbb{R}^{8} \ \forall k = 0,\dots,T, \nonumber
\end{gather}
where, the matrices $B_{xu}$ and $B_{du}$ are defined as
\begin{gather}
\label{E:bilinear2}
B_{xu}[x_k] = [ B_{xu,1}[x_k],B_{xu,2}[x_k], \dots, B_{xu,4}[x_k] ] \in \mathbb{R}^{12\times4}, \nonumber\\
B_{du}[d_k] = [ B_{du,1}[d_k],B_{du,2}[d_k], \dots, B_{du,4}[d_k] ] \in \mathbb{R}^{12\times4}\nonumber, \\
B_{xu,i} \in \R^{12\times12}, B_{du,i} \in \R^{12\times8} \ \forall i=1,2,3,4.\nonumber
\end{gather}
For this study, we assume that the disturbances are precisely known to MPC as well as DPC controller. In our future work, we will account for the uncertainties in the disturbances with an extension to Scenario approach \cite{Bernardini2009} for DPC.

\begin{figure}[t!]
	\centering
	\includegraphics[width=21pc]{figures/dpc-algo-rf.eps}
	\caption{DPC-En: At time $k$, the algorithm uses the forecast of disturbances $\tX^d_{\mathrm{k|k}}$ to select linear models $\Theta_1$ to $\Theta_t$ in the leaves of each ensemble. The linear models in each ensemble are averaged to calculate a single model represented by $\hat{\Theta}_j$ which act as constraints in the optimization problem. The optimal sequence $[\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+N-1|k}}]$, of which the first one is applied, and $\tX^d_{\mathrm{k+1|k+1}}$ is calculated to proceed to $k+1$.}
	\label{F:dpc-algo-rf}
\end{figure}

### 2017-09-17 paper/cs-mpc.tex
dumped CDC paper

We use an MPC controller with a quadratic and a linear cost for comparison.
The finite RHC approach involves optimizing a cost function subject to the dynamics of the system and the constraints, over a finite horizon of time \cite{Mayne2000}. After an optimal sequence of control inputs are computed, the first input is applied, then at the next step the optimization is solved again.

%\begin{figure}
%\centering
%\includegraphics[scale=0.8]{figures/mpc_horizon.eps}
%\caption{Finite-horizon moving window of MPC: at time $k$, the MPC optimization problem is solved for a finite length window of $N$ steps and the first control input $u_k$ is applied; the window then recedes one step forward and the process is repeated at time $k+1$.}
%\captionsetup{justification=centering}
%\label{F:mpc}
%\end{figure}

The objective of the controller is to minimize the energy usage $c^Tu$ while maintaining a desired level of thermal comfort.
Therefore, at time step $k$, we solve a continuously linearized MPC problem to determine the optimal sequence of inputs $[u_{\mathrm{k|k}},\dots,u_{\mathrm{k+N-1|k}}]$:
\begin{subequations}
\begin{align}
\text{min } & \ \ \ \ \sum_{j=1}^{N} {x}^T_{\mathrm{k+j|k}} \mathcal{Q} {x}_{\mathrm{k+j|k}} + c^Tu_{\mathrm{k+j-1}} +  \lambda\epsilon_j\\
\text{s.~t. } & \ \ x_{\mathrm{k+j|k}} =  Ax_{\mathrm{k|k}} + B u_{\mathrm{k+j-1|k}} + B_d d_{\mathrm{k+j-1|k}} \label{SE:mpc1} \\
& \ \ \ \ \ B = B_u + B_{xu}[x_{\mathrm{k|k}}] + B_{du}[d_{\mathrm{k+j-1|k}}] \label{SE:mpc2}\\
& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{u} \leq u_{\mathrm{k+j-1|k}} \leq \bar{u}\\ 
& \ \ \ \ \ \ \ \ \ \ \ \ \underline{x}-\epsilon_j \leq x_{\mathrm{k+j|k}} \leq \bar{x} + \epsilon_j\\\
& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0, \ j = 1,\dots,N,
\end{align}\label{E:mpc}
\end{subequations} 
\noindent where $\mathcal{Q} \in \R^{12 \times 12}$ has all zeros except at $\mathcal{Q}^{(1,1)}$ corresponding to the zone temperature, $c \in \R^{4}$ is proportional to cost of using each actuator and $\lambda$ penalizes the slack variables.

### 2017-09-17 paper/dpc.tex
dumped CDC paper

\section{DATA PREDICTIVE CONTROL}
\label{S:dpc}

The central idea behind DPC is to obtain control-oriented models using machine learning or black-box modeling, and formulate the control problem in a way that receding horizon control (RHC) can still be applied and the optimization problem can be solved efficiently.

Consider a black-box model given by $x_{k+1}=f(x_k,u_k,d_k)$, where $x,u,d$ represent states, inputs and disturbances, respectively. Depending upon the learning algorithm, $f$ is typically nonlinear, nonconvex and sometimes nondifferentiable (as is the case with regression trees and random forests) with no closed-form expression. 
Such functional representations learned through black-box modeling may not be directly suitable for control and optimization as the optimization problem can be computationally intractable, or due to nondifferentiabilities we may have to settle with a sub-optimal solution using evolutionary algorithms \cite{Kusiak2009}.
These problems can be eliminated by decomposing 
\begin{align}
f(x_k,u_k,d_k)= g(d_k,x_k,h(u_k)),
\label{E:sepvars}
\end{align}
where both $g$ and $h$ are learned using the data, and $h(u_k)$ is convex and differentiable, and thus suitable for optimization. DPC uses this functional decomposition or \textit{separation of variables} to overcome the aforementioned challenges with black-box optimization.

%First, as the receding horizon involves predicting states into the future, with a black-box model trained on control variables as features, the optimization problem can be computationally intractable as the functions are typically nonlinear, nonconvex and do not have a closed-form expression.
%Second, black-box models like trees and random forests are not suitable for optimization because the gradient is not defined, so we may have to settle with a sub-optimal solution using evolutionary algorithms \cite{Kusiak2009}. 
%Third, interpretability is a desirable characteristic of a predictive model which only a few black-box models satisfy.
%DPC uses regression trees and tree ensembles (in particular random forests) for closed-loop RHC which overcome both the challenges by using \textit{separation of variables} which we explain next.
%Regression trees are interpretable by design. 
%The remaining problems are solved by using \textit{separation of variables} which we explain next.

\subsection{Separation of Variables}
\label{SS:sepvar}
%Consider a multivariable dynamical system subject to external disturbances. 
We distinguish between two sets of variables: control (or manipulated) variables $\tX^c \in \R^c$ and disturbance (or non-manipulated) variables $\tX^d \in \R^d$. The union of the two sets forms the full feature set for training, i.e. $\tX \equiv \tX^c \cup \tX^d \in \R^{c+d}$.
Our goal is to replace a model-based controller with a data-driven controller, where the latter depends only on the historical sensor data. 
These measurements could directly represent one or more states in the model-based control framework. We denote these as outputs $\tY \in \R$ for training, i.e. a $\tY$ represents a particular output and we can have separate models for multiple outputs. We define the number of training samples by $|(\tX,\tY)| = n$.

Using separation of variables, the training process is divided into two steps. \textit{Step 1:} The trees and the ensembles are trained only on $\tX^d$, which eases the computational complexity. It is important to note that besides external disturbances, $\tX^d$ also contains autoregressive terms of the output $\tY$ which is the main reason for the state space explosion. \textit{Step 2:} Linear regression models are trained in the leaves (or terminal nodes) of the trees which are function of only $\tX^c$. We have validated this linear model assumption in \cite{Jain2016}. As we shall see in Sec.~\ref{SS:dpcrt} and \ref{SS:dpcrf}, the second step reduces the run-time control problem into a convex program. This process is illustrated in Fig.~\ref{F:dpc-sepvars}.
\begin{figure}[t!]
	\centering
	\includegraphics[width=20pc]{figures/dpc-sepvars.eps}
	\caption{Separation of variables. \textit{Step 1:} Tree $\mathcal{T}_1$ is trained only on the disturbances $\tX_d$ as the features. Tree $\mathcal{T}_2$ uses both the disturbances $\tX_d$ and the control variables $\tX_c$ for splitting and is thus not computationally suitable for control. \textit{Step 2:} In the leaf $R_i$ of the trees, a linear regression model parametrized by $\beta_i$ is defined as a function only of the control variables.}
	\captionsetup{justification=centering}
	\label{F:dpc-sepvars}
\end{figure}

\subsection{DPC-RT: DPC with Regression Trees}
\label{SS:dpcrt}
When the data has lots of features, which interact in complicated, nonlinear ways, assembling a single global model such as linear or polynomial regression can be difficult, and can lead to poor response predictions.
An approach to non-linear regression is to partition the data space into smaller regions, where the interactions are more manageable. 
This partition is repeated recursively until finally we get to small chunks of the data space where we can fit simple (eg. linear parametric) models. 
Therefore, in \eqref{E:sepvars}, the global model $f$ has two parts: the recursive partition $g$, and a linear (and convex) model $h$ for each cell of the partition.
%The algorithm for learning regression trees (CART) is described in~\cite{Breiman1984}.

Now, our goal is to predict the state $\tY$ at time $k$ for next $N$ time steps, i.e. $\tY_{\mathrm{k+1|k}},\dots,\tY_{\mathrm{k+N|k}}$, where $N$ is the control horizon. Applying the separation of variables, we build $N$ regression trees using CART procedure \cite{Breiman1984} such that the output $\tY_{\mathrm{k+j|k}}$ of the $j^{th}$ tree depends upon the previous $N$ disturbances:
\begin{gather}
\label{E:model_tree}
\tY_{\mathrm{k+j|k}} = \mathit{f}_{\mathrm{tree}} \left( \tX^d_{\mathrm{k+j-N|k}} ,\dots,\tX^d_{\mathrm{k+j-1|k}}  \right), \\ \nonumber
\mathsf{X}^d_{\mathrm{k+j-l|k}} \in \R^d \  \ \forall \ l,j=1,\dots,N.
\end{gather}
Then, the linear models as functions of $\tX^c$ in each leaf of the tree $\mathcal{T}_j$ are defined as
\begin{gather}
\label{E:model_leaf}
\tY_{\mathrm{k+j|k}} =  \beta^T_j [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T, \\ \nonumber
\mathsf{X}^c_{\mathrm{k+j-l|k}} \in \R^c \  \ \forall \ l,j=1,\dots,N.
\end{gather}
Note that the coefficients $\beta_j$ would be different for each leaf. Eq.~\eqref{E:model_leaf} implies that the prediction of output $\tY_{\mathrm{k+j}}$ at time $k$ is an affine combination of control inputs from time $k$ to $k+j-1$. Thus, we have managed to linearize the original model dynamics via black-box modeling. This two-step training is done offline. In run-time, given the disturbances $\tX^d_\mathrm{k|k}$ at time $k$, we can narrow down to a leaf of each tree in \eqref{E:model_tree} to retrieve the linear models in \eqref{E:model_leaf}.

In run-time, when a new control action is to be determined, each tree (prediction step) contributes to a linear constraint in the optimization as a replacement for the state dynamics in the case of MPC. Thus, the RHC optimization problem with a quadratic cost ($\mathcal{Q} \geq 0, \mathcal{R} \succeq 0$) can be formulated as:
\begin{align}
\begin{aligned}
\text{min } & \sum_{j=1}^{N} ({\tY}_{\mathrm{k+j|k}})^2 \mathcal{Q} + {\tX^c}^T_{\mathrm{k+j-1|k}} \mathcal{R} {\tX^c}_{\mathrm{k+j-1|k}} +  \lambda\epsilon_j\\
\text{s.~t. } & \ \ \ \ \ \tY_{\mathrm{k+j|k}} =  \beta^T [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tX}^c \leq \tX^c_{\mathrm{k+j-1|k}} \leq \bar{\tX}^c\\ 
& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tY}-\epsilon_j \leq \tY_{\mathrm{k+j|k}} \leq \bar{\tY} + \epsilon_j\\\
& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0, \ j = 1,\dots,N.
\end{aligned}
\label{E:dpcrt}
\end{align}
Here, $\mathcal{Q} \in \mathbb{R}$ and $\mathcal{R} \in \mathbb{R}^{c \times c}$, and the slack variables $\epsilon_j$ ensure recursive feasibility since the equality constraint on $\tY$ is relaxed. Of course, a different cost function can be chosen depending upon the application. In the current formulation, the data-driven control problem is reduced to a convex program which is much easier to solve than running an optimization directly on a black-box model trained on $\tX^c$ as features. We solve this optimization in the same manner as MPC to determine the optimal sequence of inputs $[\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+N-1|k}}]$, apply the first control input $\tX^c_{\mathrm{k|k}}$ and proceed to the next time step $k+1$. The pseudo code for DPC-RT is given in Alg.~\ref{A:dpcrt}.

\subsection{DPC-En: DPC with Ensemble Methods}
\label{SS:dpcrf}
Regression trees obtain good predictive accuracy in many domains. However, the models used in their leaves have some limitations regarding the kind of functions they are able to approximate.
The problem with trees is their high variance and that they can overfit the data easily.
A small change $\delta$ in the data can result in a different series of splits and thus violate the acceptable accuracy $\eps$, i.e. $\exists \  \hat{\tX}^d \ | \ ||\tX^d-\hat{\tX}^d|| < \delta$ \& $||\tY-\tY_{\mathrm{true}}||>\eps$.
This is the price to be paid for estimating a tree-based structure from the data.
%The main reason behind this is the hierarchical nature of the process: the effect of an error in the top split is propagated down to all of the splits below it.
%While pruning and cross validation can help reduce the over fitting partially, we use ensemble methods for growing more accurate trees.

We use ensemble methods \cite{Friedman2001} to combine the predictions of several independent regression trees in order to improve generalizability and robustness over a single estimator. 
%Random forests or \textit{tree-bagging} are a type of ensemble method which makes predictions by averaging over the predictions of several independent trees.
The essential idea is to average many noisy trees to reduce the overall variance in prediction.
We inject randomness into the tree construction in two ways. First, we randomize the features used to define splitting in each tree.
Second, we build each tree using a bootstrapped or sub-sampled data set.
In this way, each tree in the forest is trained on different data, which introduces differences between the trees. More explicitly, training features $\tX^d \in \R^p$ with $p<d$ and the in-bag samples (in-bag samples correspond to the data samples on which the tree was trained) are different for each tree in the forest i.e $|(\tX,\tY)|<n$.

The goal with DPC-En is to replace each tree in Alg.~\ref{A:dpcrt} by a forest
\begin{gather}
\label{E:model_forest}
\tY_{\mathrm{k+j|k}} = \mathit{f}_{\mathrm{forest}} \left( \tX^d_{\mathrm{k+j-N|k}} ,\dots,\tX^d_{\mathrm{k+j-1|k}}  \right), \\ \nonumber
\mathsf{X}^d_{\mathrm{k+j-l|k}} \in \R^p \  \ \forall \ l,j=1,\dots,N,
\end{gather}
which, again, is trained only on $\mathsf{X}^d$, but $\tX^d \in \R^p \subset \R^d$ for each tree,  and then fit a linear regression model using $\mathsf{X}^c$ in every leaf of every tree. We build $N$ such forests for $N$ prediction steps such that the leaf $R_i$ of forest $\mathcal{R}_j$ uses a linear model
\begin{gather}
\label{E:model_leaf_forest}
\tY_{\mathrm{k+j|k}} =  \Theta^T_{ij} [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T, \\ \nonumber
\mathsf{X}^c_{\mathrm{k+j-l|k}} \in \R^c \  \ \forall \ l,j=1,\dots,N.
\end{gather}
Here $(\mathsf{X}^c,\mathsf{Y})$ correspond to the in-bag samples for the trees.

\begin{algorithm}[t!]
	\caption{Data Predictive Control with Regression Trees}
	\label{A:dpcrt}
	\begin{algorithmic}[1]
		\State \textsc{Design Time}
		\Procedure{Model Training using Separation of Vars}{}
		\State Set $\mathsf{X}^c$ $\gets$ manipulated features
		\State Set $\mathsf{X}^d$ $\gets$ non-manipulated features
		\State Build $N$ predictive trees with $(\tY,\tX^d)$ defined in \eqref{E:model_tree}
		\ForAll{trees $\mathcal{T}_j$}
		\ForAll{regions $R_i$ at the leaves of $\mathcal{T}_j$}
		\State Fit $ \tY_{\mathrm{k+j|k}} =  \beta^T_j \left[1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} \right]^T$ as in \eqref{E:model_leaf}
		\EndFor
		\EndFor
		\EndProcedure
		\State \textsc{Run Time}
		\Procedure{Predictive Control}{}
		\While{$k< k_{\mathrm{stop}}$}
		\ForAll{trees $\mathcal{T}_j$}
		\State Determine the leaf $R_i$ using $\tX^d$ as in \eqref{E:model_tree}
		\State Obtain the linear model at $R_{i}$ trained in \eqref{E:model_leaf}
		\EndFor
		\State Solve optimization in \eqref{E:dpcrt} to determine optimal
		\State control actions $[\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+N-1|k}}]$
		\State Apply the first input $\tX^c_{\mathrm{k|k}}$
		\EndWhile
		\EndProcedure
	\end{algorithmic}
\end{algorithm}

While the offline training burden in DPC-En is slightly increased compared to DPC-RT, in the control step we exploit the better accuracy, and lower variance properties of the random forest. 
If a forest has $t$ number of trees, given the forecast of disturbances, we have $t$ sets of linear coefficients. We simply average out all the coefficients from all the trees to get one linear model represented by $\hat{\Theta}_j$ for each forest. Note that the averaging step can only be done in run-time because the leaf of each tree can be narrowed down only when the $\tX^d$ is known. Thus, for $N$ forests, we again have exactly $N$ linear equality constraints in the optimization problem below:
\begin{align}
\begin{aligned}
\text{min } & \sum_{j=1}^{N} ({\tY}_{\mathrm{k+j|k}})^2 \mathcal{Q} + {\tX^c}^T_{\mathrm{k+j-1|k}} \mathcal{R} {\tX^c}_{\mathrm{k+j-1|k}} +  \lambda\epsilon_j\\
\text{s.~t. } & \ \ \ \ \ \tY_{\mathrm{k+j|k}} =  \hat{\Theta}^T_j [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tX}^c \leq \tX^c_{\mathrm{k+j-1|k}} \leq \bar{\tX}^c\\ 
& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tY}-\epsilon_j \leq \tY_{\mathrm{k+j|k}} \leq \bar{\tY} + \epsilon_j\\\
& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0, \ j = 1,\dots,N.
\end{aligned}
\label{E:dpcrf}
\end{align}
DPC-En is graphically described in Fig.~\ref{F:dpc-algo-rf}. The ensemble data predictive control (DPC-En) is the first such method to bridge the gap between ensemble predictive models (such as random forests) and receding horizon control. In the next section, we compare DPC-RT and DPC-En to MPC for a building model.

### 2017-09-17 paper/intro.tex
dumped CDC paper

\section{INTRODUCTION}
\label{S:intro}

Machine learning and control theory are two foundational but disjoint communities. Machine learning requires data to produce models, and control systems require models to provide stability and performance guarantees to plant operations. Machine learning is widely used for regression or classification, but thus far data-driven models have not been suitable for closed-loop control of physical plants. The challenge now, with using data-driven approaches, is to close the loop for real-time control and decision making.

Consider a multivariable dynamical system subject to external disturbances. The first and foremost requirement for making any decision is to obtain the underlying control-oriented predictive model of the system. With a reasonable forecast of the external disturbances, these models should predict the state of the system in the future and thus a predictive controller based on Model Predictive Control (MPC) can act preemptively to provide a desired behavior. In particular, MPC has been proven to be very powerful for multivariable systems in the presence of input and output constraints, and forecast of the disturbances. The caveat is that MPC requires a reasonably accurate physical representation of the system. This makes MPC unsuitable for control of complex plants such as natural gas processing, oil refineries, boilers, manufacturing plants, and buildings where the user expertise, time, and associated sensor costs required to develop a model are very high \cite{Sturzenegger2016,vzavcekova2014}.

There are two main reasons for model complexity. 
(1) The prime contributor is the change in model properties over time. Even if the model is identified once via an expensive route, as the model changes with time, the system identification must be repeated to update the model. Thus, model adaptability or adaptive control is desirable for such systems. 
(2) A secondary reason is the model heterogeneity which further prohibits the use of model-based control. For example, unlike the automobile or the aircraft industry, each building is designed and used in a different way. Therefore, this modeling process must be repeated for every new building. 
Due to aforementioned reasons, the control strategies in such systems are often limited to fuzzy logic rules that are based on best practices.

%	\noindent (1) \textbf{Model complexity:} 
%	Building high fidelity models for intelligent control can be both cost and time prohibitive. For example, in the case of buildings, this requires installation of thousands of new sensors, expert domain knowledge and several months to years of efforts depending upon the size and complexity of structure of the building. We present this case-study in detail in Sec.~\ref{S:casestudy}.

The question now is, can we employ data-driven techniques to reduce the cost of modeling, and still exploit the benefits that MPC has to offer? We therefore look for automatic and data-driven approaches to control that are also adaptive, scalable and interpretable. We solve this problem with \textit{Data Predictive Control (DPC)} by bridging the gap between Machine Learning and Predictive Control.

In our previous work~\cite{JainACC2017,Jain2016}, we introduced the concept of DPC for receding horizon control.
This work has the following contributions.
(1) We first formally present two underlying algorithms: (i) DPC with regression trees, and (ii) DPC with random forests, which also ensure recursive feasibility in receding horizon control.
(2) Using a bilinear building model whose parameters were identified using experiments on a building in Switzerland, we demonstrate the strength of DPC for receding horizon control via one-to-one comparison against a benchmark MPC controller. We show DPC captures 70\% variance in MPC and offers a comparable performance.
(3) We present a practical application of DPC for Demand Response, where we apply DPC to a 6 story 22 zone building model in EnergyPlus \cite{Crawley2001} for which model-based control is not economical and practical due to extreme complexity. We show scalability and efficiency of DPC in providing financial incentives to the end-customers bypassing the need for high fidelity models. We observe that DPC provides the desired power reduction with an average error of 3\%.

%In our previous work~\cite{JainACC2017,Jain2016}, we introduced the concept of DPC for receding horizon control. We now formally present the algorithms for DPC and apply them to a practical building model identified using real weather data and sensor measurements. We present a comprehensive case study to demonstrate the applicability of DPC to demand response, especially how DPC can provide financial incentives to the end-customers bypassing the need for high fidelity models. 
%In particular, we first describe two underlying algorithms: DPC with regression trees and DPC with random forests in Sec.~\ref{S:dpc}. 
%In Sec.~\ref{S:proof}, we apply DPC to a bilinear building model, whose parameters were identified using experiments on a building in Switzerland, to show a one-to-one comparison of receding horizon control with DPC against a benchmark MPC controller. In Sec.~\ref{S:casestudy}, we present a practical application of DPC for Demand Response, where we apply DPC to a multi-story 19 zone building model in EnergyPlus for which model-based control is not economical and practical due to extreme complexity.

%\noindent(2) One of the limitations of using data-driven models for control is the quality and the quantity of data. We address this problem of experiment design by using Gaussian Processes (GP). We develop probability distributions from the available data to show how existing DPC-RT and DPC-En models can be improved when limited data is available for training.

%We begin with description of DPC-RT and DPC-En algorithms, and Gaussian Processes for DPC in Sec.~\ref{S:dpc}. Sec.~\ref{S:casestudy} presents a case study on DR using a bilinear building model. Here, we also validate the black-box models used for control, and compare DPC against MPC. We discuss the challenges associated with DPC in Sec.~\ref{SS:challenges}. We conclude the paper with a summary of the results and a brief discussion on the future work in Sec.~\ref{S:conclusion}.

### 2017-09-17 paper/casestudyReal.tex
added first results plot

@@
 
+\begin{figure}[h]
+	\subfigure[Comparison of temperature of room $3$ obtained with DPC and ON/OFF controller.]{
+		\label{F:Temperatures_small}
+		\centering
+		\includegraphics[width=22pc]{figures/Temperatures_small.eps}
+	}
+	\subfigure[Input schedules obtained from DPC and ON/OFF controller.]{
+		\label{F:Inputs_small}
+		\centering
+		\includegraphics[width=22pc]{figures/Inputs_small.eps}
+	}
+	\subfigure[Energy consumption over the simulation period obtained with DPC and ON/OFF controller.]{
+		\label{F:Energy_small}
+		\centering
+		\includegraphics[width=22pc]{figures/Energy_small.eps}
+	}
+	\caption{Comparison of DPC and ON/OFF control performance.}
+	\captionsetup{justification=centering}
+	\label{F:Comparison_small}
+\end{figure}
+
+\begin{figure}[h]
+	\subfigure[Comparison of temperature of room $3$ obtained with DPC and ON/OFF controller.]{
+		\label{F:Temperatures_all}
+		\centering
+		\includegraphics[width=22pc]{figures/Temperatures_all.eps}
+	}
+	\subfigure[Input schedules obtained from DPC and ON/OFF controller.]{
+		\label{F:Energy_all}
+		\centering
+		\includegraphics[width=22pc]{figures/Energy_all.eps}
+	}
+	\caption{Comparison of DPC and ON/OFF control performance with different violations.}
+	\captionsetup{justification=centering}
+	\label{F:Comparison_small}
+\end{figure}
 \paragraph{Energy-efficiency improvements}

### 2017-09-18 paper/casestudyReal.tex
added simulations and some text

--- 2017-09-17 paper/casestudyReal.tex
@@
-\section{Case study 3: real house} TBD
+\section{REAL CASE STUDY: OPTIMAL HEATER SCHEDULING} In this section we show a real application of DPC.
+In particular we create both an EnergyPlus and a random forest model for energy consumption, that we use to compare performance, and random forest models for power consumption and room temperatures to be used in the closed-loop simulations.
+These models are created using real data from the period March $11^{th}$ - May $15^{th}$ $2016$ of an off-grid house located in L'Aquila (Italy).
+We apply DPC to obtain an optimal scheduling policy for the heating system and compare the results with a classical ON/OFF controller.
+In the following we first describe the house and the models we created for the simulation, then we set up our optimization problem and show performance results.
 
@@
 
-\subsection{Simulation Results} TBD
+\subsection{Random forest models} We create $2$ different sets of models $S_1$ and $S_2$ using random forests.
+In each set we have a model for the power consumption of the house and $4$ models that describe the evolution of the room temperature in each of the $4$ rooms.
+Models in $S_1$ are used as plant simulator of the house and are created using the classical random forest algorithm, while models in $S_2$ are used for predictions in the DPC algorithm and are then created using the methodology provided in Section \ref{SS:dpcrf}.
+The non-manipulated features in $\tX^d$ are the disturbance data (relative humidity, atmospheric pressure, outside air temperature, sun radiation, wind, time of the day and day of the week) and the states (temperature of $4$ rooms).
+The manipulated feature in $\tX^c$ is the flow rate ($[m^3/h]$).
+All this features are used to create models in $S_1$ and the power model in $S_2$, while disturbance data, only state temperature $x_i$ and flow rate are used to identify $\hat{\Theta}^{T}_{\tT^i_j}$ for temperature models in $S_2$.
+
+\begin{table}[h!]
+	\centering
+	\caption{Model accuracy of models in $S_1$.}
+	\captionsetup{justification=centering}
+	\begin{tabular}{c|c|c|c|c|c}
+		\toprule
+		$S_1$     & Power & T. room $1$ & T. room $2$ & T. room $3$ & T. room $4$  \\ 
+		\midrule
+			$S_1$ & $num$ & $0.627$        & $158$          & $158$           & $158$\\
+			$S_2$ & $num$ & $0.673$        & $81$           & $158$           & $158$\\
+		\bottomrule
+	\end{tabular}
+	\label{T:S1accuracy}
+\end{table}
+
+\begin{table}[h!]
+	\centering
+	\caption{Model accuracy of models in $S_2$.}
+	\captionsetup{justification=centering}
+	\begin{tabular}{c|c|c|c|c|c}
+		\toprule
+		$S_1$     & Power & T. room $1$ & T. room $2$ & T. room $3$ & T. room $4$  \\ 
+		\midrule
+		$S_1$ & $num$ & $0.627$        & $158$          & $158$           & $158$\\
+		$S_2$ & $num$ & $0.673$        & $81$           & $158$           & $158$\\
+		\bottomrule
+	\end{tabular}
+	\label{T:S2accuracy}
+\end{table}
+
+\subsection{DPC and bang-bang control problems} TBD
+
+\paragraph{DPC}
+We want to optimize the ON/OFF heater schedule in order to minimize power consumption of the house while keeping temperature of room $3$ within a comfort range.
+We choose to constraint only one room temperature since in the house it is not possible to control temperature in each room.
+This is because as explained in Section TBD, the same flow controls the temperature of all rooms simultaneously.
+We also allow violations $\eps^{min},\eps^{max}$  of temperature bounds to guarantee feasibility of the algorithm.
+We include these violations in the objective function to minimize and we compare performance considering different weights.
+The problem is setup as follows:
+\begin{align}
+	\begin{aligned}
+		\min_{\tX^c,\eps_j^{min},\eps_j^{max}} & \ \ \ \sum_{j=1}^{N} Q{\tP}_{\mathrm{k+j|k}}^2 +  \lambda_{min}\parallel\eps^{min}\parallel_2 + \lambda_{max}\parallel\eps^{max}\parallel_2 \\
+		\text{s.~t. }& \ \ \ \ \ {\tP}_{\mathrm{k+j|k}} =  \hat{\Theta}^{T}_{\tP_j} [ 1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
+					 & \ \ \ \ \ \tT^1_{\mathrm{k+j|k}} =  \hat{\Theta}^{T}_{\tT^1_j} [ 1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
+					 & \ \ \ \ \ \tT^2_{\mathrm{k+j|k}} =  \hat{\Theta}^{T}_{\tT^2_j} [ 1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
+			 	 	 & \ \ \ \ \ \tT^3_{\mathrm{k+j|k}} =  \hat{\Theta}^{T}_{\tT^3_j} [ 1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
+				 	 & \ \ \ \ \ \tT^4_{\mathrm{k+j|k}} =  \hat{\Theta}^{T}_{\tT^4_j} [ 1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\					
+					 & \underline{\tT}_{\mathrm{k+j-1|k}}-\eps^{min}_j \leq \tT^3_{\mathrm{k+j|k}} \leq \overline{\tT}_{\mathrm{k+j-1|k}} + \eps^{max}_j\\
+					 & \ \ \ \ \ \ \ \ \tX^c_{\mathrm{k+j-1|k}} = 0 \lor \tX^c_{\mathrm{k+j-1|k}} = 0.35\\ 
+					 & \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \eps_j \geq 0, \ j = 1,\dots,N.
+	\end{aligned}
+	\label{E:DPCrealcase}
+\end{align}
+\paragraph{Bang-bang controller}
+This is the classical controller used to keep temperature within a range.
+It switch the heater on when the temperature goes under the temperature lower bound and switch it off when the temperature goes over the temperature upper bound.
+The advantage is that it is very simple to install but it uses more energy than what it is actually needed to achieve the task.
+\subsection{Simulation Results} To avoid to many oscillations we set up both control problems such that when the heater is activated then it must stay active for at least $20$ minutes.
+
+\begin{figure}[h!]
+	\subfigure[Comparison of temperature of room $3$ obtained with DPC and bang-bang controller.]{
+		\label{F:Temperatures_small}
+		\centering
+		\includegraphics[width=22pc]{figures/Temperatures_small.eps}
+	}
+	\subfigure[Input schedules obtained from DPC and bang-bang controller.]{
+		\label{F:Inputs_small}
+		\centering
+		\includegraphics[width=22pc]{figures/Inputs_small.eps}
+	}
+	\caption{Comparison of DPC and bang-bang control performance.}
+	\captionsetup{justification=centering}
+	\label{F:Comparison_small}
+\end{figure}
+
+\begin{figure}[h!]
+	\subfigure[Comparison of temperature of room $3$ obtained with DPC and bang-bang controller.]{
+		\label{F:Temperatures_all}
+		\centering
+		\includegraphics[width=22pc]{figures/Temperatures_all.eps}
+	}
+	\subfigure[Input schedules obtained from DPC and bang-bang controller.]{
+		\label{F:Energy_all}
+		\centering
+		\includegraphics[width=22pc]{figures/Energy_all.eps}
+	}
+	\caption{Comparison of DPC and bang-bang control performance with different violations.}
+	\captionsetup{justification=centering}
+	\label{F:Comparison_small}
+\end{figure}
 

### 2017-09-18 paper/main.tex
dumped CDC paper
added first results plot
added simulations and some text

--- 2017-09-16 paper/DPC_T&A.tex
@@
 
+\usepackage{achinDefs}
 \usepackage{lineno,hyperref}
@@
 \journal{Journal of \LaTeX\ Templates}
-
-%%%%%%%%%%%%%%%%%%%%%%%
-%% `Elsevier LaTeX' style
-\bibliographystyle{elsarticle-num}
-%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \ead{achinj@seas.unpenn.edu}
-\author[TDRaddress,EC]{Tullio De Rubeis}
+\author[TDRaddress,EC]{Tullio de Rubeis}
 \ead{tullio.derubeis@graduate.univaq.it}
@@
 
+\address[FSaddress]{Department of Information Engineering, Computer Science and Mathematics, Universit\`{a} degli Studi dell'Aquila, L'Aquila, Italy}
+\address[AJaddress]{Department of Electrical and Systems Engineering, University of Pennsylvania, Philadelphia, USA}
+\address[TDRaddress]{Department of Industrial and Information Engineering and Economics, Universit\`{a} degli Studi dell'Aquila, L'Aquila, Italy}
 \address[EC]{Equal contribution}
-\address[FSaddress]{Department of Information Engineering, Computer Science and Mathematics, Università degli Studi dell’Aquila, L’Aquila, Italy}
-\address[TDRaddress]{Department of Industrial and Information Engineering and Economics, Università degli Studi dell’Aquila, L’Aquila, Italy}
-\address[AJaddress]{Department of Electrical and Systems Engineering, University of Pennsylvania, Philadelphia, USA}
 
+%% ABSTRACT
 \begin{abstract}
-TBD
+\input{abstract.tex}
 \end{abstract}
@@
 \begin{keyword}
-Energy efficiency, predictive control, machine learning.
+Energy efficiency, demand response, predictive control, machine learning
 \end{keyword}
@@
 
-%\linenumbers
+%% SECTIONS
+\input{intro.tex}
 
-\section{Introduction}
+\input{dpc.tex}
 
-\paragraph{Motivations} TBD
+\input{casestudyBilinear.tex}
 
-\paragraph{State of the art} TBD
+\input{casestudyEPlus.tex}
+\clearpage
+\input{casestudyReal.tex}
 
-\paragraph{Contributions} TBD
+%% CONCLUSION
+\input{conclusion.tex}
 
-\paragraph{Paper organization} TBD
+%ACKNOWLEDGMENTS
+\input{ack.tex}
 
-\section{Problem definition and objectives} TBD
-
-\section{Data Predictive Control using Random Forests} TBD
-
-\subsection{Modeling} TBD
-
-\subsection{Predictive Control} TBD
-
-\section{Case study 1: bilinear model} TBD
-
-\section{Case study 2: EnergyPlus model} TBD
-
-\section{Case study 3: real house} TBD
-
-\subsection{Description of the house} TBD
-
-\subsection{EnergyPlus model} TBD
-
-\subsection{Simulation Results} TBD
-
-\paragraph{Energy-efficiency improvements}
-
-\section{Conclusion and future work} TBD
-
+%% BIBLIOGRAPHY
 \section*{References}
-
-\bibliography{mybibfile}
+\bibliographystyle{elsarticle-num}
+\bibliography{fsRefs,achinRefs}
 

### 2017-09-18 paper/casestudyReal.tex
added temporary flow scheme in images folder

@@
 
-\subsection{Random forest models} We create $2$ different sets of models $S_1$ and $S_2$ using random forests.
-In each set we have a model for the power consumption of the house and $4$ models that describe the evolution of the room temperature in each of the $4$ rooms.
-Models in $S_1$ are used as plant simulator of the house and are created using the classical random forest algorithm, while models in $S_2$ are used for predictions in the DPC algorithm and are then created using the methodology provided in Section \ref{SS:dpcrf}.
+	\subsection{Random forest models} We create $2$ different sets of models $S_1 = \{\tP^{RF}_{\mathrm{k+1|k}},\tT^{RF,1}_{\mathrm{k+1|k}},\tT^{RF,2}_{\mathrm{k+1|k}},\tT^{RF,3}_{\mathrm{k+1|k}},\tT^{RF,4}_{\mathrm{k+j|k}}\}$ and $S_2 = \{\tP_{\mathrm{k+j|k}},\tT^{1}_{\mathrm{k+j|k}},\tT^{2}_{\mathrm{k+j|k}},\tT^{3}_{\mathrm{k+j|k}},\tT^{4}_{\mathrm{k+j|k}},\ j = 1,\ldots,N\}$ using random forests.
+In each set we have a model for the power consumption of the house ($\tP^{RF}$ and $\tP$) and $4$ models that describe the evolution of the room temperature in each of the $4$ rooms ($\tT^{RF,i}$ and $\tT^i$).
+Models in $S_1$ are used as plant simulator of the house and are created using the classical random forest algorithm, since it can be seen in Table \ref{T:S1accuracy} they have better accuracy than those in $S_2$.
+Models in $S_2$ are used for predictions in the DPC algorithm and are then created using the methodology provided in Section \ref{SS:dpcrf}.
 The non-manipulated features in $\tX^d$ are the disturbance data (relative humidity, atmospheric pressure, outside air temperature, sun radiation, wind, time of the day and day of the week) and the states (temperature of $4$ rooms).
@@
 All this features are used to create models in $S_1$ and the power model in $S_2$, while disturbance data, only state temperature $x_i$ and flow rate are used to identify $\hat{\Theta}^{T}_{\tT^i_j}$ for temperature models in $S_2$.
+The accuracy of these models is shown in Table \ref{T:S1accuracy}.
 
@@
 		\midrule
-			$S_1$ & $num$ & $0.627$        & $158$          & $158$           & $158$\\
-			$S_2$ & $num$ & $0.673$        & $81$           & $158$           & $158$\\
+			$\tP^{RF}_{\mathrm{k+1|k}}$ & $num$ & $0.627$        & $158$          & $158$           & $158$\\
+			$\tP_{\mathrm{k+1|k}}$      & $num$ & $0.673$        & $81$           & $158$           & $158$\\
+			$\tT^{RF,1}_{\mathrm{k+1|k}}$ & $num$ & $0.627$        & $158$          & $158$           & $158$\\
+			$\tT^1_{\mathrm{k+1|k}}$      & $num$ & $0.673$        & $81$           & $158$           & $158$\\
+			$\tT^{RF,2}_{\mathrm{k+1|k}}$ & $num$ & $0.627$        & $158$          & $158$           & $158$\\
+			$\tT^2_{\mathrm{k+1|k}}$      & $num$ & $0.673$        & $81$           & $158$           & $158$\\
+			$\tT^{RF,3}_{\mathrm{k+1|k}}$ & $num$ & $0.627$        & $158$          & $158$           & $158$\\
+			$\tT^3_{\mathrm{k+1|k}}$      & $num$ & $0.673$        & $81$           & $158$           & $158$\\
+			$\tT^{RF,4}_{\mathrm{k+1|k}}$ & $num$ & $0.627$        & $158$          & $158$           & $158$\\
+			$\tT^4_{\mathrm{k+1|k}}$      & $num$ & $0.673$        & $81$           & $158$           & $158$\\
 		\bottomrule

### 2017-09-22 paper/casestudyEPlus.tex
added simulations and some text
completed section 5 and fixed figures problems

@@
 
-For the following simulations, we use five control variables (i) cooling set point for corridors $\mathsf{ClgSP}$, (ii) cooling set point for guest rooms $\mathsf{GuestSP}$, (iii) cooling set point for kitchen $\mathsf{KitchenSP}$, (iv) chilled water supply temperature $\mathsf{ChwSP}$, and (v) supply air temperature  $\mathsf{SupplyAirSP}$, so $\tX^c = [\mathsf{ClgSP},\mathsf{GuestClgSP},\mathsf{KitchenClgSP},\mathsf{SupplyAirSP}, \mathsf{ChwSP}]$.
+For the following simulations, we use five control variables (i) cooling set point for corridors $\mathsf{ClgSP}$, (ii) cooling set point for guest rooms $\mathsf{GuestSP}$, (iii) cooling set point for kitchen $\mathsf{KitchenSP}$, (iv) chilled water supply temperature $\mathsf{ChwSP}$, and (v) supply air temperature  $\mathsf{SupplyAirSP}$, so $\tX^c = [\mathsf{ClgSP},\mathsf{GuestClgSP},\\\mathsf{KitchenClgSP},\mathsf{SupplyAirSP}, \mathsf{ChwSP}]$.
 The power forest $\mathcal{R}_p$ is built using the total building power consumption $\tP$.
@@
 
-\begin{figure}[h]
+\begin{figure}[h!]
+	\begin{center}
 	\subfigure[Optimal inputs calculated by DPC-En.
@@
 	}
+\end{center}
 	\caption{Power management using DPC. The controller is active between 7am - 2pm. This region is marked in dashed red lines.}
@@
 Thus, DPC-En successfully tracks a given power reference signal with an average $\sim$ 3\% error for such a complex building which would require several years of efforts to develop a physics based model.
-
-\subsection{Practical Challenges and Future Work}
-\label{SS:challenges}
-\input{cs-challenges.tex}

### 2017-09-22 paper/casestudyReal.tex
completed section 5 and fixed figures problems

--- 2017-09-17 paper/casestudyReal.tex
@@
-\section{Case study 3: real house} TBD
+\section{REAL CASE STUDY: OPTIMAL HEATING SYSTEM SCHEDULING} In this section we show an application of DPC to a real case study.
+In particular, in the following we first describe the plant structure, then we provide both data-based and EnergyPlus model of such a plant to be used for control and performance comparison, and finally we show results of the closed-loop simulations.
 
-\subsection{Description of the house} TBD
+\subsection{Description of the house}\label{SS:descriptionHouse}
+The chosen case study is a detached off-grid two-story residential house, located in the outskirt of L'Aquila (coordinates $42\degree$ $16'$ latitude and $13\degree$ $32'$ longitude), Italy and is shown in Figure \ref{F:house}.
+The building, inhabited by the two owners, has a main north-south orientation and it is composed by a heated ground floor, that includes $4$ rooms, and an attic without heating system.
+Therefore, although the gross area of the house is equal to $209.5\,m^2$, the heated gross area is equal to $112.4\,m^2$.
 
-\subsection{EnergyPlus model} TBD
-
-\subsection{Simulation Results} TBD
-
-\begin{figure}[h]
-	\subfigure[Comparison of temperature of room $3$ obtained with DPC and ON/OFF controller.]{
-		\label{F:Temperatures_small}
-		\centering
-		\includegraphics[width=22pc]{figures/Temperatures_small.eps}
-	}
-	\subfigure[Input schedules obtained from DPC and ON/OFF controller.]{
-		\label{F:Inputs_small}
-		\centering
-		\includegraphics[width=22pc]{figures/Inputs_small.eps}
-	}
-	\subfigure[Energy consumption over the simulation period obtained with DPC and ON/OFF controller.]{
-		\label{F:Energy_small}
-		\centering
-		\includegraphics[width=22pc]{figures/Energy_small.eps}
-	}
-	\caption{Comparison of DPC and ON/OFF control performance.}
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[width=24pc]{figures/Vista_sud.eps}
+		\caption{Off-grid residential house.}
 		\captionsetup{justification=centering}
-	\label{F:Comparison_small}
+		\label{F:house}
+	\end{center}
 \end{figure}
 
-\begin{figure}[h]
-	\subfigure[Comparison of temperature of room $3$ obtained with DPC and ON/OFF controller.]{
-		\label{F:Temperatures_all}
+Thanks to the off-grid characteristic, the technological plants guarantee the complete energy self-sufficiency of the building.
+Indeed the house, equipped with a biomass boiler, a solar thermal plant, a stand-alone photovoltaic system, black water and rainwater reuse systems and as well for water supply, has a complete independence from the utilities.
+The bearing structure of the house is made of reinforced concrete and EPS (expanded polystyrene) insulation, while the building envelope is composed by prefabricated wood-cement blocks, shown in Figure \ref{F:houseSection}, with EPS and graphite insulation, that allow low thermal fluxes thanks to their high thermal performance.
+The thermal performance of a wood-cement block, with similar geometry, was investigated in a previous work \cite{Nardi2016}, both with experimental and numerical approach.
+
+\begin{figure}[h!]
+	\begin{center}
+		\subfigure[Walls.]{
+			\label{F:houseSection1}
 			\centering
-		\includegraphics[width=22pc]{figures/Temperatures_all.eps}
+			\includegraphics[width=24pc]{figures/blocco_parete_rev01.eps}
 		}
-	\subfigure[Input schedules obtained from DPC and ON/OFF controller.]{
-		\label{F:Energy_all}
+		\subfigure[Floor and roof.]{
+			\label{F:houseSection2}
 			\centering
-		\includegraphics[width=22pc]{figures/Energy_all.eps}
+			\includegraphics[width=24pc]{figures/blocco_solaio_rev01.eps}
 		}
-	\caption{Comparison of DPC and ON/OFF control performance with different violations.}
+	\end{center}
+	\caption{Cross-sections of the wood-cement blocks.}
 	\captionsetup{justification=centering}
-	\label{F:Comparison_small}
+	\label{F:houseSection}
 \end{figure}
-\paragraph{Energy-efficiency improvements}
+The heating system of the building, that supplies the thermal energy required in winter season, is a hydronic system consisting of a vegetable biomass boiler, with a manual ON/OFF, a constant-flow pump station, and tubular steel radiators as shown in Figure \ref{F:housePlantScheme}.
+The standard biomass boiler has an efficiency of $83.5\%$ with $16.5\ kW$ of thermal power transferred to the water, without gas-flame modulation.
+The heat transfer fluid distribution is realized through a manifold circuit, that supplies the radiators placed in the various rooms of the house.
+The energy needs for the domestic hot water (DHW) are covered by the same boiler, coupled with a solar thermal plant.
+It is worth noting that, in this work, the thermal energy needed for the domestic hot water production is neglected.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[width=26pc]{figures/Schema_di_impianto.eps}
+		\caption{Technological plant scheme of the use case.}
+		\captionsetup{justification=centering}
+		\label{F:housePlantScheme}
+	\end{center}
+\end{figure}
+For a thorough knowledge of the building thermal behavior, an in-situ analysis was carried out.
+Temperature measuring devices, completely self-produced at the “G.
+Parolini Lab” of the University of L'Aquila \cite{Pantoli2017}, mainly based on an $ATmega 2560$ microcontroller and $DS18B20$ temperature probes (temperature range from $-55.0\degree C$ to $125.0\degree C$ with an accuracy of $\pm 0.5\degree C$), were employed to acquire the ambient temperatures of the four differently oriented rooms of the house as shown in Figure \ref{F:houseFloors}.
+The probes positions were chosen to optimize the data acquisition and to minimize the discomfort of the occupants in the house.
+Furthermore, as can be seen in Figures \ref{F:housePlantScheme} and \ref{F:houseFloors} (COMMENT: NOT CLEAR WHERE THE SENSORS ARE IN THE FIGURES.
+MORE DETAILS NEEDED), a commercial heat meter (temperature range from $10.0\degree C$ to $90.0\degree C$ with an accuracy of $\pm 0.05\degree C$) was installed downstream of the biomass boiler to measure the produced thermal energy.
+All the measuring devices have been set with a data acquisition rate equal to $10$ minutes, from March $11^{th}$ $2016$ to May $15^{th}$ $2016$.
+\begin{figure}[h!]
+	\begin{center}
+		\subfigure[Ground floor.]{
+			\label{F:houseGroundFloor}
+			\centering
+			\includegraphics[width=26pc]{figures/Arch_PT_e_sonde.eps}
+		}
+		\subfigure[Attic.]{
+			\label{F:houseAttic}
+			\centering
+			\hspace{-1.6cm}
+			\includegraphics[width=22.8pc]{figures/Arch_P1.eps}
+		}
+	\end{center}
+	\caption{Layout of the house and probes placement. Legend: red circle for ambient temperature; orange circle for heat meter.}
+	\captionsetup{justification=centering}
+	\label{F:houseFloors}
+\end{figure}
+
+The collected data are used in the following to create both an EnergyPlus and a random forest model for the energy consumption of the house.
+These models will be used for performance comparison of DPC with respect to a classical bang-bang controller.
+In particular DPC will be set up to provide an optimal ON/OFF scheduling policy for the heating system in order to save energy while guaranteeing thermal comfort for the occupants.
+To this aim, we also create random forest models for power consumption and room temperatures to be used in the closed-loop simulations.
+
+\subsection{EnergyPlus model}\label{SS:energyPlusmodel}
+To assess the energy performance of the use case's heating system, the EnergyPlus engine coupled with DesignBuilder modeling environment was employed.
+Based on the weather data provided by CETEMPS – Centre of Excellence \cite{CETEMPS}, shown in Figure \ref{F:houseExternalWeather}, the performed dynamic simulation was based on a weather file specifically created for L'Aquila, that according to the K\"{o}ppen-Geiger climate classification, has a warm temperate climate (Csa, Cfa and Csb) \cite{Peel2007}.
+(COMMENT: TO BE BETTER WRITTEN)
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[width=26pc]{figures/dati_climatici_rev01.eps}
+		\caption{Weather data of L'Aquila.}
+		\captionsetup{justification=centering}
+		\label{F:houseExternalWeather}
+	\end{center}
+\end{figure}
+
+Because of the complex morphology of the blocks shown in Figure \ref{F:houseSection}, some simplified assumptions were made, in order to create the EnergyPlus virtual model.
+Considering the thermal properties of the wood-cement blocks that compose the walls shown in Figure \ref{F:houseSection1}, the thermal effects of the ribs were neglected.
+For the block employed for floor and roof shown in Figure \ref{F:houseSection2}, a less complex equivalent block, with only three layers, was evaluated in order to consider only one equivalent thermal transmittance value.
+The properties of the blocks used for the simulation model are listed in Table \ref{T:houseProperties}.
+
+\begin{table}[h!]
+	\centering
+	\begin{tabular}{ccccc}
+		\toprule
+		\multirow{3}{1.4cm}{\centering Structural member} & \multirow{3}{3.7cm}{\centering Layer description (from inside to outside)}     & Thermal     		    & Total     & Total        \\
+		                                                  &                                                                                & resistance   		    & thickness & U-value      \\
+		                                                  &                                                                                & $[(m^2K)/W]$		    & $[m]$     & $[W/(m^2K)]$ \\
+		\midrule
+		\multirow{3}*{Wall}                               & Wood-cement                                                                    & $0.308$     		    &           &              \\ 
+		                            					  & Concrete                                                                       & $0.096$    		    & $0.50 $   & $0.12 $      \\ 
+		                            					  & EPS and graphite                                                               & $6.774$    		    &           &              \\
+														  &																				   &						&			&			   \\
+		\multirow{3}*{Floor}        					  & Screed                                                                         & $0.027$     		    &           &              \\ 
+							        					  & Polystyrene                                                                    & $6.000$     		    & $0.23 $   & $0.28 $      \\ 
+							        					  & Wood-cement                                                                    & $0.308$     		    &           &              \\
+														  &																				   &						&			&			   \\
+		\multirow{5}{1.4cm}{\centering Pitched roof} 	  & \multirow{2}{3.7cm}{\centering Polyurethane resins and polyisocianurate foams} & \multirow{2}*{$4.000$} &           &              \\ 
+														  &                                                                                &                        &           &              \\
+														  & Screed                                                                         & $0.027$    		    & $0.31 $   &  $0.13 $     \\
+							        					  & Polystyrene                                                                    & $6.000$    		    &           &              \\ 
+							        					  & Wood-cement                                                                    & $0.308 $   		    &           &              \\
+		\bottomrule
+	\end{tabular}
+	\caption{Wood-cement blocks properties.}
+	\captionsetup{justification=centering}
+	\label{T:houseProperties}
+\end{table}
+
+Therefore, the dynamic simulation model of the use case, shown in Figure \ref{F:houseVirtualModel}, has been created by analyzing the fundamental characteristics of the building (orientation, geometry, structural members, heating system components, air changes with natural ventilation, activity, internal gains, air leakages) and the weather file specifically created for L'Aquila.
+The comprehensive method was chosen for modeling the heating system, once checked all the characteristics of the components.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[width=24pc]{figures/Render_rev01.eps}
+		\caption{Virtual model of the use case.}
+		\captionsetup{justification=centering}
+		\label{F:houseVirtualModel}
+	\end{center}
+\end{figure}
+
+In order to calibrate the EnergyPlus model, a comparison between simulated and measured thermal energy consumption was performed.
+Considering the off-grid characteristic of the house, the installation of a heat meter was necessary for acquiring the actual energy consumption.
+Moreover, the heat meter installation has allowed a detailed knowledge of the actual heating system scheduling, shown in Figure \ref{F:houseActualScheduling}, faithfully reproduced in the simulated model.
+The heat meter, that consists of two thermocouples for flow and return thermal fluid temperatures, a turbine flowmeter and a controller, employed Equation \eqref{Eq:energyConsumption} to calculate the real thermal energy consumption $\dot Q$ of the house.
+\begin{equation}\label{Eq:energyConsumption}
+\dot Q = 0.2777698\times10^{-3}\times\rho\times\Delta V\times c_p\times\Delta T
+\end{equation}
+In \eqref{Eq:energyConsumption} $\dot Q$ is the thermal energy consumption $[Wh]$, $\rho$ is the water density $[kg/m^3]$, $\Delta V$ is the water volume variation $[m^3]$ detected by the turbine flowmeter, $c_p = 4.186\, [kJ/(kgK)]$ is the water specific heat at constant pressure, $\Delta T$ is the difference between flow and return temperatures of the water $[K]$, $0.2777698∙10−3$ is a dimensionless conversion factor.
+The considered period goes from March $15^{th}$ $2016$ to April $15^{th}$ $2016$.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[width=28pc]{figures/Scheduling_rev02.eps}
+		\caption{Actual scheduling of the heating system, where white color means switched OFF and green color switched ON, and ($T_{ext,ave}$) is the external average temperature.
+From midnight to $8am$ the heating system is always switched off.)}
+		\captionsetup{justification=centering}
+		\label{F:houseActualScheduling}
+	\end{center}
+\end{figure}
+Following the hourly calibration proposed by the M\&V guidelines of ASHRAE \cite{USDOE}, a simulated model is calibrated when the mean bias error (MBE) and the coefficient of variation of the root mean square error [CV(RMSE)] are less than acceptable tolerances, respectively equal to $\pm 10.0\%$ and $30.0\%$.
+MBE and CV(RMSE) have been calculated by using Equations \eqref{Eq:mbe} and \eqref{Eq:cvrmse}.
+
+\begin{equation}\label{Eq:mbe}
+MBE(\%) = \frac{\sum_{Period}{(S-M)_{Interval}}}{\sum_{Period}{M_{Interval}}} \times 100
+\end{equation}
+
+where $M$ is the measured $kWh$ and $S$ is the simulated $kWh$.
+
+\begin{align}\label{Eq:cvrmse}
+\nonumber CV(RMSE_{Period}) &= \frac{RMSE_{Period}}{A_{Period}} \times 100 \\
+				            &= \sqrt{\frac{\sum{(S-M)^2_{Interval}}}{N_{Interval}}} \times \frac 1{A_{Period}} \times 100
+\end{align}
+
+where $A_{Period}$ is the mean of the measured data for the period, Equation \eqref{Eq:aperiod}, and $N_{Interval} = 4563$ is the number of time intervals in the monitoring period.
+
+\begin{equation}\label{Eq:aperiod}
+A_{Period} = \frac{\sum_{Period}{M_{Interval}}}{N_{Interval}}
+\end{equation}
+
+The comparison between numerical and experimental data is shown in Figure \ref{F:houseComparisonExperimental} and shows a quite good agreement.
+Therefore, with a MBE equal to $7.38\%$ and a CV(RMSE) equal to $8.37\%$, the EnergyPlus model of the use case can be considered well calibrated.
+
+\begin{figure}[h!]
+	\begin{center}
+		\subfigure[Thermal energy.]{
+			\label{F:houseComparisonExperimentalEnergy}
+			\centering
+			\includegraphics[width=24pc]{figures/Thermal_energy_rev01.eps}
+		}
+		\subfigure[Thermal power.]{
+			\label{F:houseComparisonExperimentalPower}
+			\centering
+			\includegraphics[width=24pc]{figures/Thermal_power.eps}
+		}
+	\end{center}
+	\caption{Comparison between numerical and experimental data.}
+	\captionsetup{justification=centering}
+	\label{F:houseComparisonExperimental}
+\end{figure}
+\subsection{Random forest models} For the closed-loop simulations with DPC we create $2$ different sets of models, $S_1 = \{\tP^{r}_{\mathrm{k+1|k}},\tT^{r_1}_{\mathrm{k+1|k}},\tT^{r_2}_{\mathrm{k+1|k}},\tT^{r_3}_{\mathrm{k+1|k}},\tT^{r_4}_{\mathrm{k+j|k}}\}$ and $S_2 = \{\tP_{\mathrm{k+j|k}},\tT^{1}_{\mathrm{k+j|k}},\tT^{2}_{\mathrm{k+j|k}},\tT^{3}_{\mathrm{k+j|k}},\\\tT^{4}_{\mathrm{k+j|k}},\ j = 1,\ldots,N\}$, using random forests.
+In each set we have a model for the power consumption of the house ($\tP^{r}$ and $\tP$) and $4$ models that describe the evolution of the room temperature in each of the $4$ rooms ($\tT^{r_i}$ and $\tT^i$).
+Models in $S_1$ are created using the classical random forests algorithm with all the features and are used as plant simulator of the house.
+For this reason, they are computed to give prediction only for time step $k+1$ and not over the whole horizon $N$.
+Models in $S_2$ are used as predictors over a horizon $N$ in the DPC algorithm and are then created using the methodology provided in Section \ref{SS:dpcrf}.
+The non-manipulated features in $\tX^d$ are the disturbance data (relative humidity, atmospheric pressure, outside air temperature, solar radiation, wind, time of the day and day of the week) and the states (temperature of the $4$ rooms).
+The manipulated feature in $\tX^c$ is the flow rate ($[m^3/h]$).
+All this features are used to create models in $S_1$ and the power model in $S_2$, while disturbance data, state temperature of room $i$ only and flow rate are used to identify $\hat{\Theta}^{T}_{\tT^i_j}$ for temperature models in $S_2$.
+All the models have been trained on the data from March $11^{th}$ $2016$ to April $26^{th}$ $2016$ and validated on the data from May $1^{st}$ $2016$ to May $15^{th}$ $2016$.
+The accuracy of these models with respect to real data is shown in Table \ref{T:S1accuracy}, based on the definition of Normalized Root Mean Square Error (NRMSE).
+
+\begin{table}[h!]
+	\centering
+	\begin{tabular}{cccccc}
+		\toprule
+		Set       & Power   & T. room $1$ & T. room $2$ & T. room $3$ & T. room $4$  \\ 
+		\midrule
+		$S_1$     & $92.44$ & $96.58$     & $96.99$     & $97.21$     & $96.62$\\
+		$S_2$     & $92.21$ & $97.38$     & $97.29$     & $96.93$     & $96.81$\\
+		\bottomrule
+	\end{tabular}
+	\caption{Models accuracy for power and temperatures models in $S_1$ and $S_2$ expressed as $\mathrm{1-NRMSE}\,(\%)$.}
+	\captionsetup{justification=centering}
+	\label{T:S1accuracy}
+\end{table}
+
+A graphical comparison is shown in Figure \ref{F:power_testing} for the power consumption and Figure \ref{F:temperature_testing} for the temperature of room $1$.
+The plots for the other rooms are omitted since they are very similar.
+
+\begin{figure}[h!]
+\begin{center}
+	\includegraphics[width=26pc]{figures/power_testing.eps}
+	\caption{Power consumption model accuracy validation.}
+	\captionsetup{justification=centering}
+	\label{F:power_testing}
+\end{center}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
+	\includegraphics[width=26pc]{figures/temperature_testing.eps}
+	\caption{Temperature model accuracy validation.}
+	\captionsetup{justification=centering}
+	\label{F:temperature_testing}
+\end{center}
+\end{figure}
+
+We use random forest model $\tP^r$ also to estimate the energy consumption of the house.
+In particular we estimated it for the same period analyzed in Section \ref{SS:energyPlusmodel} and obtained the results in Figure \ref{F:Energy_testing}, with a MBE equal to $13.0\%$ and a CV(RMSE) equal to $13.7\%$.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[width=26pc]{figures/Energy_test.eps}
+		\caption{Energy consumption accuracy.}
+		\captionsetup{justification=centering}
+		\label{F:Energy_testing}
+	\end{center}
+\end{figure}
+
+As we can see, with random forest model we get a worse MBE and CV(RMSE) for energy consumption, although the maximum error is very similar.
+This is because the EnergyPlus model, due to its internal dynamics based on the house structure, is able to compensate the error over time.
+On the other hand our curve perfectly follows the shape of the real case.
+For this reason, in Section \ref{SS:simulationResults}, we will use both models in order to compare the quality of the DPC with respect to a classical bang-bang controller.
+
+\subsection{DPC and bang-bang controllers} We set up $2$ different controllers, DPC and bang-bang controller, to obtain a scheduling policy to switch the radiators ON and OFF in order to keep the temperature of room $3$ within a comfort range.
+Since the heating system serves all of the $4$ rooms simultaneously, without giving the possibility to control the temperature of each room independently, we setup the problem defining the comfort range only for one room.
+Other rooms temperatures will follow the scheduling policy.
+Room $3$ has been chosen randomly.
+In the following we describe the $2$ controllers.
+
+\paragraph{DPC}
+We want to optimize the ON/OFF heating system schedule in order to minimize power consumption of the house while keeping temperature of room $3$ within a comfort range.
+We also allow violations $\eps^{min},\eps^{max}$  of temperature bounds to guarantee feasibility of the algorithm.
+We include these violations in the objective function to be minimized.
+The problem is set up as follows:
+\begin{align}
+	\begin{aligned}
+		\min_{\tX^c,\eps_j^{min},\eps_j^{max}} & \ \ \ \sum_{j=1}^{N} Q{\tP}_{\mathrm{k+j|k}}^2 +  \lambda_{min}\parallel\eps^{min}\parallel_2 + \lambda_{max}\parallel\eps^{max}\parallel_2 \\
+		\text{s.~t. }& \ \ \ \ \ {\tP}_{\mathrm{k+j|k}} =  \hat{\Theta}^{T}_{\tP_j} [ 1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
+					 & \ \ \ \ \ \tT^1_{\mathrm{k+j|k}} =  \hat{\Theta}^{T}_{\tT^1_j} [ 1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
+					 & \ \ \ \ \ \tT^2_{\mathrm{k+j|k}} =  \hat{\Theta}^{T}_{\tT^2_j} [ 1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
+			 	 	 & \ \ \ \ \ \tT^3_{\mathrm{k+j|k}} =  \hat{\Theta}^{T}_{\tT^3_j} [ 1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
+				 	 & \ \ \ \ \ \tT^4_{\mathrm{k+j|k}} =  \hat{\Theta}^{T}_{\tT^4_j} [ 1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\					
+					 & \underline{\tT}_{\mathrm{k+j-1|k}}-\eps^{min}_j \leq \tT^3_{\mathrm{k+j|k}} \leq \overline{\tT}_{\mathrm{k+j-1|k}} + \eps^{max}_j\\
+					 & \ \ \ \ \ \ \ \ \tX^c_{\mathrm{k+j-1|k}} = \underline{\tX}^c \lor \tX^c_{\mathrm{k+j-1|k}} = \overline{\tX}^c\\ 
+					 & \ \ \ \ \ \ \ \ \,\eps^{min}_j,\eps^{max}_j \geq 0, \ j = 1,\dots,N.
+	\end{aligned}
+	\label{E:DPCrealcase}
+\end{align}
+The choice of different weights $Q$, $\lambda_{min}$ and $\lambda_{max}$ allows the designer to give more importance to energy consumption rather than temperature comfort and viceversa.
+In Section \ref{SS:simulationResults} we will show different performance results considering different weights.
+Parameters $\underline{\tX}^c$ and $\overline{\tX}^c$ are respectively minimum and maximum values the heating system can actuate, while $\underline{\tT}_{\mathrm{k+j-1|k}}$ and $\overline{\tT}_{\mathrm{k+j-1|k}}$ are respectively time varying lower and upper bounds to keep the temperature in a desired range of comfort.
+Due to the integer variable constraint for input $\tX^c$, the problem is a Mixed Integer Quadratic Programming.
+For the implementation we use in Section \ref{SS:simulationResults} Gurobi solver \cite{Gurobi2015} through CVX \cite{cvx,gb08}.
+
+\paragraph{Bang-bang controller}
+This is the classical controller widely used in private houses to keep temperature within a comfort range.
+It switches the heating system ON when the temperature goes under the temperature lower bound and switches it OFF when the temperature goes over the temperature upper bound.
+The advantage in using this controller is that it is very simple to set up.
+On the other hand it uses more energy than actually needed to achieve the task.
+
+\subsection{Simulation Results}\label{SS:simulationResults} We simulated  DPC in \eqref{E:DPCrealcase} and the bang-bang controller, in closed-loop with the house models $\tT^{r_i}_{k+1|k},\ i=1,\ldots,4$ in $S_1$.
+We considered a sampling time of $10$ minutes and chose $N=4$ as a predictive horizon, i.e.
+$40$ minutes.
+From historical data we got that $\overline{\tX}^c = 0.35 \, m^3/h$ when the heating system is ON and obviously $\underline{\tX}^c = 0 \, m^3/h$ when the heating system is OFF.
+For the temperature comfort range we set a constant upper bound $\overline{\tT}_{\mathrm{k}} = 22.5 \, \degree C$ and a variable lower bound that is $\underline{\tT}_{\mathrm{k}} = 21 \, \degree C$ from $7am$ to $9am$, that is when people in the house wake up and go out for work, and from $6pm$ to midnight, that is when people come back from work and go to sleep, and $\underline{\tT}_{\mathrm{k}} = 20 \, \degree C$ during other hours, i.e.
+when people are either not home or sleep.
+
+We said in Section \ref{SS:descriptionHouse} that the fuel used in the house is vegetable biomass.
+Therefore it is not possible to have ON/OFF switching phases too close to each other, differently from a traditional gas boiler, due to the burning process and heat exchange.
+For this reason we set up both control problems such that when the heating system is activated then it must stay active for at least $20$ minutes.
+This operating period can be obviously adapted depending on the fuel flow rate.
+
+We ran DPC with $3$ different sets of parameters $Q$, $\lambda_{min}$ and $\lambda_{max}$.
+Each set allows a different level of temperature bounds violation.
+In particular we considered a small, a medium and a large violation configuration.
+The simulation period was from May $1^{st}$ $2016$ at $00am$ to May $5^{th}$ $2016$ at $1pm$.
+\paragraph{Result 1} For the sake of clarity, we report in Figure \ref{F:comparison_small} only the comparison for temperature and input schedule obtained allowing small bound violations in DPC.
+We can see that the temperature controlled with DPC does not exit the bounds and if it does then it is around $0.1\,\degree C$.
+Bang-bang control also present small bounds violations due to its working principle.
+We can see how the DPC control law requires the heating system to be ON for less time than the bang-bang one to keep the temperature in the comfort range.
+We will see in Figure \ref{F:comparison_all_energy} that this translates in energy saving.
+
+\begin{figure}[h!]
+	\begin{center}
+	\subfigure[Comparison of temperature of room $3$ obtained with DPC and bang-bang controller.]{
+		\label{F:temperatures_small}
+		\centering
+		\includegraphics[width=26pc]{figures/Temperatures_small.eps}
+	}
+	\subfigure[Input schedules obtained from DPC and bang-bang controller.]{
+		\label{F:inputs_small}
+		\centering
+		\includegraphics[width=26pc]{figures/Inputs_small.eps}
+	}
+	\end{center}
+	\caption{Comparison of DPC and bang-bang control performance.}
+	\captionsetup{justification=centering}
+	\label{F:comparison_small}
+\end{figure}
+
+\paragraph{Result 2} In Figure \ref{F:comparison_all_temperature} a comparison of temperature regulation obtained running the DPC with small, medium and large violations is shown.
+The results show that with the large violation configuration, that gives more importance to the power consumption minimization than to keep temperature within the bounds, temperature is almost always outside the lower bound during the period when the range is tighter.
+However the maximum violation is still lower than $1\,\degree C$.
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[width=26pc]{figures/Temperatures_all.eps}
+	\end{center}
+	\caption{Comparison of DPC control performance in terms of comfort with different violations.}
+	\label{F:comparison_all_temperature}
+\end{figure}
+In Table \ref{T:violationErrors}, MBE and CV(RMSE) violation errors are reported to quantify the bounds violation of DPC, in each of the $3$ configurations, and of bang-bang controller.
+\begin{table}[h!]
+	\centering
+%	\scalebox{0.9}{
+		\begin{tabular}{cccccc}
+			\toprule
+			CONTROLLER  & LBV $\mathrm{MBE}$  & LBV $\mathrm{RMSE}$ & UBV $\mathrm{MBE}$ & UBV $\mathrm{RMSE}$ 	\\ 
+			\midrule
+			$DPC-SV$    & $0.013$             & $0.028$  			      & $0$    				 & $0$     	  	\\
+			$DPC-MV$    & $0.165$ 			  & $0.146$     			  & $0$    				 & $0$		  	\\
+			$DPC-LV$    & $0.410$  			  & $0.224$     			  & $0$    				 & $0$	      	\\
+			$Bang-bang$ & $0.049$ 			  & $0.080$    				  & $0.0063$ 		     & $0.0120$	  	\\
+			\bottomrule
+		\end{tabular}
+%	}
+	\caption{Lower Bound Violation (LBV) and Upper Bound Violation (UBV) errors expressed as $\mathrm{MBE}\%$ and $\mathrm{CV(RMSE)}\%$ for DPC Small Violation (DPC-SV), DPC Medium Violation (DPC-MV), DPC large Violation (DPC-LV) and bang-bang controller.}
+	\captionsetup{justification=centering}
+	\label{T:violationErrors}
+\end{table}
+We can see that if we allow very small violations, DPC outperforms bang-bang controller in terms of comfort guarantees.
+
+\paragraph{Result 3} In Figure \ref{F:comparison_all_energy}, using the neregy model derived from random forest power model $\tP^r$, we show how DPC outperforms the bang-bang controller also in terms of energy consumption and how the bounds violations allows us to save more energy.
+We can see that if we want to keep the temperature within a comfort range, only allowing small violations, the use of DPC produce an energy saving of $44\, kWh$, that correspond to the $33.0\%$ over a period of $4$ days and a half.
+Instead if we allow a large violation over the same period, the energy saving is of $78\, kWh$, that correspond to the $59.0\%$.
+If we consider these results over a monthly period, i.e.
+multiplying them for $6$, we get a monthly energy saving that goes from $264\,kWh$ to $468\,kWh$, with different comfort constraint configurations.
+Considering that the equivalent price of $1\,kWh$ of vegetable biomass is $TOT$\euro{} (TULLIO MAY YOU TELL ME ABOUT THIS PART?), we have a monthly saving of $TOT$\euro{}.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[width=26pc]{figures/Energy_all.eps}
+	\end{center}
+	\caption{Comparison of DPC and bang-bang control performance with different violations using random forest model.}
+	\label{F:comparison_all_energy}
+\end{figure}
+An alternative comparison can be done considering the EnergyPlus model of the house for energy consumption described in Section \ref{SS:energyPlusmodel}.
+Results are shown in Figure \ref{F:comparison_all_energy_E+}.
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[width=26pc]{figures/Energy_all.eps}
+	\end{center}
+	\caption{SUBSTITUTE WITH TULLIO'S RESULTS Comparison of DPC and bang-bang control performance with different violations using EnergyPlus model.}
+	\label{F:comparison_all_energy_E+}
+\end{figure}
+\textbf{TO BE ADAPTED WITH TULLIO'S RESULTS We can see that, over the same period of $4$ days and a half, using DPC we saved $TOT\, kWh$ allowing small violations, that correspond to a $TOT\%$ of energy saving, and $TOT\, kWh$ allowing large violations, that correspond to a $TOT\%$ of energy saving.}

### 2017-09-22 paper/conclusion.tex
completed section 5 and fixed figures problems

@@
+\section{CONCLUSION}
+\label{S:conclusion}
+
 We present two algorithms based on trees and random forests for receding horizon control with data-driven models.
+
 We compare the performance of our Data Predictive Control to MPC on a multivariable bilinear building model.
@@
 On the other hand, DPC with regression trees suffers from practical limitations due to model overfitting.
+
 We further apply DPC with random forests to a large scale 6 story EnergyPlus model with 22 zones for which the traditional model-based control is largely unsuitable due to complex dynamics and the cost of model identification.
@@
 
+We finally apply DPC using data from a real house and compare it with the classical bang-bang controller widely used for temperature control in houses.
+We implemented DPC to guarantee energy saving while guaranteeing thermal comfort in the house.
+We obtained, with respect to a bang-bang controller, an energy saving that goes from the $33\%$, when we want the temperature to be strictly in a comfort range, to the $59\%$, if we allow a small violation of such a range.
+
 DPC has applications which go beyond buildings and energy systems, to industrial process control, and controlling large critical infrastructures like water networks, district heating \& cooling.
 DPC is immensely valuable in situations where first principles based modeling cost is extremely high.
+
+\subsection{Practical Challenges and Future Work}
+\label{SS:challenges}
+\input{cs-challenges.tex}

### 2017-09-22 paper/main.tex
completed section 5 and fixed figures problems

@@
 \usepackage{lineno,hyperref}
+\usepackage{slashbox}
 \modulolinenumbers[5]
+\usepackage{gensymb}
+\usepackage{multirow}
+\usepackage{eurosym}
 
@@
 
-\title{Data Predictive Control: Theory and Applications\tnoteref{mytitlenote}}
+\title{Data Predictive Control with Random Forests:\\Theory and Applications\tnoteref{mytitlenote}}
 \tnotetext[mytitlenote]{The research leading to these results has received funding from the Italian Government under Cipe resolution n.135 (Dec.
@@
 \begin{keyword}
-Energy efficiency, demand response, predictive control, machine learning
+Energy efficiency, demand response, predictive control, machine learning, EnergyPlus modeling, real case study
 \end{keyword}
@@
 \input{casestudyEPlus.tex}
-\clearpage
+
 \input{casestudyReal.tex}

### 2017-09-22 code/data/2 inputs/CreateDPCdata.m

@@ -1,51 +1,0 @@
-%% Convert data to adapt to dpc code
-
-clear all
-close all
-clc
-
-load('Data_PreProccessing/cetemps_interp_10min.mat')
-load('Data_PreProccessing/et_interp_10min.mat')
-load('Data_PreProccessing/ti_interp_10min.mat')
-
-sTrain = 67; % Training start index
-eTrain = 6734; % Training end index
-sTest = 7346; % Testing start index
-eTest = 9505; % Testing end index
-
-%% Creating "tod - time of day" and "dow - day of week" for the training and testing processes
-tTrain = datenum(ti_interp_10min(sTrain:eTrain,1)); % Time vector for the train dataset
-temp = datevec(tTrain);
-todTrain = temp(:,4)';
-[dowTrain ~] = weekday(tTrain);
-dowTrain = dowTrain';
-
-tTest = datenum(ti_interp_10min(sTest:eTest,1));
-temp = datevec(tTest);
-todTest = temp(:,4)';
-[dowTest ~] = weekday(tTest);
-dowTest = dowTest';
-
-clear temp
-
-%% Creates data sets
-
-% Creates training data
-
-x = cell2mat(ti_interp_10min(sTrain:eTrain,2:end))';
-d = cell2mat(cetemps_interp_10min(sTrain:eTrain,2:end))';
-d = [d;todTrain;dowTrain];
-u = cell2mat(et_interp_10min(sTrain:eTrain,[3 5]))';
-y = cell2mat(et_interp_10min(sTrain:eTrain,6))';
-
-save('train.mat','d','u','x','y');
-
-% Creates testing data
-
-x = cell2mat(ti_interp_10min(sTest:eTest,2:end))';
-d = cell2mat(cetemps_interp_10min(sTest:eTest,2:end))';
-d = [d;todTest;dowTest];
-u = cell2mat(et_interp_10min(sTest:eTest,[3 5]))';
-y = cell2mat(et_interp_10min(sTest:eTest,6))';
-
-save('test.mat','d','u','x','y');

### 2017-09-23 paper/casestudyReal.tex
small changes

@@
 \section{REAL CASE STUDY: OPTIMAL HEATING SYSTEM SCHEDULING} In this section we show an application of DPC to a real case study.
-In particular, in the following we first describe the plant structure, then we provide both data-based and EnergyPlus model of such a plant to be used for control and performance comparison, and finally we show results of the closed-loop simulations.
+In particular, in the following we first describe the plant structure, then we provide both data-based and EnergyPlus models of such a plant to be used for control and performance comparison, and finally we show results of the closed-loop simulations.
 
@@
 This is because the EnergyPlus model, due to its internal dynamics based on the house structure, is able to compensate the error over time.
-On the other hand our curve perfectly follows the shape of the real case.
+On the other hand our curve perfectly follows the time behavior of the real case.
 For this reason, in Section \ref{SS:simulationResults}, we will use both models in order to compare the quality of the DPC with respect to a classical bang-bang controller.
@@
 For the temperature comfort range we set a constant upper bound $\overline{\tT}_{\mathrm{k}} = 22.5 \, \degree C$ and a variable lower bound that is $\underline{\tT}_{\mathrm{k}} = 21 \, \degree C$ from $7am$ to $9am$, that is when people in the house wake up and go out for work, and from $6pm$ to midnight, that is when people come back from work and go to sleep, and $\underline{\tT}_{\mathrm{k}} = 20 \, \degree C$ during other hours, i.e.
-when people are either not home or sleep.
+when people are either not at home or asleep.
 
-We said in Section \ref{SS:descriptionHouse} that the fuel used in the house is vegetable biomass.
+We explained in Section \ref{SS:descriptionHouse} that the fuel used in the house is vegetable biomass.
 Therefore it is not possible to have ON/OFF switching phases too close to each other, differently from a traditional gas boiler, due to the burning process and heat exchange.
-For this reason we set up both control problems such that when the heating system is activated then it must stay active for at least $20$ minutes.
+For this reason we set up both control problems with the constraint that when the heating system is activated it must stay active for at least $20$ minutes.
 This operating period can be obviously adapted depending on the fuel flow rate.
@@
 \paragraph{Result 1} For the sake of clarity, we report in Figure \ref{F:comparison_small} only the comparison for temperature and input schedule obtained allowing small bound violations in DPC.
-We can see that the temperature controlled with DPC does not exit the bounds and if it does then it is around $0.1\,\degree C$.
-Bang-bang control also present small bounds violations due to its working principle.
+We can see that the temperature controlled with DPC does not violate the bounds and if it does then the violation is around $0.1\,\degree C$.
+Bang-bang control also presents small bounds violations due to its working principle.
 We can see how the DPC control law requires the heating system to be ON for less time than the bang-bang one to keep the temperature in the comfort range.
@@
 
-\paragraph{Result 3} In Figure \ref{F:comparison_all_energy}, using the neregy model derived from random forest power model $\tP^r$, we show how DPC outperforms the bang-bang controller also in terms of energy consumption and how the bounds violations allows us to save more energy.
+\paragraph{Result 3} In Figure \ref{F:comparison_all_energy}, using the energy model derived from random forest power model $\tP^r$, we show how DPC outperforms the bang-bang controller also in terms of energy consumption and how the bounds violations allows us to save more energy.
 We can see that if we want to keep the temperature within a comfort range, only allowing small violations, the use of DPC produce an energy saving of $44\, kWh$, that correspond to the $33.0\%$ over a period of $4$ days and a half.

### 2017-09-23 paper/abstract.tex
added abstract

@@
-Decisions on how to best operate large complex plants such as natural gas processing, oil refineries, and energy efficient buildings are becoming ever so complex that model-based predictive control (MPC) algorithms must play an important role.
-However, a key factor prohibiting the widespread adoption of MPC, is the cost, time, and effort associated with learning first-principles dynamical models of the underlying physical system.
-An alternative approach is to employ learning algorithms to build black-box models which rely only on real-time data from the sensors.
-Machine learning is widely used for regression and classification, but thus far data-driven models have not been used for closed-loop control.
-We present novel Data Predictive Control (DPC) algorithms that use Regression Trees and Random Forests for receding horizon control.
-%DPC also utilizes Gaussian Processes for smart experiment design by recommending how the accuracy of the current black-box models can be improved. 
-We demonstrate the strength of our approach with a case study on a bilinear building model identified using real weather data and sensor measurements. 
-In a one-to-one comparison, we show that DPC explains 70\% variation in the MPC controller.
-%We validate our results by a one-to-one comparison against a benchmark MPC controller. 
-We further apply DPC to a large scale multi-story EnergyPlus building model to curtail total power consumption in a Demand Response setting. 
-In such cases, when the model-based controllers fail due to modeling cost, complexity and scalability, our results show that DPC curtails the desired power usage with high confidence.
+Model-based control techniques have been widely used over the past years in control systems to improve system performance of simple controllers, such as PID, bang-bang, etc.
+Among these, one of the most popular is Model Predictive Control (MPC), due to its capacity to optimally control a system, based on the capability to predict its behavior over a future horizon.
+A key factor prohibiting the widespread adoption of MPC for complex systems is related to the difficulties (cost, time, and effort) associated with the model identification.
+Data-driven models have been intensively studied in the last year to overcome this problem.
+However, closed-loop control systems based on such models have not received much attention.
+We introduce a novel idea for predictive control based on system data, Data Predictive Control (DPC), leveraging machine learning algorithms, in particular regression trees and random forests.
+We show, using a bilinear model, that DPC has comparable performance with MPC, and apply it to the Demand Response problem of a building and to find the optimal ON/OFF scheduling of the heating system of a real house in order to minimize the power consumption while guaranteeing thermal comfort for occupants.

### 2017-09-26 paper/abstract.tex
completed abstract, section 5 and conclusion

@@
 We introduce a novel idea for predictive control based on system data, Data Predictive Control (DPC), leveraging machine learning algorithms, in particular regression trees and random forests.
-We show, using a bilinear model, that DPC has comparable performance with MPC, and apply it to the Demand Response problem of a building and to find the optimal ON/OFF scheduling of the heating system of a real house in order to minimize the power consumption while guaranteeing thermal comfort for occupants.
+We show, using a bilinear model, that DPC has comparable performance with MPC.
+We apply it to the Demand Response problem of a large scale multi-story EnergyPlus building model, and show that DPC curtails the desired power usage with high confidence.
+Finally we apply it to find the optimal ON/OFF scheduling of the heating system of a real house in order to minimize the power consumption while guaranteeing thermal comfort for occupants.
+Results show that, compared to a bang-bang controller, with DPC we obtain an energy saving in terms of $kWh$ that goes from $25,4\%$ to $49.2\%$.

### 2017-09-26 paper/casestudyReal.tex
completed abstract, section 5 and conclusion

@@
 The probes positions were chosen to optimize the data acquisition and to minimize the discomfort of the occupants in the house.
-Furthermore, as can be seen in Figures \ref{F:housePlantScheme} and \ref{F:houseFloors} (COMMENT: NOT CLEAR WHERE THE SENSORS ARE IN THE FIGURES.
-MORE DETAILS NEEDED), a commercial heat meter (temperature range from $10.0\degree C$ to $90.0\degree C$ with an accuracy of $\pm 0.05\degree C$) was installed downstream of the biomass boiler to measure the produced thermal energy.
+Furthermore, as can be seen in Figures \ref{F:housePlantScheme} and \ref{F:houseFloors}, a commercial heat meter (temperature range from $10.0\degree C$ to $90.0\degree C$ with an accuracy of $\pm 0.05\degree C$) was installed downstream of the biomass boiler to measure the produced thermal energy.
+The turbine flowmeter was installed on the return pipe, while the two thermocouples were placed inside the delivery and return pipes, respectively.
 All the measuring devices have been set with a data acquisition rate equal to $10$ minutes, from March $11^{th}$ $2016$ to May $15^{th}$ $2016$.
@@
 \subsection{EnergyPlus model}\label{SS:energyPlusmodel}
-To assess the energy performance of the use case's heating system, the EnergyPlus engine coupled with DesignBuilder modeling environment was employed.
-Based on the weather data provided by CETEMPS – Centre of Excellence \cite{CETEMPS}, shown in Figure \ref{F:houseExternalWeather}, the performed dynamic simulation was based on a weather file specifically created for L'Aquila, that according to the K\"{o}ppen-Geiger climate classification, has a warm temperate climate (Csa, Cfa and Csb) \cite{Peel2007}.
-(COMMENT: TO BE BETTER WRITTEN)
+In this section we create an EnergyPlus model of the thermal energy consumption of the house that will be used in Section \ref{SS:simulationResults} to compare the quality of the DPC with respect to a classical bang-bang controller.
+
+To assess energy performance of the use case heating system, the EnergyPlus engine together with DesignBuilder modeling environment has been employed.
+The model has been created according to the K\"{o}ppen-Geiger climate classification that, based weather data of L'Aquila shown in Figure \ref{F:houseExternalWeather}, provided by the CETEMPS – Centre of Excellence \cite{CETEMPS}, classifies Italy in Csa, Cfa and Csb climate zones, with a warm temperate climate [4], .
 
@@
 	\begin{center}
-		\includegraphics[width=26pc]{figures/dati_climatici_rev01.eps}
+		\includegraphics[width=30pc]{figures/dati_climatici_rev01.eps}
 		\caption{Weather data of L'Aquila.}
@@
 		\caption{Actual scheduling of the heating system, where white color means switched OFF and green color switched ON, and ($T_{ext,ave}$) is the external average temperature.
-From midnight to $8am$ the heating system is always switched off.)}
+From midnight to $8am$ the heating system is always switched off.}
 		\captionsetup{justification=centering}
@@
 \end{figure}
-Following the hourly calibration proposed by the M\&V guidelines of ASHRAE \cite{USDOE}, a simulated model is calibrated when the mean bias error (MBE) and the coefficient of variation of the root mean square error [CV(RMSE)] are less than acceptable tolerances, respectively equal to $\pm 10.0\%$ and $30.0\%$.
+Following the hourly calibration proposed by the M\&V guidelines of ASHRAE \cite{USDOE}, also applied in \cite{Mustafarai2014,Raftery2011}, a simulated model is calibrated when the mean bias error (MBE) and the coefficient of variation of the root mean square error [CV(RMSE)] are less than acceptable tolerances, respectively equal to $\pm 10.0\%$ and $30.0\%$.
 MBE and CV(RMSE) have been calculated by using Equations \eqref{Eq:mbe} and \eqref{Eq:cvrmse}.
@@
 
-We use random forest model $\tP^r$ also to estimate the energy consumption of the house.
+%We use random forest model $\tP^r$ also to estimate the energy consumption of the house.
 In particular we estimated it for the same period analyzed in Section \ref{SS:energyPlusmodel} and obtained the results in Figure \ref{F:Energy_testing}, with a MBE equal to $13.0\%$ and a CV(RMSE) equal to $13.7\%$.
-
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=26pc]{figures/Energy_test.eps}
-		\caption{Energy consumption accuracy.}
-		\captionsetup{justification=centering}
-		\label{F:Energy_testing}
-	\end{center}
-\end{figure}
-
-As we can see, with random forest model we get a worse MBE and CV(RMSE) for energy consumption, although the maximum error is very similar.
+%
+%\begin{figure}[h!]
+%	\begin{center}
+%		\includegraphics[width=26pc]{figures/Energy_test.eps}
+%		\caption{Energy consumption accuracy.}
+%		\captionsetup{justification=centering}
+%		\label{F:Energy_testing}
+%	\end{center}
+%\end{figure}
+%
+%As we can see, with random forest model we get a worse MBE and CV(RMSE) for energy consumption, although the maximum error is very similar.
 This is because the EnergyPlus model, due to its internal dynamics based on the house structure, is able to compensate the error over time.
@@
 In particular we considered a small, a medium and a large violation configuration.
-The simulation period was from May $1^{st}$ $2016$ at $00am$ to May $5^{th}$ $2016$ at $1pm$.
 \paragraph{Result 1} For the sake of clarity, we report in Figure \ref{F:comparison_small} only the comparison for temperature and input schedule obtained allowing small bound violations in DPC.
+For the same reason the simulation period is restricted to $4$ days and a half, from May $1^{st}$ $2016$ at $00am$ to May $5^{th}$ $2016$ at $1pm$, in order to make the plots more clear.
 We can see that the temperature controlled with DPC does not violate the bounds and if it does then the violation is around $0.1\,\degree C$.
@@
 We can see how the DPC control law requires the heating system to be ON for less time than the bang-bang one to keep the temperature in the comfort range.
-We will see in Figure \ref{F:comparison_all_energy} that this translates in energy saving.
+We will see in Figure \ref{F:comparison_all_energy_E+} that this translates in energy saving.
 
@@
 
-\paragraph{Result 3} In Figure \ref{F:comparison_all_energy}, using the energy model derived from random forest power model $\tP^r$, we show how DPC outperforms the bang-bang controller also in terms of energy consumption and how the bounds violations allows us to save more energy.
+\paragraph{Result 3} 
+%In Figure \ref{F:comparison_all_energy_E+}, using the energy model derived from random forest power model $\tP^r$, we show how DPC outperforms the bang-bang controller also in terms of energy consumption and how the bounds violations allow us to save more energy.
 We can see that if we want to keep the temperature within a comfort range, only allowing small violations, the use of DPC produce an energy saving of $44\, kWh$, that correspond to the $33.0\%$ over a period of $4$ days and a half.
@@
 
+%\begin{figure}[h!]
+%	\begin{center}
+%		\includegraphics[width=26pc]{figures/Energy_all.eps}
+%	\end{center}
+%	\caption{Comparison of DPC and bang-bang control performance with different violations using random forest model.}
+%	\label{F:comparison_all_energy}
+%\end{figure}
+%An alternative comparison can be done considering the EnergyPlus model of the house for energy consumption described in Section \ref{SS:energyPlusmodel}.
+Results are shown in Figure \ref{F:comparison_all_energy_E+}.
+In Figure \ref{F:comparison_all_energy_E+}, using the thermal energy consumption model derived in Section \ref{SS:energyPlusmodel} using EnergyPlus, we show how DPC outperforms the bang-bang controller also in terms of energy consumption and how the bounds violations allow us to save more energy.
+In this case, since the plot is clear and we are interested in showing energy saving on a long period, we ran the simulations over the testing data, i.e.
+$15$ days of May, from May $1^{st}$ $2016$ to May $15^{th}$ $2016$.
+We can see that if we want to keep the temperature within a comfort range, only allowing small violations, the use of DPC produce an energy saving of $45\, kWh$, that correspond to the $25.4\%$ over a period of $15$ days.
+Instead if we allow a large violation over the same period, the energy saving is of $87\, kWh$, that correspond to the $49.2\%$.
 \begin{figure}[h!]
 	\begin{center}
-		\includegraphics[width=26pc]{figures/Energy_all.eps}
+		\includegraphics[width=26pc]{figures/Energy_all_EnergyPlus.eps}
 	\end{center}
-	\caption{Comparison of DPC and bang-bang control performance with different violations using random forest model.}
-	\label{F:comparison_all_energy}
-\end{figure}
-An alternative comparison can be done considering the EnergyPlus model of the house for energy consumption described in Section \ref{SS:energyPlusmodel}.
-Results are shown in Figure \ref{F:comparison_all_energy_E+}.
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=26pc]{figures/Energy_all.eps}
-	\end{center}
-	\caption{SUBSTITUTE WITH TULLIO'S RESULTS Comparison of DPC and bang-bang control performance with different violations using EnergyPlus model.}
+	\caption{Comparison of DPC and bang-bang controller performance in terms of thermal energy saving with different violations using EnergyPlus model.}
 	\label{F:comparison_all_energy_E+}
 \end{figure}
-\textbf{TO BE ADAPTED WITH TULLIO'S RESULTS We can see that, over the same period of $4$ days and a half, using DPC we saved $TOT\, kWh$ allowing small violations, that correspond to a $TOT\%$ of energy saving, and $TOT\, kWh$ allowing large violations, that correspond to a $TOT\%$ of energy saving.}

### 2017-09-26 paper/conclusion.tex
completed abstract, section 5 and conclusion

--- 2017-09-17 paper/conclusion.tex
@@
 
+We finally apply DPC using data from a real house and compare it with the classical bang-bang controller widely used for temperature control in houses.
+We implemented DPC to guarantee energy saving while guaranteeing thermal comfort for the occupants.
+We obtained, with respect to a bang-bang controller, an energy saving that goes from the $25.4\%$, when we want the temperature to be strictly in a comfort range, to the $49.2\%$, if we allow little violations of such a range.
+
 DPC has applications which go beyond buildings and energy systems, to industrial process control, and controlling large critical infrastructures like water networks, district heating \& cooling.
 DPC is immensely valuable in situations where first principles based modeling cost is extremely high.
+
+\subsection{Practical Challenges and Future Work}
+\label{SS:challenges}
+\input{cs-challenges.tex}

### 2017-09-26 paper/dpc.tex
completed abstract, section 5 and conclusion

@@
 \begin{align}
-f(x_k,u_k,d_k)= g(d_k,x_k,h(u_k)),
+f(x_k,u_k,d_k)= g(x_k,h(u_k),d_k),
 \label{E:sepvars}
@@
 The problem with trees is their high variance and that they can overfit the data easily.
-A small change $\delta$ in the data can result in a different series of splits and thus violate the acceptable accuracy $\eps$, i.e.
-$\exists \  \hat{\tX}^d \ | \ ||\tX^d-\hat{\tX}^d|| < \delta$ \& $||\tY-\tY_{\mathrm{true}}||>\eps$.
+A small change $\delta$ in the data can result in a different series of splits and thus violate the acceptable accuracy $\eps$.
 This is the price to be paid for estimating a tree-based structure from the data.
@@
 \end{gather}
-which, again, is trained only on $\mathsf{X}^d$, but $\tX^d \in \R^p \subset \R^d$ for each tree,  and then fit a linear regression model using $\mathsf{X}^c$ in every leaf of every tree.
+which, again, is trained only on $\mathsf{X}^d$, with $\tX^d \in \R^p \subset \R^d$ for each tree,  and then fit a linear regression model using $\mathsf{X}^c$ in every leaf of every tree.
 We build $N$ such forests for $N$ prediction steps such that the leaf $R_i$ of forest $\mathcal{R}_j$ uses a linear model

### 2017-09-26 paper/intro.tex
completed section 5 and fixed figures problems
completed abstract, section 5 and conclusion

@@
 
-In our previous work~\cite{JainACC2017,Jain2016}, we introduced the concept of DPC for receding horizon control.
+In our previous work~\cite{Behl2016,Jain2016,JainACC2017,JainCDC2017}, we introduced the concept of DPC for receding horizon control.
 This work has the following contributions.
-(1) We first formally present two underlying algorithms: (i) DPC with regression trees, and (ii) DPC with random forests, which also ensure recursive feasibility in receding horizon control.
+(1) We first formally present two underlying algorithms: (i) DPC with regression trees, and (ii) DPC with random forests.
 (2) Using a bilinear building model whose parameters were identified using experiments on a building in Switzerland, we demonstrate the strength of DPC for receding horizon control via one-to-one comparison against a benchmark MPC controller.

### 2017-09-26 paper/Highlights.tex
added Highlights file

\documentclass[12pt]{article}

\begin{document}

\title{Highlights}
\date{}
\maketitle

\thispagestyle{empty}
\begin{itemize}
	% Remember maximum 85 characters (included spaces, points, commas, etc.)
	\item Predictive control strategy using data-based models via machine learning algorithms.
	\item Data Predictive Control shows comparable performance to Model Predictive Control.
	\item Data Predictive Control shows good performance in Demand Response curtailment.
	\item Data Predictive Control allows energy saving up to $50\%$ on the considered case study.
\end{itemize}

\end{document}

### 2017-09-26 paper/casestudyReal.tex
Modified introduction. First complete draft.

@@
-\section{REAL CASE STUDY: OPTIMAL HEATING SYSTEM SCHEDULING} In this section we show an application of DPC to a real case study.
+\section{REAL CASE STUDY: OPTIMAL HEATING SYSTEM SCHEDULING}
+\label{S:realCaseStudy}
+
+In this section we show an application of DPC to a real case study.
 In particular, in the following we first describe the plant structure, then we provide both data-based and EnergyPlus models of such a plant to be used for control and performance comparison, and finally we show results of the closed-loop simulations.
@@
 			\centering
-			\includegraphics[width=24pc]{figures/blocco_parete_rev01.eps}
+			\includegraphics[width=20pc]{figures/blocco_parete_rev01.eps}
 		}
@@
 			\centering
-			\includegraphics[width=24pc]{figures/blocco_solaio_rev01.eps}
+			\includegraphics[width=20pc]{figures/blocco_solaio_rev01.eps}
 		}
 	\end{center}
+	\vspace{-0.5cm}
 	\caption{Cross-sections of the wood-cement blocks.}
@@
 	\begin{center}
-		\includegraphics[width=26pc]{figures/Schema_di_impianto.eps}
+		\includegraphics[width=27pc]{figures/Schema_di_impianto.eps}
 		\caption{Technological plant scheme of the use case.}
@@
 	\begin{center}
-		\includegraphics[width=24pc]{figures/Render_rev01.eps}
+		\includegraphics[width=20pc]{figures/Render_rev01.eps}
 		\caption{Virtual model of the use case.}
@@
 \end{figure}
-\subsection{Random forest models} For the closed-loop simulations with DPC we create $2$ different sets of models, $S_1 = \{\tP^{r}_{\mathrm{k+1|k}},\tT^{r_1}_{\mathrm{k+1|k}},\tT^{r_2}_{\mathrm{k+1|k}},\tT^{r_3}_{\mathrm{k+1|k}},\tT^{r_4}_{\mathrm{k+j|k}}\}$ and $S_2 = \{\tP_{\mathrm{k+j|k}},\tT^{1}_{\mathrm{k+j|k}},\tT^{2}_{\mathrm{k+j|k}},\tT^{3}_{\mathrm{k+j|k}},\\\tT^{4}_{\mathrm{k+j|k}},\ j = 1,\ldots,N\}$, using random forests.
+\subsection{Random forest models} For the closed-loop simulations with DPC we create $2$ different sets of models, $S_1 = \{\tP^{r}_{\mathrm{k+1|k}},\tT^{r_1}_{\mathrm{k+1|k}},\tT^{r_2}_{\mathrm{k+1|k}},\tT^{r_3}_{\mathrm{k+1|k}},\tT^{r_4}_{\mathrm{k+j|k}}\}$ and $S_2 = \{\tP_{\mathrm{k+j|k}},\tT^{1}_{\mathrm{k+j|k}},\tT^{2}_{\mathrm{k+j|k}},\\\tT^{3}_{\mathrm{k+j|k}},\tT^{4}_{\mathrm{k+j|k}},\ j = 1,\ldots,N\}$, using random forests.
 In each set we have a model for the power consumption of the house ($\tP^{r}$ and $\tP$) and $4$ models that describe the evolution of the room temperature in each of the $4$ rooms ($\tT^{r_i}$ and $\tT^i$).
@@
 We can see that if we want to keep the temperature within a comfort range, only allowing small violations, the use of DPC produce an energy saving of $45\, kWh$, that correspond to the $25.4\%$ over a period of $15$ days.
+In case of medium violations get an energy saving of $57\, kWh$, that is $32.2\%$.
 Instead if we allow a large violation over the same period, the energy saving is of $87\, kWh$, that correspond to the $49.2\%$.

### 2017-09-26 paper/intro.tex
Modified introduction. First complete draft.

@@
 The first and foremost requirement for making any decision is to obtain the underlying control-oriented predictive model of the system.
-With a reasonable forecast of the external disturbances, these models should predict the state of the system in the future and thus a predictive controller based on Model Predictive Control (MPC) can act preemptively to provide a desired behavior.
+With a reasonable forecast of the external disturbances, these models should predict the state of the system in the future and thus Model Predictive Control (MPC) can act preemptively to provide a desired system behavior while optimizing a desired performance.
 In particular, MPC has been proven to be very powerful for multivariable systems in the presence of input and output constraints, and forecast of the disturbances.
@@
 
-The question now is, can we employ data-driven techniques to reduce the cost of modeling, and still exploit the benefits that MPC has to offer?
-We therefore look for automatic and data-driven approaches to control that are also adaptive, scalable and interpretable.
-We solve this problem with \textit{Data Predictive Control (DPC)} by bridging the gap between Machine Learning and Predictive Control.
+The question now is, can we employ data-driven techniques to reduce the cost of modeling and still exploit the benefits that MPC has to offer?
+Can we build automatic and data-driven approaches for control purposes that are also adaptive, scalable and interpretable?
+In this paper we address this problem introducing \textit{Data Predictive Control (DPC)} to bridge the gap between Machine Learning and Predictive Control.
 
+In literature several studies can be found dealing with data-based control techniques that make use of machine learning algorithms \cite{Hou2013}.
+For example in \cite{Afram2017} Artificial Neural Networks are exploited to create data-based system models to be used as plant simulator in closed-loop with a Suervisory Model Predictive Control.
+In \cite{Macarulla2017} the authors proposed a predictive control strategy based on Neural Networks, for boilers control in buildings, to decide the optimal time to switch the plan on to guarantee energy saving and thermal comfort.
+However the strategy is not easily scalable to different types of plants and does not use optimization theory in the closed-loop scheme.
+In \cite{Costanzo2016} a reinforcement learning control strategy, called Model-Assisted Batch Reinforcement Learning, is considered to provide data-driven control for the demand response problem in HVAC systems.
+Reinforcement Learning is a model-free methodology and for this reason it can not be used for on-line optimization in a closed-loop predictive control scheme.
+In \cite{Ferreira2012} the authors considered a data-driven predictive control based on Neural Networks.
+They use this approach to guarantee energy saving and  thermal comfort in public buildings.
+Neural Networks are used in the closed-loop control scheme to determine a thermal comfort index based on parameters that can be measured or estimated, but no  system dynamics with internal state are included into the optimization problem.
+More papers related to this topic can be found in literature, but to the best of the authors knowledge none of them address the problem of including data-based state models in the optimal predictive control loop, and hence allow to set up a MPC-like optimal control problem, as we do in DPC.
+
+\paragraph{Main contributions and paper organization}
 In our previous work~\cite{Behl2016,Jain2016,JainACC2017,JainCDC2017}, we introduced the concept of DPC for receding horizon control.
-This work has the following contributions.
-(1) We first formally present two underlying algorithms: (i) DPC with regression trees, and (ii) DPC with random forests.
-(2) Using a bilinear building model whose parameters were identified using experiments on a building in Switzerland, we demonstrate the strength of DPC for receding horizon control via one-to-one comparison against a benchmark MPC controller.
-We show DPC captures 70\% variance in MPC and offers a comparable performance.
-(3) We present a practical application of DPC for Demand Response, where we apply DPC to a 6 story 22 zone building model in EnergyPlus \cite{Crawley2001} for which model-based control is not economical and practical due to extreme complexity.
+In this paper we extend these results providing the following contributions:
+\begin{enumerate}
+	\item in Section \ref{S:dpc} we formally present the two following control techniques:
+	\begin{itemize}
+		\item DPC with regression trees;
+		\item DPC with random forests;
+	\end{itemize}
+	\item in Section \ref{S:proof}, using a bilinear building model whose parameters were identified using experiments on a building in Switzerland, we demonstrate the strength of DPC for receding horizon control via one-to-one comparison against a benchmark MPC controller.
+We show DPC captures 70\% variance in MPC and offers a comparable performance;
+	\item in Section \ref{S:casestudy}, we present a practical application of DPC for Demand Response, where we apply DPC to a 6 story 22 zone building model in EnergyPlus \cite{Crawley2001} for which model-based control is not economical and practical due to extreme complexity.
 We show scalability and efficiency of DPC in providing financial incentives to the end-customers bypassing the need for high fidelity models.
-We observe that DPC provides the desired power reduction with an average error of 3\%.
+We observe that DPC provides the desired power reduction with an average error of 3\%;
+	\item in Section \ref{S:realCaseStudy}, we implement DPC on real data coming from an off-grid house located in L'Aquila (Italy), to find the optimal ON/OFF scheduling for the heating system in order to save energy while guaranteeing thermal comfort for the occupants.
+We quantify the total amount of energy saved, with respect to the classical bang-bang controller widely used in houses for temperature control, using an EnergyPlus model built specifically for the house.
+We show we can perform an energy saving that goes from $25\%$, if we guarantee thermal comfort, to $50\%$, if we allow very little discomfort in terms of rooms temperature.
+\end{enumerate}
 

### 2017-09-27 paper/abstract.tex
edited abstract+intro

--- 2017-09-23 paper/abstract.tex
@@
 We introduce a novel idea for predictive control based on system data, Data Predictive Control (DPC), leveraging machine learning algorithms, in particular regression trees and random forests.
-We show, using a bilinear model, that DPC has comparable performance with MPC, and apply it to the Demand Response problem of a building and to find the optimal ON/OFF scheduling of the heating system of a real house in order to minimize the power consumption while guaranteeing thermal comfort for occupants.
+We show, using a bilinear model, that DPC has comparable performance with MPC.
+We further apply DPC to the Demand Response problem of a large scale multi-story EnergyPlus building model, and show that DPC curtails the desired power usage with high confidence.
+Finally, we use DPC to optimize ON/OFF scheduling of the heating system of a real house in order to minimize the power consumption while guaranteeing thermal comfort for occupants.
+Our results show that with DPC we obtain an energy saving in terms of $kWh$ that goes from $25.4\%$ to $49.2\%$ when compared to a bang-bang controller.

### 2017-09-27 paper/intro.tex
edited abstract+intro

@@
 
-In literature several studies can be found dealing with data-based control techniques that make use of machine learning algorithms \cite{Hou2013}.
-For example in \cite{Afram2017} Artificial Neural Networks are exploited to create data-based system models to be used as plant simulator in closed-loop with a Suervisory Model Predictive Control.
-In \cite{Macarulla2017} the authors proposed a predictive control strategy based on Neural Networks, for boilers control in buildings, to decide the optimal time to switch the plan on to guarantee energy saving and thermal comfort.
-However the strategy is not easily scalable to different types of plants and does not use optimization theory in the closed-loop scheme.
+In the literature, several studies deal with the data-driven control techniques using machine learning algorithms \cite{Hou2013}.
+For example, Artificial Neural Networks are exploited to create data-driven models to be used as plant simulator in closed-loop with a Supervisory Model Predictive Control in \cite{Afram2017}.
+In \cite{Macarulla2017} the authors proposed a predictive control strategy based on Neural Networks, for boilers control in buildings, to decide the optimal time to switch-on the plan to guarantee energy saving and thermal comfort.
+However, the strategy is not easily scalable to different types of plants and does not use optimization theory in the closed-loop scheme.
 In \cite{Costanzo2016} a reinforcement learning control strategy, called Model-Assisted Batch Reinforcement Learning, is considered to provide data-driven control for the demand response problem in HVAC systems.
 Reinforcement Learning is a model-free methodology and for this reason it can not be used for on-line optimization in a closed-loop predictive control scheme.
-In \cite{Ferreira2012} the authors considered a data-driven predictive control based on Neural Networks.
-They use this approach to guarantee energy saving and  thermal comfort in public buildings.
+In \cite{Ferreira2012} the authors considered a data-driven predictive control based on Neural Networks to guarantee energy saving and  thermal comfort in public buildings.
 Neural Networks are used in the closed-loop control scheme to determine a thermal comfort index based on parameters that can be measured or estimated, but no  system dynamics with internal state are included into the optimization problem.
-More papers related to this topic can be found in literature, but to the best of the authors knowledge none of them address the problem of including data-based state models in the optimal predictive control loop, and hence allow to set up a MPC-like optimal control problem, as we do in DPC.
+More papers related to this topic can be found in the literature, but to the best of the authors' knowledge none of them address the problem of including data-driven state models in the optimal predictive control loop, and hence allow to set up a MPC-like optimal control problem, as we do in DPC.
 
-\paragraph{Main contributions and paper organization}
-In our previous work~\cite{Behl2016,Jain2016,JainACC2017,JainCDC2017}, we introduced the concept of DPC for receding horizon control.
-In this paper we extend these results providing the following contributions:
+\subsection{Main contributions and paper organization}
+In our previous work \cite{Behl2016,Jain2016,JainACC2017,JainCDC2017}, we introduced the concept of DPC for receding horizon control.
+In this paper, we extend these results providing the following contributions:
 \begin{enumerate}
-	\item in Section \ref{S:dpc} we formally present the two following control techniques:
+	\item In Section \ref{S:dpc}, we formally present the two following control techniques:
 	\begin{itemize}
-		\item DPC with regression trees;
-		\item DPC with random forests;
+		\item DPC with regression trees, and
+		\item DPC with random forests.
 	\end{itemize}
-	\item in Section \ref{S:proof}, using a bilinear building model whose parameters were identified using experiments on a building in Switzerland, we demonstrate the strength of DPC for receding horizon control via one-to-one comparison against a benchmark MPC controller.
-We show DPC captures 70\% variance in MPC and offers a comparable performance;
-	\item in Section \ref{S:casestudy}, we present a practical application of DPC for Demand Response, where we apply DPC to a 6 story 22 zone building model in EnergyPlus \cite{Crawley2001} for which model-based control is not economical and practical due to extreme complexity.
+	\item In Section \ref{S:proof}, we demonstrate the strength of DPC for receding horizon control via one-to-one comparison against a benchmark MPC controller using a bilinear building model whose parameters were identified using experiments on a building in Switzerland.
+We show that DPC captures 70\% variance in MPC and offers a comparable performance.
+	\item Section \ref{S:casestudy} describes a practical application of DPC for Demand Response, where we apply DPC to a 6 story 22 zone building model in EnergyPlus \cite{Crawley2001} for which model-based control is not economical and practical due to extreme complexity.
 We show scalability and efficiency of DPC in providing financial incentives to the end-customers bypassing the need for high fidelity models.
-We observe that DPC provides the desired power reduction with an average error of 3\%;
-	\item in Section \ref{S:realCaseStudy}, we implement DPC on real data coming from an off-grid house located in L'Aquila (Italy), to find the optimal ON/OFF scheduling for the heating system in order to save energy while guaranteeing thermal comfort for the occupants.
+We observe that DPC provides the desired power reduction with an average error of 3\%.
+	\item In Section \ref{S:realCaseStudy}, we implement DPC on the real data from an off-grid house located in L'Aquila (Italy), to find the optimal ON/OFF scheduling for the heating system in order to save energy while guaranteeing thermal comfort for the occupants.
 We quantify the total amount of energy saved, with respect to the classical bang-bang controller widely used in houses for temperature control, using an EnergyPlus model built specifically for the house.
-We show we can perform an energy saving that goes from $25\%$, if we guarantee thermal comfort, to $50\%$, if we allow very little discomfort in terms of rooms temperature.
+We show we can perform an energy saving that goes from $25\%$, if we guarantee thermal comfort, to $50\%$, if we allow very little discomfort in terms of rooms' temperature.
 \end{enumerate}

### 2017-09-27 paper/intro.tex
Couple of small changes in abstract, intro and keywords

@@
 We quantify the total amount of energy saved, with respect to the classical bang-bang controller widely used in houses for temperature control, using an EnergyPlus model built specifically for the house.
-We show we can perform an energy saving that goes from $25\%$, if we guarantee thermal comfort, to $50\%$, if we allow very little discomfort in terms of rooms' temperature.
+We show we can perform an energy saving that goes from $25.4\%$, if we guarantee thermal comfort, to $49.2\%$, if we allow very little discomfort in terms of rooms' temperature.
 \end{enumerate}

### 2017-09-29 paper/casestudyEPlus.tex
Modified introduction. First complete draft.
Added flow scheme, motivation. Implemented comments from Italy.

@@
 
-Fig.~\ref{F:sepvars} shows the prediction accuracy for the power forest, and also explains the two level training approach introduced in Sec.~\ref{SS:sepvar}.
+Figure~\ref{F:sepvars} shows the prediction accuracy for the power forest, and also explains the two level training approach introduced in Section~\ref{SS:sepvar}.
 During S1, the forests are trained using only disturbances as the features.
@@
 	\centering
-	\hspace{-0.3cm}\includegraphics[width=22pc]{figures/eplus_validation.eps}	
+	\includegraphics[width=24pc]{figures/eplus_validation.eps}	
 	\caption{Model accuracy during training: The prediction made by forest using only $\tX^d$ (red) captures the effect due to disturbances.
@@
 
-The results are shown in Fig.~\ref{F:eplus_track}. 
+The results are shown in Figure~\ref{F:eplus_track}. 
 The DPC controller is active between 7am - 2pm. Before 7am and after 2pm, the building is using a predefined rule-based control strategy.
-The optimal control inputs from DPC-En are shown in Fig.~\ref{F:eplus_inputs}.
+The optimal control inputs from DPC-En are shown in Figure~\ref{F:eplus_inputs}.
 It is observed that, with the optimal inputs generated by DPC, we can track the reference power consumption signal closely.
 In fact, the average tracking error between 7am - 2pm is 3\%.
-The difference between the predicted power consumption and that in the closed-loop simulation in Fig.~\ref{F:eplus_power} is due to model mismatch between the EnergyPlus model and the power forest used in the optimization \eqref{E:ptrack}.
+The difference between the predicted power consumption and that in the closed-loop simulation in Figure~\ref{F:eplus_power} is due to model mismatch between the EnergyPlus model and the power forest used in the optimization \eqref{E:ptrack}.
 Due to this inaccuracy, the actual power consumption is on an average 7 kW higher.

### 2017-09-29 paper/casestudyReal.tex
awesome figure
Added flow scheme, motivation. Implemented comments from Italy.

@@
-\section{REAL CASE STUDY: OPTIMAL HEATING SYSTEM SCHEDULING}
+\section{CASE STUDY: OPTIMAL HEATING SYSTEM SCHEDULING}
 \label{S:realCaseStudy}
@@
 In this section we show an application of DPC to a real case study.
-In particular, in the following we first describe the plant structure, then we provide both data-based and EnergyPlus models of such a plant to be used for control and performance comparison, and finally we show results of the closed-loop simulations.
+The section organization is shown in Figure \ref{F:overview}.
+More precisely, we consider a real house located in L'Aquila, Italy, as a case study that we describe in Section \ref{SS:descriptionHouse}.
+Using historical data from this house, random forest models are built in Section \ref{SS:randomForestsModels} to be used either as plant or as predictive models for DPC.
+In Section \ref{SS:controllersDPCandBangbang}, DPC and bang-bang control strategies are defined to be used in the close-loop simulations.
+Finally in Section \ref{SS:simulationResults} controllers performance are compared in terms of energy saving using a thermal energy model of the house built via EnergyPlus in Section \ref{SS:energyPlusmodel}, using both historical data, and house construction layout and materials.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[width=1\linewidth]{figures/overview.eps}
+		\caption{Section \ref{S:realCaseStudy} organization scheme.}
+		\captionsetup{justification=centering}
+		\label{F:overview}
+	\end{center}
+\end{figure}
 
@@
 The chosen case study is a detached off-grid two-story residential house, located in the outskirt of L'Aquila (coordinates $42\degree$ $16'$ latitude and $13\degree$ $32'$ longitude), Italy and is shown in Figure \ref{F:house}.
-The building, inhabited by the two owners, has a main north-south orientation and it is composed by a heated ground floor, that includes $4$ rooms, and an attic without heating system.
+The building, inhabited by the two owners, has a main north-south orientation and it is composed by a heated ground floor and an attic without heating system.
 Therefore, although the gross area of the house is equal to $209.5\,m^2$, the heated gross area is equal to $112.4\,m^2$.
@@
 Thanks to the off-grid characteristic, the technological plants guarantee the complete energy self-sufficiency of the building.
-Indeed the house, equipped with a biomass boiler, a solar thermal plant, a stand-alone photovoltaic system, black water and rainwater reuse systems and as well for water supply, has a complete independence from the utilities.
+Indeed the house, equipped with a biomass boiler, a solar thermal plant, a stand-alone photovoltaic system, black water and rainwater reuse systems and a well for water supply, has a complete independence from the utilities.
 The bearing structure of the house is made of reinforced concrete and EPS (expanded polystyrene) insulation, while the building envelope is composed by prefabricated wood-cement blocks, shown in Figure \ref{F:houseSection}, with EPS and graphite insulation, that allow low thermal fluxes thanks to their high thermal performance.
@@
 Temperature measuring devices, completely self-produced at the “G.
-Parolini Lab” of the University of L'Aquila \cite{Pantoli2017}, mainly based on an $ATmega 2560$ microcontroller and $DS18B20$ temperature probes (temperature range from $-55.0\degree C$ to $125.0\degree C$ with an accuracy of $\pm 0.5\degree C$), were employed to acquire the ambient temperatures of the four differently oriented rooms of the house as shown in Figure \ref{F:houseFloors}.
+Parolini Lab” of the University of L'Aquila \cite{Pantoli2017}, mainly based on an $ATmega 2560$ microcontroller and $DS18B20$ temperature probes (temperature range from $-55.0\degree C$ to $125.0\degree C$ with an accuracy of $\pm 0.5\degree C$), were employed to acquire the ambient temperatures of $4$ differently oriented rooms of the house as shown in Figure \ref{F:houseFloors} (displayed with red dots $A1$, $A2$, $A3$ and $A4$).
 The probes positions were chosen to optimize the data acquisition and to minimize the discomfort of the occupants in the house.
-Furthermore, as can be seen in Figures \ref{F:housePlantScheme} and \ref{F:houseFloors}, a commercial heat meter (temperature range from $10.0\degree C$ to $90.0\degree C$ with an accuracy of $\pm 0.05\degree C$) was installed downstream of the biomass boiler to measure the produced thermal energy.
+Furthermore, as can be seen in Figures \ref{F:housePlantScheme} and \ref{F:houseFloors}, a commercial heat meter (temperature range from $10.0\degree C$ to $90.0\degree C$ with an accuracy of $\pm 0.05\degree C$, displayed with orange dot $G1$) was installed downstream of the biomass boiler to measure the produced thermal energy.
 The turbine flowmeter was installed on the return pipe, while the two thermocouples were placed inside the delivery and return pipes, respectively.
@@
 
-The collected data are used in the following to create both an EnergyPlus and a random forest model for the energy consumption of the house.
+The collected data are used in the following to create both an EnergyPlus and a random forest model for the energy consumption assessment of the house.
 These models will be used for performance comparison of DPC with respect to a classical bang-bang controller.
@@
 To assess energy performance of the use case heating system, the EnergyPlus engine together with DesignBuilder modeling environment has been employed.
-The model has been created according to the K\"{o}ppen-Geiger climate classification that, based weather data of L'Aquila shown in Figure \ref{F:houseExternalWeather}, provided by the CETEMPS – Centre of Excellence \cite{CETEMPS}, classifies Italy in Csa, Cfa and Csb climate zones, with a warm temperate climate [4], .
+The model has been created based  on L'Aquila weather data, shown in Figure \ref{F:houseExternalWeather}, provided by the CETEMPS – Centre of Excellence \cite{CETEMPS}.
+According to the K\"{o}ppen-Geiger climate classification, Italy is classified in Csa, Cfa and Csb climate zones, with a warm temperate climate \cite{Peel2007}.
 
@@
 		\includegraphics[width=30pc]{figures/dati_climatici_rev01.eps}
-		\caption{Weather data of L'Aquila.}
+		\caption{Weather data of L'Aquila, Italy.}
 		\captionsetup{justification=centering}
@@
 \end{figure}
-\subsection{Random forest models} For the closed-loop simulations with DPC we create $2$ different sets of models, $S_1 = \{\tP^{r}_{\mathrm{k+1|k}},\tT^{r_1}_{\mathrm{k+1|k}},\tT^{r_2}_{\mathrm{k+1|k}},\tT^{r_3}_{\mathrm{k+1|k}},\tT^{r_4}_{\mathrm{k+j|k}}\}$ and $S_2 = \{\tP_{\mathrm{k+j|k}},\tT^{1}_{\mathrm{k+j|k}},\tT^{2}_{\mathrm{k+j|k}},\\\tT^{3}_{\mathrm{k+j|k}},\tT^{4}_{\mathrm{k+j|k}},\ j = 1,\ldots,N\}$, using random forests.
-In each set we have a model for the power consumption of the house ($\tP^{r}$ and $\tP$) and $4$ models that describe the evolution of the room temperature in each of the $4$ rooms ($\tT^{r_i}$ and $\tT^i$).
+\subsection{Random forest models}
+\label{SS:randomForestsModels}
+For the closed-loop simulations with DPC we create $2$ different sets of models, $S_1 = \{\tT^{r_1}_{\mathrm{k+1|k}},\tT^{r_2}_{\mathrm{k+1|k}},\tT^{r_3}_{\mathrm{k+1|k}},\tT^{r_4}_{\mathrm{k+j|k}}\}$ and $S_2 = \{\tP_{\mathrm{k+j|k}},\tT^{1}_{\mathrm{k+j|k}},\tT^{2}_{\mathrm{k+j|k}},\tT^{3}_{\mathrm{k+j|k}},\\\tT^{4}_{\mathrm{k+j|k}},\ j = 1,\ldots,N\}$, using random forests.
+In each set we have  $4$ models that describe the room temperature evolution ($\tT^{r_i}$ and $\tT^i$, $i=1,2,3,4$) in each of the $4$ rooms equipped with temperature sensors.
+In $S_2$ we have a model for the power consumption of the house ($\tP$).
 Models in $S_1$ are created using the classical random forests algorithm with all the features and are used as plant simulator of the house.
@@
 
-\subsection{DPC and bang-bang controllers} We set up $2$ different controllers, DPC and bang-bang controller, to obtain a scheduling policy to switch the radiators ON and OFF in order to keep the temperature of room $3$ within a comfort range.
+\subsection{DPC and bang-bang controllers}
+\label{SS:controllersDPCandBangbang}
+We set up $2$ different controllers, DPC and bang-bang controller, to obtain a scheduling policy to switch the radiators ON and OFF in order to keep the temperature of room $3$, i.e.
+the living room, within a comfort range.
 Since the heating system serves all of the $4$ rooms simultaneously, without giving the possibility to control the temperature of each room independently, we setup the problem defining the comfort range only for one room.
@@
 $15$ days of May, from May $1^{st}$ $2016$ to May $15^{th}$ $2016$.
-We can see that if we want to keep the temperature within a comfort range, only allowing small violations, the use of DPC produce an energy saving of $45\, kWh$, that correspond to the $25.4\%$ over a period of $15$ days.
+We can see that to the bang-bang control strategy is associated an energy consumption approximately equal to $177\,kWh$.
+If we want to keep the temperature within a comfort range, only allowing small violations, the use of DPC produce an energy saving of $45\, kWh$, that corresponds to the $25.4\%$ over a period of $15$ days.
 In case of medium violations get an energy saving of $57\, kWh$, that is $32.2\%$.
-Instead if we allow a large violation over the same period, the energy saving is of $87\, kWh$, that correspond to the $49.2\%$.
+Instead if we allow a large violation over the same period, the energy saving is of $87\, kWh$, that corresponds to the $49.2\%$.
 \begin{figure}[h!]

### 2017-09-29 paper/cs-comparison.tex
completed section 5 and fixed figures problems
Modified introduction. First complete draft.
Added flow scheme, motivation. Implemented comments from Italy.

@@
 
-The performance is compared for 3 days in winter, i.e. January 28-31 and 3 days in summer, i.e. May 1-3. These are shown on the same plots in Fig.~\ref{F:comparison}.
+The performance is compared for 3 days in winter, i.e. January 28-31 and 3 days in summer, i.e. May 1-3. These are shown on the same plots in Figure~\ref{F:comparison}.
 The sampling time in the simulations is 1 hr.
@@
 
-The external disturbances - solar gains, internal gain due to equipment and dry-bulb temperature during the chosen periods are shown in Fig.~\ref{F:dist}.
+The external disturbances - solar gain, internal gain due to equipment and dry-bulb temperature during the chosen periods are shown in Figure~\ref{F:dist}.
 The internal gain due to occupancy was proportional to the gain due to equipment.
 The reference temperature is chosen to be 22 $\mathrm{^oC}$.
-Due to cold weather, which is evident from the dry-bulb temperature, the heater is switched on during the night to maintain the thermal comfort requirements.
-When the building is occupied during the day, due to excessive internal gains, the building requires cooling.
+Due to cold weather, which is evident from the dry-bulb temperature, the heating system is switched on during the night to maintain the thermal comfort requirements.
+When the building is occupied during the day, due to excessive internal gain, the building requires cooling.
 The lighting in the building is adjusted to meet the minimum light requirements.
-The optimal cooling usage factor and the radiator power for MPC, DPC-En and DPC-RT are shown in Fig.~\ref{F:control1} and Fig.~\ref{F:control2}, respectively.
+The optimal cooling usage factor and the radiator power for MPC, DPC-En and DPC-RT are shown in Figure~\ref{F:control1} and Figure~\ref{F:control2}, respectively.
 The control strategy with DPC-En shows a remarkable similarity to MPC, switching on/off the equipments at the same time with similar usage.
@@
 Although it seems like that adding the rate constraints to DPC-En would smoothen its behavior, this was avoided because the sampling time of the system is 1 hr which is already too high.
-The room temperature profile in Fig.~\ref{F:state} is close to the reference in the case of DPC-En as well as MPC.
-Fig.~\ref{F:obj} shows that the cumulative cost of the objective function is, as expected, minimum for MPC, and a bit higher for DPC-En.
+The room temperature profile in Figure~\ref{F:state} is close to the reference in the case of DPC-En as well as MPC.
+Figure~\ref{F:obj} shows that the cumulative cost of the objective function is, as expected, minimum for MPC, and a bit higher for DPC-En.
 The cost for DPC-RT blows up around 12 noon on 30$^{th}$ January as one of the slack variables is non-zero, which happens due to high model inaccuracy.
 
-The quantitative performance comparison is shown in Tab.~\ref{T:comparison}.
+The quantitative performance comparison is shown in Table~\ref{T:comparison}.
 MPC tracks the reference more closely at the expense of higher input costs in comparison to DPC-En.
@@
 	\centering
-	\caption{Quantitative comparison of explained variance, mean value of objective function, mean input cost $c^Tu$ and mean deviance from the reference temperature $|\mathsf{T}-\mathsf{T}_{\mathrm{ref}}|$.}
-	\captionsetup{justification=centering}
-	\begin{tabular}{c|c|c|c|c}
+	\begin{tabular}{ccccc}
 		\toprule
@@
 	\end{tabular}
+	\caption{Quantitative comparison of explained variance, mean value of objective function, mean input cost $c^Tu$ and mean deviance from the reference temperature $|\mathsf{T}-\mathsf{T}_{\mathrm{ref}}|$.}
+	\captionsetup{justification=centering}
 	\label{T:comparison}
 \end{table}
-\begin{figure}[h!]
-	\subfigure[External disturances: solar gains, internal gain due to equipment and dry-bulb temperature.]{
+
+\begin{figure}[t!]
+	\begin{center}
+	\subfigure[External disturbances: solar gain, internal gain due to equipment and dry-bulb temperature.]{
 		\label{F:dist}
 		\centering
-		\includegraphics[width=20pc]{figures/disturbances.eps}
+		\includegraphics[width=16pc]{figures/disturbances.eps}
 	}
@@
 		\centering
-		\includegraphics[width=20pc]{figures/input3.eps}
+		\includegraphics[width=16pc]{figures/input3.eps}
 	}
@@
 		\label{F:control2}
-		\includegraphics[width=20pc]{figures/input4.eps}
+		\includegraphics[width=16pc]{figures/input4.eps}
 	}
@@
 		\centering
-		\includegraphics[width=20pc]{figures/state.eps}
+		\includegraphics[width=16pc]{figures/state.eps}
 	}
+\end{center}
+\end{figure}
+\begin{figure}[h!]
+\begin{center}
 	\subfigure[Cumulative optimal cost after solving optimization. MPC serves as the benchmark with the minimum cost, followed by DPC-En and then DPC-RT.]{
@@
 		\centering
-		\includegraphics[width=20pc]{figures/cumsumcost.eps}
+		\includegraphics[width=16pc]{figures/cumsumcost.eps}
 	}
+	\end{center}
 	%\vspace{-1cm}
@@
 \end{figure}
-
 Thus, we have shown that DPC-En provides a comparable performance to MPC without using the physical model.

### 2017-09-29 paper/cs-dpc.tex
completed section 5 and fixed figures problems
completed abstract, section 5 and conclusion
Added flow scheme, motivation. Implemented comments from Italy.

@@
 For example, states in the bilinear model also include slab temperatures which require modeling of structural and material properties in detail and often we also need to install new sensors to capture additional states.
-Thus, DPC is based solely on one state of the model i.e.
+Thus, DPC is based solely on one state of the model, i.e.
 the zone temperature that can be easily measured with a thermostat.
-This serves as the output variable $\tY$ of interest for which we build $N$ trees and $N$ forests as described in Sec.~\ref{SS:dpcrt} and \ref{SS:dpcrf}, respectively.
+This serves as the output variable $\tY$ of interest for which we build $N$ trees and $N$ forests as described in Section~\ref{SS:dpcrt} and \ref{SS:dpcrf}, respectively.
 Therefore, $\tY_{\mathrm{k+j|k}}:=x_{\mathrm{k+j|k}}^1$, where $x^1$ is the first component of $x$.
@@
 \begin{align}
-\text{min } & \sum_{j=1}^{N} {\tY}_{\mathrm{k+j|k}} \mathcal{Q}^{(1,1)} {\tY}_{\mathrm{k+j|k}} + c^T \tX^c_{\mathrm{k+j-1|k}}+  \lambda\epsilon_j\\
+\text{min } & \sum_{j=1}^{N} ({\tY}_{\mathrm{k+j|k}} - x_{ref}) \mathcal{Q}^{(1,1)} ({\tY}_{\mathrm{k+j|k}} - x_{ref}) + c^T \tX^c_{\mathrm{k+j-1|k}}+  \lambda\epsilon_j\\
 \text{s.~t. } & \ \ \ \ \ \tY_{\mathrm{k+j|k}} =  \alpha_j^T \left[1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} \right]^T \label{SE:dpc}\\
@@
 
-We compare the prediction for the first time step $\tY_{\mathrm{k+1|k}}$ and the 6-hour ahead prediction $\tY_{\mathrm{k+6|k}}$ for a week in the month of May in Fig.~\ref{F:validation}.
+We compare the prediction for the first time step $\tY_{\mathrm{k+1|k}}$ and the 6-hour ahead prediction $\tY_{\mathrm{k+6|k}}$ for a week in the month of May in Figure~\ref{F:validation}.
 It is visible how trees have a high variance, and the forests are more accurate.
-Note that data from January and May was not used for training.
-The quantitative summary of the accuracy is given in Tab.~\ref{T:validation}.
+Note that data from January and May were not used for training.
+The quantitative summary of the accuracy is given in Table~\ref{T:validation}.
 We can see that the random forests are better in all respects.
@@
 	\centering
-	\caption{Quantitative comparison of root mean square error (RMSE), $R^2$ score, and explained variance (EV) for trees and forests for different predictions steps.}
-	\captionsetup{justification=centering}
-	\begin{tabular}{c|c|c|c}
+	\begin{tabular}{cccc}
 		\toprule
@@
 	\end{tabular}
+	\caption{Quantitative comparison of root mean square error (RMSE), $R^2$ score, and explained variance (EV) for trees and forests for different predictions steps.}
+	\captionsetup{justification=centering}
 	\label{T:validation}

### 2017-09-29 paper/dpc.tex
Added flow scheme, motivation. Implemented comments from Italy.

@@
 We have validated this linear model assumption in \cite{Jain2016}.
-As we shall see in Sec.~\ref{SS:dpcrt} and \ref{SS:dpcrf}, the second step reduces the run-time control problem into a convex program.
-This process is illustrated in Fig.~\ref{F:dpc-sepvars}.
+As we shall see in Section~\ref{SS:dpcrt} and \ref{SS:dpcrf}, the second step reduces the run-time control problem into a convex program.
+This process is illustrated in Figure~\ref{F:dpc-sepvars}.
 \begin{figure}[t!]
@@
 \end{align}
-DPC-En is graphically described in Fig.~\ref{F:dpc-algo-rf}.
+DPC-En is graphically described in Figure~\ref{F:dpc-algo-rf}.
 The ensemble data predictive control (DPC-En) is the first such method to bridge the gap between ensemble predictive models (such as random forests) and receding horizon control.

### 2017-09-29 paper/intro.tex
awesome figure
Added flow scheme, motivation. Implemented comments from Italy.

@@
 
+\subsection{Motivation}
 Consider a multivariable dynamical system subject to external disturbances.
@@
 For example, in the case of buildings, this requires installation of thousands of new sensors, expert domain knowledge and several months to years of efforts depending upon the size and complexity of structure of the building.
-We present this case-study in detail in Sec.~\ref{S:casestudy}.
+We present this case-study in detail in Section~\ref{S:casestudy}.
 
-The question now is, can we employ data-driven techniques to reduce the cost of modeling and still exploit the benefits that MPC has to offer?
+\paragraph{Example} As an example of such modeling complexity, consider the grey-box approach in \cite{Braun2002}.
+The scope of such approach is to predict the heat transfer rate to the air within the building.
+In particular, the authors addressed a single zone modeling problem using an equivalent $RC$ model.
+To this aim, five different types of structures are considered in the example: external walls, ceiling/roof, floor, internal walls and windows.
+Each of these elements are represented using $3$ resistances, $2$ capacitances and $2$ temperatures, except for the windows that are represented only with resistances since they have a negligible energy storage.
+A model consisting of $8$ states and $9$ inputs, including disturbances such as solar radiation and others, with the instantaneous heat gain to the building air from all surfaces as output, is created for the zone.
+To have a complete LTI model, $9$ parameters needed to be estimated through a non-trivial training process, using a collection of information associated with a physical description of the building (see \cite{Braun2002} for more details).
+In Section \ref{S:realCaseStudy} we will consider an application of the methodology proposed in this paper on a real house with $10$ different zones.
+This means that, in order to have a model of the zones as the one just described, only for what concerns the heat transfer rate within the building, and hence neglecting the energy consumption model of the house, we would need to consider a model with around $80$ states and $90$ parameters to be estimated.
+
+\paragraph{Objective} The question now is, can we employ data-driven techniques to reduce the cost of modeling and still exploit the benefits that MPC has to offer?
 Can we build automatic and data-driven approaches for control purposes that are also adaptive, scalable and interpretable?
@@
 
+\subsection{Related work}
 In the literature, several studies deal with the data-driven control techniques using machine learning algorithms \cite{Hou2013}.
@@
 In our previous work \cite{Behl2016,Jain2016,JainACC2017,JainCDC2017}, we introduced the concept of DPC for receding horizon control.
-In this paper, we extend these results providing the following contributions:
+In this paper, we extend these results providing the following contributions.
 \begin{enumerate}
 	\item In Section \ref{S:dpc}, we formally present the two following control techniques:
-	\begin{itemize}
+	\begin{enumerate}
 		\item DPC with regression trees, and
 		\item DPC with random forests.
-	\end{itemize}
+	\end{enumerate}
 	\item In Section \ref{S:proof}, we demonstrate the strength of DPC for receding horizon control via one-to-one comparison against a benchmark MPC controller using a bilinear building model whose parameters were identified using experiments on a building in Switzerland.
@@
 We present a comprehensive case study to demonstrate the applicability of DPC to demand response, especially how DPC can provide financial incentives to the end-customers bypassing the need for high fidelity models.
-%In particular, we first describe two underlying algorithms: DPC with regression trees and DPC with random forests in Sec.~\ref{S:dpc}. 
-%In Sec.~\ref{S:proof}, we apply DPC to a bilinear building model, whose parameters were identified using experiments on a building in Switzerland, to show a one-to-one comparison of receding horizon control with DPC against a benchmark MPC controller.
-In Sec.~\ref{S:casestudy}, we present a practical application of DPC for Demand Response, where we apply DPC to a multi-story 19 zone building model in EnergyPlus for which model-based control is not economical and practical due to extreme complexity.
+%In particular, we first describe two underlying algorithms: DPC with regression trees and DPC with random forests in Section~\ref{S:dpc}. 
+%In Section~\ref{S:proof}, we apply DPC to a bilinear building model, whose parameters were identified using experiments on a building in Switzerland, to show a one-to-one comparison of receding horizon control with DPC against a benchmark MPC controller.
+In Section~\ref{S:casestudy}, we present a practical application of DPC for Demand Response, where we apply DPC to a multi-story 19 zone building model in EnergyPlus for which model-based control is not economical and practical due to extreme complexity.
 
@@
 
-%We begin with description of DPC-RT and DPC-En algorithms, and Gaussian Processes for DPC in Sec.~\ref{S:dpc}.
-Sec.~\ref{S:casestudy} presents a case study on DR using a bilinear building model.
+%We begin with description of DPC-RT and DPC-En algorithms, and Gaussian Processes for DPC in Section~\ref{S:dpc}.
+Section~\ref{S:casestudy} presents a case study on DR using a bilinear building model.
 Here, we also validate the black-box models used for control, and compare DPC against MPC.
-We discuss the challenges associated with DPC in Sec.~\ref{SS:challenges}.
-We conclude the paper with a summary of the results and a brief discussion on the future work in Sec.~\ref{S:conclusion}.
+We discuss the challenges associated with DPC in Section~\ref{SS:challenges}.
+We conclude the paper with a summary of the results and a brief discussion on the future work in Section~\ref{S:conclusion}.

### 2017-09-29 paper/main.tex
Modified introduction. First complete draft.
edited abstract+intro
Couple of small changes in abstract, intro and keywords
Added flow scheme, motivation. Implemented comments from Italy.

@@
 \cortext[mycorrespondingauthor]{Corresponding author}
-\ead{francesco.smarra@univaq.it@univaq.it}
+\ead{francesco.smarra@univaq.it}
 
@@
 \author[TDRaddress]{Dario Ambrosini}
-\ead{dario.ambrosini@ing.univaq.it}
+\ead{dario.ambrosini@univaq.it}
 \author[FSaddress]{Alessandro D'Innocenzo}
@@
 
+\fntext[EC]{Equal contribution}
+
 \address[FSaddress]{Department of Information Engineering, Computer Science and Mathematics, Universit\`{a} degli Studi dell'Aquila, L'Aquila, Italy}
 \address[AJaddress]{Department of Electrical and Systems Engineering, University of Pennsylvania, Philadelphia, USA}
-\address[TDRaddress]{Department of Industrial and Information Engineering and Economics, Universit\`{a} degli Studi dell'Aquila, L'Aquila, Italy}
-\address[EC]{Equal contribution}
+\address[TDRaddress]{"G. Parolini Lab" - Department of Industrial and Information Engineering and Economics, Universit\`{a} degli Studi dell'Aquila, L'Aquila, Italy}
 
@@
 \begin{keyword}
-Energy efficiency, demand response, predictive control, machine learning, EnergyPlus modeling, real case study
+Energy efficiency, demand response, predictive control, machine learning, EnergyPlus modeling, real application
 \end{keyword}

### 2017-09-29 paper/casestudyReal.tex
small edits

@@
 
-In this section we show an application of DPC to a real case study.
-The section organization is shown in Figure \ref{F:overview}.
-More precisely, we consider a real house located in L'Aquila, Italy, as a case study that we describe in Section \ref{SS:descriptionHouse}.
-Using historical data from this house, random forest models are built in Section \ref{SS:randomForestsModels} to be used either as plant or as predictive models for DPC.
-In Section \ref{SS:controllersDPCandBangbang}, DPC and bang-bang control strategies are defined to be used in the close-loop simulations.
-Finally in Section \ref{SS:simulationResults} controllers performance are compared in terms of energy saving using a thermal energy model of the house built via EnergyPlus in Section \ref{SS:energyPlusmodel}, using both historical data, and house construction layout and materials.
+In this section, we show an application of DPC to a real house located in L'Aquila, Italy.
+Random forest models for DPC built using historical data from this house are described in Section \ref{SS:randomForestsModels}.
+In Section \ref{SS:energyPlusmodel}, we present the thermal energy model of the building in EnergyPlus, built using historical data, construction layout and materials after spending $\sim3$ months of efforts.
+In Section \ref{SS:controllersDPCandBangbang}, we set up the DPC optimization problem and describe the bang-bang control strategy.
+Finally, in Section \ref{SS:simulationResults}, the performance of the two controllers are compared in terms of energy saving using the EnergyPlus model.
+The section organization is graphically shown in Figure \ref{F:overview}.
 
@@
 		\includegraphics[width=1\linewidth]{figures/overview.eps}
-		\caption{Section \ref{S:realCaseStudy} organization scheme.}
+		\caption{Random Forest trained on all features including the control variable are used as the plant in the closed-loop simulations with DPC and bang-bang controller.
+DPC optimzation uses Random Forests with affine functions as predictive models.
+Controllers' strategies are evaluated using the EnergyPlus model to compare the actual energy consumption.}
 		\captionsetup{justification=centering}

### 2017-09-29 paper/intro.tex
small edits

@@
 In particular, the authors addressed a single zone modeling problem using an equivalent $RC$ model.
-To this aim, five different types of structures are considered in the example: external walls, ceiling/roof, floor, internal walls and windows.
+To this end, five different types of structures are considered in the example: external walls, ceiling/roof, floor, internal walls and windows.
 Each of these elements are represented using $3$ resistances, $2$ capacitances and $2$ temperatures, except for the windows that are represented only with resistances since they have a negligible energy storage.
 A model consisting of $8$ states and $9$ inputs, including disturbances such as solar radiation and others, with the instantaneous heat gain to the building air from all surfaces as output, is created for the zone.
-To have a complete LTI model, $9$ parameters needed to be estimated through a non-trivial training process, using a collection of information associated with a physical description of the building (see \cite{Braun2002} for more details).
-In Section \ref{S:realCaseStudy} we will consider an application of the methodology proposed in this paper on a real house with $10$ different zones.
-This means that, in order to have a model of the zones as the one just described, only for what concerns the heat transfer rate within the building, and hence neglecting the energy consumption model of the house, we would need to consider a model with around $80$ states and $90$ parameters to be estimated.
+To have a complete LTI model, $9$ parameters are needed to be estimated via a non-trivial training process, using a collection of information associated with a physical description of the building (see \cite{Braun2002} for more details).
+Extending this result, for the building we consider in Section \ref{S:realCaseStudy} with $10$ different zones, we would need to consider a model with approximately $80$ states and $90$ parameters to be estimated.
 

### 2017-09-29 paper/casestudyReal.tex
small changes

@@
 Each set allows a different level of temperature bounds violation.
-In particular we considered a small, a medium and a large violation configuration.
+In particular we considered a small ($Q=100$, $\lambda_{min}=3000$ and $\lambda_{max}=100$), a medium ($Q=100$, $\lambda_{min}=1000$ and $\lambda_{max}=100$) and a large ($Q=100$, $\lambda_{min}=100$ and $\lambda_{max}=100$) violation configuration.
 \paragraph{Result 1} For the sake of clarity, we report in Figure \ref{F:comparison_small} only the comparison for temperature and input schedule obtained allowing small bound violations in DPC.
 For the same reason the simulation period is restricted to $4$ days and a half, from May $1^{st}$ $2016$ at $00am$ to May $5^{th}$ $2016$ at $1pm$, in order to make the plots more clear.
-We can see that the temperature controlled with DPC does not violate the bounds and if it does then the violation is around $0.1\,\degree C$.
+We can see that the temperature controlled with DPC does not violate the bounds and if it does then the violation is approximately $0.1\,\degree C$.
 Bang-bang control also presents small bounds violations due to its working principle.
@@
 In Figure \ref{F:comparison_all_energy_E+}, using the thermal energy consumption model derived in Section \ref{SS:energyPlusmodel} using EnergyPlus, we show how DPC outperforms the bang-bang controller also in terms of energy consumption and how the bounds violations allow us to save more energy.
-In this case, since the plot is clear and we are interested in showing energy saving on a long period, we ran the simulations over the testing data, i.e.
+In this case, since the plot is clear and we are interested in showing energy saving on a long period, we ran the simulations over the whole testing period, i.e.
 $15$ days of May, from May $1^{st}$ $2016$ to May $15^{th}$ $2016$.

### 2017-09-29 paper/cs-model.tex
Added flow scheme, motivation. Implemented comments from Italy.
small changes

@@
 
-The model dynamics is given below. The bilinearity is present in both input-state, and input-disturbance.
+The model dynamics are given below. The bilinearity is present in both input-state, and input-disturbance.
 \begin{gather}
@@
 In our future work, we will account for the uncertainties in the disturbances with an extension to Scenario approach \cite{Bernardini2009} for DPC.
-
-\begin{figure}[t!]
-	\centering
-	\includegraphics[width=21pc]{figures/dpc-algo-rf.eps}
-	\caption{DPC-En: At time $k$, the algorithm uses the forecast of disturbances $\tX^d_{\mathrm{k|k}}$ to select linear models $\Theta_1$ to $\Theta_t$ in the leaves of each ensemble.
-The linear models in each ensemble are averaged to calculate a single model represented by $\hat{\Theta}_j$ which act as constraints in the optimization problem.
-The optimal sequence $[\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+N-1|k}}]$, of which the first one is applied, and $\tX^d_{\mathrm{k+1|k+1}}$ is calculated to proceed to $k+1$.}
-	\label{F:dpc-algo-rf}
-\end{figure}

### 2017-09-29 paper/dpc.tex
small changes

@@
 We solve this optimization in the same manner as MPC to determine the optimal sequence of inputs $[\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+N-1|k}}]$, apply the first control input $\tX^c_{\mathrm{k|k}}$ and proceed to the next time step $k+1$.
-The pseudo code for DPC-RT is given in Alg.~\ref{A:dpcrt}.
-
-\subsection{DPC-En: DPC with Ensemble Methods}
-\label{SS:dpcrf}
-Regression trees obtain good predictive accuracy in many domains. However, the models used in their leaves have some limitations regarding the kind of functions they are able to approximate.
-The problem with trees is their high variance and that they can overfit the data easily.
-A small change $\delta$ in the data can result in a different series of splits and thus violate the acceptable accuracy $\eps$.
-This is the price to be paid for estimating a tree-based structure from the data.
-%The main reason behind this is the hierarchical nature of the process: the effect of an error in the top split is propagated down to all of the splits below it.
-%While pruning and cross validation can help reduce the over fitting partially, we use ensemble methods for growing more accurate trees.
-
-We use ensemble methods \cite{Friedman2001} to combine the predictions of several independent regression trees in order to improve generalizability and robustness over a single estimator. 
-%Random forests or \textit{tree-bagging} are a type of ensemble method which makes predictions by averaging over the predictions of several independent trees.
-The essential idea is to average many noisy trees to reduce the overall variance in prediction.
-We inject randomness into the tree construction in two ways. First, we randomize the features used to define splitting in each tree.
-Second, we build each tree using a bootstrapped or sub-sampled data set.
-In this way, each tree in the forest is trained on different data, which introduces differences between the trees.
-More explicitly, training features $\tX^d \in \R^p$ with $p<d$ and the in-bag samples (in-bag samples correspond to the data samples on which the tree was trained) are different for each tree in the forest i.e $|(\tX,\tY)|<n$.
-
-The goal with DPC-En is to replace each tree in Alg.~\ref{A:dpcrt} by a forest
-\begin{gather}
-\label{E:model_forest}
-\tY_{\mathrm{k+j|k}} = \mathit{f}_{\mathrm{forest}} \left( \tX^d_{\mathrm{k+j-N|k}} ,\dots,\tX^d_{\mathrm{k+j-1|k}}  \right), \\ \nonumber
-\mathsf{X}^d_{\mathrm{k+j-l|k}} \in \R^p \  \ \forall \ l,j=1,\dots,N,
-\end{gather}
-which, again, is trained only on $\mathsf{X}^d$, with $\tX^d \in \R^p \subset \R^d$ for each tree,  and then fit a linear regression model using $\mathsf{X}^c$ in every leaf of every tree.
-We build $N$ such forests for $N$ prediction steps such that the leaf $R_i$ of forest $\mathcal{R}_j$ uses a linear model
-\begin{gather}
-\label{E:model_leaf_forest}
-\tY_{\mathrm{k+j|k}} =  \Theta^T_{ij} [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T, \\ \nonumber
-\mathsf{X}^c_{\mathrm{k+j-l|k}} \in \R^c \  \ \forall \ l,j=1,\dots,N.
-\end{gather}
-Here $(\mathsf{X}^c,\mathsf{Y})$ correspond to the in-bag samples for the trees.
+The pseudo code for DPC-RT is given in Algorithm~\ref{A:dpcrt}.
 
@@
 
+\subsection{DPC-En: DPC with Ensemble Methods}
+\label{SS:dpcrf}
+Regression trees obtain good predictive accuracy in many domains. However, the models used in their leaves have some limitations regarding the kind of functions they are able to approximate.
+The problem with trees is their high variance and that they can overfit the data easily.
+A small change $\delta$ in the data can result in a different series of splits and thus violate the acceptable accuracy $\eps$.
+This is the price to be paid for estimating a tree-based structure from the data.
+%The main reason behind this is the hierarchical nature of the process: the effect of an error in the top split is propagated down to all of the splits below it.
+%While pruning and cross validation can help reduce the over fitting partially, we use ensemble methods for growing more accurate trees.
+
+We use ensemble methods \cite{Friedman2001} to combine the predictions of several independent regression trees in order to improve generalizability and robustness over a single estimator. 
+%Random forests or \textit{tree-bagging} are a type of ensemble method which makes predictions by averaging over the predictions of several independent trees.
+The essential idea is to average many noisy trees to reduce the overall variance in prediction.
+We inject randomness into the tree construction in two ways. First, we randomize the features used to define splitting in each tree.
+Second, we build each tree using a bootstrapped or sub-sampled data set.
+In this way, each tree in the forest is trained on different data, which introduces differences between the trees.
+More explicitly, training features $\tX^d \in \R^p$ with $p<d$ and the in-bag samples (in-bag samples correspond to the data samples on which the tree was trained) are different for each tree in the forest i.e $|(\tX,\tY)|<n$.
+
+The goal with DPC-En is to replace each tree in Algorithm~\ref{A:dpcrt} by a forest
+\begin{gather}
+\label{E:model_forest}
+\tY_{\mathrm{k+j|k}} = \mathit{f}_{\mathrm{forest}} \left( \tX^d_{\mathrm{k+j-N|k}} ,\dots,\tX^d_{\mathrm{k+j-1|k}}  \right), \\ \nonumber
+\mathsf{X}^d_{\mathrm{k+j-l|k}} \in \R^p \  \ \forall \ l,j=1,\dots,N,
+\end{gather}
+which, again, is trained only on $\mathsf{X}^d$, with $\tX^d \in \R^p \subset \R^d$ for each tree,  and then fit a linear regression model using $\mathsf{X}^c$ in every leaf of every tree.
+We build $N$ such forests for $N$ prediction steps such that the leaf $R_i$ of forest $\mathcal{R}_j$ uses a linear model
+\begin{gather}
+\label{E:model_leaf_forest}
+\tY_{\mathrm{k+j|k}} =  \Theta^T_{ij} [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T, \\ \nonumber
+\mathsf{X}^c_{\mathrm{k+j-l|k}} \in \R^c \  \ \forall \ l,j=1,\dots,N.
+\end{gather}
+Here $(\mathsf{X}^c,\mathsf{Y})$ correspond to the in-bag samples for the trees.
+
 While the offline training burden in DPC-En is slightly increased compared to DPC-RT, in the control step we exploit the better accuracy, and lower variance properties of the random forest. 
@@
 In the next section, we compare DPC-RT and DPC-En to MPC for a building model.
+
+\begin{figure}[t!]
+	\centering
+	\includegraphics[width=21pc]{figures/dpc-algo-rf.eps}
+	\caption{DPC-En: At time $k$, the algorithm uses the forecast of disturbances $\tX^d_{\mathrm{k|k}}$ to select linear models $\Theta_1$ to $\Theta_t$ in the leaves of each ensemble.
+The linear models in each ensemble are averaged to calculate a single model represented by $\hat{\Theta}_j$ which act as constraints in the optimization problem.
+The optimal sequence $[\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+N-1|k}}]$, of which the first one is applied, and $\tX^d_{\mathrm{k+1|k+1}}$ is calculated to proceed to $k+1$.}
+	\label{F:dpc-algo-rf}
+\end{figure}

### 2017-09-29 paper/casestudyReal.tex
small changes

@@
 		\toprule
-		\multirow{3}{1.4cm}{\centering Structural member} & \multirow{3}{3.7cm}{\centering Layer description (from inside to outside)}     & Thermal     		    & Total     & Total        \\
+		\multirow{3}{1.4cm}{\centering Structural member} & \multirow{3}{3.7cm}{\centering Layer 
+															description (from inside to outside)}       & Thermal     		     & Total     & Total        \\
 		                                                  &                                             & resistance   		     & thickness & U-value      \\
 		                                                  &                                             & $[(m^2K)/W]$		     & $[m]$     & $[W/(m^2K)]$ \\
+
 		\midrule
@@
 														  &							                    &						 &		     &			    \\
-		\multirow{3}*{Floor}        					  & Screed                                                                         & $0.027$     		    &           &              \\ 
+
+		\multirow{3}*{Floor}        					  & Wood-cement                                 & $0.308$     		     &           &              \\ 
 							        					  & Polystyrene                                 & $6.000$     		     & $0.23 $   & $0.28 $      \\ 
-							        					  & Wood-cement                                                                    & $0.308$     		    &           &              \\
+							        					  & Screed                                      & $0.027$     		     &           &              \\
 														  &									    	    &					 	 &		     &			    \\
-		\multirow{5}{1.4cm}{\centering Pitched roof} 	  & \multirow{2}{3.7cm}{\centering Polyurethane resins and polyisocianurate foams} & \multirow{2}*{$4.000$} &           &              \\ 
+
+		\multirow{5}{1.4cm}{\centering Pitched roof} 	  & Wood-cement                                 & $0.308$                &           &              \\ 
+														  & Polystyrene                                 & $6.000$                &           &              \\
+							        					  & Screed                                      & $0.027$                & $0.31 $   & $0.13 $      \\ 
+							        					  & \multirow{2}{3.7cm}{\centering Polyurethane 
+							        					  	resins and polyisocianurate foams}          & \multirow{2}*{$4.000$} &           &              \\
 						        					  	  &                                             &                        &           &              \\
-														  & Screed                                                                         & $0.027$    		    & $0.31 $   &  $0.13 $     \\
-							        					  & Polystyrene                                                                    & $6.000$    		    &           &              \\ 
-							        					  & Wood-cement                                                                    & $0.308 $   		    &           &              \\
 		\bottomrule

### 2017-09-30 paper/abstract.tex
Couple of small changes in abstract, intro and keywords
Added flow scheme, motivation. Implemented comments from Italy.
some changes

@@
 A key factor prohibiting the widespread adoption of MPC for complex systems is related to the difficulties (cost, time, and effort) associated with the model identification.
-Data-driven models have been intensively studied in the last year to overcome this problem.
-However, closed-loop control systems based on such models have not received much attention.
+Due to their lack of dynamical equations to predict real systems behavior, data-driven models are not suitable for control.
 We introduce a novel idea for predictive control based on system data, Data Predictive Control (DPC), leveraging machine learning algorithms, in particular regression trees and random forests.
@@
 Finally, we use DPC to optimize ON/OFF scheduling of the heating system of a real house in order to minimize the power consumption while guaranteeing thermal comfort for occupants.
-Our results show that with DPC we obtain an energy saving in terms of $kWh$ that goes from $25.4\%$ to $49.2\%$ when compared to a bang-bang controller.
+Our results show that with DPC we obtain an energy saving in terms of primary energy up to $49.2\%$ when compared to a bang-bang controller.

### 2017-09-30 paper/casestudyReal.tex
some changes

@@
 
-\begin{figure}[h!]
+\subsection{Description of the house}\label{SS:descriptionHouse}
+The chosen case study is a detached off-grid two-story residential house, located in the outskirt of L'Aquila (coordinates $42\degree$ $16'$ latitude and $13\degree$ $32'$ longitude), Italy and is shown in Figure \ref{F:house}.
+The building, inhabited by the two owners, has a main north-south orientation and it is composed by a heated ground floor and an attic without heating system.
+Therefore, although the gross area of the house is equal to $209.5\,m^2$, the heated gross area is equal to $112.4\,m^2$.
+\begin{figure}[t!]
 	\begin{center}
@@
 \end{figure}
-
-\subsection{Description of the house}\label{SS:descriptionHouse}
-The chosen case study is a detached off-grid two-story residential house, located in the outskirt of L'Aquila (coordinates $42\degree$ $16'$ latitude and $13\degree$ $32'$ longitude), Italy and is shown in Figure \ref{F:house}.
-The building, inhabited by the two owners, has a main north-south orientation and it is composed by a heated ground floor and an attic without heating system.
-Therefore, although the gross area of the house is equal to $209.5\,m^2$, the heated gross area is equal to $112.4\,m^2$.
-
-\begin{figure}[h!]
+\begin{figure}[t!]
 	\begin{center}
@@
 \end{figure}
-
 Thanks to the off-grid characteristic, the technological plants guarantee the complete energy self-sufficiency of the building.
@@
 The thermal performance of a wood-cement block, with similar geometry, was investigated in a previous work \cite{Nardi2016}, both with experimental and numerical approach.
-
-\begin{figure}[h!]
+\begin{figure}[t!]
 	\begin{center}
@@
 
-\begin{figure}[h!]
+For a thorough knowledge of the building thermal behavior, an in-situ analysis was carried out.
+Temperature measuring devices, completely self-produced at the “G.
+Parolini Lab” of the University of L'Aquila \cite{Pantoli2017}, mainly based on an $ATmega 2560$ microcontroller and $DS18B20$ temperature probes (temperature range from $-55.0\degree C$ to $125.0\degree C$ with an accuracy of $\pm 0.5\degree C$), were employed to acquire the ambient temperatures of $4$ differently oriented rooms of the house as shown in Figure \ref{F:houseFloors} (displayed with red dots $A1$, $A2$, $A3$ and $A4$).
+The probes positions were chosen to optimize the data acquisition and to minimize the discomfort of the occupants in the house.
+Furthermore, as can be seen in Figures \ref{F:housePlantScheme} and \ref{F:houseFloors}, a commercial heat meter (temperature range from $10.0\degree C$ to $90.0\degree C$ with an accuracy of $\pm 0.05\degree C$, displayed with orange dot $G1$) was installed downstream of the biomass boiler to measure the produced thermal energy.
+The turbine flowmeter was installed on the return pipe, while the two thermocouples were placed inside the delivery and return pipes, respectively.
+All the measuring devices have been set with a data acquisition rate equal to $10$ minutes, from March $11^{th}$ $2016$ to May $15^{th}$ $2016$.
+
+The collected data are used in the following to create both an EnergyPlus and a random forest model for the energy consumption assessment of the house.
+These models will be used for performance comparison of DPC with respect to a classical bang-bang controller.
+In particular DPC will be set up to provide an optimal ON/OFF scheduling policy for the heating system in order to save energy while guaranteeing thermal comfort for the occupants.
+To this aim, we also create random forest models for power consumption and room temperatures to be used in the closed-loop simulations.
+
+\begin{figure}[t!]
 	\begin{center}
@@
 \end{figure}
-For a thorough knowledge of the building thermal behavior, an in-situ analysis was carried out.
-Temperature measuring devices, completely self-produced at the “G.
-Parolini Lab” of the University of L'Aquila \cite{Pantoli2017}, mainly based on an $ATmega 2560$ microcontroller and $DS18B20$ temperature probes (temperature range from $-55.0\degree C$ to $125.0\degree C$ with an accuracy of $\pm 0.5\degree C$), were employed to acquire the ambient temperatures of $4$ differently oriented rooms of the house as shown in Figure \ref{F:houseFloors} (displayed with red dots $A1$, $A2$, $A3$ and $A4$).
-The probes positions were chosen to optimize the data acquisition and to minimize the discomfort of the occupants in the house.
-Furthermore, as can be seen in Figures \ref{F:housePlantScheme} and \ref{F:houseFloors}, a commercial heat meter (temperature range from $10.0\degree C$ to $90.0\degree C$ with an accuracy of $\pm 0.05\degree C$, displayed with orange dot $G1$) was installed downstream of the biomass boiler to measure the produced thermal energy.
-The turbine flowmeter was installed on the return pipe, while the two thermocouples were placed inside the delivery and return pipes, respectively.
-All the measuring devices have been set with a data acquisition rate equal to $10$ minutes, from March $11^{th}$ $2016$ to May $15^{th}$ $2016$.
-\begin{figure}[h!]
+\begin{figure}[t!]
 	\begin{center}
@@
 \end{figure}
-
-The collected data are used in the following to create both an EnergyPlus and a random forest model for the energy consumption assessment of the house.
-These models will be used for performance comparison of DPC with respect to a classical bang-bang controller.
-In particular DPC will be set up to provide an optimal ON/OFF scheduling policy for the heating system in order to save energy while guaranteeing thermal comfort for the occupants.
-To this aim, we also create random forest models for power consumption and room temperatures to be used in the closed-loop simulations.
-
 \subsection{EnergyPlus model}\label{SS:energyPlusmodel}
@@
 
-To assess energy performance of the use case heating system, the EnergyPlus engine together with DesignBuilder modeling environment has been employed.
+To assess energy performance of the use case heating system, the EnergyPlus engine \cite{energyPlus} together with DesignBuilder modeling environment \cite{designBuilder} has been employed.
 The model has been created based  on L'Aquila weather data, shown in Figure \ref{F:houseExternalWeather}, provided by the CETEMPS – Centre of Excellence \cite{CETEMPS}.
@@
 
-\begin{figure}[h!]
+Because of the complex morphology of the blocks shown in Figure \ref{F:houseSection}, some simplified assumptions were made, in order to create the EnergyPlus virtual model.
+Considering the thermal properties of the wood-cement blocks that compose the walls shown in Figure \ref{F:houseSection1}, the thermal effects of the ribs were neglected.
+For the block employed for floor and roof shown in Figure \ref{F:houseSection2}, a less complex equivalent block, with only three layers, was evaluated in order to consider only one equivalent thermal transmittance value.
+The properties of the blocks used for the simulation model are listed in Table \ref{T:houseProperties}.
+\begin{figure}[t!]
 	\begin{center}
@@
 \end{figure}
-
-Because of the complex morphology of the blocks shown in Figure \ref{F:houseSection}, some simplified assumptions were made, in order to create the EnergyPlus virtual model.
-Considering the thermal properties of the wood-cement blocks that compose the walls shown in Figure \ref{F:houseSection1}, the thermal effects of the ribs were neglected.
-For the block employed for floor and roof shown in Figure \ref{F:houseSection2}, a less complex equivalent block, with only three layers, was evaluated in order to consider only one equivalent thermal transmittance value.
-The properties of the blocks used for the simulation model are listed in Table \ref{T:houseProperties}.
-
-\begin{table}[h!]
+\begin{table}[t!]
 	\centering
@@
 
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=20pc]{figures/Render_rev01.eps}
-		\caption{Virtual model of the use case.}
-		\captionsetup{justification=centering}
-		\label{F:houseVirtualModel}
-	\end{center}
-\end{figure}
-
 In order to calibrate the EnergyPlus model, a comparison between simulated and measured thermal energy consumption was performed.
@@
 The considered period goes from March $15^{th}$ $2016$ to April $15^{th}$ $2016$.
-
-\begin{figure}[h!]
+\begin{figure}[t!]
 	\begin{center}
-		\includegraphics[width=28pc]{figures/Scheduling_rev02.eps}
-		\caption{Actual scheduling of the heating system, where white color means switched OFF and green color switched ON, and ($T_{ext,ave}$) is the external average temperature.
-From midnight to $8am$ the heating system is always switched off.}
+		\includegraphics[width=20pc]{figures/Render_rev01.eps}
+		\caption{Virtual model of the use case.}
 		\captionsetup{justification=centering}
-		\label{F:houseActualScheduling}
+		\label{F:houseVirtualModel}
 	\end{center}
@@
 Therefore, with a MBE equal to $7.38\%$ and a CV(RMSE) equal to $8.37\%$, the EnergyPlus model of the use case can be considered well calibrated.
-
-\begin{figure}[h!]
+\begin{figure}[t!]
+	\begin{center}
+		\includegraphics[width=28pc]{figures/Scheduling_rev02.eps}
+		\caption{Actual scheduling of the heating system, where white color means switched OFF and green color switched ON, and ($T_{ext,ave}$) is the external average temperature.
+From midnight to $8am$ the heating system is always switched off.}
+		\captionsetup{justification=centering}
+		\label{F:houseActualScheduling}
+	\end{center}
+\end{figure}
+\begin{figure}[t!]
 	\begin{center}
@@
 
-\begin{table}[h!]
+\begin{table}[t!]
 	\centering
@@
 
-\begin{figure}[h!]
+\begin{figure}[t!]
 \begin{center}
+	\subfigure[Power consumption model accuracy validation.]{
+		\label{F:power_testing}
+		\centering
 		\includegraphics[width=26pc]{figures/power_testing.eps}
-	\caption{Power consumption model accuracy validation.}
+	}
+	\subfigure[Temperature model accuracy validation.]{
+		\label{F:temperature_testing}
+		\centering
+		\includegraphics[width=26pc]{figures/temperature_testing.eps}
+	}	
+	\end{center}
+\caption{Model accuracy validation.}
 \captionsetup{justification=centering}
-	\label{F:power_testing}
-\end{center}
-\end{figure}
-
-\begin{figure}[h!]
-\begin{center}
-	\includegraphics[width=26pc]{figures/temperature_testing.eps}
-	\caption{Temperature model accuracy validation.}
-	\captionsetup{justification=centering}
-	\label{F:temperature_testing}
-\end{center}
+\label{F:testing}
 \end{figure}
@@
 
-\begin{figure}[h!]
+\begin{figure}[t!]
 	\begin{center}
@@
 However the maximum violation is still lower than $1\,\degree C$.
-\begin{figure}[h!]
+\begin{figure}[t!]
 	\begin{center}
@@
 In Table \ref{T:violationErrors}, MBE and CV(RMSE) violation errors are reported to quantify the bounds violation of DPC, in each of the $3$ configurations, and of bang-bang controller.
-\begin{table}[h!]
-	\centering
-%	\scalebox{0.9}{
-		\begin{tabular}{cccccc}
-			\toprule
-			CONTROLLER  & LBV $\mathrm{MBE}$  & LBV $\mathrm{RMSE}$ & UBV $\mathrm{MBE}$ & UBV $\mathrm{RMSE}$ 	\\ 
-			\midrule
-			$DPC-SV$    & $0.013$             & $0.028$  			      & $0$    				 & $0$     	  	\\
-			$DPC-MV$    & $0.165$ 			  & $0.146$     			  & $0$    				 & $0$		  	\\
-			$DPC-LV$    & $0.410$  			  & $0.224$     			  & $0$    				 & $0$	      	\\
-			$Bang-bang$ & $0.049$ 			  & $0.080$    				  & $0.0063$ 		     & $0.0120$	  	\\
-			\bottomrule
-		\end{tabular}
-%	}
-	\caption{Lower Bound Violation (LBV) and Upper Bound Violation (UBV) errors expressed as $\mathrm{MBE}\%$ and $\mathrm{CV(RMSE)}\%$ for DPC Small Violation (DPC-SV), DPC Medium Violation (DPC-MV), DPC large Violation (DPC-LV) and bang-bang controller.}
-	\captionsetup{justification=centering}
-	\label{T:violationErrors}
-\end{table}
 We can see that if we allow very small violations, DPC outperforms bang-bang controller in terms of comfort guarantees.
@@
 Instead if we allow a large violation over the same period, the energy saving is of $87\, kWh$, that corresponds to the $49.2\%$.
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=26pc]{figures/Energy_all_EnergyPlus.eps}
-	\end{center}
-	\caption{Comparison of DPC and bang-bang controller performance in terms of thermal energy saving with different violations using EnergyPlus model.}
-	\label{F:comparison_all_energy_E+}
-\end{figure}
+\begin{table}[t!]
+	\centering
+	%	\scalebox{0.9}{
+	\begin{tabular}{cccccc}
+		\toprule
+		CONTROLLER  & LBV $\mathrm{MBE}$  & LBV $\mathrm{RMSE}$ & UBV $\mathrm{MBE}$ & UBV $\mathrm{RMSE}$ 	\\ 
+		\midrule
+		$DPC-SV$    & $0.013$             & $0.028$  			      & $0$    				 & $0$     	  	\\
+		$DPC-MV$    & $0.165$ 			  & $0.146$     			  & $0$    				 & $0$		  	\\
+		$DPC-LV$    & $0.410$  			  & $0.224$     			  & $0$    				 & $0$	      	\\
+		$Bang-bang$ & $0.049$ 			  & $0.080$    				  & $0.0063$ 		     & $0.0120$	  	\\
+		\bottomrule
+	\end{tabular}
+	%	}
+	\caption{Lower Bound Violation (LBV) and Upper Bound Violation (UBV) errors expressed as $\mathrm{MBE}\%$ and $\mathrm{CV(RMSE)}\%$ for DPC Small Violation (DPC-SV), DPC Medium Violation (DPC-MV), DPC large Violation (DPC-LV) and bang-bang controller.}
+	\captionsetup{justification=centering}
+	\label{T:violationErrors}
+\end{table}

### 2017-09-30 paper/conclusion.tex
some changes

@@
 We compare the performance of our Data Predictive Control to MPC on a multivariable bilinear building model.
-We establish that DPC with random forests shows a remarkable similarity to MPC in the optimal control strategies explaining 70\% variance.
+We establish that DPC with random forests shows a remarkable similarity to MPC in the optimal control strategies explaining $70\%$ variance.
 On the other hand, DPC with regression trees suffers from practical limitations due to model overfitting.
@@
 We show that DPC, relying only on the sensor data, can provide significant energy savings while maintaining thermal comfort.
-Our results demonstrate that even for such complex system, DPC tracks a reference signal with a mean error of 3\%.
+Our results demonstrate that even for such complex system, DPC tracks a reference signal with a mean error of $3\%$.
+
+\begin{figure}[t!]
+	\begin{center}
+		\includegraphics[width=26pc]{figures/Energy_all_EnergyPlus.eps}
+	\end{center}
+	\caption{Comparison of DPC and bang-bang controller performance in terms of thermal energy saving with different violations using EnergyPlus model.}
+	\label{F:comparison_all_energy_E+}
+\end{figure}
 

### 2017-09-30 paper/abstract.tex
final edits from achin

@@
 Model-based control techniques have been widely used over the past years in control systems to improve system performance of simple controllers, such as PID, bang-bang, etc.
-Among these, one of the most popular is Model Predictive Control (MPC), due to its capacity to optimally control a system, based on the capability to predict its behavior over a future horizon.
+Among these, one of the most popular is Model Predictive Control (MPC) due to its ability to predict the system's behavior over a future horizon and handle constraints in the optimal control setup.
 A key factor prohibiting the widespread adoption of MPC for complex systems is related to the difficulties (cost, time, and effort) associated with the model identification.
-Due to their lack of dynamical equations to predict real systems behavior, data-driven models are not suitable for control.
+On the other hand, data-driven models are not suitable for control because of non-physical models.
+%Due to their lack of dynamical equations to predict real systems behavior, data-driven models are not suitable for control. 
 We introduce a novel idea for predictive control based on system data, Data Predictive Control (DPC), leveraging machine learning algorithms, in particular regression trees and random forests.
-We show, using a bilinear model, that DPC has comparable performance with MPC.
-We further apply DPC to the Demand Response problem of a large scale multi-story EnergyPlus building model, and show that DPC curtails the desired power usage with high confidence.
+Using a bilinear model, we show that DPC has comparable performance with MPC.
+We further apply DPC to the Demand Response problem of a large scale multi-story EnergyPlus building building model, and show that DPC curtails the desired power usage with high confidence.
 Finally, we use DPC to optimize ON/OFF scheduling of the heating system of a real house in order to minimize the power consumption while guaranteeing thermal comfort for occupants.

### 2017-09-30 paper/casestudyReal.tex
final edits from achin

@@
 Thanks to the off-grid characteristic, the technological plants guarantee the complete energy self-sufficiency of the building.
-Indeed the house, equipped with a biomass boiler, a solar thermal plant, a stand-alone photovoltaic system, black water and rainwater reuse systems and a well for water supply, has a complete independence from the utilities.
-The bearing structure of the house is made of reinforced concrete and EPS (expanded polystyrene) insulation, while the building envelope is composed by prefabricated wood-cement blocks, shown in Figure \ref{F:houseSection}, with EPS and graphite insulation, that allow low thermal fluxes thanks to their high thermal performance.
+The house is equipped with a biomass boiler, a solar thermal plant, a stand-alone photovoltaic system, black water and rainwater reuse systems and a well for water supply, and has complete independence from the utilities.
+The bearing structure of the house is made of reinforced concrete and EPS (expanded polystyrene) insulation, while the building envelope is composed by prefabricated wood-cement blocks, shown in Figure \ref{F:houseSection}, with EPS and graphite insulation, that allow low thermal fluxes.
 The thermal performance of a wood-cement block, with similar geometry, was investigated in a previous work \cite{Nardi2016}, both with experimental and numerical approach.
@@
 
-To assess energy performance of the use case heating system, the EnergyPlus engine \cite{energyPlus} together with DesignBuilder modeling environment \cite{designBuilder} has been employed.
+To assess energy performance of the use case heating system, the EnergyPlus \cite{energyPlus} together with DesignBuilder modeling environment \cite{designBuilder} has been employed.
 The model has been created based  on L'Aquila weather data, shown in Figure \ref{F:houseExternalWeather}, provided by the CETEMPS – Centre of Excellence \cite{CETEMPS}.
@@
 In \eqref{Eq:energyConsumption} $\dot Q$ is the thermal energy consumption $[Wh]$, $\rho$ is the water density $[kg/m^3]$, $\Delta V$ is the water volume variation $[m^3]$ detected by the turbine flowmeter, $c_p = 4.186\, [kJ/(kgK)]$ is the water specific heat at constant pressure, $\Delta T$ is the difference between flow and return temperatures of the water $[K]$, $0.2777698∙10−3$ is a dimensionless conversion factor.
-The considered period goes from March $15^{th}$ $2016$ to April $15^{th}$ $2016$.
+The considered period goes from March $15$, $2016$ to April $15$, $2016$.
 \begin{figure}[t!]
@@
 All this features are used to create models in $S_1$ and the power model in $S_2$, while disturbance data, state temperature of room $i$ only and flow rate are used to identify $\hat{\Theta}^{T}_{\tT^i_j}$ for temperature models in $S_2$.
-All the models have been trained on the data from March $11^{th}$ $2016$ to April $26^{th}$ $2016$ and validated on the data from May $1^{st}$ $2016$ to May $15^{th}$ $2016$.
+All the models have been trained on the data from March $11$, $2016$ to April $26$, $2016$ and validated on the data from May $1$, $2016$ to May $15$, $2016$.
 The accuracy of these models with respect to real data is shown in Table \ref{T:S1accuracy}, based on the definition of Normalized Root Mean Square Error (NRMSE).
@@
 \end{align}
-The choice of different weights $Q$, $\lambda_{min}$ and $\lambda_{max}$ allows the designer to give more importance to energy consumption rather than temperature comfort and viceversa.
+The choice of different weights $Q$, $\lambda_{min}$ and $\lambda_{max}$ allows the designer to give more importance to energy consumption rather than temperature comfort and vice versa.
 In Section \ref{SS:simulationResults} we will show different performance results considering different weights.
@@
 From historical data we got that $\overline{\tX}^c = 0.35 \, m^3/h$ when the heating system is ON and obviously $\underline{\tX}^c = 0 \, m^3/h$ when the heating system is OFF.
-For the temperature comfort range we set a constant upper bound $\overline{\tT}_{\mathrm{k}} = 22.5 \, \degree C$ and a variable lower bound that is $\underline{\tT}_{\mathrm{k}} = 21 \, \degree C$ from $7am$ to $9am$, that is when people in the house wake up and go out for work, and from $6pm$ to midnight, that is when people come back from work and go to sleep, and $\underline{\tT}_{\mathrm{k}} = 20 \, \degree C$ during other hours, i.e.
-when people are either not at home or asleep.
+For the temperature comfort range we set a constant upper bound $\overline{\tT}_{\mathrm{k}} = 22.5 \, \degree C$ and a variable lower bound that is $\underline{\tT}_{\mathrm{k}} = 21 \, \degree C$ from $7am$ to $9am$ when people in the house wake up and go out for work, and from $6pm$ to midnight when people come back from work and go to sleep.
+During other hours when people are either not at home or asleep, we set $\underline{\tT}_{\mathrm{k}} = 20 \, \degree C$
 
 We explained in Section \ref{SS:descriptionHouse} that the fuel used in the house is vegetable biomass.
-Therefore it is not possible to have ON/OFF switching phases too close to each other, differently from a traditional gas boiler, due to the burning process and heat exchange.
+Therefore, it is not possible to have ON/OFF switching phases too close to each other, differently from a traditional gas boiler, due to the burning process and heat exchange.
 For this reason we set up both control problems with the constraint that when the heating system is activated it must stay active for at least $20$ minutes.
@@
 Each set allows a different level of temperature bounds violation.
-In particular we considered a small ($Q=100$, $\lambda_{min}=3000$ and $\lambda_{max}=100$), a medium ($Q=100$, $\lambda_{min}=1000$ and $\lambda_{max}=100$) and a large ($Q=100$, $\lambda_{min}=100$ and $\lambda_{max}=100$) violation configuration.
-\paragraph{Result 1} For the sake of clarity, we report in Figure \ref{F:comparison_small} only the comparison for temperature and input schedule obtained allowing small bound violations in DPC.
-For the same reason the simulation period is restricted to $4$ days and a half, from May $1^{st}$ $2016$ at $00am$ to May $5^{th}$ $2016$ at $1pm$, in order to make the plots more clear.
+In particular, we considered a small ($Q=100$, $\lambda_{min}=3000$ and $\lambda_{max}=100$), a medium ($Q=100$, $\lambda_{min}=1000$ and $\lambda_{max}=100$) and a large ($Q=100$, $\lambda_{min}=100$ and $\lambda_{max}=100$) violation configuration.
+\paragraph{Result 1} The comparison for temperature and input schedule obtained allowing small bound violations in DPC is shown in Figure \ref{F:comparison_small}.
+For sake of clarity, the simulation period is restricted to $4$ days and a half, from May $1$, $2016$ $00am$ to May $5$, $2016$ $1pm$.
 We can see that the temperature controlled with DPC does not violate the bounds and if it does then the violation is approximately $0.1\,\degree C$.
@@
 We can see how the DPC control law requires the heating system to be ON for less time than the bang-bang one to keep the temperature in the comfort range.
-We will see in Figure \ref{F:comparison_all_energy_E+} that this translates in energy saving.
+We will see in Figure \ref{F:comparison_all_energy_E+} that this translates to significant energy saving.
 
@@
 
-\paragraph{Result 2} In Figure \ref{F:comparison_all_temperature} a comparison of temperature regulation obtained running the DPC with small, medium and large violations is shown.
+\paragraph{Result 2} A comparison of temperature regulation obtained running the DPC with small, medium and large violations is shown in Figure \ref{F:comparison_all_temperature}.
 The results show that with the large violation configuration, that gives more importance to the power consumption minimization than to keep temperature within the bounds, temperature is almost always outside the lower bound during the period when the range is tighter.
@@
 In this case, since the plot is clear and we are interested in showing energy saving on a long period, we ran the simulations over the whole testing period, i.e.
-$15$ days of May, from May $1^{st}$ $2016$ to May $15^{th}$ $2016$.
-We can see that to the bang-bang control strategy is associated an energy consumption approximately equal to $177\,kWh$.
+$15$ days of May, from May $1$, $2016$ to May $15$, $2016$.
+We observe that the energy consumption associated to the bang-bang control strategy is approximately equal to $177\,kWh$.
 If we want to keep the temperature within a comfort range, only allowing small violations, the use of DPC produce an energy saving of $45\, kWh$, that corresponds to the $25.4\%$ over a period of $15$ days.

### 2017-09-30 paper/conclusion.tex
final edits from achin

@@
 We implemented DPC to guarantee energy saving while guaranteeing thermal comfort for the occupants.
-We obtained, with respect to a bang-bang controller, an energy saving that goes from the $25.4\%$, when we want the temperature to be strictly in a comfort range, to the $49.2\%$, if we allow little violations of such a range.
+With respect to a bang-bang controller, we obtained an energy saving that goes from $25.4\%$, when we force the temperature to be strictly in a comfort range, to $49.2\%$, if we allow small violations.
 

### 2017-09-30 paper/conclusion.tex
implemented Rahul's comments

@@
 
-We present two algorithms based on trees and random forests for receding horizon control with data-driven models.
-
-We compare the performance of our Data Predictive Control to MPC on a multivariable bilinear building model.
+To overcome the difficulties associated with the model identification in Model Predictive Control (MPC), we introduce a novel idea for predictive control using data: Data Predictive Control (DPC).
+Data-based control strategies are usually based on non-physical models, therefore the can not be integrated with most of the classical control approaches.
+The goal is to create data-driven models that are suitable for receiding horizon control.
+To this aim, we present two algorithms, based on trees and random forests, to create data-driven models to be used in the DPC strategy.
+We then apply DPC to $3$ different case study in order to prove its strength.
+\begin{enumerate}
+\item \emph{\textbf{Comparison with MPC}} We compare the performance of our Data Predictive Control to MPC on a multivariable bilinear building model.
 We establish that DPC with random forests shows a remarkable similarity to MPC in the optimal control strategies explaining $70\%$ variance.
 On the other hand, DPC with regression trees suffers from practical limitations due to model overfitting.
-
-We further apply DPC with random forests to a large scale 6 story EnergyPlus model with 22 zones for which the traditional model-based control is largely unsuitable due to complex dynamics and the cost of model identification.
+\item \emph{\textbf{Demand Response application}} We further apply DPC with random forests to a large scale 6 story EnergyPlus model with 22 zones for which the traditional model-based control is largely unsuitable due to complex dynamics and the cost of model identification.
 We show that DPC, relying only on the sensor data, can provide significant energy savings while maintaining thermal comfort.
 Our results demonstrate that even for such complex system, DPC tracks a reference signal with a mean error of $3\%$.
-
-\begin{figure}[t!]
-	\begin{center}
-		\includegraphics[width=26pc]{figures/Energy_all_EnergyPlus.eps}
-	\end{center}
-	\caption{Comparison of DPC and bang-bang controller performance in terms of thermal energy saving with different violations using EnergyPlus model.}
-	\label{F:comparison_all_energy_E+}
-\end{figure}
-
-We finally apply DPC using data from a real house and compare it with the classical bang-bang controller widely used for temperature control in houses.
+\item \emph{\textbf{Optimal heating system scheduling of a real house}} We finally apply DPC using data from a real house and compare it with the classical bang-bang controller widely used for temperature control in houses.
 We implemented DPC to guarantee energy saving while guaranteeing thermal comfort for the occupants.
 With respect to a bang-bang controller, we obtained an energy saving that goes from $25.4\%$, when we force the temperature to be strictly in a comfort range, to $49.2\%$, if we allow small violations.
-
+\end{enumerate}
 DPC has applications which go beyond buildings and energy systems, to industrial process control, and controlling large critical infrastructures like water networks, district heating \& cooling.

### 2017-12-24 paper/final/Response_to_reviewers/Response_to_reviewers.tex
first change fot the revision

\documentclass{article}
\textwidth 6in%
\textheight 7.9in%
\oddsidemargin 1cm
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}%
\usepackage[dvips]{color}
\setcounter{MaxMatrixCols}{30}
\usepackage{natbib}

\def\R{\mathcal{R}}
\def\P{\mathcal{P}}
\def\O{\mathcal{O}}
\def\L{\mathcal{L}}
\def\S{\mathcal{S}}
\def\W{\mathcal{W}}
\def\E{\mathcal{E}}
\def\M{\mathcal{M}}

\begin{document}

\title{Authors' Response \\
\begin{Large}
	New title
\end{Large}
}
\author{Francesco Smarra, Achin Jain, Tullio de Rubeis, \\ Dario Ambrosini, Alessandro D’Innocenzo, Rahul Mangharam}

\date{}
\maketitle

\bigskip

The authors would like to thank the Associate Editor and the Reviewers for their constructive and helpful comments. The following major and minor revisions have been applied (all changes in the manuscript are highlighted in \textcolor[rgb]{0.00,0.00,1.00}{blue colored fonts}):
\begin{itemize}
	\item Section 1 - INTRODUCTION
	\begin{itemize}
		\item 
		\item 
	\end{itemize}
	\item Section 2 - DATA PREDICTIVE CONTROL
	\begin{itemize}
		\item 
		\item 
		\item 
	\end{itemize}
	\item Section 3 - COMPARISON WITH MPC
	\begin{itemize}
		\item 
	\end{itemize}
	\item Section 4 - APPLICATION: DEMAND RESPONSE
	\begin{itemize}
		\item 
	\end{itemize}
	\item Section 5 - CASE STUDY: OPTIMAL HEATING SYSTEM SCHEDULING
	\begin{itemize}
		\item 
	\end{itemize}
	\item Minor changes
	\begin{itemize}
		\item 
		\item 
		\item 
		\item 
		\item 
	\end{itemize}
\end{itemize}

The following are itemised lists of the authors' response to the Reviewers and to the Associate Editor.

\section{Response to Associate Editor}

\begin{enumerate}
	\item \textbf{Associate Editor:} \textit{An updated and complete literature review should be conducted. The relevance to Applied Energy should be enhanced with the considerations of scope and readership of the Journal.  The present form has little relevance to APEN which might suit better for another journal. Authors must address and enhance the relevance to Applied Energy otherwise, they can consider to publish in another journal. }.

	\textbf{Authors:}

	\item \textbf{Associate Editor:} \textit{A proof reading by a native English speaker should be conducted to improve both language and organization quality. }

	\textbf{Authors:}

	\item \textbf{Associate Editor:} \textit{The originality of the paper needs to be further clarified. The present form does not have sufficient results to justify the novelty of a high quality journal paper. }

	\textbf{Authors:}

	\item \textbf{Associate Editor:} \textit{The results should be further elaborated to show how they could be used for the real applications.}

	\textbf{Authors:}

\end{enumerate}

\section{Response to Reviewer 1}

\begin{enumerate}
\item \textbf{Reviewer 1:} \textit{This paper presents a new framework called data predictive control which integrates machine learning and model predictive control to derive the optimal online control decisions. The framework is then applied to study several demand response problems in energy area as case studies. The problem considered in this paper is important, and the proposed approach is a good initial attempt to address the problem relying on heuristics.}.

\textbf{Authors:}

\item \textbf{Reviewer 1:} \textit{In Section 2.1, what is the relationship between these training features and labels $(X^c, X^d, Y)$ and the input, output, disturbance, and state variables in the traditional state space model? Are they the same?}

\textbf{Authors:} 
	\begin{itemize}
		\item Explain in the introduction (other than in section 2) the difference between the state-space modeling and the one we propose;
		\item Add a subsection in the "Related work" to explain the evolution of our work with respect to our previous ones;
		\item Add Appendix on regression trees and random forests;
		\item Add equations in the intro. In particular, add a section where we quantify the complexity of a physics-based state-space modeling through an example. Setup and write equation to show the complexity
		\item add another section to show how the previous point simplifies using data-driven methods
	\end{itemize}

\item \textbf{Reviewer 1:} \textit{Another point is that in traditional state space model, one cannot measure everything except input and output in most cases. How could we obtain such training data if these data are not measurable.}

\textbf{Authors:}

\item \textbf{Reviewer 1:} \textit{The key intuition of the proposed framework is Eq. (1). It would be much better if the authors can explain their framework using Eq. (1) directly first instead of using another set of notations such as $(X^c, X^d, Y)$.}

\textbf{Authors:}

\item \textbf{Reviewer 1:} \textit{The relationship between Eq. (1) and Fig. (1) is not very clear. Specifically, one would expect to estimate $g$ and $h$ using data, but how they can be trained in a two-step process as shown in Fig. 1 is not clear.}

\textbf{Authors:}

\end{enumerate}

\section{Response to Reviewer 2}

\begin{enumerate}

\item \textbf{Reviewer 2:} \textit{The paper applies a recently developed concept, data predictive control (DPC) to building energy control. Models based on decision trees and random forests are applied. The efficacy of the approach appears well validated.}

\textbf{Authors:}

\item \textbf{Reviewer 2:} \textit{The title is a bit confusing as it's very general. It should be changed to be more specifically focusing on the applications in this paper.}

\textbf{Authors:}

\item \textbf{Reviewer 2:} \textit{Also, novelties beyond what's in [9] - [12] should be clarified.}

\textbf{Authors:}

\item \textbf{Reviewer 2:} \textit{In (1) and 2.1, the description of separation of variables is vague and needs to be more precise. E.g., the line below (1) does not make sense. It is not until later seeing the examples the meaning of this part becomes clearer. Such confusion persists up to and beyond (3). Mathematically, (2) and (3) cannot be simultaneously true. While I can see the authors trying to be descriptive, the descriptions lack rigor.}

\textbf{Authors:}

\item \textbf{Reviewer 2:} \textit{The main optimization problems (4) and (7) are not clear in themselves. As you look ahead to do DPC, does each future time slot have a different linear model, depending on future non-manipulative states?}

\textbf{Authors:}

\item \textbf{Reviewer 2:} \textit{Then, do you assume you know precisely all the future non-manipulative states? If so, can you justify this? Apparently, with a little error in predicting of the future states, a completely different linear model may be chosen. Will this change the performance significantly? Even with forests this is a major concern.}

\textbf{Authors:}

	\begin{itemize}
		\item Cite study that shows that short term predictions are extremely reliable and say there is a trade-off between the prediction accuracy and the horizon N;
		\item Provide a simulation (as for example with the bilinear model) where we inject a noise into the disturbance and show that it works as well.
	\end{itemize}

\item \textbf{Reviewer 2:} \textit{In Fig. 6b, what's the difference between blue and yellow, as they seem to completely overlap?}

\textbf{Authors:}

\item \textbf{Reviewer 2:} \textit{Also, what is new in section 5 compared with section 4? Even though a real house is claimed to be used, it's still tested using simulations and energyplus, like in the previous sections.}

\textbf{Authors:}

\item \textbf{Reviewer 2:} \textit{A real world experiment on this house would be much more interesting.}

\textbf{Authors:}

\end{enumerate}
%\bibliographystyle{plain}

%\bibliography{mcnbib}
\end{document}

### 2017-12-24 paper/final/casestudyEPlus.tex
first change fot the revision

--- 2017-09-29 paper/casestudyEPlus.tex
@@
 (1) Power forests are built using output as the total building power consumption, and (2) Temperature forests with output as temperature of one of the 22 zones.
-
 The training data set contains the following types of features.
@@
 (3) The \textit{building data} include (i) cooling set points for  the guest rooms, kitchen and corridors, (ii) supply air temperature, and (iii) chilled water temperature.
-
-For the following simulations, we use five control variables (i) cooling set point for corridors $\mathsf{ClgSP}$, (ii) cooling set point for guest rooms $\mathsf{GuestSP}$, (iii) cooling set point for kitchen $\mathsf{KitchenSP}$, (iv) chilled water supply temperature $\mathsf{ChwSP}$, and (v) supply air temperature  $\mathsf{SupplyAirSP}$, so $\tX^c = [\mathsf{ClgSP},\mathsf{GuestClgSP},\\\mathsf{KitchenClgSP},\mathsf{SupplyAirSP}, \mathsf{ChwSP}]$.
+For the following simulations, we use five control variables (i) cooling set point for corridors $\mathsf{ClgSP}$, (ii) cooling set point for guest rooms $\mathsf{GuestSP}$, (iii) cooling set point for kitchen $\mathsf{KitchenSP}$, (iv) chilled water supply temperature $\mathsf{ChwSP}$, and (v) supply air temperature  $\mathsf{SupplyAirSP}$, so $\tX^c = [\mathsf{ClgSP},\mathsf{GuestClgSP},\mathsf{KitchenClgSP},\mathsf{SupplyAirSP}, \mathsf{ChwSP}]$.
 The power forest $\mathcal{R}_p$ is built using the total building power consumption $\tP$.
@@
 The temperature forest $\mathcal{R}_t$ is built with zone temperature $\tT$ as the output. Except for the lag terms corresponding to the same zone temperature, all other features are same in $\tX^d$.
-
 Figure~\ref{F:sepvars} shows the prediction accuracy for the power forest, and also explains the two level training approach introduced in Section~\ref{SS:sepvar}.
@@
 We observe how the accuracy is drastically improved after including the linear models in the predictions.
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=24pc]{figures/eplus_validation.eps}	
-	\caption{Model accuracy during training: The prediction made by forest using only $\tX^d$ (red) captures the effect due to disturbances.
-The linear models in the leaves capture the local effects (green) due to the control inputs $\tX^c$ and improve the model accuracy.}
-	\label{F:sepvars}
-	\captionsetup{justification=centering}
-\end{figure}
 
@@
 It may not be possible to have the same thermal comfort level in all the zones due to power curtailment, so we choose one zone (for example CEO's office) where the constraints must be met.
+\begin{figure}[t!]
+	\centering
+	\includegraphics[width=30pc]{figures/eplus_validation.eps}	
+	\caption{Model accuracy during training: The prediction made by forest using only $\tX^d$ (red) captures the effect due to disturbances.
+The linear models in the leaves capture the local effects (green) due to the control inputs $\tX^c$ and improve the model accuracy.}
+	\label{F:sepvars}
+	\captionsetup{justification=centering}
+\end{figure}
 This is done by solving the following optimization problem with control variables $\tX^c = [\mathsf{ClgSP},\mathsf{GuestClgSP},\mathsf{KitchenClgSP},\mathsf{SupplyAirSP}, \mathsf{ChwSP}]$ as defined before:
@@
 & \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tX}^c \leq \tX^c_{\mathrm{k+j-1|k}} \leq \bar{\tX}^c\\ 
-& \ \ \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0,\delta_j \geq 0, \ j = 1,\dots,N.
+& \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0,\delta_j \geq 0, \ j = 1,\dots,N.
 \end{aligned}
@@
 		\centering
-		\includegraphics[width=22pc]{figures/eplus_control.eps}
+		\includegraphics[width=29pc]{figures/eplus_control.eps}
 	}
@@
 		\centering
-		\includegraphics[width=22pc]{figures/eplus_closedloop.eps}
+		\includegraphics[width=29pc]{figures/eplus_closedloop.eps}
 	}

### 2017-12-24 paper/final/comparison.tex
first change fot the revision

We now run DPC and MPC controller in a closed-loop simulation with the HAMLab plant model. 
The objective function and the box constraints on input and states are exactly same in both cases. Only difference in the optimization problem is due to system dynamics. In particular, MPC solves \eqref{E:mpc},  DPC-RT solves \eqref{E:dpcrt} and DPC-En \eqref{E:dpcrf}. $\mathsf{T}_{\mathrm{ref}}$ in the optimization problems is set to $18\,\mathrm{^oC}$.

\subsection{Simulation Settings}
We test the performance of DPC controller against MPC controller on June 1, 2000. The choice of this day is arbitrary, and does not fall in the training period. For the HAMLab simulation model under consideration, the hourly weather predictions are only available from 1971-2000, however, the choice of the year does not influence the comparison. 
Heat gain due to occupancy $\mathsf{Q}_{\mathrm{oc}}$ is discrete variable which is defined to be $500\,\mathrm{W}$ from $8\,\mathrm{am}$ to $6\,\mathrm{pm}$ and $0\,\mathrm{W}$ otherwise. The disturbances on June 1, 2000 are shown in Fig.~\ref{F:dist}. The sampling time $T_s$ in MPC and model training for DPC is $5\,\mathrm{min}$, while the weather data is available after every $1\,\mathrm{h}$, so these disturbances are kept constant for 12 time steps. 
$N$ and $\delta$ are again chosen to be $6$. For solving optimization numerically, we use Gurobi Optimizer \cite{Gurobi2015}.

\subsection{Results}

Optimal control strategies for all 3 methods are shown in Fig.~\ref{F:control}. The chosen reference temperature and external disturbances require both heating and cooling at some point during the day. The solution obtained from MPC sets the benchmark that we compare to. 
Note that the MPC implementation uses the exact knowledge of the plant dynamics. Therefore, the associated control strategy is indeed the optimal strategy for the plant.
In reality the quality of data, and modeling errors also affect the performance of the MPC controller~\cite{behl2014model}.
Qualitatively, the DPC-En control input is not much different from that of MPC. Until $8\,\mathrm{am}$, when the disturbances are not changing to a great extent, MPC and DPC-En are very close. Slightly after $8\,\mathrm{am}$, when solar heat, external temperature and heat due to occupancy all increase abruptly, we observe that DPC-En cools less and differs slightly from the benchmark. Although DPC-RT also maintains the trend in the control input when the disturbances increase or decrease, the control strategy is quite different from that of MPC. Due to high variance in the predictions, DPC-RT also results in non-smooth inputs, which are not good for practical reasons of switching constraints on heating and cooling equipment in buildings.

\begin{figure}[t!]
\centering
\includegraphics[width=21pc]{figures/pred_hzn.eps}
\caption{Accuracy of predictions with snapshots in time: at time $k_1$, we make 6 predictions for time $k_1$ to $k_1+5$. The numbers represent mean accuracy in predictions for these steps for trees (top) and ensembles (bottom). This process is repeated at time $k_2, \ldots, k_5$.}
\captionsetup{justification=centering}
\label{F:pred_accuracy}
\end{figure}
\begin{figure}[t!]
\centering
\includegraphics[width=21pc]{figures/residual.eps}
\caption{Residual error $\mathsf{T}_{\mathrm{pred}} - \mathsf{T}_{\mathrm{act}}$ for the tree $\mathcal{T}_{1}$  and the forest $\mathcal{R}_1$ that predict $\mathsf{T}_{\mathrm{in,k+1}}$ shows that the residuals for the forest are more concentrated around 0, while tree predictions have a much higher variance.}
\captionsetup{justification=centering}
\label{F:residuals}
\end{figure}
Fig.~\ref{F:state} shows the plot for the zone air temperature. Again, DPC-RT shows a bigger deviation from the optimal solution while DPC-En is near-optimal. This is attributed to linear model averaging in ensemble learning which improves the model prediction. This improvement comes at the cost of reduced interpretability.

\begin{figure}[t!]
\subfigure[Hourly weather data $\mathsf{T}_{\mathrm{ex}}$ and $\mathsf{Q}_{\mathrm{sl}}$ and internal disturbance $\mathsf{Q}_{\mathrm{in}}$.]{
\label{F:dist}
\centering
\includegraphics[width=21pc]{figures/dist.eps}
}
\subfigure[Optimal control strategies $\mathsf{Q}_{\mathrm{in}}$.]{
\label{F:control}
\centering
\includegraphics[width=21pc]{figures/control.eps}
}
\subfigure[Optimal internal zone air temperature $\mathsf{T}_{\mathrm{in}}$.]{
\label{F:state}
\includegraphics[width=21pc]{figures/state.eps}
}
\subfigure[Cumulative sum of the objective function.]{
\label{F:obj}
\centering
\includegraphics[width=21pc]{figures/obj.eps}
}
%\vspace{-1cm}
\caption{Comparison of optimal performance obtained with MPC, DPCRT and DPC-En on June 1, 2000. All 3 controllers start from the same state of the model.}
\captionsetup{justification=centering}
\end{figure}
The quantitative results from the simulations are tabulated in Tab.~\ref{T:quant}. We are interested in evaluating how close DPC performs in comparison to optimal benchmark set by MPC. The coefficient of determination or the $R^2$ score shows that 95.3\% variability in MPC has been accounted for by DPC-En, while DPC-RT captures only 83.9\% variability. In other words, DPC-En and MPC control inputs are 95.3\% close. Fig.~\ref{F:obj} shows the cumulative sum of the objective function as a function of time. At the end of the day, the cumulative sum for DPC-En is 3.8\% more than MPC, and for DPC-RT it is 8.1\% more than MPC. While both DPC-En and DPC-RT are near-optimal, the value of the objective function validates that DPC-En is more closer to the optimal solution.
In the considered scenario, MPC tracks the reference temperature more closely offering better thermal comfort than DPC-RT or DPC-En by spending more energy.

### 2017-12-24 paper/final/cs-mpc.tex
first change fot the revision

--- 2017-09-17 paper/cs-mpc.tex
@@
 The finite RHC approach involves optimizing a cost function subject to the dynamics of the system and the constraints, over a finite horizon of time \cite{Mayne2000}.
-After an optimal sequence of control inputs are computed, the first input is applied, then at the next step the optimization is solved again.
+After an optimal sequence of control inputs is computed, the first input is applied, then at the next step the optimization is solved again.
 
@@
 
-The objective of the controller is to minimize the energy usage $c^Tu$ while maintaining a desired level of thermal comfort.
+The objective of the controller is to minimize the energy usage $c^Tu$ while maintaining a desired level of thermal comfort $x_{ref}$.
 Therefore, at time step $k$, we solve a continuously linearized MPC problem to determine the optimal sequence of inputs $[u_{\mathrm{k|k}},\dots,u_{\mathrm{k+N-1|k}}]$:
@@
 \begin{align}
-\text{min } & \ \ \ \ \sum_{j=1}^{N} {x}^T_{\mathrm{k+j|k}} \mathcal{Q} {x}_{\mathrm{k+j|k}} + c^Tu_{\mathrm{k+j-1}} +  \lambda\epsilon_j\\
+\text{min } & \sum_{j=1}^{N} ({x}^T_{\mathrm{k+j|k}} - x_{ref}) \mathcal{Q} ({x}_{\mathrm{k+j|k}} - x_{ref}) + c^Tu_{\mathrm{k+j-1}} +  \lambda\epsilon_j\\
 \text{s.~t. } & \ \ x_{\mathrm{k+j|k}} =  Ax_{\mathrm{k|k}} + B u_{\mathrm{k+j-1|k}} + B_d d_{\mathrm{k+j-1|k}} \label{SE:mpc1} \\

### 2017-12-24 paper/final/dpc.tex
first change fot the revision

\section{DATA PREDICTIVE CONTROL}
\label{S:dpc}
TO IMPROVE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
\begin{itemize}
	\item Add more details
	\item Review notation
	\item Improve procedure and algorithm description
	\item Add optimization setup discussion
\end{itemize}

%The central idea behind DPC is to obtain control-oriented models using machine learning or black-box modeling, and formulate the control problem in a way that receding horizon control (RHC) can still be applied and the optimization problem can be solved efficiently.
%
%Consider a black-box model given by $x_{k+1}=f(x_k,u_k,d_k)$, where $x,u,d$ represent states, inputs and disturbances, respectively. Depending upon the learning algorithm, $f$ is typically nonlinear, nonconvex and sometimes nondifferentiable (as is the case with regression trees and random forests) with no closed-form expression. 
%Such functional representations learned through black-box modeling may not be directly suitable for control and optimization as the optimization problem can be computationally intractable, or due to nondifferentiabilities we may have to settle with a sub-optimal solution using evolutionary algorithms \cite{Kusiak2009}.
%These problems can be eliminated by decomposing 
%\begin{align}
%f(x_k,u_k,d_k)= g(x_k,h(u_k),d_k),
%\label{E:sepvars}
%\end{align}
%where both $g$ and $h$ are learned using the data, and $h(u_k)$ is convex and differentiable, and thus suitable for optimization. DPC uses this functional decomposition or \textit{separation of variables} to overcome the aforementioned challenges with black-box optimization.
%
%\subsection{Separation of Variables}
%\label{SS:sepvar}
%
%We distinguish between two sets of variables: control (or manipulated) variables $\tX^c \in \R^c$ and disturbance (or non-manipulated) variables $\tX^d \in \R^d$. The union of the two sets forms the full feature set for training, i.e. $\tX \equiv \tX^c \cup \tX^d \in \R^{c+d}$.
%Our goal is to replace a model-based controller with a data-driven controller, where the latter depends only on the historical sensor data. 
%These measurements could directly represent one or more states in the model-based control framework. We denote these as outputs $\tY \in \R$ for training, i.e. a $\tY$ represents a particular output and we can have separate models for multiple outputs. We define the number of training samples by $|(\tX,\tY)| = n$.
%
%Using separation of variables, the training process is divided into two steps. \textit{Step 1:} The trees and the ensembles are trained only on $\tX^d$, which eases the computational complexity. It is important to note that besides external disturbances, $\tX^d$ also contains autoregressive terms of the output $\tY$ which is the main reason for the state space explosion. \textit{Step 2:} Linear regression models are trained in the leaves (or terminal nodes) of the trees which are function of only $\tX^c$. We have validated this linear model assumption in \cite{Jain2016}. As we shall see in Section~\ref{SS:dpcrt} and \ref{SS:dpcrf}, the second step reduces the run-time control problem into a convex program. This process is illustrated in Figure~\ref{F:dpc-sepvars}.
%
%\subsection{DPC-RT: DPC with Regression Trees}
%\label{SS:dpcrt}
%When the data has lots of features, which interact in complicated, nonlinear ways, assembling a single global model such as linear or polynomial regression can be difficult, and can lead to poor response predictions.
%An approach to non-linear regression is to partition the data space into smaller regions, where the interactions are more manageable. 
%This partition is repeated recursively until finally we get to small chunks of the data space where we can fit simple (eg. linear parametric) models. 
%Therefore, in \eqref{E:sepvars}, the global model $f$ has two parts: the recursive partition $g$, and a linear (and convex) model $h$ for each cell of the partition.
%
%\begin{figure}[t!]
%	\centering
%	\includegraphics[width=20pc]{figures/dpc-sepvars.eps}
%	\caption{Separation of variables. \textit{Step 1:} Tree $\mathcal{T}_1$ is trained only on the disturbances $\tX_d$ as the features. Tree $\mathcal{T}_2$ uses both the disturbances $\tX_d$ and the control variables $\tX_c$ for splitting and is thus not computationally suitable for control. \textit{Step 2:} In the leaf $R_i$ of the trees, a linear regression model parametrized by $\beta_i$ is defined as a function only of the control variables.}
%	\captionsetup{justification=centering}
%	\label{F:dpc-sepvars}
%\end{figure}
%
%Now, our goal is to predict the state $\tY$ at time $k$ for next $N$ time steps, i.e. $\tY_{\mathrm{k+1|k}},\dots,\tY_{\mathrm{k+N|k}}$, where $N$ is the control horizon. Applying the separation of variables, we build $N$ regression trees using CART procedure \cite{Breiman1984} such that the output $\tY_{\mathrm{k+j|k}}$ of the $j^{th}$ tree depends upon the previous $N$ disturbances:
%\begin{gather}
%\label{E:model_tree}
%\tY_{\mathrm{k+j|k}} = \mathit{f}_{\mathrm{tree}} \left( \tX^d_{\mathrm{k+j-N|k}} ,\dots,\tX^d_{\mathrm{k+j-1|k}}  \right), \\ \nonumber
%\mathsf{X}^d_{\mathrm{k+j-l|k}} \in \R^d \  \ \forall \ l,j=1,\dots,N.
%\end{gather}
%Then, the linear models as functions of $\tX^c$ in each leaf of the tree $\mathcal{T}_j$ are defined as
%\begin{gather}
%\label{E:model_leaf}
%\tY_{\mathrm{k+j|k}} =  \beta^T_j [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T, \\ \nonumber
%\mathsf{X}^c_{\mathrm{k+j-l|k}} \in \R^c \  \ \forall \ l,j=1,\dots,N.
%\end{gather}
%Note that the coefficients $\beta_j$ would be different for each leaf. Eq.~\eqref{E:model_leaf} implies that the prediction of output $\tY_{\mathrm{k+j}}$ at time $k$ is an affine combination of control inputs from time $k$ to $k+j-1$. Thus, we have managed to linearize the original model dynamics via black-box modeling. This two-step training is done offline. In run-time, given the disturbances $\tX^d_\mathrm{k|k}$ at time $k$, we can narrow down to a leaf of each tree in \eqref{E:model_tree} to retrieve the linear models in \eqref{E:model_leaf}.
%
%In run-time, when a new control action is to be determined, each tree (prediction step) contributes to a linear constraint in the optimization as a replacement for the state dynamics in the case of MPC. Thus, the RHC optimization problem with a quadratic cost ($\mathcal{Q} \geq 0, \mathcal{R} \succeq 0$) can be formulated as:
%\begin{align}
%\begin{aligned}
%\text{min } & \sum_{j=1}^{N} ({\tY}_{\mathrm{k+j|k}})^2 \mathcal{Q} + {\tX^c}^T_{\mathrm{k+j-1|k}} \mathcal{R} {\tX^c}_{\mathrm{k+j-1|k}} +  \lambda\epsilon_j\\
%\text{s.~t. } & \ \ \ \ \ \tY_{\mathrm{k+j|k}} =  \beta^T [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
%& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tX}^c \leq \tX^c_{\mathrm{k+j-1|k}} \leq \bar{\tX}^c\\ 
%& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tY}-\epsilon_j \leq \tY_{\mathrm{k+j|k}} \leq \bar{\tY} + \epsilon_j\\\
%& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0, \ j = 1,\dots,N.
%\end{aligned}
%\label{E:dpcrt}
%\end{align}
%Here, $\mathcal{Q} \in \mathbb{R}$ and $\mathcal{R} \in \mathbb{R}^{c \times c}$, and the slack variables $\epsilon_j$ ensure recursive feasibility since the equality constraint on $\tY$ is relaxed. Of course, a different cost function can be chosen depending upon the application. In the current formulation, the data-driven control problem is reduced to a convex program which is much easier to solve than running an optimization directly on a black-box model trained on $\tX^c$ as features. We solve this optimization in the same manner as MPC to determine the optimal sequence of inputs $[\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+N-1|k}}]$, apply the first control input $\tX^c_{\mathrm{k|k}}$ and proceed to the next time step $k+1$. The pseudo code for DPC-RT is given in Algorithm~\ref{A:dpcrt}.
%
%\begin{algorithm}[t!]
%	\caption{Data Predictive Control with Regression Trees}
%	\label{A:dpcrt}
%	\begin{algorithmic}[1]
%		\State \textsc{Design Time}
%		\Procedure{Model Training using Separation of Vars}{}
%		\State Set $\mathsf{X}^c$ $\gets$ manipulated features
%		\State Set $\mathsf{X}^d$ $\gets$ non-manipulated features
%		\State Build $N$ predictive trees with $(\tY,\tX^d)$ defined in \eqref{E:model_tree}
%		\ForAll{trees $\mathcal{T}_j$}
%		\ForAll{regions $R_i$ at the leaves of $\mathcal{T}_j$}
%		\State Fit $ \tY_{\mathrm{k+j|k}} =  \beta^T_j \left[1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} \right]^T$ as in \eqref{E:model_leaf}
%		\EndFor
%		\EndFor
%		\EndProcedure
%		\State \textsc{Run Time}
%		\Procedure{Predictive Control}{}
%		\While{$k< k_{\mathrm{stop}}$}
%		\ForAll{trees $\mathcal{T}_j$}
%		\State Determine the leaf $R_i$ using $\tX^d$ as in \eqref{E:model_tree}
%		\State Obtain the linear model at $R_{i}$ trained in \eqref{E:model_leaf}
%		\EndFor
%		\State Solve optimization in \eqref{E:dpcrt} to determine optimal
%		\State control actions $[\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+N-1|k}}]$
%		\State Apply the first input $\tX^c_{\mathrm{k|k}}$
%		\EndWhile
%		\EndProcedure
%	\end{algorithmic}
%\end{algorithm}
%
%\subsection{DPC-En: DPC with Ensemble Methods}
%\label{SS:dpcrf}
%Regression trees obtain good predictive accuracy in many domains. However, the models used in their leaves have some limitations regarding the kind of functions they are able to approximate.
%The problem with trees is their high variance and that they can overfit the data easily.
%A small change $\delta$ in the data can result in a different series of splits and thus violate the acceptable accuracy $\eps$.
%This is the price to be paid for estimating a tree-based structure from the data.
%
%
%We use ensemble methods \cite{Friedman2001} to combine the predictions of several independent regression trees in order to improve generalizability and robustness over a single estimator. 
%The essential idea is to average many noisy trees to reduce the overall variance in prediction.
%We inject randomness into the tree construction in two ways. First, we randomize the features used to define splitting in each tree.
%Second, we build each tree using a bootstrapped or sub-sampled data set.
%In this way, each tree in the forest is trained on different data, which introduces differences between the trees. More explicitly, training features $\tX^d \in \R^p$ with $p<d$ and the in-bag samples (in-bag samples correspond to the data samples on which the tree was trained) are different for each tree in the forest i.e $|(\tX,\tY)|<n$.
%
%The goal with DPC-En is to replace each tree in Algorithm~\ref{A:dpcrt} by a forest
%\begin{gather}
%\label{E:model_forest}
%\tY_{\mathrm{k+j|k}} = \mathit{f}_{\mathrm{forest}} \left( \tX^d_{\mathrm{k+j-N|k}} ,\dots,\tX^d_{\mathrm{k+j-1|k}}  \right), \\ \nonumber
%\mathsf{X}^d_{\mathrm{k+j-l|k}} \in \R^p \  \ \forall \ l,j=1,\dots,N,
%\end{gather}
%which, again, is trained only on $\mathsf{X}^d$, with $\tX^d \in \R^p \subset \R^d$ for each tree,  and then fit a linear regression model using $\mathsf{X}^c$ in every leaf of every tree. We build $N$ such forests for $N$ prediction steps such that the leaf $R_i$ of forest $\mathcal{R}_j$ uses a linear model
%\begin{gather}
%\label{E:model_leaf_forest}
%\tY_{\mathrm{k+j|k}} =  \Theta^T_{ij} [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T, \\ \nonumber
%\mathsf{X}^c_{\mathrm{k+j-l|k}} \in \R^c \  \ \forall \ l,j=1,\dots,N.
%\end{gather}
%Here $(\mathsf{X}^c,\mathsf{Y})$ correspond to the in-bag samples for the trees.
%
%While the offline training burden in DPC-En is slightly increased compared to DPC-RT, in the control step we exploit the better accuracy, and lower variance properties of the random forest. 
%If a forest has $t$ number of trees, given the forecast of disturbances, we have $t$ sets of linear coefficients. We simply average out all the coefficients from all the trees to get one linear model represented by $\hat{\Theta}_j$ for each forest. Note that the averaging step can only be done in run-time because the leaf of each tree can be narrowed down only when the $\tX^d$ is known. Thus, for $N$ forests, we again have exactly $N$ linear equality constraints in the optimization problem below:
%\begin{align}
%\begin{aligned}
%\text{min } & \sum_{j=1}^{N} ({\tY}_{\mathrm{k+j|k}})^2 \mathcal{Q} + {\tX^c}^T_{\mathrm{k+j-1|k}} \mathcal{R} {\tX^c}_{\mathrm{k+j-1|k}} +  \lambda\epsilon_j\\
%\text{s.~t. } & \ \ \ \ \ \tY_{\mathrm{k+j|k}} =  \hat{\Theta}^T_j [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
%& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tX}^c \leq \tX^c_{\mathrm{k+j-1|k}} \leq \bar{\tX}^c\\ 
%& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tY}-\epsilon_j \leq \tY_{\mathrm{k+j|k}} \leq \bar{\tY} + \epsilon_j\\\
%& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0, \ j = 1,\dots,N.
%\end{aligned}
%\label{E:dpcrf}
%\end{align}
%DPC-En is graphically described in Figure~\ref{F:dpc-algo-rf}. The ensemble data predictive control (DPC-En) is the first such method to bridge the gap between ensemble predictive models (such as random forests) and receding horizon control. In the next section, we compare DPC-RT and DPC-En to MPC for a building model.
%
%\begin{figure}[t!]
%	\centering
%	\includegraphics[width=26pc]{figures/dpc-algo-rf.eps}
%	\caption{DPC-En: At time $k$, the algorithm uses the forecast of disturbances $\tX^d_{\mathrm{k|k}}$ to select linear models $\Theta_1$ to $\Theta_t$ in the leaves of each ensemble. The linear models in each ensemble are averaged to calculate a single model represented by $\hat{\Theta}_j$ which act as constraints in the optimization problem. The optimal sequence $[\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+N-1|k}}]$, of which the first one is applied, and $\tX^d_{\mathrm{k+1|k+1}}$ is calculated to proceed to $k+1$.}
%	\label{F:dpc-algo-rf}
%\end{figure}

### 2017-12-24 paper/final/intro.tex
first change fot the revision

\section{INTRODUCTION}
\label{S:intro}

Machine learning and control theory are two foundational but disjoint communities. Machine learning requires data to produce models, and control systems require models to provide stability and performance guarantees to plant operations. Machine learning is widely used for regression or classification, but thus far data-driven models have not been suitable for closed-loop control of physical plants. The challenge now, with using data-driven approaches, is to close the loop for real-time control and decision making.

\subsection{Motivation}
TO IMPROVE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}

%Consider a multivariable dynamical system subject to external disturbances. The first and foremost requirement for making any decision is to obtain the underlying control-oriented predictive model of the system. With a reasonable forecast of the external disturbances, these models should predict the state of the system in the future and thus Model Predictive Control (MPC) can act preemptively to provide a desired system behavior while optimizing a desired performance. In particular, MPC has been proven to be very powerful for multivariable systems in the presence of input and output constraints, and forecast of the disturbances. The caveat is that MPC requires a reasonably accurate physical representation of the system. This makes MPC unsuitable for control of complex plants such as natural gas processing, oil refineries, boilers, manufacturing plants, and buildings where the user expertise, time, and associated sensor costs required to develop a model are very high \cite{Sturzenegger2016,vzavcekova2014}.
%
%There are two main reasons for model complexity. 
%(1) The prime contributor is the change in model properties over time. Even if the model is identified once via an expensive route, as the model changes with time, the system identification must be repeated to update the model. Thus, model adaptability or adaptive control is desirable for such systems. 
%(2) A secondary reason is the model heterogeneity which further prohibits the use of model-based control. For example, unlike the automobile or the aircraft industry, each building is designed and used in a different way. Therefore, this modeling process must be repeated for every new building. 
%Due to aforementioned reasons, the control strategies in such systems are often limited to fuzzy logic rules that are based on best practices.

%\paragraph{Example} As an example of such modeling complexity, consider the grey-box approach in \cite{Braun2002}. The scope of such approach is to predict the heat transfer rate to the air within the building. In particular, the authors addressed a single zone modeling problem using an equivalent $RC$ model. To this end, five different types of structures are considered in the example: external walls, ceiling/roof, floor, internal walls and windows. Each of these elements are represented using $3$ resistances, $2$ capacitances and $2$ temperatures, except for the windows that are represented only with resistances since they have a negligible energy storage. A model consisting of $8$ states and $9$ inputs, including disturbances such as solar radiation and others, with the instantaneous heat gain to the building air from all surfaces as output, is created for the zone. To have a complete LTI model, $9$ parameters are needed to be estimated via a non-trivial training process, using a collection of information associated with a physical description of the building (see \cite{Braun2002} for more details). Extending this result, for the building we consider in Section \ref{S:realCaseStudy} with $10$ different zones, we would need to consider a model with approximately $80$ states and $90$ parameters to be estimated.

\subsubsection{physics-based modeling example}
TO BE DONE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
\begin{itemize}
	\item Introduce example
	\item add modeling equations
	\item quantify complexity
\end{itemize}

\subsubsection{data-driven modeling example}
TO BE DONE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
\begin{itemize}
	\item show procedure
	\item quantify complexity
\end{itemize}

\subsubsection{physics-based vs data-driven modeling example}
TO BE DONE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
\begin{itemize}
	\item compare results
\end{itemize}

\paragraph{Objective} The question now is, can we employ data-driven techniques to reduce the cost of modeling and still exploit the benefits that MPC has to offer? Can we build automatic and data-driven approaches for control purposes that are also adaptive, scalable and interpretable? In this paper we address this problem introducing \textit{Data Predictive Control (DPC)} to bridge the gap between Machine Learning and Predictive Control.

\subsection{Related work}
TO IMPROVE $\rightarrow$ \textcolor[rgb]{1,0,0}{Tullio}
%In the literature, several studies deal with the data-driven control techniques using machine learning algorithms \cite{Hou2013}. For example, Artificial Neural Networks are exploited to create data-driven models to be used as plant simulator in closed-loop with a Supervisory Model Predictive Control in \cite{Afram2017}. In \cite{Macarulla2017} the authors proposed a predictive control strategy based on Neural Networks, for boilers control in buildings, to decide the optimal time to switch-on the plan to guarantee energy saving and thermal comfort. However, the strategy is not easily scalable to different types of plants and does not use optimization theory in the closed-loop scheme. In \cite{Costanzo2016} a reinforcement learning control strategy, called Model-Assisted Batch Reinforcement Learning, is considered to provide data-driven control for the demand response problem in HVAC systems. Reinforcement Learning is a model-free methodology and for this reason it can not be used for on-line optimization in a closed-loop predictive control scheme. In \cite{Ferreira2012} the authors considered a data-driven predictive control based on Neural Networks to guarantee energy saving and  thermal comfort in public buildings. Neural Networks are used in the closed-loop control scheme to determine a thermal comfort index based on parameters that can be measured or estimated, but no  system dynamics with internal state are included into the optimization problem. More papers related to this topic can be found in the literature, but to the best of the authors' knowledge none of them address the problem of including data-driven state models in the optimal predictive control loop, and hence allow to set up a MPC-like optimal control problem, as we do in DPC.

\subsection{Main contribution and paper organization}
TO CLARIFY $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
\begin{itemize}
	\item Add more details
	\item Emphasize difference between Section 4 and Section 5
	\item Cite appendix
\end{itemize}
TO ADD $\rightarrow$ \textcolor[rgb]{1,0,0}{Tullio}
\begin{itemize}
	\item Add references for weather forecast accuracy
	\item Quantify accuracy for short-term prediction for weather forcast to justify the assumption of perfect knowledge for the disturbance
\end{itemize}
TO ADD $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
\begin{itemize}
	\item Add subsection in Section 5 with simulation of DPC considering disturbance with added noise and show that we still have good performance
\end{itemize}
%In our previous work \cite{Behl2016,Jain2016,JainACC2017,JainCDC2017}, we introduced the concept of DPC for receding horizon control. In this paper, we extend these results providing the following contributions.
%\begin{enumerate}
%	\item In Section \ref{S:dpc}, we formally present the two following control techniques:
%	\begin{enumerate}
%		\item DPC with regression trees, and
%		\item DPC with random forests.
%	\end{enumerate}
%	\item In Section \ref{S:proof}, we demonstrate the strength of DPC for receding horizon control via one-to-one comparison against a benchmark MPC controller using a bilinear building model whose parameters were identified using experiments on a building in Switzerland. We show that DPC captures 70\% variance in MPC and offers a comparable performance.
%	\item Section \ref{S:casestudy} describes a practical application of DPC for Demand Response, where we apply DPC to a 6 story 22 zone building model in EnergyPlus \cite{Crawley2001} for which model-based control is not economical and practical due to extreme complexity. We show scalability and efficiency of DPC in providing financial incentives to the end-customers bypassing the need for high fidelity models. We observe that DPC provides the desired power reduction with an average error of 3\%.
%	\item In Section \ref{S:realCaseStudy}, we implement DPC on the real data from an off-grid house located in L'Aquila (Italy), to find the optimal ON/OFF scheduling for the heating system in order to save energy while guaranteeing thermal comfort for the occupants. We quantify the total amount of energy saved, with respect to the classical bang-bang controller widely used in houses for temperature control, using an EnergyPlus model built specifically for the house. We show we can perform an energy saving that goes from $25.4\%$, if we guarantee thermal comfort, to $49.2\%$, if we allow very little discomfort in terms of rooms' temperature.
%\end{enumerate}

### 2017-12-24 paper/final/main.tex
first change fot the revision

--- 2017-09-29 paper/main.tex
@@
 
-\title{Data Predictive Control with Random Forests:\\Theory and Applications\tnoteref{mytitlenote}}
-\tnotetext[mytitlenote]{The research leading to these results has received funding from the Italian Government under Cipe resolution n.135 (Dec.
-21, 2012), project \emph{INnovating City Planning through Information and Communication Technologies} (INCIPICT).}
+\title{Data Predictive Control with Random Forests:\\Theory and Applications}
 
@@
 \author[AJaddress,EC]{Achin Jain}
-\ead{achinj@seas.unpenn.edu}
+\ead{achinj@seas.upenn.edu}
 \author[TDRaddress,EC]{Tullio de Rubeis}
@@
 \begin{abstract}
-\input{abstract.tex}
+%\input{abstract.tex}
 \end{abstract}
@@
 
+\appendix
+\section{REGRESSION TREES}\label{AppendixRT}
+TO BE DONE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
+\section{RANDOM FORESTS}\label{AppendixRF}
+TO BE DONE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
 \end{document}

### 2017-12-24 paper/first submission/Cover letter.tex
first change fot the revision

\documentclass[12pt]{article}

\begin{document}

\title{Highlights}
\date{}
\maketitle 
\thispagestyle{empty}

This work contains two novel aspects with respect to existing papers in the literature. The first contribution provides a new modeling framework for a class of MIMO WNCS that takes into account the effect of industrial control protocols, such as WirelessHART and ISA-100, and the presence of packet losses. To do this we used the framework of Markov Jump Linear Systems. The second (and main) contribution of this paper provides, through the use of the Gerschgorin disk theory, an efficient sub-optimal methodology to design the parameters of the WNCS in order to optimize some performance while guaranteeing the stability of the closed-loop system. Due to the presence of packet losses the concept of stability is intended in the stochastic sense.

\section{Why the paper is relevant to the journal?}
In this paper two novel contributions are provided:
\begin{enumerate}
	\item A new modeling framework for Wireless Networked Control Systems (WNCS) affected by packet losses is proposed leveraging Markov Jump Linear Systems theory;
	\item A new control sub-optimal methodology to design network parameters to make the WNCS robust with respect to packet losses is proposed leveraging Gerschgorin disk theory.
\end{enumerate}
Both results perfectly fit with the requirements provided in the "aims and scope" web section of this journal. In particular, the first contribution provides "design techniques for uncertain linear [...] systems", while the second one provides "non-optimal method(s) of improving the robustness of uncertain systems".

\section{Why the theoretical contributions in the paper are novel?}
To the best of the authors' knowledge this is the first paper that provides a modeling framework for WNCS that takes into account both the effect of industrial control protocols, such as WirelessHart and ISA-100, and the effect of packet losses leveraging Markov Jump Linear Systems theory. Thanks to this modeling framework, a new sub-optimal methodology based on Gerschgorin disks theory to design WCNS parameters to maximize the robustness of the system is provided in the paper. To the best of the authors' knowledge this is also a novel contibution.

\section{List of keywords}
\begin{itemize}
	\item Wireless Networked Control Systems;
	\item Control of networks;
	\item Control over networks;
	\item Markov Jump Linear Systems.
\end{itemize}
\end{document}

### 2017-12-24 paper/first submission/ack.tex
first change fot the revision

\section*{ACKNOWLEDGMENT}
This work was supported partially by the Italian Government under Cipe resolution n.135 (Dec. 21, 2012), project \emph{INnovating City Planning through Information and Communication Technologies} (INCIPICT), and by TerraSwarm, one of six centers of STARnet, a Semiconductor Research Corporation program sponsored by MARCO and DARPA.

The authors would like to thank Xiaojing Zhang, a PostDoctoral Researcher at the University of California, Berkeley, for providing the bilinear building model, Iole Nardi, a PostDoctoral Researcher at the University of L'Aquila, for the help in obtaining the real data from the house, and Manfred Morari, for his feedback on DPC.

### 2017-12-24 paper/first submission/casestudyReal.tex
first change fot the revision

--- 2017-09-30 paper/casestudyReal.tex
@@
 Finally, in Section \ref{SS:simulationResults}, the performance of the two controllers are compared in terms of energy saving using the EnergyPlus model.
+In particular, we show that DPC allows an energy saving up to $49.2\%$ with respect to the bang-bang controller while guaranteeing thermal comfort for occupants.
 The section organization is graphically shown in Figure \ref{F:overview}.
-
-\subsection{Description of the house}\label{SS:descriptionHouse}
-The chosen case study is a detached off-grid two-story residential house, located in the outskirt of L'Aquila (coordinates $42\degree$ $16'$ latitude and $13\degree$ $32'$ longitude), Italy and is shown in Figure \ref{F:house}.
-The building, inhabited by the two owners, has a main north-south orientation and it is composed by a heated ground floor and an attic without heating system.
-Therefore, although the gross area of the house is equal to $209.5\,m^2$, the heated gross area is equal to $112.4\,m^2$.
-\begin{figure}[t!]
+\begin{figure}[h!]
 	\begin{center}
@@
 \end{figure}
+
+\subsection{Description of the house}\label{SS:descriptionHouse}
+The chosen case study is a detached off-grid two-story residential house, located in the outskirt of L'Aquila (coordinates $42\degree$ $16'$ latitude and $13\degree$ $32'$ longitude), Italy and is shown in Figure \ref{F:house}.
+The building, inhabited by the two owners, has a main north-south orientation and it is composed by a heated ground floor and an attic without heating system.
+Therefore, although the gross area of the house is equal to $209.5\,m^2$, the heated gross area is equal to $112.4\,m^2$.
+Thanks to the off-grid characteristic, the technological plants guarantee the complete energy self-sufficiency of the building.
+The house is equipped with a biomass boiler, a solar thermal plant, a stand-alone photovoltaic system, black water and rainwater reuse systems and a well for water supply, and has complete independence from the utilities.
+The bearing structure of the house is made of reinforced concrete and EPS (expanded polystyrene) insulation, while the building envelope is composed by prefabricated wood-cement blocks, shown in Figure \ref{F:houseSection}, with EPS and graphite insulation, that allow low thermal fluxes.
+The thermal performance of a wood-cement block, with similar geometry, was investigated in a previous work \cite{Nardi2016}, both with experimental and numerical approach.
+The heating system of the building, that supplies the thermal energy required in winter season, is a hydronic system consisting of a vegetable biomass boiler, with a manual ON/OFF, a constant-flow pump station, and tubular steel radiators as shown in Figure \ref{F:housePlantScheme}.
+The standard biomass boiler has an efficiency of $83.5\%$ with $16.5\ kW$ of thermal power transferred to the water, without gas-flame modulation.
+The heat transfer fluid distribution is realized through a manifold circuit, that supplies the radiators placed in the various rooms of the house.
+The energy needs for the domestic hot water (DHW) are covered by the same boiler, coupled with a solar thermal plant.
+It is worth noting that, in this work, the thermal energy needed for the domestic hot water production is neglected.
 \begin{figure}[t!]
 	\begin{center}
-		\includegraphics[width=24pc]{figures/Vista_sud.eps}
+		\includegraphics[width=20pc]{figures/Vista_sud.eps}
 		\caption{Off-grid residential house.}
@@
 	\end{center}
+%	\vspace{-0.5cm}
 \end{figure}
-Thanks to the off-grid characteristic, the technological plants guarantee the complete energy self-sufficiency of the building.
-The house is equipped with a biomass boiler, a solar thermal plant, a stand-alone photovoltaic system, black water and rainwater reuse systems and a well for water supply, and has complete independence from the utilities.
-The bearing structure of the house is made of reinforced concrete and EPS (expanded polystyrene) insulation, while the building envelope is composed by prefabricated wood-cement blocks, shown in Figure \ref{F:houseSection}, with EPS and graphite insulation, that allow low thermal fluxes.
-The thermal performance of a wood-cement block, with similar geometry, was investigated in a previous work \cite{Nardi2016}, both with experimental and numerical approach.
-\begin{figure}[t!]
+\begin{figure}[h!]
 	\begin{center}
@@
 			\centering
-			\includegraphics[width=20pc]{figures/blocco_parete_rev01.eps}
+			\includegraphics[width=16pc]{figures/blocco_parete_rev01.eps}
 		}
@@
 			\centering
-			\includegraphics[width=20pc]{figures/blocco_solaio_rev01.eps}
+			\includegraphics[width=17pc]{figures/blocco_solaio_rev01.eps}
 		}
@@
 \end{figure}
-The heating system of the building, that supplies the thermal energy required in winter season, is a hydronic system consisting of a vegetable biomass boiler, with a manual ON/OFF, a constant-flow pump station, and tubular steel radiators as shown in Figure \ref{F:housePlantScheme}.
-The standard biomass boiler has an efficiency of $83.5\%$ with $16.5\ kW$ of thermal power transferred to the water, without gas-flame modulation.
-The heat transfer fluid distribution is realized through a manifold circuit, that supplies the radiators placed in the various rooms of the house.
-The energy needs for the domestic hot water (DHW) are covered by the same boiler, coupled with a solar thermal plant.
-It is worth noting that, in this work, the thermal energy needed for the domestic hot water production is neglected.
-
 For a thorough knowledge of the building thermal behavior, an in-situ analysis was carried out.
@@
 The properties of the blocks used for the simulation model are listed in Table \ref{T:houseProperties}.
+
+Therefore, the dynamic simulation model of the use case, shown in Figure \ref{F:houseVirtualModel}, has been created by analyzing the fundamental characteristics of the building (orientation, geometry, structural members, heating system components, air changes with natural ventilation, activity, internal gains, air leakages) and the weather file specifically created for L'Aquila.
+The comprehensive method was chosen for modeling the heating system, once checked all the characteristics of the components.
 \begin{figure}[t!]
 	\begin{center}
-		\includegraphics[width=30pc]{figures/dati_climatici_rev01.eps}
+		\hspace{-0.7cm}
+		\includegraphics[width=28pc]{figures/dati_climatici_rev01.eps}
+		\vspace{-0.5cm}
 		\caption{Weather data of L'Aquila, Italy.}
@@
 \end{table}
-
-Therefore, the dynamic simulation model of the use case, shown in Figure \ref{F:houseVirtualModel}, has been created by analyzing the fundamental characteristics of the building (orientation, geometry, structural members, heating system components, air changes with natural ventilation, activity, internal gains, air leakages) and the weather file specifically created for L'Aquila.
-The comprehensive method was chosen for modeling the heating system, once checked all the characteristics of the components.
-
 In order to calibrate the EnergyPlus model, a comparison between simulated and measured thermal energy consumption was performed.
@@
 		\label{F:houseVirtualModel}
+	\end{center}
+\end{figure}
+\begin{figure}[t!]
+	\begin{center}
+		\includegraphics[width=28pc]{figures/Scheduling_rev02.eps}
+		\caption{Actual scheduling of the heating system, where white color means switched OFF and green color switched ON, and ($T_{ext,ave}$) is the external average temperature.
+From midnight to $8am$ the heating system is always switched off.}
+		\captionsetup{justification=centering}
+		\label{F:houseActualScheduling}
 	\end{center}
@@
 	\begin{center}
-		\includegraphics[width=28pc]{figures/Scheduling_rev02.eps}
-		\caption{Actual scheduling of the heating system, where white color means switched OFF and green color switched ON, and ($T_{ext,ave}$) is the external average temperature.
-From midnight to $8am$ the heating system is always switched off.}
-		\captionsetup{justification=centering}
-		\label{F:houseActualScheduling}
-	\end{center}
-\end{figure}
-\begin{figure}[t!]
-	\begin{center}
-		\subfigure[Thermal energy.]{
+		\subfigure[Thermal energy comparison. The accuracy error is $7.38\%$ with MBE definition and $8.37\%$ with CV(RMSE) definition]{
 			\label{F:houseComparisonExperimentalEnergy}
 			\centering
-			\includegraphics[width=24pc]{figures/Thermal_energy_rev01.eps}
+			\includegraphics[width=28pc]{figures/Thermal_energy_rev01.eps}
 		}
-		\subfigure[Thermal power.]{
+		\subfigure[Thermal power comparison.]{
 			\label{F:houseComparisonExperimentalPower}
 			\centering
-			\includegraphics[width=24pc]{figures/Thermal_power.eps}
+			\includegraphics[width=28pc]{figures/Thermal_power.eps}
 		}
@@
 
-\begin{table}[t!]
-	\centering
-	\begin{tabular}{cccccc}
-		\toprule
-		Set       & Power   & T. room $1$ & T. room $2$ & T. room $3$ & T. room $4$  \\ 
-		\midrule
-		$S_1$     & $92.44$ & $96.58$     & $96.99$     & $97.21$     & $96.62$\\
-		$S_2$     & $92.21$ & $97.38$     & $97.29$     & $96.93$     & $96.81$\\
-		\bottomrule
-	\end{tabular}
-	\caption{Models accuracy for power and temperatures models in $S_1$ and $S_2$ expressed as $\mathrm{1-NRMSE}\,(\%)$.}
-	\captionsetup{justification=centering}
-	\label{T:S1accuracy}
-\end{table}
-
 A graphical comparison is shown in Figure \ref{F:power_testing} for the power consumption and Figure \ref{F:temperature_testing} for the temperature of room $1$.
 The plots for the other rooms are omitted since they are very similar.
-
-\begin{figure}[t!]
-\begin{center}
-	\subfigure[Power consumption model accuracy validation.]{
-		\label{F:power_testing}
-		\centering
-		\includegraphics[width=26pc]{figures/power_testing.eps}
-	}
-	\subfigure[Temperature model accuracy validation.]{
-		\label{F:temperature_testing}
-		\centering
-		\includegraphics[width=26pc]{figures/temperature_testing.eps}
-	}	
-	\end{center}
-\caption{Model accuracy validation.}
-\captionsetup{justification=centering}
-\label{F:testing}
-\end{figure}
 
@@
 We include these violations in the objective function to be minimized.
+\begin{table}[t!]
+	\centering
+	\begin{tabular}{cccccc}
+		\toprule
+		Set       & Power   & T. room $1$ & T. room $2$ & T. room $3$ & T. room $4$  \\ 
+		\midrule
+		$S_1$     &         & $96.58$     & $96.99$     & $97.21$     & $96.62$\\
+		$S_2$     & $92.21$ & $97.38$     & $97.29$     & $96.93$     & $96.81$\\
+		\bottomrule
+	\end{tabular}
+	\caption{Models accuracy for power and temperatures models in $S_1$ and $S_2$ expressed as $\mathrm{1-NRMSE}\,(\%)$.}
+	\captionsetup{justification=centering}
+	\label{T:S1accuracy}
+	\vspace{-0.5cm}
+\end{table}
+\begin{figure}[t!]
+	\begin{center}
+		\subfigure[Power consumption model accuracy validation. The accuracy over the testing period expressed as $\mathrm{1-NRMSE}\,(\%)$ is $92.1\%$.]{
+			\label{F:power_testing}
+			\centering
+			\includegraphics[width=26pc]{figures/power_testing.eps}
+		}
+		\subfigure[Temperature model accuracy validation for room $1$.
+The accuracy over the testing period expressed as $\mathrm{1-NRMSE}\,(\%)$ is $96.58\%$, for the temperature model $T^{r_1}$ in $S_1$ and $97.3\%$ for the temperature model $T^1$ in $S_2$.
+The accuracy for the other rooms is very similar as can be seen in Table \ref{T:S1accuracy}.]{
+			\label{F:temperature_testing}
+			\centering
+			\includegraphics[width=26pc]{figures/temperature_testing.eps}
+		}	
+	\end{center}
+	\vspace{-0.7cm}
+	\caption{Model accuracy validation for power and temperature models.}
+	\vspace{-0.7cm}
+	\captionsetup{justification=centering}
+	\label{F:testing}
+\end{figure}
+
 The problem is set up as follows:
@@
 	\begin{center}
-	\subfigure[Comparison of temperature of room $3$ obtained with DPC and bang-bang controller.]{
+	\subfigure[Temperature variation obtained with DPC and bang-bang controller.DPC controller allows almost no violation, so guaranteeing better comfort than bang-bang controller.]{
 		\label{F:temperatures_small}
 		\centering
-		\includegraphics[width=26pc]{figures/Temperatures_small.eps}
+		\includegraphics[width=28pc]{figures/Temperatures_small.eps}
 	}
-	\subfigure[Input schedules obtained from DPC and bang-bang controller.]{
+	\subfigure[Input schedules obtained from DPC and bang-bang controller.
+DPC keeps the heating system ON for less time than bang-bang controller, hence saving energy, and guarantees better thermal comfort.]{
 		\label{F:inputs_small}
 		\centering
-		\includegraphics[width=26pc]{figures/Inputs_small.eps}
+		\includegraphics[width=28pc]{figures/Inputs_small.eps}
 	}
 	\end{center}
-	\caption{Comparison of DPC and bang-bang control performance.}
+\vspace{-0.5cm}
+	\caption{Comparison of DPC and bang-bang control performance over $5$ days of the testing period for room $3$.}
 	\captionsetup{justification=centering}
@@
 However the maximum violation is still lower than $1\,\degree C$.
-\begin{figure}[t!]
-	\begin{center}
-		\includegraphics[width=26pc]{figures/Temperatures_all.eps}
-	\end{center}
-	\caption{Comparison of DPC control performance in terms of comfort with different violations.}
-	\label{F:comparison_all_temperature}
-\end{figure}
 In Table \ref{T:violationErrors}, MBE and CV(RMSE) violation errors are reported to quantify the bounds violation of DPC, in each of the $3$ configurations, and of bang-bang controller.
@@
 Instead if we allow a large violation over the same period, the energy saving is of $87\, kWh$, that corresponds to the $49.2\%$.
+\begin{figure}[t!]
+	\begin{center}
+		\includegraphics[width=28pc]{figures/Temperatures_all.eps}
+	\end{center}
+	\caption{Comparison, in terms of comfort, of DPC control performance simulated over $5$ days of the testing period with $3$ different violation configurations: small, medium and large.
+With the large violation configuration, the temperature is almost always outside of the lower bound when this is tighter.
+However the maximum bound violation is less than $1\degree C$.
+With the small violation configuration, the temperature is always within the bound with very few exceptions.
+However, when it happens, the violation is lower than $0.1\degree C$.
+The medium violation configuration gives a temperature bound violation that is in the middle with respect to the small and the large ones.}
+	\label{F:comparison_all_temperature}
+\end{figure}
 \begin{table}[t!]
@@
 \end{table}
+\begin{figure}[t!]
+	\begin{center}
+		\includegraphics[width=28pc]{figures/Energy_all_EnergyPlus.eps}
+	\end{center}
+	\caption{Comparison of DPC and bang-bang controller performance over $15$ days of the testing period with different violation configurations, in terms of thermal energy saving using EnergyPlus model.
+Using bang-bang controller the house energy consumption after $15$ days is of $177kWh$.
+DPC with small, medium and large violation configurations allows an energy saving of $25.4\%$, $32.3\%$ and $49.2\%$ respectively.}
+	\label{F:comparison_all_energy_E+}
+\end{figure}

### 2017-12-24 paper/first submission/cs-challenges.tex
first change fot the revision

--- 2017-09-17 paper/cs-challenges.tex
@@
-\textit{Data Availability:} The main practical challenge for DPC lies in the availability of data for training and we require answers to questions like how much data (functional testing) is required, and how should the sampling be done?
+\begin{enumerate}
+\item \emph{\textbf{Data Availability:}} The main practical challenge for DPC lies in the availability of data for training and we require answers to questions like how much data (functional testing) is required, and how should the sampling be done?
 Therefore, the procedure for optimal experiment design, and model improvement with estimation of variance in predictions is one of the main focus of our ongoing work.
-
-\textit{Stability:} While the buildings are inherently stable, many other applications require stability guarantees.
+\item \emph{\textbf{Stability:}} While the buildings are inherently stable, many other applications require stability guarantees.
 In our ongoing work, we are working towards proving asymptotic stability to origin with DPC-RT and DPC-En by using concept of switched LTI systems.
 This will make DPC useful for systems with faster dynamics.
-
-\textit{Robustness:} Another direction of work is on handling uncertainties in the DPC framework, namely an extension to Scenario DPC to account for the disturbance uncertainty.
+\item \emph{\textbf{Robustness:}} Another direction of work is on handling uncertainties in the DPC framework, namely an extension to Scenario DPC to account for the disturbance uncertainty.
 This will help us in quantifying the robustness of DPC.
+\end{enumerate}

### 2018-01-12 paper/final/intro.tex
added physics-based modeling in the Intro

@@
 
-%Consider a multivariable dynamical system subject to external disturbances.
+Consider a multivariable dynamical system subject to external disturbances.
 The first and foremost requirement for making any decision is to obtain the underlying control-oriented predictive model of the system.
@@
 This makes MPC unsuitable for control of complex plants such as natural gas processing, oil refineries, boilers, manufacturing plants, and buildings where the user expertise, time, and associated sensor costs required to develop a model are very high \cite{Sturzenegger2016,vzavcekova2014}.
-%
-%There are two main reasons for model complexity. 
-%(1) The prime contributor is the change in model properties over time.
+
+There are two main reasons for model complexity. 
+(1) The prime contributor is the change in model properties over time.
 Even if the model is identified once via an expensive route, as the model changes with time, the system identification must be repeated to update the model.
 Thus, model adaptability or adaptive control is desirable for such systems.
-%(2) A secondary reason is the model heterogeneity which further prohibits the use of model-based control.
+(2) A secondary reason is the model heterogeneity which further prohibits the use of model-based control.
 For example, unlike the automobile or the aircraft industry, each building is designed and used in a different way.
 Therefore, this modeling process must be repeated for every new building.
-%Due to aforementioned reasons, the control strategies in such systems are often limited to fuzzy logic rules that are based on best practices.
+Due to aforementioned reasons, the control strategies in such systems are often limited to fuzzy logic rules that are based on best practices.
 
-%\paragraph{Example} As an example of such modeling complexity, consider the grey-box approach in \cite{Braun2002}.
+\paragraph{Example} \textcolor[rgb]{0.00,0.00,1.00}{As a first} example of such modeling complexity, consider the grey-box approach in \cite{Braun2002}.
 The scope of such approach is to predict the heat transfer rate to the air within the building.
@@
 
+\subsection{Modeling complexity: physics-based vs data-driven}
+\textcolor[rgb]{0,0,1}{In this section we provide a more detailed example that quantifies the difference in terms of complexity when modeling a building using physical laws and data.
+We compare the 2 approaches to show that the modeling technique used in our solution, i.e.
+data-driven modeling, eliminates several drawbacks that occur using physics-based models, such as the need to have very good knowledge of the building structure and the materials, time required to build a model, low availability of sensor measurements, etc..
 \subsubsection{physics-based modeling example}
-TO BE DONE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
+This example is taken from \cite{Sturzenegger2016}, where a building located in Allschwil, Switzerland, consisting of 6 floors, is considered.
+The building total conditioned floor area is of $6000\ m^2$ ca..
+The building is modeled as a bilinear system constructed from physical principles.
+Only the second floor has been modeled, and has been assumed to represent also the others.
+The modeling process was divided into 4 main points:
+\begin{enumerate}
+	\item Building geometry and construction data were used together with first-principles laws to derive the following linear model for the building's thermal dynamics:
+		\begin{equation}\label{E:RCZoneEq}
+			\dot x(t) = A_x x(t) + B_q q(t).
+		\end{equation}
+		This model describes the behavior of zone, wall, floor and ceiling temperatures.
+		Walls, floors and ceilings are considered divided into layers with different features.
+Therefore each zone was described with an RC network model (see Figure 3-10 in \cite{SturzeneggerTR} for more details), where the capacitances represent the states of the layers and the resistances represent the thermal resistance of the layers.
+%		\begin{figure}[t]
+%			\begin{center}
+%				\includegraphics[width=0.6\textwidth]{figures/RC_net.eps}
+%				\caption{RC model of a zone. $q$ represents the external heat fluxes}\label{F:RCnet}
+%			\end{center}
+%		\end{figure}
+		The heat exchange between two adjacent layers, i.e.
+layer "a" and layer "b", was modeled to be proportional to the temperature difference of the two layers and the corresponding thermal resistance $R$, as follows
+		\begin{equation}\label{E:RCHeatExchange}
+			\begin{aligned}
+				&C_a\dot x_a = \left(x_b(t) - x_a(t)\right)/R\\
+				&C_b\dot x_b = \left(x_a(t) - x_b(t)\right)/R,
+			\end{aligned}
+		\end{equation}
+		where $C_a$ and $C_b$ are the heat capacitances of the layers.
+This is done for each layer of each zone, obtaining the compact model \eqref{E:RCZoneEq}.
+The thermal parameters were derived from zones geometry and materials data.
+	\item External heat fluxes were modeled with a bilinear model of the fluxes which are both direct into the building and indirect through elements and zones:
+		\begin{equation}\label{E:RCfluxes}
+			q(t) = A_q x(t) + B_{q,u}u(t) + B_{q,d}d(t) + \sum_{i=1}^{n_u}{[\left(B_{q,du,i}d(t) + D_{q,xu,i}x(t)\right)u_i(t)]},
+		\end{equation}
+		with $u$ the inputs and $d$ the disturbances of the system.
+		Equation \eqref{E:RCfluxes} comes from a series of 20 equations analyzed in Section 3.3.1.3 of the Technical Report \cite{SturzeneggerTR}, that we do not report here for the sake of simplicity of the treatise.
+To model the fluxes several information have been considered in the modeling process:
 	\begin{itemize}
-	\item Introduce example
-	\item add modeling equations
-	\item quantify complexity
+		\item heat exchange associated with the building hull (except for windows) was modeled considering a conductive and a radiative part;
+		\item  heat flux to each Thermally Activated Building System (TABS, i.e.
+pipes buried in the concrete slabs of the floors carrying hot/cold water) layer was modeled as a fraction of the total supplied TABS heating and cooling heat fluxes;
+		\item heat flux through the windows was considered in three different parts: a radiation that directly acts with the elements in contact with the zone's air; a heat flux due to the conduction through the window; a heat flux due to the absorption of the solar radiation from the window;
+		\item internal gains due to occupants, appliances and lighting were modeled as a convecting heat acting in each zone;
+		\item effects due to the AHU were considered.
 	\end{itemize}
+	\item The system was discretized.
+	\item The resulting model had approximately 300 states for the only second floor.
+	Since the performance of the state estimation, that is needed to compute the optimal control inputs using MPC, heavily depends on the number of states, an approximated model with fewer states was needed.
+	To this aim, different averaged temperatures of the building facades and the zones was considered, obtaining an approximated model with 35 states.
+The obtained approximated model was then "suitable for use in MPC" (Section 3.3.1.4 of the Technical Report \cite{SturzeneggerTR}).
+\end{enumerate}
+An EnergyPlus model was built for this system, and model parameters of matrices $A_x$, $B_q$, $A_q$, $B_{q,u}$, $B_{q,d}$, $B_{q,du,i}$, $D_{q,xu,i}$ were derived using geometry and materials data from such EnergyPlus model.
+This was a choice of the authors, although in a previous work they used real geometry and materials data from the building.
+24 parameters needed to be estimated/taken form a datasheet/computed for the considered zone model.
+Although some of the parameters are in common between different zones, the others had to be found independently for each zone. }
+
+\textcolor[rgb]{0,0,1}{This approach suffers of several drawbacks:
+\begin{itemize}
+	\item in order to keep the model simple, the heat exchange between layers is modeled with a linear behavior as in \eqref{E:RCHeatExchange}. Hence all the nonlinearities are neglected;
+	\item in particular cases, as for example the ventilation energy, a linear model is not good enough to provide a good approximation: for this reason a bilinear model was considered.
+	However, also in this case, more complex nonlinearities have been neglected due to the already high complexity of the model.
+	For example, in \cite{Sturzenegger2016}, the author says "ideally one would want to formulate the optimization problem in terms of set points and operating modes that can be communicated directly to the BAS.
+	However, by limiting the model to a bilinear form, this was not possible";
+	\item due to the high modeling complexity, different geometries and different materials of the floors have been neglected supposing they are the same for each floor, although this is almost never true in real cases;
+	\item the resulting model was too complex, so it was further approximated to be "suitable for use in MPC" \cite{SturzeneggerTR};
+	\item differently from the building geometry, that could be rebuilt by hand, materials data of the building can be unavailable for different reasons (not provided, lost, changed in time).
+	Therefore they need to be estimated, hence further increasing the modeling complexity;
+	\item due to the high number of states and variables involved in the modeling framework, a lot of measurements are needed to use the model to predict system's behavior.
+	This can be cost expensive due to the amount of sensors needed.
+	Furthermore, when some measurements are not directly available from the sensors, observers are needed to provide variables estimations.
+	However, observability problems could limit the construction of the observers \cite{Dorf2011MCS}. 
+\end{itemize}
+}
 
 \subsubsection{data-driven modeling example}
-TO BE DONE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
+\textcolor[rgb]{0,0,1}{Data-drinve modeling helps eliminating all the aforementioned problems...}
+TO BE DONE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin}
 \begin{itemize}
@@
 
-\subsubsection{physics-based vs data-driven modeling example}
-TO BE DONE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
+\textcolor[rgb]{0,0,1}{\textbf{Important to add this remark wrt the last bullet point (to answer the reviewer) saying that:} while in the previous case we need to measure all the variables because they are needed to compute model evolution using equations, using machine learning we don't need all the variables (so if we cannot measure everything it is not a problem).
+This is because the machine learning algorithm creates his internal relation between variables, so also if a variable is missing, the information of the missing variable is present in the behavior of another one.
+For example: if you don't have the measurement of the energy produced by a photovoltaic panel, the information of the weather conditions will take car of it.}
+
+\paragraph{physics-based vs data-driven modeling example}
+TO BE DONE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin}
 \begin{itemize}

### 2018-01-13 paper/final/intro.tex
improved modeling with more details and started literature review

@@
 	\item The system was discretized.
-	\item The resulting model had approximately 300 states for the only second floor.
+	\item The resulting model had approximately 300 states for the only second floor, i.e. temperature of the zones, walls and floors. The output of the system were the zone temperatures.
 	Since the performance of the state estimation, that is needed to compute the optimal control inputs using MPC, heavily depends on the number of states, an approximated model with fewer states was needed.
@@
 The obtained approximated model was then "suitable for use in MPC" (Section 3.3.1.4 of the Technical Report \cite{SturzeneggerTR}).
+In particular, after the approximation, the resulting model had the following states, inputs and disturbance variables:
+	\begin{itemize}
+		\item 5 outputs: averaged room temperature for each group of zones (North, South, West, East, Center). The rooms are equipped with temperature sensors.
+		\item 18 inputs: TABS heating heat flux; TABS cooling heat flux; averaged transmitted solar heat flux for each group of zones (North, South, West, East), that was estimated using blinds position measurements; air massflow through the energy recovering mode; air massflow bypassing the energy recovering mode; air massflow through the air cooler; AHU heat coil heat flux; lighting power for the offices for each group of zones (North, South, West, East); radiator heat flux in the corner offices (North, South, West, East).
+		\item 7 disturbances: internal gains in the offices and internal gains in non-office zones, which were predicted using standard schedule; ambient temperature and solar radiation on facade (North, South, West, East), whose values were obtained through Kalman filtering, using measurements from the weather station placed on the roof of the building.
+This filtering is needed to take into account the shadowing of the neighboring buildings.
+	\end{itemize}
 \end{enumerate}
@@
 24 parameters needed to be estimated/taken form a datasheet/computed for the considered zone model.
-Although some of the parameters are in common between different zones, the others had to be found independently for each zone. }
+Although some of the parameters are in common between different zones, the others had to be found independently for each zone.
+To derive the EnergyPlus model using the available measurements and to use the same control implemented in the building, the EnergyPlus model was coupled with MATLAB using BCVTB software (see \cite{SturzeneggerTR} for more details).
+To identify the model a total of 294 MATLAB signals were sent to the EnergyPlus model and a total of 1081 EnergyPlus signals were sent back to Matlab.}
 
@@
 \subsubsection{data-driven modeling example}
-\textcolor[rgb]{0,0,1}{Data-drinve modeling helps eliminating all the aforementioned problems...}
+\textcolor[rgb]{0,0,1}{Data-driven modeling helps eliminating all the aforementioned problems...}
 TO BE DONE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin}
@@
 This is because the machine learning algorithm creates his internal relation between variables, so also if a variable is missing, the information of the missing variable is present in the behavior of another one.
-For example: if you don't have the measurement of the energy produced by a photovoltaic panel, the information of the weather conditions will take car of it.}
+For example: if you don't have the measurement of the energy produced by a photovoltaic panel, the information of the weather conditions will take care of it.}
+
+\textcolor[rgb]{0,0,1}{\textbf{Say also that:} since we don't need to model the state of the walls and floor layers, we don't have that many states as they are, so we don't need to approximate our model and we can consider the temperature of each room as output, without the need of considering an averaged temperature}
+
+\textcolor[rgb]{0,0,1}{\textbf{Advantages:}
+\begin{itemize}
+	\item we don't need to consider averaged temperatures
+	\item we don't need to filter weather data since all the noises are took into account by the machine learning algorithms
+	\item we don't need to estimate any parameter
+\end{itemize}}
 
@@
 In this paper we address this problem introducing \textit{Data Predictive Control (DPC)} to bridge the gap between Machine Learning and Predictive Control.
-
+\textcolor[rgb]{0,0,1}{
 \subsection{Related work}
-TO IMPROVE $\rightarrow$ \textcolor[rgb]{1,0,0}{Tullio}
-%In the literature, several studies deal with the data-driven control techniques using machine learning algorithms \cite{Hou2013}.
+This section is divided in 3 parts:
+\begin{enumerate}
+	\item In Section \ref{SSS:ModelBasedWork}
+	\item In Section \ref{SSS:DataDrivenWork}
+	\item In Section \ref{SSS:PreviousWork} we underline the differences between the results presented in this work with respect to the ones of our previous papers.
+\end{enumerate}
+\subsubsection{Model-based related work}\label{SSS:ModelBasedWork}
+\subsubsection{Data-driven related work}\label{SSS:DataDrivenWork}
+In the literature, several studies deal with the data-driven control techniques using machine learning algorithms \cite{Hou2013}.
 For example, Artificial Neural Networks are exploited to create data-driven models to be used as plant simulator in closed-loop with a Supervisory Model Predictive Control in \cite{Afram2017}.
@@
 More papers related to this topic can be found in the literature, but to the best of the authors' knowledge none of them address the problem of including data-driven state models in the optimal predictive control loop, and hence allow to set up a MPC-like optimal control problem, as we do in DPC.
+\subsubsection{Previous work}\label{SSS:PreviousWork}
+In the following, the differences with respect to our previous papers are illustrated.
+\begin{itemize}
+	\item In \cite{Behl2016} we	proposed an MPC-like control problem considering regression trees and ensemble methods algorithms for the Demand-Response problem.
+However, the proposed models allowed to optimally control the system considering only one-step lookahead prediction.
+Hence, it was not possible to use such approach to control the system make a prediction over an horizon of arbitrary length.
+	\item This problem was addressed in \cite{Jain2016}, where the regression trees algorithm was modified to allow the tree to provide multiple-output.
+	Each output of the tree was then setup to provide the prediction of the system's behavior over different steps of a predictive horizon.
+	This allowed us to setup an MPC-like control problem considering an arbitrary horizon.
+	In this paper only single regression trees were considered, while ensemble methods were not took into account.
+	However, modeling accuracy using single trees is strongly subject to the problems of the overfitting and high variance.
+	For this reason, such approach has the advantage to be extremely simple from the complexity point of view, but the application is limited to specific kind of systems.
+	\item In this work, we extend preliminary results gave in the conference paper \cite{JainCDC2017}, where an alternative approach is proposed to improve the system identification accuracy and robustness of the previous approach.
+More precisely, instead of considering a single tree with multiple output, we considered multiple trees with single output.
+Each tree is used to provide the prediction of the system's behavior over different steps of the horizon.
+With a small increase in complexity, we setup an MPC-like control problem with arbitrary horizon, that provides better results with respect to our previous work.
+In this paper we further extend this approach to one of the ensemble methods, i.e.
+random forests.
+\end{itemize}
+}
 

### 2018-01-13 paper/final/intro.tex
small edits

@@
 \begin{enumerate}
-	\item In Section \ref{SSS:ModelBasedWork}
-	\item In Section \ref{SSS:DataDrivenWork}
-	\item In Section \ref{SSS:PreviousWork} we underline the differences between the results presented in this work with respect to the ones of our previous papers.
+	\item In Section \ref{SSS:ModelBasedWork} we provide a literature review of the approaches that make use of model-based techniques for energy optimization.
+	\item In Section \ref{SSS:DataDrivenWork} we report the main works that use of data-driven methodologies for energy optimization and point out the difference with respect to our contribution.
+	\item In Section \ref{SSS:PreviousWork} we underline the differences between the results presented in this work and the ones of our previous papers.
 \end{enumerate}
 \subsubsection{Model-based related work}\label{SSS:ModelBasedWork}
-\subsubsection{Data-driven related work}\label{SSS:DataDrivenWork}
+\subsubsection{Data-driven related work}\label{SSS:DataDrivenWork}}
 In the literature, several studies deal with the data-driven control techniques using machine learning algorithms \cite{Hou2013}.
@@
 More papers related to this topic can be found in the literature, but to the best of the authors' knowledge none of them address the problem of including data-driven state models in the optimal predictive control loop, and hence allow to set up a MPC-like optimal control problem, as we do in DPC.
+
+\textcolor[rgb]{0,0,1}{
 \subsubsection{Previous work}\label{SSS:PreviousWork}
@@
 However, the proposed models allowed to optimally control the system considering only one-step lookahead prediction.
-Hence, it was not possible to use such approach to control the system make a prediction over an horizon of arbitrary length.
+	Hence, it was not possible to use such approach to control the system considering a prediction over an horizon of arbitrary length.
 	\item This problem was addressed in \cite{Jain2016}, where the regression trees algorithm was modified to allow the tree to provide multiple-output.
-	Each output of the tree was then setup to provide the prediction of the system's behavior over different steps of a predictive horizon.
+	Each output of the tree was then setup to provide the prediction of the system's behavior over different steps of the predictive horizon.
 	This allowed us to setup an MPC-like control problem considering an arbitrary horizon.
-	In this paper only single regression trees were considered, while ensemble methods were not took into account.
-	However, modeling accuracy using single trees is strongly subject to the problems of the overfitting and high variance.
-	For this reason, such approach has the advantage to be extremely simple from the complexity point of view, but the application is limited to specific kind of systems.
-	\item In this work, we extend preliminary results gave in the conference paper \cite{JainCDC2017}, where an alternative approach is proposed to improve the system identification accuracy and robustness of the previous approach.
-More precisely, instead of considering a single tree with multiple output, we considered multiple trees with single output.
+	However, in this paper only single regression trees were considered, while ensemble methods were not took into account.
+	Modeling accuracy using single trees is strongly subject to the problems of data overfitting and high variance.
+	Such approach has the advantage to be extremely simple from the complexity point of view, but its application range is limited to specific kind of systems.
+	\item In this work, we extend preliminary results gave in the conference paper \cite{JainCDC2017}.
+	An alternative approach is proposed to improve the system's identification accuracy and robustness of the previous approach.
+	More precisely, instead of considering a single tree with multiple output, we consider multiple trees with single output.
 	Each tree is used to provide the prediction of the system's behavior over different steps of the horizon.
 With a small increase in complexity, we setup an MPC-like control problem with arbitrary horizon, that provides better results with respect to our previous work.
-In this paper we further extend this approach to one of the ensemble methods, i.e.
-random forests.
+In this paper we further extend this approach to one of the ensemble methods: the random forests.
 \end{itemize}

### 2018-01-15 paper/final/intro.tex
intro edits 1

@@
 
-Machine learning and control theory are two foundational but disjoint communities.
-Machine learning requires data to produce models, and control systems require models to provide stability and performance guarantees to plant operations.
-Machine learning is widely used for regression or classification, but thus far data-driven models have not been suitable for closed-loop control of physical plants.
-The challenge now, with using data-driven approaches, is to close the loop for real-time control and decision making.
+Control-oriented predictive models of an energy system's dynamics and energy consumption, are needed for understanding and improving the overall energy efficiency and operating costs.
+With a reasonably accurate forecast of future weather and building operating conditions, dynamical models can be used to predict the energy needs of the building over a prediction horizon, as is the case with Model Predictive Control (MPC) \cite{Sturzenegger2016}.
+However, a major challenge with MPC is in accurately modeling the dynamics of the underlying physical system.
+The task is much more complicated and time consuming in case of a large building and often times, it can be even more complex and involved than the controller design itself.
+After several years of work on using first principles based models for peak power reduction, and energy optimization for buildings, multiple authors~\cite{Sturzenegger2016, vzavcekova2014} have concluded that the biggest hurdle to mass adoption of intelligent building control is the cost and effort required to capture accurate dynamical models of the buildings.
+The user expertise, time, and associated sensor costs required to develop a model of a single building is very high.
+Thus, the payback period for the upfront hardware and software installation is expected be be too high making MPC an uneconomical choice for energy management.
 
-\subsection{Motivation}
-TO IMPROVE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
+Why is physics-based modeling hard for complex systems like buildings or what are the reasons of modeling complexities?
 
-Consider a multivariable dynamical system subject to external disturbances.
-The first and foremost requirement for making any decision is to obtain the underlying control-oriented predictive model of the system.
-With a reasonable forecast of the external disturbances, these models should predict the state of the system in the future and thus Model Predictive Control (MPC) can act preemptively to provide a desired system behavior while optimizing a desired performance.
-In particular, MPC has been proven to be very powerful for multivariable systems in the presence of input and output constraints, and forecast of the disturbances.
-The caveat is that MPC requires a reasonably accurate physical representation of the system.
-This makes MPC unsuitable for control of complex plants such as natural gas processing, oil refineries, boilers, manufacturing plants, and buildings where the user expertise, time, and associated sensor costs required to develop a model are very high \cite{Sturzenegger2016,vzavcekova2014}.
+\begin{enumerate}
+	\item \textbf{Model capture} - a building modeling domain expert typically uses a software tool to create the geometry of a building from the building design and equipment layout plans, add detailed information about material properties, about equipment and operational schedules.
+There is always a gap between the modeled and the real building and the domain expert must then manually tune the model to match the measured data from the building~\cite{New2012}.
+Moreover, the modeling process also varies from building to building with the construction and types of installed equipment.
+Another major downside with physics-based modeling is that enough data is not easily available and guesses for parameter values have to be made, which also requires expert know how.
 
-There are two main reasons for model complexity. 
-(1) The prime contributor is the change in model properties over time.
-Even if the model is identified once via an expensive route, as the model changes with time, the system identification must be repeated to update the model.
+\item \textbf{Change in model properties} over time - even if the model is identified once via an expensive route as in \cite{Sturzenegger2016}, as the model changes with time, the system identification must be repeated to update the model.
 Thus, model adaptability or adaptive control is desirable for such systems.
-(2) A secondary reason is the model heterogeneity which further prohibits the use of model-based control.
+
+\item \textbf{Model heterogeneity} further prohibits the use of model-based control.
 For example, unlike the automobile or the aircraft industry, each building is designed and used in a different way.
 Therefore, this modeling process must be repeated for every new building.
-Due to aforementioned reasons, the control strategies in such systems are often limited to fuzzy logic rules that are based on best practices.
+\end{enumerate}
 
-\paragraph{Example} \textcolor[rgb]{0.00,0.00,1.00}{As a first} example of such modeling complexity, consider the grey-box approach in \cite{Braun2002}.
+Due to aforementioned reasons, the control strategies in such systems are often limited to fixed, sometimes ad-hoc, rules that are based on best practices. 
+The alternative is to use black-box, or completely data-driven modeling approaches, to obtain a realization of the system's input-output behavior. 
+The primary advantage of using data-driven methods is that it has the potential to eliminate the time and effort required to build white and grey box building models. 
+Listening to real-time data, from existing systems and interfaces, is far cheaper than unleashing hoards of on-site engineers to physically measure and model the building.
+Improved building technology and better sensing is fundamentally redefining the opportunities around smart buildings.
+Unprecedented amounts of data from millions of smart meters and thermostats installed in recent years has opened the door for systems engineers and data scientists to analyze and use the insights that data can provide, about the dynamics and power consumption patterns of these systems.
+
+The key question now is: can we employ data-driven techniques to reduce the cost of modeling, and still exploit the benefits that MPC has to offer?
+We therefore look for automatic data-driven approaches to control that are also adaptive, scalable and interpretable.
+We solve this problem by bridging the gap between Machine Learning and Predictive Control. In particular, in this paper, we present a method based on Random Forests.
+
+%\paragraph{Example} \textcolor[rgb]{0.00,0.00,1.00}{As a first} example of such modeling complexity, consider the grey-box approach in \cite{Braun2002}.
 The scope of such approach is to predict the heat transfer rate to the air within the building.
@@
 
-\subsection{Modeling complexity: physics-based vs data-driven}
-\textcolor[rgb]{0,0,1}{In this section we provide a more detailed example that quantifies the difference in terms of complexity when modeling a building using physical laws and data.
-We compare the 2 approaches to show that the modeling technique used in our solution, i.e.
-data-driven modeling, eliminates several drawbacks that occur using physics-based models, such as the need to have very good knowledge of the building structure and the materials, time required to build a model, low availability of sensor measurements, etc..
-\subsubsection{physics-based modeling example}
-This example is taken from \cite{Sturzenegger2016}, where a building located in Allschwil, Switzerland, consisting of 6 floors, is considered.
-The building total conditioned floor area is of $6000\ m^2$ ca..
-The building is modeled as a bilinear system constructed from physical principles.
-Only the second floor has been modeled, and has been assumed to represent also the others.
-The modeling process was divided into 4 main points:
+\subsection{Example for comparison of modeling complexity: physics-based vs data-driven}
+In this section, we provide a detailed example that describes the difference in terms of complexity when we model a building using physical laws and machine learning.
+We show that the data-driven modeling eliminates several drawbacks that occur using physics-based models such as the need to have very good knowledge of the building structure and the material properties, time required to build a model, and limited availability of sensor measurements.
+
+The building under consideration is taken from \cite{Sturzenegger2016}.
+It is located in Allschwil, Switzerland.
+It has 6 floors of which only the second floor is modeled, and is assumed to represent all other floors.
+The total air conditioned floor area is $6000\ \mathrm{m^2}$ ca.
+
+\subsubsection{Physics-based modeling}
+\label{SSS:physics-based}
+We describe the physics-based approach based on an RC network outlined in \cite{Sturzenegger2016}.
+The building is modeled as a bilinear system constructed from physical principles, the modeling process consisting of three steps:
 \begin{enumerate}
-	\item Building geometry and construction data were used together with first-principles laws to derive the following linear model for the building's thermal dynamics:
+	\item Building geometry and construction data are used together with first-principles to derive the following linear model for the building's thermal dynamics:
 		\begin{equation}\label{E:RCZoneEq}
@@
 		\end{equation}
-		This model describes the behavior of zone, wall, floor and ceiling temperatures.
-		Walls, floors and ceilings are considered divided into layers with different features.
-Therefore each zone was described with an RC network model (see Figure 3-10 in \cite{SturzeneggerTR} for more details), where the capacitances represent the states of the layers and the resistances represent the thermal resistance of the layers.
+		This model describes the behavior of the zone, wall, floor and ceiling temperatures.
+		Walls, floors and ceilings are considered as divided into layers with different features.
+Therefore, each zone was described with an RC network model (see Figure 3-10 in \cite{SturzeneggerTR}), where the capacitances represent the states of the layers and the resistances represent the thermal resistance of the layers.
 %		\begin{figure}[t]
@@
 		The heat exchange between two adjacent layers, i.e.
-layer "a" and layer "b", was modeled to be proportional to the temperature difference of the two layers and the corresponding thermal resistance $R$, as follows
+layer ``a" and layer ``b", is modeled to be proportional to the temperature difference of the two layers and the corresponding thermal resistance $R$ given by
 		\begin{equation}\label{E:RCHeatExchange}
 			\begin{aligned}
-				&C_a\dot x_a = \left(x_b(t) - x_a(t)\right)/R\\
-				&C_b\dot x_b = \left(x_a(t) - x_b(t)\right)/R,
+				&C_a\dot x_a = \frac{x_b(t) - x_a(t)}{R}, \ \text{and}\\
+				&C_b\dot x_b =  \frac{x_a(t) - x_b(t)}{R},
 			\end{aligned}
@@
 		where $C_a$ and $C_b$ are the heat capacitances of the layers.
-This is done for each layer of each zone, obtaining the compact model \eqref{E:RCZoneEq}.
-The thermal parameters were derived from zones geometry and materials data.
-	\item External heat fluxes were modeled with a bilinear model of the fluxes which are both direct into the building and indirect through elements and zones:
+This is done for each layer of each zone, obtaining the compact representation given in \eqref{E:RCZoneEq}.
+The thermal parameters are derived from zones geometry and material properties.
+	\item External heat fluxes, modeled as a bilinear model, affect the building directly as well as indirectly through elements and zones:
 		\begin{equation}\label{E:RCfluxes}
@@
 		\end{equation}
-		with $u$ the inputs and $d$ the disturbances of the system.
-		Equation \eqref{E:RCfluxes} comes from a series of 20 equations analyzed in Section 3.3.1.3 of the Technical Report \cite{SturzeneggerTR}, that we do not report here for the sake of simplicity of the treatise.
-To model the fluxes several information have been considered in the modeling process:
-	\begin{itemize}
-		\item heat exchange associated with the building hull (except for windows) was modeled considering a conductive and a radiative part;
-		\item  heat flux to each Thermally Activated Building System (TABS, i.e.
-pipes buried in the concrete slabs of the floors carrying hot/cold water) layer was modeled as a fraction of the total supplied TABS heating and cooling heat fluxes;
-		\item heat flux through the windows was considered in three different parts: a radiation that directly acts with the elements in contact with the zone's air; a heat flux due to the conduction through the window; a heat flux due to the absorption of the solar radiation from the window;
-		\item internal gains due to occupants, appliances and lighting were modeled as a convecting heat acting in each zone;
-		\item effects due to the AHU were considered.
-	\end{itemize}
-	\item The system was discretized.
-	\item The resulting model had approximately 300 states for the only second floor, i.e. temperature of the zones, walls and floors. The output of the system were the zone temperatures.
-	Since the performance of the state estimation, that is needed to compute the optimal control inputs using MPC, heavily depends on the number of states, an approximated model with fewer states was needed.
-	To this aim, different averaged temperatures of the building facades and the zones was considered, obtaining an approximated model with 35 states.
-The obtained approximated model was then "suitable for use in MPC" (Section 3.3.1.4 of the Technical Report \cite{SturzeneggerTR}).
-In particular, after the approximation, the resulting model had the following states, inputs and disturbance variables:
+		where $u$ are the inputs and $d$ the disturbances of the system.
+%		Equation \eqref{E:RCfluxes} comes from a series of 20 equations analyzed in Section 3.3.1.3 of the Technical Report \cite{SturzeneggerTR}, that we do not report here for the sake of simplicity. 
+Equation \label{E:RCfluxes} for the heat flux is obtained by modeling
+	\begin{enumerate}
+		\item heat exchange associated with the building hull (except for windows), both conductive and radiative part,
+		\item  heat flux to each thermally activated building system (TABS), i.e. pipes buried in the concrete slabs of the floors carrying hot/cold water,
+		\item heat flux through the windows in three different parts: radiation due to elements directly in contact with the zone air, conduction through the window, and absorption of the solar radiation through the window,
+		\item convection due to internal gains due to occupants, appliances and lighting, and
+		\item the effects of the AHU.
+	\end{enumerate}
+	\item The resulting system \eqref{E:RCZoneEq} is discretized to obtained a model suitable for finite horizon MPC.
+This model has approximately 300 states which include temperature of the zones, walls and floors on the second floor.
+The output of the system are the zone temperatures.
+	Since the performance of the state estimation, that is needed to compute the optimal control inputs using MPC, depends on the number of states, an approximated model with fewer states is needed.
+	To this aim, different averaged temperatures of the building facades and the zones are considered, obtaining an approximated model with 35 states.
+This  approximate model is then considered ``suitable for MPC" (see Section 3.3.1.4 in \cite{SturzeneggerTR}).
+\end{enumerate}
+
+	In summary, the approximate model has the following states, inputs and disturbance variables:
 	\begin{itemize}
 		\item 5 outputs: averaged room temperature for each group of zones (North, South, West, East, Center). The rooms are equipped with temperature sensors.
-		\item 18 inputs: TABS heating heat flux; TABS cooling heat flux; averaged transmitted solar heat flux for each group of zones (North, South, West, East), that was estimated using blinds position measurements; air massflow through the energy recovering mode; air massflow bypassing the energy recovering mode; air massflow through the air cooler; AHU heat coil heat flux; lighting power for the offices for each group of zones (North, South, West, East); radiator heat flux in the corner offices (North, South, West, East).
-		\item 7 disturbances: internal gains in the offices and internal gains in non-office zones, which were predicted using standard schedule; ambient temperature and solar radiation on facade (North, South, West, East), whose values were obtained through Kalman filtering, using measurements from the weather station placed on the roof of the building.
+		\item 18 inputs: TABS heating heat flux, TABS cooling heat flux, averaged transmitted solar heat flux for each group of zones (North, South, West, East) that is estimated using blinds position measurements, air massflow through the energy recovering mode, air massflow bypassing the energy recovering mode, air massflow through the air cooler, AHU heat coil heat flux, lighting power for the offices for each group of zones (North, South, West, East), and radiator heat flux in the corner offices (North, South, West, East).
+		\item 7 disturbances: internal gains in the offices and internal gains in non-office zones which are predicted using a standard schedule, ambient temperature and solar radiation on facade (North, South, West, East) whose values were obtained through Kalman filtering using measurements from the weather station placed on the roof of the building.
 This filtering is needed to take into account the shadowing of the neighboring buildings.
 	\end{itemize}
+An EnergyPlus model is built for this system, and model parameters of matrices $A_x$, $B_q$, $A_q$, $B_{q,u}$, $B_{q,d}$, $B_{q,du,i}$, $D_{q,xu,i}$ are derived using geometry and materials data from the EnergyPlus model.
+This was a choice of the authors; real geometry and materials data can also be used to estimate the model parameters. 
+For this particular building, 24 parameters are estimated/taken form a datasheet/computed for the considered zone model.
+Although some of the parameters are in common between different zones, the others are found independently for each zone.
+To derive the EnergyPlus model using the available measurements and to use the same control implemented in the building, the EnergyPlus model is coupled with MATLAB using BCVTB \cite{Wetter2015}. 
+%To identify the model a total of 294 MATLAB signals were sent to the EnergyPlus model and a total of 1081 EnergyPlus signals were sent back to Matlab.
+\\
+
+The physics-based approach, although detailed and accurate (in some cases), is cost and time prohibitive.
+The long and complex modeling procedure that is unique to every building poses a challenge in applying MPC to large scale buildings.
+
+\subsubsection{Data-driven modeling}
+The data-driven approach, on the other hand, ignores the physical details of the building operation.
+Instead, we use machine learning to learn black-box models where our objective is to find the hyperparameters (for a given structure like Random Forests) of a model that best explains the input-output relationship.
+\\
+
+The goal with the data-driven modeling is to learn a function map
+\begin{equation}
+y(t) = f(u(t),\dots,u(t-\delta_u), d(t),\dots, d(t-\delta_d)),
+\end{equation}
+where \(y\) represents the outputs, \(u\) the inputs and \(d\) the disturbances as defined earlier, and \(\delta\) is a lag variable.
+If \(y\) is the average room temperature for one of the zones, a nonlinear function \(f\) is more appropriate, for example \(f\) may belong to the class of Neural Networks or Random Forests.
+\\
+
+The data-driven modeling in the case of buildings has several advantages over the physical counterpart.
+\begin{enumerate}
+
+	\item \textbf{Domain expertise:} This procedure eliminates the need of a building expert, simplifies the modeling complexities described in Section \ref{SSS:physics-based} and ultimately reduces the cost of modeling a building.
+Further, for a new building, given the historical data from the building, the same process can be repeated to identify a control-oriented model for MPC.
+
+	\item \textbf{Linearity assumption:} During physical modeling in \eqref{E:RCHeatExchange}, in order to keep the model simple, the heat exchange between layers is modeled with a linear behavior. 
+	Hence all the nonlinearities were neglected.
+	In the case of heat flux in \eqref{E:RCfluxes}, a linear model is not good enough to provide a good approximation. For this reason a bilinear model was considered.
+	However, more complex nonlinearities have been neglected due to the already high complexity of the model \cite{Sturzenegger2016}. 
+	In data-driven modeling, we can relax these linearity assumptions as nonlinear functions can be learned rather efficiently and accurately.
+
+	\item \textbf{Other assumptions:} Due to the high modeling complexity, physical modeling neglects different geometries and different materials of the floors assuming they are the same for each floor which is never true in reality.
+	Data-driven modeling depends on the sensor data directly. 
+	Hence, such assumptions never arise.
+
+%	\item differently from the building geometry, that could be rebuilt by hand, materials data of the building can be unavailable for different reasons (not provided, lost, changed in time).
+
+%	Therefore they need to be estimated, hence further increasing the modeling complexity;
+
+	\item \textbf{Sensor cost:} Due to large number of states and variables in the physical modeling, a lot of measurements are needed to use the model to predict the system's behavior.
+	This can be cost expensive due to the quantity of sensors needed.
+	Moreover, when some measurements are not directly available from the sensors, observers are needed for state estimations.
+	However, observability problems can limit the construction of the observers \cite{Dorf2011MCS}. 
+	On the other hand, data-driven modeling relies only on direct measurements from the sensors such as thermostats and multimeters reducing the need for new installations.
 \end{enumerate}
-An EnergyPlus model was built for this system, and model parameters of matrices $A_x$, $B_q$, $A_q$, $B_{q,u}$, $B_{q,d}$, $B_{q,du,i}$, $D_{q,xu,i}$ were derived using geometry and materials data from such EnergyPlus model.
-This was a choice of the authors, although in a previous work they used real geometry and materials data from the building.
-24 parameters needed to be estimated/taken form a datasheet/computed for the considered zone model.
-Although some of the parameters are in common between different zones, the others had to be found independently for each zone.
-To derive the EnergyPlus model using the available measurements and to use the same control implemented in the building, the EnergyPlus model was coupled with MATLAB using BCVTB software (see \cite{SturzeneggerTR} for more details).
-To identify the model a total of 294 MATLAB signals were sent to the EnergyPlus model and a total of 1081 EnergyPlus signals were sent back to Matlab.}
-
-\textcolor[rgb]{0,0,1}{This approach suffers of several drawbacks:
-\begin{itemize}
-	\item in order to keep the model simple, the heat exchange between layers is modeled with a linear behavior as in \eqref{E:RCHeatExchange}. Hence all the nonlinearities are neglected;
-	\item in particular cases, as for example the ventilation energy, a linear model is not good enough to provide a good approximation: for this reason a bilinear model was considered.
-	However, also in this case, more complex nonlinearities have been neglected due to the already high complexity of the model.
-	For example, in \cite{Sturzenegger2016}, the author says "ideally one would want to formulate the optimization problem in terms of set points and operating modes that can be communicated directly to the BAS.
-	However, by limiting the model to a bilinear form, this was not possible";
-	\item due to the high modeling complexity, different geometries and different materials of the floors have been neglected supposing they are the same for each floor, although this is almost never true in real cases;
-	\item the resulting model was too complex, so it was further approximated to be "suitable for use in MPC" \cite{SturzeneggerTR};
-	\item differently from the building geometry, that could be rebuilt by hand, materials data of the building can be unavailable for different reasons (not provided, lost, changed in time).
-	Therefore they need to be estimated, hence further increasing the modeling complexity;
-	\item due to the high number of states and variables involved in the modeling framework, a lot of measurements are needed to use the model to predict system's behavior.
-	This can be cost expensive due to the amount of sensors needed.
-	Furthermore, when some measurements are not directly available from the sensors, observers are needed to provide variables estimations.
-	However, observability problems could limit the construction of the observers \cite{Dorf2011MCS}. 
-\end{itemize}
-}
-
-\subsubsection{data-driven modeling example}
-\textcolor[rgb]{0,0,1}{Data-driven modeling helps eliminating all the aforementioned problems...}
-TO BE DONE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin}
-\begin{itemize}
-	\item show procedure
-	\item quantify complexity
-\end{itemize}
 
@@
 
-\paragraph{physics-based vs data-driven modeling example}
-TO BE DONE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin}
-\begin{itemize}
-	\item compare results
-\end{itemize}
-
-\paragraph{Objective} The question now is, can we employ data-driven techniques to reduce the cost of modeling and still exploit the benefits that MPC has to offer?
-Can we build automatic and data-driven approaches for control purposes that are also adaptive, scalable and interpretable?
-In this paper we address this problem introducing \textit{Data Predictive Control (DPC)} to bridge the gap between Machine Learning and Predictive Control.
 \textcolor[rgb]{0,0,1}{
@@
 \subsection{Main contribution and paper organization}
-TO CLARIFY $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
-\begin{itemize}
-	\item Add more details
-	\item Emphasize difference between Section 4 and Section 5
-	\item Cite appendix
-\end{itemize}
-TO ADD $\rightarrow$ \textcolor[rgb]{1,0,0}{Tullio}
-\begin{itemize}
-	\item Add references for weather forecast accuracy
-	\item Quantify accuracy for short-term prediction for weather forcast to justify the assumption of perfect knowledge for the disturbance
-\end{itemize}
-TO ADD $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
-\begin{itemize}
-	\item Add subsection in Section 5 with simulation of DPC considering disturbance with added noise and show that we still have good performance
-\end{itemize}
-%In our previous work \cite{Behl2016,Jain2016,JainACC2017,JainCDC2017}, we introduced the concept of DPC for receding horizon control.
+%TO CLARIFY $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
+%\begin{itemize}
+%	\item Add more details
+%	\item Emphasize difference between Section 4 and Section 5
+%	\item Cite appendix
+%\end{itemize}
+%TO ADD $\rightarrow$ \textcolor[rgb]{1,0,0}{Tullio}
+%\begin{itemize}
+%	\item Add references for weather forecast accuracy
+%	\item Quantify accuracy for short-term prediction for weather forcast to justify the assumption of perfect knowledge for the disturbance
+%\end{itemize}
+%TO ADD $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
+%\begin{itemize}
+%	\item Add subsection in Section 5 with simulation of DPC considering disturbance with added noise and show that we still have good performance
+%\end{itemize}
+In our previous work \cite{Behl2016,Jain2016,JainACC2017,JainCDC2017}, we introduced the concept of DPC for receding horizon control.
 In this paper, we extend these results providing the following contributions.
-%\begin{enumerate}
-%	\item In Section \ref{S:dpc}, we formally present the two following control techniques:
-%	\begin{enumerate}
-%		\item DPC with regression trees, and
-%		\item DPC with random forests.
-%	\end{enumerate}
-%	\item In Section \ref{S:proof}, we demonstrate the strength of DPC for receding horizon control via one-to-one comparison against a benchmark MPC controller using a bilinear building model whose parameters were identified using experiments on a building in Switzerland.
+\begin{enumerate}
+	\item In Section \ref{S:dpc}, we formally present the two following control techniques:
+	\begin{enumerate}
+		\item DPC with regression trees, and
+		\item DPC with random forests.
+	\end{enumerate}
+	\item In Section \ref{S:proof}, we demonstrate the strength of DPC for receding horizon control via one-to-one comparison against a benchmark MPC controller using a bilinear building model whose parameters were identified using experiments on a building in Switzerland.
 We show that DPC captures 70\% variance in MPC and offers a comparable performance.
-%	\item Section \ref{S:casestudy} describes a practical application of DPC for Demand Response, where we apply DPC to a 6 story 22 zone building model in EnergyPlus \cite{Crawley2001} for which model-based control is not economical and practical due to extreme complexity.
+	\item Section \ref{S:casestudy} describes a practical application of DPC for Demand Response, where we apply DPC to a 6 story 22 zone building model in EnergyPlus \cite{Crawley2001} for which model-based control is not economical and practical due to extreme complexity.
 We show scalability and efficiency of DPC in providing financial incentives to the end-customers bypassing the need for high fidelity models.
 We observe that DPC provides the desired power reduction with an average error of 3\%.
-%	\item In Section \ref{S:realCaseStudy}, we implement DPC on the real data from an off-grid house located in L'Aquila (Italy), to find the optimal ON/OFF scheduling for the heating system in order to save energy while guaranteeing thermal comfort for the occupants.
+	\item In Section \ref{S:realCaseStudy}, we implement DPC on the real data from an off-grid house located in L'Aquila (Italy), to find the optimal ON/OFF scheduling for the heating system in order to save energy while guaranteeing thermal comfort for the occupants.
 We quantify the total amount of energy saved, with respect to the classical bang-bang controller widely used in houses for temperature control, using an EnergyPlus model built specifically for the house.
 We show we can perform an energy saving that goes from $25.4\%$, if we guarantee thermal comfort, to $49.2\%$, if we allow very little discomfort in terms of rooms' temperature.
-%\end{enumerate}
+\end{enumerate}

### 2018-01-15 paper/final/intro.tex
edits 2

@@
 Why is physics-based modeling hard for complex systems like buildings or what are the reasons of modeling complexities?
-
 \begin{enumerate}
@@
 The thermal parameters are derived from zones geometry and material properties.
-	\item External heat fluxes, modeled as a bilinear model, affect the building directly as well as indirectly through elements and zones:
+	\item External heat fluxes, modeled as a bilinear model, affect the building directly as well as indirectly through zones:
 		\begin{equation}\label{E:RCfluxes}
@@
 	\end{enumerate}
-	\item The resulting system \eqref{E:RCZoneEq} is discretized to obtained a model suitable for finite horizon MPC.
+	\item The resulting system \eqref{E:RCZoneEq} is discretized with sampling time of $15\ \mathrm{min}$.
 This model has approximately 300 states which include temperature of the zones, walls and floors on the second floor.
@@
 
-	In summary, the approximate model has the following states, inputs and disturbance variables:
+	In summary, the approximate model has the following state, input and disturbance variables:
 	\begin{itemize}
@@
 This was a choice of the authors; real geometry and materials data can also be used to estimate the model parameters. 
-For this particular building, 24 parameters are estimated/taken form a datasheet/computed for the considered zone model.
+For this particular building, 24 parameters are estimated/taken form a datasheet computed for the considered zone model.
 Although some of the parameters are in common between different zones, the others are found independently for each zone.
@@
 %To identify the model a total of 294 MATLAB signals were sent to the EnergyPlus model and a total of 1081 EnergyPlus signals were sent back to Matlab.
-\\
 
@@
 \subsubsection{Data-driven modeling}
-The data-driven approach, on the other hand, ignores the physical details of the building operation.
+The data-driven approach we propose, on the other hand, ignores the physical details of the building operation.
 Instead, we use machine learning to learn black-box models where our objective is to find the hyperparameters (for a given structure like Random Forests) of a model that best explains the input-output relationship.
-\\
 
@@
 If \(y\) is the average room temperature for one of the zones, a nonlinear function \(f\) is more appropriate, for example \(f\) may belong to the class of Neural Networks or Random Forests.
-\\
 
@@
 
-	\item \textbf{Domain expertise:} This procedure eliminates the need of a building expert, simplifies the modeling complexities described in Section \ref{SSS:physics-based} and ultimately reduces the cost of modeling a building.
+	\item \textbf{Domain expertise:} 
+	This procedure eliminates the need of a building expert, simplifies the modeling complexities described in Section \ref{SSS:physics-based} which ultimately reduces the cost of modeling a building.
 Further, for a new building, given the historical data from the building, the same process can be repeated to identify a control-oriented model for MPC.
 
-	\item \textbf{Linearity assumption:} During physical modeling in \eqref{E:RCHeatExchange}, in order to keep the model simple, the heat exchange between layers is modeled with a linear behavior. 
+	\item \textbf{Sensor cost:}
+	Due to large number of states and variables in the physical modeling, a lot of measurements are needed to use the model to predict the system's behavior.
+	This can be expensive due to necessary new sensor installations.
+	Moreover, when some measurements are not directly available from the sensors, observers are needed for state estimations.
+	However, observability problems can limit the construction of the observers \cite{Dorf2011MCS}. 
+	On the other hand, data-driven modeling obviates the need to model the state of the walls and floor layers so we do not have as many states.
+	It relies only on direct measurements from the sensors such as thermostats and multimeters reducing the need for new installations and hence the cost of modeling.
+	Further, when some of the states are not directly measurable, often adding proxy variables suffices. 
+	For example, if we do not have the measurement of the energy produced by a photovoltaic panel, the weather conditions will account for it.
+
+	\item \textbf{Linearity assumption:} 
+	During physical modeling in \eqref{E:RCHeatExchange}, in order to keep the model simple, the heat exchange between layers is modeled with a linear behavior. 
 	Hence all the nonlinearities were neglected.
@@
 
-	\item \textbf{Other assumptions:} Due to the high modeling complexity, physical modeling neglects different geometries and different materials of the floors assuming they are the same for each floor which is never true in reality.
+	\item \textbf{Other assumptions:} 
+	Due to the high modeling complexity, physical modeling neglects different geometries and different materials of the floors assuming they are the same for each floor which is never true in reality.
+	In many cases, the details of construction layout and equipments are not even available so many parameters have to be guessed.
 	Data-driven modeling depends on the sensor data directly. 
@@
 
-%	\item differently from the building geometry, that could be rebuilt by hand, materials data of the building can be unavailable for different reasons (not provided, lost, changed in time).
-
-%	Therefore they need to be estimated, hence further increasing the modeling complexity;
-
-	\item \textbf{Sensor cost:} Due to large number of states and variables in the physical modeling, a lot of measurements are needed to use the model to predict the system's behavior.
-	This can be cost expensive due to the quantity of sensors needed.
-	Moreover, when some measurements are not directly available from the sensors, observers are needed for state estimations.
-	However, observability problems can limit the construction of the observers \cite{Dorf2011MCS}. 
-	On the other hand, data-driven modeling relies only on direct measurements from the sensors such as thermostats and multimeters reducing the need for new installations.
 \end{enumerate}
 
-\textcolor[rgb]{0,0,1}{\textbf{Important to add this remark wrt the last bullet point (to answer the reviewer) saying that:} while in the previous case we need to measure all the variables because they are needed to compute model evolution using equations, using machine learning we don't need all the variables (so if we cannot measure everything it is not a problem).
-This is because the machine learning algorithm creates his internal relation between variables, so also if a variable is missing, the information of the missing variable is present in the behavior of another one.
-For example: if you don't have the measurement of the energy produced by a photovoltaic panel, the information of the weather conditions will take care of it.}
+To summarize, the data-driven modeling simplifies the modeling of buildings to a large extent, reduces the overall cost and time investment while avoiding the linear model assumptions that we make in standard physical modeling procedure.
+Unlike physical modeling, the parameters of the model (or hyperparamters) are automatically determined by the learning algorithm, the best settings determined by cross validation.
 
-\textcolor[rgb]{0,0,1}{\textbf{Say also that:} since we don't need to model the state of the walls and floor layers, we don't have that many states as they are, so we don't need to approximate our model and we can consider the temperature of each room as output, without the need of considering an averaged temperature}
-
-\textcolor[rgb]{0,0,1}{\textbf{Advantages:}
-\begin{itemize}
-	\item we don't need to consider averaged temperatures
-	\item we don't need to filter weather data since all the noises are took into account by the machine learning algorithms
-	\item we don't need to estimate any parameter
-\end{itemize}}
+However, the challenge lies in using such models for optimal control, which is exactly the focus of this paper.
+We demonstrate how the training algorithm for Random Forests can be modified to develop control oriented models that are suitable in the MPC framework.
 

### 2018-01-18 paper/final/Response_to_reviewers/Response_to_reviewers.tex
Introduction and noisy forecast

@@
 \documentclass{article}
+
 \textwidth 6in%
+
 \textheight 7.9in%
+
 \oddsidemargin 1cm
+
 \usepackage{amsmath}
+
 \usepackage{amsfonts}
+
 \usepackage{amssymb}
+
 \usepackage{graphicx}%
+
 \usepackage[dvips]{color}
+
 \setcounter{MaxMatrixCols}{30}
+
 \usepackage{natbib}
@@
 \def\R{\mathcal{R}}
+
 \def\P{\mathcal{P}}
+
 \def\O{\mathcal{O}}
+
 \def\L{\mathcal{L}}
+
 \def\S{\mathcal{S}}
+
 \def\W{\mathcal{W}}
+
 \def\E{\mathcal{E}}
+
 \def\M{\mathcal{M}}
@@
 }
+
 \author{Francesco Smarra, Achin Jain, Tullio de Rubeis, \\ Dario Ambrosini, Alessandro D’Innocenzo, Rahul Mangharam}
@@
 \date{}
+
 \maketitle
@@
 The following major and minor revisions have been applied (all changes in the manuscript are highlighted in \textcolor[rgb]{0.00,0.00,1.00}{blue colored fonts}):
+
 \begin{itemize}
+
 	\item Section 1 - INTRODUCTION
+
 	\begin{itemize}
+
 		\item
+
 		\item
+
 	\end{itemize}
+
 	\item Section 2 - DATA PREDICTIVE CONTROL
+
 	\begin{itemize}
+
 		\item
+
 		\item
+
 		\item
+
 	\end{itemize}
+
 	\item Section 3 - COMPARISON WITH MPC
+
 	\begin{itemize}
+
 		\item
+
 	\end{itemize}
+
 	\item Section 4 - APPLICATION: DEMAND RESPONSE
+
 	\begin{itemize}
+
 		\item
+
 	\end{itemize}
+
 	\item Section 5 - CASE STUDY: OPTIMAL HEATING SYSTEM SCHEDULING
+
 	\begin{itemize}
+
 		\item
+
 	\end{itemize}
+
 	\item Minor changes
+
 	\begin{itemize}
+
 		\item
+
 		\item
+
 		\item
+
 		\item
+
 		\item
+
 	\end{itemize}
+
 \end{itemize}
@@
 \begin{enumerate}
-	\item \textbf{Associate Editor:} \textit{An updated and complete literature review should be conducted.
-The relevance to Applied Energy should be enhanced with the considerations of scope and readership of the Journal.
+
+	\item \textbf{Associate Editor:} \textit{An updated and complete literature review should be conducted.}
+
+	\textbf{Authors:} We performed a deep literature review. In particular, in Table 1 we classify the related work and then emphasise the novelty of our research.
+
+	\item \textbf{Associate Editor:} \textit{The relevance to Applied Energy should be enhanced with the considerations of scope and readership of the Journal.
 The present form has little relevance to APEN which might suit better for another journal.
-Authors must address and enhance the relevance to Applied Energy otherwise, they can consider to publish in another journal.
-}.
+Authors must address and enhance the relevance to Applied Energy otherwise, they can consider to publish in another journal.}
 
-	\textbf{Authors:}
+	\textbf{Authors:} Following the Associate Editor and reviewers comments, we re-wrote the title and the introduction of the paper focusing on the relevance of the paper to APEN.
+In particular, we emphasise that...EXPLAIN HERE AS IN INTRODUCTION
 
@@
 
-	\textbf{Authors:}
+	\textbf{Authors:} We deeply revised the paper to improve both language and organization quality.
 
@@
 The present form does not have sufficient results to justify the novelty of a high quality journal paper.
-}
+The results should be further elaborated to show how they could be used for the real applications.}
 
-	\textbf{Authors:}
+	\textbf{Authors:} We clarify in the introduction and in Section ???
+the originality of the paper and how are results can be used for real applications.
+In particular, we emphasise that...EXPLAIN HERE AS IN INTRODUCTION
 
-	\item \textbf{Associate Editor:} \textit{The results should be further elaborated to show how they could be used for the real applications.}
+	\item \textbf{Associate Editor:} \textit{}
 
-	\textbf{Authors:}
+	\textbf{Authors:} ???
 
@@
 \begin{enumerate}
+
 \item \textbf{Reviewer 1:} \textit{This paper presents a new framework called data predictive control which integrates machine learning and model predictive control to derive the optimal online control decisions.
 The framework is then applied to study several demand response problems in energy area as case studies.
-The problem considered in this paper is important, and the proposed approach is a good initial attempt to address the problem relying on heuristics.}.
+The problem considered in this paper is important, and the proposed approach is a good initial attempt to address the problem relying on heuristics.}
 
-\textbf{Authors:}
+\textbf{Authors:} We thank the reviewer for the comments on the problem and approach of our paper.
 
@@
 \textbf{Authors:}
+
 	\begin{itemize}
+
 		\item Explain in the introduction (other than in section 2) the difference between the state-space modeling and the one we propose;
+
 		\item Add a subsection in the "Related work" to explain the evolution of our work with respect to our previous ones;
+
 		\item Add Appendix on regression trees and random forests;
+
 		\item Add equations in the intro.
@@
 Setup and write equation to show the complexity
+
 		\item add another section to show how the previous point simplifies using data-driven methods
+
 	\end{itemize}
@@
 
-\textbf{Authors:}
+\textbf{Authors:} We thank the reviewer for this comment, indeed our previous presentation of the paper was unclear on this fundamental aspect.
+In our revision we widely illustrate in the Introduction how in traditional state-space models part of the state is measurable (i.e.
+room temperatures, etc) and part is not measurable (i.e.
+layers/windows temperatures, etc).
+In our approach, of course, $(X^c, X^d, Y)$ are ALL measurable variables, i.e.
+variables that are present in the historical data of a building.
+The unmeasurable variables of the traditional state-space models are not assumed to be measurable in this paper, yet our prediction works well since we compensate via the regression tree/random forest approaches.
 
@@
 
-\textbf{Authors:}
+\textbf{Authors:} We agree with the reviewer.
+Following this comment, and also other comments, we re-wrote the corresponding Section to make it more clear and complete to descrive the DPC technical approach.
+In particular...
 
@@
 
-\textbf{Authors:}
+\textbf{Authors:} We thank the Reviewer for the comment on our validation of the efficacy of our approach.
 
@@
 
-\textbf{Authors:}
+\textbf{Authors:} We agree with the Reviewer.
+In the beginning, we had chosen a more general title because we believe that our approach has the potential to be applied not only to building management, but also to other application domains.
+However, as in this paper we develop our techniques focusing and experimenting in building management (which is indeed a core topic of APEN) we changed the title in "???".
 
@@
 
-\textbf{Authors:}
+\textbf{Authors:} In the introduction of the revised version we widely discuss that in this paper we improved the preliminary results obtained in [9-12] and adapted them to be applicable to the building automation use-case presented in section ???.
 
@@
 
-\textbf{Authors:}
+\textbf{Authors:}  We agree with the reviewer.
+Following this comment, and also other comments, we re-wrote the corresponding Section to make it more clear and complete to descrive the DPC technical approach.
+In particular...
 
 \item \textbf{Reviewer 2:} \textit{The main optimization problems (4) and (7) are not clear in themselves.
-As you look ahead to do DPC, does each future time slot have a different linear model, depending on future non-manipulative states?}
-
-\textbf{Authors:}
-
-\item \textbf{Reviewer 2:} \textit{Then, do you assume you know precisely all the future non-manipulative states?
+As you look ahead to do DPC, does each future time slot have a different linear model, depending on future non-manipulative states?
+Then, do you assume you know precisely all the future non-manipulative states?
 If so, can you justify this?
@@
 
-\textbf{Authors:}
+\textbf{Authors:} We thank the reviewer for rising this important issue.
+Yes, we assume that at time $k$ we know the future non-manipulative states (namely external temperature, radiation, etc.) for the Model Predictive Control time horizon, namely for the next 40 minutes.
+In the revised version of this paper, to answer the reviewer comment, we justify this assumption and show that, even with a large error in predicting the future states, the performance does not change significantly.
+Forests??
+More precisely, we:
 
 	\begin{itemize}
+
 		\item Cite study that shows that short term predictions are extremely reliable and say there is a trade-off between the prediction accuracy and the horizon N;
+
 		\item Provide a simulation (as for example with the bilinear model) where we inject a noise into the disturbance and show that it works as well.
+
 	\end{itemize}
@@
 
-\textbf{Authors:}
+\textbf{Authors:} Achin?
 
 \item \textbf{Reviewer 2:} \textit{Also, what is new in section 5 compared with section 4?
-Even though a real house is claimed to be used, it's still tested using simulations and energyplus, like in the previous sections.}
+Even though a real house is claimed to be used, it's still tested using simulations and energyplus, like in the previous sections.
+A real world experiment on this house would be much more interesting.}
 
-\textbf{Authors:}
-
-\item \textbf{Reviewer 2:} \textit{A real world experiment on this house would be much more interesting.}
-
-\textbf{Authors:}
+\textbf{Authors:} Francesco?
 
 \end{enumerate}
+
 %\bibliographystyle{plain}
@@
 %\bibliography{mcnbib}
+
 \end{document}

### 2018-01-18 paper/final/casestudyReal.tex
Introduction and noisy forecast

--- 2017-12-24 paper/first submission/casestudyReal.tex
@@
 In particular, we considered a small ($Q=100$, $\lambda_{min}=3000$ and $\lambda_{max}=100$), a medium ($Q=100$, $\lambda_{min}=1000$ and $\lambda_{max}=100$) and a large ($Q=100$, $\lambda_{min}=100$ and $\lambda_{max}=100$) violation configuration.
+
+\textcolor[rgb]{0,0,1}{We considered 2 different simulative conditions to show the robustness of our approach:
+\begin{enumerate}
+	\item in Section \ref{SSS:DisturbancePerfect} we provide our simulative results considering perfect knowledge of the weather forecast;
+	\item in Section \ref{SSS:DisturbanceUncertain}, to show the robustness of our methodology, we consider weather forecast subject to uncertainty.
+We add a gaussian noise with high variance on the perfect forecast and show that the results are close to the perfect forecast case.
+\end{enumerate}
+\subsubsection{Perfect knowledge of the weather forecast}\label{SSS:DisturbancePerfect}
+In this section we ran the simulations considering perfect knowledge of the disturbance over the horizon, obtaining the following results.}
 \paragraph{Result 1} The comparison for temperature and input schedule obtained allowing small bound violations in DPC is shown in Figure \ref{F:comparison_small}.
@@
 	%	\scalebox{0.9}{
-	\begin{tabular}{cccccc}
+	\begin{tabular}{lccccc}
 		\toprule
@@
 \end{figure}
+
+\textcolor[rgb]{0,0,1}{
+\subsubsection{Weather forecast subject to uncertainty}\label{SSS:DisturbanceUncertain}
+In \cite{Petersen2014AE} the authors investigated, through a large-scale simulation study, the effect of weather forecast uncertainty on the performance of MPC in building systems operations, and compared to the performance of a rule-base strategy.
+They considered 48 different scenarios of uncertainties on 72 hours weather forecasts.
+On such long horizon, results have shown that, with few exceptions, MPC outperforms the rule based controller in case of energy savings and/or thermal indoor environment, and is in general quite close to the perfect forecast case, despite the uncertainty in weather forecast.}
+
+\textcolor[rgb]{0,0,1}{Horizon length for the DPC algorithm presented in this paper is usually much shorter, as for example $6\ hours$ in Section \ref{S:proof}, $7\ hours$ in Section \ref{S:casestudy}, and $40\ minutes$ in Section \ref{S:realCaseStudy}.
+This means the weather forecast is more accurate than the 72 hours case presented in \cite{Petersen2014AE}, hence also control performance is even closer to the perfect knowledge forecast case.}
+
+\textcolor[rgb]{0,0,1}{Nevertheless, to show the robustness of our approach, in this section we provide simulations considering noisy weather forecast.
+To this aim, we add noises with normal distribution on the disturbance variables in $\tX^d$ used in the DPC problem \eqref{E:DPCrealcase}, i.e.
+the disturbance in input in the blue rectangle in Figure \ref{F:overview}, while we use the correct ones to simulate the process, i.e.
+the disturbance in input in the "Plant" box in the red rectangle in Figure \ref{F:overview}.
+In Table \ref{T:NoiseParameters} we summarize the mean and the deviation values of the Gaussian noises that we add to each variable.
+We obviously do not add any noise on the "Time of the day" and on the "Day of the week" since those are perfectly known.
+In the table we also report the range of the values that the variables assumed during the reference period of the historical dataset.
+This is to compare with the mean and deviation values we chose to show that we considered quite bad error on the forecast.
+In the following results we show the change in performance on bounds violation and energy consumption due to the forecast inaccuracy.}
+\begin{table}[t!]
+	\centering	
+	\textcolor[rgb]{0,0,1}{\begin{tabular}{lccc}
+		\toprule
+		Variable               & Range      & Mean & Deviation \\ 
+		\midrule
+		Outside temperature    & [0,31]     & 1	   & 0.5       \\
+		Wind                   & [0,5] 		& 0.2  & 0.25      \\
+		Atmospheric pressure   & [990,1030] & 50   & 50        \\
+		Relative Humidity      & [20,91]	& 50   & 5         \\
+		Solar Radiation        & [0,1000]   & 50   & 50        \\
+		Time of the day        & [0,23]     & 0    & 0         \\
+		Day of the week        & [1,7]      & 0    & 0         \\
+		\bottomrule
+	\end{tabular}}
+	\caption{\textcolor[rgb]{0,0,1}{Mean and deviation values of the Gaussian noises added on the weather forecast data, and the range of variation of the variables in the historical dataset.}}
+	\captionsetup{justification=centering}
+	\label{T:NoiseParameters}
+\end{table}
+
+\textcolor[rgb]{0,0,1}{\paragraph{Result 1} In Figure \ref{F:ComparisonTempNoisy} a comparison of the DPC results in terms of temperature control, between the perfect and the noisy forecasts, is shown.
+For the sake of plot's clarity, we considered 4 days and a half as simulative period and we split the plot into 3 sub-figures referring to the small, medium and large violation cases.
+We can see that in the case of small and large violation we have a small deterioration in performance in terms of bounds violation, although this is more highlighted in the large violation case.
+The medium violation cases are instead comparable.
+The maximum violation, as in the case of perfect forecast, is still lower than $1\,\degree C$.
+In particular, we report in Table \ref{T:violationErrorsNoisy} the violation errors in terms of MBE and CV(RMSE).
+If we compare them with the ones in Table \ref{T:violationErrors}, we can see that the performance of DPC with noisy forecast are still good if compared with the performance of DPC with perfect forecast, despite the extremely bad prediction error that we considered on the weather forecast.
+}
+\begin{figure}[t!]
+	\begin{center}
+		\vspace{1.1cm}
+		\subfigure{
+			\label{F:SmallNoisy}
+			\centering
+			\includegraphics[width=23pc]{figures/Temperatures_small_noisy.eps}
+		}	
+		\subfigure{
+			\label{F:MediumNoisy}
+			\centering
+			\includegraphics[width=23pc]{figures/Temperatures_medium_noisy.eps}
+		}
+		\subfigure{
+			\label{F:LargeNoisy}
+			\centering
+			\includegraphics[width=23pc]{figures/Temperatures_large_noisy.eps}
+		}	
+	\end{center}
+	\caption{\textcolor[rgb]{0,0,1}{Comparison of DPC considering perfect and noisy weather forecasts. Small, medium and large violation cases are compared.}}
+	\captionsetup{justification=centering}
+	\label{F:ComparisonTempNoisy}
+\end{figure}
+
+\begin{table}[t!]
+	\centering
+	%	\scalebox{0.9}{
+	\textcolor[rgb]{0,0,1}{\begin{tabular}{lccccc}
+		\toprule
+		CONTROLLER  & LBV $\mathrm{MBE}$  & LBV $\mathrm{RMSE}$ & UBV $\mathrm{MBE}$ & UBV $\mathrm{RMSE}$ 	\\ 
+		\midrule
+		$DPC-SV$    & $0.072$             & $0.074$  			      & $0$    				 & $0$     	  	\\
+		$DPC-MV$    & $0.168$ 			  & $0.13$       			  & $0$    				 & $0$		  	\\
+		$DPC-LV$    & $0.708$  			  & $0.343$     			  & $0$    				 & $0$	      	\\
+		\bottomrule
+	\end{tabular}}
+	%	}
+	\caption{\textcolor[rgb]{0,0,1}{Lower Bound Violation (LBV) and Upper Bound Violation (UBV) errors expressed as $\mathrm{MBE}\%$ and $\mathrm{CV(RMSE)}\%$ for DPC Small Violation (DPC-SV), DPC Medium Violation (DPC-MV), DPC large Violation (DPC-LV) and bang-bang controller, considering noisy weather forecast.}}
+	\captionsetup{justification=centering}
+	\label{T:violationErrorsNoisy}
+\end{table}
+
+\textcolor[rgb]{0,0,1}{\paragraph{Result 2} 
+In Figure \ref{F:comparison_all_energy_E+_noisy}, we show the difference in terms of energy consumption considering the perfect weather forecast and the noisy one.
+We can observe the case with imperfect forecast is very close to the perfect forecast case, and depending the conditions for which the disturbance is over or underestimated, we have more or less energy consumption.
+These results shows the strong robustness of the DPC with respect to errors in the weather forecast.}
+\begin{figure}[t!]
+	\begin{center}
+		\includegraphics[width=28pc]{figures/Energy_all_EnergyPlus_noisy.eps}
+	\end{center}
+	\caption{\textcolor[rgb]{0,0,1}{Comparison of DPC in terms of thermal energy saving using EnergyPlus model, considering perfect (full) and imperfect (dashed) weather forecast, over $15$ days of the testing period with different violation configurations.
+The performance are pretty similar, showing the potential robustness of DPC.}}
+	\label{F:comparison_all_energy_E+_noisy}
+\end{figure}

### 2018-01-18 paper/final/conclusion.tex
Introduction and noisy forecast

--- 2017-09-30 paper/conclusion.tex
@@
 \end{enumerate}
+
 DPC has applications which go beyond buildings and energy systems, to industrial process control, and controlling large critical infrastructures like water networks, district heating \& cooling.

### 2018-01-18 paper/final/intro.tex
Introduction and noisy forecast

@@
 
-Control-oriented predictive models of an energy system's dynamics and energy consumption, are needed for understanding and improving the overall energy efficiency and operating costs.
+\textcolor[rgb]{0,0,1}{Control-oriented predictive models of an energy system's dynamics and energy consumption, are needed for understanding and improving the overall energy efficiency and operating costs of a building.
 With a reasonably accurate forecast of future weather and building operating conditions, dynamical models can be used to predict the energy needs of the building over a prediction horizon, as is the case with Model Predictive Control (MPC) \cite{Sturzenegger2016}.
 However, a major challenge with MPC is in accurately modeling the dynamics of the underlying physical system.
-The task is much more complicated and time consuming in case of a large building and often times, it can be even more complex and involved than the controller design itself.
-After several years of work on using first principles based models for peak power reduction, and energy optimization for buildings, multiple authors~\cite{Sturzenegger2016, vzavcekova2014} have concluded that the biggest hurdle to mass adoption of intelligent building control is the cost and effort required to capture accurate dynamical models of the buildings.
+The task is much more complicated and time consuming in the case of a large building and often times, it can be even more complex and involved than the controller design itself.
+After several years of work on using first principles based models for peak power reduction, energy optimization and thermal comfort for buildings, multiple authors~\cite{Sturzenegger2016, vzavcekova2014} have concluded that the biggest hurdle to mass adoption of intelligent building control is the cost, time and effort required to capture accurate dynamical models of the buildings.
 The user expertise, time, and associated sensor costs required to develop a model of a single building is very high.
-Thus, the payback period for the upfront hardware and software installation is expected be be too high making MPC an uneconomical choice for energy management.
+Thus, the payback period for the upfront hardware and software installation is expected to be too high, making MPC an uneconomical choice for energy management.}
 
-Why is physics-based modeling hard for complex systems like buildings or what are the reasons of modeling complexities?
+\textcolor[rgb]{0,0,1}{Why is physics-based modeling hard for complex systems like buildings or what are the reasons of such modeling complexities?}
 \begin{enumerate}
-	\item \textbf{Model capture} - a building modeling domain expert typically uses a software tool to create the geometry of a building from the building design and equipment layout plans, add detailed information about material properties, about equipment and operational schedules.
+	\item \textcolor[rgb]{0,0,1}{\textbf{Model capture} - a building modeling domain expert typically uses a software tool to create the geometry of a building from the building design and equipment layout plans, add detailed information about material properties, about equipment and operational schedules.
 There is always a gap between the modeled and the real building and the domain expert must then manually tune the model to match the measured data from the building~\cite{New2012}.
 	Moreover, the modeling process also varies from building to building with the construction and types of installed equipment.
-Another major downside with physics-based modeling is that enough data is not easily available and guesses for parameter values have to be made, which also requires expert know how.
-
-\item \textbf{Change in model properties} over time - even if the model is identified once via an expensive route as in \cite{Sturzenegger2016}, as the model changes with time, the system identification must be repeated to update the model.
+	Another major downside with physics-based modeling is that enough data is not easily available and guesses for parameter values have to be made, which also requires expert know-how.
+	\item \textbf{Change in model properties over time} - even if the model is identified once via an expensive route as in \cite{Sturzenegger2016}, as the model changes with time, the system identification must be repeated to update the model.
 Thus, model adaptability or adaptive control is desirable for such systems.
-
 	\item \textbf{Model heterogeneity} further prohibits the use of model-based control.
@@
 Therefore, this modeling process must be repeated for every new building.
+}
 \end{enumerate}
 
-Due to aforementioned reasons, the control strategies in such systems are often limited to fixed, sometimes ad-hoc, rules that are based on best practices. 
+\textcolor[rgb]{0,0,1}{Due to aforementioned reasons, the control strategies in such systems are often limited to fixed, sometimes ad-hoc, rules that are based on best practices. 
 The alternative is to use black-box, or completely data-driven modeling approaches, to obtain a realization of the system's input-output behavior. 
@@
 Unprecedented amounts of data from millions of smart meters and thermostats installed in recent years has opened the door for systems engineers and data scientists to analyze and use the insights that data can provide, about the dynamics and power consumption patterns of these systems.
+}
 
-The key question now is: can we employ data-driven techniques to reduce the cost of modeling, and still exploit the benefits that MPC has to offer?
-We therefore look for automatic data-driven approaches to control that are also adaptive, scalable and interpretable.
-We solve this problem by bridging the gap between Machine Learning and Predictive Control. In particular, in this paper, we present a method based on Random Forests.
+\textcolor[rgb]{0,0,1}{The key question now is: can we employ data-driven techniques to reduce the cost of modeling, and still exploit the benefits that MPC has to offer?
+We therefore look for automatic data-driven approaches to control, that are also adaptive, scalable and interpretable.
+We solve this problem by bridging the gap between Machine Learning and Predictive Control. In particular, in this paper, we present a method based on Random Forests.}
 
@@
 \subsection{Example for comparison of modeling complexity: physics-based vs data-driven}
-In this section, we provide a detailed example that describes the difference in terms of complexity when we model a building using physical laws and machine learning.
-We show that the data-driven modeling eliminates several drawbacks that occur using physics-based models such as the need to have very good knowledge of the building structure and the material properties, time required to build a model, and limited availability of sensor measurements.
+\textcolor[rgb]{0,0,1}{In this section, we provide a detailed example that describes the difference in terms of complexity when we model a building using physical laws and machine learning.
+We show that the data-driven modeling eliminates several drawbacks that occur using physics-based models such as the need to have very good knowledge of the building structure and the material properties, time required to build a model, and limited availability of sensor measurements.}
 
-The building under consideration is taken from \cite{Sturzenegger2016}.
+\textcolor[rgb]{0,0,1}{The building under consideration is taken from \cite{Sturzenegger2016}.
 It is located in Allschwil, Switzerland.
 It has 6 floors of which only the second floor is modeled, and is assumed to represent all other floors.
-The total air conditioned floor area is $6000\ \mathrm{m^2}$ ca.
+The total air conditioned floor area is $6000\ \mathrm{m^2}$ ca.}
 
@@
 \label{SSS:physics-based}
-We describe the physics-based approach based on an RC network outlined in \cite{Sturzenegger2016}.
-The building is modeled as a bilinear system constructed from physical principles, the modeling process consisting of three steps:
+\textcolor[rgb]{0,0,1}{We describe the physics-based approach based on an RC network outlined in \cite{Sturzenegger2016}.
+The building is modeled as a bilinear system constructed from physical principles.
+The modeling process consists of three steps:
 \begin{enumerate}
@@
 			\begin{aligned}
-				&C_a\dot x_a = \frac{x_b(t) - x_a(t)}{R}, \ \text{and}\\
+				&C_a\dot x_a = \frac{x_b(t) - x_a(t)}{R},\\
 				&C_b\dot x_b =  \frac{x_a(t) - x_b(t)}{R},
@@
 %		Equation \eqref{E:RCfluxes} comes from a series of 20 equations analyzed in Section 3.3.1.3 of the Technical Report \cite{SturzeneggerTR}, that we do not report here for the sake of simplicity. 
-Equation \label{E:RCfluxes} for the heat flux is obtained by modeling
+Equation \eqref{E:RCfluxes} for the heat flux is obtained by modeling
 	\begin{enumerate}
@@
 	\item The resulting system \eqref{E:RCZoneEq} is discretized with sampling time of $15\ \mathrm{min}$.
-This model has approximately 300 states which include temperature of the zones, walls and floors on the second floor.
-The output of the system are the zone temperatures.
-	Since the performance of the state estimation, that is needed to compute the optimal control inputs using MPC, depends on the number of states, an approximated model with fewer states is needed.
-	To this aim, different averaged temperatures of the building facades and the zones are considered, obtaining an approximated model with 35 states.
+	This model has approximately 300 states, which include temperature of the zones, walls and floors on the second floor.
+	The outputs of the system are the zone temperatures.
+	Since the performance of the state estimation, that is needed to compute the optimal control inputs using MPC depends on the number of states, an approximated model with fewer states is needed.
+	To this aim, different averaged temperatures of the building facades (North, South, West, East) and the zones are considered, obtaining an approximated model with 35 states.
 	This  approximate model is then considered ``suitable for MPC" (see Section 3.3.1.4 in \cite{SturzeneggerTR}). 
-\end{enumerate}
+\end{enumerate}}
 
-	In summary, the approximate model has the following state, input and disturbance variables:
+\textcolor[rgb]{0,0,1}{In summary, the approximate model has the following state, input and disturbance variables:
 	\begin{itemize}
@@
 		\item 18 inputs: TABS heating heat flux, TABS cooling heat flux, averaged transmitted solar heat flux for each group of zones (North, South, West, East) that is estimated using blinds position measurements, air massflow through the energy recovering mode, air massflow bypassing the energy recovering mode, air massflow through the air cooler, AHU heat coil heat flux, lighting power for the offices for each group of zones (North, South, West, East), and radiator heat flux in the corner offices (North, South, West, East).
-		\item 7 disturbances: internal gains in the offices and internal gains in non-office zones which are predicted using a standard schedule, ambient temperature and solar radiation on facade (North, South, West, East) whose values were obtained through Kalman filtering using measurements from the weather station placed on the roof of the building.
+		\item 7 disturbances: internal gains in the offices and internal gains in non-office zones, which are predicted using a standard schedule, ambient temperature and solar radiation on facade (North, South, West, East) whose values were obtained through Kalman filtering using measurements from the weather station placed on the roof of the building.
 This filtering is needed to take into account the shadowing of the neighboring buildings.
@@
 An EnergyPlus model is built for this system, and model parameters of matrices $A_x$, $B_q$, $A_q$, $B_{q,u}$, $B_{q,d}$, $B_{q,du,i}$, $D_{q,xu,i}$ are derived using geometry and materials data from the EnergyPlus model.
-This was a choice of the authors; real geometry and materials data can also be used to estimate the model parameters. 
-For this particular building, 24 parameters are estimated/taken form a datasheet computed for the considered zone model.
+This was a choice of the authors; real geometry and materials data can also be used to estimate the model parameters, although not always available. 
+For this particular building, 24 parameters are estimated/taken form a datasheet/computed for the considered zone model.
 Although some of the parameters are in common between different zones, the others are found independently for each zone.
-To derive the EnergyPlus model using the available measurements and to use the same control implemented in the building, the EnergyPlus model is coupled with MATLAB using BCVTB \cite{Wetter2015}. 
+To derive the EnergyPlus model using the available measurements and to use the same control implemented in the building, the EnergyPlus model is coupled with MATLAB using BCVTB \cite{Wetter2015}.\textcolor[rgb]{1,0,0}{ ADD THIS IS ONLY FOR ONE FLOOR, AND THAT THE OTHER ARE ASSUMED TO BE EQUAL.
+THIS IS TO EMPHASIZE THE COMPLEXITY}}
 %To identify the model a total of 294 MATLAB signals were sent to the EnergyPlus model and a total of 1081 EnergyPlus signals were sent back to Matlab.
 
-The physics-based approach, although detailed and accurate (in some cases), is cost and time prohibitive.
-The long and complex modeling procedure that is unique to every building poses a challenge in applying MPC to large scale buildings.
+\textcolor[rgb]{0,0,1}{The physics-based approach, although detailed and accurate (in some cases), is cost and time prohibitive.
+The long and complex modeling procedure that is unique to every building poses a challenge in applying MPC to large scale buildings.}
 
 \subsubsection{Data-driven modeling}
-The data-driven approach we propose, on the other hand, ignores the physical details of the building operation.
-Instead, we use machine learning to learn black-box models where our objective is to find the hyperparameters (for a given structure like Random Forests) of a model that best explains the input-output relationship.
+\textcolor[rgb]{0,0,1}{The data-driven approach we propose, on the other hand, ignores the physical details of the building operation.
+Instead, we use machine learning to learn black-box models where our objective is to find the hyperparameters (for a given structure like Random Forests) of a model that best explains the input-output relationship.}
 
-The goal with the data-driven modeling is to learn a function map
+\textcolor[rgb]{0,0,1}{The goal with the data-driven modeling is to learn a function map
 \begin{equation}
@@
 where \(y\) represents the outputs, \(u\) the inputs and \(d\) the disturbances as defined earlier, and \(\delta\) is a lag variable.
-If \(y\) is the average room temperature for one of the zones, a nonlinear function \(f\) is more appropriate, for example \(f\) may belong to the class of Neural Networks or Random Forests.
+If \(y\) is the average room temperature for one of the zones, a nonlinear function \(f\) is more appropriate, for example \(f\) may belong to the class of Neural Networks or Random Forests.\textcolor[rgb]{1,0,0}{???????????????}}
 
-The data-driven modeling in the case of buildings has several advantages over the physical counterpart.
+\textcolor[rgb]{0,0,1}{The data-driven modeling in the case of buildings has several advantages over the physical counterpart.}
 \begin{enumerate}
 
-	\item \textbf{Domain expertise:} 
-	This procedure eliminates the need of a building expert, simplifies the modeling complexities described in Section \ref{SSS:physics-based} which ultimately reduces the cost of modeling a building.
-Further, for a new building, given the historical data from the building, the same process can be repeated to identify a control-oriented model for MPC.
+	\item \textcolor[rgb]{0,0,1}{\textbf{Domain expertise:} 
+	This procedure (\textcolor[rgb]{1,0,0}{NOT CLEAR IF EXPLAINED IN ONLY 2 LINES BEFORE}) eliminates the need of a building expert, simplifies the modeling complexities (\textcolor[rgb]{1,0,0}{HOW?}) described in Section \ref{SSS:physics-based} which ultimately reduces the cost of modeling a building.
+Further, for a new building, given the historical data from the building, the same process can be repeated to identify a control-oriented model for MPC.}
 
-	\item \textbf{Sensor cost:}
-	Due to large number of states and variables in the physical modeling, a lot of measurements are needed to use the model to predict the system's behavior.
+	\item \textcolor[rgb]{0,0,1}{\textbf{Sensor cost:}
+	Due to large number of states and variables in the physical modeling, a lot of measurements are needed to use the model to predict the system's behavior \textcolor[rgb]{1,0,0}{IF USING A MODEL-BASED APPROACH}.
 	This can be expensive due to necessary new sensor installations.
 	Moreover, when some measurements are not directly available from the sensors, observers are needed for state estimations.
-	However, observability problems can limit the construction of the observers \cite{Dorf2011MCS}. 
-	On the other hand, data-driven modeling obviates the need to model the state of the walls and floor layers so we do not have as many states.
+	However, observability problems can limit the construction of such observers \cite{Dorf2011MCS}. 
+	On the other hand, data-driven modeling obviates the need to model the state of the walls and floor layers\textcolor[rgb]{1,0,0}{AND OTHER STATES DIFFICULT TO MEASURE} so we do not have as many states.\textcolor[rgb]{1,0,0}{EXPLAIN THAT THE MACHINE LEARNING ALGORITHM IS ABLE TO COMPENSATE MISSING VARIABLES}
 	It relies only on direct measurements from the sensors such as thermostats and multimeters reducing the need for new installations and hence the cost of modeling.
-	Further, when some of the states are not directly measurable, often adding proxy variables suffices. 
-	For example, if we do not have the measurement of the energy produced by a photovoltaic panel, the weather conditions will account for it.
+	Further, when some of the states are not directly measurable, often adding proxy \textcolor[rgb]{1,0,0}{I.E. ??}variables suffices. 
+	For example, if we do not have the measurement of the energy produced by a photovoltaic panel, the weather conditions will account for it.}
 
-	\item \textbf{Linearity assumption:} 
+	\item \textcolor[rgb]{0,0,1}{\textbf{Linearity assumption:} 
 	During physical modeling in \eqref{E:RCHeatExchange}, in order to keep the model simple, the heat exchange between layers is modeled with a linear behavior. 
@@
 	However, more complex nonlinearities have been neglected due to the already high complexity of the model \cite{Sturzenegger2016}. 
-	In data-driven modeling, we can relax these linearity assumptions as nonlinear functions can be learned rather efficiently and accurately.
-
+	In data-driven modeling, we can relax these linearity assumptions as nonlinear functions can be learned rather efficiently and accurately.\textcolor[rgb]{1,0,0}{IT IS NOT CLEAR THAT THE DD ALGORITHM TAKE CARE OF IT CREATING ITS INTERNAL STRUCTURE}
 	\item \textbf{Other assumptions:} 
@@
 	Data-driven modeling depends on the sensor data directly. 
-	Hence, such assumptions never arise.
+	Hence, such assumptions never arise.}
 
@@
 
-To summarize, the data-driven modeling simplifies the modeling of buildings to a large extent, reduces the overall cost and time investment while avoiding the linear model assumptions that we make in standard physical modeling procedure.
-Unlike physical modeling, the parameters of the model (or hyperparamters) are automatically determined by the learning algorithm, the best settings determined by cross validation.
+\textcolor[rgb]{0,0,1}{To summarize, the data-driven modeling simplifies the modeling of buildings to a large extent, reduces the overall cost and time investment while avoiding the linear model assumptions that we make in standard physical modeling procedure.
+Unlike physical modeling, the parameters of the model (or hyperparamters) are automatically determined by the learning algorithm, the best settings determined by cross validation.}
 
-However, the challenge lies in using such models for optimal control, which is exactly the focus of this paper.
-We demonstrate how the training algorithm for Random Forests can be modified to develop control oriented models that are suitable in the MPC framework.
+\textcolor[rgb]{0,0,1}{However, the challenge lies in using such models for optimal predictive control, which is exactly the focus of this paper.
+We demonstrate how the training algorithm for Random Forests can be modified to develop control oriented models that are suitable in the MPC framework.}
 
@@
 \subsection{Related work}
-This section is divided in 3 parts:
+A vast amount of literature exists dealing with methodologies for several energy efficiency applications in buildings, i.e.
+Demand Response, peak power reduction, energy saving and thermal comfort, and others.
+All this approaches can be mainly classified considering 2 main features:
 \begin{enumerate}
-	\item In Section \ref{SSS:ModelBasedWork} we provide a literature review of the approaches that make use of model-based techniques for energy optimization.
-	\item In Section \ref{SSS:DataDrivenWork} we report the main works that use of data-driven methodologies for energy optimization and point out the difference with respect to our contribution.
-	\item In Section \ref{SSS:PreviousWork} we underline the differences between the results presented in this work and the ones of our previous papers.
+	\item the system's model considered:
+		\begin{itemize}
+			\item model-based, that we consider both "white-box" and "grey-box" approaches: \cite{Shakouri2017SCS,Li2014EB,Yoon2014EB,Li2016AE,Harb2016EB,Salakij2016EB,Li2016E,Li2016EB,Hou2013,Cecconi2017EB};
+			\item data-based, that is the "black-box" approach, mainly done using Neural Networks: \cite{Safa2017SCS,Neto2008EB,Magnier2010BE,Candanedo2017EB,Ascione2017E,Cecconi2017EB,Li2016AE};
+			\item the only use of simulation tools, such as EnergyPlus \cite{energyPlus} and TRNSYS \cite{trnsys2000}: \cite{Yin2016EB,Christantoni2016EB};
+		\end{itemize}
+	\item the purpose these models are created for:
+		\begin{itemize}
+			\item only model creation and identification: \cite{Safa2017SCS,Neto2008EB,Magnier2010BE,Li2014EB,Li2016AE,Harb2016EB,Li2016EB,Candanedo2017EB,Cecconi2017EB,Ascione2017E};
+			\item control, typically Predictive Control, other than a modeling part: \cite{Shakouri2017SCS,Yoon2014EB,Yin2016EB,Salakij2016EB,Li2016E,Hu2017AE}.
+		\end{itemize}
 \end{enumerate}
-\subsubsection{Model-based related work}\label{SSS:ModelBasedWork}
-\subsubsection{Data-driven related work}\label{SSS:DataDrivenWork}}
-In the literature, several studies deal with the data-driven control techniques using machine learning algorithms \cite{Hou2013}.
-For example, Artificial Neural Networks are exploited to create data-driven models to be used as plant simulator in closed-loop with a Supervisory Model Predictive Control in \cite{Afram2017}.
-In \cite{Macarulla2017} the authors proposed a predictive control strategy based on Neural Networks, for boilers control in buildings, to decide the optimal time to switch-on the plan to guarantee energy saving and thermal comfort.
+These papers are summarize in Table \ref{T:RelatedWork}, where all differences are highlighted.
+We also emphasized the case studies the results are applied to, and whether the authors used experimental data to simulate their algorithms.
+Only in 3 cases the algorithms were tested on real systems (in the table: RI - Real Implementation).
+We can see that, except for the last 6 cases, which we will discuss more in detail in the rest of this section, when model-based approaches or only tools are considered, then also control techiniques are applied.
+While with the data-driven modeling approaches, only model identification is addressed.
+This is because bridging data-driven models with control theory, especially with the predictive control, introduces several issues \cite{Hou2013}.
+Dealing with this kind of problems attracted a lot of researcher's attention in the past years.
+We consider 6 of such works (the last 6 papers of Table \ref{T:RelatedWork}) that are more related to the methodology presented in this paper.
+In particular, in \cite{Macarulla2017} the authors proposed a predictive control strategy based on Neural Networks, for boilers control in buildings, to decide the optimal time to switch-on the plan to guarantee energy saving and thermal comfort.
 However, the strategy is not easily scalable to different types of plants and does not use optimization theory in the closed-loop scheme.
 In \cite{Costanzo2016} a reinforcement learning control strategy, called Model-Assisted Batch Reinforcement Learning, is considered to provide data-driven control for the demand response problem in HVAC systems.
-Reinforcement Learning is a model-free methodology and for this reason it can not be used for on-line optimization in a closed-loop predictive control scheme.
+Reinforcement Learning is a model-free methodology and for this reason represents an alternative approach, with pros and cons, to the MPC \cite{Ernst2009TSMC}.
 In \cite{Ferreira2012} the authors considered a data-driven predictive control based on Neural Networks to guarantee energy saving and  thermal comfort in public buildings.
-Neural Networks are used in the closed-loop control scheme to determine a thermal comfort index based on parameters that can be measured or estimated, but no  system dynamics with internal state are included into the optimization problem.
-More papers related to this topic can be found in the literature, but to the best of the authors' knowledge none of them address the problem of including data-driven state models in the optimal predictive control loop, and hence allow to set up a MPC-like optimal control problem, as we do in DPC.
+Neural Networks are used in the closed-loop control scheme to determine a thermal comfort index based on parameters that can be measured or estimated, but no Neural Networks-based system state dynamics are included into the optimization problem.
+However, Neural Networks are a particular type of machine learning algorithm that provides complex nonlinear equations to describe the system's dynamics.
+For example, \cite{Afram2017} considers a data-driven state model in the MPC closed-loop optimization.
+More precisely, Artificial Neural Networks are exploited to create data-driven models to be used as plant simulator in closed-loop with a Supervisory Model Predictive Control.
+More papers related to this topic can be found in the literature.
+Unfortunately, since Neural Networks models are nonlinear, the MPC based on such models is nonlinear as well.
+This means that a global optimal solution can not be provided.
+Furthermore, due to the high number of neurons that a Neural Network usually needs to provide good models, the nonlinear MPC problem can results into a quite complex optimization problem to be solved.
+To overcome this complexity problem, we proposed regression trees-based approaches in \cite{Behl2016,Jain2017TCPS}.
+However these approaches suffer of some drawbacks.
+In this paper, we provide a new methodology based on random forests, that overcome the drawbacks of our previous works.
+It provides robustness with respect to disturbance uncertainties and has extremely good performance when compared to other approaches (both optimal and rule-based).
+In the sext section we emphasize the differences with our previous approaches.
+}
+
+\begin{table}[t!]
+	\centering
+	\textcolor[rgb]{0,0,1}{\small\begin{tabular}{llcccc}
+		\toprule
+		Ref.                             & Case study              & Exp.                & Tool                & MB/DD             & Control           \\ 
+		\midrule
+		\cite{Yin2016EB}                 & Commercial Building     & Yes                 & E+                  & None              & Yes               \\
+		\cite{Christantoni2016EB}        & Commercial building     & Yes                 & E+                  & None              & Yes               \\
+		\cite{Li2014EB}                  & Commercial building     & Yes                 & E+                  & MB                & No                \\
+		\multirow{2}*{\cite{Harb2016EB}} & 2 office buildings and  &  \multirow{2}*{Yes} & \multirow{2}*{None} & \multirow{2}*{MB} & \multirow{2}*{No} \\
+		                                 & 1 residential building  &                     &                     &                   &                   \\
+		\cite{Li2016EB}                  & 2 commercial buildings  & n/a                 & E+                  & MB                & No                \\
+		\cite{Shakouri2017SCS}           & Residential area        & No                  & None                & MB                & Yes               \\
+		\cite{Yoon2014EB}                & 2 residential buildings & Yes                 & E+                  & MB                & Yes               \\
+		\cite{Salakij2016EB}             & 3 residential buildings & No                  & E+                  & MB                & Yes               \\
+		\cite{Li2016E}                   & 6 commercial buildings  & Yes                 & E+                  & MB                & Yes               \\
+		\cite{Hu2017AE}                  & Residential building    & Yes                 & None                & MB                & Yes               \\
+		\cite{Cecconi2017EB}             & Commercial building     & No                  & E+                  & MB-DD             & No                \\
+		\cite{Li2016AE}                  & 2 commercial buildings  & Yes                 & E+                  & MB-DD             & No                \\
+		\cite{Safa2017SCS}               & Office building         & Yes                 & None                & DD                & No                \\
+		\cite{Neto2008EB}                & Office building         & Yes                 & E+                  & DD                & No                \\
+		\cite{Magnier2010BE}             & Residential house       & Yes                 & TRANSYS             & DD                & No                \\
+		\cite{Candanedo2017EB}           & Residential building    & Yes                 & None                & DD                & No                \\
+		\cite{Ascione2017E}              & Office building         & No                  & E+                  & DD                & No                \\
+		\cite{Macarulla2017}             & Commercial building     & Yes+RI              & None                & DD                & Yes               \\
+		\cite{Costanzo2016}              & Living lab (1 room)     & Yes+RI              & None                & DD                & Yes               \\
+		\cite{Ferreira2012}              & Commercial building     & Yes+RI              & None                & DD                & Yes               \\
+		\cite{Afram2017}                 & Residential house       & Yes                 & None                & DD                & Yes               \\
+		\cite{Behl2016}                  & 9 commercial buildings  & No                  & E+                  & DD                & Yes               \\
+		\cite{Jain2017TCPS}              & Commercial building     & No                  & E+                  & DD                & Yes               \\
+		\bottomrule
+	\end{tabular}\normalsize}
+	\caption{\textcolor[rgb]{0,0,1}{References ordered considering: case study they have been applied to; whether they used experimental data, other than simulated data, and if they did real implementation (RI), i.e.
+implemented the methodologies on real systems; if they used simulative tool; the type of the model considered, i.e.
+Model-Based or Data-Driven; if they used the model to do control.}}
+	\captionsetup{justification=centering}
+	\label{T:RelatedWork}
+\end{table}
 
@@
 \begin{itemize}
-	\item In \cite{Behl2016} we	proposed an MPC-like control problem considering regression trees and ensemble methods algorithms for the Demand-Response problem.
-However, the proposed models allowed to optimally control the system considering only one-step lookahead prediction.
+	\item In \cite{Behl2016} we used regression trees and ensemble methods algorithms to create data-driven models for buildings. 
+	We then setup an MPC-like control problem using the regression trees-based model to be used in the Demand-Response problem.
+	However, ensemble methods have not been considered in the control process, but they have been used for identification purposes only.
+	Furthermore, the proposed models allowed to optimally control the system considering only one-step lookahead prediction.
 	Hence, it was not possible to use such approach to control the system considering a prediction over an horizon of arbitrary length.
-	\item This problem was addressed in \cite{Jain2016}, where the regression trees algorithm was modified to allow the tree to provide multiple-output.
+	\item This problem was addressed in \cite{Jain2017TCPS}, where the regression trees algorithm was modified to allow the tree to provide multiple-output.
 	Each output of the tree was then setup to provide the prediction of the system's behavior over different steps of the predictive horizon.
@@
 	Such approach has the advantage to be extremely simple from the complexity point of view, but its application range is limited to specific kind of systems.
-	\item In this work, we extend preliminary results gave in the conference paper \cite{JainCDC2017}.
-	An alternative approach is proposed to improve the system's identification accuracy and robustness of the previous approach.
+	\item In this work, we extend preliminary results gave in the conference papers \cite{JainACC2017}, where a simplified version of this approach was compared with a model-based control approach on a trivial system, and \cite{JainCDC2017}, where a more complex model was considered, but without taking into account any experimental data or forecast inaccuracy.
+	In particular, an alternative approach is proposed to improve the system's identification accuracy and robustness of the previous approaches.
 	More precisely, instead of considering a single tree with multiple output, we consider multiple trees with single output.
 	Each tree is used to provide the prediction of the system's behavior over different steps of the horizon.
-With a small increase in complexity, we setup an MPC-like control problem with arbitrary horizon, that provides better results with respect to our previous work.
-In this paper we further extend this approach to one of the ensemble methods: the random forests.
+	Furthermore, we extend this approach considering the random forests.
+	With a small increase in complexity, we setup an MPC-like control problem with arbitrary horizon, that provides better results and improves control robustness with respect to our previous work.
 \end{itemize}

### 2018-01-18 paper/final/intro.tex
edit related work and intro

@@
 We therefore look for automatic data-driven approaches to control, that are also adaptive, scalable and interpretable.
-We solve this problem by bridging the gap between Machine Learning and Predictive Control. In particular, in this paper, we present a method based on Random Forests.}
+We solve this problem by bridging Machine Learning and Predictive Control.
+In particular, in this paper, we present a method based on Random Forests which uses historical data for receding horizon control.}
+
+\textcolor[rgb]{0,0,1}{
+Next, we present an example to illustrate how data-driven modeling reduces the cost of modeling buildings. 
+}
 
@@
 	\end{enumerate}
-	\item The resulting system \eqref{E:RCZoneEq} is discretized with sampling time of $15\ \mathrm{min}$.
-	This model has approximately 300 states, which include temperature of the zones, walls and floors on the second floor.
+	\item The resulting system \eqref{E:RCZoneEq} is discretized with a sampling time of $15\ \mathrm{min}$.
+	This model has approximately 300 states that include temperature of the zones, walls and floors on the second floor.
 	The outputs of the system are the zone temperatures.
@@
 	\end{itemize}
-An EnergyPlus model is built for this system, and model parameters of matrices $A_x$, $B_q$, $A_q$, $B_{q,u}$, $B_{q,d}$, $B_{q,du,i}$, $D_{q,xu,i}$ are derived using geometry and materials data from the EnergyPlus model.
-This was a choice of the authors; real geometry and materials data can also be used to estimate the model parameters, although not always available. 
+An EnergyPlus model is built for this system, and model parameters of matrices $A_x$, $B_q$, $A_q$, $B_{q,u}$, $B_{q,d}$, $B_{q,du,i}$, $D_{q,xu,i}$ in Equations \eqref{E:RCZoneEq} and \eqref{E:RCHeatExchange} are derived using geometry and materials data from the EnergyPlus model.
+This was a choice of the authors; real geometry and materials data can also be used to estimate the model parameters, if necessary data are available. 
 For this particular building, 24 parameters are estimated/taken form a datasheet/computed for the considered zone model.
 Although some of the parameters are in common between different zones, the others are found independently for each zone.
-To derive the EnergyPlus model using the available measurements and to use the same control implemented in the building, the EnergyPlus model is coupled with MATLAB using BCVTB \cite{Wetter2015}.\textcolor[rgb]{1,0,0}{ ADD THIS IS ONLY FOR ONE FLOOR, AND THAT THE OTHER ARE ASSUMED TO BE EQUAL.
-THIS IS TO EMPHASIZE THE COMPLEXITY}}
+To simplify, the parameters of all the other floors are assumed to be identical which is not true in reality.
+To derive the EnergyPlus model using the available measurements and to use the same control implemented in the building, the EnergyPlus model is coupled with MATLAB using BCVTB \cite{Wetter2015}.
+}
 %To identify the model a total of 294 MATLAB signals were sent to the EnergyPlus model and a total of 1081 EnergyPlus signals were sent back to Matlab.
@@
 \textcolor[rgb]{0,0,1}{The physics-based approach, although detailed and accurate (in some cases), is cost and time prohibitive.
-The long and complex modeling procedure that is unique to every building poses a challenge in applying MPC to large scale buildings.}
+The long and complex modeling procedure that is unique to every building poses practical challenges in applying MPC to large scale buildings.}
 
@@
 \begin{equation}
+\label{E:MLexample}
 y(t) = f(u(t),\dots,u(t-\delta_u), d(t),\dots, d(t-\delta_d)),
 \end{equation}
-where \(y\) represents the outputs, \(u\) the inputs and \(d\) the disturbances as defined earlier, and \(\delta\) is a lag variable.
-If \(y\) is the average room temperature for one of the zones, a nonlinear function \(f\) is more appropriate, for example \(f\) may belong to the class of Neural Networks or Random Forests.\textcolor[rgb]{1,0,0}{???????????????}}
+where \(y\) represents the outputs, \(u\) the inputs and \(d\) the disturbances as defined earlier, and \(\delta\) is a lag variable.}
 
@@
 	\item \textcolor[rgb]{0,0,1}{\textbf{Linearity assumption:}
+	If \(y\) in Equation \eqref{E:MLexample} is the average room temperature for one of the zones, a nonlinear function \(f\) is more appropriate, for example \(f\) may belong to the class of Neural Networks or Random Forests.
 	During physical modeling in \eqref{E:RCHeatExchange}, in order to keep the model simple, the heat exchange between layers is modeled with a linear behavior. 
 	Hence all the nonlinearities were neglected.
-	In the case of heat flux in \eqref{E:RCfluxes}, a linear model is not good enough to provide a good approximation. For this reason a bilinear model was considered.
+	In the case of heat flux in \eqref{E:RCfluxes}, again, a linear model is not good enough to provide a good approximation. For this reason a bilinear model was considered.
 	However, more complex nonlinearities have been neglected due to the already high complexity of the model \cite{Sturzenegger2016}. 
-	In data-driven modeling, we can relax these linearity assumptions as nonlinear functions can be learned rather efficiently and accurately.\textcolor[rgb]{1,0,0}{IT IS NOT CLEAR THAT THE DD ALGORITHM TAKE CARE OF IT CREATING ITS INTERNAL STRUCTURE}
-	\item \textbf{Other assumptions:} 
+	In data-driven modeling, we can relax these linearity assumptions as nonlinear functions can be learned rather efficiently and accurately.\textcolor[rgb]{1,0,0}{IT IS NOT CLEAR THAT THE DD ALGORITHM TAKE CARE OF IT CREATING ITS INTERNAL STRUCTURE}}
+
+	\item \textcolor[rgb]{0,0,1}{\textbf{Other assumptions:} 
 	Due to the high modeling complexity, physical modeling neglects different geometries and different materials of the floors assuming they are the same for each floor which is never true in reality.
@@
 \subsection{Related work}
-A vast amount of literature exists dealing with methodologies for several energy efficiency applications in buildings, i.e.
-Demand Response, peak power reduction, energy saving and thermal comfort, and others.
-All this approaches can be mainly classified considering 2 main features:
+A vast literature exists in building energy applications that deals with Demand Response, peak power reduction, energy saving, thermal comfort, and related topics.
+All these approaches can be classified based on two characteristics:
 \begin{enumerate}
-	\item the system's model considered:
+	\item the type of system model:
 		\begin{itemize}
@@
 			\item data-based, that is the "black-box" approach, mainly done using Neural Networks: \cite{Safa2017SCS,Neto2008EB,Magnier2010BE,Candanedo2017EB,Ascione2017E,Cecconi2017EB,Li2016AE};
-			\item the only use of simulation tools, such as EnergyPlus \cite{energyPlus} and TRNSYS \cite{trnsys2000}: \cite{Yin2016EB,Christantoni2016EB};
+			\item simulation tool-based, such as EnergyPlus \cite{energyPlus} and TRNSYS \cite{trnsys2000}: \cite{Yin2016EB,Christantoni2016EB};
 		\end{itemize}
@@
 			\item only model creation and identification: \cite{Safa2017SCS,Neto2008EB,Magnier2010BE,Li2014EB,Li2016AE,Harb2016EB,Li2016EB,Candanedo2017EB,Cecconi2017EB,Ascione2017E};
-			\item control, typically Predictive Control, other than a modeling part: \cite{Shakouri2017SCS,Yoon2014EB,Yin2016EB,Salakij2016EB,Li2016E,Hu2017AE}.
+			\item model identification and control (typically Predictive Control): \cite{Shakouri2017SCS,Yoon2014EB,Yin2016EB,Salakij2016EB,Li2016E,Hu2017AE}.
 		\end{itemize}
 \end{enumerate}
-These papers are summarize in Table \ref{T:RelatedWork}, where all differences are highlighted.
-We also emphasized the case studies the results are applied to, and whether the authors used experimental data to simulate their algorithms.
-Only in 3 cases the algorithms were tested on real systems (in the table: RI - Real Implementation).
-We can see that, except for the last 6 cases, which we will discuss more in detail in the rest of this section, when model-based approaches or only tools are considered, then also control techiniques are applied.
-While with the data-driven modeling approaches, only model identification is addressed.
-This is because bridging data-driven models with control theory, especially with the predictive control, introduces several issues \cite{Hou2013}.
-Dealing with this kind of problems attracted a lot of researcher's attention in the past years.
-We consider 6 of such works (the last 6 papers of Table \ref{T:RelatedWork}) that are more related to the methodology presented in this paper.
+These papers are summarizes in Table \ref{T:RelatedWork} highlighting the differences.
+We also emphasize the case studies the results are applied to, and whether the authors use experimental data to simulate their algorithms.
+Only in three cases the algorithms are tested on real systems (see Table: RI - Real Implementation).
+We observe that except for the last 6 cases, which we discuss in detail, either model-based approaches or only tools are considered with/without control or data-driven modeling approaches without control.
+%This is because bridging data-driven models with control theory, especially with the predictive control, introduces several issues \cite{Hou2013}.
+Data-driven control has attracted increasing attraction in the recent years.
+The last six papers of Table \ref{T:RelatedWork} that are more related to the methodology presented in this paper.
 In particular, in \cite{Macarulla2017} the authors proposed a predictive control strategy based on Neural Networks, for boilers control in buildings, to decide the optimal time to switch-on the plan to guarantee energy saving and thermal comfort.
-However, the strategy is not easily scalable to different types of plants and does not use optimization theory in the closed-loop scheme.
+However, the approach is not easily scalable to different types of plants and does not use optimization in the closed-loop scheme.
 In \cite{Costanzo2016} a reinforcement learning control strategy, called Model-Assisted Batch Reinforcement Learning, is considered to provide data-driven control for the demand response problem in HVAC systems.
-Reinforcement Learning is a model-free methodology and for this reason represents an alternative approach, with pros and cons, to the MPC \cite{Ernst2009TSMC}.
+Reinforcement Learning is a model-free methodology and is an alternative approach to MPC \cite{Ernst2009TSMC} (with pros and cons).
 In \cite{Ferreira2012} the authors considered a data-driven predictive control based on Neural Networks to guarantee energy saving and  thermal comfort in public buildings.
 Neural Networks are used in the closed-loop control scheme to determine a thermal comfort index based on parameters that can be measured or estimated, but no Neural Networks-based system state dynamics are included into the optimization problem.
-However, Neural Networks are a particular type of machine learning algorithm that provides complex nonlinear equations to describe the system's dynamics.
-For example, \cite{Afram2017} considers a data-driven state model in the MPC closed-loop optimization.
-More precisely, Artificial Neural Networks are exploited to create data-driven models to be used as plant simulator in closed-loop with a Supervisory Model Predictive Control.
+%However, Neural Networks are a particular type of machine learning algorithm that provides complex nonlinear equations to describe the system's dynamics.
+\cite{Afram2017} uses a Neural Network based data-driven state model as a plant simulator in the MPC closed-loop optimization.
 More papers related to this topic can be found in the literature.
-Unfortunately, since Neural Networks models are nonlinear, the MPC based on such models is nonlinear as well.
-This means that a global optimal solution can not be provided.
-Furthermore, due to the high number of neurons that a Neural Network usually needs to provide good models, the nonlinear MPC problem can results into a quite complex optimization problem to be solved.
-To overcome this complexity problem, we proposed regression trees-based approaches in \cite{Behl2016,Jain2017TCPS}.
-However these approaches suffer of some drawbacks.
-In this paper, we provide a new methodology based on random forests, that overcome the drawbacks of our previous works.
-It provides robustness with respect to disturbance uncertainties and has extremely good performance when compared to other approaches (both optimal and rule-based).
-In the sext section we emphasize the differences with our previous approaches.
+Unfortunately, since Neural Networks models are nonlinear, the MPC based on such models is also nonlinear.
+This means that a global optimal solution can not be guaranteed. 
+When the complexity of network is high, solving the optimization problem becomes harder due to nonlinearties.
+To overcome this complexity, we introduced regression trees-based approach in \cite{Behl2016,Jain2017TCPS}.
+In this paper, we provide a new methodology based on random forests, that overcome the drawbacks of our previous work.
+It provides robustness with respect to disturbance uncertainties and has better performance when compared to other approaches (both optimal and rule-based).
+In the next section, we emphasize the differences with our previous approaches.
 }
@@
 \subsubsection{Previous work}\label{SSS:PreviousWork}
-In the following, the differences with respect to our previous papers are illustrated.
+We highlight the differences with respect to our previous work.
 \begin{itemize}
-	\item In \cite{Behl2016} we used regression trees and ensemble methods algorithms to create data-driven models for buildings. 
-	We then setup an MPC-like control problem using the regression trees-based model to be used in the Demand-Response problem.
-	However, ensemble methods have not been considered in the control process, but they have been used for identification purposes only.
-	Furthermore, the proposed models allowed to optimally control the system considering only one-step lookahead prediction.
-	Hence, it was not possible to use such approach to control the system considering a prediction over an horizon of arbitrary length.
-	\item This problem was addressed in \cite{Jain2017TCPS}, where the regression trees algorithm was modified to allow the tree to provide multiple-output.
-	Each output of the tree was then setup to provide the prediction of the system's behavior over different steps of the predictive horizon.
-	This allowed us to setup an MPC-like control problem considering an arbitrary horizon.
-	However, in this paper only single regression trees were considered, while ensemble methods were not took into account.
-	Modeling accuracy using single trees is strongly subject to the problems of data overfitting and high variance.
-	Such approach has the advantage to be extremely simple from the complexity point of view, but its application range is limited to specific kind of systems.
-	\item In this work, we extend preliminary results gave in the conference papers \cite{JainACC2017}, where a simplified version of this approach was compared with a model-based control approach on a trivial system, and \cite{JainCDC2017}, where a more complex model was considered, but without taking into account any experimental data or forecast inaccuracy.
-	In particular, an alternative approach is proposed to improve the system's identification accuracy and robustness of the previous approaches.
-	More precisely, instead of considering a single tree with multiple output, we consider multiple trees with single output.
-	Each tree is used to provide the prediction of the system's behavior over different steps of the horizon.
-	Furthermore, we extend this approach considering the random forests.
-	With a small increase in complexity, we setup an MPC-like control problem with arbitrary horizon, that provides better results and improves control robustness with respect to our previous work.
+	\item In \cite{Behl2016} we use regression trees to setup an MPC problem and apply it to the problem of Demand-Response.
+	However, these models can be used for optimal control with only one-step lookahead prediction.
+	Hence, it is not possible to use such approach to control the system considering a prediction over an horizon of arbitrary length.
+	\item This limitation is addressed in \cite{Jain2017TCPS}, where multi-output regression trees are used instead of single output trees.
+	The different outputs correspond to different steps of the prediction horizon.
+	This allows us to setup an MPC problem with a finite horizon.
+	Modeling accuracy using single trees is strongly affected by overfitting and high variance.
+	Such approach has the advantage to be extremely simple from the complexity point of view, but the range of applications is limited.
+	\item In \cite{JainACC2017,JainCDC2017}, we take a different approach that improves the system's identification accuracy and robustness in MPC.
+	Instead of considering a single tree/forest with multiple output, we consider multiple trees/forests with single output.
+	Each tree/forest provides the prediction of the system's behavior for a different time step of the horizon.
+	However, the results are based on simulated data and we did not account for inaccuracies in the weather forecast.
 \end{itemize}

### 2018-01-18 paper/final/dpc.tex
git merge

@@
 \label{S:dpc}
-TO IMPROVE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
-\begin{itemize}
-	\item Add more details
-	\item Review notation
-	\item Improve procedure and algorithm description
-	\item Add optimization setup discussion
-\end{itemize}
 
-%The central idea behind DPC is to obtain control-oriented models using machine learning or black-box modeling, and formulate the control problem in a way that receding horizon control (RHC) can still be applied and the optimization problem can be solved efficiently.
-%
-%Consider a black-box model given by $x_{k+1}=f(x_k,u_k,d_k)$, where $x,u,d$ represent states, inputs and disturbances, respectively.
-Depending upon the learning algorithm, $f$ is typically nonlinear, nonconvex and sometimes nondifferentiable (as is the case with regression trees and random forests) with no closed-form expression.
-%Such functional representations learned through black-box modeling may not be directly suitable for control and optimization as the optimization problem can be computationally intractable, or due to nondifferentiabilities we may have to settle with a sub-optimal solution using evolutionary algorithms \cite{Kusiak2009}.
-%These problems can be eliminated by decomposing 
-%\begin{align}
-%f(x_k,u_k,d_k)= g(x_k,h(u_k),d_k),
-%\label{E:sepvars}
-%\end{align}
-%where both $g$ and $h$ are learned using the data, and $h(u_k)$ is convex and differentiable, and thus suitable for optimization.
-DPC uses this functional decomposition or \textit{separation of variables} to overcome the aforementioned challenges with black-box optimization.
-%
-%\subsection{Separation of Variables}
-%\label{SS:sepvar}
-%
-%We distinguish between two sets of variables: control (or manipulated) variables $\tX^c \in \R^c$ and disturbance (or non-manipulated) variables $\tX^d \in \R^d$.
-The union of the two sets forms the full feature set for training, i.e.
-$\tX \equiv \tX^c \cup \tX^d \in \R^{c+d}$.
+\textcolor[rgb]{0,0,1}{The central idea behind DPC is to obtain control-oriented models using machine learning methodologies, and formulate the control problem in a way that Receding Horizon Control (RHC) can still be applied and the optimization problem can be solved efficiently.
+In this paper Regression Trees and Random Forests are considered (see Appendix ???
+for details).}
+
+\textcolor[rgb]{0,0,1}{To this aim, let an historical dataset $(\X,\Y)$, created using building measurements, be given to be used in the Regression Trees and Random Forests training process.
+Let $|(\X,\Y)| = n$, be the number of samples in the dataset.
+$\X = \{(x(k),u(k),d(k))\}$, $\forall k = 1,\ldots,n$, is the set of samples $(x(k),u(k),d(k))$, measured at each time instant $k$, where $x(k)\in\mathbb{R}^{n_x}$ is the vector of the state variables, $u(k)\in\mathbb{R}^{n_u}$ is the vector of the input variables and $d(k)\in\mathbb{R}^{n_d}$ is the vector of the disturbance variables.
+$\Y = \{y(k)\},\ \forall k = 1,\ldots,n$, is the set of samples $y(k)$, measured at each time instant $k$, where $y(k)\in\mathbb{R}^{n_y}$ is the vector of the output variables.
+We consider $\X$ as the set of predictor variables (or features) and $\Y$ as the set of response variables.
+As an example, the response variables can be represented by the power consumption or the room temperatures evolution, while the predictors can be represented by the weather forecast, as the disturbance, the set-points information and building schedules, as the inputs, and room temperatures, as the states.
+Our goal is to learn a data-driven models, using Regression Trees and Random Forests, that relate the value of the response variables with the value of the predictor variables, and that can be used to set up an MPC problem.
+}
+
+\textcolor[rgb]{0,0,1}{To do this, we want to derive a model with a closed-form expression of the following form:
+\begin{equation}\label{E:GenericModel}
+	y(k)=f(x(k),u(k),d(k)),
+\end{equation}
+where $f$ is an analytical expression that describes the system's dynamics.
+\textbf{We want to remark that $y$ can be considered as both a state variable evolution, i.e.
+$y(k) = x(k+1)$, and a variable that is function of $x$, $u$ and $d$ but is not one of them, as for example is the case of the power.
+We consider this aspect in Sections \ref{S:proof}, \ref{S:casestudy} and \ref{S:realCaseStudy}.
+IMPROVE!!!!!!!!!!!!!!!}.
+The issue that arises, when using Regression Trees and Random Forests, is that $f$ does not have a closed-form expression, hence \eqref{E:GenericModel} is not suitable for control and optimization.
+One possibility could be to look for sub-optimal solutions using evolutionary algorithms \cite{Kusiak2009}, i.e. using heuristics.
+In this section we provide a new methodology to determine a closed-form expression for $f$ by manipulating the dataset, as shown starting from Section \ref{SS:sepvar}.}
+
+\subsection{Dataset splitting}
+\label{SS:sepvar}
+
+\textcolor[rgb]{0,0,1}{The dataset splitting consists of partitioning the features set $\X$ into the sets $\X^c = \{u(k)\}\subset\X,\ \forall k=1,\ldots,n$, containing the control (or manipulated) variables, and $\X^d = \{(x(k),d(k))\}\subset\X,\ \forall k=1,\ldots,n$, containing the disturbance and state (or non-manipulated) variables.
+The union of the two sets forms the full feature set oftraining, i.e. $\X \equiv \X^c \cup \X^d$.
 %Our goal is to replace a model-based controller with a data-driven controller, where the latter depends only on the historical sensor data. 
 %These measurements could directly represent one or more states in the model-based control framework.
-We denote these as outputs $\tY \in \R$ for training, i.e.
-a $\tY$ represents a particular output and we can have separate models for multiple outputs.
-We define the number of training samples by $|(\tX,\tY)| = n$.
-%
-%Using separation of variables, the training process is divided into two steps.
-\textit{Step 1:} The trees and the ensembles are trained only on $\tX^d$, which eases the computational complexity.
-It is important to note that besides external disturbances, $\tX^d$ also contains autoregressive terms of the output $\tY$ which is the main reason for the state space explosion.
-\textit{Step 2:} Linear regression models are trained in the leaves (or terminal nodes) of the trees which are function of only $\tX^c$.
-We have validated this linear model assumption in \cite{Jain2016}.
+We denote these as outputs $\Y \in \R$ for training, i.e.
+a $\Y$ represents a particular output and we can have separate models for multiple outputs.
+We define the number of training samples by $|(\X,\Y)| = n$.
+Using this splitting methodology, the training process is divided into two steps:
+	\begin{enumerate}
+		\item the trees and the forests are trained only using $\X^d$, which also eases the computational complexity.
+			  It is important to note that besides external disturbance and the state measurements, $\X^d$ can also contains autoregressive terms of the output variables in $\Y$;
+		\item in each leaf of the tree (or in each leaf of each the tree of the forest), a linear regression model is trained as a function of the inputs variables in $\X^c$.
 		As we shall see in Section~\ref{SS:dpcrt} and \ref{SS:dpcrf}, the second step reduces the run-time control problem into a convex program.
-This process is illustrated in Figure~\ref{F:dpc-sepvars}.
-%
-%\subsection{DPC-RT: DPC with Regression Trees}
-%\label{SS:dpcrt}
-%When the data has lots of features, which interact in complicated, nonlinear ways, assembling a single global model such as linear or polynomial regression can be difficult, and can lead to poor response predictions.
-%An approach to non-linear regression is to partition the data space into smaller regions, where the interactions are more manageable. 
-%This partition is repeated recursively until finally we get to small chunks of the data space where we can fit simple (eg. linear parametric) models. 
+	\end{enumerate}
+This process is illustrated in Figure~\ref{F:dpc-sepvars} (left).
+For the sake of simplicity only trees are considered, but the same holds for the forests.
+The meaning of the variable $j$ in the linear model in the figure will be clear in the next section.
+In Figure~\ref{F:dpc-sepvars} (right), the step 1 of the training process uses also the input variables, i.e. the classical training process without the dataset splitting.
+The reason why this latter approach is not suitable for control will be discussed at the end of Section \ref{SS:dpcrt}.
+}
+
+\subsection{DPC-RT: DPC with Regression Trees}
+\label{SS:dpcrt}
+When the data has lots of features, which interact in complicated, nonlinear ways, assembling a single global model such as linear or polynomial regression can be difficult, and can lead to poor response predictions.
+\textcolor[rgb]{0,0,1}{As discussed before, an approach to non-linear regression is to partition the data space into smaller regions, where the interactions are more manageable. }
+This partition is repeated recursively until we finally get to small chunks of the data space where we can fit simple (eg. linear parametric) models. 
 %Therefore, in \eqref{E:sepvars}, the global model $f$ has two parts: the recursive partition $g$, and a linear (and convex) model $h$ for each cell of the partition.
-%
-%\begin{figure}[t!]
-%	\centering
-%	\includegraphics[width=20pc]{figures/dpc-sepvars.eps}
-%	\caption{Separation of variables.
-\textit{Step 1:} Tree $\mathcal{T}_1$ is trained only on the disturbances $\tX_d$ as the features.
-Tree $\mathcal{T}_2$ uses both the disturbances $\tX_d$ and the control variables $\tX_c$ for splitting and is thus not computationally suitable for control.
-\textit{Step 2:} In the leaf $R_i$ of the trees, a linear regression model parametrized by $\beta_i$ is defined as a function only of the control variables.}
-%	\captionsetup{justification=centering}
-%	\label{F:dpc-sepvars}
-%\end{figure}
-%
-%Now, our goal is to predict the state $\tY$ at time $k$ for next $N$ time steps, i.e.
-$\tY_{\mathrm{k+1|k}},\dots,\tY_{\mathrm{k+N|k}}$, where $N$ is the control horizon.
-Applying the separation of variables, we build $N$ regression trees using CART procedure \cite{Breiman1984} such that the output $\tY_{\mathrm{k+j|k}}$ of the $j^{th}$ tree depends upon the previous $N$ disturbances:
-%\begin{gather}
-%\label{E:model_tree}
-%\tY_{\mathrm{k+j|k}} = \mathit{f}_{\mathrm{tree}} \left( \tX^d_{\mathrm{k+j-N|k}} ,\dots,\tX^d_{\mathrm{k+j-1|k}}  \right), \\ \nonumber
-%\mathsf{X}^d_{\mathrm{k+j-l|k}} \in \R^d \  \ \forall \ l,j=1,\dots,N.
-%\end{gather}
-%Then, the linear models as functions of $\tX^c$ in each leaf of the tree $\mathcal{T}_j$ are defined as
-%\begin{gather}
-%\label{E:model_leaf}
-%\tY_{\mathrm{k+j|k}} =  \beta^T_j [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T, \\ \nonumber
-%\mathsf{X}^c_{\mathrm{k+j-l|k}} \in \R^c \  \ \forall \ l,j=1,\dots,N.
-%\end{gather}
-%Note that the coefficients $\beta_j$ would be different for each leaf.
-Eq.~\eqref{E:model_leaf} implies that the prediction of output $\tY_{\mathrm{k+j}}$ at time $k$ is an affine combination of control inputs from time $k$ to $k+j-1$.
+
+\begin{figure}[t!]
+	\centering
+	\includegraphics[width=20pc]{figures/dpc-sepvars.eps}
+	\caption{Dataset splitting.
+\textit{Step 1:} Tree $\mathcal{T}_1$ is trained only with variables in $\X^d$ as the features.
+Tree $\mathcal{T}_2$ uses both the disturbances and states in $\X^d$, and the control variables in $\X_c$ for splitting.
+It is thus not computationally suitable for control.
+\textit{Step 2:} In the leaf $\ell_i$ of the trees, a linear regression model parametrized by $\beta_i$ is defined as a function only of the control variables.}
+	\captionsetup{justification=centering}
+	\label{F:dpc-sepvars}
+\end{figure}
+
+\textcolor[rgb]{0,0,1}{Now, in order to have a model that can be used for prediction over an arbitrary long horizon in an MPC problem, our goal is to predict, at time $k$, the output $y$ for next $N$ time steps, i.e.
+$y(k),\ldots,y(k+N)$, where $N$ is the control horizon, given the state measurements, i.e.
+$x(k)$, and the disturbance forecast, i.e.
+$\tilde d(k),\ldots,\tilde d(k+N)$, at time $k$.
+For the sake of simplicity and without any loss of generality, in the following, we consider only a single output, i.e. $y(k)\in\mathbb{R}$ (or $n_y = 1$).
+However, as we will see in Sections \ref{S:casestudy} and \ref{S:realCaseStudy}, multiple trees (or forests) can be built for multiple outputs.
+Applying the dataset splitting, we build $N$ regression trees using CART procedure \cite{Breiman1984} (see also Appendix for more details), such that the output $y(k+j)$ of the $j^{th}$ tree depends upon the previous $\delta_d$ disturbances, and on the state at instant $k$ and its $\delta_x$ regressive terms:
+\begin{equation}\label{E:model_tree}
+\T_j = \mathit{f}_{\mathrm{tree}} \left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right),\ j=1,\ldots,N,
+\end{equation}
+with $\left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right)\in\X^d,\ k=1,\ldots,n$, and $\mathit{f}_{\mathrm{tree}}$ representing the tree structure.
+Then, the linear models, as functions of the inputs variables $u\in\X^c$, in each leaf of the trees $\T_j,\ j=1,\ldots,N$, are defined as
+\begin{equation}\label{E:model_leaf}
+y(k+j) =  \beta_j [1,u(k),\ldots,u(k+j) ]^T,\ j=1,\ldots,N.
+\end{equation}}
+Note that the coefficients $\beta_j$ are different for each leaf.
+Equation~\eqref{E:model_leaf} implies that the prediction of output $y(k+j)$ at time $k$ is an affine combination of control inputs from time $k$ to $k+j$.
 Thus, we have managed to linearize the original model dynamics via black-box modeling.
-This two-step training is done offline.
-In run-time, given the disturbances $\tX^d_\mathrm{k|k}$ at time $k$, we can narrow down to a leaf of each tree in \eqref{E:model_tree} to retrieve the linear models in \eqref{E:model_leaf}.
-%
-%In run-time, when a new control action is to be determined, each tree (prediction step) contributes to a linear constraint in the optimization as a replacement for the state dynamics in the case of MPC.
-Thus, the RHC optimization problem with a quadratic cost ($\mathcal{Q} \geq 0, \mathcal{R} \succeq 0$) can be formulated as:
-%\begin{align}
-%\begin{aligned}
-%\text{min } & \sum_{j=1}^{N} ({\tY}_{\mathrm{k+j|k}})^2 \mathcal{Q} + {\tX^c}^T_{\mathrm{k+j-1|k}} \mathcal{R} {\tX^c}_{\mathrm{k+j-1|k}} +  \lambda\epsilon_j\\
-%\text{s.~t. } & \ \ \ \ \ \tY_{\mathrm{k+j|k}} =  \beta^T [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
-%& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tX}^c \leq \tX^c_{\mathrm{k+j-1|k}} \leq \bar{\tX}^c\\ 
-%& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tY}-\epsilon_j \leq \tY_{\mathrm{k+j|k}} \leq \bar{\tY} + \epsilon_j\\\
-%& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0, \ j = 1,\dots,N.
-%\end{aligned}
-%\label{E:dpcrt}
-%\end{align}
-%Here, $\mathcal{Q} \in \mathbb{R}$ and $\mathcal{R} \in \mathbb{R}^{c \times c}$, and the slack variables $\epsilon_j$ ensure recursive feasibility since the equality constraint on $\tY$ is relaxed.
-Of course, a different cost function can be chosen depending upon the application.
-In the current formulation, the data-driven control problem is reduced to a convex program which is much easier to solve than running an optimization directly on a black-box model trained on $\tX^c$ as features.
-We solve this optimization in the same manner as MPC to determine the optimal sequence of inputs $[\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+N-1|k}}]$, apply the first control input $\tX^c_{\mathrm{k|k}}$ and proceed to the next time step $k+1$.
+\textcolor[rgb]{0,0,1}{One of the advantages of this approach is that this two-step training is done off-line, so the time required to create the models does not affect the control execution in run-time.}
+
+\textcolor[rgb]{0,0,1}{The problem now is: \emph{how to use this modeling framework to set up an MPC problem?} In run-time, given the disturbance forecast and the state measurements at time $k$, i.e.
+$\tilde d(k+j-\delta_d),\ldots,\tilde d(k+j),x(k-\delta_x),\ldots,x(k)$ given $k$, we can narrow down to a leaf of each tree in \eqref{E:model_tree} to retrieve the linear models in \eqref{E:model_leaf} for each step $j=1,\ldots,N$ of the horizon (see first part of the pseudo code given in Algorithm \ref{A:dpcrt}), as depicted in Figure \ref{F:dpc-sepvars} (left).
+These models are used to setup the MPC Problem \ref{P:dpcrt}.
+In particular, }when a new control action has to be determined, each tree (prediction step) contributes to a linear constraint in the optimization as a replacement for the state dynamics in the case of MPC.
+Thus, the MPC optimization problem, considering a quadratic cost ($Q \succeq 0, R \succeq 0$) function and the general case of multiple outputs, i.e.
+$y(k)\in\mathbb{R}^{n_y},\ n_y\geq 1$, can be formulated as:
+%\begin{problem}\label{P:dpcrt}
+%	\begin{equation}
+%		\begin{aligned}
+%		\text{min } & \sum_{j=1}^{N} ({\Y}_{\mathrm{k+j|k}})^2 \mathcal{Q} + {\X^c}^T_{\mathrm{k+j-1|k}} \mathcal{R} {\X^c}_{\mathrm{k+j-1|k}} + \lambda\epsilon_j\\
+%		\text{s.~t. } & \ \ \ \ \ \Y_{\mathrm{k+j|k}} =  \beta^T [1,\X^c_{\mathrm{k|k}},\dots,\X^c_{\mathrm{k+j-1|k}} ]^T \\
+%		& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{\X}^c \leq \X^c_{\mathrm{k+j-1|k}} \leq \bar{\X}^c\\ 
+%		& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\Y}-\epsilon_j \leq \Y_{\mathrm{k+j|k}} \leq \bar{\Y} + \epsilon_j\\\
+%		& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0, \ j = 1,\dots,N.
+%		\end{aligned}
+%		\label{E:dpcrt}
+%	\end{equation}
+%\end{problem}
+\begin{problem}\label{P:dpcrt}
+	\begin{equation}
+	\begin{aligned}
+	& \underset{u_{k+j}}{\text{minimize}} & & \sum_{j=1}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
+	& \text{subject to }                  & & y_{k+j}     =   \beta_j [1,u_{k},\ldots,u_{k+j} ]^\top                             \\
+	&                                     & & u_{k+j}    \in  \mathcal{U}                                                        \\
+	&                                     & & |y_{k+j}|  \leq \bar{y}_{k+j} + \epsilon_j 										 \\
+	&                                     & & \epsilon_j \geq  0							                                     \\
+	&                                     & & j           =    1,\ldots,N.            									         \\
+	\end{aligned}
+	\label{E:dpcrt}
+	\end{equation}
+\end{problem}
+
+\textcolor[rgb]{0,0,1}{Here, $Q \in \mathbb{R}^{n_y\times n_y}$ and $R \in \mathbb{R}^{n_u\times n_u}$ are weight matrices used to give more importance in the minimization to either $y$ or $u$.
+This problem is solved as in the classical MPC formulation, i.e.
+at each time step $k=1,2,\ldots$ the optimal control sequence $u^*_k,\ldots,u^*_{k+N}$ is computed, and only the first input of the sequence is applied as control input to the system: $u(k) = u^*_k$.
+The slack variables $\epsilon_j$ are added to ensure recursive feasibility (since the equality constraint on $y$ is relaxed), i.e.
+to guarantee that Problem \ref{P:dpcrt} can provide a solution at each step $k=1,2,\ldots$.
+$\lambda$ is a weight used to give more importance to bounds violation with respect to $y$ than to the cost function optimization, and viceversa. 
+Of course, a different cost function can be chosen depending upon the application, i.e. it can be linear, nonlinear, etc., obviously changing the complexity of the problem.
+In the current formulation, the data-driven control problem is reduced to a convex program which is very easy and efficient to solve.}
 The pseudo code for DPC-RT is given in Algorithm~\ref{A:dpcrt}.
-%
-%\begin{algorithm}[t!]
-%	\caption{Data Predictive Control with Regression Trees}
-%	\label{A:dpcrt}
-%	\begin{algorithmic}[1]
-%		\State \textsc{Design Time}
-%		\Procedure{Model Training using Separation of Vars}{}
-%		\State Set $\mathsf{X}^c$ $\gets$ manipulated features
-%		\State Set $\mathsf{X}^d$ $\gets$ non-manipulated features
-%		\State Build $N$ predictive trees with $(\tY,\tX^d)$ defined in \eqref{E:model_tree}
-%		\ForAll{trees $\mathcal{T}_j$}
-%		\ForAll{regions $R_i$ at the leaves of $\mathcal{T}_j$}
-%		\State Fit $ \tY_{\mathrm{k+j|k}} =  \beta^T_j \left[1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} \right]^T$ as in \eqref{E:model_leaf}
-%		\EndFor
-%		\EndFor
-%		\EndProcedure
-%		\State \textsc{Run Time}
-%		\Procedure{Predictive Control}{}
-%		\While{$k< k_{\mathrm{stop}}$}
-%		\ForAll{trees $\mathcal{T}_j$}
-%		\State Determine the leaf $R_i$ using $\tX^d$ as in \eqref{E:model_tree}
-%		\State Obtain the linear model at $R_{i}$ trained in \eqref{E:model_leaf}
-%		\EndFor
-%		\State Solve optimization in \eqref{E:dpcrt} to determine optimal
-%		\State control actions $[\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+N-1|k}}]$
-%		\State Apply the first input $\tX^c_{\mathrm{k|k}}$
-%		\EndWhile
-%		\EndProcedure
-%	\end{algorithmic}
-%\end{algorithm}
-%
-%\subsection{DPC-En: DPC with Ensemble Methods}
-%\label{SS:dpcrf}
-%Regression trees obtain good predictive accuracy in many domains. However, the models used in their leaves have some limitations regarding the kind of functions they are able to approximate.
-%The problem with trees is their high variance and that they can overfit the data easily.
-%A small change $\delta$ in the data can result in a different series of splits and thus violate the acceptable accuracy $\eps$.
-%This is the price to be paid for estimating a tree-based structure from the data.
-%
-%
-%We use ensemble methods \cite{Friedman2001} to combine the predictions of several independent regression trees in order to improve generalizability and robustness over a single estimator. 
-%The essential idea is to average many noisy trees to reduce the overall variance in prediction.
-%We inject randomness into the tree construction in two ways. First, we randomize the features used to define splitting in each tree.
-%Second, we build each tree using a bootstrapped or sub-sampled data set.
-%In this way, each tree in the forest is trained on different data, which introduces differences between the trees.
-More explicitly, training features $\tX^d \in \R^p$ with $p<d$ and the in-bag samples (in-bag samples correspond to the data samples on which the tree was trained) are different for each tree in the forest i.e $|(\tX,\tY)|<n$.
-%
-%The goal with DPC-En is to replace each tree in Algorithm~\ref{A:dpcrt} by a forest
-%\begin{gather}
-%\label{E:model_forest}
-%\tY_{\mathrm{k+j|k}} = \mathit{f}_{\mathrm{forest}} \left( \tX^d_{\mathrm{k+j-N|k}} ,\dots,\tX^d_{\mathrm{k+j-1|k}}  \right), \\ \nonumber
-%\mathsf{X}^d_{\mathrm{k+j-l|k}} \in \R^p \  \ \forall \ l,j=1,\dots,N,
-%\end{gather}
-%which, again, is trained only on $\mathsf{X}^d$, with $\tX^d \in \R^p \subset \R^d$ for each tree,  and then fit a linear regression model using $\mathsf{X}^c$ in every leaf of every tree.
-We build $N$ such forests for $N$ prediction steps such that the leaf $R_i$ of forest $\mathcal{R}_j$ uses a linear model
-%\begin{gather}
-%\label{E:model_leaf_forest}
-%\tY_{\mathrm{k+j|k}} =  \Theta^T_{ij} [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T, \\ \nonumber
-%\mathsf{X}^c_{\mathrm{k+j-l|k}} \in \R^c \  \ \forall \ l,j=1,\dots,N.
-%\end{gather}
-%Here $(\mathsf{X}^c,\mathsf{Y})$ correspond to the in-bag samples for the trees.
-%
-%While the offline training burden in DPC-En is slightly increased compared to DPC-RT, in the control step we exploit the better accuracy, and lower variance properties of the random forest. 
-%If a forest has $t$ number of trees, given the forecast of disturbances, we have $t$ sets of linear coefficients.
+\textcolor[rgb]{0,0,1}{DPC algorithm is also graphically shown in Figure \ref{F:dpc-algo-rf} for the random forest case, but the same holds for single regression trees.}
+
+\textcolor[rgb]{0,0,1}{\begin{remark}
+	 If we did not consider the splitting procedure	of the dataset, and we also used input variables to learn the trees, as in the right side of Figure 1, the resulting
+	 model would not have been suitable for control.
+This is because, since u is the variable we want to optimize, we do not know its value a priori to go through the trees and determine the correct leaves to find the models to use in the RHC problem.
+\end{remark}}
+
+\begin{algorithm}[t!]
+	\caption{Data Predictive Control with Regression Trees}
+	\label{A:dpcrt}
+	\begin{algorithmic}[1]
+		\State \textsc{Design Time (Off-Line)}
+		\Procedure{Model Training using Dataset Splitting}{}
+		\State Set $\X^c$ $\gets$ manipulated features
+		\State Set $\X^d$ $\gets$ non-manipulated features
+		\State Build $N$ predictive trees with $(\X^d,\Y)$ as defined in \eqref{E:model_tree}
+		\ForAll{trees $\mathcal{T}_j$}
+		\ForAll{the leaves $\ell_i$ of $\mathcal{T}_j$}
+		\State Fit $ y(k+j) =  \beta_j \left[1,u(k),\ldots,u(k+j) \right]^T$ as in \eqref{E:model_leaf}
+		\EndFor
+		\EndFor
+		\EndProcedure
+		\State \textsc{Run Time}
+		\Procedure{Predictive Control}{}
+		\While{$k< k_{\mathrm{stop}}$}
+		\ForAll{trees $\mathcal{T}_j$}
+		\State Determine the leaf $\ell_i$ using $\X^d$ as in \eqref{E:model_tree}
+		\State Obtain the linear model at $\ell_i$ trained in \eqref{E:model_leaf}
+		\EndFor
+		\State Solve optimization problem \ref{P:dpcrt} to determine optimal
+		\State control actions $u^*_k,\ldots,u^*_{k+j}$
+		\State Apply the first input $u(k)=u^*_k$
+		\EndWhile
+		\EndProcedure
+	\end{algorithmic}
+\end{algorithm}
+
+\subsection{DPC-En: DPC with Ensemble Methods}
+\label{SS:dpcrf}
+Regression trees obtain good predictive accuracy in many domains. However, the models used in their leaves have some limitations regarding the kind of functions they are able to approximate.
+The problem with trees is their high variance and that they can overfit the data easily.
+A small change $\delta$ in the data can result in a different series of splits and thus violate the acceptable accuracy.
+This is the price to be paid for estimating a tree-based structure from the data.
+
+We use ensemble methods \cite{Friedman2001}, in particular Random Forests (see Appendix ???
+for details), to combine the predictions of several independent regression trees in order to improve generalizability and robustness over a single estimator.
+The essential idea is to average many noisy trees to reduce the overall variance in prediction.
+We inject randomness into the tree construction in two ways. First, we randomize the features used to define splitting in each tree.
+Second, we build each tree using a bootstrapped or sub-sampled data set.
+In this way, each tree in the forest is trained on different data, which introduces differences between the trees.
+\textcolor[rgb]{0,0,1}{More explicitly, training features $\bar{\X}^d_i\subset\X^d$, $i=1,2,...$ and the in-bag samples (in-bag samples correspond to the data samples on which the tree was trained) are different for each tree in the forest, i.e $|(\bar \X^d_i,\Y)|<n$.}
+
+\textcolor[rgb]{0,0,1}{The goal with DPC-En is to replace each tree in Algorithm~\ref{A:dpcrt} by a forest.
+This can be done replacing Equations \eqref{E:model_tree} and \eqref{E:model_leaf} with \eqref{E:model_forest} and \eqref{E:model_leaf_forest} in the following:
+\begin{equation}\label{E:model_forest}
+\F_j = \mathit{f}_{\mathrm{forest}} \left( d(k+j-\delta_d),\ldots,d_(k+j),x(k-\delta_x),\ldots,x(k)  \right),\ j=1,\ldots,N,
+\end{equation}
+with $\left( d_{k+j-\delta_d},\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right)\in\X^d,\ k=1,\ldots,n$, and $\mathit{f}_{\mathrm{forest}}$ representing the forest structure.
+In particular, as mentioned above, each tree $\T_i$ of each forest $\F_j$ is trained using $(\bar \X^d_i,\Y)$.
+Then the linear models, as functions of the inputs variables $u\in\X^c$, in each leaf of each tree of the forests $\F_j,\ j=1,\ldots,N$, are defined as
+\begin{equation}\label{E:model_leaf_forest}
+y(k+j) =  \Theta_{ij} [1,u(k),\ldots,u(k+j) ]^T,\ j=1,\ldots,N.
+\end{equation}
+Here $(\X^c,\Y)$ correspond to the in-bag samples for the trees.}
+
+While the off-line training burden in DPC-En is slightly increased compared to DPC-RT, in the control step we exploit the better accuracy, and lower variance properties of the random forests. 
+If a forest has $t$ number of trees, given the forecast of disturbances, we have $t$ sets of linear coefficients.
 We simply average out all the coefficients from all the trees to get one linear model represented by $\hat{\Theta}_j$ for each forest.
-Note that the averaging step can only be done in run-time because the leaf of each tree can be narrowed down only when the $\tX^d$ is known.
-Thus, for $N$ forests, we again have exactly $N$ linear equality constraints in the optimization problem below:
-%\begin{align}
-%\begin{aligned}
-%\text{min } & \sum_{j=1}^{N} ({\tY}_{\mathrm{k+j|k}})^2 \mathcal{Q} + {\tX^c}^T_{\mathrm{k+j-1|k}} \mathcal{R} {\tX^c}_{\mathrm{k+j-1|k}} +  \lambda\epsilon_j\\
-%\text{s.~t. } & \ \ \ \ \ \tY_{\mathrm{k+j|k}} =  \hat{\Theta}^T_j [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
-%& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tX}^c \leq \tX^c_{\mathrm{k+j-1|k}} \leq \bar{\tX}^c\\ 
-%& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tY}-\epsilon_j \leq \tY_{\mathrm{k+j|k}} \leq \bar{\tY} + \epsilon_j\\\
-%& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0, \ j = 1,\dots,N.
-%\end{aligned}
-%\label{E:dpcrf}
-%\end{align}
-%DPC-En is graphically described in Figure~\ref{F:dpc-algo-rf}.
+Note that the averaging step can only be done in run-time, because the leaf of each tree can be narrowed down only when the $\X^d$ is known.
+Thus, for $N$ forests, we again have exactly $N$ linear equality constraints as in the optimization problem below:
+
+\begin{problem}\label{P:dpcrf}
+	\begin{equation}
+	\begin{aligned}
+	& \underset{u_{k+j}}{\text{minimize}} & & \sum_{j=1}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
+	& \text{subject to }                  & & y_{k+j}      =  \hat{\Theta}_j [1,u_{k},\ldots,u_{k+j} ]^\top                      \\
+	&                                     & & u_{k+j}    \in  \mathcal{U}                                                        \\
+	&                                     & & |y_{k+j}|  \leq \bar{y}_{k+j} + \epsilon_j 										 \\
+	&                                     & & \epsilon_j \geq  0							                                     \\
+	&                                     & & j           =    1,\ldots,N.            									         \\
+	\end{aligned}
+	\label{E:dpcrf}
+	\end{equation}
+\end{problem}
+DPC-En is graphically described in Figure~\ref{F:dpc-algo-rf}.
+
 The ensemble data predictive control (DPC-En) is the first such method to bridge the gap between ensemble predictive models (such as random forests) and receding horizon control.
 In the next section, we compare DPC-RT and DPC-En to MPC for a building model.
-%
-%\begin{figure}[t!]
-%	\centering
-%	\includegraphics[width=26pc]{figures/dpc-algo-rf.eps}
-%	\caption{DPC-En: At time $k$, the algorithm uses the forecast of disturbances $\tX^d_{\mathrm{k|k}}$ to select linear models $\Theta_1$ to $\Theta_t$ in the leaves of each ensemble.
-The linear models in each ensemble are averaged to calculate a single model represented by $\hat{\Theta}_j$ which act as constraints in the optimization problem.
-The optimal sequence $[\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+N-1|k}}]$, of which the first one is applied, and $\tX^d_{\mathrm{k+1|k+1}}$ is calculated to proceed to $k+1$.}
-%	\label{F:dpc-algo-rf}
-%\end{figure}
+
+\begin{figure}[t!]
+	\centering
+	\includegraphics[width=26pc]{figures/dpc-algo-rf.eps}
+	\caption{\textcolor[rgb]{0,0,1}{DPC-En: At time $k$, the algorithm uses the forecast of disturbances and the state measurements in $\X^d$ at time $k$ to select linear models $\Theta_1$ to $\Theta_t$ in the leaves of each forest.
+The linear models in each forest are averaged to calculate a single model represented by $\hat{\Theta}_j$, and act as constraints in the optimization problem.
+The optimal sequence $u^*_k,\ldots,u^*_{k+N}$, of which only the first element is applied, i.e.
+$u(k)=u^*_k$, and variables in $\X^d$ at time $k+1$ are calculated to proceed to next step at $k+1$.}}
+	\label{F:dpc-algo-rf}
+\end{figure}

### 2018-01-18 paper/final/dpc.tex
Edited Section 2

@@
 to guarantee that Problem \ref{P:dpcrt} can provide a solution at each step $k=1,2,\ldots$.
-$\lambda$ is a weight used to give more importance to bounds violation with respect to $y$ than to the cost function optimization, and viceversa. 
+$\lambda$ is a weight used to give more importance to bounds violation with respect to $y$, than to the cost function optimization, and viceversa. 
 Of course, a different cost function can be chosen depending upon the application, i.e. it can be linear, nonlinear, etc., obviously changing the complexity of the problem.
@@
 
-We use ensemble methods \cite{Friedman2001}, in particular Random Forests (see Appendix ???
-for details), to combine the predictions of several independent regression trees in order to improve generalizability and robustness over a single estimator.
+We use ensemble methods \cite{Friedman2001}, in particular Random Forests, to combine the predictions of several independent regression trees in order to improve generalizability and robustness over a single estimator.
 The essential idea is to average many noisy trees to reduce the overall variance in prediction.
@@
 In this way, each tree in the forest is trained on different data, which introduces differences between the trees.
-\textcolor[rgb]{0,0,1}{More explicitly, training features $\bar{\X}^d_i\subset\X^d$, $i=1,2,...$ and the in-bag samples (in-bag samples correspond to the data samples on which the tree was trained) are different for each tree in the forest, i.e $|(\bar \X^d_i,\Y)|<n$.}
+\textcolor[rgb]{0,0,1}{More explicitly, training features $\bar{\X}^d_i\subset\X^d$, $i=1,2,...$ and the in-bag samples (in-bag samples correspond to the data samples which each tree was trained on) are different for each tree in the forest, i.e $|(\bar \X^d_i,\bar\Y_i)|<n$, with $\bar\Y_i\subset\Y$.}
 
 \textcolor[rgb]{0,0,1}{The goal with DPC-En is to replace each tree in Algorithm~\ref{A:dpcrt} by a forest.
-This can be done replacing Equations \eqref{E:model_tree} and \eqref{E:model_leaf} with \eqref{E:model_forest} and \eqref{E:model_leaf_forest} in the following:
+This can be done replacing Equations \eqref{E:model_tree} and \eqref{E:model_leaf} with the Equations \eqref{E:model_forest} and \eqref{E:model_leaf_forest} provided in the following:
 \begin{equation}\label{E:model_forest}
-\F_j = \mathit{f}_{\mathrm{forest}} \left( d(k+j-\delta_d),\ldots,d_(k+j),x(k-\delta_x),\ldots,x(k)  \right),\ j=1,\ldots,N,
+\F_j = \mathit{f}_{\mathrm{forest}} \left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right),\ j=1,\ldots,N,
 \end{equation}
-with $\left( d_{k+j-\delta_d},\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right)\in\X^d,\ k=1,\ldots,n$, and $\mathit{f}_{\mathrm{forest}}$ representing the forest structure.
-In particular, as mentioned above, each tree $\T_i$ of each forest $\F_j$ is trained using $(\bar \X^d_i,\Y)$.
-Then the linear models, as functions of the inputs variables $u\in\X^c$, in each leaf of each tree of the forests $\F_j,\ j=1,\ldots,N$, are defined as
+with $\left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right)\in\X^d,\ k=1,\ldots,n$, and $\mathit{f}_{\mathrm{forest}}$ representing the forest structure.
+In particular, as mentioned above, each tree $\T_i$ of each forest $\F_j$ is trained using $(\bar \X^d_i,\bar\Y_i)$.
+Then the linear models, as functions of the inputs variables $u\in\bar\X^c_i\subset\X_c$, in each leaf of each tree of the forests $\F_j,\ j=1,\ldots,N$, are defined as
 \begin{equation}\label{E:model_leaf_forest}
@@
 \end{equation}
-Here $(\X^c,\Y)$ correspond to the in-bag samples for the trees.}
+Here $(\bar\X^c_i,\bar\Y_i)$ correspond to the in-bag samples for the trees.}
 

### 2018-01-19 paper/final/casestudyReal.tex
small changes in Sec.5

@@
 The problem is set up as follows:
+\begin{problem}\label{P:dpcRealCase}
 \begin{align}
@@
 \end{align}
+\end{problem}
 The choice of different weights $Q$, $\lambda_{min}$ and $\lambda_{max}$ allows the designer to give more importance to energy consumption rather than temperature comfort and vice versa.
@@
 \end{figure}
+%\begin{table}[t!]
+%	\centering
+%	%	\scalebox{0.9}{
+%	\begin{tabular}{lccccc}
+%		\toprule
+%		CONTROLLER  & LBV $\mathrm{MBE}$  & LBV $\mathrm{RMSE}$ & UBV $\mathrm{MBE}$ & UBV $\mathrm{RMSE}$ 	\\ 
+%		\midrule
+%		$DPC-SV$    & $0.013$             & $0.028$  			      & $0$    				 & $0$     	  	\\
+%		$DPC-MV$    & $0.165$ 			  & $0.146$     			  & $0$    				 & $0$		  	\\
+%		$DPC-LV$    & $0.410$  			  & $0.224$     			  & $0$    				 & $0$	      	\\
+%		$Bang-bang$ & $0.049$ 			  & $0.080$    				  & $0.0063$ 		     & $0.0120$	  	\\
+%		\bottomrule
+%	\end{tabular}
+%	%	}
+%	\caption{Lower Bound Violation (LBV) and Upper Bound Violation (UBV) errors expressed as $\mathrm{MBE}\%$ and $\mathrm{CV(RMSE)}\%$ for DPC Small Violation (DPC-SV), DPC Medium Violation (DPC-MV), DPC large Violation (DPC-LV) and bang-bang controller.}
+%	\captionsetup{justification=centering}
+%	\label{T:violationErrors}
+%\end{table}
 \begin{table}[t!]
@@
 		\midrule
-		$DPC-SV$    & $0.013$             & $0.028$  			      & $0$    				 & $0$     	  	\\
-		$DPC-MV$    & $0.165$ 			  & $0.146$     			  & $0$    				 & $0$		  	\\
-		$DPC-LV$    & $0.410$  			  & $0.224$     			  & $0$    				 & $0$	      	\\
-		$Bang-bang$ & $0.049$ 			  & $0.080$    				  & $0.0063$ 		     & $0.0120$	  	\\
+		$DPC-SV$    & $0.013$             & $0.092$  			      & $0$    				 & $0$     	  	\\
+		$DPC-MV$    & $0.165$ 			  & $0.479$     			  & $0$    				 & $0$		  	\\
+		$DPC-LV$    & $0.410$  			  & $0.733$     			  & $0$    				 & $0$	      	\\
+		$Bang-bang$ & $0.0485$ 			  & $0.265$    				  & $0.0063$ 		     & $0.040$	  	\\
 		\bottomrule
@@
 \subsubsection{Weather forecast subject to uncertainty}\label{SSS:DisturbanceUncertain}
-In \cite{Petersen2014AE} the authors investigated, through a large-scale simulation study, the effect of weather forecast uncertainty on the performance of MPC in building systems operations, and compared to the performance of a rule-base strategy.
+In \cite{Petersen2014AE} the authors investigated, through a large-scale simulation study, the effect of weather forecast uncertainty on the performance of MPC in building systems operations, and compared to the performance of a rule-based strategy.
 They considered 48 different scenarios of uncertainties on 72 hours weather forecasts.
-On such long horizon, results have shown that, with few exceptions, MPC outperforms the rule based controller in case of energy savings and/or thermal indoor environment, and is in general quite close to the perfect forecast case, despite the uncertainty in weather forecast.}
+On such long horizon, results have shown that, with few exceptions, that MPC outperforms the rule-based controller in case of energy savings and/or thermal indoor environment, and is in general quite close to the perfect forecast case, despite the uncertainty in weather forecast.}
 
 \textcolor[rgb]{0,0,1}{Horizon length for the DPC algorithm presented in this paper is usually much shorter, as for example $6\ hours$ in Section \ref{S:proof}, $7\ hours$ in Section \ref{S:casestudy}, and $40\ minutes$ in Section \ref{S:realCaseStudy}.
-This means the weather forecast is more accurate than the 72 hours case presented in \cite{Petersen2014AE}, hence also control performance is even closer to the perfect knowledge forecast case.}
+This means the weather forecast is more accurate than the 72 hours case presented in \cite{Petersen2014AE}, hence also control performance is even closer to the perfect knowledge forecast case.
+Nevertheless, to show the robustness of our approach, in this section we provide simulations considering noisy weather forecast.}
 
-\textcolor[rgb]{0,0,1}{Nevertheless, to show the robustness of our approach, in this section we provide simulations considering noisy weather forecast.
-To this aim, we add noises with normal distribution on the disturbance variables in $\tX^d$ used in the DPC problem \eqref{E:DPCrealcase}, i.e.
+\textcolor[rgb]{0,0,1}{To this aim, we add noises with normal distribution on the disturbance variables in $\X^d$, used to obtain parameters $\hat \Theta$ in the DPC problem \eqref{E:DPCrealcase}, i.e.
 the disturbance in input in the blue rectangle in Figure \ref{F:overview}, while we use the correct ones to simulate the process, i.e.
@@
 We obviously do not add any noise on the "Time of the day" and on the "Day of the week" since those are perfectly known.
-In the table we also report the range of the values that the variables assumed during the reference period of the historical dataset.
-This is to compare with the mean and deviation values we chose to show that we considered quite bad error on the forecast.
-In the following results we show the change in performance on bounds violation and energy consumption due to the forecast inaccuracy.}
+We considered zero mean since we suppose there is no constant offset in the forecast, but only an error that can over or underestimate the predictions.
+In the table, we also report the range of the values that the variables assumed during the reference period of the historical dataset.
+This is to show that the deviation values we chose add a quite bad error on the forecast.
+In particular, for example, a deviation of $0.5\degree C$ on the outside air temperature means that the error on the predicted temperature lies within the range of $\pm 1.5\degree C$ with a probability of $99\%$.
+We solved DPC problem \ref{P:dpcRealCase} considering the new parameters $\hat \Theta$, obtained using the aforementioned noisy forecast, obtaining the following results, where we show the change in performance on bounds violation and energy consumption due to the forecast inaccuracy.}
 \begin{table}[t!]
@@
 		\midrule
-		Outside temperature    & [0,31]     & 1	   & 0.5       \\
-		Wind                   & [0,5] 		& 0.2  & 0.25      \\
-		Atmospheric pressure   & [990,1030] & 50   & 50        \\
-		Relative Humidity      & [20,91]	& 50   & 5         \\
-		Solar Radiation        & [0,1000]   & 50   & 50        \\
+		Outside temperature    & [0,31]     & 0	   & 0.5       \\
+		Wind                   & [0,5] 		& 0    & 0.25      \\
+		Atmospheric pressure   & [990,1030] & 0    & 50        \\
+		Relative Humidity      & [20,91]	& 0    & 5         \\
+		Solar Radiation        & [0,1000]   & 0    & 50        \\
 		Time of the day        & [0,23]     & 0    & 0         \\
@@
 
-\textcolor[rgb]{0,0,1}{\paragraph{Result 1} In Figure \ref{F:ComparisonTempNoisy} a comparison of the DPC results in terms of temperature control, between the perfect and the noisy forecasts, is shown.
-For the sake of plot's clarity, we considered 4 days and a half as simulative period and we split the plot into 3 sub-figures referring to the small, medium and large violation cases.
+\textcolor[rgb]{0,0,1}{\paragraph{Result 1} In Figure \ref{F:ComparisonTempNoisy} a comparison of the DPC results in terms of temperature control for thermal comfort, between the perfect and the noisy forecast, is shown.
+We ran the simulations over a period of 15 days, howeveror the sake of plot's clarity, we show only 4 days and a half as simulative period and we split the plot into 3 sub-figures referring to the small, medium and large violation cases.
 We can see that in the case of small and large violation we have a small deterioration in performance in terms of bounds violation, although this is more highlighted in the large violation case.
@@
 		\midrule
-		$DPC-SV$    & $0.072$             & $0.074$  			      & $0$    				 & $0$     	  	\\
-		$DPC-MV$    & $0.168$ 			  & $0.13$       			  & $0$    				 & $0$		  	\\
-		$DPC-LV$    & $0.708$  			  & $0.343$     			  & $0$    				 & $0$	      	\\
+		$DPC-SV$    & $0.034$             & $0.142$  			      & $0$    				 & $0$     	  	\\
+		$DPC-MV$    & $0.178$ 			  & $0.453$       			  & $0$    				 & $0$		  	\\
+		$DPC-LV$    & $0.478$  			  & $0.891$     			  & $0$    				 & $0$	      	\\
 		\bottomrule

### 2018-01-19 paper/final/casestudyReal.tex
Updated Sec2,3,5, and intro review by Ale

@@
 Finally, in Section \ref{SS:simulationResults}, the performance of the two controllers are compared in terms of energy saving using the EnergyPlus model.
-In particular, we show that DPC allows an energy saving up to $49.2\%$ with respect to the bang-bang controller while guaranteeing thermal comfort for occupants.
+\textcolor[rgb]{0,0,1}{In particular, in Section \ref{SSS:DisturbancePerfect}, we show that DPC allows an energy saving up to $49.2\%$ with respect to the bang-bang controller while guaranteeing thermal comfort for occupants, considering the perfect knowledge of the weather forecast.
+In Section \ref{SSS:DisturbanceUncertain}, we show that DPC is robust with respect to imperfect weather forecast.}
 The section organization is graphically shown in Figure \ref{F:overview}.
@@
 
+\textcolor[rgb]{0,0,1}{The contribution of this section with respect to Sections \ref{S:casestudy} is threefold:
+\begin{enumerate}
+	\item we consider experimental data from a lived house, hence subject to real world imperfections, such as random (non predictable) occupancy schedules, open/close windows, random (non predictable) light on/off switch, etc..
+	On the other hand, the EnergyPlus model considered in Section \ref{S:casestudy}, although provides an accurate model of a building, is still a simulated model that does not take into account the aforementioned imperfections;
+	\item we modify the DPC algorithm to take into account uncertainties on the weather forecast, so making DPC robust with respect to prediction errors;
+	\item in Section \ref{S:casestudy}, we want to cut the peak of the power in a specific time range while keeping the temperature as close as possible to a specific setpoint.
+	Differently, in this section, we show the adaptability of DPC to different problems, minimizing the energy consumption required to keep room temperatures within a specified range of comfort for an arbitrary amount of time.
+In this way we show the potential of DPC for energy saving on a long period.
+\end{enumerate}}
 \subsection{Description of the house}\label{SS:descriptionHouse}
@@
 \label{SS:randomForestsModels}
-For the closed-loop simulations with DPC we create $2$ different sets of models, $S_1 = \{\tT^{r_1}_{\mathrm{k+1|k}},\tT^{r_2}_{\mathrm{k+1|k}},\tT^{r_3}_{\mathrm{k+1|k}},\tT^{r_4}_{\mathrm{k+j|k}}\}$ and $S_2 = \{\tP_{\mathrm{k+j|k}},\tT^{1}_{\mathrm{k+j|k}},\tT^{2}_{\mathrm{k+j|k}},\tT^{3}_{\mathrm{k+j|k}},\\\tT^{4}_{\mathrm{k+j|k}},\ j = 1,\ldots,N\}$, using random forests.
+For the closed-loop simulations with DPC we create $2$ different sets of models, $S_1 = \{\tT^{r_1}_{k+1},\tT^{r_2}_{k+1},\tT^{r_3}_{k+1},\tT^{r_4}_{k+j}\}$ and $S_2 = \{\tP_{k+j},\tT^{1}_{k+j},\tT^{2}_{k+j},\tT^{3}_{k+j},\\\tT^{4}_{k+j},\ j = 1,\ldots,N\}$, using random forests.
 In each set we have  $4$ models that describe the room temperature evolution ($\tT^{r_i}$ and $\tT^i$, $i=1,2,3,4$) in each of the $4$ rooms equipped with temperature sensors.
@@
 We want to optimize the ON/OFF heating system schedule in order to minimize power consumption of the house while keeping temperature of room $3$ within a comfort range.
-We also allow violations $\eps^{min},\eps^{max}$  of temperature bounds to guarantee feasibility of the algorithm.
+We also allow violations $\epsilon^{min},\epsilon^{max}$  of temperature bounds to guarantee feasibility of the algorithm.
 We include these violations in the objective function to be minimized.
@@
 The problem is set up as follows:
+
 \begin{problem}\label{P:dpcRealCase}
@@
 		\begin{aligned}
-		\min_{\tX^c,\eps_j^{min},\eps_j^{max}} & \ \ \ \sum_{j=1}^{N} Q{\tP}_{\mathrm{k+j|k}}^2 +  \lambda_{min}\parallel\eps^{min}\parallel_2 + \lambda_{max}\parallel\eps^{max}\parallel_2 \\
-		\text{s.~t. }& \ \ \ \ \ {\tP}_{\mathrm{k+j|k}} =  \hat{\Theta}^{T}_{\tP_j} [ 1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
-					 & \ \ \ \ \ \tT^1_{\mathrm{k+j|k}} =  \hat{\Theta}^{T}_{\tT^1_j} [ 1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
-					 & \ \ \ \ \ \tT^2_{\mathrm{k+j|k}} =  \hat{\Theta}^{T}_{\tT^2_j} [ 1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
-			 	 	 & \ \ \ \ \ \tT^3_{\mathrm{k+j|k}} =  \hat{\Theta}^{T}_{\tT^3_j} [ 1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
-				 	 & \ \ \ \ \ \tT^4_{\mathrm{k+j|k}} =  \hat{\Theta}^{T}_{\tT^4_j} [ 1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\					
-					 & \underline{\tT}_{\mathrm{k+j-1|k}}-\eps^{min}_j \leq \tT^3_{\mathrm{k+j|k}} \leq \overline{\tT}_{\mathrm{k+j-1|k}} + \eps^{max}_j\\
-					 & \ \ \ \ \ \ \ \ \tX^c_{\mathrm{k+j-1|k}} = \underline{\tX}^c \lor \tX^c_{\mathrm{k+j-1|k}} = \overline{\tX}^c\\ 
-					 & \ \ \ \ \ \ \ \ \,\eps^{min}_j,\eps^{max}_j \geq 0, \ j = 1,\dots,N.
+		& \underset{u_{k+j},\epsilon_j^{min},\epsilon_j^{max}}{\text{minimize}} & & \sum_{j=1}^{N} Q{\tP}_{k+j}^2 + \lambda_{min}\parallel\epsilon^{min}\parallel_2 + \lambda_{max}\parallel\epsilon^{max}\parallel_2                                                                                                                \\
+		& \text{subject to }                             & & \tP_{k+j}   = \hat{\Theta}_{\tP_j}   [ 1,u_k,\ldots,u_{k+j-1} ]^\top                                      \\
+		&                                                & & \tT^1_{k+j} = \hat{\Theta}_{\tT^1_j} [ 1,u_k,\ldots,u_{k+j-1} ]^\top                                      \\
+		&                                                & & \tT^2_{k+j} = \hat{\Theta}_{\tT^2_j} [ 1,u_k,\ldots,u_{k+j-1} ]^\top                                      \\
+		&                                                & & \tT^3_{k+j} = \hat{\Theta}_{\tT^3_j} [ 1,u_k,\ldots,u_{k+j-1} ]^\top                                      \\
+		&                                                & & \tT^4_{k+j} = \hat{\Theta}_{\tT^4_j} [ 1,u_k,\ldots,u_{k+j-1} ]^\top                                      \\
+		&                                                & & \underline{\tT}_{k+j-1}-\epsilon^{min}_j \leq \tT^3_{k+j} \leq \overline{\tT}_{k+j-1} + \epsilon^{max}_j  \\
+		&                                                & & u_{k+j-1}   = \underline{u} \lor u_{k+j-1} = \overline{u}                                                 \\
+		&                                                & & \epsilon^{min}_j,\epsilon^{max}_j \geq 0, \ j = 1,\dots,N.
 		\end{aligned}
-	\label{E:DPCrealcase}
-\end{align}
+	\end{align}\label{E:DPCrealcase}
 \end{problem}
+
 The choice of different weights $Q$, $\lambda_{min}$ and $\lambda_{max}$ allows the designer to give more importance to energy consumption rather than temperature comfort and vice versa.
 In Section \ref{SS:simulationResults} we will show different performance results considering different weights.
-Parameters $\underline{\tX}^c$ and $\overline{\tX}^c$ are respectively minimum and maximum values the heating system can actuate, while $\underline{\tT}_{\mathrm{k+j-1|k}}$ and $\overline{\tT}_{\mathrm{k+j-1|k}}$ are respectively time varying lower and upper bounds to keep the temperature in a desired range of comfort.
-Due to the integer variable constraint for input $\tX^c$, the problem is a Mixed Integer Quadratic Programming.
+Parameters $\underline{u}$ and $\overline{u}$ are respectively minimum and maximum values the heating system can actuate, while $\underline{\tT}_{k+j-1}$ and $\overline{\tT}_{k+j-1}$ are respectively time varying lower and upper bounds to keep the temperature in a desired range of comfort.
+Due to the integer variable constraint for input $u$, the problem is a Mixed Integer Quadratic Programming.
 For the implementation we use in Section \ref{SS:simulationResults} Gurobi solver \cite{Gurobi2015} through CVX \cite{cvx,gb08}.
@@
 
-\subsection{Simulation Results}\label{SS:simulationResults} We simulated  DPC in \eqref{E:DPCrealcase} and the bang-bang controller, in closed-loop with the house models $\tT^{r_i}_{k+1|k},\ i=1,\ldots,4$ in $S_1$.
+\subsection{Simulation Results}\label{SS:simulationResults} We simulated  DPC in \eqref{E:DPCrealcase} and the bang-bang controller, in closed-loop with the house models $\tT^{r_i}_{k+1},\ i=1,\ldots,4$ in $S_1$.
 We considered a sampling time of $10$ minutes and chose $N=4$ as a predictive horizon, i.e.
 $40$ minutes.
-From historical data we got that $\overline{\tX}^c = 0.35 \, m^3/h$ when the heating system is ON and obviously $\underline{\tX}^c = 0 \, m^3/h$ when the heating system is OFF.
-For the temperature comfort range we set a constant upper bound $\overline{\tT}_{\mathrm{k}} = 22.5 \, \degree C$ and a variable lower bound that is $\underline{\tT}_{\mathrm{k}} = 21 \, \degree C$ from $7am$ to $9am$ when people in the house wake up and go out for work, and from $6pm$ to midnight when people come back from work and go to sleep.
-During other hours when people are either not at home or asleep, we set $\underline{\tT}_{\mathrm{k}} = 20 \, \degree C$
+From historical data we got that $\overline{u} = 0.35 \, m^3/h$ when the heating system is ON and obviously $\underline{u} = 0 \, m^3/h$ when the heating system is OFF.
+For the temperature comfort range we set a constant upper bound $\overline{\tT}_{k} = 22.5 \, \degree C$ and a variable lower bound that is $\underline{\tT}_{k} = 21 \, \degree C$ from $7am$ to $9am$ when people in the house wake up and go out for work, and from $6pm$ to midnight when people come back from work and go to sleep.
+During other hours when people are either not at home or asleep, we set $\underline{\tT}_{k} = 20 \, \degree C$
 
@@
 In particular, we considered a small ($Q=100$, $\lambda_{min}=3000$ and $\lambda_{max}=100$), a medium ($Q=100$, $\lambda_{min}=1000$ and $\lambda_{max}=100$) and a large ($Q=100$, $\lambda_{min}=100$ and $\lambda_{max}=100$) violation configuration.
+\textcolor[rgb]{0,0,1}{The simulation period is of $15$ days, from May $1$, $2016$ at $00am$ to May $15$, $2016$ $00am$.}
 
@@
 	\item in Section \ref{SSS:DisturbancePerfect} we provide our simulative results considering perfect knowledge of the weather forecast;
-	\item in Section \ref{SSS:DisturbanceUncertain}, to show the robustness of our methodology, we consider weather forecast subject to uncertainty.
-We add a gaussian noise with high variance on the perfect forecast and show that the results are close to the perfect forecast case.
+	\item in Section \ref{SSS:DisturbanceUncertain}, we consider weather forecast subject to uncertainty Hence we modify Problem \ref{P:dpcRealCase} by adding a gaussian noise with high variance on the perfect forecast, and show that the results are close to the perfect forecast case, so making our methodology robust with respect to disturbance uncertainties.
 \end{enumerate}
@@
 \paragraph{Result 1} The comparison for temperature and input schedule obtained allowing small bound violations in DPC is shown in Figure \ref{F:comparison_small}.
-For sake of clarity, the simulation period is restricted to $4$ days and a half, from May $1$, $2016$ $00am$ to May $5$, $2016$ $1pm$.
+For sake of the plot's clarity, the shown period is restricted to $4$ days and a half, from May $1$, $2016$ $00am$ to May $5$, $2016$ $1pm$.
+\textcolor[rgb]{0,0,1}{The whole period will be used in "Result 2" to provide the bounds violation errors, and in "Result 3" for the energy consumption comparison.}
 We can see that the temperature controlled with DPC does not violate the bounds and if it does then the violation is approximately $0.1\,\degree C$.
@@
 However the maximum violation is still lower than $1\,\degree C$.
-In Table \ref{T:violationErrors}, MBE and CV(RMSE) violation errors are reported to quantify the bounds violation of DPC, in each of the $3$ configurations, and of bang-bang controller.
+In Table \ref{T:violationErrors}, MBE and CV(RMSE)\textcolor[rgb]{0,0,1}{ (expressed in $\%$, and computed over the whole simulative period, i.e.
+15 days) }violation errors are reported to quantify the bounds violation of DPC, in each of the $3$ configurations, and of bang-bang controller.
 We can see that if we allow very small violations, DPC outperforms bang-bang controller in terms of comfort guarantees.
@@
 In \cite{Petersen2014AE} the authors investigated, through a large-scale simulation study, the effect of weather forecast uncertainty on the performance of MPC in building systems operations, and compared to the performance of a rule-based strategy.
-They considered 48 different scenarios of uncertainties on 72 hours weather forecasts.
-On such long horizon, results have shown that, with few exceptions, that MPC outperforms the rule-based controller in case of energy savings and/or thermal indoor environment, and is in general quite close to the perfect forecast case, despite the uncertainty in weather forecast.}
+They considered 48 different scenarios of uncertainties on 72 hours weather forecast.
+On such long horizon, results have shown that, with few exceptions, MPC outperforms the rule-based controller in case of energy savings and/or thermal indoor environment, and is in general quite close to the perfect forecast case, despite the uncertainty in weather forecast.}
 
@@
 \textcolor[rgb]{0,0,1}{\paragraph{Result 1} In Figure \ref{F:ComparisonTempNoisy} a comparison of the DPC results in terms of temperature control for thermal comfort, between the perfect and the noisy forecast, is shown.
-We ran the simulations over a period of 15 days, howeveror the sake of plot's clarity, we show only 4 days and a half as simulative period and we split the plot into 3 sub-figures referring to the small, medium and large violation cases.
-We can see that in the case of small and large violation we have a small deterioration in performance in terms of bounds violation, although this is more highlighted in the large violation case.
-The medium violation cases are instead comparable.
-The maximum violation, as in the case of perfect forecast, is still lower than $1\,\degree C$.
-In particular, we report in Table \ref{T:violationErrorsNoisy} the violation errors in terms of MBE and CV(RMSE).
-If we compare them with the ones in Table \ref{T:violationErrors}, we can see that the performance of DPC with noisy forecast are still good if compared with the performance of DPC with perfect forecast, despite the extremely bad prediction error that we considered on the weather forecast.
+Also here, for the sake of plot's clarity, we show only 4 days and a half of the 15 days simulative period, and we split the plot into 3 sub-figures referring to the small, medium and large violation cases.
+We can see that, except for an isolated case in the medium violation case (slightly before the "02/05"), we have a small performance deterioration in terms of bounds violation, although this is more highlighted in the large violation case.
+However, the maximum violation, as in the case of perfect forecast, is still lower than $1\,\degree C$.
+In particular, we report in Table \ref{T:violationErrorsNoisy} the violation errors in terms of MBE and CV(RMSE) (in $\%$) computed for the whole simulative period, i.e. 15 days.
+If we compare them with the ones in Table \ref{T:violationErrors}, we can see that the performance of DPC with noisy forecast are still good with respect to the performance of DPC with perfect forecast, despite the extremely bad prediction error that we considered on the weather forecast.
 }
@@
 \textcolor[rgb]{0,0,1}{\paragraph{Result 2} 
-In Figure \ref{F:comparison_all_energy_E+_noisy}, we show the difference in terms of energy consumption considering the perfect weather forecast and the noisy one.
-We can observe the case with imperfect forecast is very close to the perfect forecast case, and depending the conditions for which the disturbance is over or underestimated, we have more or less energy consumption.
-These results shows the strong robustness of the DPC with respect to errors in the weather forecast.}
+In Figure \ref{F:comparison_all_energy_E+_noisy}, we show the difference in terms of energy consumption considering the perfect weather forecast (full line) and the noisy one (dashed line).
+We can observe that the case with imperfect forecast is very close to the perfect forecast case.
+We recall that we are optimizing a weighted sum of energy and thermal comfort, so depending on the conditions for which the disturbance is over or underestimated, the MPC can require to use more or less energy to supply the prediction errors.
+In the small violation case, the energy consumption of the 2 conditions is closer than in the others.
+This is because the DPC keeps the temperature within the bounds with an extremely small error in bounds violation.
+In particular, the only exception can be seen before May $3^{rd}$ (03/05), where, due to the prediction error, the temperature violates the bounds, so the control uses more energy (than the perfect case) to bring it back above the threshold.
+In the other 2 cases, when the temperature violates the bounds, due to the fact that violation constraints are more relaxed, the DPC does not require a too big control effort, hence using less energy.}
+
+\textcolor[rgb]{0,0,1}{These results show the strong robustness of the DPC with respect to errors in the weather forecast.}
 \begin{figure}[t!]

### 2018-01-19 paper/final/cs-dpc.tex
Updated Sec2,3,5, and intro review by Ale

--- 2017-09-29 paper/cs-dpc.tex
@@
-In this section, we explain how DPC can be applied to this case study. We begin with a description of features $\tX$ and output $\tY$ used for training.
+In this section, we explain how DPC can be applied to this case study. We begin with a description of features $\X$ and output $\Y$ used for training.
 
@@
 the zone temperature that can be easily measured with a thermostat.
-This serves as the output variable $\tY$ of interest for which we build $N$ trees and $N$ forests as described in Section~\ref{SS:dpcrt} and \ref{SS:dpcrf}, respectively.
-Therefore, $\tY_{\mathrm{k+j|k}}:=x_{\mathrm{k+j|k}}^1$, where $x^1$ is the first component of $x$.
-Next, we define the non-manipulated features $\tX^d_{\mathrm{k|k}}$.
-At time $k$, for the tree $\mathcal{T}_j$ and the forest $\mathcal{R}_j$, we base these features to include weather disturbances, external disturbances due to occupancy and equipments, and autoregressive terms of the room temperature, i.e.
-$\tX^d_{\mathrm{k|k}}:=[ d_{\mathrm{k+j-N|k}} ,\dots,d_{\mathrm{k+j-1|k}}, x_{\mathrm{k|k}}^1,\dots, x_{\mathrm{k-\delta|k}}^1 ]$, where $\delta$ is the order of autoregression.
-Finally, the inputs in DPC are exactly same as in MPC. i.e. $\tX^c_{\mathrm{k+j-1|k}}:=u_{\mathrm{k+j-1|k}}$.
-
+This serves as the output variable in $\Y$ of interest for which we build $N$ trees and $N$ forests as described in Section~\ref{SS:dpcrt} and \ref{SS:dpcrf}, respectively.
+Therefore, $y_{k+j}:=x_{k+j+1}^1$, where $x^1$ is the first component of $x$.
+Next, we define the non-manipulated features in $\X^d$.
+At time $k$, for the tree $\mathcal{T}_j$ and the forest $\mathcal{F}_j$, we base these features to include weather disturbances, external disturbances due to occupancy and equipments, and autoregressive terms of the room temperature, i.e.
+$\X^d:=\{( d_{k+j-N} ,\ldots,d_{k+j-1}, x_{k}^1,\ldots, x_{k-\delta}^1 )\},\ \forall k=1,\ldots,n$, where $\delta$ is the order of state autoregression.
+Finally, the inputs in DPC are exactly same as in MPC, i.e. $\X^c:=u_k,\ \forall k=1,\ldots,n$.
 The training data in the above format was generated by simulating the bilinear model with rule-based strategies for 10 months in 2007.
 January and May were deliberately excluded for testing the DPC implementation.
-
 \subsubsection{Optimization} 
@@
 For a fair comparison with MPC, we cast DPC optimization problem as follows:
+
 \begin{subequations}
 	\begin{align}
-\text{min } & \sum_{j=1}^{N} ({\tY}_{\mathrm{k+j|k}} - x_{ref}) \mathcal{Q}^{(1,1)} ({\tY}_{\mathrm{k+j|k}} - x_{ref}) + c^T \tX^c_{\mathrm{k+j-1|k}}+  \lambda\epsilon_j\\
-\text{s.~t. } & \ \ \ \ \ \tY_{\mathrm{k+j|k}} =  \alpha_j^T \left[1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} \right]^T \label{SE:dpc}\\
-& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tX}^c \leq \tX^c_{\mathrm{k+j-1|k}} \leq \bar{\tX}^c\\ 
-& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tY}-\epsilon_j \leq \tY_{\mathrm{k+j|k}} \leq \bar{\tY} + \epsilon_j\\\
-& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0, \ j = 1,\dots,N.
-\end{align} \label{E:dpc}
+	& \underset{u_{k+j},\epsilon_j}{\text{minimize}} & & \sum_{j=1}^{N} (y_{k+j} - x_{ref}) \mathcal{Q}^{(1,1)} (y_{k+j} - x_{ref}) + c^T u_{k+j-1} + \lambda\epsilon_j \\
+	& \text{subject to }                  & & y_{k+j} =  \alpha_j \left[1,u_{k},\ldots,u_{k+j-1} \right]^\top \label{SE:dpc}                                            \\
+	&                                     & & \underline{u} \leq u_{k+j-1} \leq \bar{u} 									   	                                        \\
+	&                                     & & \underline{y} - \epsilon_j \leq y_{k+j} \leq \bar{y} + \epsilon_j                                                         \\
+	&                                     & & \epsilon_j \geq 0, \ j = 1,\dots,N,            									                 
+	\end{align}\label{E:dpc}
 \end{subequations}
-\noindent Here $\alpha = \beta$ for DPC-RT and $\alpha = \hat{\Theta}$ for DPC-En.
+
+\noindent where $\alpha = \beta$ for DPC-RT and $\alpha = \hat{\Theta}$ for DPC-En.
 Note that, \eqref{E:dpc} is DPC analog of \eqref{E:mpc}. The only difference is the state dynamics \eqref{SE:mpc1} and \eqref{SE:mpc2} are now replaced with \eqref{SE:dpc}.
@@
 
-We compare the prediction for the first time step $\tY_{\mathrm{k+1|k}}$ and the 6-hour ahead prediction $\tY_{\mathrm{k+6|k}}$ for a week in the month of May in Figure~\ref{F:validation}.
+We compare the prediction for the first time step $y_{k+1}$ and the 6-hour ahead prediction $y_{k+6}$, given $k$, for a week in the month of May in Figure~\ref{F:validation}.
 It is visible how trees have a high variance, and the forests are more accurate.
@@
 		\midrule
-		tree-$\tY_{\mathrm{k+1|k}}$    &  0.42 &  0.75 & 0.76    \\
-		tree-$\tY_{\mathrm{k+6|k}}$  & 0.64 &  0.41  & 0.42 \\
-		forest-$\tY_{\mathrm{k+1|k}}$  & 0.29 & 0.87  & 0.88 \\
-		forest-$\tY_{\mathrm{k+6|k}}$  & 0.38 & 0.78 & 0.80 \\
+		tree-$y_{k+1}$    & 0.42 & 0.75        & 0.76 \\
+		tree-$y_{k+6}$    & 0.64 & 0.41        & 0.42 \\
+		forest-$y_{k+1}$  & 0.29 & 0.87        & 0.88 \\
+		forest-$y_{k+6}$  & 0.38 & 0.78        & 0.80 \\
 		\bottomrule
@@
 	\centering
-	\includegraphics[width=20pc]{figures/validation-s1.eps}
-	\includegraphics[width=20pc]{figures/validation-s6.eps}
+	\includegraphics[width=24pc]{figures/validation-s1.eps}
+	\includegraphics[width=24pc]{figures/validation-s6.eps}
 	\caption{Temperature predictions from a tree and a forest for first step prediction (top) and the 6-hour ahead prediction (bottom). Ensemble method shows a relatively higher accuracy.}

### 2018-01-19 paper/final/cs-model.tex
Updated Sec2,3,5, and intro review by Ale

--- 2017-09-29 paper/cs-model.tex
@@
 The model dynamics are given below. The bilinearity is present in both input-state, and input-disturbance.
-\begin{gather}
-\label{E:bilinear1}
-x_{k+1} = Ax_{k}+(B_u +B_{xu}[x_k] + B_{du}[d_k]) u_k+B_dd_k \\
-x_{k} \in \mathbb{R}^{12}, u_{k} \in \mathbb{R}^{4}, d_{k} \in \mathbb{R}^{8} \ \forall k = 0,\dots,T, \nonumber
-\end{gather}
-where, the matrices $B_{xu}$ and $B_{du}$ are defined as
-\begin{gather}
-\label{E:bilinear2}
-B_{xu}[x_k] = [ B_{xu,1}[x_k],B_{xu,2}[x_k], \dots, B_{xu,4}[x_k] ] \in \mathbb{R}^{12\times4}, \nonumber\\
-B_{du}[d_k] = [ B_{du,1}[d_k],B_{du,2}[d_k], \dots, B_{du,4}[d_k] ] \in \mathbb{R}^{12\times4}\nonumber, \\
-B_{xu,i} \in \R^{12\times12}, B_{du,i} \in \R^{12\times8} \ \forall i=1,2,3,4.\nonumber
-\end{gather}
+\begin{equation}\label{E:bilinear1}
+x_{k+1} = Ax_{k}+(B_u +B_{xu}[x_k] + B_{du}[d_k]) u_k+B_dd_k,
+\end{equation}
+where $x_{k} \in \mathbb{R}^{12}, u_{k} \in \mathbb{R}^{4}, d_{k} \in \mathbb{R}^{8} \ \forall k = 0,\dots,T$, and the matrices $B_{xu}$ and $B_{du}$ are defined as
+\begin{equation}
+	\begin{aligned}
+		&B_{xu}[x_k] = [ B_{xu,1}[x_k],B_{xu,2}[x_k], \dots, B_{xu,4}[x_k] ] \in \mathbb{R}^{12\times4}, \\
+		&B_{du}[d_k] = [ B_{du,1}[d_k],B_{du,2}[d_k], \dots, B_{du,4}[d_k] ] \in \mathbb{R}^{12\times4}, \\
+	\end{aligned}
+\end{equation}
+with $B_{xu,i} \in \R^{12\times12}, B_{du,i} \in \R^{12\times8} \ \forall i=1,2,3,4$.
 For this study, we assume that the disturbances are precisely known to MPC as well as DPC controller.
-In our future work, we will account for the uncertainties in the disturbances with an extension to Scenario approach \cite{Bernardini2009} for DPC.

### 2018-01-19 paper/final/cs-mpc.tex
Updated Sec2,3,5, and intro review by Ale

@@
 
-The objective of the controller is to minimize the energy usage $c^Tu$ while maintaining a desired level of thermal comfort $x_{ref}$.
+The objective of the controller is to minimize the energy usage $c^Tu$ while maintaining a desired level of thermal comfort $x_{ref}$ (or $\tT_\mathrm{ref}$).
 Therefore, at time step $k$, we solve a continuously linearized MPC problem to determine the optimal sequence of inputs $[u_{\mathrm{k|k}},\dots,u_{\mathrm{k+N-1|k}}]$:
+
 \begin{subequations}
 	\begin{align}
-\text{min } & \sum_{j=1}^{N} ({x}^T_{\mathrm{k+j|k}} - x_{ref}) \mathcal{Q} ({x}_{\mathrm{k+j|k}} - x_{ref}) + c^Tu_{\mathrm{k+j-1}} +  \lambda\epsilon_j\\
-\text{s.~t. } & \ \ x_{\mathrm{k+j|k}} =  Ax_{\mathrm{k|k}} + B u_{\mathrm{k+j-1|k}} + B_d d_{\mathrm{k+j-1|k}} \label{SE:mpc1} \\
-& \ \ \ \ \ B = B_u + B_{xu}[x_{\mathrm{k|k}}] + B_{du}[d_{\mathrm{k+j-1|k}}] \label{SE:mpc2}\\
-& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{u} \leq u_{\mathrm{k+j-1|k}} \leq \bar{u}\\ 
-& \ \ \ \ \ \ \ \ \ \ \ \ \underline{x}-\epsilon_j \leq x_{\mathrm{k+j|k}} \leq \bar{x} + \epsilon_j\\\
-& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0, \ j = 1,\dots,N,
+		& \underset{u_{k+j},\epsilon_j}{\text{minimize}} & & \sum_{j=1}^{N} ({x}_{k+j} - x_{ref})\top Q ({x}_{k+j} - x_{ref}) + c^Tu_{k+j-1} +  \lambda\epsilon_j \\
+		& \text{subject to }                             & & x_{k+j} = Ax_{k+j-1} + B u_{k+j-1} + B_d d_{k+j-1} \label{SE:mpc1}                                   \\
+		&                                                & & B       = B_u + B_{xu}[x_{k}] + B_{du}[d_{k+j-1}] \label{SE:mpc2}                                    \\
+		&                                                & & \underline{u} \leq u_{k+j-1} \leq \bar{u} 									   	                      \\
+		&                                                & & \underline{x} - \epsilon_j \leq x_{k+j} \leq \bar{x} + \epsilon_j                                    \\
+		&                                                & & \epsilon_j \geq 0, \ j = 1,\dots,N,            									                 
 	\end{align}\label{E:mpc}
 \end{subequations}
-\noindent where $\mathcal{Q} \in \R^{12 \times 12}$ has all zeros except at $\mathcal{Q}^{(1,1)}$ corresponding to the zone temperature, $c \in \R^{4}$ is proportional to cost of using each actuator and $\lambda$ penalizes the slack variables.
+
+\noindent where $Q \in \R^{12 \times 12}$ has all zeros except at $Q^{(1,1)}$ corresponding to the zone temperature, $c \in \R^{4}$ is proportional to cost of using each actuator and $\lambda$ penalizes the slack variables.

### 2018-01-19 paper/final/intro.tex
Updated Sec2,3,5, and intro review by Ale

@@
 \section{INTRODUCTION}
+
 \label{S:intro}
@@
 With a reasonably accurate forecast of future weather and building operating conditions, dynamical models can be used to predict the energy needs of the building over a prediction horizon, as is the case with Model Predictive Control (MPC) \cite{Sturzenegger2016}.
-However, a major challenge with MPC is in accurately modeling the dynamics of the underlying physical system.
+However, a major challenge with MPC is in (accurately) modeling the dynamics of the underlying physical system.
 The task is much more complicated and time consuming in the case of a large building and often times, it can be even more complex and involved than the controller design itself.
@@
 The user expertise, time, and associated sensor costs required to develop a model of a single building is very high.
-Thus, the payback period for the upfront hardware and software installation is expected to be too high, making MPC an uneconomical choice for energy management.}
+Thus, the payback period for the upfront hardware and software installation is expected to be too high, making MPC an uneconomical choice for energy management.
+Indeed, there are several reasons why physics-based modeling is hard for complex systems like buildings:}
 
-\textcolor[rgb]{0,0,1}{Why is physics-based modeling hard for complex systems like buildings or what are the reasons of such modeling complexities?}
 \begin{enumerate}
@@
 
-\textcolor[rgb]{0,0,1}{Due to aforementioned reasons, the control strategies in such systems are often limited to fixed, sometimes ad-hoc, rules that are based on best practices. 
+\textcolor[rgb]{0,0,1}{
+In Section \ref{secModelingIssuesExample} we will present a detailed technical example to better illustrate how data-driven approaches can address the above issues and thus reduce the cost of modeling buildings.
+Indeed in practice, due to aforementioned reasons, the control strategies in such systems are often limited to fixed, sometimes ad-hoc, rules that are based on best practices.
 The alternative is to use black-box, or completely data-driven modeling approaches, to obtain a realization of the system's input-output behavior. 
@@
 We solve this problem by bridging Machine Learning and Predictive Control.
-In particular, in this paper, we present a method based on Random Forests which uses historical data for receding horizon control.}
+In particular, in this paper, we present a method based on Random Forests which uses historical data for receding horizon control.
+Before illustrating our approach, we will discuss in the next section the related literature and the novelty of our approach.}
 
-\textcolor[rgb]{0,0,1}{
-Next, we present an example to illustrate how data-driven modeling reduces the cost of modeling buildings. 
-}
+%====================================================================================================================================================================================
 
-%\paragraph{Example} \textcolor[rgb]{0.00,0.00,1.00}{As a first} example of such modeling complexity, consider the grey-box approach in \cite{Braun2002}.
-The scope of such approach is to predict the heat transfer rate to the air within the building.
-In particular, the authors addressed a single zone modeling problem using an equivalent $RC$ model.
-To this end, five different types of structures are considered in the example: external walls, ceiling/roof, floor, internal walls and windows.
-Each of these elements are represented using $3$ resistances, $2$ capacitances and $2$ temperatures, except for the windows that are represented only with resistances since they have a negligible energy storage.
-A model consisting of $8$ states and $9$ inputs, including disturbances such as solar radiation and others, with the instantaneous heat gain to the building air from all surfaces as output, is created for the zone.
-To have a complete LTI model, $9$ parameters are needed to be estimated via a non-trivial training process, using a collection of information associated with a physical description of the building (see \cite{Braun2002} for more details).
-Extending this result, for the building we consider in Section \ref{S:realCaseStudy} with $10$ different zones, we would need to consider a model with approximately $80$ states and $90$ parameters to be estimated.
-
-\subsection{Example for comparison of modeling complexity: physics-based vs data-driven}
-\textcolor[rgb]{0,0,1}{In this section, we provide a detailed example that describes the difference in terms of complexity when we model a building using physical laws and machine learning.
-We show that the data-driven modeling eliminates several drawbacks that occur using physics-based models such as the need to have very good knowledge of the building structure and the material properties, time required to build a model, and limited availability of sensor measurements.}
-
-\textcolor[rgb]{0,0,1}{The building under consideration is taken from \cite{Sturzenegger2016}.
-It is located in Allschwil, Switzerland.
-It has 6 floors of which only the second floor is modeled, and is assumed to represent all other floors.
-The total air conditioned floor area is $6000\ \mathrm{m^2}$ ca.}
-
-\subsubsection{Physics-based modeling}
-\label{SSS:physics-based}
-\textcolor[rgb]{0,0,1}{We describe the physics-based approach based on an RC network outlined in \cite{Sturzenegger2016}.
-The building is modeled as a bilinear system constructed from physical principles.
-The modeling process consists of three steps:
-\begin{enumerate}
-	\item Building geometry and construction data are used together with first-principles to derive the following linear model for the building's thermal dynamics:
-		\begin{equation}\label{E:RCZoneEq}
-			\dot x(t) = A_x x(t) + B_q q(t).
-		\end{equation}
-		This model describes the behavior of the zone, wall, floor and ceiling temperatures.
-		Walls, floors and ceilings are considered as divided into layers with different features.
-Therefore, each zone was described with an RC network model (see Figure 3-10 in \cite{SturzeneggerTR}), where the capacitances represent the states of the layers and the resistances represent the thermal resistance of the layers.
-%		\begin{figure}[t]
-%			\begin{center}
-%				\includegraphics[width=0.6\textwidth]{figures/RC_net.eps}
-%				\caption{RC model of a zone. $q$ represents the external heat fluxes}\label{F:RCnet}
-%			\end{center}
-%		\end{figure}
-		The heat exchange between two adjacent layers, i.e.
-layer ``a" and layer ``b", is modeled to be proportional to the temperature difference of the two layers and the corresponding thermal resistance $R$ given by
-		\begin{equation}\label{E:RCHeatExchange}
-			\begin{aligned}
-				&C_a\dot x_a = \frac{x_b(t) - x_a(t)}{R},\\
-				&C_b\dot x_b =  \frac{x_a(t) - x_b(t)}{R},
-			\end{aligned}
-		\end{equation}
-		where $C_a$ and $C_b$ are the heat capacitances of the layers.
-This is done for each layer of each zone, obtaining the compact representation given in \eqref{E:RCZoneEq}.
-The thermal parameters are derived from zones geometry and material properties.
-	\item External heat fluxes, modeled as a bilinear model, affect the building directly as well as indirectly through zones:
-		\begin{equation}\label{E:RCfluxes}
-			q(t) = A_q x(t) + B_{q,u}u(t) + B_{q,d}d(t) + \sum_{i=1}^{n_u}{[\left(B_{q,du,i}d(t) + D_{q,xu,i}x(t)\right)u_i(t)]},
-		\end{equation}
-		where $u$ are the inputs and $d$ the disturbances of the system.
-%		Equation \eqref{E:RCfluxes} comes from a series of 20 equations analyzed in Section 3.3.1.3 of the Technical Report \cite{SturzeneggerTR}, that we do not report here for the sake of simplicity. 
-Equation \eqref{E:RCfluxes} for the heat flux is obtained by modeling
-	\begin{enumerate}
-		\item heat exchange associated with the building hull (except for windows), both conductive and radiative part,
-		\item  heat flux to each thermally activated building system (TABS), i.e. pipes buried in the concrete slabs of the floors carrying hot/cold water,
-		\item heat flux through the windows in three different parts: radiation due to elements directly in contact with the zone air, conduction through the window, and absorption of the solar radiation through the window,
-		\item convection due to internal gains due to occupants, appliances and lighting, and
-		\item the effects of the AHU.
-	\end{enumerate}
-	\item The resulting system \eqref{E:RCZoneEq} is discretized with a sampling time of $15\ \mathrm{min}$.
-	This model has approximately 300 states that include temperature of the zones, walls and floors on the second floor.
-	The outputs of the system are the zone temperatures.
-	Since the performance of the state estimation, that is needed to compute the optimal control inputs using MPC depends on the number of states, an approximated model with fewer states is needed.
-	To this aim, different averaged temperatures of the building facades (North, South, West, East) and the zones are considered, obtaining an approximated model with 35 states.
-	This  approximate model is then considered ``suitable for MPC" (see Section 3.3.1.4 in \cite{SturzeneggerTR}). 
-\end{enumerate}}
-
-\textcolor[rgb]{0,0,1}{In summary, the approximate model has the following state, input and disturbance variables:
-	\begin{itemize}
-		\item 5 outputs: averaged room temperature for each group of zones (North, South, West, East, Center). The rooms are equipped with temperature sensors.
-		\item 18 inputs: TABS heating heat flux, TABS cooling heat flux, averaged transmitted solar heat flux for each group of zones (North, South, West, East) that is estimated using blinds position measurements, air massflow through the energy recovering mode, air massflow bypassing the energy recovering mode, air massflow through the air cooler, AHU heat coil heat flux, lighting power for the offices for each group of zones (North, South, West, East), and radiator heat flux in the corner offices (North, South, West, East).
-		\item 7 disturbances: internal gains in the offices and internal gains in non-office zones, which are predicted using a standard schedule, ambient temperature and solar radiation on facade (North, South, West, East) whose values were obtained through Kalman filtering using measurements from the weather station placed on the roof of the building.
-This filtering is needed to take into account the shadowing of the neighboring buildings.
-	\end{itemize}
-An EnergyPlus model is built for this system, and model parameters of matrices $A_x$, $B_q$, $A_q$, $B_{q,u}$, $B_{q,d}$, $B_{q,du,i}$, $D_{q,xu,i}$ in Equations \eqref{E:RCZoneEq} and \eqref{E:RCHeatExchange} are derived using geometry and materials data from the EnergyPlus model.
-This was a choice of the authors; real geometry and materials data can also be used to estimate the model parameters, if necessary data are available. 
-For this particular building, 24 parameters are estimated/taken form a datasheet/computed for the considered zone model.
-Although some of the parameters are in common between different zones, the others are found independently for each zone.
-To simplify, the parameters of all the other floors are assumed to be identical which is not true in reality.
-To derive the EnergyPlus model using the available measurements and to use the same control implemented in the building, the EnergyPlus model is coupled with MATLAB using BCVTB \cite{Wetter2015}.
-}
-%To identify the model a total of 294 MATLAB signals were sent to the EnergyPlus model and a total of 1081 EnergyPlus signals were sent back to Matlab.
-
-\textcolor[rgb]{0,0,1}{The physics-based approach, although detailed and accurate (in some cases), is cost and time prohibitive.
-The long and complex modeling procedure that is unique to every building poses practical challenges in applying MPC to large scale buildings.}
-
-\subsubsection{Data-driven modeling}
-\textcolor[rgb]{0,0,1}{The data-driven approach we propose, on the other hand, ignores the physical details of the building operation.
-Instead, we use machine learning to learn black-box models where our objective is to find the hyperparameters (for a given structure like Random Forests) of a model that best explains the input-output relationship.}
-
-\textcolor[rgb]{0,0,1}{The goal with the data-driven modeling is to learn a function map
-\begin{equation}
-\label{E:MLexample}
-y(t) = f(u(t),\dots,u(t-\delta_u), d(t),\dots, d(t-\delta_d)),
-\end{equation}
-where \(y\) represents the outputs, \(u\) the inputs and \(d\) the disturbances as defined earlier, and \(\delta\) is a lag variable.}
-
-\textcolor[rgb]{0,0,1}{The data-driven modeling in the case of buildings has several advantages over the physical counterpart.}
-\begin{enumerate}
-
-	\item \textcolor[rgb]{0,0,1}{\textbf{Domain expertise:} 
-	This procedure (\textcolor[rgb]{1,0,0}{NOT CLEAR IF EXPLAINED IN ONLY 2 LINES BEFORE}) eliminates the need of a building expert, simplifies the modeling complexities (\textcolor[rgb]{1,0,0}{HOW?}) described in Section \ref{SSS:physics-based} which ultimately reduces the cost of modeling a building.
-Further, for a new building, given the historical data from the building, the same process can be repeated to identify a control-oriented model for MPC.}
-
-	\item \textcolor[rgb]{0,0,1}{\textbf{Sensor cost:}
-	Due to large number of states and variables in the physical modeling, a lot of measurements are needed to use the model to predict the system's behavior \textcolor[rgb]{1,0,0}{IF USING A MODEL-BASED APPROACH}.
-	This can be expensive due to necessary new sensor installations.
-	Moreover, when some measurements are not directly available from the sensors, observers are needed for state estimations.
-	However, observability problems can limit the construction of such observers \cite{Dorf2011MCS}. 
-	On the other hand, data-driven modeling obviates the need to model the state of the walls and floor layers\textcolor[rgb]{1,0,0}{AND OTHER STATES DIFFICULT TO MEASURE} so we do not have as many states.\textcolor[rgb]{1,0,0}{EXPLAIN THAT THE MACHINE LEARNING ALGORITHM IS ABLE TO COMPENSATE MISSING VARIABLES}
-	It relies only on direct measurements from the sensors such as thermostats and multimeters reducing the need for new installations and hence the cost of modeling.
-	Further, when some of the states are not directly measurable, often adding proxy \textcolor[rgb]{1,0,0}{I.E. ??}variables suffices. 
-	For example, if we do not have the measurement of the energy produced by a photovoltaic panel, the weather conditions will account for it.}
-
-	\item \textcolor[rgb]{0,0,1}{\textbf{Linearity assumption:}
-	If \(y\) in Equation \eqref{E:MLexample} is the average room temperature for one of the zones, a nonlinear function \(f\) is more appropriate, for example \(f\) may belong to the class of Neural Networks or Random Forests.
-	During physical modeling in \eqref{E:RCHeatExchange}, in order to keep the model simple, the heat exchange between layers is modeled with a linear behavior. 
-	Hence all the nonlinearities were neglected.
-	In the case of heat flux in \eqref{E:RCfluxes}, again, a linear model is not good enough to provide a good approximation. For this reason a bilinear model was considered.
-	However, more complex nonlinearities have been neglected due to the already high complexity of the model \cite{Sturzenegger2016}. 
-	In data-driven modeling, we can relax these linearity assumptions as nonlinear functions can be learned rather efficiently and accurately.\textcolor[rgb]{1,0,0}{IT IS NOT CLEAR THAT THE DD ALGORITHM TAKE CARE OF IT CREATING ITS INTERNAL STRUCTURE}}
-
-	\item \textcolor[rgb]{0,0,1}{\textbf{Other assumptions:} 
-	Due to the high modeling complexity, physical modeling neglects different geometries and different materials of the floors assuming they are the same for each floor which is never true in reality.
-	In many cases, the details of construction layout and equipments are not even available so many parameters have to be guessed.
-	Data-driven modeling depends on the sensor data directly. 
-	Hence, such assumptions never arise.}
-
-\end{enumerate}
-
-\textcolor[rgb]{0,0,1}{To summarize, the data-driven modeling simplifies the modeling of buildings to a large extent, reduces the overall cost and time investment while avoiding the linear model assumptions that we make in standard physical modeling procedure.
-Unlike physical modeling, the parameters of the model (or hyperparamters) are automatically determined by the learning algorithm, the best settings determined by cross validation.}
-
-\textcolor[rgb]{0,0,1}{However, the challenge lies in using such models for optimal predictive control, which is exactly the focus of this paper.
-We demonstrate how the training algorithm for Random Forests can be modified to develop control oriented models that are suitable in the MPC framework.}
+%====================================================================================================================================================================================
 
@@
 A vast literature exists in building energy applications that deals with Demand Response, peak power reduction, energy saving, thermal comfort, and related topics.
-All these approaches can be classified based on two characteristics:
-\begin{enumerate}
-	\item the type of system model:
-		\begin{itemize}
-			\item model-based, that we consider both "white-box" and "grey-box" approaches: \cite{Shakouri2017SCS,Li2014EB,Yoon2014EB,Li2016AE,Harb2016EB,Salakij2016EB,Li2016E,Li2016EB,Hou2013,Cecconi2017EB};
-			\item data-based, that is the "black-box" approach, mainly done using Neural Networks: \cite{Safa2017SCS,Neto2008EB,Magnier2010BE,Candanedo2017EB,Ascione2017E,Cecconi2017EB,Li2016AE};
-			\item simulation tool-based, such as EnergyPlus \cite{energyPlus} and TRNSYS \cite{trnsys2000}: \cite{Yin2016EB,Christantoni2016EB};
-		\end{itemize}
-	\item the purpose these models are created for:
-		\begin{itemize}
-			\item only model creation and identification: \cite{Safa2017SCS,Neto2008EB,Magnier2010BE,Li2014EB,Li2016AE,Harb2016EB,Li2016EB,Candanedo2017EB,Cecconi2017EB,Ascione2017E};
-			\item model identification and control (typically Predictive Control): \cite{Shakouri2017SCS,Yoon2014EB,Yin2016EB,Salakij2016EB,Li2016E,Hu2017AE}.
-		\end{itemize}
-\end{enumerate}
-These papers are summarizes in Table \ref{T:RelatedWork} highlighting the differences.
-We also emphasize the case studies the results are applied to, and whether the authors use experimental data to simulate their algorithms.
-Only in three cases the algorithms are tested on real systems (see Table: RI - Real Implementation).
-We observe that except for the last 6 cases, which we discuss in detail, either model-based approaches or only tools are considered with/without control or data-driven modeling approaches without control.
-%This is because bridging data-driven models with control theory, especially with the predictive control, introduces several issues \cite{Hou2013}.
-Data-driven control has attracted increasing attraction in the recent years.
-The last six papers of Table \ref{T:RelatedWork} that are more related to the methodology presented in this paper.
-In particular, in \cite{Macarulla2017} the authors proposed a predictive control strategy based on Neural Networks, for boilers control in buildings, to decide the optimal time to switch-on the plan to guarantee energy saving and thermal comfort.
-However, the approach is not easily scalable to different types of plants and does not use optimization in the closed-loop scheme.
-In \cite{Costanzo2016} a reinforcement learning control strategy, called Model-Assisted Batch Reinforcement Learning, is considered to provide data-driven control for the demand response problem in HVAC systems.
-Reinforcement Learning is a model-free methodology and is an alternative approach to MPC \cite{Ernst2009TSMC} (with pros and cons).
-In \cite{Ferreira2012} the authors considered a data-driven predictive control based on Neural Networks to guarantee energy saving and  thermal comfort in public buildings.
-Neural Networks are used in the closed-loop control scheme to determine a thermal comfort index based on parameters that can be measured or estimated, but no Neural Networks-based system state dynamics are included into the optimization problem.
-%However, Neural Networks are a particular type of machine learning algorithm that provides complex nonlinear equations to describe the system's dynamics.
-\cite{Afram2017} uses a Neural Network based data-driven state model as a plant simulator in the MPC closed-loop optimization.
-More papers related to this topic can be found in the literature.
-Unfortunately, since Neural Networks models are nonlinear, the MPC based on such models is also nonlinear.
-This means that a global optimal solution can not be guaranteed. 
-When the complexity of network is high, solving the optimization problem becomes harder due to nonlinearties.
-To overcome this complexity, we introduced regression trees-based approach in \cite{Behl2016,Jain2017TCPS}.
-In this paper, we provide a new methodology based on random forests, that overcome the drawbacks of our previous work.
-It provides robustness with respect to disturbance uncertainties and has better performance when compared to other approaches (both optimal and rule-based).
-In the next section, we emphasize the differences with our previous approaches.
 }
 
+\textcolor[rgb]{0,0,1}{
 \begin{table}[t!]
@@
 \end{table}
+}
 
 \textcolor[rgb]{0,0,1}{
-\subsubsection{Previous work}\label{SSS:PreviousWork}
-We highlight the differences with respect to our previous work.
+All these approaches can be classified based on two characteristics:
+\begin{enumerate}
+	\item the type of system model:
 		\begin{itemize}
-	\item In \cite{Behl2016} we use regression trees to setup an MPC problem and apply it to the problem of Demand-Response.
-	However, these models can be used for optimal control with only one-step lookahead prediction.
-	Hence, it is not possible to use such approach to control the system considering a prediction over an horizon of arbitrary length.
-	\item This limitation is addressed in \cite{Jain2017TCPS}, where multi-output regression trees are used instead of single output trees.
-	The different outputs correspond to different steps of the prediction horizon.
-	This allows us to setup an MPC problem with a finite horizon.
+			\item model-based, that we consider both "white-box" and "grey-box" approaches: \cite{Shakouri2017SCS,Li2014EB,Yoon2014EB,Li2016AE,Harb2016EB,Salakij2016EB,Li2016E,Li2016EB,Hou2013,Cecconi2017EB};
+			\item data-based, that is the "black-box" approach, mainly done using Neural Networks: \cite{Safa2017SCS,Neto2008EB,Magnier2010BE,Candanedo2017EB,Ascione2017E,Cecconi2017EB,Li2016AE};
+			\item simulation tool-based, such as EnergyPlus \cite{energyPlus} and TRNSYS \cite{trnsys2000}: \cite{Yin2016EB,Christantoni2016EB};
+		\end{itemize}
+	\item the purpose these models are created for:
+		\begin{itemize}
+			\item only model creation and identification: \cite{Safa2017SCS,Neto2008EB,Magnier2010BE,Li2014EB,Li2016AE,Harb2016EB,Li2016EB,Candanedo2017EB,Cecconi2017EB,Ascione2017E};
+			\item model identification and control (typically Predictive Control): \cite{Shakouri2017SCS,Yoon2014EB,Yin2016EB,Salakij2016EB,Li2016E,Hu2017AE}.
+		\end{itemize}
+\end{enumerate}
+}
+
+\textcolor[rgb]{0,0,1}{
+These papers are summarizes in Table \ref{T:RelatedWork} highlighting the differences.
+We also emphasize the case studies the results are applied to, and whether the authors use experimental data to simulate their algorithms.
+Only in three cases the algorithms are tested on real systems (see Table: RI - Real Implementation).
+We observe that except for the last 6 cases, which we discuss in detail, either model-based approaches or only tools are considered with/without control or data-driven modeling approaches without control.
+%This is because bridging data-driven models with control theory, especially with the predictive control, introduces several issues \cite{Hou2013}.
+Indeed data-driven control has attracted increasing attraction only in the recent years.
+The last six papers of Table \ref{T:RelatedWork} are more related to the methodology presented in this paper.
+In particular, in \cite{Macarulla2017} the authors proposed a predictive control strategy based on Neural Networks, for boilers control in buildings, to decide the optimal time to switch-on the plan to guarantee energy saving and thermal comfort.
+However, the approach is not easily scalable to different types of plants and does not use optimization in the closed-loop scheme.
+In \cite{Costanzo2016} a reinforcement learning control strategy, called Model-Assisted Batch Reinforcement Learning, is considered to provide data-driven control for the demand response problem in HVAC systems.
+Reinforcement Learning is a model-free methodology and is an alternative approach to MPC \cite{Ernst2009TSMC} (with pros and cons).
+In \cite{Ferreira2012} the authors considered a data-driven predictive control based on Neural Networks to guarantee energy saving and  thermal comfort in public buildings.
+Neural Networks are used in the closed-loop control scheme to determine a thermal comfort index based on parameters that can be measured or estimated, but no Neural Networks-based system state dynamics are included into the optimization problem.
+%However, Neural Networks are a particular type of machine learning algorithm that provides complex nonlinear equations to describe the system's dynamics.
+\cite{Afram2017} uses a Neural Network based data-driven state model as a plant simulator in the MPC closed-loop optimization.
+More papers related to this topic can be found in the literature.
+Unfortunately, since Neural Networks models are nonlinear, the MPC based on such models is also nonlinear.
+This means that a global optimal solution can not be guaranteed and, when the complexity of network is high, that solving the optimization problem becomes computationally harder due to nonlinearties.
+\\
+To overcome this complexity we initially introduced the regression trees-based approach in \cite{Behl2016,Jain2017TCPS}.
+In particular, in \cite{Behl2016} we used regression trees to setup an MPC problem and apply it to the problem of Demand-Response.
+However, these models can be used for optimal control with only one-step lookahead prediction.Hence, it is not possible to use such approach to control the system considering a prediction over an horizon of arbitrary length.
+This limitation was addressed in \cite{Jain2017TCPS}, where multi-output regression trees are used instead of single output trees.
+The different outputs correspond to different steps of the prediction horizon, which allows us to setup an MPC problem with a finite horizon.
 Modeling accuracy using single trees is strongly affected by overfitting and high variance.
 Such approach has the advantage to be extremely simple from the complexity point of view, but the range of applications is limited.
-	\item In \cite{JainACC2017,JainCDC2017}, we take a different approach that improves the system's identification accuracy and robustness in MPC.
-	Instead of considering a single tree/forest with multiple output, we consider multiple trees/forests with single output.
+In \cite{JainACC2017,JainCDC2017} we took a different approach that improves the system's identification accuracy: instead of considering a single tree/forest with multiple output, we considered multiple trees/forests with single output.
 Each tree/forest provides the prediction of the system's behavior for a different time step of the horizon.
-	However, the results are based on simulated data and we did not account for inaccuracies in the weather forecast.
-\end{itemize}
+However, the results were based on simulated data and we did not account for inaccuracies in the weather forecast.\\
+In this paper we provide a new methodology based on random forests that overcomes the drawbacks of all our previous work, and more precisely: we obtain better performance when compared to other approaches (both optimal and rule-based), we provide and validate robustness with respect to disturbance uncertainties and, last but not least, we test our method using a real data set.
+In the next section, we will illustrate in detail such contributions.
 }
 
+%====================================================================================================================================================================================
+
+%====================================================================================================================================================================================
+
+\textcolor[rgb]{0,0,1}{
 \subsection{Main contribution and paper organization}
@@
 %\end{itemize}
-In our previous work \cite{Behl2016,Jain2016,JainACC2017,JainCDC2017}, we introduced the concept of DPC for receding horizon control.
+%In our previous work \cite{Behl2016,Jain2016,JainACC2017,JainCDC2017}, we introduced the concept of DPC for receding horizon control.
 In this paper, we extend these results providing the following contributions.
+The presentation of the paper contributions is organised as follows:
 \begin{enumerate}
-	\item In Section \ref{S:dpc}, we formally present the two following control techniques:
+	\item In Section \ref{S:dpc} we formally describe the data-predictive control (DPC) techniques.
+For simplicity of presentation we first illustrate DPC based regression trees, then we introduce DPC based on random forests.
+	\item In Section \ref{S:proof} we demonstrate that DPC can provide comparable performance compared to MPC applied to a perfectly known mathematical model.
+More precisely, we first consider a benchmark MPC controller using a bilinear building model whose parameters were identified using experiments on a building in Switzerland: then we apply DPC to a data-set simulated from such bilinear model and extract a controller, which is based only on such data and, of course, ignores the mathematical model.
+We show that DPC captures 70\% variance in MPC and offers a comparable performance.
+	\item In Section \ref{S:casestudy} we demonstrate scalability and efficiency of DPC describing a practical application of DPC to a Demand Response problem based in a simulative benchmark.
+More precisely, we apply DPC to a 6 story 22 zone building model in EnergyPlus \cite{Crawley2001} for which model-based control is not economical and practical due to extreme complexity.
+We observe that DPC provides the desired power reduction with an average error of 3\%: this shows scalability and efficiency of DPC in providing financial incentives to the end-customers bypassing the need for high fidelity models.
+	\item In Section \ref{S:realCaseStudy} we implement and test DPC basing on real data of an off-grid house located in L'Aquila (Italy).
+More precisely, we use DPC to design the optimal ON/OFF scheduling for the heating system in order to save energy while guaranteeing thermal comfort for the occupants.
+We quantify the total amount of energy saved, with respect to the classical bang-bang controller widely used in houses for temperature control, using an EnergyPlus model built specifically for the house.
+We show that we can perform an energy saving that goes from $25.4\%$ (if we guarantee thermal comfort by strictly respecting the desired temperature ranges in the rooms) to $49.2\%$ (if we allow small violation of the desired temperature ranges in the rooms).
+In addition, we test the robustness of our method with respect to substantial prediction errors in the weather forecast.
+\end{enumerate}
+}
+%\paragraph{Example} \textcolor[rgb]{0.00,0.00,1.00}{As a first} example of such modeling complexity, consider the grey-box approach in \cite{Braun2002}.
+The scope of such approach is to predict the heat transfer rate to the air within the building.
+In particular, the authors addressed a single zone modeling problem using an equivalent $RC$ model.
+To this end, five different types of structures are considered in the example: external walls, ceiling/roof, floor, internal walls and windows.
+Each of these elements are represented using $3$ resistances, $2$ capacitances and $2$ temperatures, except for the windows that are represented only with resistances since they have a negligible energy storage.
+A model consisting of $8$ states and $9$ inputs, including disturbances such as solar radiation and others, with the instantaneous heat gain to the building air from all surfaces as output, is created for the zone.
+To have a complete LTI model, $9$ parameters are needed to be estimated via a non-trivial training process, using a collection of information associated with a physical description of the building (see \cite{Braun2002} for more details).
+Extending this result, for the building we consider in Section \ref{S:realCaseStudy} with $10$ different zones, we would need to consider a model with approximately $80$ states and $90$ parameters to be estimated.
+
+%====================================================================================================================================================================================
+
+%====================================================================================================================================================================================
+
+\section{COMPLEXITY ISSUES IN MODELING BUILDINGS: PHYSICS-BASED VS DATA-DRIVEN}\label{secModelingIssuesExample}
+\textcolor[rgb]{0,0,1}{In this section we provide a detailed example to emphasise differences in terms of complexity when modeling a building using physical laws vs.
+machine learning.
+We show that the data-driven modeling eliminates several drawbacks that occur using physics-based models such as the need to have very good knowledge of the building structure and the material properties, time required to build a model and limited availability of sensors.
+The example is based on a real building taken from \cite{Sturzenegger2016}, which is located in Allschwil (Switzerland).
+The building consists of 6 floors, with a total air conditioned floor area of around $6000\ \mathrm{m^2}$.
+We will first illustrate the physics-based modeling approach in \cite{Sturzenegger2016}, then we will emphasise some drawbacks and we finally illustrate how data-driven approaches can overcome such issues.
+}
+
+%====================================================================================================================================================================================
+
+\subsection{Physics-based modeling}
+\label{SSS:physics-based}
+
+\textcolor[rgb]{0,0,1}{We describe the physics-based approach based on an RC network outlined in \cite{Sturzenegger2016}, which derives a bilinear model (constructed from physical principles) of the second floor of the building.
+The model of the whole building is derived in \cite{Sturzenegger2016} assuming that such floor is identical to all other floors.
+The modeling process of the second floor consists of three steps:
+}
+
+\textcolor[rgb]{0,0,1}{(1) Building geometry and construction data are used together with first-principles to derive the following linear model for the building's thermal dynamics:
+		\begin{equation}\label{E:RCZoneEq}
+			\dot x(t) = A_x x(t) + B_q q(t).
+		\end{equation}
+		This model describes the behavior of the zone, wall, floor and ceiling temperatures.
+		Walls, floors and ceilings are considered as divided into layers with different features.
+Therefore, each zone was described with an RC network model (see Figure 3-10 in \cite{SturzeneggerTR}), where the capacitances represent the states of the layers and the resistances represent the thermal resistance of the layers.
+%		\begin{figure}[t]
+%			\begin{center}
+%				\includegraphics[width=0.6\textwidth]{figures/RC_net.eps}
+%				\caption{RC model of a zone. $q$ represents the external heat fluxes}\label{F:RCnet}
+%			\end{center}
+%		\end{figure}
+		The heat exchange between two adjacent layers, i.e.
+layer ``a" and layer ``b", is modeled to be proportional to the temperature difference of the two layers and the corresponding thermal resistance $R$ given by
+		\begin{equation}\label{E:RCHeatExchange}
+			\begin{aligned}
+				&C_a\dot x_a = \frac{x_b(t) - x_a(t)}{R},\\
+				&C_b\dot x_b =  \frac{x_a(t) - x_b(t)}{R},
+			\end{aligned}
+		\end{equation}
+		where $C_a$ and $C_b$ are the heat capacitances of the layers.
+This is done for each layer of each zone, obtaining the compact representation given in \eqref{E:RCZoneEq}.
+The thermal parameters are derived from zones geometry and material properties.
+}
+
+\textcolor[rgb]{0,0,1}{(2) External heat fluxes, modeled as a bilinear model, affect the building directly as well as indirectly through zones:
+		\begin{equation}\label{E:RCfluxes}
+			q(t) = A_q x(t) + B_{q,u}u(t) + B_{q,d}d(t) + \sum_{i=1}^{n_u}{[\left(B_{q,du,i}d(t) + D_{q,xu,i}x(t)\right)u_i(t)]},
+		\end{equation}
+		where $u$ are the inputs and $d$ the disturbances of the system.
+%		Equation \eqref{E:RCfluxes} comes from a series of 20 equations analyzed in Section 3.3.1.3 of the Technical Report \cite{SturzeneggerTR}, that we do not report here for the sake of simplicity. 
+Equation \eqref{E:RCfluxes} for the heat flux is obtained by modeling:
+	\begin{itemize}
+		\item heat exchange associated with the building hull (except for windows), both conductive and radiative part;
+		\item  heat flux to each thermally activated building system (TABS), i.e. pipes buried in the concrete slabs of the floors carrying hot/cold water;
+		\item heat flux through the windows in three different parts: radiation due to elements directly in contact with the zone air, conduction through the window, and absorption of the solar radiation through the window;
+		\item convection due to internal gains due to occupants, appliances and lighting;
+		\item the effects of the AHU.
+	\end{itemize}
+}
+
+\textcolor[rgb]{0,0,1}{(3) The resulting system \eqref{E:RCZoneEq} is discretized with a sampling time of $15\ \mathrm{min}$.
+This model has approximately 300 states that include temperature of the zones, walls and floors on the second floor.
+The outputs of the system are the zone temperatures.
+Since the performance of the state estimation, needed to compute the optimal control inputs using MPC, badly scales with the number of states, an approximated model with fewer states is derived in \cite{SturzeneggerTR}.
+In particular, although the rooms are equipped with temperature sensors, different averaged temperatures of the building facades (North, South, West, East) and of the zones are considered, obtaining a coarser model with only 35 internal variables.
+Among these 35 internal variables only 5 are measurable output variables, i.e.
+the averaged room temperature for each group of zones (North, South, West, East, Center).
+The system has 18 input variables: TABS heating heat flux, TABS cooling heat flux, averaged transmitted solar heat flux for each group of zones (North, South, West, East) that is estimated using blinds position measurements, air mass-flow through the energy recovering mode, air mass-flow bypassing the energy recovering mode, air massflow through the air cooler, AHU heat coil heat flux, lighting power for the offices for each group of zones (North, South, West, East), and radiator heat flux in the corner offices (North, South, West, East).
+Finally, 7 disturbance signals are considered: internal gains in the offices and internal gains in non-office zones, which are predicted using a standard schedule, ambient temperature and solar radiation on facade (North, South, West, East) whose values were obtained through Kalman filtering using measurements from the weather station placed on the roof of the building.
+This filtering is needed to take into account the shadowing of the neighboring buildings.
+This approximate model is then considered ``suitable for MPC" (see Section 3.3.1.4 in \cite{SturzeneggerTR}).\\
+An EnergyPlus model is built for this system, and model parameters of matrices $A_x$, $B_q$, $A_q$, $B_{q,u}$, $B_{q,d}$, $B_{q,du,i}$, $D_{q,xu,i}$ in Equations \eqref{E:RCZoneEq} and \eqref{E:RCHeatExchange} are derived using geometry and materials data from the EnergyPlus model.
+This was a choice of the authors: real geometry and materials data can also be used to estimate the model parameters, if necessary data are available.
+For this particular building, 24 parameters are estimated/taken form a datasheet/computed for the considered zone model.
+Although some of the parameters are in common between different zones, the others are found independently for each zone.
+As already discussed, in \cite{SturzeneggerTR} the parameters of all the other floors are assumed to be identical to the second floor, which potentially introduces substantial modeling uncertainties.
+To derive the EnergyPlus model using the available measurements and to use the same control implemented in the building, the EnergyPlus model is coupled with MATLAB using BCVTB \cite{Wetter2015}.\textcolor[rgb]{1,0,0}{Ale: THIS PARAGRAPH IS UNCLEAR}\\
+The physics-based approach described above, although detailed and accurate (in some cases), is clearly cost and time prohibitive, and as the building characteristics changes with time the system identification must be repeated to update the model.
+Moreover, such expensive, time-consuming and complex modeling procedure is unique (and not repeatable) for other buildings.
+For all these reasons physics-based modeling suffers practical challenges when the objective is applying MPC to large scale buildings.
+}
+
+%To identify the model a total of 294 MATLAB signals were sent to the EnergyPlus model and a total of 1081 EnergyPlus signals were sent back to Matlab.
+
+%====================================================================================================================================================================================
+
+\subsection{Data-driven modeling}
+
+\textcolor[rgb]{0,0,1}{The goal with the data-driven modeling is to learn, based on historical measurable data and without requiring physical details of a building, a function map
+\begin{equation}\label{E:MLexample}
+	y(t) = f(u(t),\dots,u(t-\delta_u), d(t),\dots, d(t-\delta_d)),
+\end{equation}
+where \(y\) represents the outputs, \(u\) the inputs and \(d\) the disturbances as defined earlier, and \(\delta\) is a memory variable.
+Note that the internal unmeasurable variables used in the previous section to model the building are not considered here.
+In this paper we use machine learning to learn such black-box models, with the objective of finding the hyperparameters (for a given structure like Random Forests \textcolor[rgb]{1,0,0}{Ale: referring to a structure like random forests here is confusing}) of a model that best explain the input-output relationship within the measured variables, compensating the effect of the internal unmeasurable variables.
+The advantages of data-driven modeling with respect to physics-based modeling in the case of buildings can be summarised as follows:}
+\textcolor[rgb]{1,0,0}{Ale: we need to improve the following list: first we can consider the list of drawbacks of the physics-based modeling that Francesco introduced in a previous version and comment for each the corresponding improvement via data-driven approach; then we also need to improve much more the text, which sometimes is unclear and verbose.}
+
 \begin{enumerate}
-		\item DPC with regression trees, and
-		\item DPC with random forests.
+	\item \textcolor[rgb]{0,0,1}{\textbf{Domain expertise:} 
+	This procedure (\textcolor[rgb]{1,0,0}{NOT CLEAR IF EXPLAINED IN ONLY 2 LINES BEFORE}) eliminates the need of a building expert, simplifies the modeling complexities (\textcolor[rgb]{1,0,0}{HOW?}) described in Section \ref{SSS:physics-based} which ultimately reduces the cost of modeling a building.
+Further, for a new building, given the historical data from the building, the same process can be repeated to identify a control-oriented model for MPC.}
+	\item \textcolor[rgb]{0,0,1}{\textbf{Sensor cost:}
+	Due to large number of states and variables in the physical modeling, a lot of measurements are needed to use the model to predict the system's behavior \textcolor[rgb]{1,0,0}{IF USING A MODEL-BASED APPROACH}.
+	This can be expensive due to necessary new sensor installations.
+	Moreover, when some measurements are not directly available from the sensors, observers are needed for state estimations.
+	However, observability problems can limit the construction of such observers \cite{Dorf2011MCS}. 
+	On the other hand, data-driven modeling obviates the need to model the state of the walls and floor layers\textcolor[rgb]{1,0,0}{AND OTHER STATES DIFFICULT TO MEASURE} so we do not have as many states.\textcolor[rgb]{1,0,0}{EXPLAIN THAT THE MACHINE LEARNING ALGORITHM IS ABLE TO COMPENSATE MISSING VARIABLES}
+	It relies only on direct measurements from the sensors such as thermostats and multimeters reducing the need for new installations and hence the cost of modeling.
+	Further, when some of the states are not directly measurable, often adding proxy \textcolor[rgb]{1,0,0}{I.E. ??}variables suffices. 
+	For example, if we do not have the measurement of the energy produced by a photovoltaic panel, the weather conditions will account for it.}
+	\item \textcolor[rgb]{0,0,1}{\textbf{Linearity assumption:}
+	If \(y\) in Equation \eqref{E:MLexample} is the average room temperature for one of the zones, a nonlinear function \(f\) is more appropriate, for example \(f\) may belong to the class of Neural Networks or Random Forests.
+	During physical modeling in \eqref{E:RCHeatExchange}, in order to keep the model simple, the heat exchange between layers is modeled with a linear behavior. 
+	Hence all the nonlinearities were neglected.
+	In the case of heat flux in \eqref{E:RCfluxes}, again, a linear model is not good enough to provide a good approximation. For this reason a bilinear model was considered.
+	However, more complex nonlinearities have been neglected due to the already high complexity of the model \cite{Sturzenegger2016}. 
+	In data-driven modeling, we can relax these linearity assumptions as nonlinear functions can be learned rather efficiently and accurately.\textcolor[rgb]{1,0,0}{IT IS NOT CLEAR THAT THE DD ALGORITHM TAKE CARE OF IT CREATING ITS INTERNAL STRUCTURE}}
+	\item \textcolor[rgb]{0,0,1}{\textbf{Other assumptions:} 
+	Due to the high modeling complexity, physical modeling neglects different geometries and different materials of the floors assuming they are the same for each floor which is never true in reality.
+	In many cases, the details of construction layout and equipments are not even available so many parameters have to be guessed.
+	Data-driven modeling depends on the sensor data directly. 
+	Hence, such assumptions never arise.}
 \end{enumerate}
-	\item In Section \ref{S:proof}, we demonstrate the strength of DPC for receding horizon control via one-to-one comparison against a benchmark MPC controller using a bilinear building model whose parameters were identified using experiments on a building in Switzerland.
-We show that DPC captures 70\% variance in MPC and offers a comparable performance.
-	\item Section \ref{S:casestudy} describes a practical application of DPC for Demand Response, where we apply DPC to a 6 story 22 zone building model in EnergyPlus \cite{Crawley2001} for which model-based control is not economical and practical due to extreme complexity.
-We show scalability and efficiency of DPC in providing financial incentives to the end-customers bypassing the need for high fidelity models.
-We observe that DPC provides the desired power reduction with an average error of 3\%.
-	\item In Section \ref{S:realCaseStudy}, we implement DPC on the real data from an off-grid house located in L'Aquila (Italy), to find the optimal ON/OFF scheduling for the heating system in order to save energy while guaranteeing thermal comfort for the occupants.
-We quantify the total amount of energy saved, with respect to the classical bang-bang controller widely used in houses for temperature control, using an EnergyPlus model built specifically for the house.
-We show we can perform an energy saving that goes from $25.4\%$, if we guarantee thermal comfort, to $49.2\%$, if we allow very little discomfort in terms of rooms' temperature.
-\end{enumerate}
+
+\textcolor[rgb]{0,0,1}{To summarize the data-driven approaches have the potential to simplify the modeling of buildings to a large extent reducing the overall cost and time investment, while avoiding several assumptions that are generally made in standard physics-based modeling procedures and that reduce the accuracy.
+Also, unlike physics-based modeling, the parameters (or hyperparamters) of the model are automatically determined by the learning algorithm to compensate the effect of unobservable variables on the input-output relation, and the best settings determined by cross validation.
+Granted the advantages of data-driven modeling the challenge lies now in using such models for optimal predictive control, which is exactly the focus of this paper.
+In the next sections we will demonstrate how the training algorithm for Random Forests can be modified to develop control oriented models that enable Model Predictive Control with better performance when compared to other approaches and robustness with respect to disturbance uncertainties.}

### 2018-01-20 paper/final/casestudyEPlus.tex
small changes

@@
 (3) The \textit{building data} include (i) cooling set points for  the guest rooms, kitchen and corridors, (ii) supply air temperature, and (iii) chilled water temperature.
-For the following simulations, we use five control variables (i) cooling set point for corridors $\mathsf{ClgSP}$, (ii) cooling set point for guest rooms $\mathsf{GuestSP}$, (iii) cooling set point for kitchen $\mathsf{KitchenSP}$, (iv) chilled water supply temperature $\mathsf{ChwSP}$, and (v) supply air temperature  $\mathsf{SupplyAirSP}$, so $\tX^c = [\mathsf{ClgSP},\mathsf{GuestClgSP},\mathsf{KitchenClgSP},\mathsf{SupplyAirSP}, \mathsf{ChwSP}]$.
-The power forest $\mathcal{R}_p$ is built using the total building power consumption $\tP$.
-Its features $\tX^d$ include the weather variables, their lag terms and their forecast over the horizon, the schedule variables, and finally the lag terms of the power consumption.
-The temperature forest $\mathcal{R}_t$ is built with zone temperature $\tT$ as the output. Except for the lag terms corresponding to the same zone temperature, all other features are same in $\tX^d$.
+For the following simulations, we use five control variables (i) cooling set point for corridors $\mathsf{ClgSP}$, (ii) cooling set point for guest rooms $\mathsf{GuestSP}$, (iii) cooling set point for kitchen $\mathsf{KitchenSP}$, (iv) chilled water supply temperature $\mathsf{ChwSP}$, and (v) supply air temperature  $\mathsf{SupplyAirSP}$, so $u := [\mathsf{ClgSP},\mathsf{GuestClgSP},\mathsf{KitchenClgSP},\mathsf{SupplyAirSP}, \mathsf{ChwSP}]^\top$.
+The power forest $\mathcal{F}_p$ is built using the total building power consumption $\tP$.
+Its features in $\X^d$ include the weather variables, their lag terms and their forecast over the horizon, the schedule variables, and finally the lag terms of the power consumption.
+The temperature forest $\mathcal{R}_t$ is built with zone temperature $\tT$ as the output.
+Except for the autoregressive terms corresponding to the same zone temperature, all other features are same in $\X^d$.
 Figure~\ref{F:sepvars} shows the prediction accuracy for the power forest, and also explains the two level training approach introduced in Section~\ref{SS:sepvar}.
@@
 	\includegraphics[width=30pc]{figures/eplus_validation.eps}	
-	\caption{Model accuracy during training: The prediction made by forest using only $\tX^d$ (red) captures the effect due to disturbances.
-The linear models in the leaves capture the local effects (green) due to the control inputs $\tX^c$ and improve the model accuracy.}
+	\caption{Model accuracy during training: The prediction made by forest using only $\X^d$ (red) captures the effect due to disturbances.
+The linear models in the leaves capture the local effects (green) due to the control inputs in $\X^c$ and improve the model accuracy.}
 	\label{F:sepvars}
@@
 \end{figure}
-This is done by solving the following optimization problem with control variables $\tX^c = [\mathsf{ClgSP},\mathsf{GuestClgSP},\mathsf{KitchenClgSP},\mathsf{SupplyAirSP}, \mathsf{ChwSP}]$ as defined before:
+This is done by solving the following optimization problem, with control variables defined before:
 \begin{align}
 \begin{aligned}
-\text{min } & \ \ \ \ \ \ \ \sum_{j=1}^{N} ({\tP}_{\mathrm{k+j|k}} - \tP_{\mathrm{ref}})^2 +  \lambda\epsilon_j + \nu \delta_j \\
-\text{s.~t. } & \ \ \ \ \ \tP_{\mathrm{k+j|k}} =  \hat{\Theta}^T_{\tP_j} [ 1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
-& \ \ \ \ \ \tT_{\mathrm{k+j|k}} =  \hat{\Theta}^T_{\tT_j} [ 1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
-& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tP}-\epsilon_j \leq \tP_{\mathrm{k+j|k}} \leq \bar{\tP} + \epsilon_j\\
-& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tT}-\delta_j \leq \tT_{\mathrm{k+j|k}} \leq \bar{\tT} + \delta_j\\
-& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tX}^c \leq \tX^c_{\mathrm{k+j-1|k}} \leq \bar{\tX}^c\\ 
-& \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0,\delta_j \geq 0, \ j = 1,\dots,N.
-\end{aligned}
-\label{E:ptrack}
+& \underset{u_{k+j-1},\epsilon_j,\delta_j}{\mathrm{minimize}} & & \sum_{j=1}^{N} ({\tP}_{k+j} - \tP_{ref})^2 +  \lambda\epsilon_j + \nu \delta_j \\
+& \mathrm{subject\ to }                                       & & \tP_{k+j} =  \hat{\Theta}_{\tP_j} [ 1,u^\top_{k},\ldots,u^\top_{k+j-1} ]^\top  \\
+&                                                             & & \tT_{k+j} =  \hat{\Theta}_{\tT_j} [ 1,u^\top_{k},\ldots,u^\top_{k+j-1} ]^\top  \\
+&                                                             & & \underline{\tP} - \epsilon_j \leq \tP_{k+j} \leq \bar{\tP} + \epsilon_j        \\
+&                                                             & & \underline{\tT} - \delta_j   \leq \tT_{k+j} \leq \bar{\tT} + \delta_j          \\
+&                                                             & & \underline{u}                \leq u_{k+j-1} \leq \bar{u}                       \\
+&                                                             & & \epsilon_j \geq 0,\delta_j \geq 0, \ j = 1,\dots,N.
+\end{aligned}\label{E:ptrack}
 \end{align}

### 2018-01-20 paper/final/casestudyReal.tex
small changes

@@
 \label{SS:randomForestsModels}
-For the closed-loop simulations with DPC we create $2$ different sets of models, $S_1 = \{\tT^{r_1}_{k+1},\tT^{r_2}_{k+1},\tT^{r_3}_{k+1},\tT^{r_4}_{k+j}\}$ and $S_2 = \{\tP_{k+j},\tT^{1}_{k+j},\tT^{2}_{k+j},\tT^{3}_{k+j},\\\tT^{4}_{k+j},\ j = 1,\ldots,N\}$, using random forests.
+For the closed-loop simulations with DPC we create $2$ different sets of models, $S_1 = \{\tT^{r_1}_{k+j},\tT^{r_2}_{k+j},\tT^{r_3}_{k+j},\tT^{r_4}_{k+j}\}$ and $S_2 = \{\tP_{k+j},\tT^{1}_{k+j},\tT^{2}_{k+j},\tT^{3}_{k+j},\tT^{4}_{k+j},\ j = 1,\ldots,N\}$, using random forests.
 In each set we have  $4$ models that describe the room temperature evolution ($\tT^{r_i}$ and $\tT^i$, $i=1,2,3,4$) in each of the $4$ rooms equipped with temperature sensors.
@@
 Models in $S_2$ are used as predictors over a horizon $N$ in the DPC algorithm and are then created using the methodology provided in Section \ref{SS:dpcrf}.
-The non-manipulated features in $\tX^d$ are the disturbance data (relative humidity, atmospheric pressure, outside air temperature, solar radiation, wind, time of the day and day of the week) and the states (temperature of the $4$ rooms).
-The manipulated feature in $\tX^c$ is the flow rate ($[m^3/h]$).
-All this features are used to create models in $S_1$ and the power model in $S_2$, while disturbance data, state temperature of room $i$ only and flow rate are used to identify $\hat{\Theta}^{T}_{\tT^i_j}$ for temperature models in $S_2$.
+The non-manipulated features in $\X^d$ are the disturbance data (relative humidity, atmospheric pressure, outside air temperature, solar radiation, wind, time of the day and day of the week) and the states (temperature of the $4$ rooms).
+The manipulated feature in $\X^c$ is the flow rate ($[m^3/h]$).
+All this features are used to create models in $S_1$ and the power model in $S_2$, while disturbance data, state temperature of room $i$ only and flow rate are used to identify $\hat{\Theta}_{\tT^i_j}$ for temperature models in $S_2$.
 All the models have been trained on the data from March $11$, $2016$ to April $26$, $2016$ and validated on the data from May $1$, $2016$ to May $15$, $2016$.
@@
 \end{table}
-\begin{figure}[t!]
-	\begin{center}
-		\subfigure[Power consumption model accuracy validation. The accuracy over the testing period expressed as $\mathrm{1-NRMSE}\,(\%)$ is $92.1\%$.]{
-			\label{F:power_testing}
-			\centering
-			\includegraphics[width=26pc]{figures/power_testing.eps}
-		}
-		\subfigure[Temperature model accuracy validation for room $1$.
-The accuracy over the testing period expressed as $\mathrm{1-NRMSE}\,(\%)$ is $96.58\%$, for the temperature model $T^{r_1}$ in $S_1$ and $97.3\%$ for the temperature model $T^1$ in $S_2$.
-The accuracy for the other rooms is very similar as can be seen in Table \ref{T:S1accuracy}.]{
-			\label{F:temperature_testing}
-			\centering
-			\includegraphics[width=26pc]{figures/temperature_testing.eps}
-		}	
-	\end{center}
-	\vspace{-0.7cm}
-	\caption{Model accuracy validation for power and temperature models.}
-	\vspace{-0.7cm}
-	\captionsetup{justification=centering}
-	\label{F:testing}
-\end{figure}
 
@@
 		\begin{aligned}
-		& \underset{u_{k+j},\epsilon_j^{min},\epsilon_j^{max}}{\text{minimize}} & & \sum_{j=1}^{N} Q{\tP}_{k+j}^2 + \lambda_{min}\parallel\epsilon^{min}\parallel_2 + \lambda_{max}\parallel\epsilon^{max}\parallel_2                                                                                                                \\
-		& \text{subject to }                             & & \tP_{k+j}   = \hat{\Theta}_{\tP_j}   [ 1,u_k,\ldots,u_{k+j-1} ]^\top                                      \\
+		& \underset{u_{k+j-1},\epsilon_j^{min},\epsilon_j^{max}}{\mathrm{minimize}} & & \sum_{j=1}^{N} Q{\tP}_{k+j}^2 + \lambda_{min}\parallel\epsilon^{min}\parallel_2 + \lambda_{max}\parallel\epsilon^{max}\parallel_2                            \\
+		& \mathrm{subject\ to } & & \tP_{k+j}   = \hat{\Theta}_{\tP_j}   [ 1,u_k,\ldots,u_{k+j-1} ]^\top                         \\
 		&                    & & \tT^1_{k+j} = \hat{\Theta}_{\tT^1_j} [ 1,u_k,\ldots,u_{k+j-1} ]^\top                            \\
@@
 		&                    & & \tT^4_{k+j} = \hat{\Theta}_{\tT^4_j} [ 1,u_k,\ldots,u_{k+j-1} ]^\top                            \\
-		&                                                & & \underline{\tT}_{k+j-1}-\epsilon^{min}_j \leq \tT^3_{k+j} \leq \overline{\tT}_{k+j-1} + \epsilon^{max}_j  \\
+		&                    & & \underline{\tT}_{k+j}-\epsilon^{min}_j\leq\tT^3_{k+j}\leq\overline{\tT}_{k+j}+\epsilon^{max}_j  \\
 		&                    & & u_{k+j-1}   = \underline{u} \lor u_{k+j-1} = \overline{u}                                       \\
@@
 \end{problem}
+
+\begin{figure}[t!]
+	\begin{center}
+		\includegraphics[width=26pc]{figures/power_testing.eps}
+		\caption{Power consumption model accuracy validation. The accuracy over the testing period expressed as $\mathrm{1-NRMSE}\,(\%)$ is $92.1\%$.}
+		\label{F:power_testing}			
+	\end{center}
+\end{figure}
+
+\begin{figure}[t!]
+	\begin{center}
+		\includegraphics[width=26pc]{figures/temperature_testing.eps}
+		\caption{Temperature model accuracy validation for room $1$.
+The accuracy over the testing period expressed as $\mathrm{1-NRMSE}\,(\%)$ is $96.58\%$, for the temperature model $T^{r_1}$ in $S_1$ and $97.3\%$ for the temperature model $T^1$ in $S_2$.
+The accuracy for the other rooms is very similar as can be seen in Table \ref{T:S1accuracy}.}
+		\label{F:temperature_testing}	
+	\end{center}
+\end{figure}
 

### 2018-01-20 paper/final/cs-comparison.tex
small changes

--- 2017-09-29 paper/cs-comparison.tex
@@
 	\end{tabular}
-	\caption{Quantitative comparison of explained variance, mean value of objective function, mean input cost $c^Tu$ and mean deviance from the reference temperature $|\mathsf{T}-\mathsf{T}_{\mathrm{ref}}|$.}
+	\vspace{0.2cm}
+	\caption{Quantitative comparison of explained variance, mean value of objective function, mean input cost $c^\top u$ and mean deviance from the reference temperature $|\mathsf{T}-\mathsf{T}_{\mathrm{ref}}|$.}
 	\captionsetup{justification=centering}
@@
 	\begin{center}
-	\subfigure[External disturbances: solar gain, internal gain due to equipment and dry-bulb temperature.]{
+	\vspace{1.1cm}
+	\includegraphics[width=23pc]{figures/disturbances.eps}
+	\caption{External disturbances: solar gain, internal gain due to equipment and dry-bulb temperature.}
 	\label{F:dist}
-		\centering
-		\includegraphics[width=16pc]{figures/disturbances.eps}
-	}
-	\subfigure[Optimal control input: cooling usage factor $\mathsf{C}$ with  $0 \leq \mathsf{C} \leq 1$. DPC-En generates a control strategy very simular to MPC.]{
-		\label{F:control1}
-		\centering
-		\includegraphics[width=16pc]{figures/input3.eps}
-	}
-	\subfigure[Optimal control input: radiator heat $\mathsf{H}$ with  $0 \leq \mathsf{H} \leq 23 \ \mathrm{W/m^2}$. Again, DPC-En generates a control strategy very simular to MPC.]{
-		\label{F:control2}
-		\includegraphics[width=16pc]{figures/input4.eps}
-	}
-	\subfigure[Room temperature has time varying bounds.
-When the building is occuped the constraints are relaxed, else $19(20) \leq \mathsf{T}_{\mathrm{in}} \leq 25(26) \ \mathrm{^oC}$ in January(May).
-MPC and DPC-En are able to track the reference temperature ($22 \ \mathrm{^oC}$) closely.]{
-		\label{F:state}
-		\centering
-		\includegraphics[width=16pc]{figures/state.eps}
-	}
 \end{center}
@@
 \begin{center}
+		\subfigure[Optimal cooling factor $\mathsf{C}$. DPC-En control strategy very similar to MPC.]{
+		\label{F:control1}
+		\centering
+		\includegraphics[width=23pc]{figures/input3.eps}
+	}
+
+	\subfigure[Optimal radiator heat $\mathsf{H}$. DPC-En control strategy very similar to MPC.]{
+		\label{F:control2}
+		\includegraphics[width=23pc]{figures/input4.eps}
+	}
+
+	\subfigure[Room temp. has time varying bounds. When building is occupied, constraints are relaxed. MPC and DPC-En track the ref. temp. ($22 \ \mathrm{^oC}$) closely.]{
+		\label{F:state}
+		\centering
+		\includegraphics[width=23pc]{figures/state.eps}
+	}
+
 	\subfigure[Cumulative optimal cost after solving optimization. MPC serves as the benchmark with the minimum cost, followed by DPC-En and then DPC-RT.]{
@@
 		\centering
-		\includegraphics[width=16pc]{figures/cumsumcost.eps}
+		\includegraphics[width=23pc]{figures/cumsumcost.eps}
 	}
 	\end{center}
-	%\vspace{-1cm}
+	\vspace{-0.6cm}
 	\caption{Comparison of optimal performance obtained with MPC, DPC-En and DPC-RT for 3 days in January and 3 days in May.}
@@
 	\captionsetup{justification=centering}
+	\vspace{-0.3cm}
 \end{figure}

### 2018-01-20 paper/final/cs-dpc.tex
small changes

@@
 This serves as the output variable in $\Y$ of interest for which we build $N$ trees and $N$ forests as described in Section~\ref{SS:dpcrt} and \ref{SS:dpcrf}, respectively.
-Therefore, $y_{k+j}:=x_{k+j+1}^1$, where $x^1$ is the first component of $x$.
+Therefore, $y(k+j):=x(k+j+1)^1$, where $x^1$ is the first component of $x$.
 Next, we define the non-manipulated features in $\X^d$.
 At time $k$, for the tree $\mathcal{T}_j$ and the forest $\mathcal{F}_j$, we base these features to include weather disturbances, external disturbances due to occupancy and equipments, and autoregressive terms of the room temperature, i.e.
-$\X^d:=\{( d_{k+j-N} ,\ldots,d_{k+j-1}, x_{k}^1,\ldots, x_{k-\delta}^1 )\},\ \forall k=1,\ldots,n$, where $\delta$ is the order of state autoregression.
-Finally, the inputs in DPC are exactly same as in MPC, i.e. $\X^c:=u_k,\ \forall k=1,\ldots,n$.
+$\X^d:=\{( d(k+j-N) ,\ldots,d(k+j-1), x(k)^1,\ldots, x(k-\delta)^1 )\},\ \forall k=1,\ldots,n$, where $\delta$ is the order of state autoregression.
+Finally, the inputs in DPC are exactly same as in MPC, i.e. $\X^c:=\{u(k)\},\ \forall k=1,\ldots,n$.
 The training data in the above format was generated by simulating the bilinear model with rule-based strategies for 10 months in 2007.
@@
 	\begin{align}
-	& \underset{u_{k+j},\epsilon_j}{\text{minimize}} & & \sum_{j=1}^{N} (y_{k+j} - x_{ref}) \mathcal{Q}^{(1,1)} (y_{k+j} - x_{ref}) + c^T u_{k+j-1} + \lambda\epsilon_j \\
-	& \text{subject to }                  & & y_{k+j} =  \alpha_j \left[1,u_{k},\ldots,u_{k+j-1} \right]^\top \label{SE:dpc}                                            \\
-	&                                     & & \underline{u} \leq u_{k+j-1} \leq \bar{u} 									   	                                        \\
+	& \underset{u_{k+j},\epsilon_j}{\mathrm{minimize}} & & \sum_{j=0}^{N} (y_{k+j}-x_{ref})\mathcal{Q}^{(1,1)} (y_{k+j}-x_{ref})+c^\top u_{k+j}+\lambda\epsilon_j \\
+	& \mathrm{subject\ to }                              & & y_{k+j} =  \alpha_j \left[1,u^\top_{k},\ldots,u^\top_{k+j} \right]^\top \label{SE:dpc}               \\
+	&                                                    & & \underline{u} \leq u_{k+j} \leq \bar{u} 	                                                          \\
 	&                                                    & & \underline{y} - \epsilon_j \leq y_{k+j} \leq \bar{y} + \epsilon_j                                    \\
-	&                                     & & \epsilon_j \geq 0, \ j = 1,\dots,N,            									                 
+	&                                                    & & \epsilon_j \geq 0, \ j = 0,\dots,N,            									                 
 	\end{align}\label{E:dpc}
@@
 
-We compare the prediction for the first time step $y_{k+1}$ and the 6-hour ahead prediction $y_{k+6}$, given $k$, for a week in the month of May in Figure~\ref{F:validation}.
+We compare the prediction for the first time step $y(k+1)$ and the 6-hour ahead prediction $y(k+6)$, given $k$, for a week in the month of May in Figure~\ref{F:validation}.
 It is visible how trees have a high variance, and the forests are more accurate.
@@
 	\centering
-	\begin{tabular}{cccc}
+	\begin{tabular}{lccc}
 		\toprule
@@
 		\midrule
-		tree-$y_{k+1}$    & 0.42 & 0.75        & 0.76 \\
-		tree-$y_{k+6}$    & 0.64 & 0.41        & 0.42 \\
-		forest-$y_{k+1}$  & 0.29 & 0.87        & 0.88 \\
-		forest-$y_{k+6}$  & 0.38 & 0.78        & 0.80 \\
+		tree - $y(k+1)$    & 0.42 & 0.75        & 0.76 \\
+		tree - $y(k+6)$    & 0.64 & 0.41        & 0.42 \\
+		forest - $y(k+1)$  & 0.29 & 0.87        & 0.88 \\
+		forest - $y(k+6)$  & 0.38 & 0.78        & 0.80 \\
 		\bottomrule

### 2018-01-20 paper/final/cs-model.tex
small changes

@@
 The bilinear model has 12 internal states including the inside zone temperature $\mathsf{T}_{\mathrm{in}}$, the slab temperatures $\mathsf{T}_{\mathrm{sb}}$, the inner wall $\mathsf{T}_{\mathrm{iw}}$ and the outside wall temperature $\mathsf{T}_{\mathrm{ow}}$.
-The state vector is defined as $x:=[\mathsf{T}_{\mathrm{in}}, \mathsf{T}_{\mathrm{sb}}^{(1:5)}, \mathsf{T}_{\mathrm{ef}}^{(1:3)}, \mathsf{T}_{\mathrm{in}}^{(1:3)}]^T$.
+The state vector is defined as $x:=[\mathsf{T}_{\mathrm{in}}, \mathsf{T}_{\mathrm{sb}}^{(1:5)}, \mathsf{T}_{\mathrm{ef}}^{(1:3)}, \mathsf{T}_{\mathrm{in}}^{(1:3)}]^\top$.
 
-There are 4 control inputs including the blind position $\mathsf{B}$, the gains due to electric lighting $\mathsf{L}$, the evaporative cooling usage factor $\mathsf{C}$, and the heat from the radiator $\mathsf{H}$ such that $u:=[\mathsf{B},\mathsf{L},\mathsf{H},\mathsf{C}]^T$.
+There are 4 control inputs including the blind position $\mathsf{B}$, the gains due to electric lighting $\mathsf{L}$, the evaporative cooling usage factor $\mathsf{C}$, and the heat from the radiator $\mathsf{H}$ such that $u:=[\mathsf{B},\mathsf{L},\mathsf{H},\mathsf{C}]^\top$.
 $\mathsf{B}$ and $\mathsf{L}$ affect both room illuminance and temperature due to heat transfer whereas $\mathsf{C}$ and $\mathsf{H}$ affect only temperature.
@@
 Other disturbances come from the internal gains due to occupancy $\mathsf{Q}_{\mathrm{io}}$ and due to equipments $\mathsf{Q}_{\mathrm{ie}}$ which were assumed as per the Swiss standards \cite{Merkblatt2006}.
-We define $d:=[\mathsf{Q}_{\mathrm{sc}},\mathsf{Q}_{\mathrm{so}},\mathsf{I}_{\mathrm{o}},\mathsf{Q}_{\mathrm{io}},\mathsf{Q}_{\mathrm{ie}},\mathsf{T}_{\mathrm{db}},\mathsf{T}_{\mathrm{wb}}]^T$.
+We define $d:=[\mathsf{Q}_{\mathrm{sc}},\mathsf{Q}_{\mathrm{so}},\mathsf{I}_{\mathrm{o}},\mathsf{Q}_{\mathrm{io}},\mathsf{Q}_{\mathrm{ie}},\mathsf{T}_{\mathrm{db}},\mathsf{T}_{\mathrm{wb}}]^\top$.
 For further details, we refer the reader to \cite{Oldewurtel2011}.
@@
 \begin{equation}\label{E:bilinear1}
-x_{k+1} = Ax_{k}+(B_u +B_{xu}[x_k] + B_{du}[d_k]) u_k+B_dd_k,
+x(k+1) = Ax(k)+(B_u +B_{xu}[x(k)] + B_{du}[d(k)]) u(k)+B_d d(k),
 \end{equation}
@@
 	\begin{aligned}
-		&B_{xu}[x_k] = [ B_{xu,1}[x_k],B_{xu,2}[x_k], \dots, B_{xu,4}[x_k] ] \in \mathbb{R}^{12\times4}, \\
-		&B_{du}[d_k] = [ B_{du,1}[d_k],B_{du,2}[d_k], \dots, B_{du,4}[d_k] ] \in \mathbb{R}^{12\times4}, \\
+		&B_{xu}[x(k)] = [ B_{xu,1}x(k),B_{xu,2}x(k), B_{xu,3}x(k), B_{xu,4}x(k) ] \in \mathbb{R}^{12\times4}, \\
+		&B_{du}[d(k)] = [ B_{du,1}d(k),B_{du,2}d(k), B_{xu,3}x(k), B_{du,4}d(k) ] \in \mathbb{R}^{12\times4}, \\
 	\end{aligned}

### 2018-01-20 paper/final/cs-mpc.tex
small changes

@@
 
-The objective of the controller is to minimize the energy usage $c^Tu$ while maintaining a desired level of thermal comfort $x_{ref}$ (or $\tT_\mathrm{ref}$).
-Therefore, at time step $k$, we solve a continuously linearized MPC problem to determine the optimal sequence of inputs $[u_{\mathrm{k|k}},\dots,u_{\mathrm{k+N-1|k}}]$:
+The objective of the controller is to minimize the energy usage $c^\top u$ while maintaining a desired level of thermal comfort $x_{ref}$ (or $\tT_\mathrm{ref}$).
+Therefore, at time step $k$, we solve a continuously linearized MPC problem to determine the optimal sequence of inputs $u^*_{k},\dots,u^*_{k+N-1}$:
 
@@
 	\begin{align}
-		& \underset{u_{k+j},\epsilon_j}{\text{minimize}} & & \sum_{j=1}^{N} ({x}_{k+j} - x_{ref})\top Q ({x}_{k+j} - x_{ref}) + c^Tu_{k+j-1} +  \lambda\epsilon_j \\
-		& \text{subject to }                             & & x_{k+j} = Ax_{k+j-1} + B u_{k+j-1} + B_d d_{k+j-1} \label{SE:mpc1}                                   \\
+		& \underset{u_{k+j-1},\epsilon_j}{\text{minimize}} & & \sum_{j=1}^{N+1} ({x}_{k+j} - x_{ref})^\top Q ({x}_{k+j} - x_{ref}) + c^\top u_{k+j-1} +  \lambda\epsilon_j                                                                                                                          \\
+		& \text{subject to }                               & & x_{k+j} = Ax_{k+j-1} + B u_{k+j-1} + B_d d_{k+j-1}\label{SE:mpc1}                   \\
 		&                                                  & & B       = B_u + B_{xu}[x_{k}] + B_{du}[d_{k+j-1}] \label{SE:mpc2}                   \\
@@
 		&                                                  & & \underline{x} - \epsilon_j \leq x_{k+j} \leq \bar{x} + \epsilon_j                   \\
-		&                                                & & \epsilon_j \geq 0, \ j = 1,\dots,N,            									                 
+		&                                                  & & x_k = x(k)                                                                          \\
+		&                                                  & & \epsilon_j \geq 0, \ j = 1,\dots,N+1,            									                 
 	\end{align}\label{E:mpc}

### 2018-01-20 paper/final/dpc.tex
small changes

@@
 \begin{equation}\label{E:model_tree}
-\T_j = \mathit{f}_{\mathrm{tree}} \left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right),\ j=1,\ldots,N,
+\T_j = \mathit{f}_{\mathrm{tree}} \left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right),\ j=0,\ldots,N,
 \end{equation}
 with $\left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right)\in\X^d,\ k=1,\ldots,n$, and $\mathit{f}_{\mathrm{tree}}$ representing the tree structure.
-Then, the linear models, as functions of the inputs variables $u\in\X^c$, in each leaf of the trees $\T_j,\ j=1,\ldots,N$, are defined as
+Then, the linear models, as functions of the inputs variables $u\in\X^c$, in each leaf of the trees $\T_j,\ j=0,\ldots,N$, are defined as
 \begin{equation}\label{E:model_leaf}
-y(k+j) =  \beta_j [1,u(k),\ldots,u(k+j) ]^T,\ j=1,\ldots,N.
+y(k+j) =  \beta_j [1,u(k),\ldots,u(k+j) ]^T,\ j=0,\ldots,N.
 \end{equation}}
@@
 \textcolor[rgb]{0,0,1}{The problem now is: \emph{how to use this modeling framework to set up an MPC problem?} In run-time, given the disturbance forecast and the state measurements at time $k$, i.e.
-$\tilde d(k+j-\delta_d),\ldots,\tilde d(k+j),x(k-\delta_x),\ldots,x(k)$ given $k$, we can narrow down to a leaf of each tree in \eqref{E:model_tree} to retrieve the linear models in \eqref{E:model_leaf} for each step $j=1,\ldots,N$ of the horizon (see first part of the pseudo code given in Algorithm \ref{A:dpcrt}), as depicted in Figure \ref{F:dpc-sepvars} (left).
+$\tilde d(k+j-\delta_d),\ldots,\tilde d(k+j),x(k-\delta_x),\ldots,x(k)$ given $k$, we can narrow down to a leaf of each tree in \eqref{E:model_tree} to retrieve the linear models in \eqref{E:model_leaf} for each step $j=0,\ldots,N$ of the horizon (see first part of the pseudo code given in Algorithm \ref{A:dpcrt}), as depicted in Figure \ref{F:dpc-sepvars} (left).
 These models are used to setup the MPC Problem \ref{P:dpcrt}.
@@
 $y(k)\in\mathbb{R}^{n_y},\ n_y\geq 1$, can be formulated as:
-%\begin{problem}\label{P:dpcrt}
-%	\begin{equation}
-%		\begin{aligned}
-%		\text{min } & \sum_{j=1}^{N} ({\Y}_{\mathrm{k+j|k}})^2 \mathcal{Q} + {\X^c}^T_{\mathrm{k+j-1|k}} \mathcal{R} {\X^c}_{\mathrm{k+j-1|k}} + \lambda\epsilon_j\\
-%		\text{s.~t. } & \ \ \ \ \ \Y_{\mathrm{k+j|k}} =  \beta^T [1,\X^c_{\mathrm{k|k}},\dots,\X^c_{\mathrm{k+j-1|k}} ]^T \\
-%		& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{\X}^c \leq \X^c_{\mathrm{k+j-1|k}} \leq \bar{\X}^c\\ 
-%		& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\Y}-\epsilon_j \leq \Y_{\mathrm{k+j|k}} \leq \bar{\Y} + \epsilon_j\\\
-%		& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0, \ j = 1,\dots,N.
-%		\end{aligned}
-%		\label{E:dpcrt}
-%	\end{equation}
-%\end{problem}
+
 \begin{problem}\label{P:dpcrt}
@@
 	\begin{aligned}
-	& \underset{u_{k+j}}{\text{minimize}} & & \sum_{j=1}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
-	& \text{subject to }                  & & y_{k+j}     =   \beta_j [1,u_{k},\ldots,u_{k+j} ]^\top                             \\
+	& \underset{u_{k+j}}{\mathrm{minimize}} & & \sum_{j=0}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
+	& \mathrm{subject\ to }                 & & y_{k+j}     =   \beta_j [1,u^\top_{k},\ldots,u^\top_{k+j} ]^\top                   \\
 	&                                       & & u_{k+j}    \in  \mathcal{U}                                                        \\
@@
 	&                                       & & \epsilon_j \geq  0							                                       \\
-	&                                     & & j           =    1,\ldots,N.            									         \\
+	&                                       & & j           =    0,\ldots,N.            									       \\
 	\end{aligned}
@@
 \begin{equation}\label{E:model_forest}
-\F_j = \mathit{f}_{\mathrm{forest}} \left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right),\ j=1,\ldots,N,
+\F_j = \mathit{f}_{\mathrm{forest}} \left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right),\ j=0,\ldots,N,
 \end{equation}
@@
 In particular, as mentioned above, each tree $\T_i$ of each forest $\F_j$ is trained using $(\bar \X^d_i,\bar\Y_i)$.
-Then the linear models, as functions of the inputs variables $u\in\bar\X^c_i\subset\X_c$, in each leaf of each tree of the forests $\F_j,\ j=1,\ldots,N$, are defined as
+Then the linear models, as functions of the inputs variables $u\in\bar\X^c_i\subset\X_c$, in each leaf of each tree of the forests $\F_j,\ j=0,\ldots,N$, are defined as
 \begin{equation}\label{E:model_leaf_forest}
-y(k+j) =  \Theta_{ij} [1,u(k),\ldots,u(k+j) ]^T,\ j=1,\ldots,N.
+y(k+j) =  \Theta_{ij} [1,u(k),\ldots,u(k+j) ]^T,\ j=0,\ldots,N.
 \end{equation}
@@
 	\begin{aligned}
-	& \underset{u_{k+j}}{\text{minimize}} & & \sum_{j=1}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
-	& \text{subject to }                  & & y_{k+j}      =  \hat{\Theta}_j [1,u_{k},\ldots,u_{k+j} ]^\top                      \\
+	& \underset{u_{k+j}}{\mathrm{minimize}} & & \sum_{j=0}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
+	& \mathrm{subject\ to }                 & & y_{k+j}      =  \hat{\Theta}_j [1,u_{k},\ldots,u_{k+j} ]^\top                      \\
 	&                                       & & u_{k+j}    \in  \mathcal{U}                                                        \\
@@
 	&                                       & & \epsilon_j \geq  0							                                     \\
-	&                                     & & j           =    1,\ldots,N.            									         \\
+	&                                       & & j           =    0,\ldots,N.            									         \\
 	\end{aligned}

### 2018-01-20 paper/final/dpc.tex
small changes on DPC

@@
 \textcolor[rgb]{0,0,1}{The central idea behind DPC is to obtain control-oriented models using machine learning methodologies, and formulate the control problem in a way that Receding Horizon Control (RHC) can still be applied and the optimization problem can be solved efficiently.
-In this paper Regression Trees and Random Forests are considered (see Appendix ???
-for details).}
+In this paper, Regression Trees and Random Forests are considered (see Appendix ??? for details).}
 
-\textcolor[rgb]{0,0,1}{To this aim, let an historical dataset $(\X,\Y)$, created using building measurements, be given to be used in the Regression Trees and Random Forests training process.
+\textcolor[rgb]{0,0,1}{To this aim, let an historical dataset $(\X,\Y)$, created using building measurements, be given to be used in the Regression Trees and Random Forests learning process.
 Let $|(\X,\Y)| = n$, be the number of samples in the dataset.
-$\X = \{(x(k),u(k),d(k))\}$, $\forall k = 1,\ldots,n$, is the set of samples $(x(k),u(k),d(k))$, measured at each time instant $k$, where $x(k)\in\mathbb{R}^{n_x}$ is the vector of the state variables, $u(k)\in\mathbb{R}^{n_u}$ is the vector of the input variables and $d(k)\in\mathbb{R}^{n_d}$ is the vector of the disturbance variables.
-$\Y = \{y(k)\},\ \forall k = 1,\ldots,n$, is the set of samples $y(k)$, measured at each time instant $k$, where $y(k)\in\mathbb{R}^{n_y}$ is the vector of the output variables.
-We consider $\X$ as the set of predictor variables (or features) and $\Y$ as the set of response variables.
-As an example, the response variables can be represented by the power consumption or the room temperatures evolution, while the predictors can be represented by the weather forecast, as the disturbance, the set-points information and building schedules, as the inputs, and room temperatures, as the states.
-Our goal is to learn a data-driven models, using Regression Trees and Random Forests, that relate the value of the response variables with the value of the predictor variables, and that can be used to set up an MPC problem.
+The set of predictor variables (or features) $\X = \{(x(k),u(k),d(k))\}$, $\forall k = 1,\ldots,n$, is the set of samples $(x(k),u(k),d(k))$, measured at each time instant $k$, where $x(k)\in\mathbb{R}^{n_x}$ is the vector of the \emph{measurable} state variables, $u(k)\in\mathbb{R}^{n_u}$ is the vector of the input variables and $d(k)\in\mathbb{R}^{n_d}$ is the vector of the \emph{measurable and predictable} disturbance variables.
+We recall that, as explained in Section \ref{secModelingIssuesExample}, here we do not use non-measurable states, such as temperature of wall layers, ceiling, floor, etc., since the data-driven methodology is able to compensate them.
+As an example, the variables in $\X$ can be represented by the weather forecast, as the disturbances, the set-points information and building schedules, as the inputs, and the room temperatures and power consumption, as the states.
+The set of response variables $\Y = \{y(k)\},\ \forall k = 1,\ldots,n$, is the set of samples $y(k)\in\mathbb{R}^{n_y}$.
+The response variables are represented by the state variables at the next time step, i.e. $y(k) = x(k+1)$.}
+
+\textcolor[rgb]{0,0,1}{Our goal is to learn data-driven models, using Regression Trees and Random Forests, that relate the value of the response variables with the value of the predictor variables, and that can be used to set up an MPC problem.
 }
@@
 where $f$ is an analytical expression that describes the system's dynamics.
-\textbf{We want to remark that $y$ can be considered as both a state variable evolution, i.e.
-$y(k) = x(k+1)$, and a variable that is function of $x$, $u$ and $d$ but is not one of them, as for example is the case of the power.
-We consider this aspect in Sections \ref{S:proof}, \ref{S:casestudy} and \ref{S:realCaseStudy}.
-IMPROVE!!!!!!!!!!!!!!!}.
 The issue that arises, when using Regression Trees and Random Forests, is that $f$ does not have a closed-form expression, hence \eqref{E:GenericModel} is not suitable for control and optimization.
 One possibility could be to look for sub-optimal solutions using evolutionary algorithms \cite{Kusiak2009}, i.e. using heuristics.
-In this section we provide a new methodology to determine a closed-form expression for $f$ by manipulating the dataset, as shown starting from Section \ref{SS:sepvar}.}
+In this section we provide a new methodology to determine a closed-form expression for $f$, by manipulating the dataset as shown starting from Section \ref{SS:sepvar}.}
 
@@
 \textcolor[rgb]{0,0,1}{The dataset splitting consists of partitioning the features set $\X$ into the sets $\X^c = \{u(k)\}\subset\X,\ \forall k=1,\ldots,n$, containing the control (or manipulated) variables, and $\X^d = \{(x(k),d(k))\}\subset\X,\ \forall k=1,\ldots,n$, containing the disturbance and state (or non-manipulated) variables.
-The union of the two sets forms the full feature set oftraining, i.e. $\X \equiv \X^c \cup \X^d$.
+The union of the two sets forms the full feature set of training, i.e. $\X \equiv \X^c \cup \X^d$.
 %Our goal is to replace a model-based controller with a data-driven controller, where the latter depends only on the historical sensor data. 
 %These measurements could directly represent one or more states in the model-based control framework.
-We denote these as outputs $\Y \in \R$ for training, i.e.
-a $\Y$ represents a particular output and we can have separate models for multiple outputs.
+We denote these as responses $\Y \in \R$ for training, i.e.
+a $\Y$ represents a particular response and we can have separate models for multiple responses.
 We define the number of training samples by $|(\X,\Y)| = n$.
@@
 		\item the trees and the forests are trained only using $\X^d$, which also eases the computational complexity.
-			  It is important to note that besides external disturbance and the state measurements, $\X^d$ can also contains autoregressive terms of the output variables in $\Y$;
-		\item in each leaf of the tree (or in each leaf of each the tree of the forest), a linear regression model is trained as a function of the inputs variables in $\X^c$.
-		As we shall see in Section~\ref{SS:dpcrt} and \ref{SS:dpcrf}, the second step reduces the run-time control problem into a convex program.
+			  It is important to note that besides external disturbance and the state measurements, $\X^d$ can also contains autoregressive terms of such variables;
+		\item in each leaf of the tree (or in each leaf of each tree of the forest), a linear regression model is trained as a function of the inputs variables in $\X^c$.
+		As we shall see in Section~\ref{SS:dpcrt} and \ref{SS:dpcrf}, this second step reduces the run-time control problem into a convex program.
 	\end{enumerate}
@@
 
-\textcolor[rgb]{0,0,1}{Now, in order to have a model that can be used for prediction over an arbitrary long horizon in an MPC problem, our goal is to predict, at time $k$, the output $y$ for next $N$ time steps, i.e.
-$y(k),\ldots,y(k+N)$, where $N$ is the control horizon, given the state measurements, i.e.
-$x(k)$, and the disturbance forecast, i.e.
-$\tilde d(k),\ldots,\tilde d(k+N)$, at time $k$.
-For the sake of simplicity and without any loss of generality, in the following, we consider only a single output, i.e. $y(k)\in\mathbb{R}$ (or $n_y = 1$).
-However, as we will see in Sections \ref{S:casestudy} and \ref{S:realCaseStudy}, multiple trees (or forests) can be built for multiple outputs.
-Applying the dataset splitting, we build $N$ regression trees using CART procedure \cite{Breiman1984} (see also Appendix for more details), such that the output $y(k+j)$ of the $j^{th}$ tree depends upon the previous $\delta_d$ disturbances, and on the state at instant $k$ and its $\delta_x$ regressive terms:
-\begin{equation}\label{E:model_tree}
+\textcolor[rgb]{0,0,1}{Now, in order to have a model that can be used for prediction in an MPC problem, with an horizon of arbitrary length $N$, our goal is to predict, at time $k$, the response $y$ for the next $N$ time steps, i.e.
+$y(k),\ldots,y(k+N)$, given the state measurements and the disturbance forecast at time $k$, i.e.
+$x(k)$ and $\tilde d(k),\ldots,\tilde d(k+N)$.
+For the sake of simplicity and without any loss of generality, in the following, we consider only a scalar response variable, i.e. $y(k)\in\mathbb{R}$ (or $n_y = 1$).
+However, as we will see in Sections \ref{S:casestudy} and \ref{S:realCaseStudy}, multiple trees (or forests) can be built for $n_y > 1$.}
+
+\textcolor[rgb]{0,0,1}{Applying the dataset splitting, we build $N$ regression trees such that the response $y(k+j)$ of the $j^{th}$ tree depends upon $\delta_d$ autoregressive terms of the disturbances, and on the state at instant $k$ and its $\delta_x$ autoregressive terms:}
+
+\textcolor[rgb]{0,0,1}{\begin{equation}\label{E:model_tree}
 \T_j = \mathit{f}_{\mathrm{tree}} \left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right),\ j=0,\ldots,N,
-\end{equation}
-with $\left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right)\in\X^d,\ k=1,\ldots,n$, and $\mathit{f}_{\mathrm{tree}}$ representing the tree structure.
-Then, the linear models, as functions of the inputs variables $u\in\X^c$, in each leaf of the trees $\T_j,\ j=0,\ldots,N$, are defined as
-\begin{equation}\label{E:model_leaf}
+\end{equation}}
+
+\noindent \textcolor[rgb]{0,0,1}{with $\left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right)\in\X^d,\ k=1,\ldots,n$, and $\mathit{f}_{\mathrm{tree}}$ representing the tree structure.
+Then, the linear models, as functions of the inputs variables $u\in\X^c$, in each leaf of the trees $\T_j,\ j=0,\ldots,N$, are defined as}
+
+\textcolor[rgb]{0,0,1}{\begin{equation}\label{E:model_leaf}
 y(k+j) =  \beta_j [1,u(k),\ldots,u(k+j) ]^T,\ j=0,\ldots,N.
 \end{equation}}
-Note that the coefficients $\beta_j$ are different for each leaf.
-Equation~\eqref{E:model_leaf} implies that the prediction of output $y(k+j)$ at time $k$ is an affine combination of control inputs from time $k$ to $k+j$.
+
+\noindent \textcolor[rgb]{0,0,1}{Note that the coefficients $\beta_j$ are different for each leaf.
+Equation~\eqref{E:model_leaf} implies that the prediction of response $y(k+j)$ at time $k$ is an affine combination of control inputs from time $k$ to $k+j$.
 Thus, we have managed to linearize the original model dynamics via black-box modeling.
-\textcolor[rgb]{0,0,1}{One of the advantages of this approach is that this two-step training is done off-line, so the time required to create the models does not affect the control execution in run-time.}
+One of the advantages of this approach is that this two-step training is done off-line, so the time required to create the models does not affect the control execution in run-time.}
 
@@
 $\tilde d(k+j-\delta_d),\ldots,\tilde d(k+j),x(k-\delta_x),\ldots,x(k)$ given $k$, we can narrow down to a leaf of each tree in \eqref{E:model_tree} to retrieve the linear models in \eqref{E:model_leaf} for each step $j=0,\ldots,N$ of the horizon (see first part of the pseudo code given in Algorithm \ref{A:dpcrt}), as depicted in Figure \ref{F:dpc-sepvars} (left).
-These models are used to setup the MPC Problem \ref{P:dpcrt}.
-In particular, }when a new control action has to be determined, each tree (prediction step) contributes to a linear constraint in the optimization as a replacement for the state dynamics in the case of MPC.
-Thus, the MPC optimization problem, considering a quadratic cost ($Q \succeq 0, R \succeq 0$) function and the general case of multiple outputs, i.e.
+These models are used to setup the MPC Problem \ref{P:dpcrt}.}
+
+\textcolor[rgb]{0,0,1}{In particular, }when a new control action has to be determined, each tree (prediction step) contributes to a linear constraint in the optimization as a replacement for the state dynamics in the case of MPC.
+Thus, the MPC optimization problem, considering a quadratic cost ($Q \succeq 0, R \succeq 0$) function and the general case of multiple responses, i.e.
 $y(k)\in\mathbb{R}^{n_y},\ n_y\geq 1$, can be formulated as:
@@
 This problem is solved as in the classical MPC formulation, i.e.
-at each time step $k=1,2,\ldots$ the optimal control sequence $u^*_k,\ldots,u^*_{k+N}$ is computed, and only the first input of the sequence is applied as control input to the system: $u(k) = u^*_k$.
+at each time step $k=1,2,\ldots$ the optimal control sequence $u^*_k,\ldots,u^*_{k+N}$ is computed, and only the first input of the sequence is applied to the system: $u(k) = u^*_k$.
 The slack variables $\epsilon_j$ are added to ensure recursive feasibility (since the equality constraint on $y$ is relaxed), i.e.
 to guarantee that Problem \ref{P:dpcrt} can provide a solution at each step $k=1,2,\ldots$.
-$\lambda$ is a weight used to give more importance to bounds violation with respect to $y$, than to the cost function optimization, and viceversa. 
+$\lambda$ is a weight used to give more importance to bounds violation, than to the quadratic term, and viceversa. 
 Of course, a different cost function can be chosen depending upon the application, i.e. it can be linear, nonlinear, etc., obviously changing the complexity of the problem.
@@
 	\includegraphics[width=26pc]{figures/dpc-algo-rf.eps}
-	\caption{\textcolor[rgb]{0,0,1}{DPC-En: At time $k$, the algorithm uses the forecast of disturbances and the state measurements in $\X^d$ at time $k$ to select linear models $\Theta_1$ to $\Theta_t$ in the leaves of each forest.
-The linear models in each forest are averaged to calculate a single model represented by $\hat{\Theta}_j$, and act as constraints in the optimization problem.
+	\caption{\textcolor[rgb]{0,0,1}{DPC-En: At time $k$, the algorithm uses the forecast of disturbances and the state measurements in $\X^d$ given at time $k$ (that with a small abuse of notation we indicate only in this figure with $\X^d_k$), to select linear models $\Theta_{l1}$ to $\Theta_{lt}$ in each leaf $l$ of the $t$ trees of each forest.
+The linear models in each forest are averaged to calculate single models represented by $\hat{\Theta}_j$, and act as constraints in the optimization problem.
 The optimal sequence $u^*_k,\ldots,u^*_{k+N}$, of which only the first element is applied, i.e.
-$u(k)=u^*_k$, and variables in $\X^d$ at time $k+1$ are calculated to proceed to next step at $k+1$.}}
+$u(k)=u^*_k$, and variables in $\X^d_{k+1}$ are calculated to proceed to next step at $k+1$ and repeat the process.}}
 	\label{F:dpc-algo-rf}

### 2018-01-20 paper/final/intro.tex
intro edits

@@
 In Section \ref{secModelingIssuesExample} we will present a detailed technical example to better illustrate how data-driven approaches can address the above issues and thus reduce the cost of modeling buildings.
-Indeed in practice, due to aforementioned reasons, the control strategies in such systems are often limited to fixed, sometimes ad-hoc, rules that are based on best practices.
+In practice, due to aforementioned reasons, the control strategies in such systems are often limited to fixed, sometimes ad-hoc, rules that are based on best practices.
 The alternative is to use black-box, or completely data-driven modeling approaches, to obtain a realization of the system's input-output behavior. 
@@
 In particular, in this paper, we present a method based on Random Forests which uses historical data for receding horizon control.
-Before illustrating our approach, we will discuss in the next section the related literature and the novelty of our approach.}
+Before illustrating our approach, in the next section, we discuss the related literature and the novelty of our approach.}
 
@@
 	\end{tabular}\normalsize}
-	\caption{\textcolor[rgb]{0,0,1}{References ordered considering: case study they have been applied to; whether they used experimental data, other than simulated data, and if they did real implementation (RI), i.e.
-implemented the methodologies on real systems; if they used simulative tool; the type of the model considered, i.e.
-Model-Based or Data-Driven; if they used the model to do control.}}
+	\caption{\textcolor[rgb]{0,0,1}{References ordered considering: case study they are applied to; whether they use experimental data, other than simulated data, and if they do real implementation (RI), i.e.
+implement the methodologies on real systems; if they use simulative tool; the type of the model considered, i.e.
+Model-Based or Data-Driven; if the models are used for control.}}
 	\captionsetup{justification=centering}
@@
 		\begin{itemize}
-			\item model-based, that we consider both "white-box" and "grey-box" approaches: \cite{Shakouri2017SCS,Li2014EB,Yoon2014EB,Li2016AE,Harb2016EB,Salakij2016EB,Li2016E,Li2016EB,Hou2013,Cecconi2017EB};
-			\item data-based, that is the "black-box" approach, mainly done using Neural Networks: \cite{Safa2017SCS,Neto2008EB,Magnier2010BE,Candanedo2017EB,Ascione2017E,Cecconi2017EB,Li2016AE};
+			\item model-based, such as ``white-box" and ``grey-box" approaches: \cite{Shakouri2017SCS,Li2014EB,Yoon2014EB,Li2016AE,Harb2016EB,Salakij2016EB,Li2016E,Li2016EB,Hou2013,Cecconi2017EB};
+			\item data-based, i.e.~``black-box" approach, mainly done using Neural Networks: \cite{Safa2017SCS,Neto2008EB,Magnier2010BE,Candanedo2017EB,Ascione2017E,Cecconi2017EB,Li2016AE};
 			\item simulation tool-based, such as EnergyPlus \cite{energyPlus} and TRNSYS \cite{trnsys2000}: \cite{Yin2016EB,Christantoni2016EB};
@@
 		\begin{itemize}
-			\item only model creation and identification: \cite{Safa2017SCS,Neto2008EB,Magnier2010BE,Li2014EB,Li2016AE,Harb2016EB,Li2016EB,Candanedo2017EB,Cecconi2017EB,Ascione2017E};
-			\item model identification and control (typically Predictive Control): \cite{Shakouri2017SCS,Yoon2014EB,Yin2016EB,Salakij2016EB,Li2016E,Hu2017AE}.
+			\item only model identification: \cite{Safa2017SCS,Neto2008EB,Magnier2010BE,Li2014EB,Li2016AE,Harb2016EB,Li2016EB,Candanedo2017EB,Cecconi2017EB,Ascione2017E};
+			\item model identification and control, mainly Predictive Control: \cite{Shakouri2017SCS,Yoon2014EB,Yin2016EB,Salakij2016EB,Li2016E,Hu2017AE}.
 		\end{itemize}
@@
 \textcolor[rgb]{0,0,1}{
-These papers are summarizes in Table \ref{T:RelatedWork} highlighting the differences.
+These references are summarized in Table \ref{T:RelatedWork} highlighting the key differences.
 We also emphasize the case studies the results are applied to, and whether the authors use experimental data to simulate their algorithms.
-Only in three cases the algorithms are tested on real systems (see Table: RI - Real Implementation).
+Only in three cases the algorithms are tested on real systems (see Table \ref{T:RelatedWork}: RI - Real Implementation).
 We observe that except for the last 6 cases, which we discuss in detail, either model-based approaches or only tools are considered with/without control or data-driven modeling approaches without control.
@@
 The last six papers of Table \ref{T:RelatedWork} are more related to the methodology presented in this paper.
-In particular, in \cite{Macarulla2017} the authors proposed a predictive control strategy based on Neural Networks, for boilers control in buildings, to decide the optimal time to switch-on the plan to guarantee energy saving and thermal comfort.
+In particular, the authors in \cite{Macarulla2017} proposed a predictive control strategy based on Neural Networks, for boilers control in buildings, to decide the optimal time to switch-on the plan to guarantee energy savings and thermal comfort.
 However, the approach is not easily scalable to different types of plants and does not use optimization in the closed-loop scheme.
-In \cite{Costanzo2016} a reinforcement learning control strategy, called Model-Assisted Batch Reinforcement Learning, is considered to provide data-driven control for the demand response problem in HVAC systems.
+In \cite{Costanzo2016}, an approach based on reinforcement learning, called Model-Assisted Batch Reinforcement Learning, is considered to provide data-driven control for the demand response problem in HVAC systems.
 Reinforcement Learning is a model-free methodology and is an alternative approach to MPC \cite{Ernst2009TSMC} (with pros and cons).
-In \cite{Ferreira2012} the authors considered a data-driven predictive control based on Neural Networks to guarantee energy saving and  thermal comfort in public buildings.
+The authors in \cite{Ferreira2012}considered a data-driven predictive control based on Neural Networks to guarantee energy saving and  thermal comfort in public buildings.
 Neural Networks are used in the closed-loop control scheme to determine a thermal comfort index based on parameters that can be measured or estimated, but no Neural Networks-based system state dynamics are included into the optimization problem.
@@
 Unfortunately, since Neural Networks models are nonlinear, the MPC based on such models is also nonlinear.
-This means that a global optimal solution can not be guaranteed and, when the complexity of network is high, that solving the optimization problem becomes computationally harder due to nonlinearties.
+This means that a global optimal solution cannot be guaranteed and solving the optimization problem becomes computationally harder due to nonlinearties when the complexity of network is high.
 \\
-To overcome this complexity we initially introduced the regression trees-based approach in \cite{Behl2016,Jain2017TCPS}.
-In particular, in \cite{Behl2016} we used regression trees to setup an MPC problem and apply it to the problem of Demand-Response.
-However, these models can be used for optimal control with only one-step lookahead prediction.Hence, it is not possible to use such approach to control the system considering a prediction over an horizon of arbitrary length.
+To overcome this complexity, we initially introduced the regression trees-based approach in \cite{Behl2016,Jain2017TCPS}. 
+In particular, in \cite{Behl2016}, we used regression trees to setup an MPC problem and apply it to the problem of Demand-Response. 
+However, these models can be used for optimal control with only one-step lookahead prediction.
+Hence, it is not possible to use such approach to control the system considering a prediction over an horizon of arbitrary length. 
 This limitation was addressed in \cite{Jain2017TCPS}, where multi-output regression trees are used instead of single output trees. 
-The different outputs correspond to different steps of the prediction horizon, which allows us to setup an MPC problem with a finite horizon.
+The different outputs correspond to different steps of the prediction horizon which allows us to setup an MPC problem with a finite horizon. 
 Modeling accuracy using single trees is strongly affected by overfitting and high variance. 
 Such approach has the advantage to be extremely simple from the complexity point of view, but the range of applications is limited. 
-In \cite{JainACC2017,JainCDC2017} we took a different approach that improves the system's identification accuracy: instead of considering a single tree/forest with multiple output, we considered multiple trees/forests with single output.
+In \cite{JainACC2017,JainCDC2017} we took a different approach that improves the system's identification accuracy -- instead of considering a single tree/forest with multiple output, we considered multiple trees/forests with single output.
 Each tree/forest provides the prediction of the system's behavior for a different time step of the horizon. 
-However, the results were based on simulated data and we did not account for inaccuracies in the weather forecast.\\
-In this paper we provide a new methodology based on random forests that overcomes the drawbacks of all our previous work, and more precisely: we obtain better performance when compared to other approaches (both optimal and rule-based), we provide and validate robustness with respect to disturbance uncertainties and, last but not least, we test our method using a real data set.
-In the next section, we will illustrate in detail such contributions.
+However, the results are based on simulated data and we did not account for inaccuracies in the weather forecast.
 }
@@
 \textcolor[rgb]{0,0,1}{
-\subsection{Main contribution and paper organization}
+\subsection{Main contribution}
 %TO CLARIFY $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
@@
 In this paper, we extend these results providing the following contributions.
-The presentation of the paper contributions is organised as follows:
+In this paper, we provide a new methodology based on random forests that overcomes the drawbacks of all our previous work, and more precisely: we obtain better performance when compared to other approaches (both optimal and rule-based), we provide and validate robustness with respect to disturbance uncertainties
+Last but not the least, we test our method using a real data set. 
+More precisely, the paper makes the following contributions:
 \begin{enumerate}
-	\item In Section \ref{S:dpc} we formally describe the data-predictive control (DPC) techniques.
-For simplicity of presentation we first illustrate DPC based regression trees, then we introduce DPC based on random forests.
-	\item In Section \ref{S:proof} we demonstrate that DPC can provide comparable performance compared to MPC applied to a perfectly known mathematical model.
-More precisely, we first consider a benchmark MPC controller using a bilinear building model whose parameters were identified using experiments on a building in Switzerland: then we apply DPC to a data-set simulated from such bilinear model and extract a controller, which is based only on such data and, of course, ignores the mathematical model.
+	\item In Section \ref{S:dpc}, we formally describe our data-driven Model Predictive Control approach i.e.~Data Predictive Control (DPC).
+For clarity, we first describe DPC based on regression trees and then we introduce DPC based on random forests.
+	\item In Section \ref{S:proof}, we demonstrate that DPC can provide comparable performance compared to MPC (using a physical model) applied to a perfectly known mathematical model.
+More precisely, we first consider a benchmark MPC controller using a bilinear building model whose parameters were identified using experiments on a building in Switzerland: then we apply DPC to a data-set simulated from such bilinear model and derive a controller based only on the data.
 We show that DPC captures 70\% variance in MPC and offers a comparable performance.
-	\item In Section \ref{S:casestudy} we demonstrate scalability and efficiency of DPC describing a practical application of DPC to a Demand Response problem based in a simulative benchmark.
-More precisely, we apply DPC to a 6 story 22 zone building model in EnergyPlus \cite{Crawley2001} for which model-based control is not economical and practical due to extreme complexity.
-We observe that DPC provides the desired power reduction with an average error of 3\%: this shows scalability and efficiency of DPC in providing financial incentives to the end-customers bypassing the need for high fidelity models.
-	\item In Section \ref{S:realCaseStudy} we implement and test DPC basing on real data of an off-grid house located in L'Aquila (Italy).
-More precisely, we use DPC to design the optimal ON/OFF scheduling for the heating system in order to save energy while guaranteeing thermal comfort for the occupants.
-We quantify the total amount of energy saved, with respect to the classical bang-bang controller widely used in houses for temperature control, using an EnergyPlus model built specifically for the house.
-We show that we can perform an energy saving that goes from $25.4\%$ (if we guarantee thermal comfort by strictly respecting the desired temperature ranges in the rooms) to $49.2\%$ (if we allow small violation of the desired temperature ranges in the rooms).
-In addition, we test the robustness of our method with respect to substantial prediction errors in the weather forecast.
+	\item In Section \ref{S:casestudy}, we demonstrate scalability and efficiency of DPC in a Demand Response problem -- DPC is used to provide financial incentives to the end-customers bypassing the need for expensive high fidelity models.
+We apply DPC to a 6 story 22 zone building model in EnergyPlus for which model-based control is not economical and practical due to extreme complexity.
+We observe that DPC provides the desired power curtailment with an average error of 3\%.
+	\item In Section \ref{S:realCaseStudy}, we implement and test DPC basing on real data from an off-grid house located in L'Aquila, Italy.
+We design the optimal ON/OFF scheduling for the heating system in order to save energy while guaranteeing thermal comfort for the occupants.
+We compare the total amount of energy saved with respect to the classical bang-bang controller (widely used in houses for temperature control) using an EnergyPlus model built specifically for the house.
+We show that we can perform an energy saving that ranges from $25.4\%$ (if we guarantee thermal comfort i.e.~strictly respect the desired temperature range in the rooms) to $49.2\%$ (if we allow small violation in the desired temperature).
+Further, we test the robustness of our method due to uncertainties in the weather forecast.
 \end{enumerate}
@@
 \section{COMPLEXITY ISSUES IN MODELING BUILDINGS: PHYSICS-BASED VS DATA-DRIVEN}\label{secModelingIssuesExample}
-\textcolor[rgb]{0,0,1}{In this section we provide a detailed example to emphasise differences in terms of complexity when modeling a building using physical laws vs.
-machine learning.
-We show that the data-driven modeling eliminates several drawbacks that occur using physics-based models such as the need to have very good knowledge of the building structure and the material properties, time required to build a model and limited availability of sensors.
+\textcolor[rgb]{0,0,1}{In this section we provide a detailed example to emphasize differences in terms of complexity when modeling a building using physical laws versus machine learning.
+We show that the data-driven modeling eliminates several drawbacks that occur using physics-based models such as the need to have good knowledge of the building structure and the material properties, time required to build a model and limited availability of sensors.
 The example is based on a real building taken from \cite{Sturzenegger2016}, which is located in Allschwil (Switzerland).

### 2018-01-20 paper/final/appendixRegTrees.tex
Appendix Reg Trees

\section{Regression Trees and Random Forests}
\label{A:RegTrees}

In this appendix we first start explaining how Regression Trees are built using an example adapted from~\cite{hastie2009elements}, and then we introduce Random Forests.

\subsection{Regression Trees}
Tree-based methods partition the feature space into a set of rectangles (more formally, hyper-rectangles) and then fit a simple model in each one. They are conceptually simple yet powerful.
Let us consider a regression problem with continuous response $\Y = \{Y\}$ and and 2 predictors $\X = \{X_1,X_2\}$, each taking values in the unit interval. The top left plot of Figure~\ref{fig:friedman} shows a partition of the feature space by lines that are parallel to the coordinate axes. In each partition element, we can model $Y$ with a different constant.
\begin{figure}
  \centering
  \includegraphics[width=0.7\columnwidth]{figures/friedman}
  \vspace{-8pt}
  \caption{Top right: 2D feature space by recursive binary splitting. Top left: partition that cannot be obtained from recursive binary splitting. Bottom left: tree corresponding to the partition. Bottom right:  perspective plot of the prediction surface.}
  \label{fig:friedman}
   \vspace{-10pt}
\end{figure} 
However, there is a problem: although each partitioning line has a simple description like $X_1 = k$, some of the resulting regions are complicated to describe.
To simplify things, we can restrict ourselves to only consider recursive binary partitions, like the ones shown in the top right plot of Figure~\ref{fig:friedman}.
We first split the space into two regions, and model the response by the mean of $Y$ in each region. 
We choose the variable and split-point to achieve the best prediction for $Y$. 
Then one or both of these regions are split into two more regions, and this process is continued, until some stopping rule is applied. This is the "recursive partitioning" part of the  algorithm. 
For example, in the top right plot of Figure~\ref{fig:friedman}, we first split at $X_1 = t_1$. 
Then the region $X_1 \leq t_1$ is split at $X_2 = t_2$ and the region $X_1 > t_1$ is split at $X_1 = t_3$.
Finally, the region $X_1 > t_3$ is split at $X_2 = t_4$. 
The result of this process is a partition of the data-space into the five regions (or leaves) $R_1, R_2,\cdots, R_5$.
The corresponding regression tree model, $\T$, predicts $Y$ with a constant, $c_i$, in region $R_i$ i.e.,
\begin{equation}
\T = \mathit{f}_{\mathrm{tree}}(X_1,X_2) = \sum_{i=1}^{5} c_i I\left\lbrace(X_1,X_2) \in R_i\right\rbrace,
\label{eq:response}
\end{equation}
where $I\{X\in R\}$ is a function that is equal to $1$, if $X\in R$, and $0$, otherwise.
This same model can be represented by the binary tree shown in the bottom left of Figure~\ref{fig:friedman}. 
The full data-set sits at the top of the tree. 
Observations satisfying the condition at each node are assigned to the left branch, and the others to the right branch. The terminal nodes or leaves of the tree correspond to the regions $R_1,R_2,...,R_5$. 
The bottom right plot of Figure~\ref{fig:friedman}, shows the perspective plot of the regression surface obtained as a result of building a regression tree with the 5 constants $c_i,\ i=1,2,3,4,5$. 
\paragraph{Node splitting criteria}
Now, a first question is: \emph{how to grow a regression tree?}
Suppose our dataset, $(\X,\Y)$, consisting of $p$ features, i.e. $\X = \{X_1,X_2,\ldots,X_p\}$, and one response variable, i.e. $\Y = \{Y\}$. Suppose we have $|(\X,\Y)| = n$ observations (samples): $(x_i,y_i)$, $i=1,2,\ldots,n$, with $x_i=(x_{i1},x_{i2},\ldots,x_{ip})$.
For regression trees we adopt the sum of squares as our splitting criteria, i.e. a variable at a node will be split if it minimizes the following sum of squares between the predicted response and the actual output variable:
\begin{equation}
\sum_i (y_i - \T(x_i))^2.
\end{equation}
The best response $c_i$ (from equation~\ref{eq:response} for the partition $R_i$), is just the average of output samples in the region $R_i$, i.e.
\begin{equation}
c_i = avg(y_i|x_i \in R_i).
\label{eq:average}
\end{equation}
Finding the best binary partition in terms of minimum sum of squares is generally computationally infeasible. 
A greedy algorithm is used instead. Starting with all of the data, consider a splitting variable $j$ and split point $s$, and define the following pair of left ($R_L$) and right ($R_R$) half-planes
\begin{equation}
\begin{aligned}
R_L(j,s) = \left\lbrace X|X_j \leq s \right\rbrace ,\\
R_R(j,s) = \left\lbrace X|X_j > s \right\rbrace
\end{aligned}
\end{equation}
The splitting variable $j$ and the split point $s$ is obtained by solving the following minimization:
\begin{equation}
\begin{aligned}
\underset{j,s}{\text{min}}\left[\underset{c_L}{\text{min}} \sum_{x_i\in R_L(j,s)} (y_i - c_L)^2 + \underset{c_R}{\text{min}} \sum_{x_i\in R_R(j,s)} (y_i - c_R)^2 \right]  
\end{aligned}
\label{eq:split}
\end{equation}
where, for any choice of $j$ and $s$, the inner minimization in equation~\ref{eq:split} is solved using
\begin{equation}
\begin{aligned}
c_L = \text{avg}(y_i|x_i \in R_L(j,s)),\\
c_R = \text{avg}(y_i|x_i \in R_R(j,s)).
\end{aligned}
\end{equation}
For each splitting variable $X_j$, the determination of the split point $s$ can be done very quickly and hence by scanning through all of the inputs ($X_i$'s), the determination of the best pair $(j, s)$ is feasible.
Having found the best split, we partition the data into the two resulting regions and repeat the splitting process on each of the two regions. Then this process is repeated on all of the resulting regions.

Rather than splitting each node into just two regions at each stage, we might consider multiway splits into more than two groups. 
While this can sometimes be useful, it is not a good general strategy. The problem is that multiway splits fragment the data too quickly, leaving insufficient data at the next level down. 
Hence we would want to use such splits only when needed. 
Also multiway splits can be achieved by a series of binary splits.

\paragraph{Stopping criteria and pruning}
At this point, the second question is: \emph{How large should we grow the tree?}
Every recursive algorithm needs to know when it's done, i.e. it requires a stopping criteria. 
For regression trees this means when to stop splitting the nodes. 
A very large tree might over fit the data, while a small tree might not capture the important structure.
Tree size is a tuning parameter governing the model’s complexity, and the optimal tree size should be adaptively chosen from the data. 
One approach is to split tree nodes only if the decrease in sum-of-squares due to the split exceeds some threshold. 
However, this strategy is myopic, since a seemingly worthless split might lead to a very good split below it.
A preferred, strategy is to grow a large tree, stopping the splitting process only when some minimum number of data points at a node (\texttt{MinLeaf}) is reached. Then this large tree is pruned using cost-complexity pruning methods.

Define a subtree $\T_{sub} \subset \T$ to be any tree that can be obtained by pruning $\T$, i.e. collapsing any number of its non-terminal nodes. 
Let node $i$ corresponding to the partition $R_i$. $|\T_{sub}|$ denotes the number of terminal nodes in $\T_{sub}$
Define,
\begin{equation}
\begin{aligned}
N_i = \# \left\lbrace x_i \in R_i \right\rbrace,\\
c_i = \frac{1}{N_i} \sum_{x_i \in R_i} y_i, \\
Q_i(T) = \frac{1}{N_i} \sum_{x_i \in R_i} (y_i - c_i)^2
\end{aligned}
\end{equation}
where $N_i$ is the number of samples in the partition $R_i$, $c_i$ is the estimate of $Y$ within $R_i$ and $Q_i(T)$ is the mean square error of the estimate $c_i$. The cost complexity criteria is then defined as:
\begin{equation}
C_{\alpha}(\T_{sub}) = \sum_{i=1}^{|\T_{sub}|} N_i Q_i(T) + \alpha |\T_{sub}|
\end{equation}
The goal is to find, for each $\alpha$, the subtree $\T_{\alpha} \subset \T$ to minimize $C_{\alpha}(\T_{sub})$. 
The tuning parameter $\alpha \geq 0$ governs the trade off between tree size and its goodness of fit to the data.
For each $\alpha$ one can show that there is a unique smallest subtree $\T_{\alpha}$ that minimizes $C_{\alpha}(\T_{sub})$~\cite{Ripley96}. Estimation of $\alpha$ is achieved by cross-validation.

\subsection{Random Forests}

### 2018-01-20 paper/final/main.tex
Introduction and noisy forecast
git merge
Appendix Reg Trees

@@
 \usepackage{eurosym}
+\usepackage{longtable}
+
+\newtheorem{problem}{Problem}
+\newtheorem{remark}{Remark}
 
@@
 
-\title{Data Predictive Control with Random Forests:\\Theory and Applications}
+\title{\textcolor[rgb]{0,0,1}{Data-Driven Model Predictive Control for Energy and Thermal Comfort Optimization in Building Energy Systems using Random Forests}}
 
@@
 \appendix
-\section{REGRESSION TREES}\label{AppendixRT}
-TO BE DONE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
-\section{RANDOM FORESTS}\label{AppendixRF}
-TO BE DONE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
+\input{appendixRegTrees.tex}
 \end{document}

### 2018-01-20 paper/final/intro.tex
data-driven

@@
 In this paper, we extend these results providing the following contributions.
-In this paper, we provide a new methodology based on random forests that overcomes the drawbacks of all our previous work, and more precisely: we obtain better performance when compared to other approaches (both optimal and rule-based), we provide and validate robustness with respect to disturbance uncertainties
-Last but not the least, we test our method using a real data set. 
-More precisely, the paper makes the following contributions:
+In this paper, we provide a new methodology based on random forests that overcomes the drawbacks of all our previous work, and more precisely: we obtain better performance when compared to other approaches (both optimal and rule-based), we provide and validate robustness with respect to disturbance uncertainties.
+Further, we test our method using a real data set. 
+More precisely, the paper makes the following contributions.
 \begin{enumerate}
@@
 \section{COMPLEXITY ISSUES IN MODELING BUILDINGS: PHYSICS-BASED VS DATA-DRIVEN}\label{secModelingIssuesExample}
-\textcolor[rgb]{0,0,1}{In this section we provide a detailed example to emphasize differences in terms of complexity when modeling a building using physical laws versus machine learning.
+\textcolor[rgb]{0,0,1}{In this section we provide a detailed example to emphasize the differences in terms of complexity when modeling a building using physical laws versus machine learning.
 We show that the data-driven modeling eliminates several drawbacks that occur using physics-based models such as the need to have good knowledge of the building structure and the material properties, time required to build a model and limited availability of sensors.
@@
 
-\textcolor[rgb]{0,0,1}{The goal with the data-driven modeling is to learn, based on historical measurable data and without requiring physical details of a building, a function map
+\textcolor[rgb]{0,0,1}{The goal with the data-driven modeling is to learn, based on historical measurable data and without modeling physical details of a building, a function map
 \begin{equation}\label{E:MLexample}
@@
 \end{equation}
-where \(y\) represents the outputs, \(u\) the inputs and \(d\) the disturbances as defined earlier, and \(\delta\) is a memory variable.
-Note that the internal unmeasurable variables used in the previous section to model the building are not considered here.
-In this paper we use machine learning to learn such black-box models, with the objective of finding the hyperparameters (for a given structure like Random Forests \textcolor[rgb]{1,0,0}{Ale: referring to a structure like random forests here is confusing}) of a model that best explain the input-output relationship within the measured variables, compensating the effect of the internal unmeasurable variables.
-The advantages of data-driven modeling with respect to physics-based modeling in the case of buildings can be summarised as follows:}
-\textcolor[rgb]{1,0,0}{Ale: we need to improve the following list: first we can consider the list of drawbacks of the physics-based modeling that Francesco introduced in a previous version and comment for each the corresponding improvement via data-driven approach; then we also need to improve much more the text, which sometimes is unclear and verbose.}
-
+where \(y\) represents the outputs, \(u\) the inputs and \(d\) the disturbances, and \(\delta\) is a memory variable which captures the dynamical behavior.
+While the inputs and external disturbances are exactly the same as discussed earlier in Section \ref{SSS:physics-based}, the outputs in \eqref{E:MLexample} include only the variables that are directly measurable through already installed sensors like thermostats and multimeters.
+Therefore, many internal states like the temperatures of different layers -- interior, middle and exterior -- of the walls, the floors and the ceilings are not required for black-box modeling which reduces the order of complexity.
+We use machine learning to learn such black-box models with the objective of finding the hyperparameters (for a given structure like Random Forests) of a model that best explain the input-output relationship within the measured variables, compensating the effect of the internal unmeasurable variables.
+The data-driven modeling addresses the following drawbacks of the physical counterpart.}
 \begin{enumerate}
-	\item \textcolor[rgb]{0,0,1}{\textbf{Domain expertise:} 
-	This procedure (\textcolor[rgb]{1,0,0}{NOT CLEAR IF EXPLAINED IN ONLY 2 LINES BEFORE}) eliminates the need of a building expert, simplifies the modeling complexities (\textcolor[rgb]{1,0,0}{HOW?}) described in Section \ref{SSS:physics-based} which ultimately reduces the cost of modeling a building.
-Further, for a new building, given the historical data from the building, the same process can be repeated to identify a control-oriented model for MPC.}
-	\item \textcolor[rgb]{0,0,1}{\textbf{Sensor cost:}
-	Due to large number of states and variables in the physical modeling, a lot of measurements are needed to use the model to predict the system's behavior \textcolor[rgb]{1,0,0}{IF USING A MODEL-BASED APPROACH}.
+	\item \textcolor[rgb]{0,0,1}{
+	During physical modeling in \eqref{E:RCHeatExchange}, in order to keep the model simple, the heat exchange between layers is assumed to exhibit a linear behavior. 
+	In the case of heat flux in \eqref{E:RCfluxes}, again, a bilinear model ignores complex nonlinearities.
+	Hence, much nonlinearities are neglected to simplify the modeling complexities.
+	We avoid such assumptions in data-driven modeling where a nonlinear function such as Neural Network or Random Forest that represents the dynamics of zone temperatures or power consumption is learned rather efficiently and accurately.
+	For example, if \(y\) is the zone temperature in Equation \eqref{E:MLexample}, we learn a nonlinear function \(f\) which depends on current and previous inputs and disturbances. 
+	The hyperparameters of \(f\) are trained automatically depending upon the learning algorithm.}
+	\item \textcolor[rgb]{0,0,1}{
+	Physical modeling neglects different geometries and different materials of the floors assuming they are the same for each floor which is never true in reality.
+	Moreover, in many cases, the details of the construction layout and equipments are not even available so many parameters have to be guessed making physical modeling a difficult choice.
+	On the other hand, data-driven modeling automatically captures the interaction with the environment while optimizing the hyperparameters.
+Thus, the data from building's construction/materials/equipments are not required explicitly.}
+	\item \textcolor[rgb]{0,0,1}{
+	Tuning physical parameters in \eqref{E:RCZoneEq}, \eqref{E:RCHeatExchange}, and \eqref{E:RCfluxes} requires expert know-how which adds to the cost and time of modeling.
+	Data-driven approach reduces both cost and time by an order of magnitude as we directly work with the senor data without explicitly modeling internal states.
+	Further, for a new building, given the historical data from the building, data-driven method is scalable as the same process can be repeated to identify a control-oriented model for MPC.}
+	\item \textcolor[rgb]{0,0,1}{
+	Physical modeling large number of states and variables.
+	Thus, many more measurements are needed to use the physical model to predict the system's behavior.
 	This can be expensive due to necessary new sensor installations.
@@
 	However, observability problems can limit the construction of such observers \cite{Dorf2011MCS}. 
-	On the other hand, data-driven modeling obviates the need to model the state of the walls and floor layers\textcolor[rgb]{1,0,0}{AND OTHER STATES DIFFICULT TO MEASURE} so we do not have as many states.\textcolor[rgb]{1,0,0}{EXPLAIN THAT THE MACHINE LEARNING ALGORITHM IS ABLE TO COMPENSATE MISSING VARIABLES}
-	It relies only on direct measurements from the sensors such as thermostats and multimeters reducing the need for new installations and hence the cost of modeling.
-	Further, when some of the states are not directly measurable, often adding proxy \textcolor[rgb]{1,0,0}{I.E. ??}variables suffices. 
+	On the other hand, data-driven modeling obviates the need to model the internal states of the walls, floors and ceilings so we do not have as many states. 
+	It relies only on direct measurements from the sensors such as thermostats and multimeters, and weather data, reducing the need for new installations and hence the cost of modeling.
+	The behavior of the internal/unobservable/missing states is captured in the parameters of function \(f\).
 	For example, if we do not have the measurement of the energy produced by a photovoltaic panel, the weather conditions will account for it.}
-	\item \textcolor[rgb]{0,0,1}{\textbf{Linearity assumption:}
-	If \(y\) in Equation \eqref{E:MLexample} is the average room temperature for one of the zones, a nonlinear function \(f\) is more appropriate, for example \(f\) may belong to the class of Neural Networks or Random Forests.
-	During physical modeling in \eqref{E:RCHeatExchange}, in order to keep the model simple, the heat exchange between layers is modeled with a linear behavior. 
-	Hence all the nonlinearities were neglected.
-	In the case of heat flux in \eqref{E:RCfluxes}, again, a linear model is not good enough to provide a good approximation. For this reason a bilinear model was considered.
-	However, more complex nonlinearities have been neglected due to the already high complexity of the model \cite{Sturzenegger2016}. 
-	In data-driven modeling, we can relax these linearity assumptions as nonlinear functions can be learned rather efficiently and accurately.\textcolor[rgb]{1,0,0}{IT IS NOT CLEAR THAT THE DD ALGORITHM TAKE CARE OF IT CREATING ITS INTERNAL STRUCTURE}}
-	\item \textcolor[rgb]{0,0,1}{\textbf{Other assumptions:} 
-	Due to the high modeling complexity, physical modeling neglects different geometries and different materials of the floors assuming they are the same for each floor which is never true in reality.
-	In many cases, the details of construction layout and equipments are not even available so many parameters have to be guessed.
-	Data-driven modeling depends on the sensor data directly. 
-	Hence, such assumptions never arise.}
 \end{enumerate}
 
-\textcolor[rgb]{0,0,1}{To summarize the data-driven approaches have the potential to simplify the modeling of buildings to a large extent reducing the overall cost and time investment, while avoiding several assumptions that are generally made in standard physics-based modeling procedures and that reduce the accuracy.
-Also, unlike physics-based modeling, the parameters (or hyperparamters) of the model are automatically determined by the learning algorithm to compensate the effect of unobservable variables on the input-output relation, and the best settings determined by cross validation.
-Granted the advantages of data-driven modeling the challenge lies now in using such models for optimal predictive control, which is exactly the focus of this paper.
-In the next sections we will demonstrate how the training algorithm for Random Forests can be modified to develop control oriented models that enable Model Predictive Control with better performance when compared to other approaches and robustness with respect to disturbance uncertainties.}
+\textcolor[rgb]{0,0,1}{
+	To summarize, the data-driven approaches have the potential to simplify the modeling of buildings to a large extent reducing the overall cost and time investment, while avoiding several assumptions that are generally made in standard physics-based modeling procedures and that reduce the accuracy.
+	Also, unlike physics-based modeling, the parameters (or hyperparamters) of the model are automatically determined by the learning algorithm to compensate for the effect of unobservable variables on the input-output relation; the best settings being determined by cross validation.\\
+	Given the advantages of data-driven modeling, the challenge now lies in using such models for optimal predictive control.
+	This is exactly the focus of this paper. 
+	In the next sections, we will demonstrate how the training algorithm for Random Forests can be modified to develop control oriented models that enable Model Predictive Control, and provide better performance when compared to other approaches and robustness with respect to disturbance uncertainties.}

### 2018-01-21 paper/final/casestudyReal.tex
disturbance section edits

@@
 \subsubsection{Weather forecast subject to uncertainty}\label{SSS:DisturbanceUncertain}
-In \cite{Petersen2014AE} the authors investigated, through a large-scale simulation study, the effect of weather forecast uncertainty on the performance of MPC in building systems operations, and compared to the performance of a rule-based strategy.
-They considered 48 different scenarios of uncertainties on 72 hours weather forecast.
-On such long horizon, results have shown that, with few exceptions, MPC outperforms the rule-based controller in case of energy savings and/or thermal indoor environment, and is in general quite close to the perfect forecast case, despite the uncertainty in weather forecast.}
+The authors in \cite{Petersen2014AE} investigate the effect of uncertainty in the weather forecast on the performance of MPC in building systems operations through a large-scale simulation study, and compare against a rule-based strategy.
+They consider 48 different scenarios of uncertainties for 72 hours weather forecast.
+With such a long horizon, results have shown that (with a few exceptions) MPC outperforms the rule-based controller in providing energy savings, and is in general quite close to the perfect forecast case despite the uncertainty in weather forecast.}
 
-\textcolor[rgb]{0,0,1}{Horizon length for the DPC algorithm presented in this paper is usually much shorter, as for example $6\ hours$ in Section \ref{S:proof}, $7\ hours$ in Section \ref{S:casestudy}, and $40\ minutes$ in Section \ref{S:realCaseStudy}.
-This means the weather forecast is more accurate than the 72 hours case presented in \cite{Petersen2014AE}, hence also control performance is even closer to the perfect knowledge forecast case.
-Nevertheless, to show the robustness of our approach, in this section we provide simulations considering noisy weather forecast.}
+\textcolor[rgb]{0,0,1}{The length of horizon for the DPC algorithm presented in this paper is usually much shorter, for example $6$ hours in Section \ref{S:proof}, $7$ hours in Section \ref{S:casestudy}, and $40$ minutes in Section \ref{S:realCaseStudy}.
+%This means the weather forecast is more accurate than the 72 hours case presented in \cite{Petersen2014AE}, hence also control performance is even closer to the perfect knowledge forecast case.
+We next show the robustness of our approach considering noisy weather forecast.}
 
-\textcolor[rgb]{0,0,1}{To this aim, we add noises with normal distribution on the disturbance variables in $\X^d$, used to obtain parameters $\hat \Theta$ in the DPC problem \eqref{E:DPCrealcase}, i.e.
-the disturbance in input in the blue rectangle in Figure \ref{F:overview}, while we use the correct ones to simulate the process, i.e.
-the disturbance in input in the "Plant" box in the red rectangle in Figure \ref{F:overview}.
-In Table \ref{T:NoiseParameters} we summarize the mean and the deviation values of the Gaussian noises that we add to each variable.
-We obviously do not add any noise on the "Time of the day" and on the "Day of the week" since those are perfectly known.
-We considered zero mean since we suppose there is no constant offset in the forecast, but only an error that can over or underestimate the predictions.
-In the table, we also report the range of the values that the variables assumed during the reference period of the historical dataset.
-This is to show that the deviation values we chose add a quite bad error on the forecast.
-In particular, for example, a deviation of $0.5\degree C$ on the outside air temperature means that the error on the predicted temperature lies within the range of $\pm 1.5\degree C$ with a probability of $99\%$.
-We solved DPC problem \ref{P:dpcRealCase} considering the new parameters $\hat \Theta$, obtained using the aforementioned noisy forecast, obtaining the following results, where we show the change in performance on bounds violation and energy consumption due to the forecast inaccuracy.}
+\textcolor[rgb]{0,0,1}{We add noise with normal distribution to the disturbance variables in $\X^d$ which is used to obtain parameters $\hat \Theta$ in the DPC problem \eqref{E:DPCrealcase} i.e.~the disturbance in the input in the blue rectangle in Figure \ref{F:overview}, while we use the true values to simulate the process i.e.~the disturbance in the input in the box ``Plant" in the red rectangle in Figure \ref{F:overview}.
+We summarize the mean and the standard deviation of the Gaussian noises that we add to each variable in Table \ref{T:NoiseParameters}.
+Since ``Time of the day" and ``Day of the week" are perfectly known, we do not add any noise to these variables.
+We assume the noise to have zero mean.
+Table \ref{T:NoiseParameters} also lists the range of the variables in the historical dataset.
+This is to show that the standard deviation we choose adds significant error to the forecast.
+For example, a standard deviation of $0.5\degree C$ on the outside air temperature means that the error on the predicted temperature lies within the range of $\pm 1.5\degree C$ with a probability of $99\%$.
+We solve the DPC problem \ref{P:dpcRealCase} considering the $\hat \Theta$ obtained using the noisy forecast.
+We obtain the following results. We show the change in the performance on bounds violation and energy consumption due to inaccurate forecast.}
 \begin{table}[t!]
@@
 	\end{tabular}}
-	\caption{\textcolor[rgb]{0,0,1}{Mean and deviation values of the Gaussian noises added on the weather forecast data, and the range of variation of the variables in the historical dataset.}}
+	\caption{\textcolor[rgb]{0,0,1}{Mean and standard deviation of the Gaussian noise added on the weather forecast data, and the range of the variables in the historical dataset.}}
 	\captionsetup{justification=centering}
@@
 
-\textcolor[rgb]{0,0,1}{\paragraph{Result 1} In Figure \ref{F:ComparisonTempNoisy} a comparison of the DPC results in terms of temperature control for thermal comfort, between the perfect and the noisy forecast, is shown.
-Also here, for the sake of plot's clarity, we show only 4 days and a half of the 15 days simulative period, and we split the plot into 3 sub-figures referring to the small, medium and large violation cases.
-We can see that, except for an isolated case in the medium violation case (slightly before the "02/05"), we have a small performance deterioration in terms of bounds violation, although this is more highlighted in the large violation case.
-However, the maximum violation, as in the case of perfect forecast, is still lower than $1\,\degree C$.
-In particular, we report in Table \ref{T:violationErrorsNoisy} the violation errors in terms of MBE and CV(RMSE) (in $\%$) computed for the whole simulative period, i.e. 15 days.
-If we compare them with the ones in Table \ref{T:violationErrors}, we can see that the performance of DPC with noisy forecast are still good with respect to the performance of DPC with perfect forecast, despite the extremely bad prediction error that we considered on the weather forecast.
+\textcolor[rgb]{0,0,1}{\paragraph{Result 1} Figure \ref{F:ComparisonTempNoisy} shows a comparison of the DPC results in terms of temperature control for thermal comfort between the perfect and the noisy forecast.
+For the sake of clarity, we show only \(4\frac{1}{2}\) days of the 15 days simulation period. We split the plot into 3 sub-figures referring to the small, medium and large violation cases.
+We see that, except for an isolated case in the medium violation case (slightly before ``02/05"), we have a small performance deterioration in terms of bounds violation.
+The effect is more prominent in the large violation case.
+However, the maximum violation is still lower than $1\,\degree C$, as in the case of perfect forecast.
+We report the violation errors in terms of MBE and CV(RMSE) (in $\%$) computed for the whole simulative period (15 days) in Table \ref{T:violationErrorsNoisy}.
+Comparing them with the values obtained earlier in Table \ref{T:violationErrors}, we see that the performance of DPC is robust to noisy predictions despite a large error in the weather forecast.
 }
@@
 	\end{center}
-	\caption{\textcolor[rgb]{0,0,1}{Comparison of DPC considering perfect and noisy weather forecasts. Small, medium and large violation cases are compared.}}
+	\caption{\textcolor[rgb]{0,0,1}{Comparison of DPC considering perfect and noisy weather forecast. Small, medium and large violation cases are compared.}}
 	\captionsetup{justification=centering}
@@
 	%	}
-	\caption{\textcolor[rgb]{0,0,1}{Lower Bound Violation (LBV) and Upper Bound Violation (UBV) errors expressed as $\mathrm{MBE}\%$ and $\mathrm{CV(RMSE)}\%$ for DPC Small Violation (DPC-SV), DPC Medium Violation (DPC-MV), DPC large Violation (DPC-LV) and bang-bang controller, considering noisy weather forecast.}}
+	\caption{\textcolor[rgb]{0,0,1}{Lower Bound Violation (LBV) and Upper Bound Violation (UBV) errors expressed as $\mathrm{MBE}\%$ and $\mathrm{CV(RMSE)}\%$ for DPC Small Violation (DPC-SV), DPC Medium Violation (DPC-MV), DPC large Violation (DPC-LV) and bang-bang controller considering noisy weather forecast.}}
 	\captionsetup{justification=centering}
@@
 \textcolor[rgb]{0,0,1}{\paragraph{Result 2} 
-In Figure \ref{F:comparison_all_energy_E+_noisy}, we show the difference in terms of energy consumption considering the perfect weather forecast (full line) and the noisy one (dashed line).
-We can observe that the case with imperfect forecast is very close to the perfect forecast case.
-We recall that we are optimizing a weighted sum of energy and thermal comfort, so depending on the conditions for which the disturbance is over or underestimated, the MPC can require to use more or less energy to supply the prediction errors.
-In the small violation case, the energy consumption of the 2 conditions is closer than in the others.
+In Figure \ref{F:comparison_all_energy_E+_noisy}, we show the difference in terms of energy consumption considering the perfect weather forecast (solid line) and the noisy forecast (dashed line).
+We observe that the energy usage is very close in both cases.
+Recall, we optimize a weighted sum of energy and thermal comfort, so depending on the conditions for which the disturbance is over or underestimated, the MPC solution changes depending upon the prediction errors.
+In the small violation case, the energy consumption of the two conditions is closer than in the others.
 This is because the DPC keeps the temperature within the bounds with an extremely small error in bounds violation.
 In particular, the only exception can be seen before May $3^{rd}$ (03/05), where, due to the prediction error, the temperature violates the bounds, so the control uses more energy (than the perfect case) to bring it back above the threshold.
-In the other 2 cases, when the temperature violates the bounds, due to the fact that violation constraints are more relaxed, the DPC does not require a too big control effort, hence using less energy.}
+In the other two cases, when the temperature violates the bounds, due to the fact that violation constraints are more relaxed, the DPC does not require much larger control effort, hence using lesser energy.}
 
-\textcolor[rgb]{0,0,1}{These results show the strong robustness of the DPC with respect to errors in the weather forecast.}
+\textcolor[rgb]{0,0,1}{The above results show the robustness of the DPC with respect to errors in the weather forecast.}
 \begin{figure}[t!]
@@
 	\end{center}
-	\caption{\textcolor[rgb]{0,0,1}{Comparison of DPC in terms of thermal energy saving using EnergyPlus model, considering perfect (full) and imperfect (dashed) weather forecast, over $15$ days of the testing period with different violation configurations.
-The performance are pretty similar, showing the potential robustness of DPC.}}
+	\caption{\textcolor[rgb]{0,0,1}{Comparison of DPC in terms of thermal energy savings, considering perfect (full) and imperfect (dashed) weather forecast, over $15$ days of the testing period with different violation settings.
+The performance is similar, showing the robustness of DPC.}}
 	\label{F:comparison_all_energy_E+_noisy}

### 2018-01-21 paper/final/abstract.tex
first change fot the revision
small changes

--- 2017-09-30 paper/abstract.tex
@@
 Model-based control techniques have been widely used over the past years in control systems to improve system performance of simple controllers, such as PID, bang-bang, etc.
-Among these, one of the most popular is Model Predictive Control (MPC) due to its ability to predict the system's behavior over a future horizon and handle constraints in the optimal control setup.
+Among these, one of the most popular is Model Predictive Control (MPC), due to its ability to predict the system's behavior over a future horizon and handle constraints in the optimal control setup.
 A key factor prohibiting the widespread adoption of MPC for complex systems is related to the difficulties (cost, time, and effort) associated with the model identification. 
-On the other hand, data-driven models are not suitable for control because of non-physical models.
-%Due to their lack of dynamical equations to predict real systems behavior, data-driven models are not suitable for control. 
-We introduce a novel idea for predictive control based on system data, Data Predictive Control (DPC), leveraging machine learning algorithms, in particular regression trees and random forests.
+\textcolor[rgb]{0,0,1}{To overcome this problem, machine learning can be used to create models that are data-driven.
+However, data-driven models introduce several issues that limit their use for MPC.
+We introduce a novel idea for predictive control based on system data, Data-driven model Predictive Control (DPC)}, leveraging machine learning algorithms, in particular regression trees and random forests.
 Using a bilinear model, we show that DPC has comparable performance with MPC.
 We further apply DPC to the Demand Response problem of a large scale multi-story EnergyPlus building building model, and show that DPC curtails the desired power usage with high confidence.
-Finally, we use DPC to optimize ON/OFF scheduling of the heating system of a real house in order to minimize the power consumption while guaranteeing thermal comfort for occupants.
+Further, we use DPC to optimize ON/OFF scheduling of the heating system of a real house in order to minimize the power consumption while guaranteeing thermal comfort for occupants.
 Our results show that with DPC we obtain an energy saving in terms of primary energy up to $49.2\%$ when compared to a bang-bang controller.
+\textcolor[rgb]{0,0,1}{Finally, we show the robustness of DPC under disturbance uncertainties.}

### 2018-01-21 paper/final/appendixRegTrees.tex
small changes

@@
-\section{Regression Trees and Random Forests}
+\section{Regression Trees}
 \label{A:RegTrees}
-
-In this appendix we first start explaining how Regression Trees are built using an example adapted from~\cite{hastie2009elements}, and then we introduce Random Forests.
-
-\subsection{Regression Trees}
+\textcolor[rgb]{0,0,1}{
+In this appendix we explain how Regression Trees are built using an example adapted from~\cite{hastie2009elements}.
 Tree-based methods partition the feature space into a set of rectangles (more formally, hyper-rectangles) and then fit a simple model in each one. They are conceptually simple yet powerful.
@@
   \centering
-  \includegraphics[width=0.7\columnwidth]{figures/friedman}
+  \includegraphics[width=0.7\columnwidth]{figures/friedman1.eps}
   \vspace{-8pt}
@@
 The terminal nodes or leaves of the tree correspond to the regions $R_1,R_2,...,R_5$.
-The bottom right plot of Figure~\ref{fig:friedman}, shows the perspective plot of the regression surface obtained as a result of building a regression tree with the 5 constants $c_i,\ i=1,2,3,4,5$. 
-\paragraph{Node splitting criteria}
+The bottom right plot of Figure~\ref{fig:friedman}, shows the perspective plot of the regression surface obtained as a result of building a regression tree with the 5 constants $c_i,\ i=1,2,3,4,5$. }
+
+\textcolor[rgb]{0,0,1}{
+\paragraph{\textbf{Node splitting criteria}}
 Now, a first question is: \emph{how to grow a regression tree?}
@@
 Having found the best split, we partition the data into the two resulting regions and repeat the splitting process on each of the two regions.
-Then this process is repeated on all of the resulting regions.
+Then this process is repeated on all of the resulting regions.}
 
+\textcolor[rgb]{0,0,1}{
 Rather than splitting each node into just two regions at each stage, we might consider multiway splits into more than two groups. 
@@
 Hence we would want to use such splits only when needed. 
-Also multiway splits can be achieved by a series of binary splits.
+Also multiway splits can be achieved by a series of binary splits.}
 
-\paragraph{Stopping criteria and pruning}
+\textcolor[rgb]{0,0,1}{
+\paragraph{\textbf{Stopping criteria and pruning}}
 At this point, the second question is: \emph{How large should we grow the tree?}
@@
 A preferred, strategy is to grow a large tree, stopping the splitting process only when some minimum number of data points at a node (\texttt{MinLeaf}) is reached.
-Then this large tree is pruned using cost-complexity pruning methods.
+Then this large tree is pruned using cost-complexity pruning methods.}
 
+\textcolor[rgb]{0,0,1}{
 Define a subtree $\T_{sub} \subset \T$ to be any tree that can be obtained by pruning $\T$, i.e. collapsing any number of its non-terminal nodes. 
@@
 The tuning parameter $\alpha \geq 0$ governs the trade off between tree size and its goodness of fit to the data.
-For each $\alpha$ one can show that there is a unique smallest subtree $\T_{\alpha}$ that minimizes $C_{\alpha}(\T_{sub})$~\cite{Ripley96}. Estimation of $\alpha$ is achieved by cross-validation.
-
-\subsection{Random Forests}
+For each $\alpha$ one can show that there is a unique smallest subtree $\T_{\alpha}$ that minimizes $C_{\alpha}(\T_{sub})$~\cite{Ripley96}. Estimation of $\alpha$ is achieved by cross-validation.}

### 2018-01-21 paper/final/conclusion.tex
small changes

@@
 
-To overcome the difficulties associated with the model identification in Model Predictive Control (MPC), we introduce a novel idea for predictive control using data: Data Predictive Control (DPC).
+To overcome the difficulties associated with the model identification in Model Predictive Control (MPC), we introduce a novel idea for predictive control using data: Data-driven model Predictive Control (DPC).
 Data-based control strategies are usually based on non-physical models, therefore the can not be integrated with most of the classical control approaches.
@@
 To this aim, we present two algorithms, based on trees and random forests, to create data-driven models to be used in the DPC strategy.
-We then apply DPC to $3$ different case study in order to prove its strength.
+We then apply DPC to $3$ different case studies in order to prove its strength.
 \begin{enumerate}
-\item \emph{\textbf{Comparison with MPC}} We compare the performance of our Data Predictive Control to MPC on a multivariable bilinear building model.
+\item \emph{\textbf{Comparison with MPC.}} We compare the performance of our DPC to MPC on a multivariable bilinear building model.
 We establish that DPC with random forests shows a remarkable similarity to MPC in the optimal control strategies explaining $70\%$ variance.
 On the other hand, DPC with regression trees suffers from practical limitations due to model overfitting.
-\item \emph{\textbf{Demand Response application}} We further apply DPC with random forests to a large scale 6 story EnergyPlus model with 22 zones for which the traditional model-based control is largely unsuitable due to complex dynamics and the cost of model identification.
+\item \emph{\textbf{Demand Response application.}} We further apply DPC with random forests to a large scale 6 story EnergyPlus model with 22 zones for which the traditional model-based control is largely unsuitable due to complex dynamics and the cost of model identification.
 We show that DPC, relying only on the sensor data, can provide significant energy savings while maintaining thermal comfort.
 Our results demonstrate that even for such complex system, DPC tracks a reference signal with a mean error of $3\%$.
-\item \emph{\textbf{Optimal heating system scheduling of a real house}} We finally apply DPC using data from a real house and compare it with the classical bang-bang controller widely used for temperature control in houses.
+\item \emph{\textbf{\textcolor[rgb]{0,0,1}{Optimal heating system scheduling of a real house.}} We finally apply DPC using experimental data from a real house and compare it with the classical bang-bang controller widely used for temperature control in houses.
 We implemented DPC to guarantee energy saving while guaranteeing thermal comfort for the occupants.
-With respect to a bang-bang controller, we obtained an energy saving that goes from $25.4\%$, when we force the temperature to be strictly in a comfort range, to $49.2\%$, if we allow small violations.
+With respect to a bang-bang controller, we obtained an energy saving that goes from $25.4\%$, when we force the temperature to be strictly in a comfort range, to $49.2\%$, if we allow small violations of such range.
+Furthermore, we also demonstrate the robustness of DPC considering uncertainties on weather forecast.}
 \end{enumerate}

### 2018-01-21 paper/final/cs-challenges.tex
small changes

--- 2017-12-24 paper/first submission/cs-challenges.tex
@@
 \begin{enumerate}
-\item \emph{\textbf{Data Availability:}} The main practical challenge for DPC lies in the availability of data for training and we require answers to questions like how much data (functional testing) is required, and how should the sampling be done?
-Therefore, the procedure for optimal experiment design, and model improvement with estimation of variance in predictions is one of the main focus of our ongoing work.
+\item \emph{\textbf{Data Availability:}} The main practical challenge for DPC lies in the availability of data for training, and we require answers to questions like \emph{how much data (functional testing) is required, and how should the sampling be done?} Therefore, the procedure for optimal experiment design, and model improvement with estimation of variance in predictions is one of the main focus of our ongoing work.
 \item \emph{\textbf{Stability:}} While the buildings are inherently stable, many other applications require stability guarantees.
@@
 This will make DPC useful for systems with faster dynamics.
-\item \emph{\textbf{Robustness:}} Another direction of work is on handling uncertainties in the DPC framework, namely an extension to Scenario DPC to account for the disturbance uncertainty.
-This will help us in quantifying the robustness of DPC.
 \end{enumerate}

### 2018-01-21 paper/final/intro.tex
disturbance section edits
small changes

@@
 
-\textcolor[rgb]{0,0,1}{Control-oriented predictive models of an energy system's dynamics and energy consumption, are needed for understanding and improving the overall energy efficiency and operating costs of a building.
+\textcolor[rgb]{0,0,1}{Control-oriented models of energy system's dynamics and energy consumption, are needed for understanding and improving the overall energy efficiency and operating costs of a building.
 With a reasonably accurate forecast of future weather and building operating conditions, dynamical models can be used to predict the energy needs of the building over a prediction horizon, as is the case with Model Predictive Control (MPC) \cite{Sturzenegger2016}.
@@
 The task is much more complicated and time consuming in the case of a large building and often times, it can be even more complex and involved than the controller design itself.
-After several years of work on using first principles based models for peak power reduction, energy optimization and thermal comfort for buildings, multiple authors~\cite{Sturzenegger2016, vzavcekova2014} have concluded that the biggest hurdle to mass adoption of intelligent building control is the cost, time and effort required to capture accurate dynamical models of the buildings.
+After several years of work on using first principles based models for peak power reduction, energy optimization and thermal comfort for buildings, multiple authors~\cite{Sturzenegger2016, vzavcekova2014} have concluded that the biggest hurdle to mass adoption of MPC in intelligent building control is the cost, time and effort required to capture accurate dynamical models of the buildings.
 The user expertise, time, and associated sensor costs required to develop a model of a single building is very high.
 Thus, the payback period for the upfront hardware and software installation is expected to be too high, making MPC an uneconomical choice for energy management.
+This is probably the main reason why rule-based control strategy have been widely used so far.
 Indeed, there are several reasons why physics-based modeling is hard for complex systems like buildings:}
@@
 \begin{enumerate}
-	\item \textcolor[rgb]{0,0,1}{\textbf{Model capture} - a building modeling domain expert typically uses a software tool to create the geometry of a building from the building design and equipment layout plans, add detailed information about material properties, about equipment and operational schedules.
-There is always a gap between the modeled and the real building and the domain expert must then manually tune the model to match the measured data from the building~\cite{New2012}.
+	\item \textcolor[rgb]{0,0,1}{\textbf{Model capture} - a building modeling domain expert typically uses a software tool to create the geometry of a building from the building design and equipment layout plans, and add detailed information about material properties, equipment and operational schedules.
+	There is always a gap between the modeled and the real building, and the domain expert must then manually tune the model to match the measured data from the building~\cite{New2012}. 
 	Moreover, the modeling process also varies from building to building with the construction and types of installed equipment.
@@
 In practice, due to aforementioned reasons, the control strategies in such systems are often limited to fixed, sometimes ad-hoc, rules that are based on best practices.
-The alternative is to use black-box, or completely data-driven modeling approaches, to obtain a realization of the system's input-output behavior. 
+The alternative is to use black-box, or completely data-driven, modeling approaches, to obtain a realization of the system's input-output behavior. 
 The primary advantage of using data-driven methods is that it has the potential to eliminate the time and effort required to build white and grey box building models. 
@@
 
-\textcolor[rgb]{0,0,1}{The key question now is: can we employ data-driven techniques to reduce the cost of modeling, and still exploit the benefits that MPC has to offer?
+\textcolor[rgb]{0,0,1}{The key question now is: \emph{can we employ data-driven techniques to reduce the cost of modeling, and still exploit the benefits that MPC has to offer?}
 We therefore look for automatic data-driven approaches to control, that are also adaptive, scalable and interpretable.
@@
 In particular, in this paper, we present a method based on Random Forests which uses historical data for receding horizon control.
-Before illustrating our approach, in the next section, we discuss the related literature and the novelty of our approach.}
+Before illustrating our approach, in the next section, we discuss the related literature and the novelty of our contribution.}
 
@@
 A vast literature exists in building energy applications that deals with Demand Response, peak power reduction, energy saving, thermal comfort, and related topics.
-}
+Among them, we selected the ones that we believe are more related to our work.}
 
@@
 			\item model-based, such as ``white-box" and ``grey-box" approaches: \cite{Shakouri2017SCS,Li2014EB,Yoon2014EB,Li2016AE,Harb2016EB,Salakij2016EB,Li2016E,Li2016EB,Hou2013,Cecconi2017EB};
-			\item data-based, i.e.~``black-box" approach, mainly done using Neural Networks: \cite{Safa2017SCS,Neto2008EB,Magnier2010BE,Candanedo2017EB,Ascione2017E,Cecconi2017EB,Li2016AE};
+			\item data-based, i.e.~``black-box" approaches, mainly done using Neural Networks: \cite{Safa2017SCS,Neto2008EB,Magnier2010BE,Candanedo2017EB,Ascione2017E,Cecconi2017EB,Li2016AE};
 			\item simulation tool-based, such as EnergyPlus \cite{energyPlus} and TRNSYS \cite{trnsys2000}: \cite{Yin2016EB,Christantoni2016EB};
@@
 Only in three cases the algorithms are tested on real systems (see Table \ref{T:RelatedWork}: RI - Real Implementation).
-We observe that except for the last 6 cases, which we discuss in detail, either model-based approaches or only tools are considered with/without control or data-driven modeling approaches without control.
+We observe that, except for the last six cases \cite{Macarulla2017,Costanzo2016,Ferreira2012,Afram2017,Behl2016,Jain2017TCPS}, which we discuss in detail, either model-based approaches or only tools are considered with/without control, or data-driven modeling approaches are considered only without control.
 %This is because bridging data-driven models with control theory, especially with the predictive control, introduces several issues \cite{Hou2013}.
-Indeed data-driven control has attracted increasing attraction only in the recent years.
 The last six papers of Table \ref{T:RelatedWork} are more related to the methodology presented in this paper.
@@
 Unfortunately, since Neural Networks models are nonlinear, the MPC based on such models is also nonlinear.
-This means that a global optimal solution cannot be guaranteed and solving the optimization problem becomes computationally harder due to nonlinearties when the complexity of network is high.
+This means that a global optimal solution cannot be guaranteed and solving the optimization problem becomes computationally harder due to nonlinearties when the complexity of the neural network is high.
 \\
-To overcome this complexity, we initially introduced the regression trees-based approach in \cite{Behl2016,Jain2017TCPS}. 
+To overcome this complexity, we introduced the regression trees-based approach in \cite{Behl2016,Jain2017TCPS}. 
 In particular, in \cite{Behl2016}, we used regression trees to setup an MPC problem and apply it to the problem of Demand-Response. 
-However, these models can be used for optimal control with only one-step lookahead prediction.
-Hence, it is not possible to use such approach to control the system considering a prediction over an horizon of arbitrary length. 
+However, this approach can be used for optimal control with only one-step lookahead prediction.
+Hence, it is not possible to use it to control the system considering a prediction over an horizon of arbitrary length. 
 This limitation was addressed in \cite{Jain2017TCPS}, where multi-output regression trees are used instead of single output trees. 
 The different outputs correspond to different steps of the prediction horizon which allows us to setup an MPC problem with a finite horizon. 
-Modeling accuracy using single trees is strongly affected by overfitting and high variance. 
-Such approach has the advantage to be extremely simple from the complexity point of view, but the range of applications is limited. 
-In \cite{JainACC2017,JainCDC2017} we took a different approach that improves the system's identification accuracy -- instead of considering a single tree/forest with multiple output, we considered multiple trees/forests with single output.
+However, modeling accuracy using single trees is strongly affected by overfitting and high variance. 
+On the other hand, such approach has the advantage to be extremely simple from the complexity point of view, although the range of applications is limited. 
+In the conference papers \cite{JainACC2017,JainCDC2017}, we took a different approach that improves the system's identification accuracy -- instead of considering a single tree with multiple output, we considered multiple trees and forests with single output.
 Each tree/forest provides the prediction of the system's behavior for a different time step of the horizon. 
@@
 In this paper, we extend these results providing the following contributions.
-In this paper, we provide a new methodology based on random forests that overcomes the drawbacks of all our previous work, and more precisely: we obtain better performance when compared to other approaches (both optimal and rule-based), we provide and validate robustness with respect to disturbance uncertainties.
-Further, we test our method using a real data set. 
+In this paper, we provide a new methodology based on random forests that overcomes the drawbacks of all our previous work, and more precisely: we obtain better performance when compared to other approaches (both optimal and rule-based), and we provide and validate robustness with respect to disturbance uncertainties.
+Further, we test our method using an experimental data set. 
 More precisely, the paper makes the following contributions.
@@
 	\item In Section \ref{S:dpc}, we formally describe our data-driven Model Predictive Control approach i.e.~Data Predictive Control (DPC).
-For clarity, we first describe DPC based on regression trees and then we introduce DPC based on random forests.
-	\item In Section \ref{S:proof}, we demonstrate that DPC can provide comparable performance compared to MPC (using a physical model) applied to a perfectly known mathematical model.
-More precisely, we first consider a benchmark MPC controller using a bilinear building model whose parameters were identified using experiments on a building in Switzerland: then we apply DPC to a data-set simulated from such bilinear model and derive a controller based only on the data.
+For clarity, we first modify the regression trees algorithm to apply DPC, and then we introduce DPC based on random forests.
+	\item In Section \ref{S:proof}, we demonstrate that DPC can provide comparable performance when compared to MPC (using a physical model) applied to a perfectly known mathematical model.
+More precisely, we first consider a benchmark MPC controller using a bilinear building model whose parameters were identified using experiments on a building in Switzerland, then we apply DPC to a data-set simulated from such bilinear model and derive a controller based only on the data.
 We show that DPC captures 70\% variance in MPC and offers a comparable performance.
-	\item In Section \ref{S:casestudy}, we demonstrate scalability and efficiency of DPC in a Demand Response problem -- DPC is used to provide financial incentives to the end-customers bypassing the need for expensive high fidelity models.
+	\item In Section \ref{S:casestudy}, we demonstrate scalability and efficiency of DPC in a Demand Response problem: DPC is used to provide financial incentives to the end-customers bypassing the need for expensive high fidelity models.
 We apply DPC to a 6 story 22 zone building model in EnergyPlus for which model-based control is not economical and practical due to extreme complexity.
@@
 We compare the total amount of energy saved with respect to the classical bang-bang controller (widely used in houses for temperature control) using an EnergyPlus model built specifically for the house.
-We show that we can perform an energy saving that ranges from $25.4\%$ (if we guarantee thermal comfort i.e.~strictly respect the desired temperature range in the rooms) to $49.2\%$ (if we allow small violation in the desired temperature).
+We show that we can perform an energy saving that ranges from $25.4\%$ (if we guarantee thermal comfort i.e.~strictly respect the desired temperature range in the rooms) to $49.2\%$ (if we allow small violation in the desired temperature range).
 Further, we test the robustness of our method due to uncertainties in the weather forecast.
@@
 		The heat exchange between two adjacent layers, i.e.
-layer ``a" and layer ``b", is modeled to be proportional to the temperature difference of the two layers and the corresponding thermal resistance $R$ given by
+layer ``a" and layer ``b", is modeled to be proportional to the temperature difference of the two layers and the corresponding thermal resistance $R$, and is given by
 		\begin{equation}\label{E:RCHeatExchange}
@@
 This filtering is needed to take into account the shadowing of the neighboring buildings.
-This approximate model is then considered ``suitable for MPC" (see Section 3.3.1.4 in \cite{SturzeneggerTR}).\\
-An EnergyPlus model is built for this system, and model parameters of matrices $A_x$, $B_q$, $A_q$, $B_{q,u}$, $B_{q,d}$, $B_{q,du,i}$, $D_{q,xu,i}$ in Equations \eqref{E:RCZoneEq} and \eqref{E:RCHeatExchange} are derived using geometry and materials data from the EnergyPlus model.
-This was a choice of the authors: real geometry and materials data can also be used to estimate the model parameters, if necessary data are available.
+This approximate model is then considered ``suitable for MPC" (see Section 3.3.1.4 in \cite{SturzeneggerTR}).}
+
+\textcolor[rgb]{0,0,1}{To identify model parameters of matrices $A_x$, $B_q$, $A_q$, $B_{q,u}$, $B_{q,d}$, $B_{q,du,i}$, $D_{q,xu,i}$ in Equations \eqref{E:RCZoneEq} and \eqref{E:RCHeatExchange}, the authors built an EnergyPlus model of the building to get geometry and materials data.
+This was a choice of the authors, but if necessary data are available, real geometry and materials data can also be used to estimate the model parameters.
 For this particular building, 24 parameters are estimated/taken form a datasheet/computed for the considered zone model.
-Although some of the parameters are in common between different zones, the others are found independently for each zone.
+Although some of the parameters are in common among different zones, the others are found independently for each zone.
 As already discussed, in \cite{SturzeneggerTR} the parameters of all the other floors are assumed to be identical to the second floor, which potentially introduces substantial modeling uncertainties.
-To derive the EnergyPlus model using the available measurements and to use the same control implemented in the building, the EnergyPlus model is coupled with MATLAB using BCVTB \cite{Wetter2015}.\textcolor[rgb]{1,0,0}{Ale: THIS PARAGRAPH IS UNCLEAR}\\
-The physics-based approach described above, although detailed and accurate (in some cases), is clearly cost and time prohibitive, and as the building characteristics changes with time the system identification must be repeated to update the model.
+To derive the EnergyPlus model using the available measurements and to use the same control implemented in the building, the EnergyPlus model is coupled with MATLAB using BCVTB \cite{Wetter2015}.}
+
+\textcolor[rgb]{0,0,1}{The physics-based approach described above, although detailed and accurate (in some cases), is clearly cost and time prohibitive, and as the building characteristics changes with time, the system identification must be repeated to update the model.
 Moreover, such expensive, time-consuming and complex modeling procedure is unique (and not repeatable) for other buildings.
-For all these reasons physics-based modeling suffers practical challenges when the objective is applying MPC to large scale buildings.
+For all these reasons, physics-based modeling suffers practical challenges when the objective is applying MPC to large scale buildings.
 }
@@
 While the inputs and external disturbances are exactly the same as discussed earlier in Section \ref{SSS:physics-based}, the outputs in \eqref{E:MLexample} include only the variables that are directly measurable through already installed sensors like thermostats and multimeters.
-Therefore, many internal states like the temperatures of different layers -- interior, middle and exterior -- of the walls, the floors and the ceilings are not required for black-box modeling which reduces the order of complexity.
+Therefore, many internal states like the temperatures of different layers -- interior, middle and exterior -- of the walls, the floors and the ceilings are not required for black-box modeling, which reduces the order of complexity.
 We use machine learning to learn such black-box models with the objective of finding the hyperparameters (for a given structure like Random Forests) of a model that best explain the input-output relationship within the measured variables, compensating the effect of the internal unmeasurable variables.
@@
 	Hence, much nonlinearities are neglected to simplify the modeling complexities.
-	We avoid such assumptions in data-driven modeling where a nonlinear function such as Neural Network or Random Forest that represents the dynamics of zone temperatures or power consumption is learned rather efficiently and accurately.
+	We avoid such assumptions in data-driven modeling where a nonlinear function such as Neural Network or Random Forest, that represents the dynamics of zone temperatures or power consumption, is learned rather efficiently and accurately.
 	For example, if \(y\) is the zone temperature in Equation \eqref{E:MLexample}, we learn a nonlinear function \(f\) which depends on current and previous inputs and disturbances. 
@@
 	\item \textcolor[rgb]{0,0,1}{
-	Physical modeling neglects different geometries and different materials of the floors assuming they are the same for each floor which is never true in reality.
+	Physical modeling neglects different geometries and different materials of the floors assuming they are the same for each floor, which is never true in reality.
 	Moreover, in many cases, the details of the construction layout and equipments are not even available so many parameters have to be guessed making physical modeling a difficult choice.
@@
 	\item \textcolor[rgb]{0,0,1}{
-	Tuning physical parameters in \eqref{E:RCZoneEq}, \eqref{E:RCHeatExchange}, and \eqref{E:RCfluxes} requires expert know-how which adds to the cost and time of modeling.
+	Tuning physical parameters in \eqref{E:RCZoneEq}, \eqref{E:RCHeatExchange}, and \eqref{E:RCfluxes} requires expert know-how, which adds to the cost and time of modeling.
 	Data-driven approach reduces both cost and time by an order of magnitude as we directly work with the senor data without explicitly modeling internal states.
-	Further, for a new building, given the historical data from the building, data-driven method is scalable as the same process can be repeated to identify a control-oriented model for MPC.}
+	Further, for a different building, given the historical data from the building, data-driven method is scalable as the same process can be repeated to identify a control-oriented model for MPC.}
 	\item \textcolor[rgb]{0,0,1}{
-	Physical modeling large number of states and variables.
+	Physical modeling has large number of states and variables.
 	Thus, many more measurements are needed to use the physical model to predict the system's behavior.
@@
 	This is exactly the focus of this paper. 
-	In the next sections, we will demonstrate how the training algorithm for Random Forests can be modified to develop control oriented models that enable Model Predictive Control, and provide better performance when compared to other approaches and robustness with respect to disturbance uncertainties.}
+	In the next sections, we will demonstrate how the training algorithm for Random Forests can be modified to develop control oriented models that enable Model Predictive Control, and provide better performance, when compared to other approaches, and robustness with respect to disturbance uncertainties.}

### 2018-01-21 paper/final/main.tex
small changes

@@
 
-\title{\textcolor[rgb]{0,0,1}{Data-Driven Model Predictive Control for Energy and Thermal Comfort Optimization in Building Energy Systems using Random Forests}}
+\title{\textcolor[rgb]{0,0,1}{Data-Driven Model Predictive Control using Random Forests for Energy and Thermal Comfort Optimization in Buildings Energy Systems using Random Forests}}
 
@@
 \begin{abstract}
-%\input{abstract.tex}
+\input{abstract.tex}
 \end{abstract}

### 2018-01-21 paper/final/Response_to_reviewers/Response_to_reviewers.tex
typo in the title corrected

@@
 \documentclass{article}
-
 \textwidth 6in%
-
 \textheight 7.9in%
-
 \oddsidemargin 1cm
@@
 \usepackage{amsmath}
-
 \usepackage{amsfonts}
-
 \usepackage{amssymb}
-
 \usepackage{graphicx}%
-
 \usepackage[dvips]{color}
-
 \setcounter{MaxMatrixCols}{30}
-
 \usepackage{natbib}
-
-\def\R{\mathcal{R}}
-
-\def\P{\mathcal{P}}
-
-\def\O{\mathcal{O}}
-
-\def\L{\mathcal{L}}
-
-\def\S{\mathcal{S}}
-
-\def\W{\mathcal{W}}
-
-\def\E{\mathcal{E}}
-
-\def\M{\mathcal{M}}
 
@@
 	\item Section 1 - INTRODUCTION
+	\textbf{
+	\begin{itemize}
+		\item The introduction has been completely re-organized and rewritten in most of its parts, with a more detailed explanation of the paper contribution;
+		\item Literature review and related work have been updated;
+	\end{itemize}}
 
+	\item Section 2 - COMPLEXITY ISSUES IN MODELING BUILDINGS: PHYSICS-BASED VS DATA-DRIVEN
+	\textbf{
 	\begin{itemize}
+		\item This section has been added to highlight the importance of our contribution, that is data-driven, with respect to model-based approaches;
+	\end{itemize}}
 
-		\item
+	\item Section 3 - DATA PREDICTIVE CONTROL
+	\textbf{
+	\begin{itemize}
+		\item We rewrote the section to make it more clear;
+		\item We provided additional details to make the mathematical model description more clear and self-contained;
+	\end{itemize}}
 
-		\item
-
-	\end{itemize}
-
-	\item Section 2 - DATA PREDICTIVE CONTROL
-
+	\item Section 6 - CASE STUDY: OPTIMAL HEATING SYSTEM SCHEDULING
+	\textbf{
 	\begin{itemize}
-
-		\item
-
-		\item
-
-		\item
-
-	\end{itemize}
-
-	\item Section 3 - COMPARISON WITH MPC
-
-	\begin{itemize}
-
-		\item
-
-	\end{itemize}
-
-	\item Section 4 - APPLICATION: DEMAND RESPONSE
-
-	\begin{itemize}
-
-		\item
-
-	\end{itemize}
-
-	\item Section 5 - CASE STUDY: OPTIMAL HEATING SYSTEM SCHEDULING
-
-	\begin{itemize}
-
-		\item
-
-	\end{itemize}
+		\item We highlighted the differences with respect to Section 5;
+		\item We added new results that account for weather forecast uncertainties and show the robustness of the proposed approach;
+	\end{itemize}}
 
 	\item Minor changes
-
+	\textbf{
 	\begin{itemize}
-
-		\item
-
-		\item
-
-		\item
-
-		\item
-
-		\item
-
-	\end{itemize}
+		\item We updated the abstract to make it more complete and clear;
+		\item We reviewed the english;
+		\item We improved the notation and corrected some mistakes;
+	\end{itemize}}
 
@@
 \textbf{Authors:} We thank the reviewer for this comment, indeed our previous presentation of the paper was unclear on this fundamental aspect.
-In our revision we widely illustrate in the Introduction how in traditional state-space models part of the state is measurable (i.e.
+In our revision we widely illustrate \textbf{in Section 2,} how in traditional state-space models part of the state is measurable (i.e.
 room temperatures, etc) and part is not measurable (i.e.
 layers/windows temperatures, etc).
-In our approach, of course, $(X^c, X^d, Y)$ are ALL measurable variables, i.e.
+In our approach, of course, $(\mathcal{X}^c, \mathcal{X}^d, \mathcal{Y})$ are ALL measurable variables, i.e.
 variables that are present in the historical data of a building.
-The unmeasurable variables of the traditional state-space models are not assumed to be measurable in this paper, yet our prediction works well since we compensate via the regression tree/random forest approaches.
+The unmeasurable variables of the traditional state-space models are not assumed to be measurable in this paper \textbf{and are neglegted in the proposed models}, yet our prediction works well since we compensate \textbf{them} via the regression tree/random forest approaches.
 
@@
 \textbf{Authors:} We agree with the reviewer.
-Following this comment, and also other comments, we re-wrote the corresponding Section to make it more clear and complete to descrive the DPC technical approach.
+Following this comment, and also other comments, we re-wrote the corresponding Section to make it more clear and complete to describe the DPC technical approach.
 In particular...

### 2018-01-21 paper/final/main.tex
typo in the title corrected

@@
 
-\title{\textcolor[rgb]{0,0,1}{Data-Driven Model Predictive Control using Random Forests for Energy and Thermal Comfort Optimization in Buildings Energy Systems using Random Forests}}
+\title{\textcolor[rgb]{0,0,1}{Data-Driven Model Predictive Control using Random Forests for Energy and Thermal Comfort Optimization in Buildings}}
 

### 2018-01-21 paper/final/Response_to_reviewers/Response_to_reviewers.tex
Sec 3 - DPC reviewed, and partially updated response to reviewers

@@
 \begin{Large}
-	New title
+	Data-Driven Model Predictive Control using Random Forests for Energy and Thermal Comfort Optimization in Buildings
 \end{Large}
@@
 	\item Section 1 - INTRODUCTION
-	\textbf{
 	\begin{itemize}
@@
 		\item Literature review and related work have been updated;
-	\end{itemize}}
+	\end{itemize}
 
 	\item Section 2 - COMPLEXITY ISSUES IN MODELING BUILDINGS: PHYSICS-BASED VS DATA-DRIVEN
-	\textbf{
 	\begin{itemize}
 		\item This section has been added to highlight the importance of our contribution, that is data-driven, with respect to model-based approaches;
-	\end{itemize}}
+	\end{itemize}
 
 	\item Section 3 - DATA PREDICTIVE CONTROL
-	\textbf{
 	\begin{itemize}
@@
 		\item We provided additional details to make the mathematical model description more clear and self-contained;
-	\end{itemize}}
+	\end{itemize}
 
 	\item Section 6 - CASE STUDY: OPTIMAL HEATING SYSTEM SCHEDULING
-	\textbf{
 	\begin{itemize}
@@
 		\item We added new results that account for weather forecast uncertainties and show the robustness of the proposed approach;
-	\end{itemize}}
+	\end{itemize}
 
 	\item Minor changes
-	\textbf{
 	\begin{itemize}
@@
 		\item We improved the notation and corrected some mistakes;
-	\end{itemize}}
+	\end{itemize}
 
@@
 \begin{enumerate}
+
+	\item \textbf{Associate Editor:} \textit{The authors should address the points and clarifications in the reviews, and also specifically address the novelty of the paper with respect to prior work.}
+
+	\textbf{Authors:}
 
@@
 In particular, we emphasise that...EXPLAIN HERE AS IN INTRODUCTION
-
-	\item \textbf{Associate Editor:} \textit{}
-
-	\textbf{Authors:} ???
 
@@
 \textbf{Authors:} We thank the reviewer for this comment, indeed our previous presentation of the paper was unclear on this fundamental aspect.
-In our revision we widely illustrate \textbf{in Section 2,} how in traditional state-space models part of the state is measurable (i.e.
+In our revision we widely illustrate in Section 2, how in traditional state-space models part of the state is measurable (i.e.
 room temperatures, etc) and part is not measurable (i.e.
@@
 variables that are present in the historical data of a building.
-The unmeasurable variables of the traditional state-space models are not assumed to be measurable in this paper \textbf{and are neglegted in the proposed models}, yet our prediction works well since we compensate \textbf{them} via the regression tree/random forest approaches.
+The unmeasurable variables of the traditional state-space models are not assumed to be measurable in this paper and are neglegted in the proposed models, yet our prediction works well since we compensate them via the regression tree/random forest approaches.
 
@@
 In the beginning, we had chosen a more general title because we believe that our approach has the potential to be applied not only to building management, but also to other application domains.
-However, as in this paper we develop our techniques focusing and experimenting in building management (which is indeed a core topic of APEN) we changed the title in "???".
+However, as in this paper we develop our techniques focusing and experimenting in building management (which is indeed a core topic of APEN) we changed the title in "Data-Driven Model Predictive Control using Random Forests for Energy and Thermal Comfort Optimization in Buildings".
 
@@
 
-\textbf{Authors:} In the introduction of the revised version we widely discuss that in this paper we improved the preliminary results obtained in [9-12] and adapted them to be applicable to the building automation use-case presented in section ???.
+\textbf{Authors:} In the introduction of the revised version, we widely discuss that in this paper we provide new results with respect to [9-12] (that now are [25,26,31,32]) and adapted them to be applicable to the building automation use-case presented in Section 6.
 
@@
 \textbf{Authors:}  We agree with the reviewer.
-Following this comment, and also other comments, we re-wrote the corresponding Section to make it more clear and complete to descrive the DPC technical approach.
-In particular...
+Following this comment, and also other comments, we completely re-wrote the corresponding Section (that now is Section 3) to make it more clear and complete to describe the DPC technical approach.
 
@@
 \textbf{Authors:} We thank the reviewer for rising this important issue.
-Yes, we assume that at time $k$ we know the future non-manipulative states (namely external temperature, radiation, etc.) for the Model Predictive Control time horizon, namely for the next 40 minutes.
-In the revised version of this paper, to answer the reviewer comment, we justify this assumption and show that, even with a large error in predicting the future states, the performance does not change significantly.
-Forests??
+Yes, we assume that at time $k$ we know the future non-manipulative states (namely external temperature, radiation, etc.) for the Model Predictive Control time horizon.
+In the revised version of this paper, to answer the reviewer comment, we account for disturbance uncertainties and show that, even with a large error in predicting the future states, the performance does not change significantly.
 More precisely, we:
-
 	\begin{itemize}
-
-		\item Cite study that shows that short term predictions are extremely reliable and say there is a trade-off between the prediction accuracy and the horizon N;
-
-		\item Provide a simulation (as for example with the bilinear model) where we inject a noise into the disturbance and show that it works as well.
-
+		\item Cite a study that shows, despite considering uncertainties on weather forecast on a long horizon of 72 hours, that MPC outperforms the rule-based controllers in providing energy savings, and is in general quite close to the perfect forecast case.
+This is due mainly to 2 main features:
+		\begin{itemize}
+			\item weather forecast are in general quite accurate on a shot-term horizon (especially on the length of horizon that we usually consider for DPC, for example 6 hours in Section 4, 7 hours in Section 5 and 40 minutes in Section 6);
+			\item MPC strategy computes an optimal input sequence over an horizon of finite length, and then apply only the first input.
+At the next step, with the new EXACT measurements of the system state and of the disturbance conditions, the algorithm is applied again.
+This is how MPC uses feedback law to correct modeling errors;
+	\end{itemize}
+		\item Provide a new case study in Section 6, where we inject a noise into the disturbance and show the robustness of our approach.
 	\end{itemize}
@@
 \item \textbf{Reviewer 2:} \textit{Also, what is new in section 5 compared with section 4?
-Even though a real house is claimed to be used, it's still tested using simulations and energyplus, like in the previous sections.
-A real world experiment on this house would be much more interesting.}
+Even though a real house is claimed to be used, it's still tested using simulations and energyplus, like in the previous sections.}
 
-\textbf{Authors:} Francesco?
+\textbf{Authors:} We thank the reviewer for this comment, indeed the contribution of this section was not well emphasized.
+We improved this aspect at the beginning of the section.
+In particular, Section 6 provides the following differences with respect to Section 5:
+\begin{itemize}
+	\item In Section 5 we use the EnergyPlus model both to create the Random Forests model and to test the DPC performance.
+	However, although the EnergyPlus model is quite accurate, it still provides simulated data, that differently from experimental data do not suffer of non-predictable imperfections such as random occupancy schedules, open/close windows, random light on/off switch, etc..
+In section 6, instead, experimental data are used to build the Random Forests models, while the EnergyPlus model (created ad-hoc for the house) is ONLY used to make a fair comparison.
+Otherwise, using Random Forests models to compare DPC (that would have used the same model) and the bang-bang control would have been unfair;
+	\item The optimization problems are different: in Section 5, we cut the peak of the power in a limited amount of time and require the temperature to be close to a desired value, while in Section 6 we minimize the energy usage while keeping temperature in a comfort range.
+Furthermore, in Section 6 we consider an on/off input, i.e.
+integer input variable, instead of a continuous one, showing that DPC can adapt to different problems;
+	\item We introduce new simulations accounting for weather forecast uncertainties, other than the ones introduced with the experimental data, that show the robustness of DPC against disturbance errors.
+\end{itemize}
+
+\item \textbf{Reviewer 2:} \textit{A real world experiment on this house would be much more interesting.}
+
+\textbf{Authors:} We thank the reviewer for this comment, indeed the contribution of this section was not well emphasized.
+We improved this aspect at the beginning of the section.
+In particular, Section 6 provides the following differences with respect to Section 5:
 

### 2018-01-21 paper/final/dpc.tex
Sec 3 - DPC reviewed, and partially updated response to reviewers

@@
 \section{DATA PREDICTIVE CONTROL}
+
 \label{S:dpc}
 
-\textcolor[rgb]{0,0,1}{The central idea behind DPC is to obtain control-oriented models using machine learning methodologies, and formulate the control problem in a way that Receding Horizon Control (RHC) can still be applied and the optimization problem can be solved efficiently.
-In this paper, Regression Trees and Random Forests are considered (see Appendix ??? for details).}
+\textcolor[rgb]{0,0,1}{The central idea behind DPC is to obtain control-oriented models using machine learning methodologies on historical datasets of buildings and formulate the optimal control problem in a way that Receding Horizon Control (RHC) can be solved efficiently.
+Let an historical dataset $(\X,\Y)$ be given.
+We define $\X = \{(x(k),u(k),d(k))\}_{k=1}^n$ the set of predictor variables (or features), i.e.
+the set of samples $(x(k),u(k),d(k))$ measured a time instants $k=1,\ldots,n$, where $x(k)\in\mathbb{R}^{n_x}$ is the vector of the \emph{measurable} state variables (e.g.
+the rooms temperatures and power consumption of the building), $u(k)\in\mathbb{R}^{n_u}$ is the vector of the \emph{measurable} input variables (e.g.
+the set-points and schedule information of the building) and $d(k)\in\mathbb{R}^{n_d}$ is the vector of the \emph{measurable and predictable} disturbance variables (e.g.
+weather historical data).
+We define $\Y = \{y(k)\}_{k=1}^{n}$ the set of response variables, i.e.
+the set of measured samples $y(k)\in\mathbb{R}^{n_y}$ representing the variables we wish to predict with our model.
+In the most general case, when we want to predict all measured variables, the response variables are represented by the state variables at the next time step, i.e.
+$y(k) = x(k+1)$.
+However, in general, we only wish to predict a subset $y(k) = \bar x(k+1) \subset \mathbb{R}^{n_x}$ of variables, e.g.
+only room temperatures and power consumption.
+Clearly, $|\X| = |\Y| = n$ is the number of time samples of the dataset.\\
+We remark that, as explained in Section \ref{secModelingIssuesExample}, the dataset $(\X,\Y)$ does not contain non-measurable states such as temperature of wall layers, ceiling, floor, etc.: nevertheless, our simulative and experimental results show that our data-driven methodology is able to compensate the absence of such variables while providing excellent prediction accuracy.}
 
-\textcolor[rgb]{0,0,1}{To this aim, let an historical dataset $(\X,\Y)$, created using building measurements, be given to be used in the Regression Trees and Random Forests learning process.
-Let $|(\X,\Y)| = n$, be the number of samples in the dataset.
-The set of predictor variables (or features) $\X = \{(x(k),u(k),d(k))\}$, $\forall k = 1,\ldots,n$, is the set of samples $(x(k),u(k),d(k))$, measured at each time instant $k$, where $x(k)\in\mathbb{R}^{n_x}$ is the vector of the \emph{measurable} state variables, $u(k)\in\mathbb{R}^{n_u}$ is the vector of the input variables and $d(k)\in\mathbb{R}^{n_d}$ is the vector of the \emph{measurable and predictable} disturbance variables.
-We recall that, as explained in Section \ref{secModelingIssuesExample}, here we do not use non-measurable states, such as temperature of wall layers, ceiling, floor, etc., since the data-driven methodology is able to compensate them.
-As an example, the variables in $\X$ can be represented by the weather forecast, as the disturbances, the set-points information and building schedules, as the inputs, and the room temperatures and power consumption, as the states.
-The set of response variables $\Y = \{y(k)\},\ \forall k = 1,\ldots,n$, is the set of samples $y(k)\in\mathbb{R}^{n_y}$.
-The response variables are represented by the state variables at the next time step, i.e. $y(k) = x(k+1)$.}
+\textcolor[rgb]{0,0,1}{Our goal is to learn data-driven models, using Regression Trees and Random Forests, that relate the value of the response variables (possibly for a certain future time horizon) with the value of the predictor variables and can be used to set up an MPC problem.
+To this aim, in the simplest example of one time-step prediction, we need to derive a model with a closed-form expression of the following form:
+\begin{equation}\label{E:GenericModel}
+	y(k)=x(k+1)=f(x(k),u(k),d(k)).
+\end{equation}
+However classical Regression Tree and Random Forest algorithms do not provide a closed-form expression for $f$, hence they cannot be used for MPC.}
 
-\textcolor[rgb]{0,0,1}{Our goal is to learn data-driven models, using Regression Trees and Random Forests, that relate the value of the response variables with the value of the predictor variables, and that can be used to set up an MPC problem.
-}
+\textcolor[rgb]{0,0,1}{In the next section we provide a new methodology (DPC) that adapts the classical Regression Tree and Random Forest algorithms to determine a closed-form expression for $f$ that is efficiently applicable to MPC.
+For simplicity of presentation we first describe DPC using Regression Trees and then DPC using Random Forests.}
 
-\textcolor[rgb]{0,0,1}{To do this, we want to derive a model with a closed-form expression of the following form:
-\begin{equation}\label{E:GenericModel}
-	y(k)=f(x(k),u(k),d(k)),
-\end{equation}
-where $f$ is an analytical expression that describes the system's dynamics.
-The issue that arises, when using Regression Trees and Random Forests, is that $f$ does not have a closed-form expression, hence \eqref{E:GenericModel} is not suitable for control and optimization.
-One possibility could be to look for sub-optimal solutions using evolutionary algorithms \cite{Kusiak2009}, i.e. using heuristics.
-In this section we provide a new methodology to determine a closed-form expression for $f$, by manipulating the dataset as shown starting from Section \ref{SS:sepvar}.}
+%==============================================================================================================
 
-\subsection{Dataset splitting}
-\label{SS:sepvar}
+\subsection{DPC-RT: DPC with Regression Trees}
 
-\textcolor[rgb]{0,0,1}{The dataset splitting consists of partitioning the features set $\X$ into the sets $\X^c = \{u(k)\}\subset\X,\ \forall k=1,\ldots,n$, containing the control (or manipulated) variables, and $\X^d = \{(x(k),d(k))\}\subset\X,\ \forall k=1,\ldots,n$, containing the disturbance and state (or non-manipulated) variables.
-The union of the two sets forms the full feature set of training, i.e. $\X \equiv \X^c \cup \X^d$.
+\label{SS:dpcrt}
+
+\textcolor[rgb]{0,0,1}{In order to have a model that can be used for prediction in an MPC problem with a future horizon of arbitrary length $N$	 we need to predict, at time $k$, the response $y$ for the next $N$ time steps, i.e.
+$y(k),\ldots,y(k+N)$.
+For the sake of simplicity and without loss of generality we consider in this section only a scalar response variable, i.e.
+$y(k)\in\mathbb{R}$ ($n_y = 1$).
+We will show in Sections \ref{S:casestudy} and \ref{S:realCaseStudy} that multiple trees (or forests) can be easily built to account for the case when $n_y > 1$.}
+
+\textcolor[rgb]{0,0,1}{Assume that, at time $k$, we want to predict the response variable at time $k+j$, i.e.
+$y(k+j),\ j\in \{1,\ldots,N\}$: when the data has lots of features interacting in complicated, nonlinear ways, assembling a single global model such as linear or polynomial regression can be difficult, and can lead to poor response predictions.}
+
+\textcolor[rgb]{0,0,1}{As discussed before an approach to non-linear regression is to partition the dataset into smaller regions where the interactions are more manageable.
+This partition can be obtained by recursively splitting the dataset via an adaptation we propose to the Regression Tree algorithm (see Appendix 1 for details), and is repeated recursively until we finally get to small chunks of the dataset (i.e.
+the leaves of the regression tree) where we can fit simple (e.g.
+linear or affine) parametric models.}
+
+%Therefore, in \eqref{E:sepvars}, the global model $f$ has two parts: the recursive partition $g$, and a linear (and convex) model $h$ for each cell of the partition.
+
+\begin{figure}[t!]
+
+	\centering
+
+	\includegraphics[width=20pc]{figures/dpc-sepvars.eps}
+
+	\caption{\textit{Step 1:} Tree $\mathcal{T}_j$ trained only with variables in $\X^d$ using adapted RT algorithm.
+Tree $\mathcal{T}'_j$ trained with variables in $\X^d$ and control variables in $\X^c$ using classical RT algorithm.
+\textit{Step 2:} In the leaf $\ell_i$ of the tree an affine model $\beta_{ij}$ is defined as a function only of the control variables dataset corresponding to leaf $\ell_i$.}
+
+	\captionsetup{justification=centering}
+
+	\label{F:dpc-sepvars}
+
+\end{figure}
+
+\textcolor[rgb]{0,0,1}{Our modification of the Regression Tree algorithm first partitions the features set $\X$ into the sets $\X^c = \{u(k)\}_{k=1}^n\subset\X$, containing the control (or manipulated) variables, and $\X^d = \{(x(k),d(k))\}_{k=1}^n\subset\X$, containing the disturbance and state (or non-manipulated) variables.
+The  union of the two non-intersecting sets forms the full feature set of training $\X \equiv \X^c \cup \X^d$.
 %Our goal is to replace a model-based controller with a data-driven controller, where the latter depends only on the historical sensor data. 
@@
 We define the number of training samples by $|(\X,\Y)| = n$.
-Using this splitting methodology, the training process is divided into two steps:
+Then the training process is divided into the following two steps, which generate as output a tree $\mathcal{T}_j$ as illustrated in Figure~\ref{F:dpc-sepvars} (left):
 \begin{enumerate}
-		\item the trees and the forests are trained only using $\X^d$, which also eases the computational complexity.
-			  It is important to note that besides external disturbance and the state measurements, $\X^d$ can also contains autoregressive terms of such variables;
-		\item in each leaf of the tree (or in each leaf of each tree of the forest), a linear regression model is trained as a function of the inputs variables in $\X^c$.
-		As we shall see in Section~\ref{SS:dpcrt} and \ref{SS:dpcrf}, this second step reduces the run-time control problem into a convex program.
+	\item the splitting of the dataset only partitions $\X^d$ (see Appendix 1 and \cite{hastie2009elements} for technical details): this choice is necessary to make our model suitable for control, as we will clarify later on, and also reduces the computational complexity.
+To each leaf $\ell_i$ will correspond an equivalence class of data samples $\X^d_i$ of the partition of $\X^d$, with $\X^d_i\subset \X^d$: as a consequence at any time $k$, given $\delta_d$ autoregressive terms of the disturbances and $\delta_x$ autoregressive terms of the state, we can associate $( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k) )$, to the corresponding leaf $\ell_i$
+	\begin{equation}\label{E:model_tree}
+		\ell_i = \mathit{f}_{\mathcal{T}_j} \left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right)
+	\end{equation}
+	if and only if $( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)) \in \X^d_i$.
+	\item in each leaf $\ell_i$ of the tree $\mathcal{T}_j$ we derive, solving a convex program over the data samples in the partition $\X^d_i$ associated to leaf $\ell_i$, an affine function, with coefficients $\beta_{ij}$, that relates the response $y(k+j)$ to the previous control inputs in $\X^c$:
+	\begin{equation}\label{E:model_leaf}
+		y(k+j) =  \beta_{ij} [1,u(k),\ldots,u(k+j) ]^T
+	\end{equation}
+	Clearly the coefficients $\beta_{ij}$ are different for each leaf $\ell_i$ as they are derived from different sets of samples.
 \end{enumerate}
-This process is illustrated in Figure~\ref{F:dpc-sepvars} (left).
-For the sake of simplicity only trees are considered, but the same holds for the forests.
-The meaning of the variable $j$ in the linear model in the figure will be clear in the next section.
-In Figure~\ref{F:dpc-sepvars} (right), the step 1 of the training process uses also the input variables, i.e. the classical training process without the dataset splitting.
-The reason why this latter approach is not suitable for control will be discussed at the end of Section \ref{SS:dpcrt}.
-}
+Applying the above procedure for $j=0,\ldots,N$ we build $N$ regression trees $\mathcal{T}_0,\ldots,\mathcal{T}_N$: thus, we have managed to linearize the original model dynamics via black-box modeling.
+Our two-step training procedure, described by the pseudo code of lines 1-11 of Algorithm~\ref{A:dpcrt}, can be computed off-line: this is an important advantage because the time required to create the model does not affect the control execution in run-time.}
 
-\subsection{DPC-RT: DPC with Regression Trees}
-\label{SS:dpcrt}
-When the data has lots of features, which interact in complicated, nonlinear ways, assembling a single global model such as linear or polynomial regression can be difficult, and can lead to poor response predictions.
-\textcolor[rgb]{0,0,1}{As discussed before, an approach to non-linear regression is to partition the data space into smaller regions, where the interactions are more manageable. }
-This partition is repeated recursively until we finally get to small chunks of the data space where we can fit simple (eg. linear parametric) models. 
-%Therefore, in \eqref{E:sepvars}, the global model $f$ has two parts: the recursive partition $g$, and a linear (and convex) model $h$ for each cell of the partition.
-
-\begin{figure}[t!]
-	\centering
-	\includegraphics[width=20pc]{figures/dpc-sepvars.eps}
-	\caption{Dataset splitting.
-\textit{Step 1:} Tree $\mathcal{T}_1$ is trained only with variables in $\X^d$ as the features.
-Tree $\mathcal{T}_2$ uses both the disturbances and states in $\X^d$, and the control variables in $\X_c$ for splitting.
-It is thus not computationally suitable for control.
-\textit{Step 2:} In the leaf $\ell_i$ of the trees, a linear regression model parametrized by $\beta_i$ is defined as a function only of the control variables.}
-	\captionsetup{justification=centering}
-	\label{F:dpc-sepvars}
-\end{figure}
-
-\textcolor[rgb]{0,0,1}{Now, in order to have a model that can be used for prediction in an MPC problem, with an horizon of arbitrary length $N$, our goal is to predict, at time $k$, the response $y$ for the next $N$ time steps, i.e.
-$y(k),\ldots,y(k+N)$, given the state measurements and the disturbance forecast at time $k$, i.e.
-$x(k)$ and $\tilde d(k),\ldots,\tilde d(k+N)$.
-For the sake of simplicity and without any loss of generality, in the following, we consider only a scalar response variable, i.e. $y(k)\in\mathbb{R}$ (or $n_y = 1$).
-However, as we will see in Sections \ref{S:casestudy} and \ref{S:realCaseStudy}, multiple trees (or forests) can be built for $n_y > 1$.}
-
-\textcolor[rgb]{0,0,1}{Applying the dataset splitting, we build $N$ regression trees such that the response $y(k+j)$ of the $j^{th}$ tree depends upon $\delta_d$ autoregressive terms of the disturbances, and on the state at instant $k$ and its $\delta_x$ autoregressive terms:}
-
-\textcolor[rgb]{0,0,1}{\begin{equation}\label{E:model_tree}
-\T_j = \mathit{f}_{\mathrm{tree}} \left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right),\ j=0,\ldots,N,
-\end{equation}}
-
-\noindent \textcolor[rgb]{0,0,1}{with $\left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right)\in\X^d,\ k=1,\ldots,n$, and $\mathit{f}_{\mathrm{tree}}$ representing the tree structure.
-Then, the linear models, as functions of the inputs variables $u\in\X^c$, in each leaf of the trees $\T_j,\ j=0,\ldots,N$, are defined as}
-
-\textcolor[rgb]{0,0,1}{\begin{equation}\label{E:model_leaf}
-y(k+j) =  \beta_j [1,u(k),\ldots,u(k+j) ]^T,\ j=0,\ldots,N.
-\end{equation}}
-
-\noindent \textcolor[rgb]{0,0,1}{Note that the coefficients $\beta_j$ are different for each leaf.
-Equation~\eqref{E:model_leaf} implies that the prediction of response $y(k+j)$ at time $k$ is an affine combination of control inputs from time $k$ to $k+j$.
-Thus, we have managed to linearize the original model dynamics via black-box modeling.
-One of the advantages of this approach is that this two-step training is done off-line, so the time required to create the models does not affect the control execution in run-time.}
-
-\textcolor[rgb]{0,0,1}{The problem now is: \emph{how to use this modeling framework to set up an MPC problem?} In run-time, given the disturbance forecast and the state measurements at time $k$, i.e.
-$\tilde d(k+j-\delta_d),\ldots,\tilde d(k+j),x(k-\delta_x),\ldots,x(k)$ given $k$, we can narrow down to a leaf of each tree in \eqref{E:model_tree} to retrieve the linear models in \eqref{E:model_leaf} for each step $j=0,\ldots,N$ of the horizon (see first part of the pseudo code given in Algorithm \ref{A:dpcrt}), as depicted in Figure \ref{F:dpc-sepvars} (left).
-These models are used to setup the MPC Problem \ref{P:dpcrt}.}
-
-\textcolor[rgb]{0,0,1}{In particular, }when a new control action has to be determined, each tree (prediction step) contributes to a linear constraint in the optimization as a replacement for the state dynamics in the case of MPC.
-Thus, the MPC optimization problem, considering a quadratic cost ($Q \succeq 0, R \succeq 0$) function and the general case of multiple responses, i.e.
-$y(k)\in\mathbb{R}^{n_y},\ n_y\geq 1$, can be formulated as:
+\textcolor[rgb]{0,0,1}{The next problem now is: \emph{how to use this modeling framework to set up an MPC problem?} We setup the MPC optimization problem in the general case of multiple responses, i.e.
+$y(k)\in\mathbb{R}^{n_y},\ n_y\geq 1$, as follows:}
 
@@
 		\begin{aligned}
-	& \underset{u_{k+j}}{\mathrm{minimize}} & & \sum_{j=0}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
-	& \mathrm{subject\ to }                 & & y_{k+j}     =   \beta_j [1,u^\top_{k},\ldots,u^\top_{k+j} ]^\top                   \\
+		& \underset{u_{k+j},\epsilon_j}{\mathrm{minimize}} & & \sum_{j=0}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
+		& \mathrm{subject\ to }                 & & y_{k+j}     =   \beta_{ij} [1,u^\top_{k},\ldots,u^\top_{k+j} ]^\top                   \\
 		&                                       & & u_{k+j}    \in  \mathcal{U}                                                        \\
@@
 
-\textcolor[rgb]{0,0,1}{Here, $Q \in \mathbb{R}^{n_y\times n_y}$ and $R \in \mathbb{R}^{n_u\times n_u}$ are weight matrices used to give more importance in the minimization to either $y$ or $u$.
-This problem is solved as in the classical MPC formulation, i.e.
-at each time step $k=1,2,\ldots$ the optimal control sequence $u^*_k,\ldots,u^*_{k+N}$ is computed, and only the first input of the sequence is applied to the system: $u(k) = u^*_k$.
-The slack variables $\epsilon_j$ are added to ensure recursive feasibility (since the equality constraint on $y$ is relaxed), i.e.
-to guarantee that Problem \ref{P:dpcrt} can provide a solution at each step $k=1,2,\ldots$.
-$\lambda$ is a weight used to give more importance to bounds violation, than to the quadratic term, and viceversa. 
-Of course, a different cost function can be chosen depending upon the application, i.e. it can be linear, nonlinear, etc., obviously changing the complexity of the problem.
-In the current formulation, the data-driven control problem is reduced to a convex program which is very easy and efficient to solve.}
-The pseudo code for DPC-RT is given in Algorithm~\ref{A:dpcrt}.
-\textcolor[rgb]{0,0,1}{DPC algorithm is also graphically shown in Figure \ref{F:dpc-algo-rf} for the random forest case, but the same holds for single regression trees.}
+\textcolor[rgb]{0,0,1}{Here, $Q \succeq 0 \in \mathbb{R}^{n_y\times n_y}$ and $R \succeq 0 \in \mathbb{R}^{n_u\times n_u}$ are weight matrices used to tradeoff the importance we provide in the minimization of $y$ versus $u$.
+The slack variables $\epsilon_j$ are added to ensure recursive feasibility: we relax the equality constraint on $y$ allowing violation up to $\epsilon_j$ to guarantee that Problem \ref{P:dpcrt} can provide a solution at each step $k \geq 0$.
+The weight $\lambda$ is then used to tradeoff the importance we provide in the optimal solution to the constraints violation versus the quadratic term.
+For example in the following sections we will use $\epsilon_j$ and $\lambda$ to define our tolerance to the violation of prescribed bounds by the rooms' temperature.\\
+Clearly different cost functions can be chosen depending upon the application, i.e.
+they can be linear, nonlinear, etc., obviously changing the complexity of the problem.
+In the current formulation, the data-driven control problem, is reduced to a convex program which is very easy and efficient to solve in C, C++, Python and Matlab, and thus is easily integrable in SCADA systems.
+Indeed Problem \ref{P:dpcrt} is solved as in the classical MPC formulation, i.e.
+at each time step $k=1,2,\ldots$ the optimal control sequence $u^*_k,\ldots,u^*_{k+N}$ is computed, and only the first input of the sequence is applied to the system: $u(k) = u^*_k$.\\
+Note that each tree $\T_j$ contributes to Problem  \ref{P:dpcrt} with the linear constraint $y_{k+j} = \beta_{ij} [1,u^\top_{k},\ldots,u^\top_{k+j} ]^\top$ as a replacement for the state dynamics in the classical MPC formulation.
+As a consequence, when solving Problem \ref{P:dpcrt} at time $k$, we need to determine the affine functions $\beta_{ij}$: it follows by Equation \eqref{E:model_tree} that, to determine $\beta_{ij}$ at time $k$, the knowledge of the state and disturbance measurements $\left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k) \right)$ is needed, however, values of $d(k+j-\delta_d),\ldots,d(k+j),\ j=0,\ldots,N$ are not available at time $k$, so we need to use disturbance forecast $\tilde d(k+N),\ldots,\tilde d(k)$.
+Using this information we can narrow down to a leaf in each tree using \eqref{E:model_tree} and thus retrieve the affine model with $\beta_{ij}$ in \eqref{E:model_leaf} for each step $j=0,\ldots,N$ of the horizon.
+The run-time solution of Problem \ref{P:dpcrt} illustrated above is described by the pseudo code of lines 12-23 of Algorithm~\ref{A:dpcrt}.}
 
 \textcolor[rgb]{0,0,1}{\begin{remark}
-	 If we did not consider the splitting procedure	of the dataset, and we also used input variables to learn the trees, as in the right side of Figure 1, the resulting
-	 model would not have been suitable for control.
-This is because, since u is the variable we want to optimize, we do not know its value a priori to go through the trees and determine the correct leaves to find the models to use in the RHC problem.
+It is now easy to understand that using the classical Regression Tree algorithm, e.g.
+using also the input variable $u$ in the data splitting procedure to create the trees as in the right side of Figure~\ref{F:dpc-sepvars}, the resulting model would not be suitable for control.
+Indeed, since $u$ is the variable we want to optimize in Problem \ref{P:dpcrt}, at time $k$ we have still not chosen its value at times $k, \ldots, k+N$: as a consequence the affine functions \emph{parameters }$\beta_{ij}$ needed to set up Problem \ref{P:dpcrt} cannot be determined at time $k$.
 \end{remark}}
 
-\begin{algorithm}[t!]
+\textcolor[rgb]{0,0,1}{The pseudo code for the whole DPC-RT procedure (i.e.
+Off-line and Run-time) is given in Algorithm~\ref{A:dpcrt}.
+Our procedure is also graphically described in Figure \ref{F:dpc-algo-rf} for the Random Forest case, providing a good intuition also for the Regression Tree case.}
+
+\textcolor[rgb]{0,0,1}{\begin{algorithm}[ht!]
 	\caption{Data Predictive Control with Regression Trees}
@@
 		\ForAll{trees $\mathcal{T}_j$}
-		\ForAll{the leaves $\ell_i$ of $\mathcal{T}_j$}
-		\State Fit $ y(k+j) =  \beta_j \left[1,u(k),\ldots,u(k+j) \right]^T$ as in \eqref{E:model_leaf}
+		\ForAll{leaves $\ell_i$ of $\mathcal{T}_j$}
+		\State Fit $ y(k+j) =  \beta_{ij} \left[1,u(k),\ldots,u(k+j) \right]^T$ as in \eqref{E:model_leaf}
 		\EndFor
@@
 		\EndFor
-		\State Solve optimization problem \ref{P:dpcrt} to determine optimal
+		\State Solve Problem \ref{P:dpcrt} to determine optimal
 		\State control actions $u^*_k,\ldots,u^*_{k+j}$
@@
 	\end{algorithmic}
-\end{algorithm}
+\end{algorithm}}
+
+%==============================================================================================================
 
 \subsection{DPC-En: DPC with Ensemble Methods}
+
 \label{SS:dpcrf}
-Regression trees obtain good predictive accuracy in many domains. However, the models used in their leaves have some limitations regarding the kind of functions they are able to approximate.
-The problem with trees is their high variance and that they can overfit the data easily.
-A small change $\delta$ in the data can result in a different series of splits and thus violate the acceptable accuracy.
-This is the price to be paid for estimating a tree-based structure from the data.
 
-We use ensemble methods \cite{Friedman2001}, in particular Random Forests, to combine the predictions of several independent regression trees in order to improve generalizability and robustness over a single estimator.
+\textcolor[rgb]{0,0,1}{Regression trees obtain good predictive accuracy in many domains.
+However, the models used in their leaves have some limitations regarding the kind of functions they are able to approximate.
+The problem with trees is their high variance and that they can overfit the data easily, and a small change in the data can result in a different series of splits and thus affect the prediction accuracy.
+This is the price to be paid for estimating a tree-based structure from the data.}
+
+\textcolor[rgb]{0,0,1}{To address the above problems we use ensemble methods \cite{Friedman2001}, in particular Random Forests, to combine the predictions of several independent regression trees in order to improve generalizability and robustness over a single estimator.
 The essential idea is to average many noisy trees to reduce the overall variance in prediction.
-We inject randomness into the tree construction in two ways. First, we randomize the features used to define splitting in each tree.
-Second, we build each tree using a bootstrapped or sub-sampled data set.
-In this way, each tree in the forest is trained on different data, which introduces differences between the trees.
-\textcolor[rgb]{0,0,1}{More explicitly, training features $\bar{\X}^d_i\subset\X^d$, $i=1,2,...$ and the in-bag samples (in-bag samples correspond to the data samples which each tree was trained on) are different for each tree in the forest, i.e $|(\bar \X^d_i,\bar\Y_i)|<n$, with $\bar\Y_i\subset\Y$.}
+We inject randomness into the tree construction in two ways: first, we randomize the features used to define splitting in each tree; second, we build each tree using a bootstrapped or sub-sampled data set.
+As a consequence each tree in the forest is trained on different data, which introduces differences between the predictive models of the trees.}
 
-\textcolor[rgb]{0,0,1}{The goal with DPC-En is to replace each tree in Algorithm~\ref{A:dpcrt} by a forest.
-This can be done replacing Equations \eqref{E:model_tree} and \eqref{E:model_leaf} with the Equations \eqref{E:model_forest} and \eqref{E:model_leaf_forest} provided in the following:
+\textcolor[rgb]{0,0,1}{More precisely, in DPC-En we replace each tree in Algorithm~\ref{A:dpcrt} by a forest $\F_j$ of $t$ trees $\T_{1j},\ldots,\T_{tj}$.
+Each tree $\T_{\kappa j}$ is trained on the basis of a different subset of features $\bar{\X}^d_{\kappa j} \subset \X^d$.
+As discussed in the previous section, at any time $k$, given $\delta_d$ autoregressive terms of the disturbances and $\delta_x$ autoregressive terms of the state $\left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k) \right)$, we can associate the corresponding leaf $\ell_i$ of $\T_{\kappa j}$
 \begin{equation}\label{E:model_forest}
-\F_j = \mathit{f}_{\mathrm{forest}} \left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right),\ j=0,\ldots,N,
+	\ell_i = \mathit{f}_{\mathcal{T}_{\kappa j}} \left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right).
 \end{equation}
-with $\left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right)\in\X^d,\ k=1,\ldots,n$, and $\mathit{f}_{\mathrm{forest}}$ representing the forest structure.
-In particular, as mentioned above, each tree $\T_i$ of each forest $\F_j$ is trained using $(\bar \X^d_i,\bar\Y_i)$.
-Then the linear models, as functions of the inputs variables $u\in\bar\X^c_i\subset\X_c$, in each leaf of each tree of the forests $\F_j,\ j=0,\ldots,N$, are defined as
+Also, in each leaf $\ell_i$ of tree $\mathcal{T}_{\kappa j}$ we derive an affine function $\Theta_{i \kappa j}$, that relates the response $y(k+j)$ to the previous control inputs in $\X^c$:
 \begin{equation}\label{E:model_leaf_forest}
-y(k+j) =  \Theta_{ij} [1,u(k),\ldots,u(k+j) ]^T,\ j=0,\ldots,N.
+	y(k+j) =  \Theta_{i \kappa j} [1,u(k),\ldots,u(k+j) ]^T.
 \end{equation}
-Here $(\bar\X^c_i,\bar\Y_i)$ correspond to the in-bag samples for the trees.}
+}
 
-While the off-line training burden in DPC-En is slightly increased compared to DPC-RT, in the control step we exploit the better accuracy, and lower variance properties of the random forests. 
-If a forest has $t$ number of trees, given the forecast of disturbances, we have $t$ sets of linear coefficients.
-We simply average out all the coefficients from all the trees to get one linear model represented by $\hat{\Theta}_j$ for each forest.
-Note that the averaging step can only be done in run-time, because the leaf of each tree can be narrowed down only when the $\X^d$ is known.
-Thus, for $N$ forests, we again have exactly $N$ linear equality constraints as in the optimization problem below:
+%In particular, as mentioned above, each tree $\T_i$ of each forest $\F_j$ is trained using $(\bar \X^d_i,\bar\Y_i)$.
 
+%Here $(\bar\X^c_i,\bar\Y_i)$ correspond to the in-bag samples for the trees.
+
+\textcolor[rgb]{0,0,1}{We can now set up the MPC problem for DPC-En as follows:}
 \begin{problem}\label{P:dpcrf}
@@
 		\begin{aligned}
-	& \underset{u_{k+j}}{\mathrm{minimize}} & & \sum_{j=0}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
-	& \mathrm{subject\ to }                 & & y_{k+j}      =  \hat{\Theta}_j [1,u_{k},\ldots,u_{k+j} ]^\top                      \\
+		& \underset{u_{k+j},\epsilon_j}{\mathrm{minimize}} & & \sum_{j=0}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
+		& \mathrm{subject\ to }                 & & y_{k+j}      =  \hat{\Theta}_{ij} [1,u_{k},\ldots,u_{k+j} ]^\top                      \\
 		&                                       & & u_{k+j}    \in  \mathcal{U}                                                        \\
@@
 \end{problem}
-DPC-En is graphically described in Figure~\ref{F:dpc-algo-rf}.
-
-The ensemble data predictive control (DPC-En) is the first such method to bridge the gap between ensemble predictive models (such as random forests) and receding horizon control.
-In the next section, we compare DPC-RT and DPC-En to MPC for a building model.
+\textcolor[rgb]{0,0,1}{Note that, as in the DPC-RT, the crucial problem when solving Problem \ref{P:dpcrf} is determining all parameters $\hat{\Theta}_{ij}$ at time $k$.
+We first proceed similarly to DPC-RT by using forecast of disturbances and the state measurements to determine for each tree $\T_{\kappa j}$ the corresponding leaf $\ell_i$ and thus the affine model $\Theta_{i \kappa j}$, with $\kappa = 1,\ldots,t$ and $j = 0,\ldots,N$.
+In DPC-En, however, we have $t$ affine models for each prediction step $j$, thus we define $\hat{\Theta}_{ij} = \frac{1}{t}\sum\limits_{\kappa = 1}^{t} \Theta_{i \kappa j}$ averaging the affine models associated to all trees of forest $\F_j$.
+Once we are able to determine at each time $k$ the parameters $\hat{\Theta}_{ij}$ of the constraints of Problem \ref{P:dpcrf}, the optimal input $u^*_k$ can be computed as described in the previous section.
+The overall procedure is sketched in Figure \ref{F:dpc-algo-rf}.}
 
@@
 	\includegraphics[width=26pc]{figures/dpc-algo-rf.eps}
-	\caption{\textcolor[rgb]{0,0,1}{DPC-En: At time $k$, the algorithm uses the forecast of disturbances and the state measurements in $\X^d$ given at time $k$ (that with a small abuse of notation we indicate only in this figure with $\X^d_k$), to select linear models $\Theta_{l1}$ to $\Theta_{lt}$ in each leaf $l$ of the $t$ trees of each forest.
-The linear models in each forest are averaged to calculate single models represented by $\hat{\Theta}_j$, and act as constraints in the optimization problem.
-The optimal sequence $u^*_k,\ldots,u^*_{k+N}$, of which only the first element is applied, i.e.
-$u(k)=u^*_k$, and variables in $\X^d_{k+1}$ are calculated to proceed to next step at $k+1$ and repeat the process.}}
+	\caption{\textcolor[rgb]{0,0,1}{Graphical description of the procedure for solving Problem \ref{P:dpcrf} at each time $k$}}
 	\label{F:dpc-algo-rf}
 \end{figure}
+
+\textcolor[rgb]{0,0,1}{The ensemble data predictive control (DPC-En) is the first such method to bridge the gap between ensemble predictive models (such as random forests) and receding horizon control.
+Reasoning on complexity, we remark that the off-line training computation in DPC-En is increased compared to DPC-RT as we need to train $t$ trees for each of the $N$ prediction steps.
+The run-time computation is also slightly increased as we need to derive the average $\hat{\Theta}_{ij} = \frac{1}{t}\sum\limits_{\kappa = 1}^{t} \Theta_{i \kappa j}$ for each of the $N$ prediction steps.
+However, as shown in the following sections, it's worth the price as we obtain much better accuracy and lower variance properties with a marginal increase of the run-time computation time (the increase of off-line computation time is not a big issue as it must be run very seldom, i.e.
+only when a new model of the building needs to be created).}

### 2018-01-21 paper/final/dpc.tex
small changes

@@
 		& \underset{u_{k+j},\epsilon_j}{\mathrm{minimize}} & & \sum_{j=0}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
-		& \mathrm{subject\ to }                 & & y_{k+j}      =  \hat{\Theta}_{ij} [1,u_{k},\ldots,u_{k+j} ]^\top                      \\
+		& \mathrm{subject\ to }                 & & y_{k+j}      =  \hat{\Theta}_{j} [1,u_{k},\ldots,u_{k+j} ]^\top                      \\
 		&                                       & & u_{k+j}    \in  \mathcal{U}                                                        \\
@@
 \end{problem}
-\textcolor[rgb]{0,0,1}{Note that, as in the DPC-RT, the crucial problem when solving Problem \ref{P:dpcrf} is determining all parameters $\hat{\Theta}_{ij}$ at time $k$.
+\textcolor[rgb]{0,0,1}{Note that, as in the DPC-RT, the crucial problem when solving Problem \ref{P:dpcrf} is determining all parameters $\hat{\Theta}_{j}$ at time $k$.
 We first proceed similarly to DPC-RT by using forecast of disturbances and the state measurements to determine for each tree $\T_{\kappa j}$ the corresponding leaf $\ell_i$ and thus the affine model $\Theta_{i \kappa j}$, with $\kappa = 1,\ldots,t$ and $j = 0,\ldots,N$.
-In DPC-En, however, we have $t$ affine models for each prediction step $j$, thus we define $\hat{\Theta}_{ij} = \frac{1}{t}\sum\limits_{\kappa = 1}^{t} \Theta_{i \kappa j}$ averaging the affine models associated to all trees of forest $\F_j$.
-Once we are able to determine at each time $k$ the parameters $\hat{\Theta}_{ij}$ of the constraints of Problem \ref{P:dpcrf}, the optimal input $u^*_k$ can be computed as described in the previous section.
+In DPC-En, however, we have $t$ affine models for each prediction step $j$, thus we define $\hat{\Theta}_{j} = \frac{1}{t}\sum\limits_{\kappa = 1}^{t} \Theta_{i \kappa j}$ averaging the affine models associated to all trees of forest $\F_j$.
+Once we are able to determine at each time $k$ the parameters $\hat{\Theta}_{j}$ of the constraints of Problem \ref{P:dpcrf}, the optimal input $u^*_k$ can be computed as described in the previous section.
 The overall procedure is sketched in Figure \ref{F:dpc-algo-rf}.}
@@
 Reasoning on complexity, we remark that the off-line training computation in DPC-En is increased compared to DPC-RT as we need to train $t$ trees for each of the $N$ prediction steps.
-The run-time computation is also slightly increased as we need to derive the average $\hat{\Theta}_{ij} = \frac{1}{t}\sum\limits_{\kappa = 1}^{t} \Theta_{i \kappa j}$ for each of the $N$ prediction steps.
+The run-time computation is also slightly increased as we need to derive the average $\hat{\Theta}_{j} = \frac{1}{t}\sum\limits_{\kappa = 1}^{t} \Theta_{i \kappa j}$ for each of the $N$ prediction steps.
 However, as shown in the following sections, it's worth the price as we obtain much better accuracy and lower variance properties with a marginal increase of the run-time computation time (the increase of off-line computation time is not a big issue as it must be run very seldom, i.e.

### 2018-01-21 paper/final/Response_to_reviewers/Response_to_reviewers.tex
first draft response to the reviewers

@@
 
-	\textbf{Authors:}
+	\textbf{Authors:} We addressed all the reviewers' points as discussed in the following of this document.
+We also emphasized the contribution of the paper with respect to our previous works in "Related work" subsection of the introduction.
 
@@
 
-	\textbf{Authors:} We performed a deep literature review. In particular, in Table 1 we classify the related work and then emphasise the novelty of our research.
+	\textbf{Authors:} We performed a deep literature review. In particular, in Table 1 we classify the related work and then emphasize the novelty of our research.
 
@@
 	\textbf{Authors:} Following the Associate Editor and reviewers comments, we re-wrote the title and the introduction of the paper focusing on the relevance of the paper to APEN.
-In particular, we emphasise that...EXPLAIN HERE AS IN INTRODUCTION
+\textbf{FOR RAHUL: should we say here that this paper was invited specifically for our methodology?}
 
@@
 
-	\textbf{Authors:} We clarify in the introduction and in Section ???
-the originality of the paper and how are results can be used for real applications.
-In particular, we emphasise that...EXPLAIN HERE AS IN INTRODUCTION
+	\textbf{Authors:} In the "Related work" and "Main contribution" subsections of the Introduction, we clarified the originality of the paper to existing results and emphasized the contributions provided.
+In Sections 2 and 6 we showed how our results can be used for real applications.
+More precisely, through a practical case study (in Section 2), we highlighted the typical drawbacks of the model-based approaches, which limit the use of advanced control optimization technique for energy optimization, such as MPC, and showed (in Sections 2 and 6) how these limitation are not a problem in the data-driven MPC approach that we propose.
+In particular, in Section 6, we used experimental data from a real house and considered noisy weather forecasts, to show that our approach is robust with respect to all sort of imperfections typical of the real world, i.e.
+error in weather predictions, unpredictable house schedules (due to people, appliancies, etc.), noisy sensor measurements (due to real experimental data), etc..
+This shows that our approach can be used for real implementation.
 
@@
 
-\textbf{Authors:}
-
-	\begin{itemize}
-
-		\item Explain in the introduction (other than in section 2) the difference between the state-space modeling and the one we propose;
-
-		\item Add a subsection in the "Related work" to explain the evolution of our work with respect to our previous ones;
-
-		\item Add Appendix on regression trees and random forests;
-
-		\item Add equations in the intro.
-In particular, add a section where we quantify the complexity of a physics-based state-space modeling through an example.
-Setup and write equation to show the complexity
-
-		\item add another section to show how the previous point simplifies using data-driven methods
-
-	\end{itemize}
+\textbf{Authors:} We thank the reviewer for this comment, indeed this very important part of the paper was not clear.
+We added a section (Section 2) where we reported a traditional state-space modeling of a building (Section 2.1), and compared with the data-driven modeling approach (Section 2.2).
+We highlighted the differences in the modeling techniques, and emphasized that the features needed in the data-driven approaches are all the measurable variables (hence a subset of all the variables) used in the model-based approach, since the model learning procedure is able to compensate the internal effect of the non-measurable ones.
+This is also further clarified in Section 3, where we added more mathematical details.
 
@@
 It would be much better if the authors can explain their framework using Eq.
-(1) directly first instead of using another set of notations such as $(X^c, X^d, Y)$.}
-
-\textbf{Authors:} We agree with the reviewer.
-Following this comment, and also other comments, we re-wrote the corresponding Section to make it more clear and complete to describe the DPC technical approach.
-In particular...
-
-\item \textbf{Reviewer 1:} \textit{The relationship between Eq.
+(1) directly first instead of using another set of notations such as $(X^c, X^d, Y)$.
+The relationship between Eq.
 (1) and Fig.
@@
 \textbf{Authors:} We agree with the reviewer.
-Following this comment, and also other comments, we re-wrote the corresponding Section to make it more clear and complete to descrive the DPC technical approach.
-In particular...
+Following this comment, and also other comments, we re-wrote the corresponding Section to make it more clear and complete to describe the DPC technical approach.
+In particular, we removed Eq.
+1 (of the old paper), since it was confusing, and explain our goal in the beginning of Section 3.
+Then, in Section 3.1, we introduced the two-steps process that modifies the Regression Tree learning algorithm, to obtain models that are suitable for MPC.
+We clarified the relationship between the two-steps process and Fig.
+1 (left), and we added a remark that explain the meaning of Fig.
+1 (right).
 
@@
 
-\textbf{Authors:} We thank the reviewer for this comment, indeed the contribution of this section was not well emphasized.
-We improved this aspect at the beginning of the section.
-In particular, Section 6 provides the following differences with respect to Section 5:
+\textbf{Authors:} We agree with the reviewer that a real world experiment would be interesting.
+However, it would not allow to compare different approaches, since they could not be ran simultaneously, hence making the comparison unfair.
+We are currently working to apply our methodology on a real university building, but this is out of the scope of this paper.
 
@@
 
-%\bibliographystyle{plain}
-
-%\bibliography{mcnbib}
-
 \end{document}

### 2018-01-22 paper/final/Response_to_reviewers/Response_to_reviewers.tex
prefinal edits

@@
 \begin{Large}
-	Data-Driven Model Predictive Control using Random Forests for Energy and Thermal Comfort Optimization in Buildings
+	Data-Driven Model Predictive Control using Random Forests for Building Energy Optimization and Climate Control
 \end{Large}
@@
 
-\textbf{Authors:} Achin?
+\textbf{Authors:} We have clarified this point in the text and the Figure caption.
+Shown in yellow is the tracking signal, in blue is the response when the optimal input is applied to the power Random Forest and in red is the response when the same input is applied to the plant (EnergyPlus building).
+The difference in closed-loop simulation and prediction is due to model mismatch.
 

### 2018-01-22 paper/final/abstract.tex
prefinal edits

--- 2017-09-30 paper/abstract.tex
@@
 A key factor prohibiting the widespread adoption of MPC for complex systems is related to the difficulties (cost, time, and effort) associated with the model identification. 
-On the other hand, data-driven models are not suitable for control because of non-physical models.
-%Due to their lack of dynamical equations to predict real systems behavior, data-driven models are not suitable for control. 
-We introduce a novel idea for predictive control based on system data, Data Predictive Control (DPC), leveraging machine learning algorithms, in particular regression trees and random forests.
-Using a bilinear model, we show that DPC has comparable performance with MPC.
-We further apply DPC to the Demand Response problem of a large scale multi-story EnergyPlus building building model, and show that DPC curtails the desired power usage with high confidence.
-Finally, we use DPC to optimize ON/OFF scheduling of the heating system of a real house in order to minimize the power consumption while guaranteeing thermal comfort for occupants.
+\textcolor[rgb]{0,0,1}{To overcome this problem, we introduce a novel idea for predictive control based on historical building data leveraging machine learning algorithms like regression trees and random forests.
+We call this approach Data-driven model Predictive Control (DPC).
+Using a physics-based bilinear model of a real building, we show that DPC provides comparable performance with MPC.
+We further apply DPC to a large scale multi-story EnergyPlus building model in the problem of Demand Response, and show that DPC curtails the desired power usage with high confidence.}
+Further, we use DPC to optimize ON/OFF scheduling of the heating system of a real house in order to minimize the power consumption while guaranteeing thermal comfort for the occupants.
 Our results show that with DPC we obtain an energy saving in terms of primary energy up to $49.2\%$ when compared to a bang-bang controller.
+\textcolor[rgb]{0,0,1}{Finally, we show the robustness of DPC under disturbance uncertainties.}

### 2018-01-22 paper/final/casestudyEPlus.tex
prefinal edits

@@
 	}
-	\subfigure[Power tracking by DPC-En at 1.1 MW: The difference in closed-loop simulation and prediction is due to model mismatch.]{
+	\subfigure[Power tracking by DPC-En at 1.1 MW.
+In blue is the response when the optimal input is applied to the power Random Forest and in red is the response when the same input is applied to the plant (EnergyPlus building).
+The difference in closed-loop simulation and prediction is due to model mismatch.]{
 		\label{F:eplus_power}
@@
 In fact, the average tracking error between 7am - 2pm is 3\%.
-The difference between the predicted power consumption and that in the closed-loop simulation in Figure~\ref{F:eplus_power} is due to model mismatch between the EnergyPlus model and the power forest used in the optimization \eqref{E:ptrack}.
+\textcolor[rgb]{0,0,1}{The results from the closed-loop simulation are shown in Figure~\ref{F:eplus_power}.
+Marked in blue is the response when the optimal input is applied to the power Random Forest and in red is the response when the same input is applied to the plant (EnergyPlus building).
+The difference between the predicted power consumption and that in the closed-loop simulation is due to model mismatch between the EnergyPlus model and the power forest used in the optimization \eqref{E:ptrack}.}
 Due to this inaccuracy, the actual power consumption is on an average 7 kW higher.

### 2018-01-22 paper/final/casestudyReal.tex
prefinal edits

@@
 In Section \ref{SS:controllersDPCandBangbang}, we set up the DPC optimization problem and describe the bang-bang control strategy.
-Finally, in Section \ref{SS:simulationResults}, the performance of the two controllers are compared in terms of energy saving using the EnergyPlus model.
-\textcolor[rgb]{0,0,1}{In particular, in Section \ref{SSS:DisturbancePerfect}, we show that DPC allows an energy saving up to $49.2\%$ with respect to the bang-bang controller while guaranteeing thermal comfort for occupants, considering the perfect knowledge of the weather forecast.
+Finally, in Section \ref{SS:simulationResults}, the performance of the two controllers are compared in terms of energy savings using the EnergyPlus model.
+\textcolor[rgb]{0,0,1}{In particular, in Section \ref{SSS:DisturbancePerfect}, we show that DPC provides up to $49.2\%$ energy savings with respect to the bang-bang controller while guaranteeing thermal comfort for the occupants, considering the perfect knowledge of the weather forecast.
 In Section \ref{SSS:DisturbanceUncertain}, we show that DPC is robust with respect to imperfect weather forecast.}
@@
 
-\textcolor[rgb]{0,0,1}{The contribution of this section with respect to Sections \ref{S:casestudy} is threefold:
+\textcolor[rgb]{0,0,1}{The contribution of this section is threefold.
 \begin{enumerate}
-	\item we consider experimental data from a lived house, hence subject to real world imperfections, such as random (non predictable) occupancy schedules, open/close windows, random (non predictable) light on/off switch, etc..
-	On the other hand, the EnergyPlus model considered in Section \ref{S:casestudy}, although provides an accurate model of a building, is still a simulated model that does not take into account the aforementioned imperfections;
-	\item we modify the DPC algorithm to take into account uncertainties on the weather forecast, so making DPC robust with respect to prediction errors;
-	\item in Section \ref{S:casestudy}, we want to cut the peak of the power in a specific time range while keeping the temperature as close as possible to a specific setpoint.
-	Differently, in this section, we show the adaptability of DPC to different problems, minimizing the energy consumption required to keep room temperatures within a specified range of comfort for an arbitrary amount of time.
-In this way we show the potential of DPC for energy saving on a long period.
+	\item We consider experimental data from an occupied house, hence subject to real world imperfections, such as random (non predictable) occupancy schedules, open/close windows, random (non predictable) light on/off switch, etc.
+	This is different from Section \ref{S:casestudy} where we used data simulated from an EnergyPlus model.
+	\item We modify the DPC algorithm to take into account uncertainties In the weather forecast, making DPC robust with respect to prediction errors.
+	\item While in Section \ref{S:casestudy} we follow a reference DR signal while maintaining thermal comfort, in this section we show the adaptability of DPC to different problems where we minimize the energy consumption required to keep the room temperatures within a specified range of comfort.
+In this way, we show the potential of DPC in providing energy savings for a long period.
 \end{enumerate}}
@@
 In particular, we considered a small ($Q=100$, $\lambda_{min}=3000$ and $\lambda_{max}=100$), a medium ($Q=100$, $\lambda_{min}=1000$ and $\lambda_{max}=100$) and a large ($Q=100$, $\lambda_{min}=100$ and $\lambda_{max}=100$) violation configuration.
-\textcolor[rgb]{0,0,1}{The simulation period is of $15$ days, from May $1$, $2016$ at $00am$ to May $15$, $2016$ $00am$.}
+\textcolor[rgb]{0,0,1}{The simulation period is of $15$ days, from $00$ hrs on May $1$, $2016$ to $00$ hrs on May $15$, $2016$.}
 
-\textcolor[rgb]{0,0,1}{We considered 2 different simulative conditions to show the robustness of our approach:
+\textcolor[rgb]{0,0,1}{We considered two different simulation conditions to show the robustness of our approach.
 \begin{enumerate}
-	\item in Section \ref{SSS:DisturbancePerfect} we provide our simulative results considering perfect knowledge of the weather forecast;
-	\item in Section \ref{SSS:DisturbanceUncertain}, we consider weather forecast subject to uncertainty Hence we modify Problem \ref{P:dpcRealCase} by adding a gaussian noise with high variance on the perfect forecast, and show that the results are close to the perfect forecast case, so making our methodology robust with respect to disturbance uncertainties.
+	\item In Section \ref{SSS:DisturbancePerfect}, we provide the results considering perfect knowledge of the weather forecast.
+	\item In Section \ref{SSS:DisturbanceUncertain}, we consider weather forecast subject to uncertainties.
+Hence, we modify Problem \ref{P:dpcRealCase} by adding a Gaussian noise with high variance to the perfect forecast, and show that the results are close to the perfect forecast case.
+	This verifies our methodology is robust with respect to disturbance uncertainties.
 \end{enumerate}
 \subsubsection{Perfect knowledge of the weather forecast}\label{SSS:DisturbancePerfect}
-In this section we ran the simulations considering perfect knowledge of the disturbance over the horizon, obtaining the following results.}
+In this section, we ran the simulations considering perfect knowledge of the disturbance over the horizon, obtaining the following results.}
 \paragraph{Result 1} The comparison for temperature and input schedule obtained allowing small bound violations in DPC is shown in Figure \ref{F:comparison_small}.
-For sake of the plot's clarity, the shown period is restricted to $4$ days and a half, from May $1$, $2016$ $00am$ to May $5$, $2016$ $1pm$.
-\textcolor[rgb]{0,0,1}{The whole period will be used in "Result 2" to provide the bounds violation errors, and in "Result 3" for the energy consumption comparison.}
+For sake of the plot's clarity, the shown period is restricted to $4$ days and a half, from $00$ hrs on May $1$, $2016$ to  $13$ hrs on May $5$, $2016$.
+\textcolor[rgb]{0,0,1}{The whole period will be used in ``Result 2" to provide the bounds violation errors, and in ``Result 3" for the energy consumption comparison.}
 We can see that the temperature controlled with DPC does not violate the bounds and if it does then the violation is approximately $0.1\,\degree C$.

### 2018-01-22 paper/final/conclusion.tex
prefinal edits

@@
 To overcome the difficulties associated with the model identification in Model Predictive Control (MPC), we introduce a novel idea for predictive control using data: Data-driven model Predictive Control (DPC).
-Data-based control strategies are usually based on non-physical models, therefore the can not be integrated with most of the classical control approaches.
-The goal is to create data-driven models that are suitable for receiding horizon control.
-To this aim, we present two algorithms, based on trees and random forests, to create data-driven models to be used in the DPC strategy.
-We then apply DPC to $3$ different case studies in order to prove its strength.
+\textcolor[rgb]{0,0,1}{Data-driven control is based on non-physical (black-box) models, therefore they can not be integrated with most of the classical control approaches.} The goal is to create data-driven models that are suitable for receiding horizon control.
+\textcolor[rgb]{0,0,1}{To this aim, we present two algorithms, based on trees and random forests, to create control-oriented models for DPC.
+We then apply DPC to three different case studies to demonstrate its strength.}
 \begin{enumerate}
@@
 Our results demonstrate that even for such complex system, DPC tracks a reference signal with a mean error of $3\%$.
-\item \emph{\textbf{\textcolor[rgb]{0,0,1}{Optimal heating system scheduling of a real house.}} We finally apply DPC using experimental data from a real house and compare it with the classical bang-bang controller widely used for temperature control in houses.
-We implemented DPC to guarantee energy saving while guaranteeing thermal comfort for the occupants.
-With respect to a bang-bang controller, we obtained an energy saving that goes from $25.4\%$, when we force the temperature to be strictly in a comfort range, to $49.2\%$, if we allow small violations of such range.
-Furthermore, we also demonstrate the robustness of DPC considering uncertainties on weather forecast.}
+\item \textcolor[rgb]{0,0,1}{\emph{\textbf{Optimal heating system scheduling of a real house.}} We finally apply DPC using experimental data from a real house and compare it with the classical bang-bang controller widely used for temperature control.
+We show that DPC provides energy savings while guaranteeing thermal comfort for the occupants.
+With respect to a bang-bang controller, we obtain energy savings that go from $25.4\%$, when we force the temperature to be strictly in a comfort range, to $49.2\%$, if we allow small violations in the thermal comfort level.
+Furthermore, we demonstrate the robustness of DPC considering uncertainties in the weather forecast.}
 \end{enumerate}

### 2018-01-22 paper/final/cs-challenges.tex
prefinal edits

@@
 \begin{enumerate}
-\item \emph{\textbf{Data Availability:}} The main practical challenge for DPC lies in the availability of data for training, and we require answers to questions like \emph{how much data (functional testing) is required, and how should the sampling be done?} Therefore, the procedure for optimal experiment design, and model improvement with estimation of variance in predictions is one of the main focus of our ongoing work.
+\item \emph{\textbf{Data Availability:}} \textcolor[rgb]{0,0,1}{The main practical challenge for DPC lies in the availability of data for training, and we require answers to questions like \emph{how much data is required, and how should the sampling (functional testing) be done?} Therefore, the procedure for optimal experiment design, and model improvement with estimation of variance in predictions is one of the main focus of our ongoing work \cite{JainICCPS2018}}.
 \item \emph{\textbf{Stability:}} While the buildings are inherently stable, many other applications require stability guarantees.

### 2018-01-22 paper/final/dpc.tex
small changes
prefinal edits

@@
 
-\textcolor[rgb]{0,0,1}{The central idea behind DPC is to obtain control-oriented models using machine learning methodologies on historical datasets of buildings and formulate the optimal control problem in a way that Receding Horizon Control (RHC) can be solved efficiently.
+\textcolor[rgb]{0,0,1}{The central idea behind DPC is to obtain control-oriented models using machine learning on historical datasets of buildings and formulate the optimal control problem in a way that Receding Horizon Control (RHC) can be solved efficiently.
 Let an historical dataset $(\X,\Y)$ be given.
@@
 Clearly, $|\X| = |\Y| = n$ is the number of time samples of the dataset.\\
-We remark that, as explained in Section \ref{secModelingIssuesExample}, the dataset $(\X,\Y)$ does not contain non-measurable states such as temperature of wall layers, ceiling, floor, etc.: nevertheless, our simulative and experimental results show that our data-driven methodology is able to compensate the absence of such variables while providing excellent prediction accuracy.}
+We remark that, as explained in Section \ref{secModelingIssuesExample}, the dataset $(\X,\Y)$ does not contain non-measurable states such as temperature of wall layers, ceiling, floor, etc.: nevertheless, our simulated and experimental results show that our data-driven methodology is able to compensate the absence of such variables while providing excellent prediction accuracy.}
 
 \textcolor[rgb]{0,0,1}{Our goal is to learn data-driven models, using Regression Trees and Random Forests, that relate the value of the response variables (possibly for a certain future time horizon) with the value of the predictor variables and can be used to set up an MPC problem.
-To this aim, in the simplest example of one time-step prediction, we need to derive a model with a closed-form expression of the following form:
+To this aim, in the simplest example of one time-step prediction, we need to derive a model with a closed-form expression of the following form
 \begin{equation}\label{E:GenericModel}
@@
 		\State Set $\X^d$ $\gets$ non-manipulated features
-		\State Build $N$ predictive trees with $(\X^d,\Y)$ as defined in \eqref{E:model_tree}
+		\State Build $N$ predictive trees with $(\X^d,\Y)$ using Regression Trees algorithm
 		\ForAll{trees $\mathcal{T}_j$}
 		\ForAll{leaves $\ell_i$ of $\mathcal{T}_j$}
-		\State Fit $ y(k+j) =  \beta_{ij} \left[1,u(k),\ldots,u(k+j) \right]^T$ as in \eqref{E:model_leaf}
+		\State Compute parameters $\beta_{ij}$ in \eqref{E:model_leaf} using convex programming
 		\EndFor

### 2018-01-22 paper/final/main.tex
prefinal edits

@@
 
-\title{\textcolor[rgb]{0,0,1}{Data-Driven Model Predictive Control using Random Forests for Energy and Thermal Comfort Optimization in Buildings}}
+\title{\textcolor[rgb]{0,0,1}{Data-Driven Model Predictive Control using Random Forests for Building Energy Optimization and Climate Control}}
 
@@
 \begin{keyword}
-Energy efficiency, demand response, predictive control, machine learning, EnergyPlus modeling, real application
+Demand response, energy efficiency, predictive control, machine learning, EnergyPlus modeling, real application
 \end{keyword}

### 2018-01-22 paper/final/abstract.tex
reviewed abstract

@@
-Model-based control techniques have been widely used over the past years in control systems to improve system performance of simple controllers, such as PID, bang-bang, etc. 
-Among these, one of the most popular is Model Predictive Control (MPC) due to its ability to predict the system's behavior over a future horizon and handle constraints in the optimal control setup.
-A key factor prohibiting the widespread adoption of MPC for complex systems is related to the difficulties (cost, time, and effort) associated with the model identification. 
-\textcolor[rgb]{0,0,1}{To overcome this problem, we introduce a novel idea for predictive control based on historical building data leveraging machine learning algorithms like regression trees and random forests.
-We call this approach Data-driven model Predictive Control (DPC).
-Using a physics-based bilinear model of a real building, we show that DPC provides comparable performance with MPC.
-We further apply DPC to a large scale multi-story EnergyPlus building model in the problem of Demand Response, and show that DPC curtails the desired power usage with high confidence.}
-Further, we use DPC to optimize ON/OFF scheduling of the heating system of a real house in order to minimize the power consumption while guaranteeing thermal comfort for the occupants.
-Our results show that with DPC we obtain an energy saving in terms of primary energy up to $49.2\%$ when compared to a bang-bang controller.
-\textcolor[rgb]{0,0,1}{Finally, we show the robustness of DPC under disturbance uncertainties.}
+\textcolor[rgb]{0,0,1}{Model Predictive Control (MPC) is a model-based technique widely and successfully used over the past years to improve control systems performance.
+A key factor prohibiting the widespread adoption of MPC for complex systems such as buildings is related to the difficulties (cost, time and effort) associated with the identification of a predictive model of a building.\\
+To overcome this problem, we introduce a novel idea for predictive control based on historical building data leveraging machine learning algorithms like regression trees and random forests.
+We call this approach Data-driven model Predictive Control (DPC), and we apply it to three different case studies to demonstrate its \textit{performance}, \textit{scalability} and \textit{robustness}.\\
+In the first case study we consider a benchmark MPC controller using a bilinear building model, then we apply DPC to a data-set simulated from such bilinear model and derive a controller based only on the data.
+Our results demonstrate that DPC can provide comparable performance with respect to MPC applied to a perfectly known mathematical model.\\
+In the second case study we apply DPC to a 6 story 22 zone building model in EnergyPlus, for which model-based control is not economical and practical due to extreme complexity, and address a Demand Response problem.
+Our results demonstrate scalability and efficiency of DPC showing that DPC provides the desired power curtailment with an average error of 3\%.\\
+In the third case study  we implement and test DPC on real data from an off-grid house located in L'Aquila, Italy.
+%We design the optimal ON/OFF scheduling for the heating system in order to save energy while guaranteeing thermal comfort for the occupants.
+We compare the total amount of energy saved with respect to the classical bang-bang controller, showing that we can perform an energy saving up
+%that ranges from $25.4\%$ (if we guarantee thermal comfort i.e.~strictly respect the desired temperature range in the rooms)
+to $49.2\%$.
+%(if we allow small violation in the desired temperature range).
+Our results demonstrate robustness of our method to uncertainties both in real data acquisition and weather forecast.}

### 2018-01-22 paper/final/appendixRegTrees.tex
Sec 3 - DPC reviewed, and partially updated response to reviewers
prefinal edits
reviewed Intro, Sec 2,3,4,5, conclusion and appendix

@@
 Tree-based methods partition the feature space into a set of rectangles (more formally, hyper-rectangles) and then fit a simple model in each one. They are conceptually simple yet powerful.
-Let us consider a regression problem with continuous response $\Y = \{Y\}$ and and 2 predictors $\X = \{X_1,X_2\}$, each taking values in the unit interval.
+Let us consider a regression problem with continuous response $\Y = \{Y\}$ and 2 predictors $\X = \{X_1,X_2\}$, each taking values in the unit interval.
 The top left plot of Figure~\ref{fig:friedman} shows a partition of the feature space by lines that are parallel to the coordinate axes.
@@
   \centering
-  \includegraphics[width=0.7\columnwidth]{figures/friedman1.eps}
+  \includegraphics[width=0.7\columnwidth]{figures/friedman.eps}
   \vspace{-8pt}
@@
 \begin{equation}
-\T = \mathit{f}_{\mathrm{tree}}(X_1,X_2) = \sum_{i=1}^{5} c_i I\left\lbrace(X_1,X_2) \in R_i\right\rbrace,
+\mathit{f}_{\T}(X_1,X_2) = \sum_{i=1}^{5} c_i I\left\lbrace(X_1,X_2) \in R_i\right\rbrace,
 \label{eq:response}
@@
 \begin{equation}
-\sum_i (y_i - \T(x_i))^2.
+\sum_i (y_i - \mathit{f}_{\T}(x_i))^2.
 \end{equation}
@@
 However, this strategy is myopic, since a seemingly worthless split might lead to a very good split below it.
-A preferred, strategy is to grow a large tree, stopping the splitting process only when some minimum number of data points at a node (\texttt{MinLeaf}) is reached.
+A preferred strategy is to grow a large tree, stopping the splitting process only when some minimum number of data points at a node (\texttt{MinLeaf}) is reached.
 Then this large tree is pruned using cost-complexity pruning methods.}

### 2018-01-22 paper/final/casestudyEPlus.tex
reviewed Intro, Sec 2,3,4,5, conclusion and appendix

@@
-\section{APPLICATION: DEMAND RESPONSE}
+\section{APPLICATION TO DEMAND RESPONSE}
 \label{S:casestudy}
@@
 DPC builds predictive models of a building based on historical weather, schedule, set-points and electricity consumption data, while also learning from the actions of the building operator.
-These models are then used for synthesizing recommendations about the control actions that the operator needs to take, during a DR event, to obtain a given load curtailment while providing guarantees on occupant comfort and operations.
+These models are then used for synthesising recommendations about the control actions that the operator needs to take, during a DR event, to obtain a given load curtailment while providing guarantees on occupant comfort and operations.
 
@@
 Its features in $\X^d$ include the weather variables, their lag terms and their forecast over the horizon, the schedule variables, and finally the lag terms of the power consumption.
-The temperature forest $\mathcal{R}_t$ is built with zone temperature $\tT$ as the output.
+The temperature forest $\mathcal{F}_t$ is built with zone temperature $\tT$ as the output.
 Except for the autoregressive terms corresponding to the same zone temperature, all other features are same in $\X^d$.
-Figure~\ref{F:sepvars} shows the prediction accuracy for the power forest, and also explains the two level training approach introduced in Section~\ref{SS:sepvar}.
+Figure~\ref{F:sepvars} shows the prediction accuracy for the power forest, and also explains the two-steps training approach introduced in Section~\ref{S:dpc}.
 During S1, the forests are trained using only disturbances as the features.
@@
 	\subfigure[Power tracking by DPC-En at 1.1 MW.
-In blue is the response when the optimal input is applied to the power Random Forest and in red is the response when the same input is applied to the plant (EnergyPlus building).
+In blue is the response when the optimal input is applied to the power predictive model of the Random Forest and in red is the response when the same input is applied to the power predictive model of EnergyPlus).
 The difference in closed-loop simulation and prediction is due to model mismatch.]{
@@
 \textcolor[rgb]{0,0,1}{The results from the closed-loop simulation are shown in Figure~\ref{F:eplus_power}.
-Marked in blue is the response when the optimal input is applied to the power Random Forest and in red is the response when the same input is applied to the plant (EnergyPlus building).
-The difference between the predicted power consumption and that in the closed-loop simulation is due to model mismatch between the EnergyPlus model and the power forest used in the optimization \eqref{E:ptrack}.}
+Marked in blue is the response when the optimal input is applied to the power predictive model of the Random Forest \eqref{E:ptrack} and in red is the response when the same input is applied to the power predictive model of EnergyPlus.
+Since the optimal input is computed using the power predictive model of the Random Forest the blue trajectory perfectly follows the tracking signal.
+The red trajectory, as expected, is characterised by a (small) tracking error because of the model mismatch between the predictive model of the Random Forest (used to compute the input) and that of EnergyPlus (used to simulate the closed-loop system).}
 Due to this inaccuracy, the actual power consumption is on an average 7 kW higher.

### 2018-01-22 paper/final/conclusion.tex
reviewed Intro, Sec 2,3,4,5, conclusion and appendix

@@
 On the other hand, DPC with regression trees suffers from practical limitations due to model overfitting.
-\item \emph{\textbf{Demand Response application.}} We further apply DPC with random forests to a large scale 6 story EnergyPlus model with 22 zones for which the traditional model-based control is largely unsuitable due to complex dynamics and the cost of model identification.
+\item \emph{\textbf{Application to Demand Response.}} We further apply DPC with random forests to a large scale 6 story EnergyPlus model with 22 zones for which the traditional model-based control is largely unsuitable due to complex dynamics and the cost of model identification.
 We show that DPC, relying only on the sensor data, can provide significant energy savings while maintaining thermal comfort.
 Our results demonstrate that even for such complex system, DPC tracks a reference signal with a mean error of $3\%$.
-\item \textcolor[rgb]{0,0,1}{\emph{\textbf{Optimal heating system scheduling of a real house.}} We finally apply DPC using experimental data from a real house and compare it with the classical bang-bang controller widely used for temperature control.
-We show that DPC provides energy savings while guaranteeing thermal comfort for the occupants.
-With respect to a bang-bang controller, we obtain energy savings that go from $25.4\%$, when we force the temperature to be strictly in a comfort range, to $49.2\%$, if we allow small violations in the thermal comfort level.
-Furthermore, we demonstrate the robustness of DPC considering uncertainties in the weather forecast.}
+\item \textcolor[rgb]{0,0,1}{\emph{\textbf{Application to optimal heating system scheduling.}} We demonstrate robustness of our method to uncertainties due to real data acquisition and weather forecast inaccuracies by implementing and testing DPC on historical data from an off-grid house located in L'Aquila, Italy.
+We derive a predictive model on such real data and design the optimal ON/OFF scheduling for the heating system in order to save energy while guaranteeing thermal comfort for the occupants.
+We compare the total amount of energy saved with respect to the classical bang-bang controller (widely used in houses for temperature control) using an EnergyPlus model built specifically for the house.
+We show that we can perform an energy saving that ranges from $25.4\%$ (if we guarantee thermal comfort i.e.~strictly respect the desired temperature range in the rooms) to $49.2\%$ (if we allow small violation in the desired temperature range).
+Finally, we test the robustness of our method to uncertainties in the real data acquisition and weather forecast.}
 \end{enumerate}
@@
 DPC has applications which go beyond buildings and energy systems, to industrial process control, and controlling large critical infrastructures like water networks, district heating \& cooling.
-DPC is immensely valuable in situations where first principles based modeling cost is extremely high.
+In general, DPC is immensely valuable in situations where first principles based modeling cost is extremely high.
 

### 2018-01-22 paper/final/dpc.tex
reviewed Intro, Sec 2,3,4,5, conclusion and appendix

@@
 the set of samples $(x(k),u(k),d(k))$ measured a time instants $k=1,\ldots,n$, where $x(k)\in\mathbb{R}^{n_x}$ is the vector of the \emph{measurable} state variables (e.g.
-the rooms temperatures and power consumption of the building), $u(k)\in\mathbb{R}^{n_u}$ is the vector of the \emph{measurable} input variables (e.g.
+the rooms temperatures, power consumption of the building, and others), $u(k)\in\mathbb{R}^{n_u}$ is the vector of the \emph{measurable} input variables (e.g.
 the set-points and schedule information of the building) and $d(k)\in\mathbb{R}^{n_d}$ is the vector of the \emph{measurable and predictable} disturbance variables (e.g.
@@
 \textcolor[rgb]{0,0,1}{Assume that, at time $k$, we want to predict the response variable at time $k+j$, i.e.
-$y(k+j),\ j\in \{1,\ldots,N\}$: when the data has lots of features interacting in complicated, nonlinear ways, assembling a single global model such as linear or polynomial regression can be difficult, and can lead to poor response predictions.}
+$y(k+j),\ j\in \{1,\ldots,N\}$: when the data have lots of features interacting in complicated, nonlinear ways, assembling a single global model such as linear or polynomial regression can be difficult, and can lead to poor response predictions.}
 
@@
 	\item the splitting of the dataset only partitions $\X^d$ (see Appendix 1 and \cite{hastie2009elements} for technical details): this choice is necessary to make our model suitable for control, as we will clarify later on, and also reduces the computational complexity.
-To each leaf $\ell_i$ will correspond an equivalence class of data samples $\X^d_i$ of the partition of $\X^d$, with $\X^d_i\subset \X^d$: as a consequence at any time $k$, given $\delta_d$ autoregressive terms of the disturbances and $\delta_x$ autoregressive terms of the state, we can associate $( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k) )$, to the corresponding leaf $\ell_i$
+To each leaf $\ell_i$ will correspond an equivalence class of data samples $\X^d_i$ of the partition of $\X^d$, with $\X^d_i\subset \X^d$: as a consequence at any time $k$, given $\delta_d$ autoregressive terms of the disturbances and $\delta_x$ autoregressive terms of the state, we can associate $( d(k-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k) )$, to the corresponding leaf $\ell_i$
 	\begin{equation}\label{E:model_tree}
-		\ell_i = \mathit{f}_{\mathcal{T}_j} \left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right)
+		\ell_i = \mathit{g}_{\mathcal{T}_j} \left( d(k-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right)
 	\end{equation}
-	if and only if $( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)) \in \X^d_i$.
+	if and only if $( d(k-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)) \in \X^d_i$.
 	\item in each leaf $\ell_i$ of the tree $\mathcal{T}_j$ we derive, solving a convex program over the data samples in the partition $\X^d_i$ associated to leaf $\ell_i$, an affine function, with coefficients $\beta_{ij}$, that relates the response $y(k+j)$ to the previous control inputs in $\X^c$:
@@
 \end{enumerate}
-Applying the above procedure for $j=0,\ldots,N$ we build $N$ regression trees $\mathcal{T}_0,\ldots,\mathcal{T}_N$: thus, we have managed to linearize the original model dynamics via black-box modeling.
-Our two-step training procedure, described by the pseudo code of lines 1-11 of Algorithm~\ref{A:dpcrt}, can be computed off-line: this is an important advantage because the time required to create the model does not affect the control execution in run-time.}
+As an example, the tree $\mathcal{T}_0$ defines the function $f$ in \eqref{E:GenericModel} to predict $y(k)$ as follows: assuming for simplicity that $\delta_x = \delta_d = 0$, given the measurement of the variables $(x(k), d(k))$ at time $k$, it is possible to determine by \eqref{E:model_tree} the corresponding leaf $\ell_i$ of $\T_0$ and the associated $\beta_{i0}$.
+The prediction of $y(k)$ is provided by \eqref{E:model_leaf} as an affine function of the control variable $u(k)$.\\
+Applying the above procedure for $j=0,\ldots,N$, we build $N$ regression trees $\mathcal{T}_0,\ldots,\mathcal{T}_N$: thus, we have managed to linearize the original model dynamics via black-box modeling.\\
+Our two-steps training procedure, described by the pseudo code of lines 1-11 of Algorithm~\ref{A:dpcrt}, can be computed off-line: this is an important advantage because the time required to create the model does not affect the control execution in run-time.}
 
@@
 		& \underset{u_{k+j},\epsilon_j}{\mathrm{minimize}} & & \sum_{j=0}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
-		& \mathrm{subject\ to }                 & & y_{k+j}     =   \beta_{ij} [1,u^\top_{k},\ldots,u^\top_{k+j} ]^\top                   \\
+		& \mathrm{subject\ to }                 & & y_{k+j}     =   \hat \beta_{j} [1,u^\top_{k},\ldots,u^\top_{k+j} ]^\top                       \\
 		&                                       & & u_{k+j}    \in  \mathcal{U}                                                                   \\
@@
 		&                                       & & \epsilon_j \geq  0							                                                  \\
-		&                                       & & j           =    0,\ldots,N.            									       \\
+		&                                       & & j           =    0,\ldots,N,               									                  \\
 		\end{aligned}
@@
 
-\textcolor[rgb]{0,0,1}{Here, $Q \succeq 0 \in \mathbb{R}^{n_y\times n_y}$ and $R \succeq 0 \in \mathbb{R}^{n_u\times n_u}$ are weight matrices used to tradeoff the importance we provide in the minimization of $y$ versus $u$.
+\noindent \textcolor[rgb]{0,0,1}{where $\hat \beta_j$ is defined later on.
+Here, $Q \succeq 0 \in \mathbb{R}^{n_y\times n_y}$ and $R \succeq 0 \in \mathbb{R}^{n_u\times n_u}$ are weight matrices used to trade-off the importance we provide in the minimization to $y$ versus $u$.
 The slack variables $\epsilon_j$ are added to ensure recursive feasibility: we relax the equality constraint on $y$ allowing violation up to $\epsilon_j$ to guarantee that Problem \ref{P:dpcrt} can provide a solution at each step $k \geq 0$.
@@
 In the current formulation, the data-driven control problem, is reduced to a convex program which is very easy and efficient to solve in C, C++, Python and Matlab, and thus is easily integrable in SCADA systems.
-Indeed Problem \ref{P:dpcrt} is solved as in the classical MPC formulation, i.e.
+Indeed, Problem \ref{P:dpcrt} is solved as in the classical MPC formulation, i.e.
 at each time step $k=1,2,\ldots$ the optimal control sequence $u^*_k,\ldots,u^*_{k+N}$ is computed, and only the first input of the sequence is applied to the system: $u(k) = u^*_k$.\\
-Note that each tree $\T_j$ contributes to Problem  \ref{P:dpcrt} with the linear constraint $y_{k+j} = \beta_{ij} [1,u^\top_{k},\ldots,u^\top_{k+j} ]^\top$ as a replacement for the state dynamics in the classical MPC formulation.
-As a consequence, when solving Problem \ref{P:dpcrt} at time $k$, we need to determine the affine functions $\beta_{ij}$: it follows by Equation \eqref{E:model_tree} that, to determine $\beta_{ij}$ at time $k$, the knowledge of the state and disturbance measurements $\left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k) \right)$ is needed, however, values of $d(k+j-\delta_d),\ldots,d(k+j),\ j=0,\ldots,N$ are not available at time $k$, so we need to use disturbance forecast $\tilde d(k+N),\ldots,\tilde d(k)$.
-Using this information we can narrow down to a leaf in each tree using \eqref{E:model_tree} and thus retrieve the affine model with $\beta_{ij}$ in \eqref{E:model_leaf} for each step $j=0,\ldots,N$ of the horizon.
+Note that each tree $\T_j$ contributes to Problem  \ref{P:dpcrt} with the linear constraint $y_{k+j} = \hat \beta_{j} [1,u^\top_{k},\ldots,u^\top_{k+j} ]^\top$ as a replacement for the state dynamics in the classical MPC formulation.
+As a consequence, when solving Problem \ref{P:dpcrt} at time $k$, we need to determine the affine functions parameters $\hat \beta_{j}$: it follows by Equation \eqref{E:model_tree} that, to determine $\beta_{ij}, j=0,\ldots,N$ at time $k$, the knowledge of the state and disturbance measurements $\left( d(k-\delta_d),\ldots,d(k+N),x(k-\delta_x),\ldots,x(k) \right)$ is needed.
+However, the values of $d(k+1),\ldots,d(k+N)$ are not available at time $k$, so we need to use disturbance forecast $\tilde d(k+1),\ldots,\tilde d(k+N)$.
+Using this information we can narrow down to a leaf in each tree using \eqref{E:model_tree} and thus retrieve the affine model with $\beta_{ij}$ in \eqref{E:model_leaf} for each step $j=0,\ldots,N$ of the horizon, and associate $\hat \beta_j := \beta_{ij}$.
 The run-time solution of Problem \ref{P:dpcrt} illustrated above is described by the pseudo code of lines 12-23 of Algorithm~\ref{A:dpcrt}.}
@@
 		\ForAll{leaves $\ell_i$ of $\mathcal{T}_j$}
-		\State Compute parameters $\beta_{ij}$ in \eqref{E:model_leaf} using convex programming
+		\State Compute parameters $\hat\beta_j:=\beta_{ij}$ in \eqref{E:model_leaf} using convex programming
 		\EndFor
@@
 Each tree $\T_{\kappa j}$ is trained on the basis of a different subset of features $\bar{\X}^d_{\kappa j} \subset \X^d$.
-As discussed in the previous section, at any time $k$, given $\delta_d$ autoregressive terms of the disturbances and $\delta_x$ autoregressive terms of the state $\left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k) \right)$, we can associate the corresponding leaf $\ell_i$ of $\T_{\kappa j}$
+As discussed in the previous section, at any time $k$, given $\delta_d$ autoregressive terms of the disturbances and $\delta_x$ autoregressive terms of the state, we can associate $( d(k-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k) )$, to the corresponding leaf $\ell_i$ of $\T_{\kappa j}$
 \begin{equation}\label{E:model_forest}
-	\ell_i = \mathit{f}_{\mathcal{T}_{\kappa j}} \left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right).
+	\ell_i = \mathit{g}_{\mathcal{T}_{\kappa j}} \left( d(k-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right).
 \end{equation}
@@
 		&                                       & & \epsilon_j \geq  0							                                     \\
-		&                                       & & j           =    0,\ldots,N.            									         \\
+		&                                       & & j           =    0,\ldots,N,            									         \\
 		\end{aligned}
@@
 \end{problem}
-\textcolor[rgb]{0,0,1}{Note that, as in the DPC-RT, the crucial problem when solving Problem \ref{P:dpcrf} is determining all parameters $\hat{\Theta}_{j}$ at time $k$.
+\noindent \textcolor[rgb]{0,0,1}{where $\hat{\Theta}_{j}$ is defined later on.
+Note that, as in the DPC-RT, the crucial problem when solving Problem \ref{P:dpcrf} is determining all parameters $\hat{\Theta}_{j}$ at time $k$.
 We first proceed similarly to DPC-RT by using forecast of disturbances and the state measurements to determine for each tree $\T_{\kappa j}$ the corresponding leaf $\ell_i$ and thus the affine model $\Theta_{i \kappa j}$, with $\kappa = 1,\ldots,t$ and $j = 0,\ldots,N$.

### 2018-01-22 paper/final/intro.tex
reviewed Intro, Sec 2,3,4,5, conclusion and appendix

@@
 \textcolor[rgb]{0,0,1}{Control-oriented models of energy system's dynamics and energy consumption, are needed for understanding and improving the overall energy efficiency and operating costs of a building.
-With a reasonably accurate forecast of future weather and building operating conditions, dynamical models can be used to predict the energy needs of the building over a prediction horizon, as is the case with Model Predictive Control (MPC) \cite{Sturzenegger2016}.
+With a reasonably accurate forecast of future weather and building operating conditions, dynamical models can be used to predict the energy needs of the building over a prediction horizon, and use them to determine optimal control actions to save energy and guarantee thermal comfort, as is the case with Model Predictive Control (MPC) \cite{Sturzenegger2016}.
 However, a major challenge with MPC is in (accurately) modeling the dynamics of the underlying physical system.
-The task is much more complicated and time consuming in the case of a large building and often times, it can be even more complex and involved than the controller design itself.
+The task is much more complicated and time consuming in the case of a large buildings and often times, it can be even more complex and involved than the controller design itself.
 After several years of work on using first principles based models for peak power reduction, energy optimization and thermal comfort for buildings, multiple authors~\cite{Sturzenegger2016, vzavcekova2014} have concluded that the biggest hurdle to mass adoption of MPC in intelligent building control is the cost, time and effort required to capture accurate dynamical models of the buildings.
@@
 \begin{enumerate}
-	\item \textcolor[rgb]{0,0,1}{\textbf{Model capture} - a building modeling domain expert typically uses a software tool to create the geometry of a building from the building design and equipment layout plans, and add detailed information about material properties, equipment and operational schedules.
-	There is always a gap between the modeled and the real building, and the domain expert must then manually tune the model to match the measured data from the building~\cite{New2012}. 
+	\item \textcolor[rgb]{0,0,1}{\textbf{Model capture} - a building modeling domain expert typically uses a software tool to create a model to reproduce the geometry of a building from the building design and equipment layout plans, and add detailed information about material properties, equipment and operational schedules.
+	However, there is always a gap between the modeled and the real building, and the domain expert must then manually tune the model to match the measured data~\cite{New2012}. 
 	Moreover, the modeling process also varies from building to building with the construction and types of installed equipment.
-	Another major downside with physics-based modeling is that enough data is not easily available and guesses for parameter values have to be made, which also requires expert know-how.
+	Another major downside with physics-based modeling is that enough data is not easily available, so guesses for parameter values have to be made, which also requires expert know-how.
 	\item \textbf{Change in model properties over time} - even if the model is identified once via an expensive route as in \cite{Sturzenegger2016}, as the model changes with time, the system identification must be repeated to update the model.
@@
 The primary advantage of using data-driven methods is that it has the potential to eliminate the time and effort required to build white and grey box building models. 
-Listening to real-time data, from existing systems and interfaces, is far cheaper than unleashing hoards of on-site engineers to physically measure and model the building.
+Listening to real data, from existing systems and interfaces, is far cheaper than unleashing hoards of on-site engineers to physically measure and model the building.
 Improved building technology and better sensing is fundamentally redefining the opportunities around smart buildings.
@@
 \textcolor[rgb]{0,0,1}{The key question now is: \emph{can we employ data-driven techniques to reduce the cost of modeling, and still exploit the benefits that MPC has to offer?}
-We therefore look for automatic data-driven approaches to control, that are also adaptive, scalable and interpretable.
+We therefore look for automatic data-driven approaches for control, that are also adaptive, scalable and interpretable.
 We solve this problem by bridging Machine Learning and Predictive Control.
@@
 	\caption{\textcolor[rgb]{0,0,1}{References ordered considering: case study they are applied to; whether they use experimental data, other than simulated data, and if they do real implementation (RI), i.e.
-implement the methodologies on real systems; if they use simulative tool; the type of the model considered, i.e.
-Model-Based or Data-Driven; if the models are used for control.}}
+implement the methodologies on real systems; if they use simulative tools; the type of the model considered, i.e.
+Model-Based or Data-Driven or both; if the models are used for control.}}
 	\captionsetup{justification=centering}
@@
 These references are summarized in Table \ref{T:RelatedWork} highlighting the key differences.
-We also emphasize the case studies the results are applied to, and whether the authors use experimental data to simulate their algorithms.
+We also emphasize the case studies the results are applied to, and whether the authors used experimental data to simulate their algorithms.
 Only in three cases the algorithms are tested on real systems (see Table \ref{T:RelatedWork}: RI - Real Implementation).
-We observe that, except for the last six cases \cite{Macarulla2017,Costanzo2016,Ferreira2012,Afram2017,Behl2016,Jain2017TCPS}, which we discuss in detail, either model-based approaches or only tools are considered with/without control, or data-driven modeling approaches are considered only without control.
+We observe that, except for the last six cases \cite{Macarulla2017,Costanzo2016,Ferreira2012,Afram2017,Behl2016,Jain2017TCPS}, which we discuss in detail, either model-based approaches or only tools are considered with/without control, or data-driven modeling approaches are considered \emph{only without} control.
 %This is because bridging data-driven models with control theory, especially with the predictive control, introduces several issues \cite{Hou2013}.
-The last six papers of Table \ref{T:RelatedWork} are more related to the methodology presented in this paper.
-In particular, the authors in \cite{Macarulla2017} proposed a predictive control strategy based on Neural Networks, for boilers control in buildings, to decide the optimal time to switch-on the plan to guarantee energy savings and thermal comfort.
+The last six papers of Table \ref{T:RelatedWork} are more related to the methodology presented in this paper, since they address both data-driven modeling and control.
+In particular, the authors in \cite{Macarulla2017} proposed a predictive control strategy based on Neural Networks, for boilers control in buildings, to decide the optimal time to switch-on the plant to guarantee energy savings and thermal comfort.
 However, the approach is not easily scalable to different types of plants and does not use optimization in the closed-loop scheme.
@@
 In the conference papers \cite{JainACC2017,JainCDC2017}, we took a different approach that improves the system's identification accuracy -- instead of considering a single tree with multiple output, we considered multiple trees and forests with single output.
-Each tree/forest provides the prediction of the system's behavior for a different time step of the horizon. 
-However, the results are based on simulated data and we did not account for inaccuracies in the weather forecast.
+Each tree/forest provides the prediction of the system's behavior for a different time steps of the horizon. 
+However, the results are based on simulated data and we did not account for inaccuracies in the weather forecast and data acquisition.
 }
@@
 In this paper, we extend these results providing the following contributions.
-In this paper, we provide a new methodology based on random forests that overcomes the drawbacks of all our previous work, and more precisely: we obtain better performance when compared to other approaches (both optimal and rule-based), and we provide and validate robustness with respect to disturbance uncertainties.
-Further, we test our method using an experimental data set. 
-More precisely, the paper makes the following contributions.
+In this paper, we provide a new methodology based on random forests that overcomes the drawbacks of all our previous works, and more precisely: we obtain better performance and scalability when compared to other approaches (both optimal and rule-based), and we provide and validate robustness with respect to uncertainties due to real data acquisition and weather forecast inaccuracies.
+More in detail, the paper contribution is organised as follows.
 \begin{enumerate}
-	\item In Section \ref{S:dpc}, we formally describe our data-driven Model Predictive Control approach i.e.~Data Predictive Control (DPC).
+	\item In Section \ref{S:dpc}, we formally describe our Data-driven model Predictive Control approach, i.e.~Data Predictive Control (DPC).
 For clarity, we first modify the regression trees algorithm to apply DPC, and then we introduce DPC based on random forests.
-	\item In Section \ref{S:proof}, we demonstrate that DPC can provide comparable performance when compared to MPC (using a physical model) applied to a perfectly known mathematical model.
-More precisely, we first consider a benchmark MPC controller using a bilinear building model whose parameters were identified using experiments on a building in Switzerland, then we apply DPC to a data-set simulated from such bilinear model and derive a controller based only on the data.
+	\item In Section \ref{S:proof}, we demonstrate that DPC can provide comparable \textit{performance} when compared to MPC (using a physical model) applied to a perfectly known mathematical model.
+More precisely, we first consider a benchmark MPC controller using a bilinear building model whose parameters were identified using experiments on a building in Switzerland, then we apply DPC to a dataset simulated from such bilinear model and derive a controller based only on the data.
 We show that DPC captures 70\% variance in MPC and offers a comparable performance.
-	\item In Section \ref{S:casestudy}, we demonstrate scalability and efficiency of DPC in a Demand Response problem: DPC is used to provide financial incentives to the end-customers bypassing the need for expensive high fidelity models.
+	\item In Section \ref{S:casestudy}, we demonstrate scalability and efficiency of DPC in a Demand Response problem with the aim of enabling financial incentives for the end-customers bypassing the need for expensive high fidelity models.
 We apply DPC to a 6 story 22 zone building model in EnergyPlus for which model-based control is not economical and practical due to extreme complexity.
 We observe that DPC provides the desired power curtailment with an average error of 3\%.
-	\item In Section \ref{S:realCaseStudy}, we implement and test DPC basing on real data from an off-grid house located in L'Aquila, Italy.
-We design the optimal ON/OFF scheduling for the heating system in order to save energy while guaranteeing thermal comfort for the occupants.
+	\item In Section \ref{S:realCaseStudy}, we demonstrate the robustness of our method to uncertainties due to real data acquisition and weather forecast inaccuracies by implementing and testing DPC on historical data from an off-grid house located in L'Aquila, Italy.
+We derive a predictive model on such real data and design the optimal ON/OFF scheduling for the heating system in order to save energy while guaranteeing thermal comfort for the occupants.
 We compare the total amount of energy saved with respect to the classical bang-bang controller (widely used in houses for temperature control) using an EnergyPlus model built specifically for the house.
 We show that we can perform an energy saving that ranges from $25.4\%$ (if we guarantee thermal comfort i.e.~strictly respect the desired temperature range in the rooms) to $49.2\%$ (if we allow small violation in the desired temperature range).
-Further, we test the robustness of our method due to uncertainties in the weather forecast.
+Finally, we test the robustness of our method to uncertainties in data acquisition and weather forecast.
 \end{enumerate}
@@
 
-\textcolor[rgb]{0,0,1}{(1) Building geometry and construction data are used together with first-principles to derive the following linear model for the building's thermal dynamics:
+\textcolor[rgb]{0,0,1}{\begin{enumerate}
+		\item Building geometry and construction data are used together with first-principles to derive the following linear model for the building's thermal dynamics:
 		\begin{equation}\label{E:RCZoneEq}
@@
 		This model describes the behavior of the zone, wall, floor and ceiling temperatures.
-		Walls, floors and ceilings are considered as divided into layers with different features.
+		Walls, floors and ceilings are considered as divided into 3 layers with different features.
 Therefore, each zone was described with an RC network model (see Figure 3-10 in \cite{SturzeneggerTR}), where the capacitances represent the states of the layers and the resistances represent the thermal resistance of the layers.
@@
 The thermal parameters are derived from zones geometry and material properties.
-}
-
-\textcolor[rgb]{0,0,1}{(2) External heat fluxes, modeled as a bilinear model, affect the building directly as well as indirectly through zones:
+		\item External heat fluxes, modeled as a bilinear model, affect the building directly as well as indirectly through zones:
 		\begin{equation}\label{E:RCfluxes}
@@
 	\end{itemize}
-}
-
-\textcolor[rgb]{0,0,1}{(3) The resulting system \eqref{E:RCZoneEq} is discretized with a sampling time of $15\ \mathrm{min}$.
-This model has approximately 300 states that include temperature of the zones, walls and floors on the second floor.
+	\item The resulting system \eqref{E:RCZoneEq} is discretized with a sampling time of $15\ \mathrm{min}$.
+This model has approximately 300 states, that include temperature of the zones, walls and floors on the second floor.
 The outputs of the system are the zone temperatures.
@@
 This filtering is needed to take into account the shadowing of the neighboring buildings.
-This approximate model is then considered ``suitable for MPC" (see Section 3.3.1.4 in \cite{SturzeneggerTR}).}
+This approximate model is then considered ``suitable for MPC" (see Section 3.3.1.4 in \cite{SturzeneggerTR}).
+\end{enumerate}}
 
@@
 \begin{equation}\label{E:MLexample}
-	y(t) = f(u(t),\dots,u(t-\delta_u), d(t),\dots, d(t-\delta_d)),
+	y(t) = f(x(k),\ldots,x(k-\delta_x), u(k),\dots,u(k-\delta_u), d(k),\dots, d(k-\delta_d)),
 \end{equation}
-where \(y\) represents the outputs, \(u\) the inputs and \(d\) the disturbances, and \(\delta\) is a memory variable which captures the dynamical behavior.
-While the inputs and external disturbances are exactly the same as discussed earlier in Section \ref{SSS:physics-based}, the outputs in \eqref{E:MLexample} include only the variables that are directly measurable through already installed sensors like thermostats and multimeters.
+where \(y\) represents the variables we wish to predict, \(x\) the measurable variables of the system, \(u\) the measurable inputs and \(d\) the measurable disturbances, with $\delta_x, \delta_u, \delta_d$ memory indices of the variables which capture the dynamical behavior.
+Compared to the variables in the example presented above in Section \ref{SSS:physics-based}, all variables $y,x,u,d$ in \eqref{E:MLexample} only include variables that are directly measurable through already installed sensors like thermostats and multimeters.
 Therefore, many internal states like the temperatures of different layers -- interior, middle and exterior -- of the walls, the floors and the ceilings are not required for black-box modeling, which reduces the order of complexity.
@@
 	In the case of heat flux in \eqref{E:RCfluxes}, again, a bilinear model ignores complex nonlinearities.
-	Hence, much nonlinearities are neglected to simplify the modeling complexities.
-	We avoid such assumptions in data-driven modeling where a nonlinear function such as Neural Network or Random Forest, that represents the dynamics of zone temperatures or power consumption, is learned rather efficiently and accurately.
+	Hence, many nonlinearities are neglected to simplify the modeling complexities.
+	We avoid such assumptions in data-driven modeling where a nonlinear function such as Random Forest, that represents the dynamics of zone temperatures or power consumption, is learned rather efficiently and accurately.
 	For example, if \(y\) is the zone temperature in Equation \eqref{E:MLexample}, we learn a nonlinear function \(f\) which depends on current and previous inputs and disturbances. 
@@
 	Physical modeling neglects different geometries and different materials of the floors assuming they are the same for each floor, which is never true in reality.
-	Moreover, in many cases, the details of the construction layout and equipments are not even available so many parameters have to be guessed making physical modeling a difficult choice.
+	Moreover, in many cases, the details of the construction layout and equipments are not even available, so many parameters have to be guessed making physical modeling a difficult choice.
 	On the other hand, data-driven modeling automatically captures the interaction with the environment while optimizing the hyperparameters.
@@
 	The behavior of the internal/unobservable/missing states is captured in the parameters of function \(f\).
-	For example, if we do not have the measurement of the energy produced by a photovoltaic panel, the weather conditions will account for it.}
+%	For example, if we do not have the measurement of the energy produced by a photovoltaic panel, the weather conditions will account for it.
+}
 \end{enumerate}
@@
 	To summarize, the data-driven approaches have the potential to simplify the modeling of buildings to a large extent reducing the overall cost and time investment, while avoiding several assumptions that are generally made in standard physics-based modeling procedures and that reduce the accuracy.
-	Also, unlike physics-based modeling, the parameters (or hyperparamters) of the model are automatically determined by the learning algorithm to compensate for the effect of unobservable variables on the input-output relation; the best settings being determined by cross validation.\\
+	Also, unlike physics-based modeling, the parameters (or hyperparamters) of the model are automatically determined by the learning algorithm to compensate for the effect of unobservable variables on the input-output relation.\\
 	Given the advantages of data-driven modeling, the challenge now lies in using such models for optimal predictive control.
 	This is exactly the focus of this paper. 
-	In the next sections, we will demonstrate how the training algorithm for Random Forests can be modified to develop control oriented models that enable Model Predictive Control, and provide better performance, when compared to other approaches, and robustness with respect to disturbance uncertainties.}
+	In the next sections, we will demonstrate how the training algorithm for Random Forests can be modified to develop control oriented models that enable Model Predictive Control, and provide better performance, when compared to other approaches, and robustness with respect to disturbance and to real data acquisition uncertainties.}

### 2018-01-22 paper/final/Highlights.tex
review of Sec 6 and remark added in Sec3

--- 2017-09-26 paper/Highlights.tex
@@
 \maketitle 
+\thispagestyle{empty}
 
-\thispagestyle{empty}
 \begin{itemize}
 	% Remember maximum 85 characters (included spaces, points, commas, etc.)
-	\item Predictive control strategy using data-based models via machine learning algorithms.
+	\item Predictive control strategy using data-driven models via Random Forests.
 	\item Data Predictive Control shows comparable performance to Model Predictive Control.
 	\item Data Predictive Control shows good performance in Demand Response curtailment.
-	\item Data Predictive Control allows energy saving up to $50\%$ on the considered case study.
+	\item Data Predictive Control allows energy saving up to $50\%$ in considered case study.
+	\item Data Predictive Control is robust to data and weather forecast uncertainties.
 \end{itemize}

### 2018-01-22 paper/final/casestudyReal.tex
review of Sec 6 and remark added in Sec3

@@
-\section{CASE STUDY: OPTIMAL HEATING SYSTEM SCHEDULING}
+\section{APPLICATION TO OPTIMAL HEATING SYSTEM SCHEDULING}
 \label{S:realCaseStudy}
 
-In this section, we show an application of DPC to a real house located in L'Aquila, Italy.
+In this section we show an application of DPC to a real house located in L'Aquila, Italy.
 Random forest models for DPC built using historical data from this house are described in Section \ref{SS:randomForestsModels}.
@@
 In Section \ref{SSS:DisturbanceUncertain}, we show that DPC is robust with respect to imperfect weather forecast.}
-The section organization is graphically shown in Figure \ref{F:overview}.
 \begin{figure}[h!]
@@
 		\caption{Random Forest trained on all features including the control variable are used as the plant in the closed-loop simulations with DPC and bang-bang controller.
-DPC optimzation uses Random Forests with affine functions as predictive models.
+DPC optimization uses Random Forests with affine functions as predictive models.
 Controllers' strategies are evaluated using the EnergyPlus model to compare the actual energy consumption.}
@@
 
-\textcolor[rgb]{0,0,1}{The contribution of this section is threefold.
+\textcolor[rgb]{0,0,1}{The contribution of this section, graphically shown in Figure \ref{F:overview}, is threefold.
 \begin{enumerate}
-	\item We consider experimental data from an occupied house, hence subject to real world imperfections, such as random (non predictable) occupancy schedules, open/close windows, random (non predictable) light on/off switch, etc.
-	This is different from Section \ref{S:casestudy} where we used data simulated from an EnergyPlus model.
-	\item We modify the DPC algorithm to take into account uncertainties In the weather forecast, making DPC robust with respect to prediction errors.
+	\item Differently from Section \ref{S:casestudy}, where we used data simulated from an EnergyPlus model, we consider here real data from an occupied house, hence subject to real world imperfections such as random (non predictable) occupancy schedules, open/close windows, random (non predictable) light on/off switch, etc.
 	\item While in Section \ref{S:casestudy} we follow a reference DR signal while maintaining thermal comfort, in this section we show the adaptability of DPC to different problems where we minimize the energy consumption required to keep the room temperatures within a specified range of comfort.
 In this way, we show the potential of DPC in providing energy savings for a long period.
+	\item We consider uncertainties in the weather forecast, showing that DPC is robust with respect to prediction inaccuracies.
 \end{enumerate}}
@@
 \begin{enumerate}
-	\item In Section \ref{SSS:DisturbancePerfect}, we provide the results considering perfect knowledge of the weather forecast.
+	\item In Section \ref{SSS:DisturbancePerfect}, we provide the results considering perfect knowledge of the weather forecast and only focus on robustness with respect to uncertainties due to real data acquisition.
 	\item In Section \ref{SSS:DisturbanceUncertain}, we consider weather forecast subject to uncertainties.
-Hence, we modify Problem \ref{P:dpcRealCase} by adding a Gaussian noise with high variance to the perfect forecast, and show that the results are close to the perfect forecast case.
-	This verifies our methodology is robust with respect to disturbance uncertainties.
+We modify Problem \ref{P:dpcRealCase} by adding Gaussian noise to the perfect forecast and show that the results are close to the perfect weather forecast case.
 \end{enumerate}
@@
 
-\textcolor[rgb]{0,0,1}{
 \subsubsection{Weather forecast subject to uncertainty}\label{SSS:DisturbanceUncertain}
-The authors in \cite{Petersen2014AE} investigate the effect of uncertainty in the weather forecast on the performance of MPC in building systems operations through a large-scale simulation study, and compare against a rule-based strategy.
-They consider 48 different scenarios of uncertainties for 72 hours weather forecast.
-With such a long horizon, results have shown that (with a few exceptions) MPC outperforms the rule-based controller in providing energy savings, and is in general quite close to the perfect forecast case despite the uncertainty in weather forecast.}
 
-\textcolor[rgb]{0,0,1}{The length of horizon for the DPC algorithm presented in this paper is usually much shorter, for example $6$ hours in Section \ref{S:proof}, $7$ hours in Section \ref{S:casestudy}, and $40$ minutes in Section \ref{S:realCaseStudy}.
-%This means the weather forecast is more accurate than the 72 hours case presented in \cite{Petersen2014AE}, hence also control performance is even closer to the perfect knowledge forecast case.
-We next show the robustness of our approach considering noisy weather forecast.}
+\textcolor[rgb]{0,0,1}{In this section we study the robustness of our approach to noisy weather forecast.
+To this aim, we add noise with a zero-mean Gaussian distribution to the prediction variables $\tilde d(k+1),\ldots,\tilde d(k+N)$ used to obtain the parameters $\hat \Theta_{\cdot_j}$, $j=0,\ldots,N$ in the DPC Problem \eqref{E:DPCrealcase} (i.e.~the disturbance in the input in the blue rectangle in Figure \ref{F:overview}) while we use the true values to simulate the process (i.e.~the disturbance in the input in the box ``Plant" in the red rectangle in Figure \ref{F:overview}).\\
+We summarize the mean and the standard deviation of the Gaussian noises that we add to each variable in Table \ref{T:NoiseParameters}.
+For example, a standard deviation of $0.5\degree C$ on the outside air temperature means that the error on the predicted temperature lies within the range of $\pm 1.5\degree C$ with a probability of $99\%$.
+Since ``Time of the day" and ``Day of the week" are perfectly known, we clearly do not add any noise to these variables.
+Table \ref{T:NoiseParameters} also provides the range of values of the variables in the historical disturbances dataset to emphasise that the standard deviation we choose adds significant error to the forecast.
+}
 
-\textcolor[rgb]{0,0,1}{We add noise with normal distribution to the disturbance variables in $\X^d$ which is used to obtain parameters $\hat \Theta$ in the DPC problem \eqref{E:DPCrealcase} i.e.~the disturbance in the input in the blue rectangle in Figure \ref{F:overview}, while we use the true values to simulate the process i.e.~the disturbance in the input in the box ``Plant" in the red rectangle in Figure \ref{F:overview}.
-We summarize the mean and the standard deviation of the Gaussian noises that we add to each variable in Table \ref{T:NoiseParameters}.
-Since ``Time of the day" and ``Day of the week" are perfectly known, we do not add any noise to these variables.
-We assume the noise to have zero mean.
-Table \ref{T:NoiseParameters} also lists the range of the variables in the historical dataset.
-This is to show that the standard deviation we choose adds significant error to the forecast.
-For example, a standard deviation of $0.5\degree C$ on the outside air temperature means that the error on the predicted temperature lies within the range of $\pm 1.5\degree C$ with a probability of $99\%$.
-We solve the DPC problem \ref{P:dpcRealCase} considering the $\hat \Theta$ obtained using the noisy forecast.
-We obtain the following results. We show the change in the performance on bounds violation and energy consumption due to inaccurate forecast.}
 \begin{table}[t!]
@@
 
-\textcolor[rgb]{0,0,1}{\paragraph{Result 1} Figure \ref{F:ComparisonTempNoisy} shows a comparison of the DPC results in terms of temperature control for thermal comfort between the perfect and the noisy forecast.
+\textcolor[rgb]{0,0,1}{The following paragraphs provide comments on the change in performance due to inaccurate forecast on bounds violation and on energy consumption.\\
+\paragraph{Result 1} Figure \ref{F:ComparisonTempNoisy} shows a comparison of the DPC results in terms of temperature control for thermal comfort between the perfect and the noisy forecast.
 For the sake of clarity, we show only \(4\frac{1}{2}\) days of the 15 days simulation period. We split the plot into 3 sub-figures referring to the small, medium and large violation cases.
@@
 In Figure \ref{F:comparison_all_energy_E+_noisy}, we show the difference in terms of energy consumption considering the perfect weather forecast (solid line) and the noisy forecast (dashed line).
-We observe that the energy usage is very close in both cases.
-Recall, we optimize a weighted sum of energy and thermal comfort, so depending on the conditions for which the disturbance is over or underestimated, the MPC solution changes depending upon the prediction errors.
-In the small violation case, the energy consumption of the two conditions is closer than in the others.
-This is because the DPC keeps the temperature within the bounds with an extremely small error in bounds violation.
-In particular, the only exception can be seen before May $3^{rd}$ (03/05), where, due to the prediction error, the temperature violates the bounds, so the control uses more energy (than the perfect case) to bring it back above the threshold.
-In the other two cases, when the temperature violates the bounds, due to the fact that violation constraints are more relaxed, the DPC does not require much larger control effort, hence using lesser energy.}
+We first recall that our optimization function is a weighted sum of energy and thermal comfort: when the weather forecast is affect to inaccuracies the MPC solution (and thus the control input) changes depending upon the prediction errors.\\
+We immediately observe that the energy usage in the ideal and perturbed cases is very close, especially in the small violation case.
+Our interpretation for this is that the DPC, in the small violation case, keeps the temperature within the bounds with an extremely small error in bounds violation.
+The only exception can be seen before May $3^{rd}$ (03/05) where, due to a large prediction error, the temperature violates the bounds: as a consequence the controller uses much more energy (with respect to the perfect forecast knowledge case) to bring the temperature in the bounds, as evidenced by the blacked lines of Figure \ref{F:comparison_all_energy_E+_noisy}.
+In the medium and large violation cases, instead, when the temperature violates the bounds the DPC does not strongly increase the energy usage as the violation tolerance is more relaxed: this produces a general reduction of energy usage in the "noisy" simulations, as evidenced by the purple and red lines of Figure \ref{F:comparison_all_energy_E+_noisy}.
+}
 

### 2018-01-22 paper/final/dpc.tex
review of Sec 6 and remark added in Sec3

@@
 \section{DATA PREDICTIVE CONTROL}
-
 \label{S:dpc}
@@
 \subsection{DPC-RT: DPC with Regression Trees}
-
 \label{SS:dpcrt}
@@
 \begin{figure}[t!]
-
 	\centering
-
 	\includegraphics[width=20pc]{figures/dpc-sepvars.eps}
-
 	\caption{\textit{Step 1:} Tree $\mathcal{T}_j$ trained only with variables in $\X^d$ using adapted RT algorithm.
@@
 \textit{Step 2:} In the leaf $\ell_i$ of the tree an affine model $\beta_{ij}$ is defined as a function only of the control variables dataset corresponding to leaf $\ell_i$.}
-
 	\captionsetup{justification=centering}
-
 	\label{F:dpc-sepvars}
-
 \end{figure}
@@
 $y(k)\in\mathbb{R}^{n_y},\ n_y\geq 1$, as follows:}
-
 \begin{problem}\label{P:dpcrt}
@@
 The run-time solution of Problem \ref{P:dpcrt} illustrated above is described by the pseudo code of lines 12-23 of Algorithm~\ref{A:dpcrt}.}
+
+\textcolor[rgb]{0,0,1}{\begin{remark}
+The authors in \cite{Petersen2014AE} investigate the effect of uncertainty in the weather forecast on the performance of MPC in building systems operations through a large-scale simulation study, and compare against a rule-based strategy.
+They consider 48 different scenarios of uncertainties for 72 hours weather forecast.
+With such a long horizon, results have shown that (with a few exceptions) MPC outperforms the rule-based controller in providing energy savings, and is in general quite close to the perfect forecast case despite the uncertainty in weather forecast.
+The length of horizon for the DPC algorithm presented in this paper is usually much shorter, for example $6$ hours in Section \ref{S:proof}, $7$ hours in Section \ref{S:casestudy}, and $40$ minutes in Section \ref{S:realCaseStudy}.
+As a consequence, we can reasonably presume that our approach will be robust to weather forecast inaccuracies.
+Indeed, in Section \ref{S:realCaseStudy} we test the robustness of our approach to noisy weather forecast and show that the control performance is very close to the ideal case.
+%This means the weather forecast is more accurate than the 72 hours case presented in \cite{Petersen2014AE}, hence also control performance is even closer to the perfect knowledge forecast case.
+\end{remark}}
 
@@
 \subsection{DPC-En: DPC with Ensemble Methods}
-
 \label{SS:dpcrf}
@@
 %In particular, as mentioned above, each tree $\T_i$ of each forest $\F_j$ is trained using $(\bar \X^d_i,\bar\Y_i)$.
-
 %Here $(\bar\X^c_i,\bar\Y_i)$ correspond to the in-bag samples for the trees.
@@
 \end{problem}
+
 \noindent \textcolor[rgb]{0,0,1}{where $\hat{\Theta}_{j}$ is defined later on.
@@
 The overall procedure is sketched in Figure \ref{F:dpc-algo-rf}.}
-
 \begin{figure}[t!]

### 2018-01-22 paper/final/casestudyReal.tex
corrected typos

@@
 \textcolor[rgb]{0,0,1}{In this section we study the robustness of our approach to noisy weather forecast.
-To this aim, we add noise with a zero-mean Gaussian distribution to the prediction variables $\tilde d(k+1),\ldots,\tilde d(k+N)$ used to obtain the parameters $\hat \Theta_{\cdot_j}$, $j=0,\ldots,N$ in the DPC Problem \eqref{E:DPCrealcase} (i.e.~the disturbance in the input in the blue rectangle in Figure \ref{F:overview}) while we use the true values to simulate the process (i.e.~the disturbance in the input in the box ``Plant" in the red rectangle in Figure \ref{F:overview}).\\
+To this aim, we add noise with a zero-mean Gaussian distribution to the prediction variables $\tilde d(k+1),\ldots,\tilde d(k+N)$ used to obtain the parameters $\hat \Theta_{\cdot_j}$, $j=0,\ldots,N$ in the DPC Problem \eqref{E:DPCrealcase} (i.e.~the disturbance in input in the blue rectangle in Figure \ref{F:overview}), while we use the true values to simulate the process (i.e.~the disturbance in input in the box ``Plant" in the red rectangle in Figure \ref{F:overview}).\\
 We summarize the mean and the standard deviation of the Gaussian noises that we add to each variable in Table \ref{T:NoiseParameters}.
@@
 In Figure \ref{F:comparison_all_energy_E+_noisy}, we show the difference in terms of energy consumption considering the perfect weather forecast (solid line) and the noisy forecast (dashed line).
-We first recall that our optimization function is a weighted sum of energy and thermal comfort: when the weather forecast is affect to inaccuracies the MPC solution (and thus the control input) changes depending upon the prediction errors.\\
+We first recall that our optimization function is a weighted sum of energy and thermal comfort: when the weather forecast is affected by inaccuracies, the MPC solution (and thus the control input) changes depending upon the prediction errors.\\
 We immediately observe that the energy usage in the ideal and perturbed cases is very close, especially in the small violation case.
 Our interpretation for this is that the DPC, in the small violation case, keeps the temperature within the bounds with an extremely small error in bounds violation.
-The only exception can be seen before May $3^{rd}$ (03/05) where, due to a large prediction error, the temperature violates the bounds: as a consequence the controller uses much more energy (with respect to the perfect forecast knowledge case) to bring the temperature in the bounds, as evidenced by the blacked lines of Figure \ref{F:comparison_all_energy_E+_noisy}.
+The only exception can be seen before May $3^{rd}$ (03/05) where, due to a large prediction error, the temperature violates the bounds: as a consequence the controller uses more energy (with respect to the perfect forecast knowledge case) to bring the temperature in the bounds, as evidenced by the blacked lines of Figure \ref{F:comparison_all_energy_E+_noisy}.
 In the medium and large violation cases, instead, when the temperature violates the bounds the DPC does not strongly increase the energy usage as the violation tolerance is more relaxed: this produces a general reduction of energy usage in the "noisy" simulations, as evidenced by the purple and red lines of Figure \ref{F:comparison_all_energy_E+_noisy}.

### 2018-01-22 paper/final/Cover letter.tex
small changes

\documentclass[12pt]{article}

\begin{document}

\title{Cover Letter}
\date{}
\author{}
\maketitle 
\thispagestyle{empty}

In this paper we introduce a novel idea for predictive control based on historical building data, leveraging machine learning algorithms like regression trees and random forests. We call this approach Data-driven model Predictive Control (DPC), and we apply it to three different case studies to demonstrate its performance, scalability and robustness. This is the first time that Random Forests are bridged towards Predictive Control theory, and used to optimize energy consumption, thermal comfort, and peak power reduction in buildings. This approach overcomes the complexity drawbacks of the classical model-based approaches, that limit their widespread in real applications, and provides better results with respect to the current used rule-based approaches, while being extremely simple to be implemented in C, C++, Python and Matlab, and thus is easily integrable in SCADA systems.

\paragraph{What is the novelty of this work?}
We introduce a novel idea for predictive control based on historical building data leveraging machine learning algorithms like regression trees and random forests. We call this approach Data-driven model Predictive Control (DPC). This is the first time Random Forests based algorithms are bridged with the Predictive Control theory. This approach has shown its performance, scalability and robustness, with respect to uncertainties due to real data acquisition and weather forecast inaccuracies, and under different scenarios of energy and thermal comfort optimization problems, i.e. Demand Response and peak power curtailment, energy and thermal comfort optimization, in buildings.

\paragraph{Is the paper appealing to a popular or scientific audience?}
This paper is appealing for a vast audience (taken from the list available on the Applied Energy's guideline):
\begin{itemize}
	\item Mechanical, and alternative renewable energy sources including biomass: since, (1) a real house has been considered making a study on the thermal properties of the structure, and (2) a thermal plant (based on biomass) schedule has been optimized for energy efficiency;
	\item Building, policy and decision makers in energy conservation and conversion, and sustainable energy systems : since the whole paper focuses on a new methodology, that is data-driven, for energy and thermal comfort optimization, and peak power reduction in buildings;
	\item Researchers: in the paper we create a new methodology to bridge machine learning and control, so the work is appealing also for these two communities.

\end{itemize}

\paragraph{Why the authors think the paper is important and why the journal should publish it?}
The paper introduce a novel methodology for energy and thermal comfort optimization in buildings. In particular, this methodology provides a new data-driven control technique, Data-driven model Predictive Control (DPC), that offer many advantages in terms of practical applicability with respect to the classical model-based approaches, and improves the performance of the rule-based approaches, while being extremely simple to be implemented in real applications.
\paragraph{Has the article been checked by a native tongue speaker with expertise in the field?}
Yes
\paragraph{Are you available as a reviewer for at least three other articles for Applied Energy during the current year?}
Yes

\end{document}

### 2018-01-22 paper/final/main.tex
reviewed Intro, Sec 2,3,4,5, conclusion and appendix
small changes

@@
 
-\author[FSaddress,EC]{Francesco Smarra\corref{mycorrespondingauthor}}
+\author[FSaddress,AJaddress,EC]{Francesco Smarra\corref{mycorrespondingauthor}}
 \cortext[mycorrespondingauthor]{Corresponding author}
 \ead{francesco.smarra@univaq.it}
+\ead{fsmarra@seas.upenn.edu}
 
@@
 \begin{keyword}
-Demand response, energy efficiency, predictive control, machine learning, EnergyPlus modeling, real application
+Energy efficiency, energy saving, energy system analysis, energy system simulation, energy system modeling, energy system and automation, sustainable buildings.
 \end{keyword}
@@
 
+\appendix
+\input{appendixRegTrees.tex}
+
 %% BIBLIOGRAPHY
@@
 
-\appendix
-\input{appendixRegTrees.tex}
 \end{document}

### 2018-01-22 paper/final/Response_to_reviewers/Response_to_reviewers.tex
response to reviewers completed

--- 2018-01-21 paper/final/Response_to_reviewers/Response_to_reviewers.tex
@@
 \documentclass{article}
+
 \textwidth 6in%
+
 \textheight 7.9in%
+
 \oddsidemargin 1cm
@@
 \usepackage{amsmath}
+
 \usepackage{amsfonts}
+
 \usepackage{amssymb}
+
 \usepackage{graphicx}%
+
 \usepackage[dvips]{color}
+
 \setcounter{MaxMatrixCols}{30}
+
 \usepackage{natbib}
@@
 \maketitle
-
 \bigskip
 
-The authors would like to thank the Associate Editor and the Reviewers for their constructive and helpful comments.
-The following major and minor revisions have been applied (all changes in the manuscript are highlighted in \textcolor[rgb]{0.00,0.00,1.00}{blue colored fonts}):
+The authors would like to thank the Editors, the Associate Editor and the Reviewers for their constructive and helpful comments, which helped us to improve both the quality and the presentation of the paper.
+We performed a major revision taking into account all their comments (changes in the manuscript are highlighted in \textcolor[rgb]{0.00,0.00,1.00}{blue colored fonts}), and in particular:
 
@@
 	\item Section 1 - INTRODUCTION
+
 	\begin{itemize}
-		\item The introduction has been completely re-organized and rewritten in most of its parts, with a more detailed explanation of the paper contribution;
-		\item Literature review and related work have been updated;
+
+		\item The introduction has been completely re-organized and rewritten in most of its parts, with a more precise and detailed explanation of the paper contribution;
+
+		\item Literature review and related work have been updated and extended;
+
 	\end{itemize}
@@
 	\item Section 2 - COMPLEXITY ISSUES IN MODELING BUILDINGS: PHYSICS-BASED VS DATA-DRIVEN
+
 	\begin{itemize}
-		\item This section has been added to highlight the importance of our contribution, that is data-driven, with respect to model-based approaches;
+
+		\item This section has been added to the paper to highlight in detail the advantages of data-based modeling with respect to physics-based modeling for buildings;
+
 	\end{itemize}
@@
 	\item Section 3 - DATA PREDICTIVE CONTROL
+
 	\begin{itemize}
-		\item We rewrote the section to make it more clear;
-		\item We provided additional details to make the mathematical model description more clear and self-contained;
+
+		\item We rewrote the section to make the presentation more clear;
+
+		\item We provided additional details to make the mathematical  description more clear and rigorous;
+
 	\end{itemize}
 
-	\item Section 6 - CASE STUDY: OPTIMAL HEATING SYSTEM SCHEDULING
+	\item Section 6 - APPLICATION TO OPTIMAL HEATING SYSTEM SCHEDULING
+
 	\begin{itemize}
-		\item We highlighted the differences with respect to Section 5;
-		\item We added new results that account for weather forecast uncertainties and show the robustness of the proposed approach;
+
+		\item We highlighted the differences in contribution with respect to Section 5;
+
+		\item We added a new section with new results that account for weather forecast uncertainties and show the robustness of the proposed approach;
+
 	\end{itemize}
 
-	\item Minor changes
+	\item Other changes
+
 	\begin{itemize}
+
 		\item We updated the abstract to make it more complete and clear;
-		\item We reviewed the english;
-		\item We improved the notation and corrected some mistakes;
+
+		\item We performed a deep review to improve both language and organisation quality;
+
+		\item We improved the notation and corrected some typos;
+		\item We added Appendix A to provide further technical details describing the classical Regression Tree algorithm.
+
 	\end{itemize}
@@
 
-The following are itemised lists of the authors' response to the Reviewers and to the Associate Editor.
+The following sections consist of itemized lists of the detailed authors' response to the Reviewers, to the Editors and to the Associate Editor.
 
-\section{Response to Associate Editor}
+%================================================================================
+
+\section{Response to the Associate Editor}
 
@@
 
-	\textbf{Authors:}
+	\textbf{Authors:} We addressed all the reviewers' points, as discussed in the following of this document.
+We also emphasized the novelty of the paper with respect to prior work in the "Related work" subsection of the introduction.
 
-	\item \textbf{Associate Editor:} \textit{An updated and complete literature review should be conducted.}
+\end{enumerate}
 
-	\textbf{Authors:} We performed a deep literature review. In particular, in Table 1 we classify the related work and then emphasise the novelty of our research.
+\section{Response to the Editors}
 
-	\item \textbf{Associate Editor:} \textit{The relevance to Applied Energy should be enhanced with the considerations of scope and readership of the Journal.
+\begin{enumerate}
+
+	\item \textbf{Editor:} \textit{An updated and complete literature review should be conducted.}
+
+	\textbf{Authors:} We conducted and updated and complete literature review.
+In particular, in Table 1 we classify the related work and then emphasize the novelty of our research with respect to each prior work.
+
+	\item \textbf{Editor:} \textit{The relevance to Applied Energy should be enhanced with the considerations of scope and readership of the Journal.
 The present form has little relevance to APEN which might suit better for another journal.
@@
 
-	\textbf{Authors:} Following the Associate Editor and reviewers comments, we re-wrote the title and the introduction of the paper focusing on the relevance of the paper to APEN.
-In particular, we emphasise that...EXPLAIN HERE AS IN INTRODUCTION
+	\textbf{Authors:} In the initial submission we had chosen a more general title and discussion in the introduction because we believe that our approach has the potential to be applied not only to building management, but also to other application domains.
+However, as in this paper we develop our techniques focusing and experimenting in building management we changed the title in "Data-Driven Model Predictive Control using Random Forests for Building Energy Optimization and Climate Control" and re-wrote the introduction of the paper focusing on the relevance of the paper contribution to building management.\\
+	Now that we better emphasised the relevance of the paper to building management, and since "\textit{Applied Energy provides a forum for information on innovation, research, development and demonstration in the areas of energy conversion and conservation, the optimal use of energy resources, analysis and optimization of energy processes, mitigation of environmental pollutants, and sustainable energy systems" (https://www.journals.elsevier.com/applied-energy)}, we really hope that our paper will be considered relevant to APEN, especially with respect to the topics "\textit{optimal use of energy resources}" and "\textit{analysis and optimization of energy processes}".
+We also hope that our paper will be considered related to the APEN keywords \textit{Energy efficiency and energy saving: strategies, applications and implications}, \textit{Energy system analysis, simulation and modeling}, \textit{Implementation of energy R\&D strategies and policies}, \textit{Energy system and automation} and \textit{Sustainable buildings}.
 
-	\item \textbf{Associate Editor:} \textit{A proof reading by a native English speaker should be conducted to improve both language and organization quality. }
+	\item \textbf{Editor:} \textit{A proof reading by a native English speaker should be conducted to improve both language and organization quality.}
 
@@
 
-	\item \textbf{Associate Editor:} \textit{The originality of the paper needs to be further clarified.
+	\item \textbf{Editor:} \textit{The originality of the paper needs to be further clarified.
 The present form does not have sufficient results to justify the novelty of a high quality journal paper.
@@
 
-	\textbf{Authors:} We clarify in the introduction and in Section ???
-the originality of the paper and how are results can be used for real applications.
-In particular, we emphasise that...EXPLAIN HERE AS IN INTRODUCTION
+	\textbf{Authors:} In the "Related work" and "Main contribution" subsections of the Introduction, we clarified the novelty of the paper with respect to existing results and emphasized the contributions provided.\\
+		In the newly added Section 2 we highlight the typical drawbacks of the model-based approaches, which limit the use of advanced control techniques for energy optimization such as MPC due to the difficulties (cost, time and effort) related to the model creation, and show how the data-driven approach that we propose (DPC) overcomes these limitation.\\
+		In Sections 4,5 and 6 we show how our DPC methodology can be used for real applications.
+In particular, in Section 4 we demonstrate that DPC can provide comparable performance with respect to classical MPC applied to a perfectly known mathematical model.
+In Section 5 we apply DPC to a 6 story 22 zone building model in EnergyPlus, for which model-based control is not economical and practical due to extreme complexity, and address a Demand Response problem.
+Our results demonstrate scalability and efficiency of DPC showing that it provides the desired power curtailment with an average error of 3\%.
+In Section 6 we implement and test DPC on real data from an off-grid house located in L'Aquila, Italy.
+We compare the total amount of energy saved with respect to the classical bang-bang controller, showing that DPC can perform an energy saving up to $49.2\%$.
+We also demonstrate the robustness of our method to uncertainties both in real data acquisition (affected by unpredictable house schedules due to people, appliances, etc., noisy sensor measurements due to real experimental data, etc.) and weather forecast.
+\end{enumerate}
 
-\end{enumerate}
+%================================================================================
 
@@
 
-\textbf{Authors:}
-
-	\begin{itemize}
-
-		\item Explain in the introduction (other than in section 2) the difference between the state-space modeling and the one we propose;
-
-		\item Add a subsection in the "Related work" to explain the evolution of our work with respect to our previous ones;
-
-		\item Add Appendix on regression trees and random forests;
-
-		\item Add equations in the intro.
-In particular, add a section where we quantify the complexity of a physics-based state-space modeling through an example.
-Setup and write equation to show the complexity
-
-		\item add another section to show how the previous point simplifies using data-driven methods
-
-	\end{itemize}
+\textbf{Authors:} We thank the reviewer for this comment, this very important part of the paper was not clear in the initial submission.
+We added a new section (Section 2) where we reported a traditional physics-based state-space modeling of a building (Section 2.1) and compared it with the data-driven modeling approach (Section 2.2).
+We highlighted the differences in the modeling techniques and emphasized that the features needed in the data-driven approaches are ALL and ONLY the measurable variables (hence a subset of all the variables) used in the model-based approach, since the model learning procedure is able to compensate the effect of the non-measurable variables.
+This is also further discussed in Section 3, where we added more mathematical details to make the presentation more clear and rigorous.
 
@@
 
-\textbf{Authors:} We thank the reviewer for this comment, indeed our previous presentation of the paper was unclear on this fundamental aspect.
+\textbf{Authors:} We thank the reviewer for this comment, indeed our previous presentation of the paper was unclear on this fundamental aspect, we apologise for this.
 In our revision we widely illustrate in Section 2, how in traditional state-space models part of the state is measurable (i.e.
@@
 layers/windows temperatures, etc).
-In our approach, of course, $(\mathcal{X}^c, \mathcal{X}^d, \mathcal{Y})$ are ALL measurable variables, i.e.
+In our approach, of course, $(\mathcal{X}, \mathcal{Y})$ are ALL measurable variables, i.e.
 variables that are present in the historical data of a building.
-The unmeasurable variables of the traditional state-space models are not assumed to be measurable in this paper and are neglegted in the proposed models, yet our prediction works well since we compensate them via the regression tree/random forest approaches.
+The unmeasurable variables of the traditional state-space models are not assumed to be measurable in this paper, yet our prediction works well since we compensate their effect via the regression tree/random forest approaches.
 
@@
 \textbf{Authors:} We agree with the reviewer.
-Following this comment, and also other comments, we re-wrote the corresponding Section to make it more clear and complete to descrive the DPC technical approach.
-In particular...
+Following this comment, and also other comments, we re-wrote the corresponding Section 3 (describing the DPC technical approach) to make the presentation and notation more clear and the mathematical description more rigorous.
+We first replaced Eq.(1) of the initial submission with Eq.(5), i.e.
+$y(k)=f(x(k),u(k),d(k))$, of the current version.
+In Section 3.1 we re-wrote in a more clear and detailed way the two-steps process (based on an adaptation of the Regression Tree algorithm) used to construct from the historical data the predictive model $f$, and we also clarified the relationship between such two-steps process and Fig.
+1 (left).
+We also added a remark that better explains Fig.
+1 (right) and Appendix A to provide further technical details describing the classical Regression Tree algorithm.
 
 \end{enumerate}
+
+%================================================================================
 
@@
 \textbf{Authors:} We agree with the Reviewer.
-In the beginning, we had chosen a more general title because we believe that our approach has the potential to be applied not only to building management, but also to other application domains.
-However, as in this paper we develop our techniques focusing and experimenting in building management (which is indeed a core topic of APEN) we changed the title in "Data-Driven Model Predictive Control using Random Forests for Energy and Thermal Comfort Optimization in Buildings".
+In the initial submission we had chosen a more general title (and discussion in the introduction) because we believe that our approach has the potential to be applied not only to building management, but also to other application domains.
+However, as in this paper we develop our techniques focusing and experimenting in building management (which is indeed a core topic of APEN) we changed the title in "Data-Driven Model Predictive Control using Random Forests for Building Energy Optimization and Climate Control" (and re-wrote the introduction of the paper focusing on the relevance of the paper contribution to building management).
 
@@
 
-\textbf{Authors:} In the introduction of the revised version, we widely discuss that in this paper we provide new results with respect to [9-12] (that now are [25,26,31,32]) and adapted them to be applicable to the building automation use-case presented in Section 6.
+\textbf{Authors:} In subsection "1.1 Related work" of the revised version we discuss in detail the novelty of our paper beyond [9,10,11,12] (which are numbered [25,26,31,32] in the revised version).
+We summarise the novel contributions of our paper in the initial paragraph of subsection "1.2 Main contribution":
+\textit{In this paper, we provide a new methodology based on random forests that overcomes the drawbacks of all our previous works, and more precisely: we obtain better performance and scalability when compared to other approaches (both optimal and rule-based), and we provide and validate robustness with respect to uncertainties due to real data acquisition and weather forecast inaccuracies.}
+After the above paragraph in the paper, we describe our contributions in detail.
 
@@
 \textbf{Authors:} We agree with the reviewer.
-Following this comment, and also other comments, we completely re-wrote the corresponding Section (that now is Section 3) to make it more clear and complete to describe the DPC technical approach.
+Following this comment, and also other comments, we re-wrote the corresponding Section 3 (describing the DPC technical approach) to make the presentation and notation more clear and the mathematical description more rigorous.
+We also added Appendix A to provide further technical details describing the classical Regression Tree algorithm.
 
@@
 \textbf{Authors:} We thank the reviewer for rising this important issue.
-Yes, we assume that at time $k$ we know the future non-manipulative states (namely external temperature, radiation, etc.) for the Model Predictive Control time horizon.
-In the revised version of this paper, to answer the reviewer comment, we account for disturbance uncertainties and show that, even with a large error in predicting the future states, the performance does not change significantly.
-More precisely, we:
+As discussed in the previous answer, in this revision we re-wrote Section 3 and improved the description of the main optimization problems.
+Answering to the specific questions of the reviewer: yes, as we look ahead to do DPC each future time slot have a different linear model, depending on future non-manipulative states; yes, we assume that at time $k$ we know the future non-manipulative states (namely external temperature, radiation, etc.) for the Model Predictive Control time horizon.
+In the revised version of this paper, to justify this assumption, we first provide a reference paper [57] that investigates the effect of uncertainty in the weather forecast on the performance of MPC in building systems operations through a large-scale simulation study, and compares against a rule-based strategy.
+They consider 48 different scenarios of uncertainties for 72 hours weather forecast.
+With such a long horizon, results have shown that (with a few exceptions) MPC outperforms the rule-based controller in providing energy savings, and is in general quite close to the perfect forecast case despite the uncertainty in weather forecast.
+Since the length of horizon for the DPC algorithm presented in this paper is usually much shorter, for example $6$ hours in Section 4, $7$ hours in Section 5, and $40$ minutes in Section 6, we can reasonably presume that our approach will be robust to weather forecast inaccuracies.
+In our opinion, robustness of DPC to weather forecast inaccuracies is due to 2 main features:
+
 	\begin{itemize}
-		\item Cite a study that shows, despite considering uncertainties on weather forecast on a long horizon of 72 hours, that MPC outperforms the rule-based controllers in providing energy savings, and is in general quite close to the perfect forecast case.
-This is due mainly to 2 main features:
-		\begin{itemize}
-			\item weather forecast are in general quite accurate on a shot-term horizon (especially on the length of horizon that we usually consider for DPC, for example 6 hours in Section 4, 7 hours in Section 5 and 40 minutes in Section 6);
-			\item MPC strategy computes an optimal input sequence over an horizon of finite length, and then apply only the first input.
-At the next step, with the new EXACT measurements of the system state and of the disturbance conditions, the algorithm is applied again.
-This is how MPC uses feedback law to correct modeling errors;
+
+	\item weather forecast are in general quite accurate on a short-term horizon;
+
+	\item the MPC strategy computes an optimal input sequence over an horizon of finite length, and then applies only the first input.
+At the next step, with the new measurements of the system state and of the disturbance conditions, the algorithm is applied again.
+For this reason MPC is in general very robust to uncertainties;
+
 	\end{itemize}
-		\item Provide a new case study in Section 6, where we inject a noise into the disturbance and show the robustness of our approach.
-	\end{itemize}
+	Finally, in Section 6 we add a new subsection where we test the robustness of our approach by perturbing the weather forecast with Gaussian noise with large variance.
+The results show that the control performance is indeed very close to the ideal case even with a large error in the weather forecast.
 
@@
 
-\textbf{Authors:} Achin?
+\textbf{Authors:} We have clarified this point in the text and the Figure caption.
+In particular, marked in blue is the response when the optimal input is applied to the power predictive model of the Random Forest and in red is the response when the same input is applied to the power predictive model of EnergyPlus.
+Since the optimal input is computed using the power predictive model of the Random Forest the blue trajectory perfectly follows the tracking signal.
+The red trajectory, as expected, is characterised by a (small) tracking error because of the model mismatch between the predictive model of the Random Forest (used to compute the input) and that of EnergyPlus (used to simulate the closed-loop system).
 
 \item \textbf{Reviewer 2:} \textit{Also, what is new in section 5 compared with section 4?
-Even though a real house is claimed to be used, it's still tested using simulations and energyplus, like in the previous sections.}
+Even though a real house is claimed to be used, it's still tested using simulations and energyplus, like in the previous sections.
+A real world experiment on this house would be much more interesting.}
 
-\textbf{Authors:} We thank the reviewer for this comment, indeed the contribution of this section was not well emphasized.
-We improved this aspect at the beginning of the section.
-In particular, Section 6 provides the following differences with respect to Section 5:
+\textbf{Authors:} We thank the reviewer for this comment, indeed the contribution of Section 6 (i.e.
+Section 5 in the initial submission) was not well emphasized.
+We now emphasize the difference between Section 5 (i.e.
+Section 4 in the initial submission) and Section 6 (i.e.
+Section 5 in the initial submission):
 \begin{itemize}
-	\item In Section 5 we use the EnergyPlus model both to create the Random Forests model and to test the DPC performance.
+
+	\item As the main difference, in Section 5 we use the EnergyPlus model both to create the predictive model via Random Forests and to test the DPC performance.
 However, although the EnergyPlus model is quite accurate, it still provides simulated data, that differently from experimental data do not suffer of non-predictable imperfections such as random occupancy schedules, open/close windows, random light on/off switch, etc..
-In section 6, instead, experimental data are used to build the Random Forests models, while the EnergyPlus model (created ad-hoc for the house) is ONLY used to make a fair comparison.
-Otherwise, using Random Forests models to compare DPC (that would have used the same model) and the bang-bang control would have been unfair;
-	\item The optimization problems are different: in Section 5, we cut the peak of the power in a limited amount of time and require the temperature to be close to a desired value, while in Section 6 we minimize the energy usage while keeping temperature in a comfort range.
+In Section 6, instead, real data are used to build the Random Forests models, while the EnergyPlus model (created ad-hoc for the house) is ONLY used to compare the total amount of energy can be saved applying DPC with respect to the classical bang-bang controller.
+
+	\item The optimization problems are different: in Section 5 we cut the peak of power consumption in a limited amount of time and require the temperature to be close to a desired value, while in Section 6 we minimize the energy usage while keeping temperature in a comfort range.
 Furthermore, in Section 6 we consider an on/off input, i.e.
-integer input variable, instead of a continuous one, showing that DPC can adapt to different problems;
-	\item We introduce new simulations accounting for weather forecast uncertainties, other than the ones introduced with the experimental data, that show the robustness of DPC against disturbance errors.
+an integer input variable, instead of a continuous one, showing that  DPC can easily adapt to different environments in practice;
+	\item Last but not least, we add in the revised version a new subsection of Section 6 adding simulations accounting for weather forecast uncertainties, other than the ones introduced with the experimental data, that show the robustness of DPC against weather forecast inaccuracies.
+
 \end{itemize}
-
-\item \textbf{Reviewer 2:} \textit{A real world experiment on this house would be much more interesting.}
-
-\textbf{Authors:} We thank the reviewer for this comment, indeed the contribution of this section was not well emphasized.
-We improved this aspect at the beginning of the section.
-In particular, Section 6 provides the following differences with respect to Section 5:
+Finally, we agree with the reviewer that a real world experiment would be interesting.
+However, the aim of Section 6 was to compare the impact in energy saving of DPC and of the classical bang-bang controller: since they could not be ran experimentally simultaneously our simulative approach appears reasonable.
+Now that we have demonstrated in this paper \textit{performance}, \textit{scalability} and \textit{robustness} of DPC we are currently working to apply it on a real building of the University of L'Aquila (Italy), but this study is out of the scope of this paper.
 

### 2018-01-22 paper/final/Response_to_reviewers/Response_to_reviewers.tex
small changes

@@
 	\textbf{Authors:} In the "Related work" and "Main contribution" subsections of the Introduction, we clarified the novelty of the paper with respect to existing results and emphasized the contributions provided.\\
-		In the newly added Section 2 we highlight the typical drawbacks of the model-based approaches, which limit the use of advanced control techniques for energy optimization such as MPC due to the difficulties (cost, time and effort) related to the model creation, and show how the data-driven approach that we propose (DPC) overcomes these limitation.\\
+		In the newly added Section 2, we highlighted the typical drawbacks of the model-based approaches, which limit the use of advanced control techniques for energy optimization such as MPC due to the difficulties (cost, time and effort) related to the model creation, and show how the data-driven approach that we propose (DPC) overcomes these limitations.\\
 		In Sections 4,5 and 6 we show how our DPC methodology can be used for real applications.
@@
 
-\textbf{Authors:} We thank the reviewer for this comment, this very important part of the paper was not clear in the initial submission.
+\textbf{Authors:} We thank the reviewer for this comment, indeed this very important part of the paper was not clear in the initial submission.
 We added a new section (Section 2) where we reported a traditional physics-based state-space modeling of a building (Section 2.1) and compared it with the data-driven modeling approach (Section 2.2).
@@
 
-\textbf{Authors:} We thank the reviewer for this comment, indeed our previous presentation of the paper was unclear on this fundamental aspect, we apologise for this.
+\textbf{Authors:} We thank the reviewer for this comment, indeed our previous presentation of the paper was unclear on this fundamental aspect, we apologize for this.
 In our revision we widely illustrate in Section 2, how in traditional state-space models part of the state is measurable (i.e.
@@
 In our approach, of course, $(\mathcal{X}, \mathcal{Y})$ are ALL measurable variables, i.e.
-variables that are present in the historical data of a building.
+variables that are present in the historical dataset of a building.
 The unmeasurable variables of the traditional state-space models are not assumed to be measurable in this paper, yet our prediction works well since we compensate their effect via the regression tree/random forest approaches.
@@
 $y(k)=f(x(k),u(k),d(k))$, of the current version.
-In Section 3.1 we re-wrote in a more clear and detailed way the two-steps process (based on an adaptation of the Regression Tree algorithm) used to construct from the historical data the predictive model $f$, and we also clarified the relationship between such two-steps process and Fig.
+In Section 3.1 we re-wrote in a more clear and detailed way the two-steps process (based on an adaptation of the Regression Tree algorithm) used to construct, from the historical dataset, the predictive model $f$, and we also clarified the relationship between such two-steps process and Fig.
 1 (left).
 We also added a remark that better explains Fig.
-1 (right) and Appendix A to provide further technical details describing the classical Regression Tree algorithm.
+1 (right), and Appendix A to provide further technical details describing the classical Regression Tree algorithm.
 
@@
 In the initial submission we had chosen a more general title (and discussion in the introduction) because we believe that our approach has the potential to be applied not only to building management, but also to other application domains.
-However, as in this paper we develop our techniques focusing and experimenting in building management (which is indeed a core topic of APEN) we changed the title in "Data-Driven Model Predictive Control using Random Forests for Building Energy Optimization and Climate Control" (and re-wrote the introduction of the paper focusing on the relevance of the paper contribution to building management).
+However, as in this paper we develop our techniques focusing and experimenting in building management (which is indeed a core topic of APEN), we changed the title in "Data-Driven Model Predictive Control using Random Forests for Building Energy Optimization and Climate Control" (and re-wrote the introduction of the paper focusing on the relevance of the paper contribution to building management).
 
@@
 As discussed in the previous answer, in this revision we re-wrote Section 3 and improved the description of the main optimization problems.
-Answering to the specific questions of the reviewer: yes, as we look ahead to do DPC each future time slot have a different linear model, depending on future non-manipulative states; yes, we assume that at time $k$ we know the future non-manipulative states (namely external temperature, radiation, etc.) for the Model Predictive Control time horizon.
+Answering to the specific questions of the reviewer: yes, as we look ahead to do DPC, each future time slot has a different linear model, depending on future non-manipulative disturbances; yes, we assume that, at time $k$, we know the future non-manipulative disturbances (namely external temperature, radiation, etc.) for the Model Predictive Control time horizon.
 In the revised version of this paper, to justify this assumption, we first provide a reference paper [57] that investigates the effect of uncertainty in the weather forecast on the performance of MPC in building systems operations through a large-scale simulation study, and compares against a rule-based strategy.
@@
 	\end{itemize}
-	Finally, in Section 6 we add a new subsection where we test the robustness of our approach by perturbing the weather forecast with Gaussian noise with large variance.
+	Finally, in Section 6 we add a new subsection where we test the robustness of our approach by perturbing the weather forecast with Gaussian noises with large variance.
 The results show that the control performance is indeed very close to the ideal case even with a large error in the weather forecast.
@@
 In particular, marked in blue is the response when the optimal input is applied to the power predictive model of the Random Forest and in red is the response when the same input is applied to the power predictive model of EnergyPlus.
-Since the optimal input is computed using the power predictive model of the Random Forest the blue trajectory perfectly follows the tracking signal.
+Since the optimal input is computed using the power predictive model of the Random Forest, the blue trajectory perfectly follows the tracking signal.
 The red trajectory, as expected, is characterised by a (small) tracking error because of the model mismatch between the predictive model of the Random Forest (used to compute the input) and that of EnergyPlus (used to simulate the closed-loop system).
@@
 	\item As the main difference, in Section 5 we use the EnergyPlus model both to create the predictive model via Random Forests and to test the DPC performance.
-However, although the EnergyPlus model is quite accurate, it still provides simulated data, that differently from experimental data do not suffer of non-predictable imperfections such as random occupancy schedules, open/close windows, random light on/off switch, etc..
+However, although the EnergyPlus model is quite accurate, it still provides simulated data, that differently from real experimental data do not suffer of non-predictable imperfections such as random occupancy schedules, open/close windows, random light on/off switch, etc..
 In Section 6, instead, real data are used to build the Random Forests models, while the EnergyPlus model (created ad-hoc for the house) is ONLY used to compare the total amount of energy can be saved applying DPC with respect to the classical bang-bang controller.
 
-	\item The optimization problems are different: in Section 5 we cut the peak of power consumption in a limited amount of time and require the temperature to be close to a desired value, while in Section 6 we minimize the energy usage while keeping temperature in a comfort range.
+	\item The optimization problems are different: in Section 5 we cut the peak of power consumption in a limited amount of time and require the temperature to be close to a desired value, while in Section 6 we minimize the energy usage while keeping the temperature within a comfort range.
 Furthermore, in Section 6 we consider an on/off input, i.e.
 an integer input variable, instead of a continuous one, showing that  DPC can easily adapt to different environments in practice;
-	\item Last but not least, we add in the revised version a new subsection of Section 6 adding simulations accounting for weather forecast uncertainties, other than the ones introduced with the experimental data, that show the robustness of DPC against weather forecast inaccuracies.
+	\item Last but not least, we added in the revised version a new subsection in Section 6, providing simulations accounting for weather forecast uncertainties, other than the ones introduced with the experimental data, that show the robustness of DPC against weather forecast inaccuracies.
 

### 2018-01-23 paper/final/Response_to_reviewers/Response_to_reviewers.tex
Minor edits

@@
 	\textbf{Authors:} In the initial submission we had chosen a more general title and discussion in the introduction because we believe that our approach has the potential to be applied not only to building management, but also to other application domains.
-However, as in this paper we develop our techniques focusing and experimenting in building management we changed the title in "Data-Driven Model Predictive Control using Random Forests for Building Energy Optimization and Climate Control" and re-wrote the introduction of the paper focusing on the relevance of the paper contribution to building management.\\
-	Now that we better emphasised the relevance of the paper to building management, and since "\textit{Applied Energy provides a forum for information on innovation, research, development and demonstration in the areas of energy conversion and conservation, the optimal use of energy resources, analysis and optimization of energy processes, mitigation of environmental pollutants, and sustainable energy systems" (https://www.journals.elsevier.com/applied-energy)}, we really hope that our paper will be considered relevant to APEN, especially with respect to the topics "\textit{optimal use of energy resources}" and "\textit{analysis and optimization of energy processes}".
+However, as in this paper we develop our techniques focusing and experimenting in building management we changed the title in ``Data-Driven Model Predictive Control using Random Forests for Building Energy Optimization and Climate Control" and re-wrote the introduction of the paper focusing on the relevance of the paper contribution to building energy management, bridging machine learning and control for energy flexibility.\\\\
+	Now that we better emphasized the relevance of the paper to building energy management, and since "\textit{Applied Energy provides a forum for information on innovation, research, development and demonstration in the areas of energy conversion and conservation, the optimal use of energy resources, analysis and optimization of energy processes, mitigation of environmental pollutants, and sustainable energy systems" (https://www.journals.elsevier.com/applied-energy)}, we really hope that our paper will be considered relevant to APEN, especially with respect to the topics "\textit{optimal use of energy resources}" and "\textit{analysis and optimization of energy processes}".
 We also hope that our paper will be considered related to the APEN keywords \textit{Energy efficiency and energy saving: strategies, applications and implications}, \textit{Energy system analysis, simulation and modeling}, \textit{Implementation of energy R\&D strategies and policies}, \textit{Energy system and automation} and \textit{Sustainable buildings}.
@@
 
-	\textbf{Authors:} We deeply revised the paper to improve both language and organization quality.
+	\textbf{Authors:} We thoroughly revised and proofread the paper to improve both language and organization quality.
 
@@
 
-	\textbf{Authors:} In the "Related work" and "Main contribution" subsections of the Introduction, we clarified the novelty of the paper with respect to existing results and emphasized the contributions provided.\\
+	\textbf{Authors:} In the ``Related work" and ``Main contribution" subsections of the Introduction, we clarified the novelty of the paper with respect to existing results and emphasized the contributions provided.\\
+
 		In the newly added Section 2, we highlighted the typical drawbacks of the model-based approaches, which limit the use of advanced control techniques for energy optimization such as MPC due to the difficulties (cost, time and effort) related to the model creation, and show how the data-driven approach that we propose (DPC) overcomes these limitations.\\
+
 		In Sections 4,5 and 6 we show how our DPC methodology can be used for real applications.
 In particular, in Section 4 we demonstrate that DPC can provide comparable performance with respect to classical MPC applied to a perfectly known mathematical model.
-In Section 5 we apply DPC to a 6 story 22 zone building model in EnergyPlus, for which model-based control is not economical and practical due to extreme complexity, and address a Demand Response problem.
+In Section 5, we apply DPC to a 6 story 22 zone building model in EnergyPlus, for which model-based control is not economical and practical due to extreme complexity, and address a Demand Response problem.
 Our results demonstrate scalability and efficiency of DPC showing that it provides the desired power curtailment with an average error of 3\%.
@@
 \textbf{Authors:} We thank the reviewer for this comment, indeed our previous presentation of the paper was unclear on this fundamental aspect, we apologize for this.
-In our revision we widely illustrate in Section 2, how in traditional state-space models part of the state is measurable (i.e.
+In our revision we clearly illustrate in Section 2, how in traditional state-space models part of the state is measurable (i.e.
 room temperatures, etc) and part is not measurable (i.e.
-layers/windows temperatures, etc).
+layers/windows temperatures, time of day, day of week, holidays, etc).
 In our approach, of course, $(\mathcal{X}, \mathcal{Y})$ are ALL measurable variables, i.e.
 variables that are present in the historical dataset of a building.
-The unmeasurable variables of the traditional state-space models are not assumed to be measurable in this paper, yet our prediction works well since we compensate their effect via the regression tree/random forest approaches.
+The unmeasurable variables of the traditional state-space models are not assumed to be measurable in this paper, yet our prediction works well since we (implicitly) account for their effect via the regression tree/random forest approaches.
 
@@
 \textbf{Authors:} We agree with the Reviewer.
-In the initial submission we had chosen a more general title (and discussion in the introduction) because we believe that our approach has the potential to be applied not only to building management, but also to other application domains.
-However, as in this paper we develop our techniques focusing and experimenting in building management (which is indeed a core topic of APEN), we changed the title in "Data-Driven Model Predictive Control using Random Forests for Building Energy Optimization and Climate Control" (and re-wrote the introduction of the paper focusing on the relevance of the paper contribution to building management).
+In the initial submission we had chosen a more general title (and discussion in the introduction) because we believe that our approach has the potential to be applied not only to building energy management, but also to other application domains.
+However, as in this paper we develop our techniques focusing and experimenting in building management (which is indeed a core topic of APEN), we changed the title to "Data-Driven Model Predictive Control using Random Forests for Building Energy Optimization and Climate Control" (and re-wrote the introduction of the paper focusing on the relevance of the paper contribution to building energy management).
 
@@
 \textbf{Authors:} In subsection "1.1 Related work" of the revised version we discuss in detail the novelty of our paper beyond [9,10,11,12] (which are numbered [25,26,31,32] in the revised version).
-We summarise the novel contributions of our paper in the initial paragraph of subsection "1.2 Main contribution":
+We summarize the novel contributions of our paper in the initial paragraph of subsection "1.2 Main contribution":
 \textit{In this paper, we provide a new methodology based on random forests that overcomes the drawbacks of all our previous works, and more precisely: we obtain better performance and scalability when compared to other approaches (both optimal and rule-based), and we provide and validate robustness with respect to uncertainties due to real data acquisition and weather forecast inaccuracies.}
@@
 
-	\item weather forecast are in general quite accurate on a short-term horizon;
+	\item weather forecasts, are in general, quite accurate on a short-term horizon;
 
@@
 \item \textbf{Reviewer 2:} \textit{Also, what is new in section 5 compared with section 4?
-Even though a real house is claimed to be used, it's still tested using simulations and energyplus, like in the previous sections.
+Even though a real house is claimed to be used, it's still tested using simulations and EnergyPlus, like in the previous sections.
 A real world experiment on this house would be much more interesting.}
@@
 	\item As the main difference, in Section 5 we use the EnergyPlus model both to create the predictive model via Random Forests and to test the DPC performance.
-However, although the EnergyPlus model is quite accurate, it still provides simulated data, that differently from real experimental data do not suffer of non-predictable imperfections such as random occupancy schedules, open/close windows, random light on/off switch, etc..
-In Section 6, instead, real data are used to build the Random Forests models, while the EnergyPlus model (created ad-hoc for the house) is ONLY used to compare the total amount of energy can be saved applying DPC with respect to the classical bang-bang controller.
+Although the EnergyPlus model is quite accurate, it still provides simulated data, that differs from real experimental data, which does not suffer from non-predictable imperfections such as random occupancy schedules, open/close windows, random light on/off switch, etc.
+In Section 6, instead, real data is used to build the Random Forests models, while the EnergyPlus model (created ad-hoc for the house) is ONLY used to compare the total amount of energy can be saved applying DPC with respect to the classical bang-bang controller.
 
